### 2021-03-23 thesis/test/response_2.tex
add tex files

% LaTeX rebuttal letter example. 
% 
% Copyright 2019 Friedemann Zenke, fzenke.net

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}

\usepackage{xifthen}
\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}
\newcounter{subpoint}[reviewer]
\setcounter{subpoint}{0}

% command declarations for reviewer points and our responses
\newcommand{\reviewersection}{\stepcounter{reviewer} \bigskip \hrule
                  \section*{Reviewer \thereviewer}}

\newenvironment{point}
   {\refstepcounter{point} \bigskip \noindent {Point \textbf{\arabic{point}} -- } \ }
   {\par }

\newenvironment{subpoint}
   {\refstepcounter{subpoint} \bigskip \noindent {Point \textbf{\arabic{point}.\arabic{subpoint}} -- } \ }
   {\par }

\newenvironment{action}
   {\medskip \noindent \begin{sf}\textbf{Action}:\ }
   {\medskip \end{sf}}

\newenvironment{reply}
   {\medskip \noindent \begin{sf}\textbf{Reply}:\ }
   {\medskip \end{sf}}

\begin{document}

\section*{Response to the reviewers}
% General intro text goes here
We thank the reviewers for their critical assessment of our work. 
In the following we address their concerns point by point.

% Let's start point-by-point with Reviewer 1
\reviewersection
The paper presents an analytical model of energy consumption by an application running on a multicore processor. The parameters of the model are the frequency, the number of active cores, and the problem size. The model is experimentally validated for several applications on a computer node equipped with two Xeon Phi E5-2698 16-core processors.
The submission has a potential for publication. However, the following reservations do not allow me to recommend it for publication in the current form:

% Point one description 
\begin{point}
\end{point}

\begin{subpoint}
The main difference of the presented model from other frequency-based models is the inclusion of the workload in the model (inclusion of the number of active cores is a minor difference, hardly worth of publication). However, this main contribution is practically not studied. The corresponding parameter N is neither motivated nor explained. It is not clear how to determine N for your application.
\end{subpoint}

\begin{reply}
%The main difference between the model presented model and other frequencies based on frequency is the inclusion of the workload (the number of active cores, we agree that makes a small difference, hardly worthy of being published). However, this main contribution, besides considering all parameters, is hardly available.
Besides the inclusion of the workload that can be considered the main contribution (the number of active cores, we agree that makes a small difference, hardly worthy of being published). The inclusion of all parameters is hardly available in fact we couldn't find any that considered all.

%We understand that the workload must be defined in a manner making possible to quantify N for all applications uniformly.  For that, we clarified the discussion about the input size in several parts of the paper, by providing a definition from our point of view (other variants are still possible, for instance when considering counting memory access operations). In section IV.B a general definition of input size was provided:
We clarified the input size in several parts of the paper by defining our perspective (other variants are still possible, for instance, when considering counting memory access operations). In section IV.B, a general definition of input size was provided:
"Finally, to fully characterize the application, a parameter representing the application's workload, called input size $N$, is introduced, representing the number of basic operations need to complete a problem \cite{Kumar1994AnalyzingArchitectures}"

\end{reply}

\begin{subpoint}
In the experimental subsection V.C, the authors say that for any application in their experiments they vary N from 1 to 5. What does it mean? What may that mean, for example, for matrix multiplication or an FFT if I want to use their model? Moreover, as N is the major new parameter, the proposed model should be compared with other energy models using workload as a model parameter, both theoretically and experimentally (for example, Shahid, Arsalan, et al. "A comparative study of techniques for energy predictive modelling using performance monitoring counters on modern multicore cpus." IEEE Access (2020)).
\end{subpoint}

\begin{reply}
The N parameter is a measure of the problem's size, defined as the number of basic operations to complete a task \cite{Kumar1994AnalyzingArchitectures}. We can say that the basics operations are the CPU instructions. Because of some CPU architecture characteristics as the pipeline, the instructions are executed approximately the same average time. Thus, it is possible to estimate the problem size by looking at the execution time, allowing us to divide a large problem size into several smaller ones as done in the work of Oliveira \cite{Oliveira2018ApplicationCores}.

The unit of measurement should accompany the definition. However, for simplicity, we assigned numbers from 1 the smallest size to 5 the largest, increasing the problem linearly, allowing us to interpolate any input in between these values.

Mapping the application input to problem size can be accomplished in many different ways. For instance, we could create a function to map one to another or use an correlation of some system event to estimate the problem size, but this is another major topic and cannot be covered in this journal.

We thank the author for the reference, but our journal focuses on the inclusion of all parameters as mentioned previously. Thus we do not go deep into each of the models since they are already validated in previous work.

We included this explanation of the input size division in section V.C:
"For the input size, by assuming that all CPU instructions are executed at approximately the same average time, the number of basic operations will be directly correlated with the time. Thus, we can estimate the problem size by looking at the execution time, allowing us to divide a large problem size into several smaller ones knowing their relationship as done in the work of Oliveira \cite{Oliveira2018ApplicationCores}.
The unit of measurement should accompany the definition. However, for simplicity, we assigned numbers from 1 the smallest size to 10 the largest, increasing the problem linearly, that way allow us also to interpolate any input in between these values."
\end{reply}

\begin{subpoint}
The authors do need to compare the accuracy of their model to the accuracy of the state-of-the-art models. They can set p to 32, fix the frequency, f, and build their model and other models, accounting for the workload size, for a wide range of workload sizes, and compare their accuracy. In the submitted paper, there is no single graph where the problem size would be a variable.
\end{subpoint}

\begin{reply}
To complement the results as requested by the reviewer, we also added figures in section V.D showing the model fitting regarding the input size.
\end{reply}

\begin{point}
The experimental part is rather poor lacking many important details. Most importantly, it is not clear how the energy consumption is measured. Without detailed description of the methodology, I cannot trust the experimental results or reproduce them.
\end{point}

\begin{reply}
The experiments related to power and energy consumption using traditional measures served to validate by correlation the samples collected using IPMI before its adoption (see previous works). Nevertheless, as requested, we complement the explanation in section V.C:
"For each configuration, samples were collected using IPMI every 1 second. This sampling rate was chosen because the order of magnitude of the applications' mean run time was minutes. Therefore, this rate provides enough samples to measure average power. Additionally, timestamps and the total run time were collected. The total energy spent on each configuration is estimated by first interpolating the power samples using the first-order method, and then integrating this function over time."
\end{reply}

\begin{point}
 The authors’ criticism of the performance counters as model parameters is shallow and does not reflect the state of the art. There are recent comprehensive studies on this subject, which they are unaware of (for example, Fahad, Muhammad, et al. "A comparative study of methods for measurement of energy of computing." Energies 12.11 (2019): 2204).
\end{point}

\begin{reply}
We thank the author for the reference, which we use as an example in the revised text. However, our criticisms about performance counters still holds. In fact we can find the same opinion in many other journal papers as in \cite{Weaver2008, Weaver2013a, Das2019SoK:Security, McGuire2009, Ramos2019AnCounters}. Some of the main points are the following:

%All architectures do not equally support performance counters and their accuracy varies with their number too. Regarding accuracy, that is not a real problem as the measurements can be used in a comparative manner instead of an absolute one
All architectures do not equally support performance counters and their accuracy varies with their number too. Indeed, many models are based on performance counters that do not need to rely on the counter's precise data to provide useful energy prediction. However, generally, they are specific for an architecture of a set of architectures, yet this implies other limitations. The number of counters that we can read simultaneously (generally limited to 5 in hardware) leads to another problem of finding a subset of the counters \cite{Silva-De-Souza2020Containergy-aWorkloads}. Also, claiming the system counters to our driver making no other application would be able to use neither the user for debugging.

That said, our final goal is to provide a generic driver capable of running on most Linux devices, and for that, we believe using performance counters is not the right way to go, at least for now while PMUs are not reliable and standardized. For those reasons, we preferred to use a more generic method, already validated in our previous work \cite{Silva2018Energy-optimalApplications}.
\end{reply}

\bibliographystyle{unsrt}
\bibliography{references.bib}

\end{document}

### 2021-03-23 thesis/test/response_letter.tex
add tex files

% LaTeX rebuttal letter example. 
% 
% Copyright 2019 Friedemann Zenke, fzenke.net

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}

\usepackage{xifthen}
\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}

% command declarations for reviewer points and our responses
\newcommand{\reviewersection}{\stepcounter{reviewer} \bigskip \hrule
                  \section*{Reviewer \thereviewer}}

\newenvironment{point}
   {\refstepcounter{point} \bigskip \noindent {Point \textbf{\arabic{point}} -- } \ }
   {\par }

\newenvironment{action}
   {\medskip \noindent \begin{sf}\textbf{Action}:\ }
   {\medskip \end{sf}}

\newenvironment{reply}
   {\medskip \noindent \begin{sf}\textbf{Reply}:\ }
   {\medskip \end{sf}}

\begin{document}

\section*{Response to the reviewers}
% General intro text goes here
We thank the reviewers for their critical assessment of our work. 
In the following we address their concerns point by point.

% Let's start point-by-point with Reviewer 1
\reviewersection
In this work, the authors propose an analytical model to estimate energy-optimal software configurations and by using DVFS and DPM techniques to decrease the energy consumption of the high-performance computing applications. This paper is well-written. However, it is still can be improved from some perspectives.

% Point one description 
\begin{point}
The Related Work section is too long. Moreoever, the authors fail to concisely indicate that comparing with the existing works, what is the outstanding contribution made by the authors in such a long section. 
\end{point}

% Our reply
\begin{action}
This section was reformulated and split into the theoretical background in section II, and related work in section III where we give a more clear distinction from other studies.
\end{action}

\begin{point}
eq. (2) and eq. (20) should be Eq. (2) and Eq. (20).
\end{point}

\begin{reply}
Thanks for noticing this.
\end{reply}

\begin{action}
Corrected.
\end{action}

\begin{point}
What is the unit of "s" in Eq. (4)? Why "speed" multiplied by "time" is "workload"?
\end{point}

\begin{reply}
The unity of $s$ is instructions/seconds, so we have instructions/seconds*seconds, resulting in instructions, which is something that can represent the workload.
\end{reply}

\begin{action}
The unity of $s$ was added in the text section II.B "…and s is the speed of execution in instructions/seconds…".
\end{action}

\begin{point}
Below Eq. (6), the $"T_{total}"$ seems weird.
\end{point}

\begin{reply}
Indeed, the total should have been subscribed.
\end{reply}

\begin{action}
We reformulated complex nomenclature to simplify the equations.
\end{action}

\begin{point}
Did you see the difference of the "dynamic" in Eq. (7) and in Eq. (8)?
\end{point}

\begin{reply}
They are the same, Eq. (8) is just the definition of it.
\end{reply}

\begin{action}
The Eq. (7) was removed and replaced by the textural explanation "there are three main components of power dissipation,  namely, static power, dynamic power, and leakage power, that accumulated compose the total power draw…" in section IV.A.
\end{action}

\begin{point}
The captions of some FIGUREs are ended by period, some are ended without period.
\end{point}

\begin{reply}
Thanks for noticing this.
\end{reply}

\begin{action}
Corrected.
\end{action}

\begin{point}
FIGURE 11 should be explained more.
\end{point}

\begin{reply}
Indeed, the paragraph that explained the figure was too concise and unclear.
\end{reply}

\begin{action}
This paragraph and the figure caption were reformulated. The following paragraph was added: "By default, operating systems do not implement DPM at the core level, and in HPC, generally, the user explicitly chooses the number of cores to runs his job. To give a better idea of how much DPM at the core level can impact energy consumption, we analyze the choice of the number of cores during one year in the UFRN's HPC center. The result is plotted in figure…" in section V.F.
\end{action}

\begin{point}
The energy model and workload model also can be modeled/estimated by Markov Process. The authors may want to refer to and cite the following article. X. Liu, T. Han and N. Ansari, "Intelligent battery management for cellular networks with hybrid energy supplies," 2016 IEEE Wireless Communications and Networking Conference, Doha, 2016, pp. 1-6.
\end{point}

\begin{reply}
Indeed, we had not considered this type of model  in our related works.
\end{reply}

\begin{action}
We have included this reference in the discussion of the related work.
\end{action}

% Begin a new reviewer section
\reviewersection

This work presents the modeling of 13 parallel applications employed to define energy-optimal configurations for a number of different problem sizes. The results show that approach generates about 10 times less energy overhead for the same level of accuracy.

\begin{point}
A massive number of symbols are introduced in the formation. It will help understanding the optimization by providing a proper nomenclature at the beginning of this paper or somewhere and clearly distinguish control variables, problem parameters and constants.
\end{point}

\begin{action}
All symbols are defined and better explained in the document, as suggested by the reviewer. Abbreviations have been added in the appendix.
\end{action}

\begin{point}
The pseudo codes of algorithms need to be clarified, and the current description is not clear.
\end{point}

\begin{action}
Pseudo codes are defined and better explained in the document as suggested by the reviewer.
\end{action}

\begin{point}
The section of related work is too general, and more studies should be added to comparatively show the contributions of this paper.
\end{point}

\begin{reply}
Thank you for the remark.
\end{reply}

\begin{action}
The related work section has been reformulated and split into an introductory theoretical background, in Section II, and related work, in Section III, where we more clearly distinguish our contribution from other studies.
\end{action}

\begin{point}
Please clarify whether the assumptions of the model are reasonable in reality.
\end{point}

\begin{reply}
As suggested, although the hypotheses are well accepted in the literature, their validation was clarified in the results evaluation section.
\end{reply}

\begin{point}
The authors are suggested to use some real-life datasets (e.g., in the above suggested references) to evaluate the proposed method.
\end{point}

\begin{reply}
We thank the reviewer for the suggestion.
\end{reply}

\begin{action}
Section V "Experimental validation" extends benchmarks to real-life data.
\end{action}

\begin{point}
The authors should discuss in more detail how the algorithm parameters are computed/selected and their dependence on an actual problem. The super-parameter settings for all tested algorithms need more discussions as regards how they affect the results and how sensitive are the results to these settings.
\end{point}

\begin{action}
A discussion on the selection of parameters, and their impact on the model, was incremented in section V.C.
\end{action}

\begin{point}
The complexity of the proposed algorithm should be analyzed.
\end{point}

\begin{reply}
We propose a model based on parameters characterizing any algorithm, rather than a specific algorithm. 
\end{reply}

% Begin a new reviewer section
\reviewersection

This work proposes an analytical model that can be used to estimate energy-optimal software configurations and provide knowledgeable hints to improve DVFS and DPM techniques for single-node HPC applications. However, there are many problems and drawbacks that affect the quality of this paper:

\begin{point}
Please give more details about how DPM optimization problems are solved to obtain both frequency and number of active cores. 
\end{point}

\begin{action}
DPM problems can be modeled in the broadest way, and so can optimization methods. As suggested, we provide an example in section V.F "DVFS and DPM Optimization" showing how our model contributes to the DPM optimization problem with just a search algorithm.
\end{action}

\begin{point}
Many symbols, notations are not well designed, and should be simplified. Single-letters (English or Greek) should be used, e.g., "Ttotal" may be replaced by "bar over T" or big Gamma.
\end{point}

\begin{action}
As suggested, we simplify symbols most of the symbols we could to single letters except the ones already used in the literature, to follow the same notation as the original author.
\end{action}

\begin{point}
Please clearly give the references of many equations, e.g., (12) and (13) if they were proposed by others
\end{point}

\begin{reply}
The eq 12 is proposed by us and that was mentioned in the paragraph. "Thus, the proposed model for one processing core of a multi-core processor is derived by using..." in section IV.A. The eq 13 is a simple deduction that we can do with the described parameters.
\end{reply}

\begin{action}
As requested, we added the citations to clarify if it was proposed by us or not.
\end{action}

\begin{point}
The comparison between the proposed model and SVR is not convincing. There are many advanced prediction methods, e.g., "Dendritic neuron model with effective learning algorithms for classification, approximation and prediction," IEEE Transactions on Neural Networks and Learning Systems, 30(2), pp. 601 - 614, Feb. 2019.
\end{point}

\begin{reply}
Thank you for the remark; the document was updated accordingly. Indeed, we compare it not only to SVR, but to many models, MLP, KNN, Decision Tree for regression, and now we include the reference cited above. Nevertheless, our comparison's primary goal is to show that it can achieve better accuracy with considerably less training and data. All those models mentioned can, in fact, achieve better accuracy but with a lot more data and fine-tuning. This makes it unfeasible to our problem, where to construct an extensive database will end up consuming a massive amount of time and energy.
\end{reply}

\begin{action}
We include the reference cited above and explain in the text why we chose SVR for the comparison in the first paragraph of section V.
\end{action}

\begin{point}
There are many studies proposed to solve similar energy optimization problems, e.g., TTSA: An Effective Scheduling Approach for Delay Bounded Tasks in Hybrid Clouds; Application-Aware Dynamic Fine-Grained Resource Provisioning for Virtualized Cloud Data Centers. Authors could also give some comments on them or add some discussion in the future work on how to apply them to improve the current method.
\end{point}

\begin{reply}
Indeed, the above-mentioned approaches can profit from our model to compute the most effective scheduling. Our model can fit any algorithm that needs a real-world model of application energy.
\end{reply}

\begin{action}
As suggested by the reviewer, those works will be taken into account in future work, as we apply the models to modern DVFS and DPM algorithms.
\end{action}

\begin{point}
After the first sentence in Sec. II, please cite some references, e.g., "Predicting the results of RNA molecular specific hybridization using machine learning," in IEEE/CAA Journal of Automatica Sinica, vol. 6, no. 6, pp. 1384-1396, November 2019; "Forecasting of software reliability using neighborhood fuzzy particle swarm optimization based novel neural network," in IEEE/CAA Journal of Automatica Sinica, vol. 6, no. 6, pp. 1365-1383, November 2019
\end{point}

\begin{action}
As requested, we include these references.
\end{action}

\begin{point}
English presentation needs much improvement, e.g., Abstract, Line -3, "to a standard machine-learning modeling" => "with a standard machine-learning model". Line -2, "about 10x less energy overhead" => "about 10 times less energy overhead".
\end{point}

\begin{action}
We double-checked the grammar.
\end{action}

\end{document}

### 2021-03-23 thesis/ch1/ch1.tex
add tex files

\section{Energy-Optimal Configurations for Single-Node HPC Applications\\
%{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used}
%\thanks{Identify applicable funding agency here. If none, delete this.}
}

% Energy efficiency is a growing concern for modern computing, especially for High-Performance Computing (HPC) due to operational costs and the environmental impact. We propose a methodology to find the best energy-efficient configurations to run single-node applications in an HPC environment using an application-agnostic power model of the architecture and an architecture-aware performance model of the application. The operating frequency and number of active cores are the configuration parameters. This method differs from frequency scaling algorithms that do not have any previous information on the performance behavior of the application or the power consumption profile of the architecture. We characterize the application performance using Support Vector Regression. The power consumption is estimated by modeling the dynamic and static power consumed by CMOS transistors using a computation-intensive program to stress the architecture. The estimated energy-optimal configuration is computed by minimizing the product of the power model and the performance model's outcomes. The results obtained using four PARSEC applications with five different inputs show that the proposed approach was able to find configurations that used about 14$\times$ less energy when compared to the worst case of the default Linux governor for voltage and frequency scaling. When compared to the best case of this governor, the proposed approach saved up to 23\% on energy consumption, with an overall average of 6\% less energy.
% 150 words abstract
Energy efficiency is a growing concern for modern computing, especially for HPC due to operational costs and the environmental impact, considering that processors have an important role in this energy consumption. In this work, we propose a methodology to find energy-optimal frequency and number of active cores to run single-node HPC applications using an application-agnostic power model of the architecture and an architecture-aware performance model of the application. We characterize the application performance using machine learning, specifically the "Support Vector Regression" algorithm. Besides that, the power consumption is estimated by modeling CMOS dynamic and static power without knowledge of the application. So, The energy-optimal configuration is estimated by minimizing the product these two models outcomes, the power model and the performance model. Then, the final model can be used to find better frequency and number of cores to aim energy efficiency application execution. Results were obtained for four PARSEC applications and, with five different inputs shows that the proposed approach used substantially less energy when compared to the DVFS governor, in best cases and worst cases.

\section{Introduction} \label{sec:introduction}
Processors are the main contributor to the power consumption of High-Performance Computing (HPC) servers. They contribute between 20 and 40\% to the total server’s power draw~\cite{Fan2007}. Google's servers showed that during peak utilization processors consumed about 42\% of the overall server’s power consumption~\cite{Barroso2007}. Reducing processor power consumption is an effective approach to reduce the whole system's power consumption.  Therefore, modern processors incorporate several features for power management such as independent processing cores that can be disabled by the operating system~\cite{Rotem2012}, clock gating techniques for reducing the dynamic power dissipation of synchronous circuits~\cite{Srinivasan2015} and Dynamic Voltage and Frequency Scaling (DVFS)~\cite{Mittal2014}.

DVFS has been demonstrated to be a very effective technique for reducing the power consumption of processors \cite{Hackenberg2015, Dzhagaryan2014, Hahnel2012, Basmadjian2012, Travers2015, Miyoshi2002, Anghel2011, Pietri2014}. The technique tries to optimize power consumption by adjusting the frequency according to the current load of the processor. Generally, the frequency scales with the intensity of the load and the voltage scales to the minimum value that enables the selected frequency. Among other aspects, DVFS helps in reducing energy consumption because it allows memory-bounded programs to be executed more efficiently \cite{Spiga2006}. Nonetheless, aspects such as load variability may compromise the effectiveness of DVFS. Another important aspect that is typically not taken into account is the number of processing cores to be used by a parallel program. This choice is left to the user, which often is not trivial as shown in this paper.% the power consumption but increasing the execution time. This way is not guaranteed that the energy consumption will be the lower since energy depends on the power and time, there might be a configuration with higher power consumption that can lead to less energy consumed because of the execution time is lower.

We propose a methodology to find the operating frequency and number of active cores that minimize the total energy used to execute an HPC application on a single shared-memory HPC node.

The methodology uses an application-agnostic power model and an architecture-specific application characterization to model performance. The power model is based on the modeling of Complementary Metal-Oxide-Semiconductor (CMOS) logic in the function of the operating frequency \cite{Sarwar1997}. It models both dynamic and static power. Besides operating frequency, the power model is also parametric to the number of active sockets and the number of active cores per socket.

Performance is modeled by characterizing the application on the target architecture. The idea is to predict the performance of the application at any given configuration. The model takes as inputs the operating frequency, the number of active cores and the input size. The modeling is done using a supervised learning method for regression called Support Vector Regression (SVR)~\cite{Smola2004}.

To find the optimal-energy configurations, the algorithm minimizes the product of outcomes of the power and performance models. This approach was validated on four PARSEC applications~\cite{Bienia2008} and compared to the \emph{Ondemand} governor, which is the default DVFS scheme for the Linux operating system. The results show that the proposed approach was able to find configurations that used about 14$\times$ less energy when compared to the worst case of the Ondemand governor. When compared to the best case of this DVFS scheme, i.e. when the user guesses the optimal number of cores to be used, the proposed approach was able to find configurations that used as much as 18.7\% less energy to execute the target application. The overall average energy saving reached 5.7\% for the proposed approach when compared to the best case and 88.8\% when compared to the worst case.

% TODO: Terminar de detalhar quando fechar a versão final das seções

The rest of this paper is organized as follows. Section~\ref{sec:models} presents the proposed models for power, performance, and energy. The experimental setup and the fitting of the models are described in Section~\ref{sec:experimentalsetup}. In Section~\ref{sec:experimentsresults}, the results of applying the proposed approach to four PARSEC applications are presented. Related works are presented in Section~\ref{sec:relatedwork}. Finally, conclusions are drawn and future work is proposed in Section\ref{sec:conclusion}.

% \section{Theoretical background} \label{Theoretical_background}
% %In the following subsections we present a background on how the DVFS control works and describe how to monitor of power information in real time.

% \subsection{DVFS} \label{DVFS}
% The DVFS implementation complies with the Advanced Configuration and Power Interface (ACPI) \cite{Wikipedia2010} standard which is adopted by operating systems to configure hardware components related to power management. In this standard two states are defined, the C and P states, which are important to the processor frequency and voltage control.

% The C state optimizes the power consumption when the processor is not executing any instructions. This state has several levels. The processor initially starts at C0 level, where the Central Processing Unit (CPU) is fully activated. If the CPU is idle for a specified time, then it moves to the next level, C1, where some CPU functionalities are disabled. After spending some time on the C1 level, the processor passes to the next level where more functionalities are disabled, and so on. ACPI defines four levels, but, manufacturers can define additional ones.

% The P state optimizes the voltage and the frequency during the active execution. It starts at the P0 level, with the highest frequency and voltage possible.  At the P1 level, both the frequency and voltage are decreased, and this goes on until the last level is reached with the lowest possible values. The ACPI standard does not specify the number of the P levels, so this is a manufacturers' choice.

% The P and C states can be controlled by interfaces mediated by drivers available on the operating system. The Linux kernel has many drivers available developed by the CPU manufacturers and the community \cite{Brown2005}. The default driver is the "acpi-cpufreq'' that uses policies implemented by so-called governors that dynamically decide the frequency values. Some of the governors available are Performance, Powersave, Ondemand, Conservative and Userspace. Performance and Powersave are static, and they set the frequency to the maximum and minimum allowed values, respectively. Ondemand and Conservative implementing algorithms to estimate the CPU required capacity and adjust the processor frequency accordingly. Finally, Userspace allows the user to specify the frequency.

% \subsection{IPMI} \label{IPMI}

% The Intelligent Platform Management Interface(IPMI) is a set of interfaces used by HPC system administrators. These allow for out-of-band management of computer systems and platform-status monitoring through local network \cite{November2013}. It can monitor variables and resources such as the system's temperature, voltage, fans and power supplies, with independent sensors attach to the hardware.

% The system consists of the main controller, called the baseboard management controller (BMC), and other management controllers distributed among different system modules, as shown in figure \ref{fig:ipmi_diagram}.

% \begin{figure}[H]
% \centerline{\includegraphics[width=\columnwidth]{figures/IPMI-Block-Diagram.png}}
% \caption{IPMI block diagram, this image was taken from http://pt.wikipedia.org }
% \label{fig:ipmi_diagram}
% \end{figure}

% %The BMC operate independent of the operating system and his access can be through HTTP protocol or using a tool provided by the manufacturer.

\section{Models} \label{sec:models}
In this Section, we present the proposed power and performance models that are used to estimate the minimum-energy consumption configuration.

\subsection{Power Model} \label{sec:powermodel}
Some of the main factors that contribute to the CPU power consumption are the dynamic power consumption, the short-circuit power consumption, and the power loss due to the current leakage of transistors, \cite{Rauber2014, Goel2016, Du2017, Gonzalez1997}. The complexity of the circuits of modern processors makes it very difficult to model their power consumption accurately. A viable approach for modeling the CPU's power draw is to model their building components, which are mainly made out of CMOS logic gates. Thus, modeling the power consumption for one logic gate and multiplying this by the total number of gates reduces the complexity of modeling the internal circuits but still provides the sufficient accuracy needed for making optimization decisions.

%The modern technology to construct logic gates is the CMOS, and t
There are three main components of power dissipation in digital CMOS circuits,
\begin{equation}
P_{total}=P_{static}+P_{leak}+P_{dynamic} \label{eq_totalPower},
\end{equation}
namely, static power $P_{static}$, dynamic power $P_{dynamic}$, and leakage power $P_{leak}$.
According to~\cite{Sarwar1997, Butzen2007}, the dynamic power, and leakage power behavior can be approximated by:
\begin{equation}
P_{dynamic}=CV^2f \quad \text{and} \quad P_{leak} \propto V
\label{eq_draws}
\end{equation}
% and
% \begin{equation}
% \end{equation}
where $C$ is the CMOS capacitance, $V$ the voltage applied to the circuit and $f$ the switching frequency.

Another common approximation is to expect a linear relationship between the voltage and the applied frequency~\cite{Usman2013} $ f \propto V \label{eq_fapoxV} $. The static power in our model represents the energy consumed by other components of the CPU. Thus, the proposed model for one processing core of a multi-core processor is derived by using (\ref{eq_draws}) and (\ref{eq_fapoxV}) to rewrite (\ref{eq_totalPower}) as follows:

\begin{equation}
P_{total}(f)= c_1f^3+c_2f+c_3 \label{eq_fitting},
\end{equation}
where $c_1$, $c_2$, and $c_3$ are the model's parameters. When we include the number of active cores $p$, the estimation of the power consumption of the whole processor becomes:

%done \todo[inline]{You do not say what p is. I guess is processors or active cores?}

\begin{equation}
P_{total}(f,p)= p(c_1f^3+c_2f)+c_3 \label{eq_power_core},
\end{equation}
where the dynamic and leakage power parts are multiplied by the number of cores and the static power that is the consumption of other parts of the CPU remains the same. For systems that have more than one processor sockets, the power cost of enabling each socket can be considered. Adding the number of sockets $s$ to the equation gives the final version of the power model used in this work:
% done \todo[inline]{Is s the number of sockets rather than the power cost??}
\begin{equation}
P_{total}(f,p,s)= p(c_1f^3+c_2f)+c_3+c_4s \label{eq_power_final},
\end{equation}
with $c_4$ being the model parameter for the number of sockets.

\subsection{Performance Model} \label{sec:performancemodel}
The performance model aims to estimate the application's execution time for a given target architecture based on a given operating frequency, number of active cores and input size.

The performance was modeled by sampling the execution time of the application for several combinations of discrete values of frequency, number of active cores and input size. The samples were used as a training set for a Support Vector Regression (SVR); a version of the Support Vector Machine (SVM) algorithm for regression proposed in~\cite{Drucker1997}. Training the SVR means minimize the weights $w$ subject to $|y_i-\langle w,x_i\rangle-b| \leq \varepsilon$.

% \[ \begin{cases} 
%       y_i-\langle w,x_i\rangle-b \leq \varepsilon \\
%       \langle w,x_i\rangle+b-y_i \leq \varepsilon \\
%   \end{cases}
% \]

In our model $x_i$ is a vector with the frequency, the number of active cores and input size, $y_i$ is the execution time measured. $\langle w,x_i\rangle+b-y_i$ is the predicted output time and $\varepsilon$ is a free parameter that serves as a threshold.

\subsection{Energy Model} \label{sec:energymodel}
By combining outcome of the power model described in~\cref{sec:powermodel} and the SVR characterization of the application performance described in \cref{sec:performancemodel}, we can estimate the total energy used by the application as follows:
\begin{equation}
E(f,p,s,N)=P(f,p,s)\times{\rm SVR}(f,p,N) \label{eq_energy},
\end{equation}
where $P(f,p,s)$ is the total power modeled by~(\cref{eq_totalPower}), ${\rm SVM}(f,p,N)$ is the execution time estimated by the SVR characterization of the application, $f$ is the frequency, $p$ is the number of active cores, $s$ is the number of sockets, and $N$ is the input size.

%\todo[inline]{it is not obvious why you can estimate energy consumption rapidly}
With (\ref{eq_energy}), it is possible to calculate energy consumption estimations for every possible configuration. Then, the configuration that minimizes energy consumption for a given input can be selected. It is also possible to apply constraints on the execution time, frequency, and the number of active cores although this is not considered in this work.

%To calculate the minimum of this function any numeric minimization method can be used and the domain to search is well defined by the limitations of the system frequency, number of cores and sockets

\section{Experimental Setup} \label{sec:experimentalsetup}
%As discussed in the previous section, prior calculate the minimal energy, its necessary find the following models: Power model and Performance Model. Therefore, the experiments start obtaining these models of applications on platform chosen.   
In the following subsections we present the software and hardware experimental setup used to validate the proposed approach.

\subsection{Case-Study Applications} \label{sec:casestudyapplication}
% Ok \todo[inline]{no need to list all the areas PARSEC is covering as you only cope with 4 case studies I believe}
Four applications from the PARSEC parallel benchmark suite, version 3.0~\cite{Bienia2008}, were used as case studies. 
%These, cover an ample range of areas such as financial analysis, computer vision, engineering, enterprise storage, animation, similarity search, data mining, machine learning, and media processing, 
This suite focuses on emerging workloads and was designed to be representative of the next generation shared-memory programs for chip-multiprocessors. The four applications used in this work were chosen for being relatively straightforward to devise smaller input sizes from the standard native inputs. 
% , although the benchmark already provides different inputs they division doesn't scale in an exact proportional rate and some of them were too small for our experiments. 
These are Fuidanimate, Raytrace, Swaptions, and Blackscholes.
Since the other provided inputs were meant to evaluate simulated architectures, they are too small for measuring execution time in real machines.
% A short description of each one follows.

% \subsubsection{Blackscholes} calculates the prices for a portfolio of European options analytically using the Black-Scholes partial differential equation. There is no closed-form expression for the Black-Scholes equation and as such it must be computed numerically. The program's inputs are the number of threads, the input file containing the options data, and the output file name. %\todo{what do the input and output file contain?}

% \subsubsection{Fuidanimate} uses an extension of the Smoothed Particle Hydrodynamics (SPH) method to simulate an incompressible fluid for interactive animation purposes. The inputs are the number of threads, the number of frames, and an input file with information of all fluid particles and his proprieties.%\todo{again what does the input file has?}.

% \subsubsection{Raytrace} is a version of the raytracing method that is typically employed by real-time animations such as the ones used in computer games. It is optimized for speed rather than realism. The computational complexity of the algorithm depends on the resolution of the output image and the scene. The inputs used on this applications was the number of threads, the number of frames, a 3D object and the display resolution.

% \subsubsection{Swaptions} Uses the Heath-Jarrow-Morton (HJM) framework to price a portfolio of swaptions. Swaptions employs Monte Carlo (MC) simulation to compute the prices. The input to this program are the number of threads, number of swaptions and the number of trials.

\subsection{Case-Study Architecture} \label{sec:casestudyarchitecture}
% \todo[inline]{is the first sentence really necessary? if yes, NPAD and UFRN needs definition}
%In the experiments performed in this work, we used the compute nodes of the High-Performance Computing Center (NPAD/UFRN)\footnote{NPAD is the name of the High Performance Computing Center (from Portuguese: Núcleo de Processamento de Alto Desempenho) of the Universidade Federal do Rio Grande do Norte (UFRN).}. This HPC nodes consist 
In the experiments performed in this work, we used compute nodes that consists of two Intel Xeon E5-2698 v3 processors with sixteen cores each and two hardware threads for each core. The maximum non-turbo frequency is 2.3GHz, and the total physical memory of the node is 128GB (8$\times$16GB). Turbo frequency and hardware multi-threading were disabled during all experiments. The operating system used is Linux CentOS 6.5, kernel 2.6.32.

The Linux kernel has many drivers available developed by the CPU manufacturers and the community \cite{Brown2005}. The default driver is the "acpi-cpufreq'' that uses policies implemented by so-called governors that dynamically decide the frequency values. Some of the governors available are Performance, Powersave, Ondemand, Conservative and Userspace. Performance and Powersave are static, and they set the frequency to the maximum and minimum allowed values, respectively. Ondemand and Conservative implement algorithms to estimate the CPU required capacity and adjust the processor frequency accordingly. Finally, Userspace allows the user to specify the frequency.

In this work, changing the frequency of the cores was done using the Linux "acpi-cpufreq'' driver. The number of active cores was changed by modifying the appropriate Linux virtual files. Both changes require root privileges. In practice, this approach can be brought into production by allowing the resource manager to perform these changes for the user using pre- and post-scripts for job submissions with energy consumption requirements. We also assume that all processors run at the same frequency.

%To execute the experiments was necessary to change the frequency of cores and, also, turn off the cores not used on each step of execution. This was done by Operating System drivers, as acpi-cpufreq, and the Linux virtual files. On systems where not is possible change the frequency by Operating System, it is necessary to enable the frequency scaling by software on the BIOS.

\subsection{Fitting the Power Model} \label{sec:powerfitting}
To fit the power-model equation, the CPU was stressed up to 100\% with a special benchmark made for this purpose similar to the Linux tool stress. Power information was acquired from the Intelligent Platform Management Interface (IPMI) sensors which measures the energy consumption of the entire system at about one sample per second. IPMI provides information about variables and resources such as the system's temperature, voltage, fans, and power supplies; using independent sensors attach to the hardware.

% The system consists of the main controller, called the baseboard management controller (BMC), and other management controllers distributed among different system modules, as shown in figure \ref{fig:ipmi_diagram}.

% \begin{figure}[H]
% \centerline{\includegraphics[width=\columnwidth]{figures/IPMI-Block-Diagram.png}}
% \caption{IPMI block diagram, this image was taken from http://pt.wikipedia.org }
% \label{fig:ipmi_diagram}
% \end{figure}

% %The BMC operate independent of the operating system and his access can be through HTTP protocol or using a tool provided by the manufacturer.

The power was collected for all combinations of frequency --- starting from 1.2~GHz and increasing by 100 MHz each time until 2.2~GHz is reached, and possible numbers of active cores --- from 1 to 32. Between each test, the CPU was left idle until it cooled down to avoid interference on the next test.

The coefficients of (\ref{eq_power_final}), $c_1$, $c_2$, $c_3$ and $c_4$, were found by performing multi-linear regression on the data collected. The retrieved fitting can be seen on Fig.~\ref{fig:fitting}.

\begin{figure}[H]
\centerline{\includegraphics[width=\columnwidth]{ch1/figures/power.png}}
\caption{Power model fitting. The dots represent real power measurements and the solid lines represents the modeled power.}
\label{fig:fitting}
\end{figure}

The fitted equation for estimating the power in the target architecture was:
\begin{equation} 
P_{total}(f,p,s)=p(0.29f^3+0.97f)+198.59+9.18s, \label{eq:fittedpower}
\end{equation}
where the unit for frequency is GHz.

To validate this model was calculated the absolute percentage error, i.e. the mean of the perceptual error on each point. This metric was chosen because of the significant difference between the smallest and the biggest values and it is calculated as follows:

\begin{equation}
\sum_{i}^{\rm \# samples}{\frac{|y_i-y_{\rm model}|}{y_i}}.
\label{mpe}
\end{equation}

The resulting minimum, mean, and maximum absolute percentage errors were 0.007\%, 0.75\%, and 2.6\% respectively. While the root mean squared error was 2.38 W.

\subsection{Performance Characterization} \label{sec:performancecharacterization}

To characterize an application, we ran it for all different numbers of active cores in the range of $1<=p<=32$, for all the frequencies in the range of $1.2<=f<=2.2$ using 100MHz steps, and for 5 different input sizes.

%depending on the application, the number of cores was different because some applications can only run with the power of 2 and others can only run with an even number of threads. 
% done, please check \todo[inline]{is this empirically chosen then? What is the reasoning of making such choices?not clear why this selection is useful}
The input sizes were chosen in such a way that the average execution time was in the order of minutes. The sampled power information, on every second, was used to calculate the real energy usage. The total time to complete the characterization varied between one and two days, depending on the application.

The SVR model was built using the collected data. A grid search was used to tune the model parameters. In this case, a Radial Base Function (RBF) kernel and the penalty for the wrong term of $10\times10^3$ and gamma 0.5 \cite{scikit-learn}. To train the SVR, the data collected was divided into two parts, 90\% for training and 10\% to test the accuracy.

The model was validated also using cross-validation $k$-fold with $k$ equal to 10, using the Mean Absolute Error (MAE) and Percentage Absolute Error (PAE) as metrics. The average results of the cross-validation can be seen in Table~\ref{tab:svr_evaluation}.

\begin{table}[H]
\centering
\caption{Performance-Model's Cross validation Errors}
\label{tab:svr_evaluation}
\begin{tabular}{|l|l|l|l|l|}
\hline
Application  & MAE & PAE \\ \hline
\hline
Blackscholes & 2.01  & 4.6\% \\ \hline
Fluidanimate & 6.65  & 1.89\% \\ \hline
Raytrace  & 3.77  & 0.87\% \\ \hline
Swaptions & 2.29  & 2.56\% \\ \hline
\end{tabular}
\end{table}

One of the results of the characterization can be seen in Figs.~\ref{fig:svr_time2}.
% ,~\ref{fig:svr_time3},~\ref{fig:svr_time4}, and ~\ref{fig:svr_time1}.

\begin{figure}[H]
	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/time_fluid_4.png}}
    \caption{Fluidanimate's performance model. The dots represent real performance measurements and the solid lines represent the modeled performance for various numbers of active cores and frequencies when running for input size 3.}
	\label{fig:svr_time2}
\end{figure}

%\begin{figure}[H]
%\subfigure[Fluidanimate's performance model. The dots represent real performance measurements and the solid lines represent the modeled performance.]{%
%	\label{fig:svr_time2}
%	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/time_fluid_4.png}}
%}%
%\end{figure}

% \begin{figure}[H]
% 	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/time_rtview_3.png}}
% 	\caption{Raytrace's performance model. The dots represent real performance measurements and the solid lines represent the modeled performance for various numbers of active cores and frequencies when running for input size 3.}
% 	\label{fig:svr_time3}
% \end{figure}

%\begin{figure}[H]
%\subfigure[Raytrace's performance model. The dots represent real performance measurements and the solid lines represent the modeled performance.]{%
%	\label{fig:svr_time3}
%	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/time_rtview_3.png}}
%}%
%\end{figure}

% \begin{figure}[H]
% 	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/time_swap_3.png}}
% 	\caption{Swaptions's performance model. The dots represent real performance measurements and the solid lines represent the modeled performance for various numbers of active cores and frequencies when running for input size 3.}
% 	\label{fig:svr_time4}
% \end{figure}

%\begin{figure}[H]
%\subfigure[Swaptions performance model. The dots represent real performance measurements and the solid lines represent the modeled performance.]{%
%	\label{fig:svr_time4}
%	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/time_swap_3.png}}
%}%
%\end{figure}

% \begin{figure}[H]
% 	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/time_black_4.png}}
%     \caption{Blackscholes's performance model. The dots represent real performance measurements and the solid lines represent the modeled performance for various numbers of active cores and frequencies when running for input size 3.}
% 	\label{fig:svr_time1}
% \end{figure}

%\begin{figure}[H]
%\subfigure[Blackscholes performance model. The dots represent real performance measurements and the solid lines represent the modeled performance.]{%
%	\label{fig:svr_time1}
%	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/time_black_4.png}}
%}%
%\caption{SVR performance modeling plots for four PARSEC applications. The plots present the modeled and the measured performances for various numbers of active cores and frequencies when running for the mid-size input.}
%\end{figure}
%\todo{mid-size input?}

\section{Experimental Results} \label{sec:experimentsresults}
%As mentioned previously, the aim of this work is to obtain the optimal energy configuration on a parallel application running on a homogeneous platform. In general, the actual solutions are using the DVFS techniques to minimize energy consumption. The following shows the results obtained with the proposed approach.
In this Section, we present results for the energy model that we introduced in Section~\ref{sec:models} based on the parameter fitting described in Sections~\ref{sec:powerfitting} and~\ref{sec:performancecharacterization}. First, we compare and comment on the model in contrast with the actual energy measurements. Finally, we evaluate the effectiveness of the proposed approach by comparing it to the Linux default Ondemand DVFS governor.

\subsection{Measured versus Modeled Energy} \label{sec:measuredversusmodeledenergy}
The energy measurements were obtained by integrating the power measurements over the total execution time of the application.
The power measurements were made using the IPMI sensors with a sampling rate of about one sample per second.
% \todo[inline]{move mode details about IPMI here}

%From energy equation (\ref{eq_energy}), running the parsec applications and using the proposed model, it can be observed a compensation of energy consumption, depending on how application scales, between the growth of the number of cores and the associated decrease on execution time due to the parallelization. So, the results shows that the power consumption does not vary very much, as can see on figures 
Figs. \ref{fig:rt_s3}, and~\ref{fig:swap_s3}
%\ref{fig:fluid_s3}, \ref{fig:black_s3}
plot the measured and modeled energy consumption for Raytrace and Swaptions, respectfully, for varying the number of active cores and operating frequency, running with the mid-size input.

% \begin{figure}[H]
% 	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/fluid_4.png}}
%     \caption{Fluidanimate's energy measurements versus modeled energy consumption  varying the number of active cores and operating frequency, running with the input size 3.}
% 	\label{fig:fluid_s3}
% \end{figure}

%\begin{figure}[H]
%\subfigure[Fluidanimate's energy measurements versus modeled energy consumption]{%
%	\label{fig:fluid_s3}
%	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/fluid_4.png}}
%}%
%\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/rtview_3.png}}
    \caption{Raytrace's energy measurements versus modeled energy consumption  varying the number of active cores and operating frequency, running with the input size 3.}
	\label{fig:rt_s3}
\end{figure}

%\begin{figure}[H]
%\subfigure[Raytrace's energy measurements versus modeled energy consumption]{%
%	\label{fig:rt_s3}
%	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/rtview_3.png}}
%}%
%\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/swap_3.png}}
    \caption{Swaptions's energy measurements versus modeled energy consumption  varying the number of active cores and operating frequency, running with the input size 3.}
	\label{fig:swap_s3}
\end{figure}

%\begin{figure}[H]
%\subfigure[Swaptions' energy measurements versus modeled energy consumption]{%
%	\label{fig:swap_s3}
%	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/swap_3.png}}
%}%
%\end{figure}

% \begin{figure}[H]
% 	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/black_4.png}}
%     \caption{Blackscholes's energy measurements versus modeled energy consumption varying the number of active cores and operating frequency, running with the input size 3.}
% 	\label{fig:black_s3}
% \end{figure}

%\begin{figure}[H]
%\subfigure[Blackscholes' energy measurements versus modeled energy consumption]{%
%	\label{fig:black_s3}
%	\centerline{\includegraphics[width=\columnwidth]{ch1/figures/black_4.png}}
%}%
%\caption{Measured and modeled energy consumption for four PARSEC applications for varying the number of active cores and operating frequency, running with the mid-size input.}
%\end{figure}

%Looking at the same frequency, the energy consumption has greater significance when the program runs with less active cores, but, even the difference being not very significant, the total energy of the best configuration always tend to happen on higher frequency because execution time decreases faster than power increases when raising the frequency. This relation can change depending on the power consumption of the system.
In general, for the case-study applications and case-study architecture, the optimal-energy configurations tend to be the ones using the highest frequency, which characterizes a race-to-idle rather than a pace-to-idle optimal behavior~\cite{kim2015racing}. This can be explained by the large static power observed in the considered architecture, evidenced by the large $c_3$ parameter in (\ref{eq_power_final}) that was fitted in (\ref{eq:fittedpower}). With a large static power, using a pace-to-idle strategy, i.e. the use of frequencies lower than the maximum, is expected to be effective only if the sum of the leakage and the dynamic power parcels is larger than the static power parcel. 
Based on the fitted power model, this would never happen, i.e. the sum of leakage and dynamic power is always less than the static power,
\begin{equation*}
p(0.29f^3+0.97f)+9.18s < 198.59,
\label{ineq:race_to_idle}
\end{equation*}
even if we use the maximum number of cores, $p=32$ and $s=2$, and the maximum frequency, $f=2.2$.

Nevertheless, race-to-idle is not always the best strategy for all systems because energy scales with the execution time, which in turn scales inversely with the number of active cores and the operating frequency, and because power scales linearly with the number of cores, but cubic with the frequency. For this reason, our proposal found optimal strategy for Raytrace and Blackscholes that are not the race-to-idle configuration, as we present in the next Subsection.

The optimal number of active cores also depends on the parallel scalability of the application. The more scalable the application, the more cores it requires to minimize energy. A scalable application can increasingly exchange the speedup of more cores with lower frequencies in order to spend less energy. This is because of the linear relationship between power and number of cores and the exponential relationship between power and frequency.

\subsection{Proposed Approach versus Ondemand Linux Governor}
\label{sec:proposedapproach}
We have compared the energy consumption of the four case-study applications using the energy-optimal configurations provided by the proposed approach to the energy consumption resulted by use of the Linux default DVFS governor, Ondemand. Since the governor does not choose the number of active cores, we executed each application using 1, 2, 4, 6, 8, $\cdots$, 28, 30, and 32 cores, accounting for the best and the worst cases of energy consumption. Tables \ref{tab:fluidfreq}, \ref{tab:raytracefreq}, \ref{tab:swapfreq} and \ref{tab:blackfreq} present these results for Fuidanimate, Raytrace, Swaptions, and Blackscholes, respectively. the maximum and minimum savings of the proposed approach in comparison with the Ondemand governor. The savings are calculated as follows: 
\begin{equation}
    \frac{100*({\rm Proposed}-{\rm Ondemand})}{\rm Ondemand}.
\end{equation}

\begin{table}[H]
\caption{Fluidanimate Minimal energy}
\label{tab:fluidfreq}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{l|l|l|l|l|l|l|ll}
Input & \rot{\begin{tabular}[c]{@{}l@{}}Mean Freq. \\ in GHz \\ (\#Cores) \end{tabular}} & \rot{Energy in KJ} & \rot{\begin{tabular}[c]{@{}l@{}}Mean Freq. \\ in GHz \\ (\#Cores) \end{tabular}} & \rot{Energy in KJ} & \rot{\begin{tabular}[c]{@{}l@{}} Freq. \\ in GHz \\ (\#Cores) \end{tabular}} & \rot{Energy in KJ} & \multicolumn{1}{l|}{\rot{Min. Save(\%)}} & \rot{Max. Save(\%)} \\ \hline
1     & 1.85 (32)                                                                        & 4.85               & 2.29 (1)                                                                         & 32.38              & 2.0 (32)                                                                     & 4.15               & \multicolumn{1}{l|}{14.5}                & 87.2                \\ \hline
2     & 1.88  (32)                                                                       & 9.35               & 2.29 (1)                                                                         & 66.77              & 2.0 (32)                                                                     & 7.89               & \multicolumn{1}{l|}{15.7}                & 88.2                \\ \hline
3     & 1.89  (32)                                                                       & 18.82              & 2.30 (1)                                                                         & 135.00             & 2.0 (32)                                                                     & 16.98              & \multicolumn{1}{l|}{9.8}                 & 87.4                \\ \hline
4     & 2.08  (32)                                                                       & 37.80              & 2.30 (1)                                                                         & 272.55             & 2.1 (32)                                                                     & 33.20              & \multicolumn{1}{l|}{12.2}                & 87.8                \\ \hline
5     & 2.00  (32)                                                                       & 76.28              & 2.30 (1)                                                                         & 546.84             & 2.2 (32)                                                                     & 66.83              & \multicolumn{1}{l|}{12.4}                & 87.8                \\ \hline
      & \multicolumn{2}{l|}{Ondemand Min.}                                                                    & \multicolumn{2}{l|}{Ondemand Max.}                                                                    & \multicolumn{2}{l|}{Proposed}                                                                     &                                          &                    
\end{tabular}
}
\end{table}

\begin{table}[H]
\caption{Raytrace Minimal energy}
\label{tab:raytracefreq}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{l|l|l|l|l|l|l|ll}
Input & \rot{\begin{tabular}[c]{@{}l@{}}Mean Freq. \\ in GHz \\ (\#Cores) \end{tabular}} & \rot{Energy in KJ} & \rot{\begin{tabular}[c]{@{}l@{}}Mean Freq. \\ in GHz \\ (\#Cores) \end{tabular}} & \rot{Energy in KJ} & \rot{\begin{tabular}[c]{@{}l@{}} Freq. \\ in GHz \\ (\#Cores) \end{tabular}} & \rot{Energy in KJ} & \multicolumn{1}{l|}{\rot{Save Min.(\%)}} & \rot{Save Max.(\%)} \\ \hline
1     & 1.30 (4)                                                                         & 38.56              & 2.29 (1)                                                                         & 60.29              & 2.2 (6)                                                                      & 37.92              & \multicolumn{1}{l|}{1.7}                 & 37.1                \\ \hline
2     & 1.32 (8)                                                                         & 43.59              & 2.30 (1)                                                                         & 98.11              & 2.2 (10)                                                                     & 39.93              & \multicolumn{1}{l|}{8.4}                 & 59.3                \\ \hline
3     & 1.65 (16)                                                                        & 49.40              & 2.30 (1)                                                                         & 168.82             & 2.2 (14)                                                                     & 45.77              & \multicolumn{1}{l|}{7.4}                 & 72.9                \\ \hline
4     & 1.62 (32)                                                                        & 55.61              & 2.30 (1)                                                                         & 299.83             & 2.2 (22)                                                                     & 52.99              & \multicolumn{1}{l|}{4.7}                 & 82.3                \\ \hline
5     & 1.77 (32)                                                                        & 69.33              & 2.30 (1)                                                                         & 520.34             & 2.2 (26)                                                                     & 67.28              & \multicolumn{1}{l|}{3.0}                 & 87.1                \\ \hline
      & \multicolumn{2}{l|}{Ondemand Min.}                                                                    & \multicolumn{2}{l|}{Ondemand Max.}                                                                    & \multicolumn{2}{l|}{Proposed}                                                                     &                                          &                    
\end{tabular}
}
\end{table}

\begin{table}[H]
\caption{Swaptions Minimal energy}
\label{tab:swapfreq}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{l|l|l|l|l|l|l|ll}
Input & \rot{\begin{tabular}[c]{@{}l@{}}Mean Freq. \\ in GHz \\ (\#Cores) \end{tabular}} & \rot{Energy in KJ} & \rot{\begin{tabular}[c]{@{}l@{}}Mean Freq. \\ in GHz \\ (\#Cores) \end{tabular}} & \rot{Energy in KJ} & \rot{\begin{tabular}[c]{@{}l@{}} Freq. \\ in GHz \\ (\#Cores) \end{tabular}} & \rot{Energy in KJ} & \multicolumn{1}{l|}{\rot{Min. Save(\%)}} & \rot{Max. Save(\%)} \\ \hline
1     & 2.15 (32)                                                                        & 5.88               & 2.29 (1)                                                                         & 80.08              & 2.2 (32)                                                                     & 5.73               & \multicolumn{1}{l|}{2.5}                 & 92.8                \\ \hline
2     & 2.00 (32)                                                                        & 9,21               & 2.30 (1)                                                                         & 106.84             & 2.2 (32)                                                                     & 7,81               & \multicolumn{1}{l|}{15.2}                & 92.7                \\ \hline
3     & 2.22 (32)                                                                        & 10.37              & 2.30 (1)                                                                         & 133.41             & 2.0 (32)                                                                     & 9.90               & \multicolumn{1}{l|}{4.5}                 & 92.6                \\ \hline
4     & 2.02 (32)                                                                        & 14.29              & 2.30 (1)                                                                         & 160.34             & 2.0 (32)                                                                     & 12.33              & \multicolumn{1}{l|}{13.8}                & 92.3                \\ \hline
5     & 2.08 (32)                                                                        & 15.82              & 2.30 (1)                                                                         & 186.39             & 1.9 (32)                                                                     & 14.45              & \multicolumn{1}{l|}{8.7}                 & 92.2                \\ \hline
      & \multicolumn{2}{l|}{Ondemand Min.}                                                                    & \multicolumn{2}{l|}{Ondemand Max.}                                                                    & \multicolumn{2}{l|}{Proposed}                                                                     &                                          &                    
\end{tabular}
}
\end{table}
\begin{table}[H]
\caption{Balckschoels Minimal energy}
\label{tab:blackfreq}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{l|l|l|l|l|l|l|ll}
Input & \rot{\begin{tabular}[c]{@{}l@{}}Mean Freq. \\ in GHz \\ (\#Cores) \end{tabular}} & \rot{Energy in KJ} & \rot{\begin{tabular}[c]{@{}l@{}}Mean Freq. \\ in GHz \\ (\#Cores) \end{tabular}} & \rot{Energy in KJ} & \rot{\begin{tabular}[c]{@{}l@{}} Freq. \\ in GHz \\ (\#Cores) \end{tabular}} & \rot{Energy in KJ} & \multicolumn{1}{l|}{\rot{Min. Save (\%)}} & \rot{Max. Save(\%)} \\ \hline
1     & 1.57 (32)                                                                        & 1.36               & 2.27 (1)                                                                         & 16.35              & 2.2 (30)                                                                     & 1.69               & \multicolumn{1}{l|}{-23.9}                & 89.7                \\ \hline
2     & 2.09 (32)                                                                        & 2.93               & 2.24  (1)                                                                        & 33.16              & 1.8 (32)                                                                     & 3.36               & \multicolumn{1}{l|}{-14.7}                & 89.9                \\ \hline
3     & 1.82 (32)                                                                        & 8.08               & 2.23  (1)                                                                        & 65.97              & 2.2 (30)                                                                     & 6.55               & \multicolumn{1}{l|}{18.9}                 & 90.1                \\ \hline
4     & 2.01 (32)                                                                        & 12.59              & 2.14  (1)                                                                        & 131.85             & 2.2 (26)                                                                     & 13.64              & \multicolumn{1}{l|}{-8.3}                 & 89.7                \\ \hline
5     & 1.97 (32)                                                                        & 25.29              & 1.57  (1)                                                                        & 263.89             & 2.2 (28)                                                                     & 26.52              & \multicolumn{1}{l|}{-4.8}                 & 90.0                \\ \hline
      & \multicolumn{2}{l|}{Ondemand Min.}                                                                    & \multicolumn{2}{l|}{Ondemand Max.}                                                                    & \multicolumn{2}{l|}{Proposed}                                                                     &                                           &                    
\end{tabular}
}
\end{table}

In most cases, the proposed approach obtained better results than the best cases of the Ondemand governor. For Blackscholes, the proposed approach was only better than the Ondemand best case for input number 3. On average, the proposed method was 5.7\% better than the best case of the Ondemand governor.
%This can be explained by the linear speedup characteristic of this application. Thus, it is to be expected that on constant efficiency, the better solution will be with more cores.

%The other results on the other applications show that the proposed approach, even in cases on the bests configurations for DVFS, demonstrate that can obtain energy saving of up 18\%, comparing to the Ondemand governor. 
In all cases, the method proposed here outperformed the worst case of the Ondemand governor. On average, the difference in energy consumption was about 88.8\%, being 92.8\% the maximum difference and 37.1\% the minimum. In general, the energy consumption of the DFVS scheme was larger for smaller numbers of cores. Nonetheless, it was not always the case that the best number of cores for this scheme was the maximum, i.e. 32 cores. Possibly, for architectures with larger number of cores, choosing the exact number the minimizes energy consumption would be less evident.

Fig. \ref{fig:energy_results} shows the behavior of the energy consumption for some tested cases of the Ondemand governor. The presented values are normalized to the energy consumption of the proposed approach.

\begin{figure}[H]
    \centering
    \subfloat[Fluidanimate]{
        \includegraphics[width=\columnwidth]{ch1/figures/fluid.png}
        \label{fig:com_fluid}
    }
	\\
    \subfloat[Raytrace]{
        \includegraphics[width=\columnwidth]{ch1/figures/rtview.png}
        \label{fig:com_rtview}
    }
	\\
    \subfloat[Swaptions]{
        \includegraphics[width=\columnwidth]{ch1/figures/swap.png}
        \label{fig:com_swap}
    }
	\\
    \subfloat[BlackScholes]{
        \includegraphics[width=\columnwidth]{ch1/figures/black.png}
        \label{fig:com_black}
    }
    \caption{Energy consumption of the Ondemand governor for power-of-2 numbers of cores and the proposed approach. The values are relative to the energy of the proposed approach.}
    \label{fig:energy_results}
\end{figure}

%%%% moved to the conclusions.
% A weak of this approach is that the model of performance and power needed to be trained with previously known input sizes. This weakness can be treated with a future solution that can incorporate a prediction of input size during execution of the application.

% Although, the results demonstrate to be very interesting because it is achieved better results for all inputs sizes on 3 out of 4 applications as can see on \ref{fig:com_fluid}, \ref{fig:com_rtview} and \ref{fig:com_swap}. Also for Fluidanimate and Raytrace the mean energy saving compared to the best result of Ondemand was higher then 10\%. On the overall mean, we save 6\% of energy.

\section{Related Work} \label{sec:relatedwork}
DVFS is the most common technique employed to obtain energy savings on multi-core systems. Thus, the technique has been extensively researched with the aim of providing strategies for selecting the optimal voltage and frequency for a specific application and architecture. In \cite{Anghel2011} the authors utilized two algorithms for scaling the frequency of the processors: a human-immune system inspired algorithm to monitor the server's power and performance states; and fuzzy logic based algorithm for changing the server's performance state. \cite{Cochran2011} introduced a scaling method for determining the system's optimal operation points for the number of threads and DVFS settings.

In \cite{DaCosta2015}, an approach that considers instantaneous system activity states was proposed. In this case, the memory and network activity were used to generate a DVFS management setting.

Performance counters have also been used to perform effective DVFS. In \cite{Spiliopoulos2011}, the authors used a Continuous Adaptive DVFS based on a performance model of the processor. The model was based on sampling the hardware's performance counters at regular intervals to predict performance/energy workloads. Base on these predictions appropriate voltage, and frequency settings were selected.

In this work, we introduce a power and a performance model to find energy-optimal operating frequency and number of active cores for applications running on specific multi-core platforms. Our approach does not use the DVFS manager to control the processor voltage and frequency settings. This new approach can obtain better results than DVFS strategies as was shown in \cref{sec:experimentsresults}.

The success obtained from this approach is possible due to the fact that the use previous knowledge of the application's performance on the target architecture can expose sufficiently relevant information, such as parallel speedups, that is harder to guess in runtime techniques based on DVFS.

The use of an application-agnostic power modeling for the target architecture helps to make the technique portable to other applications. That is, to estimate the energy-optimal frequency and number of active cores for a new application, only a performance characterization is needed.
%\todo[inline]{I have edited the related work section. The last two sentence starting from "Our approach..." need to be edited. You need to put why our technique is different and why is better. Saying that our results are better is not wise, with out providing the savings achieved by the other techniques.}

\section{Conclusion and future work} \label{sec:conclusion}
In this paper, we propose a new approach to optimize the energy efficiency of single-node batch HPC applications. In contrast to existing scheduling algorithms, our technique utilizes the application's runtime profile, and a power model of the compute node to predict the optimal frequency and number of cores to be used. This proven effective in reducing the energy consumption of applications.

In contrast to conventional DVFS schemes, out approach chooses not only the operating frequency, but also the number of cores to run an application. DVFS schemes leave to the user the choice of how many cores to use, which becomes harder take with the increase in processor core counts.

Results from four parallel PARSEC applications running on an HPC node with two sixteen-core processors show that the novel approach outperforms the default Linux DVFS scheme on its best case with an average of 5.7\% energy savings. In its worst case, the savings were about 88.8\%, on average. The former represents the case when the user would have an oracle to disclosure the optimal number of cores to run the application. The latter often falls into the single-core usage, which is a common choice of many regular users of batch single-node jobs in HPC.

% In high-performance computing systems shared by multiple users normally on Universities and research centers. The user demands his necessary resources needed for the job. With our approach, the users can choose to finish the job while saving energy and the scheduler will choose the necessary resources. For example, if we assume that the choice of the number of cores varies with a random uniform distribution when compared with our method we could save about 18\%. Another common usage of the system is to use only one core which tends to be the worst case as shown in \ref{fig:energy_results}, in this case, we could save around 88.8\%, depending on the user's profile.

A weakness of the proposed technique is the need for information about the input size of the application before execution. A possible solution would be to use performance counters, present in all modern HPC processors, to guess the input size based on previously trained data.

Future work will improve the proposed energy model by taking into account more relevant information, such as the percentage of CPU utilization. This can enable the identification of different phases of the target program and thus, it will enable more fine-grained changes of the frequency and, perhaps, the number of active cores, to further improve the results presented here.
%Furthermore, since the Ondemand governor is not able to select the number of cores, this choice is left to the user, if the application allows. In contrast, the proposed approach automatically uses the number of cores and frequency to reduce energy consumption.

% \newpage
% \appendix

% This appendix shows the steps necessaries to aim reproduce the experiments of this work.

% \section*{Artefact Description}
% Our artifact provides scripts that can reproduce the experiments in this paper. You can download them at a public repository on GitHub. There is a Python Virtual environment and 4 scripts to use.

% The requirements necessary to reproduce this results is a system with a Linux kernel, super user access configured, the acpi-cpufreq driver and the IPMI system and the PARSEC 3.0 Benchmark installed on system. To setup this environment, you can run:

% \begin{lstlisting}[language=bash,caption={Setup}]
% git clone https://github.com/Teste899/SC18
% cd SC18/
% source sc18/bin/activate$
% \end{lstlisting}

% The scripts available are:
% \begin{enumerate}
% \item{$power\_monitor:$ collect power information using IPMI and a program to stress the CPU to the maximum.}
% \item{$power\_fitting:$ calculates the power model constants of the system and compute the errors using data from the $power\_monitor$.}
% \item{$performance\_monitor:$ makes the complete modeling from the collection of execution times to the calculation of the SVR.}
% \item{$performance\_fitting:$ use the SVR model and the power model to compute the energy and find the minimal energy configuration.}
% \end{enumerate}

% Each script has its own parameters that are detailed using -h option

% \begin{lstlisting}[language=bash,caption={Scripts}]
% python power_monitor.py -h
% python power_fitting.py -h
% python performance_monitor.py -h
% python performance_fitting.py -h
% \end{lstlisting}

% The input arguments used on the applications was:

% \begin{lstlisting}[language=bash,caption={Applications input arguments}]
% Blackscholes
% '_nt_', 'in_625K.txt', 'out_625K.txt'
% '_nt_', 'in_1M.txt', 'out_1M.txt'
% '_nt_', 'in_2M.txt', 'out_2M.txt' 
% '_nt_', 'in_5M.txt', 'out_5M.txt'
% '_nt_', 'in_10M.txt', 'out_10M.txt'

% Fluidanimate
% '_nt_', '250', 'in_500K.fluid'
% '_nt_', '500', 'in_500K.fluid'
% '_nt_', '1000', 'in_500K.fluid'
% '_nt_', '2000', 'in_500K.fluid'
% '_nt_', '4000', 'in_500K.fluid'

% Raytrace
% 'thai_statue.obj','-automove', '-nthreads',
% '_nt_', '-frames', '10000', '-res', '140', '140'
% 'thai_statue.obj','-automove', '-nthreads',
% '_nt_', '-frames', '10000', '-res', '196', '196'
% 'thai_statue.obj','-automove', '-nthreads',
% '_nt_', '-frames', '10000', '-res', '274', '274'
% 'thai_statue.obj','-automove', '-nthreads',
% '_nt_', '-frames', '10000', '-res', '384', '384'
% 'thai_statue.obj','-automove', '-nthreads',
% '_nt_', '-frames', '10000', '-res', '537', '537'

% Swaptions
% '-ns', '64', '-sm', '3000000', '-nt', '_nt_'
% '-ns', '64', '-sm', '4000000', '-nt', '_nt_'
% '-ns', '64', '-sm', '5000000', '-nt', '_nt_'
% '-ns', '64', '-sm', '6000000', '-nt', '_nt_'
% '-ns', '64', '-sm', '7000000', '-nt', '_nt_'
% \end{lstlisting}

% The PARSEC 3.0 Benchmark and its necessary input files are available at http://parsec.cs.princeton.edu/download.htm.

% \section*{Execution order}
% The monitoring scripts must first be run to collect the necessary data for be used by fitting scripts. A execution example of the power script is:

% \begin{lstlisting}[language=bash,caption={Power scripts}]
% python power_monitor.py 1,2,4,8,10,12 \ 
%                         16 60 30 power_model
% python power_fitting.py power_model
% \end{lstlisting}

% This script will collect power information and to perform the model fitting and its validation.

% \begin{lstlisting}[language=bash,caption={Performance scripts}]
% python2.7 performance_monitor.py cpu_load 
% "16,32" 32 "nt_ 5,_nt_ 10" 1 10
% python2.7 performance_fitting.py cpu_load 1
% \end{lstlisting}

% The performance scripts will also collect data of the power consumption and execution time. From the execution time it is going to calculate the SVR estimation and perform a cross validation. The energy equation its also calculated there using the power model previously calculated and all the results can be dynamically observed.

### 2021-03-23 thesis/ch7/ch7.tex
add tex files

\section{Common approaches to the optimization problems}

To build the optimization problem it's necessary to model the system power and application, and controllable variables. Common approaches to the optimization problems are:

Minimize the energy function in function of our control variables, subject to constraints in the control variable.

\begin{equation}
\begin{aligned}
\textrm{min} \quad & P(s_1, s_2, ...)T(s_1,s_2,...)\\
\textrm{subject to} \quad & b_1<s_1<b_2\\
\quad & b_3<s_2<b_4\\
\quad & \vdots\\
\end{aligned}
\end{equation}

Another way is to minimize the total energy, with the constraint of to finish the work.

\begin{equation}
\begin{aligned}
\textrm{min} \quad & \sum{P_it_i}\\
\textrm{subject to} \quad & W_{tot} = \sum w_i\\
\quad & \vdots\\
\end{aligned}
\end{equation}

This kind of problem can be seen in multiple ways, considering an application as the total workload and choosing different speeds for each phase of the application, or could also treat each workload as a different application and create schedulers both CPU and cluster level. In fact the question of on each level of optimizer produces a better result is not well explored in the literature yet. The habitual approach it's to tackle each problem at once and combine the strategies resulting in a chain of schedulers.

The complexity of this problem also varies, depending on the choice of power function it can result in linear programming \cite{Kim2015RacingHeuristics}, quadratic programming \cite{Horyath2008Multi-modeClusters}, until NP-hard problems \cite{Fu2018RaceMinimization}.

\subsection{DVFS optimization} \label{sec:dvfs_optmin}
The effectiveness of the proposed approach for optimization was evaluated with a simple algorithm that finds the optimal frequency and number of active cores from the equation proposed. The results were then compared to the Linux default choices for power management.

With the equation developed \ref{eq:en_final}, it is possible to calculate energy consumption estimations for every possible configuration since there is a finite interval of possible values for the frequency and cores. Then, the configuration that minimizes energy consumption for a given input can be selected. It is also possible to apply constraints on the execution time, frequency, and the number of active cores.

Current HPC managers leave to the user the choice of how many cores to use. Because of that three situations were analyzed with relation to the number of cores:

\begin{enumerate}
	\item Worst choice: number of cores that maximize the total energy consumed
	\item Random choice: energy consumed for a random choice of number of cores
	\item Best choice: number of cores that minimize the total energy consumed (oracle)
\end{enumerate}

The default option for the Linux governor is the Ondemand, and by default, it does not have a DPM control for the number of active cores. Since Ondemand only does DVFS, to allow the comparison, each application was executed with all number of cores available in the system 1 to 32.

Figures \ref{fig:energy_worst_case}, \ref{fig:energy_mean_case} and \ref{fig:energy_best_case}, shows the energy savings with respect to Ondemand in the following way $\frac{Ondemand-Model_{min}}{Ondemand}$ with all three cases described above. The list below shows the savings and losses for each case:

\begin{figure}[ht]
	\centering
	\includegraphics[width=\columnwidth,height=8.5cm]{models/figures/dvfs_cmp_max.png}
	\caption{Energy saving comparison with the proposed model and for the worst case}
	\label{fig:energy_worst_case}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{models/figures/dvfs_cmp_mean.png}
	\caption{Energy saving comparison with the proposed model and for the random case}
	\label{fig:energy_mean_case}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{models/figures/dvfs_cmp_32.png}
	\caption{Energy saving comparison with the proposed model and for the best case}
	\label{fig:energy_best_case}
\end{figure}

The list below shows the savings and losses for each case:

\begin{enumerate}
	\item Worst choice: save 69.88\% on average
	\item Random choice: save 12.04\% on average
	\item Best choice: lost 14.06\% on average
\end{enumerate}

%This can be explained by the linear speedup characteristic from this application. Thus, it is to be expected that on constant efficiency, the better solution will be with more cores.

%The other results on the other applications show that the proposed approach, even in cases on the bests configurations for DVFS, demonstrate that can obtain energy saving of up 18\%, comparing to the Ondemand governor.

%\begin{figure}[H]
%    \centering
%    \includegraphics[width=\columnwidth]{dvfs/completo_black_ondemand_nolimit_2_10.png}
%    \caption{Caption}
%    \label{fig:my_label}
%\end{figure}
%
%In general, for the case-study applications and case-study architecture, the optimal-energy configurations tend to be the ones using the highest frequency, which characterizes a race-to-idle rather than a pace-to-idle optimal behavior~\cite{Kim2015RacingHeuristics}. This can be explained by the large static power observed in the considered architecture, evidenced by the large $c_3$ parameter in (\ref{eq_power_final}) that was fitted in (\ref{eq:fittedpower}). With a large static power, using a pace-to-idle strategy, i.e. the use of frequencies lower than the maximum, is expected to be effective only if the sum of the leakage and the dynamic power parcels is larger than the static power parcel. 
%Based on the fitted power model, this would never happen, i.e. the sum of leakage and dynamic power is always less than the static power,
%\begin{equation*}
%p(0.29f^3+0.97f)+9.18s < 198.59, 
%\end{equation*}
%even if we use the maximum number of cores, $p=32$ and $s=2$, and the maximum frequency, $f=2.2$.
%Nevertheless, race-to-idle was not always the best strategy because energy scales with the execution time, which in turn scales inversely with the number of active cores and the operating frequency, and because power scales linearly with the number of cores, but exponentially with the frequency.
%
%The optimal number of active cores depends on the parallel scalability of the application. The more scalable the application, the more cores it requires to minimize energy. A scalable application can increasingly exchange the speedup of more cores with lower frequencies in order to spend less energy. This is because of the linear relationship between power and number of cores and the exponential relationship between power and frequency.
%
%In all cases, the method proposed here outperformed the worst case of the Ondemand governor. On average, the difference in energy consumption was about 790\%, being 1298\% the maximum difference and 59\% the minimum. In general, the energy consumption of the DFVS scheme was larger for smaller numbers of cores. Nonetheless, it was not always the case that the best number of cores for this scheme was the maximum, i.e. 32 cores. Possibly, for architectures with larger number of cores, choosing the exact number the minimizes energy consumption would be less evident.

Since the number of cores in HPC is requested by the user, the number of CPU requests during one year in UFRN's cluster was analyzed. The result is plotted in the figure \ref{fig:cpu_requests}.
\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{models/figures/cpu_request.png}
	\caption{Number of CPUs requested by the users during 2016-2017}
	\label{fig:cpu_requests}
\end{figure}

The worst choice was always with a single core and is the most common choice of many regular users. The best choice was quite often 32 cores, which is the third most popular choice of the users but it is $72\times$ less than 1 core. This can point a direction of how much energy can be saved by using this technique for DPM or more advanced optimization algorithms that can be derived from our model.

In practice, this approach can be brought into production by allowing the resource manager to perform these changes for the user using pre-scripts and post-scripts for job submissions with energy consumption requirements. This approach is simple to be implemented and can work in almost any Linux system.

\section{Conclusion} \label{sec:conclusion}
This work proposes an energy model based on the frequency and number of cores for a full shared memory system. This model can serve as a base for DVFS and DPM optimization problems that include both frequency and active cores, as well for analysis of the contribution of each parameter (ex: parallelism level) to the energy consumption.

Results from 3 HPC benchmarks running in one cluster demonstrate the potential of the proposed novel model. While consuming less energy than traditional machine learning approaches, it can serve as bases from DVFS and DPM algorithm as shown in the \ref{sec:dvfs_optmin} in an average case saving about 12\% up to 69\%. The previous knowledge of the application's performance can expose sufficiently relevant information, such as parallel speedups, that is harder to guess in run-time techniques based on DVFS.

A weakness of the proposed model is the need for information about the input size of the application that can be complex to derive. A possible solution would be to precisely define what is input size, given an definition in function of common variable for all applications like throughput for example. Future work will demonstrate all the possible analysis that is possible to archive with the equation as long as more advanced DVFS models that can be derived using the equation. For instance identification of different phases of the target program and thus, it will enable more fine-grained changes of the frequency and, perhaps, the number of active cores, to further improve the results presented here.

Another important aspect that is typically not taken into account is the number of processing cores to be used by a parallel program. This choice is left to the user, which often is not trivial as shown in this paper.

%%%% moved to the conclusions.
% A weak of this approach is that the model of performance and power needed to be trained with previously known input sizes. This weakness can be treated with a future solution that can incorporate a prediction of input size during execution of the application.

### 2021-03-23 thesis/experiments/experiments.tex
add tex files

\section{Frequency control} \label {sec: controle_freq}

In modern systems, control over the processor frequency can be done by hardware with independent circuits as well as by software. For that, there is the Advanced Configuration and Power Interface (ACPI) an open standard adopted by operating systems to configure hardware components related to power management.

In the ACPI two important states for the DVFS are defined that optimize the energy consumption. They are "C", which is activated when the processor is not executing any instructions, and "P", which is activated while the processor is operating. These states have several levels and at each level the frequency and voltage are changed.

State P starts at level P0, where frequency and voltage are the maximum possible, then P1, where both decrease, until reaching the last state, Pn, where frequency and voltage are the lowest possible. The change of state depends on the level of utilization of the processor. To stay in each state, the level of processor usage must be within specific limits. After exceeding these limits for a certain time, the state will change to the next state corresponding to that new level of processor usage. The number of possible states depends on each manufacturer.

After an idle time, the processor begins to activate C states, starting with C0 where it is still fully active, then moving to C1, where some features are disabled, up to Cn, where all possible features are disabled. The ACPI standard establishes the functionalities that can be disabled between level C1 and level C3, as seen in Table \ref {tab: states_c}. The other levels are specific to each manufacturer.

\begin {table} [H]
\centering
\begin {tabular} {| l | l | l |}
\hline
Mode & Name & Functionality \\ \hline
C0 & operating state & Active processor \\ \hline
C1 & Halt & Stop executing instructions \\ \hline
C2 & Stop-Clock & Disable the internal clock \\ \hline
C3 & Sleep & Disable cache coherence \\ \hline
\end {tabular}
\caption {C states}
\label {tab: states_c}
\end {table}

In state C, the higher the level the greater the energy savings, but returning to the fully functional level is more difficult. In states P, there is a trade-off between performance and energy savings. Figure \ref {fig: p_state} best illustrates the change of states, in which we can see which parts of the circuits are deactivated in states C, the latency to return to the active state, power consumption and also shows the relationship of states P with often.

\begin {figure} [H]
\centering
\begin{subfigure}[t]{0.5\textwidth}
\centering
\includegraphics[height= 7.5cm]{intro/Imagens/c_states}
\caption {C states}
\end {subfigure}%
~
\begin{subfigure}[t]{0.5\textwidth}
\centering
\includegraphics[height= 7.5cm]{intro/Imagens/p_states}
\caption {States P}
\end {subfigure}
\caption {Illustration of states C and P} {Altered image from \protect \url {https://www.thomas-krenn.com/en/wiki/Processor_P-states_and_C-states}}
\label {fig: p_state}
\end {figure}

% Are states C and P orthogonal, do they operate independently?

For this management, the operating system provides in the user space a way to control the frequency. This work used an operating system that has Linux as its core.

Linux is compatible with several modern architectures and is widely used in servers, smartphones and supercomputers. It was based on the UNIX system which has the philosophy of treating everything on the system as a file, including settings and input and output devices, such as keyboard, mouse and hard drive. Another important feature is that it is modular and parts of the system can be loaded or removed during execution.

On Linux there are several frequency management options \cite {Brown2005}. The main ones are acpi-cpufreq, Intel P-state, AMD powernow. In this work, acpi-cpufreq is used, which is standard and allows direct control of frequency through system files. Acpi-cpufreq is a Linux module that uses implemented policies that dynamically decide the frequency to be used. Some of these policies are:

\begin {itemize}
\item Performance - configured as often as possible
\item Powersave - configured as low as possible
\item Userspace - the user chooses the frequency to be used
\item Ondemand - controls the frequency depending on the processor load. When the load increases the frequency also increases accordingly.
\item Conservative - similar to Ondemand but more smoothly, the frequency increase is continuous instead of jumping.
\end {itemize}

\section {Power consumption monitoring} \label {sec: monitoring}
The Running Average Power Limit (RAPL) and Intelligent Platform Management Interface (IPMI) interfaces were used to measure the power consumed. described in \cite{November2013}.

\subsection {IPMI}
IPMI \cite {November2013} is a set of specifications for autonomous subsystems that provides processor, firmware and operating system independent management and monitoring. The use of IPMI allows system administrators to avoid having to travel to the server location, which is often far away, to perform their tasks. Also, servers are located in places with low temperature and with a lot of noise due to the ventilation system, and one should avoid spending too much time in these places. With remote management, it is possible to turn the system on and off, remotely access the (BIOS) and reinstall the system in case of any serious failure.

\begin {figure} [H]
\centering
\includegraphics [height = 7.5cm] {intro/Imagens/IPMI-Block-Diagram.png}
\caption {IMPI diagram} {Image taken from \protect \url {https://pt.wikipedia.org/wiki/Intelligent_Platform_Management_Interface}, the main components of IPMI and how they communicate are shown}
\label {fig: IPMI}
\end {figure}

Access to the IPMI network can be done using the HTTP protocol or a tool made available by the manufacturer (ipmitool), which also performs access via the network. It is also used to monitor the status of the platform with a set of sensors coupled with system temperatures, voltages, fans and power supplies.

\subsection {RAPL}

Modern Intel microprocessors, based on the SandyBridge architecture, include the RAPL \cite {Rotem2012, Hahnel2012, Hackenberg2015} interface designed to limit the use of energy on a chip while ensuring maximum performance. This interface supports energy measurement capabilities through an integrated circuit that estimates energy use based on a model driven by architectural event counters for all components. It also provides temperature readings and current leak models. Estimates are made available in model-specific registers (MSR), updated in milliseconds. The energy estimates offered by RAPL were validated by Intel, which showed excellent results.

\section{Verifying models hypothesis}

\subsection{Frequency voltage ratio}

\begin{table}[H]
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|}
		\hline
		Freq (GHz) & 2.2   & 2.1   & 2.0   & 1.9   & 1.8   & 1.7   & 1.6   & 1.5   & 1.4   & 1.3   & 1.2   \\ \hline
		V          & 0.77  & 0.76  & 0.75  & 0.74  & 0.73  & 0.72  & 0.71  & 0.70  & 0.69  & 0.68  & 0.67  \\ \hline
		aperf      & 2.199 & 2.099 & 2.000 & 1.899 & 1.799 & 1.699 & 1.599 & 1.500 & 1.397 & 1.297 & 1.200 \\ \hline
	\end{tabular}
\end{table}

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{experiments/figures/freq_volt_rel.png}
	\caption{Frequency voltage relation}
	\label{fig:freq_volt_rel}
\end{figure}

\subsection{Instructions executed}

\begin{table}[H]
	\caption{Cores variation}
	\begin{tabular}{|c|c|c|c|}
		\hline
		Application  & Instructions & Dev.     & Dev. (\%) \\ \hline
		Vip          & 7.97e+11     & 7.16e+06 & 0.00\%    \\ \hline
		Openmc       & 8.17e+07     & 1.65e+04 & 0.02\%    \\ \hline
		Rtview       & 9.91e+12     & 1.55e+09 & 0.02\%    \\ \hline
		X264         & 4.52e+11     & 5.81e+07 & 0.01\%    \\ \hline
		Bodytrack    & 1.86e+12     & 3.95e+10 & 2.13\%    \\ \hline
		Fluidanimate & 2.09e+12     & 8.44e+10 & 4.04\%    \\ \hline
		Xhpl         & 1.14e+08     & 1.24e+05 & 0.11\%    \\ \hline
		Blackschole  & 3.75e+12     & 1.40e+09 & 0.04\%    \\ \hline
		Dedup        & 1.02e+11     & 5.74e+07 & 0.06\%    \\ \hline
		Swapti       & 2.43e+12     & 8.87e+08 & 0.04\%    \\ \hline
		Canneal      & 1.19e+11     & 4.46e+07 & 0.04\%    \\ \hline
		Freqmine     & 1.27e+12     & 4.78e+08 & 0.04\%    \\ \hline
		Ferret       & 4.76e+11     & 7.04e+07 & 0.01\%    \\ \hline
	\end{tabular}
\end{table}

\begin{table}[H]
	\caption{Frequency Variation}
	\begin{tabular}{|c|c|c|c|}
		\hline
		Application  & Instructions & Dev.     & Dev. (\%) \\ \hline
		Vip          & 7.97e+11     & 1.16e+06 & 0.00\%    \\ \hline
		Openmc       & 8.17e+07     & 4.52e+03 & 0.01\%    \\ \hline
		Rtview       & 9.91e+12     & 6.64e+05 & 0.00\%    \\ \hline
		X264         & 4.52e+11     & 1.54e+05 & 0.00\%    \\ \hline
		Bodytrack    & 1.84e+12     & 2.54e+05 & 0.00\%    \\ \hline
		Fluidanimate & 2.38e+12     & 1.70e+09 & 0.07\%    \\ \hline
		Xhpl         & 1.14e+08     & 5.95e+03 & 0.01\%    \\ \hline
		Blackschole  & 3.75e+12     & 4.36e+05 & 0.00\%    \\ \hline
		Dedup        & 1.02e+11     & 8.32e+07 & 0.08\%    \\ \hline
		Swapti       & 2.43e+12     & 1.48e+05 & 0.00\%    \\ \hline
		Canneal      & 1.19e+11     & 3.01e+05 & 0.00\%    \\ \hline
		Freqmine     & 1.27e+12     & 3.70e+08 & 0.03\%    \\ \hline
		Ferret       & 4.76e+11     & 5.63e+07 & 0.01\%    \\ \hline
	\end{tabular}
\end{table}

\subsection{Input size and instructions}

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/Blackschole.png}
	\caption{Caption}
	\label{fig:my_label}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/Canneal.png}
	\caption{Caption}
	\label{fig:my_label}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/Blackschole.png}
	\caption{Caption}
	\label{fig:my_label}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/Canneal.png}
	\caption{Caption}
	\label{fig:my_label}
\end{figure}

\section{Model experimental validation} \label{sec:experimentalvalidation}
In this section, the models presented in~\cref{sec:powermodel} and~\cref{sec:performancemodel} were validated with a benchmark specific for multi-core architectures. Additionally, in order to assess the modeling overhead and accuracy, our proposal was then compared to machine learning approaches. We compared against Support Vector Regression (SVR)~\cite{Smola2004}, Decision Tree \cite{Kitts2006RegressionLecture}, k-nearest neighbors \cite{Altman1992AnRegression}, Multilayer perceptron \cite{Murtagh1991MultilayerRegression}, and some new methods such as Gao et al. \cite{Gao2019DendriticPrediction}. However, SVR was chosen as the most representative because in our tests it performed best without an aggressive fine-tuning.

\subsection{Case-Study Applications} \label{sec:casestudyapplication}
The PARSEC parallel benchmark suite, version 3.0~\cite{Bienia2008}, OpenMC \cite{Romano2015OpenMC:Development} and LINPACK (HPL) \cite{Dongarra1988TheExplanation}, were chosen as case studies. The PARSEC benchmark focused on emerging workloads and was designed to represent the next-generation shared-memory programs for chip-multiprocessors. It covers an ample range of areas such as financial analysis, computer vision, engineering, enterprise storage, animation, similarity search, data mining, machine learning, and media processing. The OpenMC and the LINPACK are two classic HPC programs.

\subsection{Case-Study Architecture} \label{sec:casestudyarchitecture}
The experiments were executed in one computer node equipped with two Intel Xeon E5-2698 v3 processors with sixteen cores each and two hardware threads for each core. The overall view of the architecture is shown in Figure \ref{fig:architecture}. The maximum non-turbo frequency is 2.3GHz, and the total physical memory of the node is 128GB (8$\times$16GB). Turbo frequency and hardware multi-threading were disabled during all experiments. The operating system used was Linux CentOS 6.5, kernel 4.16.

\begin{figure}[htb!]
	\centering
	\includegraphics[width=\columnwidth]{models/figures/architecture.png}
	\caption{Node architecture (the image was made with the lstop application).}
	\label{fig:architecture}
\end{figure}

The Linux kernel has many different policies for power management, depending on the driver. In the default driver, the acpi-cpufreq, the options are:
\begin{itemize}
	\item Powersave
	\item Performance
	\item Ondemand
	\item Conservative
	\item Userspace
\end{itemize}
In this work, the frequency control was performed using the Userspace governor, and the core control was accomplished by modifying the appropriate system files with the default CPU-hotplug driver.

The architecture is equipped with the Intelligent Platform Management Interface (IPMI), a set of interfaces allowing out-of-band management of computer systems and platform-status monitoring via the local network~\cite{Schwenkler2006IntelligentInterface}. It can monitor variables and resources such as the system's temperature, voltage, fans, and power supplies, with independent sensors attached to the hardware.

\subsection{Fitting the Models} \label{sec:fitting}
To find the parameters of the \cref{eq:en_final}, 10 uniformly random configurations of frequencies ($f$), cores ($p$) and inputs ($N$) were chosen from the range $1<=p<=32$, $1.2<=f<=2.2$ and $1<=N<=5$ respectively. The application was executed for each chosen configuration, and the measured  energy and time values were collected. For the input size if we assume that all CPU instructions are executed at approximately the same time, the number of basic operations will be directly correlated with the time. Thus, we can estimate the problem size by looking at the execution time, allowing us to divide a large problem size into several smaller ones knowing their relationship as done in the work of Oliveira ~\cite{Oliveira2018ApplicationCores}. The unity can also vary depending on the definition. In our case for simplicity, we assign numbers from 1 the smallest size to 10 the largest, increasing the problem linearly, that way its also possible to interpolate any input in between these values.

%The application's input was chosen to increase the workload linearly, from smallest to largest, so that most of the input space is covered~\cite{Oliveira2018ApplicationCores}.

For each configuration, samples of the power were collected using IPMI every 1 second. This sampling rate was chosen because the order of magnitude of the mean run time of the applications is minutes. Therefore, this rate provides enough samples to measure average power. Additionally, timestamps and the total run time were collected. The total energy spent on each configuration is estimated by first interpolation the power samples using a first order method and then integrating this function in the time.

From the sampled data, the parameters of the model can be estimated. This results in an optimization problem of finding the parameters that minimize the distance between estimated and the measured values. To solve this minimization problem, non-linear least squares have been applied.

The python library scikit-learning was used to build the SVR model~\cite{PedregosaFandVaroquauxGandGramfortAandMichel2011Scikit-learn:Python}. The SVR was trained using the same data used for parameters estimation of \cref{eq:en_final} with a grid search used to find the best kernel function and the best values for the hyper-parameters penalty for the wrong ($C$) and ($\gamma$). For this data, the best function was the Radial Base Function (RBF), and the hyper-parameters were $C=10^4$ and $\gamma=0.5$.

\subsection{Measured versus Modeled Energy}
\label{sec:measuredversusmodeledenergy}

To validate the proposed energy model, all possible configurations were tested by varying the cores in a range of $1<=p<=32$, the frequency in $1.2<=f<=2.2$ and the input in $1<=N<=5$, which varies from 400 to over 1000 configurations depending on the application as some applications have restrictions on the number of cores. The mean percentage error was computed as the difference between the estimated and measured values according to the following equation:
\begin{equation}
MPE = \frac{1}{N} \sum_i^N \frac{|y_{\rm estimated}-y_{\rm measured}|}{y_{\rm measured}}.
\label{eq:mpe}
\end{equation}

\subsubsection{Frequency x Cores}
Figures \ref{fig:en_eq_black}, \ref{fig:en_eq_canneal} plot the measured and modeled energy consumption for some of the applications modeled. They  show some of the possible shapes that the model can take while varying the number of active cores, operating frequency, and input size.
\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.48\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_cores/completo_black_5.png}}
		\caption{Blackscholes.}
		\label{fig:en_eq_black}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.48\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_cores/completo_canneal_1.png}}
		\caption{Canneal.}
		\label{fig:en_eq_canneal}
	\end{subfigure}
	\caption{Example fit for a specific input size: Blackscholes (a) and Canneal (b).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
	}
\end{figure}
\subsubsection{Frequency x Input}
Figures \ref{fig:en_eq_black}, \ref{fig:en_eq_canneal} plot the measured and modeled energy consumption for some of the applications modeled. They  show some of the possible shapes that the model can take while varying the number of active cores, operating frequency, and input size.
\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.48\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_inps/completo_black_5.png}}
		\caption{Blackscholes.}
		\label{fig:en_eq_black}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.48\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_inps/completo_canneal_1.png}}
		\caption{Canneal.}
		\label{fig:en_eq_canneal}
	\end{subfigure}
	\caption{Example fit for a specific input size: Blackscholes (a) and Canneal (b).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
	}
\end{figure}
\subsubsection{Cores x Input}
Figures \ref{fig:en_eq_black}, \ref{fig:en_eq_canneal} plot the measured and modeled energy consumption for some of the applications modeled. They  show some of the possible shapes that the model can take while varying the number of active cores, operating frequency, and input size.
\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.48\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/cores_inps/completo_black_5.png}}
		\caption{Blackscholes.}
		\label{fig:en_eq_black}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.48\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/cores_inps/completo_canneal_1.png}}
		\caption{Canneal.}
		\label{fig:en_eq_canneal}
	\end{subfigure}
	\caption{Example fit for a specific input size: Blackscholes (a) and Canneal (b).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
	}
\end{figure}

\subsubsection{Validation}
The validation results for each application trained with 10 random configurations are displayed in figure \ref{fig:mpe_svr_eq} and the raw MPE values in  table~\ref{tab:mpe_svr_eq}.
\begin{figure}[htb!]
	\includegraphics[width=\columnwidth]{models/figures/mpe_svr_eq.png}
	\caption{Comparison of the Mean Percentage Error between the proposed model and SVR. "Model mean" and "SVR mean" are the average of all MPE values for all applications.
	}
	\label{fig:mpe_svr_eq}
\end{figure}
\begin{table}[htb!]
	\centering
	\begin{tabular}{|c|c|c|}
		\hline
		Application  & Model & SVR   \\ \hline
		Ferret       & 5.25     & 12.49  \\ \hline
		Raytrace     & 6.36     & 11.95  \\ \hline
		Fluianimate  & 2.44     & 22.90  \\ \hline
		x264         & 8.28     & 15.33  \\ \hline
		Vips         & 7.54     & 10.80  \\ \hline
		Swaptions    & 6.54     & 18.57  \\ \hline
		Canneal      & 3.12     & 6.13   \\ \hline
		Dedup        & 8.85     & 13.70  \\ \hline
		Freqmine     & 2.44     & 3.24   \\ \hline
		Blackscholes & 2.18     & 11.00  \\ \hline
		HPL          & 7.47     & 12.75  \\ \hline
		Bodytrack    & 16.98    & 34.12  \\ \hline
		Openmc       & 11.15    & 24.34  \\ \hline
	\end{tabular}
	\caption{Comparison of the Mean Percentage Error between the proposed model and SVR: raw values.}
	\label{tab:mpe_svr_eq}
\end{table}

Figure \ref{fig:mpe_svr_eq} shows that the proposed model always performed better, with a lower MPE, than SVR when we were limited to 10 training points. This result will be further explored in \cref{sec:overhead} where we compare with different training sizes.

\subsection{Overheads on training} \label{sec:overhead}
It is known that machine learning is data-driven, in that sense the SVR model obtained using only 10 configurations could be improved, but what about the analytical model? 
To answer that question, the proposed model and the SVR were also trained with a varying  number of configurations.
We then compared the MPE and the amount of energy spent to create each model. 
This accuracy-energy trade-off is crucial since the energy overhead when building models defeats the primary goal of saving power when running applications.

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_ferret_4.png}}
		\caption{MPE for Ferret.}
		\label{fig:overhead_ferret}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_vips_4.png}}
		\caption{MPE for Vips.}
		\label{fig:overhead_vips}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_x264_4.png}}
		\caption{MPE for x264.}
		\label{fig:overhead_x264}
	\end{subfigure}
	\caption{MPE case studies: Ferret (a), Vips (b) and x264 (c): model (almost constant) versus SVR (lowest MPE beyond train-size 35 on case (b)).}
	\label{fig:overheadFerretVips}
\end{figure}

Figure \ref{fig:overheadFerretVips} shows the comparisons of MPE and energy spent to create each model for two selected applications. According to the results, the analytical model is found to be very stable, not changing much as more data is added, while the SVR keeps reshaping to adapt to the data. 
The error of the analytical model is almost constant but that of  the SVR, initially very high, drops at some point.
\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.44\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/overall.png}}
		\caption{Average energy spent on all applications during model creation. The two curves are identical because the same data were used to adjust the SVR and the model.}
		\label{fig:overall_overhead}
	\end{subfigure}
	%
	\quad
	\begin{subfigure}[b]{0.44\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/overall_mpe.png}}
		\caption{
			MPE all applications: SVR needs 10 times more data to have an MPE lower than the model.}
		\label{fig:overall_MPE}
	\end{subfigure}
	\caption{Overall results for energy and MPE for each training size.}
	\label{fig:overall_train}
\end{figure}
Figure \ref{fig:overall_train} presents the overall results, with the mean energy overhead and MPE for all applications.
The meeting point of the MPE for the SVR and the proposed model can be extracted from  \ref{fig:overall_MPE}.
There it shows that around 90 configurations the SVR starts to have a smaller error. The cost of that is the linear increase in energy spent on training. The increase in energy, about 10 times more, can be observed in Figure \ref{fig:overall_overhead}.

\section{PMCs}

\section{Energy per instruction}

\begin{lstlisting}
xor rcx, rcx
mov rax, 1
mov rdx, 0

loop:
	targ_inst(*arg)
	targ_inst(*arg)
	targ_inst(*arg)
	targ_inst(*arg)
	targ_inst(*arg)
	targ_inst(*arg)
	targ_inst(*arg)
	targ_inst(*arg)
	targ_inst(*arg)

add rcx, 1
cmp rcx, 9999999
jne loop
\end{lstlisting}

Estimating the energy of the instruction from this benchmark

$E_{total}=9999999(\frac{10}{13}inst+\frac{3}{13}loop)$

$E_{total}=7692306*inst+constant$

\subsection{Generic}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{experiments/figures/inst_en_args_generic.png}
	\caption{Energy per instruction argument}
	\label{fig:experiment_en1}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{experiments/figures/inst_mean_en_generic.png}
	\caption{Mean energy per instruction over all arguments}
	\label{fig:experiment_en2}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{experiments/figures/inst_std_en_generic.png}
	\caption{Standard deviation energy per instruction over all arguments}
	\label{fig:experiment_en3}
\end{figure}

\subsection{SSE}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{experiments/figures/inst_en_args_sse.png}
	\caption{Energy per instruction argument (sse)}
	\label{fig:experiment_en4}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{experiments/figures/inst_mean_en_sse.png}
	\caption{Mean energy per instruction over all arguments (sse)}
	\label{fig:experiment_en5}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{experiments/figures/inst_std_en_sse.png}
	\caption{Standard deviation energy per instruction over all arguments (sse)}
	\label{fig:experiment_en6}
\end{figure}

\section{SSE and Generic}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{experiments/figures/inst_en_cmp_sse_generic.png}
	\caption{Comparing SSE and generic instructions energy consumption}
	\label{fig:experiment_en7}
\end{figure}

\section{Power}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{experiments/figures/inst_pw_generic.png}
	\caption{Instructions power draw}
	\label{fig:experiment_pw1}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{experiments/figures/inst_pw_args.png}
	\caption{Instructions power draw by argument}
	\label{fig:experiment_pw2}
\end{figure}

\section{Tracing application methods}

\begin{outline}
	\1 Linux ptrace
		\2 Breakpoint on each instruction	
	\1 Intel pin
		\2 JIT
		\2 Not open source
	\1 qemu user
		\2 TCG accelerator (transform target to host instruction) similar to JIT
		\2 Partial emulation
\end{outline}

We choose the less intrusive and open-source alternative qemu user, with an small modificaion we are able to trace an application instructions. Generating an table as follows:

\begin{table}[H]
	\centering
	\begin{tabular}{|c|c|c|}
		\hline
		opcode         & mnemonic & args                          \\ \hline
		48 89 e7       & movq     & \%rsp, \%rdi\textbackslash{}n \\ \hline
		48 89 e7       & movq     & \%rsp, \%rdi\textbackslash{}n \\ \hline
		e8 08 0e 00 00 & callq    & 0x4000a24ea0\textbackslash{}n \\ \hline
		55             & pushq    & \%rbp\textbackslash{}n        \\ \hline
		48 89 e5       & movq     & \%rsp, \%rbp\textbackslash{}n \\ \hline
		41 57          & pushq    & \%r15\textbackslash{}n        \\ \hline
		41 56          & pushq    & \%r14\textbackslash{}n        \\ \hline
		41 55          & pushq    & \%r13\textbackslash{}n        \\ \hline
		& \vdots & \\ \hline
	\end{tabular}
\end{table}

\section{CPU load}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=1\columnwidth]{experiments/figures/pw_freq_load.png}
		\caption{Instructions power draw varing CPU load}
		\label{fig:experiment_pw_load}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\centering
		\includegraphics[width=1\columnwidth]{experiments/figures/pw_load.png}
		\caption{Instructions power draw varing CPU load}
		\label{fig:experiment_pw_load}
	\end{subfigure}
\end{figure}

### 2021-03-23 thesis/fingerprint/fingerprint.tex
add tex files

%\author{ \parbox{3 in}{\centering Huibert Kwakernaak*
%         \thanks{*Use the $\backslash$thanks command to put information here}\\
%         Faculty of Electrical Engineering, Mathematics and Computer Science\\
%         University of Twente\\
%         7500 AE Enschede, The Netherlands\\
%         {\tt\small h.kwakernaak@autsubmit.com}}
%         \hspace*{ 0.5 in}
%         \parbox{3 in}{ \centering Pradeep Misra**
%         \thanks{**The footnote marks may be inserted manually}\\
%        Department of Electrical Engineering \\
%         Wright State University\\
%         Dayton, OH 45435, USA\\
%         {\tt\small pmisra@cs.wright.edu}}
%}

\section{An Accurate Tool for Modeling, Fingerprinting, Comparison, and Clustering of Parallel Applications Based on Performance Counters}
The analysis of application performance is essential to better exploit its potential on High-Performance Computing (HPC) architectures. Access to performance counters, available in modern processors, allows collecting key information about program behavior to provide the most appropriate HPC execution strategy.
In this context, we have developed an accurate tool based on performance counters, which facilitates modeling, fingerprinting, behavior comparison and clustering of applications.
It provides a high-level Python API for accessing and configuring performance counters; while execution and counters data gathering is  performed by a C++ module to reduce overhead. 
Indeed, the accuracy of this multiplatform tool was also compared to existing alternatives.  
Key features, such as performance counters collection, post-processing, and comparison, enable fingerprinting of applications, an important step in understanding program behavior for later classification and optimization according to the parameters characterizing the target HPC platform.
For demonstration purposes, the tool was used in the clustering of Polybench applications, a frequently used benchmark set for kernels monitoring. 
%%, a frequently used benchmark for compiler optimization and testing. 
This clustering facilitated the identification of applications with similar and comparable behaviors in terms of input size, data access and transfer, resource utilization and computation, which facilitates the creation of test sets for a given environment based on specific measurement parameters.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{INTRODUCTION}

Hardware Performance Counters are special registers available on most modern processors capable of counting micro-architectural events such as instructions executed, cache-hit, branches miss-predicted, energy estimation and much more.  In new architectures, there are hundreds of hardware events that can be monitored, and more are added to each new generation. 
Performance counters were initially introduced for debugging, but since then they have provided a lot of useful information about running applications without slowing down the execution.
They have been used in several other areas, such as software profiling \cite{Melo2010, Kufrin2005, Knupfer2011}, CPU power modeling \cite{Zamani2012ASystems}, dynamic frequency and voltage scaling, vulnerability research and malware defense \cite{Demme2013OnCounters}.

%\cite{Geimer2010, Geimer2010TheArchitecture, Shende2006TheSystem}

Exploiting Performance Monitoring Units (PMU) requires an intimate knowledge of the micro-architecture and kernel API, as well as an awareness of an ever increasing complexity. 
Otherwise, the measurement performance and accuracy will be seriously affected. Although many tools have been developed using performance counters, programmable interfaces capable of providing good accuracy are still lacking, especially for high-level programming languages. Indeed, apart from PAPI \cite{Weaver2013, Mucci1999} and Perfmon \cite{EranianPerfmon2Interface, Eranian2005TheSpecification} there are only a few APIs allowing access to these counters, and many others are poorly documented, unstable, or designer for a specific purpose.

Performance metrics may have different definitions and programming interfaces on different platforms. 
Therefore, besides gathering information, post-processing modules are also needed. 
Such modules will overcome the lack of precision of counters on some architectures, as indicated in \cite{Weaver2008,Weaver2013a,Das2019SoK:Security}. 
Events that must be precise and deterministic (such as retired instructions) show a variation on run-to-run and over-count on x86\_64 machines, even in strictly controlled environments. These effects are almost always non-intuitive to casual users and pose problems when strict determinism is desirable.

To meet the above mentioned requirements, our strategy combines low-level, efficient and accurate access to PMUs, facilitated by high-level programmable interfaces. This allows the user to perform all configurations and post-processing in Python,  while the underlined architecture details and information gathering are supported by a C++ module, thus preserving accuracy with limited overhead. 
In order to understand the behavior of programs for future comparisons and classification, the proposed tool makes another contribution: the ability to fingerprint and cluster applications.

The definition of reference parameters, such as input size of programs, and performance measures uniformization, allow us to cluster benchmarks such as PolyBench \cite{PolyBench/C3.2}, a collection of numerical computations with static control flow extracted from various application domains, with interesting results. 
In addition to contributing to the standardization of kernel execution and monitoring, this clustering has identified applications with similar and comparable behavior in terms of input size, data transfer and access, resources used and computation; which facilitates the creation of test sets for a given environment, according to specific measurement parameters.

The rest of this article is organized as follows. subsection 2 provides the related work regarding tools available and requirements. subsection 3 presents the performance counters and motivation. subsection 4 presents our clustering and application analysis tool, architecture and components. subsection 5 shows evaluation results: the comparison of existing API and the clustering of the PolyBench benchmarks. Finally, subsection 8 concludes this article.

\subsection{RELATED WORK}

There are only a few APIs allowing access to performance counters.
PAPI \cite{Weaver2013, Mucci1999}, one of the most used libraries for accessing hardware performance counters, was originally developed to provide portable access to the counters found on a diverse collection of modern microprocessors. Rather than learning and writing a new performance infrastructure every time it is ported to a new machine. Measurement code can be written in the PAPI API, which hides the underlying interface.  
PAPI was developed on C and a few non-official libraries were ported to Python. The main problem we found using PAPI was the Python version of has a considerable overhead, it also does not have an easy way to create raw events or low-level control without having to use a special driver. And as our tests will show later, counters sampling over time does not produce good results either.

There are also available a set of interfaces using their own drivers, mainly because the counters are only accessible in kernel mode (ring 0) to control the events for which the counter must be started or stopped. Some events are fixed and others require the development of a dedicated kernel driver.

Perfctr \cite{MikaelPettersson.TheInterface.} supports per-kernel-thread and system-wide monitoring for most major processor architectures. It is distributed as a stand-alone kernel patch. The interface is mostly used by tools built on top of the PAPI performance toolkit.

The Intel VTUNE \cite{IntelCorp.TheAnalyzer.} performance analyzer comes with its own kernel interface, implemented by an open-source driver. The interface supports system-wide monitoring only and is very specific to the needs of the tool.

The problem with the approach of a tool and its own kernel interface is dangerous because, as mentioned on \cite{EranianPerfmon2Interface}, there is clearly code duplication, but more importantly, there is no coordination between the various interfaces that may coexist sharing access to the same PMU resource. To solve this problem, Perfmon2 \cite{Eranian2005TheSpecification} offers a standard interface that all tools can use. Unfortunately, it has not been widely adopted, just supported by a few architectures like the IA64. Instead, Linux comes up with a performance counters subsystem which provides a complete set of configurations.

\subsection{READING PERFORMANCE COUNTERS}

Although hundreds of events are available for monitoring, only a limited number of counters can be used simultaneously.
Therefore, the events to be monitored should be carefully selected and configured using the available counters.

The number of counters available varies between processor architectures, e.g., modern Intel CPUs \cite{Intel2013IntelGuide} support three fixed and four programmable counters per core. Fixed counters monitor events such as Instruction Retired (how many instructions were completely executed), logical cycles, and reference cycles, while programmable counters can be configured to monitor architectural and non-architectural events. 
However, if additional counters are needed, the available ones must be multiplexed.
The configuration of the counters is done by writing in Model-Specific Registers (MSR), only accessible in ring 0 (kernel mode) as indicated previously.

% \begin{itemize}
%     \item rdmsr - Reads the contents of a 64-bit model specific register (MSR) specified in the ECX register into registers EDX:EAX. This instruction must be executed at privilege level 0 or in real-address mode
%     \item rdpmc - Is slightly faster that the equivalent rdmsr instruction. rdpmc can also be configured to allow access to the counters from userspace, without being privileged.
% \end{itemize}

Operating systems provide an abstraction of these hardware capabilities to access counters and MSRs. 
On the Linux system, where our work was developed, there is a performance monitoring subsystem which provides per-task and per CPU counters, counter groups, and related event features. 
All events are seen as 64-bit virtual counters, regardless of the width of the underlying hardware counters. 
They are accessible via special file descriptors, one file descriptor per virtual counter, opened via the perf\_event\_open() system call.  %These system call do not use rdpmc, but rdpmc is not necessarily faster than other methods of reading event values.
Counter events can be processed by interrupt, polling or on time. The interrupt operates by hooking a user-defined function to a specified event, such as a counter overflow, and whenever this event happens, a signal will be generated passing the control to the designated handler function. With polling, whenever an event occurs on the system, the counter value is queued by the operating system and the user can read from this queue using a system call. The last option is to sample over time reading counters every n second.

Ideal hardware performance counters provide exact deterministic results. Real-world PMU implementations do not always live up to this ideal \cite{Weaver2008, Weaver2013a, Das2019SoK:Security, McGuire2009}. Events that should be exact and deterministic (such as the number of executed instructions) show run-to-run variations and over counts on x86 64 machines, even when running in strictly controlled environments. 
These effects are non-intuitive to casual users and cause difficulties when strict determinism is desirable, such as when implementing deterministic replay or deterministic threading libraries. 
Because of that, we have implemented a methodology to reduce the noise and over counts on performance counters.

\subsection{IMPLEMENTATION}

%This module provide a high-level abstraction API to Linux perf events without overhead while executing

This tool is composed of 5 modules the are Profiler, Events, Workload, Analyzer and libpfm4. 
The Workload and libpfm4 module are developed on C/C++ and interfaced with python using Python C API with SWIG (a software development tool to connects programs written in C/C++ with a variety of high-level programming languages). 
The libpfm4 module, developed by \cite{Eranian2008}, is used as an auxiliary library.

%The main features that this library provide is a precise synchronous start the application and counter the event, a low overhead on sampling the events on time and an easy way to find and configure event in groups or standalone.

\subsubsection{ARCHITECTURE}

In figure \ref{fig:achitecture} we can see how these modules interact with each other.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{fingerprint/figures/architecture.png}
    \caption{Modules interconnection}
    \label{fig:achitecture}
\end{figure}

The Profiler is the user interface for configuring and creating events.
It is also responsible for calling the Workload module to run the application and retrieve the data after the execution is complete.

The Workload module, developed on C++, is the core of the library.
It is responsible for creating the application and the sampling process.
It provides a precisely synchronized start, it halts the program before the execution of the first instruction, using the debug interface on Linux (place), and launches the application once the counters have been properly reset and ready to run.
This module, built-in as a Python module using the Python C API, defines a set of functions, macros, and variables to access most aspects of the Python run-time system.
% The workload run simplified code
% \begin{lstlisting}[language=c++]
% reset_counters();
% start_counters();
% start_application();
% while(application.is_running()){
%      sleep(secounds);
%      sample_counters();
% }
% return sampled data;
% \end{lstlisting}
The Event module provides a description of events, events parameters, configurations and PMUs available.

System calls for reading and event creation are done directly by the Workload module or through the libpfm4 Python link. 
The libpfm4 is also used to find events encodings and convert an event name (expressed as a string) to the corresponding event encoding, either as a raw event number (as documented by the hardware vendor) or the OS-specific encoding.
In the latter case, the library is able to prepare the OS-specific data structures needed by the kernel to setup the event.

The Analyzer module is responsible for the post-processing of the data. 
It takes the data from several runs of the application and provides a set of functions to remove outlines, interpolate, filter and compare.
% \subsubsection{FUNCTIONALITY}
% This subsection is going to describe the main functionality of the API.
% Profiler module:
% \begin{lstlisting}[language=Python]
% Profiler(events_groups, program_args=None)
% \end{lstlisting}
% This constructor creates a profiler object from a list of events groups with their names, optionally can pass the application that will be executed.
% \\
% \begin{lstlisting}[language=Python]
% set_program(program_args)
% \end{lstlisting}
% Set the program and his arguments to be executed from a list.
% \\
% \begin{lstlisting}[language=Python]
% start_counters(pid=0)
% \end{lstlisting}
% Monitor events by pid. If pid its set to 0, monitor the entire system.
% \\
% \begin{lstlisting}[language=Python]
% enable_events()
% disable_events()
% reset_events()
% read_events()
% \end{lstlisting}
% Control events
% \\
% \begin{lstlisting}[language=Python]
% run(sample_period, reset_on_sample=False)
% \end{lstlisting}
% Run the application and sample the events on the time, it receives the sample time and if a flat that controls the reset on the sample. It returns the data with the events sampled. 
% \\
% \begin{lstlisting}[language=Python]
% run_background()
% \end{lstlisting}
% Run the application in background
% \begin{lstlisting}[language=Python]
% run_program(...)
% save_data(data, name)
% \end{lstlisting}
% Simple functions to run a program multiple times and store the results in a file.
% \\
% \\
% Events module:
% \begin{lstlisting}[language=Python]
% get_supported_pmus()
% get_supported_events(name)
% get_event_description(name)
% get_event_attrs(name)
% \end{lstlisting}
% These functions return a list with a description for the PMU, event, and attributes.
% \\
% \\
% Analyser module:
% \begin{lstlisting}[language=Python]
% load_data(name)
% \end{lstlisting}
% Load the data capture into an Analyser object.
% \\
% \begin{lstlisting}[language=Python]
% process(verbose=False)
% \end{lstlisting}
% Process the data, using the method described on the post-processing subsection \ref{sec:posprocessing}.
% \\
% \begin{lstlisting}[language=Python]
% compare(a1, a2, feature, npoints)
% \end{lstlisting}
% Compare two objects on a specific feature.

\subsubsection{POST-PROCESSING}
\label{sec:posprocessing}
% With this API we create clusterize the applications of Polybench applications.

With the data collected from multiple runs of the application, the first goal is to obtain a single curve that minimizes the noise caused by the operating system and inaccuracies of the counters. 
In figure \ref{fig:multiple_exec} we can see the raw result of multiple runs of the Polybench 2mm program measuring the number of executed instructions. 
For better visualization, the horizontal axis has been normalized over the interval from 0 to 100. 
This implies that we no longer analyze the program on the time scale, but in the interval in which it took place.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{fingerprint/figures/PERF_COUNT_HW_INSTRUCTIONS.png}
    \caption{Multiple executions of the Polybench 2mm program with different input sizes}
    \label{fig:multiple_exec}
\end{figure}

We apply a median filter to each set of runs, sorting the values and removing edge values.
Then we calculate the average curve, whose final result can be observed on the figure \ref{fig:single_curve}.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{fingerprint/figures/PERF_COUNT_HW_INSTRUCTIONS_1.png}
    \caption{Single instance representation}
    \label{fig:single_curve}
\end{figure}

 After that, we interpolate the curve using the B-spline \cite{Hang2017CubicApplications} algorithm to get the same number of points to all curves. 
 The result of this step can be observed in figure \ref{fig:interporlation}.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{fingerprint/figures/PERF_COUNT_HW_INSTRUCTIONS_2.png}
    \caption{Interpolation}
    \label{fig:interporlation}
\end{figure}

Finally, we use the Savgol filter \cite{Luo2005PropertiesDifferentiators} in order to smooth the data to increase the signal-to-noise ratio without too much distortion. The result is shown in figure \ref{fig:filtering}.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{fingerprint/figures/PERF_COUNT_HW_INSTRUCTIONS_3.png}
    \caption{Filtering}
    \label{fig:filtering}
\end{figure}

\subsection{RESULTS}

In this subsection, we first show the comparison between our tool and others already established, as well as the clustering process.

\subsubsection{ACCURACY COMPARISON}

% \begin{table*}[h]
% \centering
% \caption{Average}
% \begin{tabular}{|c|c|c|c|c|c|c|}
% \hline
% Counter                                 & Pined values & Linux API   & PAPI      & PAPI Python & Perf tool   & MyPerf      \\ \hline
% INSTRUCTIONS\_RETIRED                  & 226990030    & 227000691   & 227000620 & 225901249   & 227000572   & 227000650   \\ \hline
% BRANCH\_INSTRUCTIONS\_RETIRED          & 9240000      & 9250617     & 9250566   & 9239617     & 9250552     & 9250501     \\ \hline
% BR\_INST\_RETIRED:CONDITIONAL          & 8220000      & 8220000     & 8220000   & 8209717     & 8220000     & 8220000     \\ \hline
% MEM\_UOP\_RETIRED:ANY\_LOADS           &              & 2484182672  &           &             & 2484383940  & 2484155029  \\ \hline
% MEM\_UOP\_RETIRED:ANY\_STORES          &              & 189962002   &           &             & 189961539   & 189961321   \\ \hline
% UOPS\_RETIRED:ANY                      &              & 12291082129 &           &             & 12290811038 & 12290901997 \\ \hline
% PARTIAL\_RAT\_STALLS:MUL\_SINGLE\_UOP  &              & 600878      &           &             & 600151      & 600330      \\ \hline
% ARITH:FPU\_DIV                         &              & 5801446     &           &             & 5801000     & 5800977     \\ \hline
% FP\_COMP\_OPS\_EXE:X87                 &              & 48785528    &           &             & 48784834    & 48786021    \\ \hline
% INST\_RETIRED:X87                      &              & 17200008    &           &             & 17200008    & 17200007    \\ \hline
% FP\_COMP\_OPS\_EXE:SSE\_SCALAR\_DOUBLE &              & 5401694     &           &             & 5401842     & 5401679     \\ \hline
% \end{tabular}
% \label{tab:mean}
% \end{table*}

% \begin{table*}[h]
% \centering
% \caption{Standard deviation}
% \begin{tabular}{|c|c|c|c|c|c|}
% \hline
% Counter                                    & Linux API & PAPI & PAPI Python & Perf tool & MyPerf \\ \hline
% INSTRUCTIONS\_RETIRED                  & 396       & 133  & 337763      & 110       & 175    \\ \hline
% BRANCH\_INSTRUCTIONS\_RETIRED          & 297       & 208  & 8485        & 379       & 91     \\ \hline
% BR\_INST\_RETIRED:CONDITIONAL          & 0         & 0    & 3383        & 0         & 0      \\ \hline
% MEM\_UOP\_RETIRED:ANY\_LOADS           & 37399     &      &             & 39217     & 38953  \\ \hline
% MEM\_UOP\_RETIRED:ANY\_STORES          & 1513      &      &             & 1035      & 687    \\ \hline
% UOPS\_RETIRED:ANY                      & 345246    &      &             & 335832    & 333298 \\ \hline
% PARTIAL\_RAT\_STALLS:MUL\_SINGLE\_UOP  & 1222      &      &             & 252       & 521    \\ \hline
% ARITH:FPU\_DIV                         & 1760      &      &             & 1621      & 1544   \\ \hline
% FP\_COMP\_OPS\_EXE:X87                 & 1283      &      &             & 1920      & 3311   \\ \hline
% INST\_RETIRED:X87                      & 4         &      &             & 4         & 3      \\ \hline
% FP\_COMP\_OPS\_EXE:SSE\_SCALAR\_DOUBLE & 1547      &      &             & 3259      & 2097   \\ \hline
% \end{tabular}
% \label{tab:std}
% \end{table*}

\begin{table*}[h]
\centering
\caption{Comparison}
\resizebox{\textwidth}{!}{%
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\hline
\multicolumn{6}{|c|}{Average*$10^{-6}$} & \multicolumn{4}{c|}{Standard deviation}\\ \hline
Counters & \begin{tabular}{c}Pined\\values\end{tabular} & \begin{tabular}{c}Linux\\API\end{tabular} & PAPI & \begin{tabular}{c}PAPI\\Python\end{tabular} & MyPerf  & \begin{tabular}{c}Linux\\API\end{tabular} & PAPI & \begin{tabular}{c}PAPI\\Python\end{tabular} & MyPerf \\ \hline
INSTRUCTIONS\_RETIRED                  & 226.99       & 227       & 227  & 225.9       & 227     & 396       & 133  & 337763      & 175    \\ \hline
BRANCH\_INSTRUCTIONS\_RETIRED          & 9.24         & 9.25      & 9.25 & 9.24        & 9.25    & 297       & 208  & 8485        & 91     \\ \hline
BR\_INST\_RETIRED:CONDITIONAL          & 8.22         & 8.22      & 8.22 & 8.21        & 8.22    & 0         & 0    & 3383        & 0      \\ \hline
MEM\_UOP\_RETIRED:ANY\_LOADS           &              & 2484.18   &      &             & 2484.16 & 37399     &      &             & 38953  \\ \hline
MEM\_UOP\_RETIRED:ANY\_STORES          &              & 189.96    &      &             & 189.96  & 1513      &      &             & 687    \\ \hline
UOPS\_RETIRED:ANY                      &              & 12291.08  &      &             & 12290.9 & 345246    &      &             & 333298 \\ \hline
PARTIAL\_RAT\_STALLS:MUL\_SINGLE\_UOP  &              & 0.6       &      &             & 0.6     & 1222      &      &             & 521    \\ \hline
ARITH:FPU\_DIV                         &              & 5.8       &      &             & 5.8     & 1760      &      &             & 1544   \\ \hline
FP\_COMP\_OPS\_EXE:X87                 &              & 48.79     &      &             & 48.79   & 1283      &      &             & 3311   \\ \hline
INST\_RETIRED:X87                      &              & 17.2      &      &             & 17.2    & 4         &      &             & 3      \\ \hline
FP\_COMP\_OPS\_EXE:SSE\_SCALAR\_DOUBLE &              & 5.4       &      &             & 5.4     & 1547      &      &             & 2097   \\ \hline
\end{tabular}
}
\label{tab:counters}
\end{table*}

To validate the tool, we compared the results of the counters obtained with different APIs. 
We used the hand-crafted assembly benchmark from \cite{Weaver2013a}, designed to test determinism and accuracy of PMUs.
We compared the values obtained from the Linux API, PAPI on C and PAPI on Python. 
The events used for this comparison were instructions retired, branch instructions, memory read, memory load, and arithmetic operations.
We ran the benchmark 30 times and calculated the mean and standard deviation as shown in table \ref{tab:counters}. Some events could not be measured using PAPI because the tool does not accept raw events and there are no equivalent events.

Since the benchmark was hand-crafted with assembly, we know exactly the value for some counter events. 
For this reason, the number of instructions, branch instructions, and conditional branch are pin. However, some other events are architecture-specific and there is no pined value. 
In the latter case, we can still compare to the Linux API, which should be closest to the reality.

The differences using the Linux low-level API, PAPI, and our tool are negligible (the average percentage distance is less than 0.01\% in all the cases). 
As expected, PAPI on Python had the largest difference (with an average distance of 0.25\%) mainly due to an unsynchronized start that resulted in the loss of some instructions at the beginning of the execution. 
This can be an important problem if the application contains a small number of instructions.

The standard deviation of the 30 executions shows that our tool has the smallest variation on a run-to-run on most events. On the contrary, PAPI on Python shows a big variation compared to the others.

\subsubsection{CLUSTERING}

To cluster applications first we need to have a way to compare two programs, for that we define a new variable that tries to compute a fingerprint to the program, this variable has to look similar when we execute the same program with different conditions and inputs. Thinking in the simplest model of a program as a Turing machine as everything can be done with a tap of memory and set of rules we empirically define the variable input size as described on the equation \ref{eq:input_size}. On real computers this is not too far from reality, most of the computations and input and output operations somehow pass through the memory, so analyzing the relationship of the total number of instructions and memory instructions can give us a good fingerprint.

\begin{equation}
    I_{sz} = \frac{I}{I_{m}} \\
    \label{eq:input_size}
\end{equation}

Where $I_{sz}$ we called input size, $I$ the number of instructions executed, $I_m$ number of memory instructions execute. We observe that this variable demonstrate to have the proprieties that we are looking for, produce similar results to the same program with different input size, and environment. This can be used to identify programs but also to see the similarities between different applications.

To compute the distance between two programs we use the Canberra metric \cite{Jurman2009CanberraLists}, described on the equation \ref{eq:canberra}.

\begin{equation}
    d(p,q)= \sum_{i=1}^{n}{\frac{|p_i-q_i|}{|p_i|+|q_i|}}
    \label{eq:canberra}
\end{equation}

Where $p$ and $q$ are n-dimensional vectors.

We compute the input size for all 30 applications of the Polybench with 3 different inputs. Running each program 15 times, with a sampling rate of 0.01 seconds collecting the total number of instruction, number of loads and write instructions, number of floating point operations, besides some software counters. After applying the post-processing to the data collected we compute the input size and perform a Hierarchical clustering using the linkage method of Ward \cite{Murtagh2011WardsAlgorithm}, that minimizes the total within-cluster variance. The results of the clustering can be seen on the figure \ref{fig:sping_force_is} and on the dendrogram on figure \ref{fig:dendograma_input_size}.

%The exact HPCs was:
%PERF_COUNT_HW_INSTRUCTIONS,
%MEM_UOPS_RETIRED:ALL_LOADS,
%MEM_UOPS_RETIRED:ALL_STORES,
%FP_ARITH_INST_RETIRED:SCALAR

\begin{figure*}[h]
    \centering
    \includegraphics[width=\textwidth]{fingerprint/figures/dendograma_input_size.png}
    \caption{Dendrogram}
    \label{fig:dendograma_input_size}
\end{figure*}

From this dendrogram, we can have an idea of how close two applications are. We choose the number of clusters that maximize the number of hits of the same program with different input sizes on the same cluster, in this case, was 5 clusters.

It was observed that as input size grows the program behavior tends to a specific curve, but for small input sizes some have variation, so in some cases, the same program has been classified in more than one cluster. This can also happen if specific parts of the program are triggered with specific inputs, in which case it will also belong to more than one cluster.

In order to have an overall classification of each program, we can pick up the frequency in which each appeared in the clusters and classified it in the cluster in which it appeared more often. In this case, the clusters are:

\begin{itemize}
    \item Cluster 1: 2mm, 3mm, cholesky, correlation, covariance, floyd-warshall, gemm, gramschmidt, lu, ludcmp, nussinov, symm

    \item Cluster 2: deriche, doitgen, syrk

    \item Cluster 3: adi, fdtd-2d, jacobi-2d, syr2k

    \item Cluster 4: atax, bicg, durbin, gemver, gesummv, mvt, trisolv, trmm

    \item Cluster 5: heat-3d, seidel-2d
\end{itemize}

On the figure, \ref{fig:sping_force_is} we display the clusters using the spring force algorithm where each program with a specific input is a node and the edge weight is the Canberra distance. To a better visualization, the name of the inputs was replaced by numbers, the EXTRALARGE is 3, LARGE 2 and MEDIUM 1.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{fingerprint/figures/graph_input_size.png}
    \caption{Spring force of Canberra distance}
    \label{fig:sping_force_is}
\end{figure}

The figure above shows a different way to observe in a reduced space the distance between clusters and applications and how they are organized. From this graph, we see that clustering it's well partitioned and we can clearly separate each cluster. It is also interesting to note that the applications of the cluster formed by the circle symbol are more separated, which may indicate that they were classified in this way because they did not fit into any other cluster.

% \begin{figure*}[t]
%     \begin{subfigure}{\textwidth}
%         \includegraphics[width=\textwidth]{fingerprint/figures/dendograma_input_size.png}
%     \end{subfigure}
%     \\
%     \begin{subfigure}{\textwidth}
%         \includegraphics[width=\textwidth]{fingerprint/figures/dendograma_floating.png}
%     \end{subfigure}
%     %and so on
% \end{figure*}

To have an idea of how the behavior of the variable input size for each clusters is, it was plotted for the applications of clusters 1 and 2 shown on the figures \ref{fig:c1_input_size_0}, \ref{fig:c1_input_size_1}.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{fingerprint/figures/cluster_input_0.png}
    \caption{Input size - Cluster 1}
    \label{fig:c1_input_size_0}
\end{figure}

From these figures, we can have an idea of what behavior was classified as the same class. On cluster 1 on the figure \ref{fig:c1_input_size_0} all applications showed the same behavior with approximately the same amplitude.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{fingerprint/figures/cluster_input_1.png}
    \caption{Input size - Cluster 2}
    \label{fig:c1_input_size_1}
\end{figure}

On figure \ref{fig:c1_input_size_1} we observe that curves that have similar shape regardless of scale on vertical and horizontal axis have also been classified as the same clusters. This is the wanted comportment for the classification because we are only interested in the overall shape of the curve.

% \begin{figure}[H]
%     \centering
%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_input_4.png}
%     \caption{Input size - Cluster 3}
%     \label{fig:c1_input_size_2}
% \end{figure}

% To conclude this study we also clustering applications by its floating point behavior. For that we only use the counter correspondent to floating point operations. The figure \ref{fig:dendogram_fp} shows the dendrogram and the \ref{fig:sping_force_fp} the spring force graph plot.

% \begin{figure*}[h]
%     \includegraphics[width=\textwidth]{fingerprint/figuresdendograma_floating.png}
%     \caption{Dendrogram}
%     \label{fig:dendogram_fp}
% \end{figure*}

% \begin{figure}[H]
%     \centering
%     \includegraphics[width=\textwidth]{fingerprint/figures/graph_floating2.png}
%     \caption{Spring force of Canberra distance}
%     \label{fig:sping_force_fp}
% \end{figure}

% The figures \ref{fig:cluster_fp1},\ref{fig:cluster_fp2},\ref{fig:cluster_fp3} show the behavior on the same cluster for the float point operations. For this classification the number of clusters was 8.

% \begin{figure}[H]
%     \centering
%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_fp_0.png}
%     \caption{Floating point - Cluster 1}
%     \label{fig:cluster_fp1}
% \end{figure}

% \begin{figure}[H]
%     \centering
%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_fp_4.png}
%     \caption{Floating point - Cluster 2}
%     \label{fig:cluster_fp2}
% \end{figure}

% \begin{figure}[H]
%     \centering
%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_fp_5.png}
%     \caption{Floating point - Cluster 3}
%     \label{fig:cluster_fp3}
% \end{figure}

% Here we can also observed again that applications with the same shape were classified as the same clusters.

\subsection{CONCLUSIONS}

%\cite{Processors2012, Gregg2017}
From the results, we observe that the API present overhead similar or lower to other low-level APIs, with the advantage of being in high abstraction and simplified configuration with a few lines of code, is possible to configure and gather counter data.

The tool developed provide also provided a way to fingerprint programs and compute similarities between different programs or the same program with different inputs. This can be useful to reduce applications spaces for benchmarks as was done in Polybench clustering but also to analyze the behavior of a parameter providing insights to the programmer to find a possible bottleneck.

We also provide a precise definition to input size that made possible fingerprint the programs, but it also can help programmers of benchmark applications to better create inputs with more precise growth of a particular parameter.

%\subsection{FUTURE WORK}
%
%We pretend to use this tool to create a data set of applications behavior and automatically classify any program as belonging to a cluster or a set of clusters. The idea is to have a set of clusters that can describe most applications in this way we can know specific behaviors of the applications. This can be applied to various areas such as voltage scaling and frequency of the processor, once knowing the behavior of a particular program we can have an idea of what is the best strategy to control the frequency in order to save energy or increase performance.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{APPENDIX}

% \begin{figure*}[h]
%     \centering
%     \includegraphics[width=\textwidth]{dendograma_input_size.png}
%     \caption{Dendogram input size}
%     \label{fig:dendograma_input_size}
% \end{figure*}

% \begin{figure*}[h]
%     \centering
%     \includegraphics[width=\textwidth]{dendograma_floating.png}
%     \caption{Dendogram floating}
%     \label{fig:dendograma_floating}
% \end{figure*}

This tool and all the analysis are available on github: \url{github.com/VitorRamos/performance_features}

% \begin{lstlisting}[language=bash]
% pip install performance-features
% \end{lstlisting}

%\subsection{ACKNOWLEDGMENT}
%
%This research was developed in the super-computing centers of the University of MONS in collaboration with the center of the Federal University of Rio Grande do Norte.

### 2021-03-23 thesis/intro/Capitulos/ApendiceA.tex
add tex files

% Apêndice
\apendice
\chapter{Primeiro apêndice}

\section{Controle de frequência}

O controle de frequência foi feito utilizando o driver  acpi-cpufreq com o governador userspace. O numero de cores ativos foi alterando modificando os arquivos apropriados nos sistemas de arquivos virtuais do Linux. Na prática, essa abordagem pode ser trazida à produção, permitindo que o gerente de recursos execute essas alterações para o usuário usando scripts pré e pós para envios de tarefas com requisitos de consumo de energia. E para as medições de energia foram utilizados tanto IPMI como RAPL.

Os passos seguidos para isso foram:

\begin{enumerate}

	\item Ativar o controle por hardware na BIOS do sistema utilizando IPMI

	\item Desativar o driver do fabricante (Intel p-state) passando um parâmetro na linha de inicialização do sistema operacional (boot), isso pode ser feito adicionando o comando GRUB\_CMDLINE\_LINUX\_DEFAULT="intel\_pstate=disable"  no arquivo de configuração do grub que é o gerenciador de inicialização "/boot/grub/grub.conf".

	\item Retirar cpufreq da lista negra de drivers, em alguns sistemas o cpufreq está na lista negra para evitar conflito com outros drivers que gerenciam a frequência. Os drivers da lista negra podem ser encontrados em "/etc/modprobe.d"  para retirar basta remover o arquivo com o nome do driver.

	\item Carregar o driver cpufreq, como ele é um módulo podemos utilizar o comando "modprobe acpi-cpufreq" que carrega um módulo no kernel em tempo de execução.

	\item Carregar governadores para cpufreq. Dependendo do sistema nem todos os governadores são carregados precisando serem carregados manualmente. Os governadores disponíveis podem ser encontrados no diretório "/lib/modules/\$(uname -r)/kernel/drivers/cpufreq" e para carregá-los pode ser utilizado o comando "modprobe cpufreq\_governor".

\end{enumerate}

Uma vez o modulo acpi-cpufreq carregado, a alteração de frequência é feita através da interface de arquivos de sistema (sysfs). No diretório

$"/sys/devices/system/cpu/cpu*/cpufreq/"$

encontramos diversos arquivos de controle que comandam diferentes funcionalidades como:

\begin{itemize}
	\item scaling\_governor : Governador atual
	\item scaling\_available\_governors : Lista governadores disponiveis
	\item scaling\_available\_frequencies: Lista frequências disponiveis
	\item scaling\_max\_freq : Configura maior frequência que pode ser utilizada
	\item scaling\_min\_freq : Configura menor frequência que pode ser utilizada
	\item scaling\_setspeed : Configura frequência para governador userspace 
	\item scaling\_cur\_freq : Frequência que o sistema escolheu
	\item cpuinfo\_cur\_freq : Frequência real do sistema
\end{itemize}

Estas variáveis pode ser especificadas para cada núcleo do processador podendo ter diferentes configurações por núcleo.

Para ativar e desativar os núcleos do processador existe o arquivo de controle

$"/sys/devices/system/cpu/cpu*/online".$

\section{Monitoramento}

\subsection{IPMI}
Para fazer o monitoramento com IPMI também foi desenvolvido um módulo python que acessa a interface web do IPMI faz as requisições de dados necessárias e decodifica esses dados.

O primeiro passo foi analisar a comunicação do interface web com o navegador para extrair as requisições necessárias para obter os dados. Primeiramente foi feita requisições manuais do protocolo HTTP, mas o site utiliza de vários javascripts para compor sua interface dificulta análise. Como no sistema utilizado o único acesso é por Secure Shell (ssh) onde tudo é feito pelo terminal de comando, foi feito um tunelamento para redirecionar o conteúdo para outro computador que possui uma interface gráfica.

$ssh -fN -L 8080:service3-bmc:80 admin$

Neste comando todo o tráfego da porta 8080 no localhost será redirecionado para o service3-bmc na porta 80 ( endereço e porta do IPMI web service).

Utilizado o navegador para acessando a interface e as ferramentas de desenvolvedor para analisar o comunicação com site. Foram acessadas as informações de interesse, e foi observado requisições de arquivo XML contendo as informações de consumo de potência. Sabendo qual a requisição necessária o próximo passo foi decodificar os dados do arquivo. Para isso foi feita uma análise do javascript utilizado no site.

\subsection{RAPL}
Para utilizar o RAPL é preciso carregar o module msr  isso pode ser feito utilizando o comando "modprobe msr"

Existem três maneira de ler os valores do RAPL:

Os registradores MSRs podem ser acessados pelo sistema operacional, como o módulo do kernel msr no Linux. Para ler os valores de energia podem ser utilizada a interface do próprio sistema disponível em

$/sys/class/powercap/intel-rapl/intel-rapl:0/energy_uj$

Ou lendo um registrador especifico através do próprio dispositivo em $/dev/msr$.
Além de fornecer informação de energia existem diversas dados sobre parâmetros de performance, por exemplo, acerto de cache, frequência do processador e instruções de branch. Também é possível especificar o limites de potencia.

Todos os códigos utilizados neste trabalho podem ser encontrados no github neste link \url{https://github.com/Teste899/SC18}.

### 2021-03-23 thesis/intro/Capitulos/Conclusao.tex
add tex files

% Considerações finais
\chapter{Conclusão}

Neste trabalho, foi proposta uma nova abordagem para otimizar a eficiência de energia de aplicativos HPC em lote de nó único. Em contraste com os algoritmos de DVFS existentes, nossa técnica utiliza o perfil de tempo de execução do aplicativo e um modelo de energia do nó de cálculo para prever a frequência ideal e o número de núcleos a serem usados. Isso se mostrou eficaz na redução do consumo de energia das aplicações, possivelmente devido ao fato de que o uso do conhecimento prévio do desempenho do aplicativo na arquitetura de destino pode expor informações suficientemente relevantes, como acelerações paralelas, que é mais difícil de encontrar em técnicas de tempo de execução baseadas em DVFS.

O uso de uma modelagem de energia agnóstica do aplicativo para a arquitetura de destino ajuda a tornar a técnica portátil para outros aplicativos. Ou seja, para estimar a frequência ótima de energia e o número de núcleos ativos para uma nova aplicação, apenas uma caracterização de desempenho é necessária.

As principais fraquezas da técnica proposta são, a necessidade de informações sobre o tamanho de entrada do aplicativo antes da execução, o modelo de potencia não leva em consideração a variação ao longo do tempo e o treinamento da SVR consume muita energia.

Para trabalhos futuros, possíveis soluções para os problemas encontrados são, utilizar contadores de desempenho, presentes em todos os processadores HPC modernos, para descobrir o tamanho da entrada com base em dados previamente coletados, incluir a porcentagem de utilização do processador no modelo de potência e encontrar uma formulação matemática para o modelo de desempenho que além de resolver o problema do consumo de energia no treinamento, ainda simplifica a forma de encontrar o mínimo global, podendo ser calculado analiticamente.

### 2021-03-23 thesis/intro/Capitulos/Introducao.tex
add tex files

% Introdução
\chapter{Introdução} \label{cap:introducao}

Consumo de energia é uma preocupação na computação moderna, tanto em dispositivos pessoais, smartphones e notebooks, visando economia da bateria do dispositivo, como em grandes servidores, nuvens e sistemas computacionais de alto desempenho devido aos custos de manutenção e impacto ambiental que eles podem gerar.

Em computação de alto desempenho \abrv[HPC --High Performance Computing]{(HPC)} isso é mais complicado, pois, facilmente um sistema chega a consumir centenas de KiloWatts e sistemas com maior poder computacional podem chegar até a dezenas de MegaWatts \cite{Top500}. Isso pode inviabilizar projetos ou até mesmo descontinuar trabalhos existentes por falta de recursos. Além disso, a próxima geração de sistemas HPC com capacidade de processamento de exaFLOPS, isto é, bilhões de bilhões de cálculos por segundo, terão seu uso restrito pelo seu consumo de energia \cite{Alfonso2013}.

Investir em novas maneiras de reduzir o consumo de potência é crucial para esses sistemas se tornarem viáveis economicamente, além da redução do impacto ambiental causado pelo alto consumo deles. Há muitas pesquisas nesta área, e portanto, diversas formas de reduzir o consumo de energia. Uma das principais é através do consumo de energia do processador. Pesquisas do Google \cite{Fan2007} mostraram que ele é o principal componente responsável pelo consumo de energia em servidores, chegando no pico a 57\% da energia total gasta pelo sistema \cite{Barroso2007}.

Por isso, processadores atuais incluem diversos componentes para minimizar a potência utilizada, como, unidades independentes que podem ser desabilitadas \cite{Rotem2012}, clock gating: técnica utilizada em circuitos síncronos para diminuir o consumo de energia \cite{Srinivasan2015}, Dynamic Voltage and Frequency Scaling \cite{Mittal2014} \abrv[DVFS -- Dynamic Voltage and Frequency Scaling]{(DVFS)}, entre outras.

A técnica de DVFS \cite{Hackenberg2015, Dzhagaryan2014, Hahnel2012, Basmadjian2012, Travers2015, Miyoshi2002, Anghel2011, Pietri2014, Spiga2006} é umas das mais eficientes, pois, o consumo do processador está altamente correlacionado com a sua frequência de operação. Esta consiste em controlar a frequência e a voltagem de acordo com a utilização do dispositivo. Este controle pode ser feito tanto em software através do sistema operacional, quanto em hardware.

Este trabalho propõe uma metodologia para encontrar a frequência operacional e o número de núcleos ativos que minimizam a energia total usada para executar uma aplicação HPC em um único nó com memória compartilhada. Essa metodologia é diferente dos existentes no Linux, pois utiliza uma modelagem do consumo de potência junto com uma caracterização do tempo de execução da aplicação para melhor encontrar a configuração que minimiza a energia total utilizada.

O modelo de potência baseia-se na modelagem da potência do circuito lógico CMOS em função da frequência de operação \cite{Sarwar1997}, incluindo a potência dinâmica, estática e de fuga. Além da frequência de operação, o modelo de energia também é parametrizado para o número de soquetes ativos e de núcleos ativos por soquete.

A modelagem do desempenho é feita pela caracterização da aplicação na arquitetura de destino. A ideia é prever o desempenho da aplicação em qualquer configuração específica. O modelo utiliza a frequência de operação, o número de núcleos ativos e o tamanho da entrada. Essa modelagem utiliza o método de aprendizado supervisionado para regressão chamado Support Vector Regression \abrv[SVR -- Support Vector Regression]{(SVR)} \cite{Ventura2009, Smola2004}.

Para encontrar as configurações de energia ideal, o algoritmo minimiza o produto dos resultados dos modelos de potência e desempenho. Essa abordagem foi validada em quatro aplicações do PARSEC, que é um benchmark para arquiteturas paralelas, \cite{Bienia2008} e comparada a politica \emph{Ondemand}, que é o esquema padrão do DVFS para o sistema operacional Linux e a politica do P-state que é o padrão nos processadores da Intel. Os resultados mostram que a abordagem proposta foi capaz de encontrar configurações que consumiram até 30 \% menos energia quando comparado aos métodos citados anteriormente. A média geral de economia de energia foi de 5,5\%.

%\section{Organização do trabalho}

%No capitulo \ref{cap:referencial_teorico} será dado um referencial teórico das técnicas utilizadas neste trabalho será discutido sobre o controle de frequência na seção \ref{sec:controle_freq} e também sobre o monitoramento de energia nos sistemas em \ref{sec:monitoramento}. No próximo capitulo \ref{cap:modelos} será abordado os modelos de potencia \ref{sec:potencia}, performance  \ref{sec:performance} e energia \ref{sec:energia} detalhando seus desenvolvimentos. A metodologia utilizada para obter os ajustes dos modelos e encontrasse no capitulo \ref{cap:metodologia} junto com a configuração utilizada \ref{sec:config} e as aplicações \ref{sec:apps}. Os resultados estão no capitulo \ref{cap:resultados} onde é mostrado o ajuste dos modelos \ref{sec:ajuste_pw}, \ref{sec:ajuste_perf} e \ref{sec:ajuste_en}. E também onde é feita a comparação \ref{sec:comp}.

### 2021-03-23 thesis/intro/Capitulos/Metodologia.tex
add tex files

% Capítulo 3
\chapter{Metodologia} \label{cap:metodologia}

\section{Configuração} \label{sec:config}
Nos experimentos realizados neste trabalho, foi utilizado um nó de computação que consiste em dois processadores Intel Xeon E5-2698 v3 com dezesseis núcleos e duas threads de hardware para cada núcleo. A frequência máxima não turbo de 2.3 GHz e a memória física total do nó é de 128 GB (8 $ \times $ 16 GB) com o sistema operacional CentOS 6.5 e kernel Linux  versão 4.16. A frequência turbo e o hyper-thread foram desativados durante todos os experimentos.

\section{Aplicações} \label{sec:apps}

O processo de caracterização foi realizado utilizando aplicações do PARSEC benchmark que possui uma suíte de aplicações paralelas que foi projetado para ser representativo da próxima geração de programas de memória compartilhada em chips com múltiplos processadores. As aplicações do PARSEC cobrem as mais diversas áreas, tais como, análise de finanças, visão computacional, mineração de dados, aprendizagem de máquina e processamento de mídia, variando no tipo de paralelização, intensidade de comunicação e tamanho do problema de entrada. Neste trabalho foram utilizadas as aplicações Blackscholes, Fuidanimate, Raytrace e Swaptions da versão 3.0.

\begin{itemize}

	\item Blackscholes: calcula os preços de uma carteira de opções européias analiticamente com a equação diferencial parcial de Black-Scholes. Não há expressão em forma fechada para a equação de Black-Scholes e, como tal, deve ser calculada numericamente. Suas entradas são número de threads, arquivo de entrada contendo os dados das opções e nome do arquivo de saída.

	\item Fuidanimate: usa uma extensão do método de hidrodinâmica de partículas suavizadas (SPH) para simular um fluido incompressível para fins de animação interativa. Suas entradas são o número de threads do número de quadros e um arquivo de entrada com informações de todas as partículas fluidas e suas propriedades.

	\item O Raytrace: uma versão do método de raytracing que normalmente seria empregada para animações em tempo real, como jogos de computador. Está otimizado para a velocidade e não para o realismo. A complexidade computacional do algoritmo depende da resolução da imagem de saída e da cena. As entradas usadas nesses aplicativos são o número de threads, o número de quadros, um objeto 3D e a resolução da tela.

	\item Swaptions: utiliza o framework Heath-Jarrow-Morton para calcular os preços do portifolio swaptions. Swaptions emprega simulações de Monte Carlo para computar os preços. A entrada para este programa é o número de threads, swaptions e tentativas.

\end{itemize}

\section{Ajuste do modelo de potência} \label{sec:met_ajuste_pw}

Para encontrar as constantes, foram desenvolvidas aplicações que utilizam todo o tempo do processador com diversas instruções diferentes, pois, dependendo da instrução executada, pode variar o consumo do processador.

Com essas aplicações foram coletadas amostras de frequência com RAPL e IPMI durante sua execução, variando a frequência de 1.2 GHz até 2.3 GHz, com passos de 100 MHz, e o número de núcleos de 2 até 32, com um passo de 2. A taxa de amostragem foi de aproximadamente 1 segundo e o tempo de execução de 2 a 5 minutos. Entre cada teste foi esperado um tempo de 30 segundos para as próximas medições não serem afetadas pelas anteriores. Com esses dados foi feita uma regressão multi-linear com o método dos mínimos quadrados para descobrir os coeficientes da equação de potência descrito na Seção \ref{sec:potencia}. Que se resume a resolver a equação $\vec{c}=(A^TA)^{-1}A^TB$, onde $\vec{c}$ é o vetor dos coeficientes da equação $A\vec{c}=B$, onde:

\begin{equation}
A=
\begin{bmatrix}
p_1f_1^3 & p_1f_1 & 1 & s_1 \\
p_2f_2^3 & p_1f_2 & 1 & s_1 \\
& \vdots \\
p_nf_m^3 & p_nf_m & 1 & s_2 \\
\end{bmatrix}
\quad e\quad B= 
\begin{bmatrix}
P_1 \\
P_2 \\
\vdots \\
P_k
\end{bmatrix}
\end{equation}

Para cada valor de frequência $f_i$ e número de núcleos $p_i$ associados a potencia $P_i$.

Para validar este modelo foi calculado a média do erro percentual \abrv[MEP -- Média do erro percentual]{(MEP)} e o erro médio absoluto \abrv[MEA -- Média do erro absoluto]{(MEA)}. Essas métricas foram escolhida devido à diferença significativa entre o menor e maior valor. Elas são descritas por (\ref{eq:mpe}).

\begin{equation}
MEP=\frac{1}{n_{amostras}}\sum_{i}^{n_{amostras}}{\frac{|y_i-y_{\rm model}|}{y_i}} \quad MEA=\frac{1}{n_{amostras}}\sum_{i}^{n_{amostras}}{|y_i-y_{\rm model}|}
\label{eq:mpe}
\end{equation}

\section{Treinamento do modelo de performance} \label{sec:met_ajuste_perf}

Para caracterizar a aplicação, foram coletados o tempo de execução para diferentes configurações de números de núcleos ativos dentro do intervalo de $ 1 <= p <= 32 $, e frequência no intervalo de $ 1,2 <= f <= 2,2 $ GHz, com passos de 100 MHz para 5 tamanhos de entradas diferentes, que por sua vez foram escolhidos de forma a dobrar o tamanho do problema a cada entrada, seguindo a equação:

\begin{equation}
N(x)=\frac{M}{2^{p-x}}, 0<x\leq p
\end{equation}

Em que M é o maior problema e p é o numero de divisões. Assim a complexidade resultante do tempo do algoritmo é dado por $O(N(x))$. O tempo médio de execução foi da ordem de minutos. E ao mesmo tempo que era registrado o tempo de execução, também eram coletadas amostras de energia a cada segundo, para futuras comparações com o modelo proposto.

Foi utilizado o método busca em grid para encontrar os melhores parâmetros para o modelo. Ele procura entre vários valores especificados os que apresentam melhores resultados. Neste caso, um kernel de função de base radial \abrv[RBF -- Radial Base Funtion]{(RBF)} dado pela Equação $\phi(r)=e^{-(\gamma r)^2}$ com $\gamma=0.5$. A penalidade por termo errado foi $C=10^4$. A implementação utilizada foi a da biblioteca scikit-learn \cite{scikit-learn}. Para treinar o SVR, os dados coletados foram divididos em duas partes, 90 \% para treinamento e 10 \% para teste. O tempo total para coletar os dados da caracterização variou entre algumas horas e 1 dia, dependendo da aplicação.

% \begin{equation}
% \label{eq:rbf}
% \phi(r)=e^{-(\gamma r)^2}
% \end{equation}

O modelo foi validado com uma validação cruzada do tipo fold com 10 divisões, usando MEA e o MEP como métricas.

### 2021-03-23 thesis/intro/Capitulos/Modelos.tex
add tex files

% Capítulo 4
\chapter{Modelos} \label{cap:modelos}

\section{Potência}  \label{sec:potencia}

Os principais fatores que contribuem para o consumo de energia do processador são consumo dinâmico, por curto-circuito e a perda de energia devido ao vazamento de corrente \cite{Rauber2014, Goel2016, Du2017, Gonzalez1997}, sendo que, a complexidade dos circuitos dos processadores modernos torna muito difícil modelar seu consumo de energia fisicamente acurado. Uma abordagem viável para modelar o consumo de energia do processador é modelar seus componentes principais, que são as portas lógicas. Assim, modelar o consumo de energia para uma porta lógica e multiplicando pelo número total de portas, reduz a complexidade da modelagem e fornecendo precisão suficiente para tomar decisões de otimização, que pode ser observado na seção \ref{cap:resultados}.

A modelagem do consumo de energia do processador utilizada baseia-se no circuito \abrv[CMOS -- Complementary Metal Oxide Semiconductor]{CMOS} \cite{Sarwar1997, Butzen2007} que é uma tecnologia empregada na fabricação de circuitos lógicos. Esta modelagem é uma aproximação que leva em consideração que o processador é basicamente constituído destes circuitos. Sendo assim, o consumo total de energia é proporcional ao número de circuitos lógicos multiplicado pelo consumo de um circuito lógico.

Existem 3 componentes principais para a dissipação de energia nos circuitos CMOS como visto na Equação (\ref{eq_totalPower}).

\begin{equation}
P_{total}=P_{static}+P_{leak}+P_{dynamic} \label{eq_totalPower}
\end{equation}

Onde $P_{total}$ é a potência total dissipada, $P_{dynamic}$ é a dinâmica, devido ao chaveamento do circuito, $P_{static}$ a potência estática e $P_{leak}$ e a potência  perdida devido ao vazamento de corrente no transistor.

Na Figura \ref{fig:CMOS} é mostrado um circuito CMOS inversor que é composto por dois transistores um PMOS e un NMOS, e mais um capacitor.

\begin{figure}[H]
\centering
\includegraphics[height=7.5cm]{Imagens/CMOS.jpg}
\caption{Diagrama eletrico do circuito do CMOS inversor}{Imagem retirada de \protect\url{https://www.hindawi.com/journals/vlsi/2012/505983/}}
\label{fig:CMOS}
\end{figure}

A dissipação de potência dinâmica é causada pela carga do capacitor no circuito durante cada transição de alto para baixo na tensão de entrada, onde parte da energia da fonte é dissipada no PMOS e a outra parte é armazenada no capacitor que é descarregada no NMOS na transição de baixo para alta, na tensão de entrada.

Considerando o CMOS inversor mostrado na Figura \ref{fig:CMOS}, assumindo que a tensão de entrada tem a forma de uma onda quadrada com período de ciclo de trabalho T a energia consumida na transição baixo para alto pode ser derivada calculando a potência média consumida pela fonte o que é dado por (\ref{eq:pwdyn}).

\begin{equation}
P_{dynamic}=\frac{1}{\delta T}\int_{t_1}^{t_2}P(t)dt=\frac{1}{T}\int_{0}^{\infty}i_{vdd}(t)V_{dd}dt \label{eq:pwdyn}
\end{equation}

Sabendo que a corrente em um capacitor é dada por \ref{eq:corrent_cap}.

\begin{equation}
i_{vdd}=C\frac{dV}{dt} \label{eq:corrent_cap}
\end{equation}

Assim temos (\ref{eq:pwdyn_2}).

\begin{equation}
P_{dynamic}=\frac{V_{dd}}{T}\int_{0}^{\infty}C\frac{dV}{dt}dt=\frac{CV_{dd}^2}{T}=CV_{dd}^2f \label{eq:pwdyn_2}
\end{equation}

A corrente vazada no circuito é devido principalmente a diversos fatores como visto em \cite{Butzen2007}, mas como a variação da corrente de vazamento é pequena para simplificar consideraremos que a corrente vazada é praticamente constante simplificando para (\ref{eq:pwleak}):

\begin{equation}
P_{leak}=I_{leak}V_{dd}\alpha V_{dd} \label{eq:pwleak}
\end{equation}

A potência estática é devido a corrente que vai da fonte diretamente para o terra que é aproximadamente constante.

Embora seja possível alterar a frequência e voltagem de forma independente, na prática isso não acontece \cite{Usman2013}, pois dado uma voltagem existe uma frequência máxima em que o circuito pode chegar, para a qual acima deste valor os resultados já não são mais confiáveis. Essa frequência é limitada pelo caminho crítico do circuito, que é a maior distancia entre uma entrada e saída. Também, a frequência nunca é colocada em um valor muito baixo pois resultaria em desperdício de energia. Logo os circuitos tendem a seguir a relação (\ref{eq:fpropv_1}).

\begin{equation}
f \alpha \frac{(V-V_{th})^{\eta}}{V} \label{eq:fpropv_1}
\end{equation}
%\simb[$\alpha$ Proporcional]{$\alpha$}

Onde $V_{th}$ é a voltagem de transição, $\eta$ é uma constante que depende da resposta em frequência do transistor. Para os transistor que não saturam com velocidade podemos considerar $V>>V_{th}$ e aproximar $n=2$ segundo \cite{Gonzalez1997}.

\begin{equation}
f \approx \frac{(V-V_{th})^{\eta}}{V} \approx \frac{V^2}{V} \approx V  \label{eq:fpropv_2}
\end{equation}

Assim o modelo final do processador considerando que ele tem n circuitos lógicos é (\ref{eq:final_1}).

\begin{equation}
P_{total}(f)= (CV_{dd}^2f+I_lV_{dd}+P_s)n \label{eq:final_1}
\end{equation}

Simplificando as constantes e colocando tudo em função da frequência temos (\ref{eq:final_fit}).

\begin{equation}
P_{total}(f)= c_1f^3+c_2f+c_3 \label{eq:final_fit}
\end{equation}

Ampliando este modelo de forma eurística para sistemas multicores com p cores, foi definido (\ref{eq:final_cores}).

\begin{equation}
P_{total}(f,p)= p(c_1f^3+c_2f)+c_3 \label{eq:final_cores}
\end{equation}

Finalmente, considerando sistemas com mais de um processador, onde foi incluído mais um termo estático devido a energia para ativar cada um, resultando em (\ref{eq:final_cores_sock}).

\begin{equation}
P_{total}(f,p,s)= p(c_1f^3+c_2f)+c_3+c_4s \label{eq:final_cores_sock}
\end{equation}

Onde $s$ é o número total de processadores, $p$ é o numero de núcleos, $f$ a frequência e $c_1$,$c_2$,$c_3$ e $c_4$ são constantes.

\section{Performance} \label{sec:performance}

O modelo de performance tem como objetivo estimar o tempo de execução da aplicação a partir da frequência, número de cores e tamanho da entrada. Para isso foi utilizado a SVR como visto na Figura \ref{fig:svr}, que é uma versão da técnica de aprendizagem da máquina supervisionada baseada em vetor de suporte para regressão.

\begin{figure}[H]
\centering
\includegraphics[height=2cm]{Imagens/svr_diagrama.png}
\caption{Diagrama da SVR}
\label{fig:svr}
\end{figure}

A SVR \cite{Drucker1997} consiste em encontrar uma função $f(x)$ de forma que o erro máximo para o alvo seja de $\epsilon$ para todos os pontos do treinamento \cite{Smola2004}. Ela funciona da seguinte forma: Dado o conjunto de treinamento, um vetor de entrada $x_i$ com saída $y_i$, é resolvido o seguinte problema de otimização (\ref{eq:svr_primal}).

\begin{equation}
\label{eq:svr_primal}
\min_{w,\xi_i,\xi_i^*}\frac{1}{2}w^Tw+C\sum_{i}^{n}{\xi_i+\xi_i^*}
\end{equation}
Sujeito a:
\[ 
\begin{cases} 
      y_i-w^T\phi(x_i)-b \leq \varepsilon+\xi_i \\
      w^T\phi(x_i)+b-y_i \leq \varepsilon+\xi_i^* \\
      \xi_i,\xi_i^* \geq 0, i=0,..,n
\end{cases}
\]

Os valores que queremos encontrar são $w$ que é o vetor de suporte. $\xi$ é uma variável de folga que permite que o modelo tenha um erro maior que $\epsilon$ em alguns casos, garantindo a convergência onde existem pontos muito fora da curva e b que é o bias.

Os parâmetros desse modelo são, $\phi(x)$, $\epsilon$ e $C$. O $\phi(x)$ é uma função kernel que mapeia a entrada em outro espaço de maior dimensão permitindo que o algoritmo se ajuste ao hiperplano de margem máxima em um espaço transformado. Isso permite a não linearidade desse algoritmo ao utilizar uma função kernel não linear. As funções kernel mais comuns são polinomial, função de base radial de Gauss (RBF) e tangente hiperbólica. O $\epsilon$ é o desvio máximo para o alvo $y_i$. A constante $C$ determina a tolerância para erros maiores que $\epsilon$.

A resolução deste problema de otimização pode ser feita eficientemente em sua forma dual de Lagrange, que é obtida através da função objetivo de Lagrange utilizando multiplicadores não negativos. Ao adicionar as restrições de otimização e resolver para as variáveis do problema original que minimiza esta função ficando com as variáveis do problema original em função dos multiplicadores de Lagrange, que são chamados de variáveis duais, assim o novo problema é maximizar a função objetivo em função das variáveis duais incluindo as restrições de não negatividade. O novo problema se torna:

\begin{equation}
\label{eq:svr_dual}
\max_{\alpha,\alpha^*}{-\frac{1}{2}\sum_{i,j}{(\alpha_i-\alpha_i^*)(\alpha_j-\alpha_j^*)k(x_i,x_j)}-\epsilon \sum_{i}{(\alpha_i-\alpha_i^*)}+\sum_{i}{y_i(\alpha_i-\alpha_i^*)}}
\end{equation}
Sujeito a:
\[
\begin{cases} 
      \sum_{i}{(\alpha_i-\alpha_i^*)}=0\\
      \alpha_i,\alpha_i^* \in [0,C] \\
   \end{cases}
\]

Onde $k(x,x')=\langle \phi(x), \phi(x') \rangle$ e $\alpha_i,\alpha_i^*$ são os multiplicadores de Lagrange e $f(x)$,$w$ e $b$ são:

\begin{equation}
\label{eq:svr_fx}
f(x)= \sum_{i}{(\alpha_i-\alpha_i^*)k(x_i,x)}+b
\end{equation}

\begin{equation}
\label{eq:svr_w}
w= \sum_{i}{(\alpha_i-\alpha_i^*)\phi(x_i)}
\end{equation}

\begin{equation}
\label{eq:svr_b}
b= y_i-\langle w,x_i \rangle-\epsilon
\end{equation}

Os valores de $\alpha_i,\alpha_i^*$ por sua vez podem ser encontrados de forma iterativa e eficiente utilizando o algorítimo "Sequential minimal optimization" \cite{Platt1998}.

Pela equação (\ref{eq:svr_fx}) podemos ver que a complexidade do algorítimo é linearmente dependente do número de amostras do treinamento assim como os vetores de suporte. A grande vantagem desse método é que pelas restrições do problema de otimização serem estritamente convexas, é garantida uma solução única. Isto significa que algorítimos SVs não tem problemas de mínimos locais como Redes Neurais.

\section{Energia} \label{sec:energia}

Combinando o modelo de energia da seção anterior e a caracterização SVR do tempo podemos estimar a energia total consumida com a equação da energia média descrita por:

\begin{equation}
%E_m=\int_{t1}^{t2}P_mdt=P_m\delta T
E_m=P_m\Delta T
\end{equation}

Substituindo os modelos descritos nas Seções \ref{sec:potencia} e \ref{sec:performance}, obtemos a equação (\ref{eq:energia}).

\begin{equation}
E(f,p,s,N)=P(f,p,s)SVR(f,p,N)
\label{eq:energia}
\end{equation}

Onde $f$ é a frequência, $p$ o número de cores ativos, $s$ o número de processadores ativos e $N$ o tamanho da entrada.

Este modelo possui algumas restrições, pois, não prevê a variação do consumo de potência ao longo do tempo, devido à variação da utilização do processador. Assim é necessário escolher uma carga para modelar a equação de potência. Por causa disso a energia real consumida será diferente da estimada por um fator multiplicativo em alguns casos, mas, a tendência geral será a mesma, o que possibilita encontrar o mínimo.

Com esta equação é possível encontrar a configuração onde a energia é minimizada para uma determinada entrada, mas, também encontrar a configuração mínima de energia dada uma restrição de tempo, frequência ou número de núcleos ativos.

### 2021-03-23 thesis/intro/Capitulos/Referencial_teorico.tex
add tex files

% Capítulo 2
\chapter{Referencial Teórico} \label{cap:referencial_teorico}

\section{Controle de frequência} \label{sec:controle_freq}

Nos sistemas modernos o controle sobre a frequência do processador pode ser feita por hardware com circuitos independentes como também por software. Para isso existe o Advanced Configuration and Power Interface \abrv[ACPI -- Advanced Configuration and Power Interface]{(ACPI)} um padrão aberto adotado por sistemas operacionais para configurar componentes de hardware relacionados a gerenciamento de energia.

No ACPI são definidos dois estados importantes para o DVFS que otimizam o consumo de energia. São eles o "C", que é ativado quando o processador não está executando nenhuma instrução, e o "P", ativado enquanto o processador está operando. Esses estados possuem diversos níveis e em cada nível a frequência e tensão são alterados.

O estado P começa no nível P0, onde a frequência e tensão são as máximas possíveis, depois o P1, onde diminui ambos, até chegar no último estado, Pn, onde a frequência e tensão são as menores possíveis. A mudança de estado depende do nível de utilização do processador. Para se manter em cada estado, o nível de utilização do processador deve está dentro de limites específicos. Apos ultrapassar esses limites por um determinado tempo, o estado irá mudar para o próximo estado correspondente a esse novo nível de utilização do processador. O número de estados possíveis depende de cada fabricante.

Após um tempo ocioso, o processador começa a ativar os estados C, começando pelo C0 onde ele ainda está completamente ativo, depois passando para o C1, onde alguns recursos são desabilitados, até o Cn, em que todos os recursos possíveis são desabilitados. No padrão ACPI são estabelecidas as funcionalidades que podem ser desabilitadas entre o nivel C1 e o nível C3, como vistas na Tabela \ref{tab:estados_c}. Os demais níveis são específicos de cada fabricante.

\begin{table}[H]
	\centering
	\begin{tabular}{|l|l|l|}
		\hline
		Modo & Nome            & Funcionalidade                \\ \hline
		C0   & operating state & Processador ativo             \\ \hline
		C1   & Halt            & Para de executar instruções   \\ \hline
		C2   & Stop-Clock      & Desabilita o clock interno    \\ \hline
		C3   & Sleep           & Desabilita coerência de cache \\ \hline
	\end{tabular}
	\caption{Estados C}
	\label{tab:estados_c}
\end{table}

No estado C, quanto maior o nível maior é a economia de energia, porém o retorno ao nível totalmente funcional é mais difícil. Já nos estados P, existe troca entre desempenho e economia de energia. A Figura \ref{fig:p_state} ilustra melhor a troca de estados, nela podemos ver quais partes dos circuitos são desativadas nos estados C, a latência para voltar a o estado ativo, o consumo de potência e também mostra a relação dos estados P com a frequência.

\begin{figure}[H]
	\centering
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[height=7.5cm]{Imagens/C-states.png}
		\caption{Estados C}
	\end{subfigure}%
	~ 
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[height=7.5cm]{Imagens/P-states.jpg}
		\caption{Estados P}
	\end{subfigure}	
	\caption{Ilustração dos estados C e P}{Imagem alterada de \protect\url{https://www.thomas-krenn.com/en/wiki/Processor_P-states_and_C-states}}
	\label{fig:p_state}
\end{figure}

%Os estados C e P são ortogonais, eles operam de forma independente?

Para esse gerenciamento, o sistema operacional disponibiliza no espaço de usuário uma forma de controlar a frequência. Neste trabalho foi utilizado um sistema operacional que tem como núcleo o Linux.

O Linux é compatível com diversas arquiteturas modernas e amplamente utilizado em servidores, smartphones e supercomputadores. Ele foi baseado no sistema UNIX que tem como filosofia tratar tudo no sistema como um arquivo, incluindo configurações e dispositivos de entrada e saída, como teclado, mouse e disco rígido. Outra característica importante é que ele é modular e partes do sistema podem ser carregadas ou removidas durante sua execução.

No Linux existem diversas opções de gerenciamento de frequência \cite{Brown2005}. Os principais são o acpi-cpufreq, Intel P-state, AMD powernow. Neste trabalho foi utilizado o acpi-cpufreq que é padrão e permite o controle direto da frequência através dos arquivos de sistema. O acpi-cpufreq é um módulo do Linux que utiliza de políticas implementadas que decidem dinamicamente a frequência a ser utilizada. Algumas dessas politicas são:

\begin{itemize}
\item Performance - configurado para a maior frequência possível
\item Powersave - configurado para a menor frequência possível
\item Userspace - o usuário escolhe a frequência a ser utilizada 
\item Ondemand -  controla a frequência dependendo da carga do processador. Quando a carga aumenta a frequência também aumenta de acordo.
\item Conservative - semelhante ao Ondemand mas de forma mais suave, o incremento da frequência é continuo ao invés de ser em saltos.
\end{itemize}

\section{Monitoramento do consumo de potência} \label{sec:monitoramento}
Para medir a potência consumida foram utilizadas as interfaces Running Average Power Limit \abrv[RAPL -- Running Average Power Limit]{(RAPL)} e Intelligent Platform Management Interface \abrv[IPMI -- Intelligent Platform Management Interface]{(IPMI)} descritas em \cite{November2013}.

\subsection{IPMI}
IPMI \cite{November2013} é um conjunto de especificações para subsistemas autônomos que fornece gerenciamento e monitoramento independente de processador, firmware e sistema operacional. A utilização do IPMI permite que os administradores de sistema não precisem se deslocar até o local do servidor, que muitas vezes é distante, para realizar suas tarefas. Também, servidores ficam em locais com baixa temperatura e com muito ruído sonoro devido ao sistema de ventilação, e deve-se evitar passar muito tempo nesses locais. Com o gerenciamento remoto é possível ligar e desligar o sistema, acessar remotamente a \abrv[BIOS -- Basic Input Output System]{(BIOS)} e reinstalar o sistema em caso de alguma falha grave.

\begin{figure}[H]
\centering
\includegraphics[height=7.5cm]{Imagens/IPMI-Block-Diagram.png}
\caption{Diagrama do IMPI}{Imagem retirada de \protect\url{https://pt.wikipedia.org/wiki/Intelligent_Platform_Management_Interface}, são mostrado os principais componentes do IPMI e como eles se comunicam}
\label{fig:IPMI}
\end{figure}

O  acesso a rede do IPMI pode ser feito pelo protocolo HTTP ou por uma ferramenta disponibilizada pelo fabricante (ipmitool), que também realiza o acesso via rede. Ele também é utilizado para monitorar o status da plataforma com um conjunto de sensores acoplados como temperaturas do sistema, tensões, ventiladores e fontes de alimentação.

\subsection{RAPL}

Microprocessadores Intel modernos, a partir da arquitetura SandyBridge, incluem a interface RAPL \cite{Rotem2012, Hahnel2012, Hackenberg2015} projetada para limitar o uso de energia em um chip ao mesmo tempo que garante o máximo desempenho. Essa interface suporta recursos de medição de energia através de um circuito integrado que estima o uso de energia com base em um modelo conduzido por contadores de eventos arquiteturais de todos os componentes. Além disso também fornece leituras de temperatura e modelos de vazamento de corrente. As estimativas são disponibilizadas em registradores específico de modelo \abrv[MSR -- Model Specific Register]{(MSR)}, sendo atualizado na ordem de milissegundos. As estimativas de energia oferecidas pela RAPL foram validadas pela Intel que mostrou ótimos resultados.

### 2021-03-23 thesis/intro/Capitulos/Resultados.tex
add tex files

% Capítulo 5
\chapter{Resultados} \label{cap:resultados}

Nesta sessão serão mostrados os ajustes dos modelos e o resultado do modelo final de energia.

\section{Modelo de potência}

O modelo de potencia ajustado pode ser visto na Figura \ref{fig:pot_ipmi} e os parâmetros da equação em (\ref{eq:fittedpower}) com a unidade de frequência em GHz.

\begin{figure}[H]
\centering
\includegraphics[height=8cm]{Imagens/potencia_ipmi.png}
\caption{Ajuste Modelo de Potencia}{Onde os pontos representam os valores reais de potência medidos com o IPMI e a grade a curva obtida}
\label{fig:pot_ipmi}
\end{figure}

\begin{equation} 
P_{total}(f,p,s)=p(0.29f^3+0.97f)+198.59+9.18s \label{eq:fittedpower}
\end{equation}

O MEP resultante foi de 0.75\% e MEA foi de 1.89W.

\section{Modelo de performance} \label{sec:ajuste_perf}

Alguns resultados da caracterização podem ser vistos na Figura \ref{fig:tempos} onde é mostrada a curva obtida pela SVR e os valores reais medidos.

\begin{figure}[H]
	\centering
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=\columnwidth]{Imagens/time/black_3.png}
		\caption{Tempo Blackscholes}
	\end{subfigure}%
	~ 
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=\columnwidth]{Imagens/time/fluid_3.png}
		\caption{Tempo Fluidanimate}
	\end{subfigure}
	\\
	\centering
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=\columnwidth]{Imagens/time/swap_3.png}
		\caption{Tempo Swaptions}
	\end{subfigure}%
	~
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=\columnwidth]{Imagens/time/rtview_3.png}
		\caption{Tempo Raytrace}
	\end{subfigure}

	\caption{Tempo para modelo de performance}
	\label{fig:tempos}
\end{figure}

A media dos  resultados da validação cruzada podem ser vistos na Tabela \ref{tab:svr_evaluation}.

\begin{table}[H]
\centering
\begin{tabular}{|l|l|l|l|l|}
\hline
\hline
Application  & MEA & MEP \\ \hline
Blackscholes & 2.01  & 4.6\% \\ \hline
Fluidanimate & 6.65  & 1.89\% \\ \hline
Raytrace  & 3.77  & 0.87\% \\ \hline
Swaptions & 2.29  & 2.56\% \\ \hline
\end{tabular}
\caption{Performance-Model's Cross validation Errors}
\label{tab:svr_evaluation}
\end{table}

\section{Energia} \label{sec:ajuste_en}

As Figuras \ref{fig:energia_black}, \ref{fig:energia_fluid}, \ref{fig:energia_rtview} e \ref{fig:energia_swaptions} mostram os valores de energia medidos junto com a superfície ajustada do modelo. As medições de energia foram obtidas integrando a potência ao longo do tempo total de execução da aplicação.

\begin{figure}[H]
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/black_1.png}
        \caption{Entrada 1}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/black_2.png}
        \caption{Entrada 2}
    \end{subfigure}
    \\
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/black_3.png}
        \caption{Entrada 3}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/black_4.png}
        \caption{Entrada 4}
    \end{subfigure}
    \\
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/black_5.png}
        \caption{Entrada 5}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/black_6.png}
        \caption{Entrada 6}
    \end{subfigure}
	\caption{Modelo de Energia do Blackscholes}
	\label{fig:energia_black}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/fluid_1.png}
        \caption{Entrada 1}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/fluid_2.png}
        \caption{Entrada 2}
    \end{subfigure}
    \\
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/fluid_3.png}
        \caption{Entrada 3}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/fluid_4.png}
        \caption{Entrada 4}
    \end{subfigure}
    \\
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/fluid_5.png}
        \caption{Entrada 5}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/fluid_6.png}
        \caption{Entrada 6}
    \end{subfigure}

    \caption{Modelo de Energia do Fluidanimate}
    \label{fig:energia_fluid}
\end{figure}

O FluidAnimate possui uma restrição quanto ao número de núcleos que precisa ser potência de 2, assim não temos as mesmas variações das aplicações anteriores mas apresentou comportamento similar, sendo que o decaimento no consumo de energia foi mais gradual com o número de núcleos.

Podemos observar um decaimento do tipo $\frac{1}{p}$ com o numero de núcleos ativos devido ao tempo paralelo seguir essa relação com o tempo serial $T_{paralelo} \alpha \frac{T_{serial}}{p}$, isso mostra a importância da paralelização na economia de energia.

\begin{figure}[H]
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/swap_1.png}
        \caption{Entrada 1}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/swap_2.png}
        \caption{Entrada 2}
    \end{subfigure}
    \\
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/swap_3.png}
        \caption{Entrada 3}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/swap_4.png}
        \caption{Entrada 4}
    \end{subfigure}
    \\
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/swap_5.png}
        \caption{Entrada 5}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/swap_6.png}
        \caption{Entrada 6}
    \end{subfigure}

    \caption{Modelo de Energia do Swaptions}
    \label{fig:energia_swaptions}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/rtview_1.png}
        \caption{Entrada 1}
	    \label{fig:energia_rtview_e1}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/rtview_2.png}
        \caption{Entrada 2}
        \label{fig:energia_rtview_e2}
    \end{subfigure}
    \\
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/rtview_3.png}
        \caption{Entrada 3}
        \label{fig:energia_rtview_e3}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/rtview_4.png}
        \caption{Entrada 4}
        \label{fig:energia_rtview_e4}
    \end{subfigure}
    \\
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/rtview_5.png}
        \caption{Entrada 5}
        \label{fig:energia_rtview_e5}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\columnwidth]{Imagens/energy/rtview_6.png}
        \caption{Entrada 6}
        \label{fig:energia_rtview_e6}
    \end{subfigure}

    \caption{Modelo de Energia do RayTrace}
    \label{fig:energia_rtview}
\end{figure}

Podemos perceber que a tendencia foi seguida corretamente, tendo uma maior distorção apenas nas primeiras entradas do Raytrace visto nas figuras \ref{fig:energia_rtview_e1},\ref{fig:energia_rtview_e2},\ref{fig:energia_rtview_e3}.

\section{Comparação de Energia} \label{sec:comp}

Para descobrir a verdadeira economia do método proposto ele foi comparando o consumo de energia das quatro aplicações de estudo de caso. Usando as configurações ótimas de energia fornecidas pela abordagem proposta com o DVFS padrão do Linux, o Ondemand e com o Intel P-state, variando o número de núcleos ativos com os valores de 1 a 32 de forma exponencial e expondo os melhores casos dos 3 métodos, para termos uma melhor ideia de comparação, são mostrados nas tabelas os valores absolutos de consumo, a frequência media e a economia com relação ao modelo proposto.

\begin{table}[H]
	\resizebox{\columnwidth}{!}{
		\begin{tabular}{l|l|l|l|l|l|l|ll}
			\multicolumn{1}{l|}{\rot{Entrada}} & \rot{\begin{tabular}[c]{@{}l@{}}Freq. Media \\ em GHz \\ (\#Cores) \end{tabular}}  & \rot{Energia em KJ}   &  \rot{\begin{tabular}[c]{@{}l@{}}Freq. Media \\ em GHz \\ (\#Cores) \end{tabular}} & \rot{Energia em KJ}  & \rot{\begin{tabular}[c]{@{}l@{}} Freq. \\ em GHz \\ (\#Cores) \end{tabular}} & \rot{Energia em KJ}        & \multicolumn{1}{l|}{\rot{\begin{tabular}[c]{@{}l@{}} Economia \\ Ondemand \end{tabular}}} & \multicolumn{1}{l}{\rot{\begin{tabular}[c]{@{}l@{}} Economia \\ Intel\end{tabular}}} \\ \hline
			\multicolumn{1}{l|}{1} & 1.31 (32) & 0.66 & 1.27 (32) & 0.72 & 1.50 (32)  & 0.97 & \multicolumn{1}{l|}{-32.18} & \multicolumn{1}{l}{-25.89} \\ \hline
			\multicolumn{1}{l|}{2} & 1.20 (32) & 2.18 & 2.80 (32) & 1.67 & 2.10 (28)  & 1.67 & \multicolumn{1}{l|}{30.68} & \multicolumn{1}{l}{0.19} \\ \hline
			\multicolumn{1}{l|}{3} & 2.33 (32) & 3.22 & 2.09 (32) & 3.53 & 2.20 (32)  & 2.97 & \multicolumn{1}{l|}{8.40} & \multicolumn{1}{l}{18.66} \\ \hline
			\multicolumn{1}{l|}{4} & 1.42 (32) & 7.44 & 2.00 (32) & 6.46 & 2.20 (28)  & 6.52 & \multicolumn{1}{l|}{14.11} & \multicolumn{1}{l}{-0.84} \\ \hline
			\multicolumn{1}{l|}{5} & 1.20 (32) & 15.40 & 2.42 (32) & 12.57 & 2.20 (28)  & 13.33 & \multicolumn{1}{l|}{15.48} & \multicolumn{1}{l}{-5.69} \\ \hline
			\multicolumn{1}{l|}{6} & 1.54 (32) & 26.98 & 1.77 (32) & 26.48 & 2.20 (30)  & 25.73 & \multicolumn{1}{l|}{4.85} & \multicolumn{1}{l}{2.89} \\ \hline
			& \multicolumn{2}{l|}{Ondemand Min.} & \multicolumn{2}{l|}{Intel Min.} & \multicolumn{2}{l|}{Proposta} & Economia em \% & \\ 
		\end{tabular}
	}
	\caption{Comparação energia miníma Blackscholes}{Comparação do método proposto com Ondemand e P-state, mostrando a porcentagem de economia com relação aos melhores valores encontrados, onde positivo significa economia.}
	\label{tab:Blackscholesfreq}
\end{table}

Na Tabela \ref{tab:Blackscholesfreq} podemos ver que na maioria das vezes foi obtida configurações que consumiu menos energia, chegando a economia de até 30\%. Apenas para a primeira entrada onde tem o menor tempo de execução a configuração proposta não compensou. A número de núcleos ativos na melhor configuração sempre foi próximo do máximo, ou seja, 32, mesmo que a aplicação não tenha um decaimento no tempo significativo com o aumento no número de núcleos, isso confirma a importância da paralelização nas aplicações para se obter economia de energia. Outra observação importante e que nem sempre a menor frequência levou ao menor consumo de energia, como podemos ver pelo Ondemand que em para todas as entradas teve um média de frequência menor e foi o que consumiu mais energia dos 3 métodos comparados.

\begin{table}[H]
	\resizebox{\columnwidth}{!}{
		\begin{tabular}{l|l|l|l|l|l|l|ll}
			\multicolumn{1}{l|}{\rot{Entrada}} & \rot{\begin{tabular}[c]{@{}l@{}}Freq. Media \\ em GHz \\ (\#Cores) \end{tabular}}  & \rot{Energia em KJ}   &  \rot{\begin{tabular}[c]{@{}l@{}}Freq. Media \\ em GHz \\ (\#Cores) \end{tabular}} & \rot{Energia em KJ}  & \rot{\begin{tabular}[c]{@{}l@{}} Freq. \\ em GHz \\ (\#Cores) \end{tabular}} & \rot{Energia em KJ}        & \multicolumn{1}{l|}{\rot{\begin{tabular}[c]{@{}l@{}} Economia \\ Ondemand \end{tabular}}} & \multicolumn{1}{l}{\rot{\begin{tabular}[c]{@{}l@{}} Economia \\ Intel\end{tabular}}} \\ \hline
			\multicolumn{1}{l|}{1} & 2.15 (32) & 2.28 & 2.61 (32) & 2.61 & 2.10 (32)  & 2.35 & \multicolumn{1}{l|}{-3.03} & \multicolumn{1}{l}{11.05} \\ \hline
			\multicolumn{1}{l|}{2} & 2.01 (32) & 4.36 & 1.74 (32) & 5.13 & 2.00 (32)  & 4.15 & \multicolumn{1}{l|}{5.00} & \multicolumn{1}{l}{23.71} \\ \hline
			\multicolumn{1}{l|}{3} & 1.68 (32) & 9.00 & 2.22 (32) & 10.01 & 2.00 (32)  & 7.89 & \multicolumn{1}{l|}{14.13} & \multicolumn{1}{l}{26.93} \\ \hline
			\multicolumn{1}{l|}{4} & 1.93 (32) & 16.87 & 2.51 (32) & 18.69 & 2.00 (32)  & 16.98 & \multicolumn{1}{l|}{-0.63} & \multicolumn{1}{l}{10.09} \\ \hline
			\multicolumn{1}{l|}{5} & 1.98 (32) & 33.29 & 2.63 (32) & 37.68 & 2.10 (32)  & 33.20 & \multicolumn{1}{l|}{0.26} & \multicolumn{1}{l}{13.47} \\ \hline
			\multicolumn{1}{l|}{6} & 1.43 (32) & 68.43 & 2.54 (32) & 77.30 & 2.10 (32)  & 66.67 & \multicolumn{1}{l|}{2.65} & \multicolumn{1}{l}{15.95} \\ \hline
			& \multicolumn{2}{l|}{Ondemand Min.} & \multicolumn{2}{l|}{Intel Min.} & \multicolumn{2}{l|}{Proposta} & Economia em \% & \\ 
		\end{tabular}
	}
	\caption{Comparação energia miníma Fluidanimate}{Comparação do método proposto com Ondemand e P-state, mostrando a porcentagem de economia com relação aos melhores valores encontrados, onde positivo significa economia.}
	\label{tab:Fluidanimatefreq}
\end{table}

A comparação do Fluidanimate visto na Tabela \ref{tab:Fluidanimatefreq} foi o que apresentou melhor resultado tendo economia em todos as entradas comparado ao P-state e na maioria contra o Ondemand. A frequência do Ondemand e método proposto foram bem próximas e mesmo assim foi obtido economia isso pode ser explicado pela variação de frequência que existe no Ondemand levando a um maior tempo de execução.

\begin{table}[H]
	\resizebox{\columnwidth}{!}{
		\begin{tabular}{l|l|l|l|l|l|l|ll}
			\multicolumn{1}{l|}{\rot{Entrada}} & \rot{\begin{tabular}[c]{@{}l@{}}Freq. Media \\ em GHz \\ (\#Cores) \end{tabular}}  & \rot{Energia em KJ}   &  \rot{\begin{tabular}[c]{@{}l@{}}Freq. Media \\ em GHz \\ (\#Cores) \end{tabular}} & \rot{Energia em KJ}  & \rot{\begin{tabular}[c]{@{}l@{}} Freq. \\ em GHz \\ (\#Cores) \end{tabular}} & \rot{Energia em KJ}        & \multicolumn{1}{l|}{\rot{\begin{tabular}[c]{@{}l@{}} Economia \\ Ondemand \end{tabular}}} & \multicolumn{1}{l}{\rot{\begin{tabular}[c]{@{}l@{}} Economia \\ Intel\end{tabular}}} \\ \hline
			\multicolumn{1}{l|}{1} & 2.77 (32) & 4.11 & 2.77 (32) & 4.14 & 1.80 (32)  & 4.52 & \multicolumn{1}{l|}{-9.00} & \multicolumn{1}{l}{-8.49} \\ \hline
			\multicolumn{1}{l|}{2} & 2.77 (32) & 6.37 & 2.76 (32) & 6.30 & 2.20 (32)  & 5.73 & \multicolumn{1}{l|}{11.18} & \multicolumn{1}{l}{10.01} \\ \hline
			\multicolumn{1}{l|}{3} & 2.77 (32) & 8.49 & 2.76 (32) & 8.51 & 2.20 (32)  & 7.81 & \multicolumn{1}{l|}{8.73} & \multicolumn{1}{l}{8.91} \\ \hline
			\multicolumn{1}{l|}{4} & 2.58 (32) & 10.59 & 2.71 (32) & 10.61 & 2.00 (32)  & 9.90 & \multicolumn{1}{l|}{6.92} & \multicolumn{1}{l}{7.14} \\ \hline
			\multicolumn{1}{l|}{5} & 2.57 (32) & 12.68 & 2.60 (32) & 12.68 & 2.10 (32)  & 12.01 & \multicolumn{1}{l|}{5.63} & \multicolumn{1}{l}{5.62} \\ \hline
			\multicolumn{1}{l|}{6} & 2.15 (32) & 14.86 & 2.30 (32) & 14.78 & 1.90 (32)  & 14.45 & \multicolumn{1}{l|}{2.86} & \multicolumn{1}{l}{2.29} \\ \hline
			& \multicolumn{2}{l|}{Ondemand Min.} & \multicolumn{2}{l|}{Intel Min.} & \multicolumn{2}{l|}{Proposta} & Economia em \% & \\ 
		\end{tabular}
	}
	\caption{Comparação energia miníma Swaptions}{Comparação do método proposto com Ondemand e P-state, mostrando a porcentagem de economia com relação aos melhores valores encontrados, onde positivo significa economia.}
	\label{tab:Swaptionsfreq}
\end{table}

As configurações ótimas para a aplicação Swaptions mostram resultados de economia similares as aplicações mostradas anteriormente, como visto na Tabela \ref{tab:Swaptionsfreq} podemos ver que na maioria das vezes também foi obtida configurações que consumiu menos energia, e apenas para a primeira entrada a configuração proposta não compensou. Nessa aplicação as frequências médias do Ondemand e P-state foram bem maiores do que o método proposto diferente da aplicação anterior, mesmo assim em ambas as aplicações houve economia de energia no método proposto isso mostra que cada aplicação tem suas configurações ideais não tendo uma configuração que economize em todas.

\begin{table}[H]
	\resizebox{\columnwidth}{!}{
		\begin{tabular}{l|l|l|l|l|l|l|ll}
			\multicolumn{1}{l|}{\rot{Entrada}} & \rot{\begin{tabular}[c]{@{}l@{}}Freq. Media \\ em GHz \\ (\#Cores) \end{tabular}}  & \rot{Energia em KJ}   &  \rot{\begin{tabular}[c]{@{}l@{}}Freq. Media \\ em GHz \\ (\#Cores) \end{tabular}} & \rot{Energia em KJ}  & \rot{\begin{tabular}[c]{@{}l@{}} Freq. \\ em GHz \\ (\#Cores) \end{tabular}} & \rot{Energia em KJ}        & \multicolumn{1}{l|}{\rot{\begin{tabular}[c]{@{}l@{}} Economia \\ Ondemand \end{tabular}}} & \multicolumn{1}{l}{\rot{\begin{tabular}[c]{@{}l@{}} Economia \\ Intel\end{tabular}}} \\ \hline
			\multicolumn{1}{l|}{1} & 1.45 (8) & 32.62 & 3.60 (2) & 31.47 & 2.20 (4)  & 34.37 & \multicolumn{1}{l|}{-5.08} & \multicolumn{1}{l}{-8.43} \\ \hline
			\multicolumn{1}{l|}{2} & 1.54 (16) & 37.06 & 1.78 (16) & 34.39 & 2.20 (6)  & 37.92 & \multicolumn{1}{l|}{-2.26} & \multicolumn{1}{l}{-9.31} \\ \hline
			\multicolumn{1}{l|}{3} & 1.85 (8) & 41.53 & 1.33 (8) & 39.65 & 2.20 (10)  & 39.93 & \multicolumn{1}{l|}{3.99} & \multicolumn{1}{l}{-0.72} \\ \hline
			\multicolumn{1}{l|}{4} & 1.88 (16) & 47.04 & 2.24 (16) & 44.44 & 2.20 (14)  & 45.77 & \multicolumn{1}{l|}{2.78} & \multicolumn{1}{l}{-2.91} \\ \hline
			\multicolumn{1}{l|}{5} & 1.68 (32) & 51.88 & 2.42 (32) & 50.34 & 2.20 (22)  & 52.99 & \multicolumn{1}{l|}{-2.09} & \multicolumn{1}{l}{-5.00} \\ \hline
			\multicolumn{1}{l|}{6} & 2.06 (32) & 65.02 & 2.78 (32) & 65.84 & 2.20 (26)  & 67.28 & \multicolumn{1}{l|}{-3.36} & \multicolumn{1}{l}{-2.14} \\ \hline
			& \multicolumn{2}{l|}{Ondemand Min.} & \multicolumn{2}{l|}{Intel Min.} & \multicolumn{2}{l|}{Proposta} & Economia em \% & \\ 
		\end{tabular}
	}
	\caption{Comparação energia miníma Raytrace}{Comparação do método proposto com Ondemand e P-state, mostrando a porcentagem de economia com relação aos melhores valores encontrados, onde positivo significa economia.}
	\label{tab:Raytracefreq}
\end{table}

O Raytrace foi o que obteve pior resultado, como pode ser visto na Tabela \ref{tab:Raytracefreq} não existiu economia quando comparado ao P-state e em poucas entradas para o Ondemand. Uma característica desta aplicação é que diferente das outras e que o sua carga do processador inicialmente é bem baixa e posteriormente vai para 100 \% aplicações com esse perfil mostraram não ter bons resultados com esse método, enquanto que aplicações com alta variação na carga do processador como o FluidAnimate apresentam ótimos resultados.

Para termos uma ideia da economia geral as Figuras \ref{fig:comp_black},\ref{fig:comp_swap},\ref{fig:comp_raytrace} e \ref{fig:comp_fluid} mostram o consumo de energia em porcentagem relativo ao proposto agora mostrando todas as variações, representando os melhores e os piores casos de consumo de energia, bem como a média geral.

\begin{figure}[H]
	\centering
	\begin{subfigure}[t]{\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Imagens/comp/relative_black_ondemand.png}
		\caption{Ondemand}
	\end{subfigure}%
	\\
	\begin{subfigure}[t]{\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Imagens/comp/relative_black_intel.png}
		\caption{Intel P-state}
	\end{subfigure}%
	\caption{Comparação relativa de energia do Blackscholes}{Comparação mostrando o consumo de energia relativa ao modelo proposto para todas as variações de núcleos.}
	\label{fig:comp_black}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[t]{\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Imagens/comp/relative_fluid_ondemand.png}
		\caption{Ondemand}
	\end{subfigure}%
	\\
	\begin{subfigure}[t]{\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Imagens/comp/relative_raytrace_intel.png}
		\caption{Intel P-state}
	\end{subfigure}%
	\caption{Comparação relativa de energia FluidAnimate}{Comparação mostrando o consumo de energia relativa ao modelo proposto para todas as variações de núcleos.}
	\label{fig:comp_fluid}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[t]{\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Imagens/comp/relative_swap_ondemand.png}
		\caption{Ondemand}
	\end{subfigure}%
	\\
	\begin{subfigure}[t]{\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Imagens/comp/relative_swap_intel.png}
		\caption{Intel P-state}
	\end{subfigure}%
	\caption{Comparação relativa de energia Swaptions}{Comparação mostrando o consumo de energia relativa ao modelo proposto para todas as variações de núcleos.}
	\label{fig:comp_swap}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[t]{\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Imagens/comp/relative_raytrace_ondemand.png}
		\caption{Ondemand}
	\end{subfigure}%
	\\
	\begin{subfigure}[t]{\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Imagens/comp/relative_raytrace_intel.png}
		\caption{Intel P-state}
	\end{subfigure}%
	\caption{Comparação relativa de energia Raytrace}{Comparação mostrando o consumo de energia relativa ao modelo proposto para todas as variações de núcleos.}
	\label{fig:comp_raytrace}
\end{figure}

Em geral, o consumo de energia dos esquemas padrões de DFVS, P-state e Ondemand, foi maior para números menores de núcleos isso mostra que a paralelização também é um fator importante para o consumo de energia. No entanto, nem sempre foi o caso de o melhor número de núcleos para este esquema ser o máximo, ou seja, 32 núcleos. Possivelmente, para arquiteturas com maior número de núcleos, a escolha do número exato que minimiza o consumo de energia seja menos evidente.

Em todos os casos, o método proposto superou o pior caso dos esquemas padrões, chegando a consumir 14 vezes menos, de acordo com o que pode ser observado na Figura \ref{fig:comp_swap} entrada 2. Já considerando os melhores casos apresentados nas tabelas, na maioria das vezes foi atingido economia chegando a até 30.68\% em \ref{tab:Blackscholesfreq} na entrada 2.

As Figuras \ref{fig:comp_black32},\ref{fig:comp_swap32},\ref{fig:comp_raytrace32} e \ref{fig:comp_fluid32} mostram a energia consumida em joules para cada entrada comparando a configuração proposta com o Ondemand e o P-state no seu estado padrão, isto é, com todos os núcleos ativos.

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{Imagens/comp/relative_black_side_by_size_energia.png}
	\caption{Comparação Blackscholes}
	\label{fig:comp_black32}
\end{figure}%

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{Imagens/comp/relative_fluid_side_by_size_energia.png}
	\caption{Comparação com 32 núcleos ativos para o FluidAnimate}
	\label{fig:comp_fluid32}
\end{figure}%

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{Imagens/comp/relative_swap_side_by_size_energia.png}
	\caption{Comparação com 32 núcleos ativos para o Swpations}
	\label{fig:comp_swap32}
\end{figure}%

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{Imagens/comp/relative_raytrace_side_by_size_energia.png}
	\caption{Comparação com 32 núcleos ativos para o Raytrace}
	\label{fig:comp_raytrace32}
\end{figure}%

É possível observar que na maioria dos casos a proposta apresentada obteve melhores resultados no estado padrão do Ondemand e P-state, para melhor concluir a tabela \ref{tab:dif_energia} é um resumo dos gráficos acima onde pode ser observada a diferença do somatório do consumo em todas as aplicações entre o modelo proposto e os já estabelecidos. Na média geral este método obteve 5.7\% de economia de energia.

\begin{table}[H]
	\centering
	\begin{tabular}{c|c|c|c|c|}
		%\cline{2-5}
		& Kj              & \%          & Kj            & \%         \\ \hline
		\multicolumn{1}{c|}{blackscholes} & 4682.96         & 9.15        & 237.06        & 0.46       \\ \hline
		\multicolumn{1}{c|}{swaptions}    & 2689.88         & 4.94        & 2599.10       & 4.78       \\ \hline
		\multicolumn{1}{c|}{raytrace}     & 13863.12        & 4.98        & 3624.61       & 1.30       \\ \hline
		\multicolumn{1}{c|}{fluidanimate} & 2997.39         & 2.28        & 20188.61      & 15.38      \\ \hline
		& \multicolumn{2}{c|}{Ondemand} & \multicolumn{2}{c|}{Intel} \\% \cline{2-5} 
	\end{tabular}
	\caption{Diferença total energia}
	\label{tab:dif_energia}
\end{table}

\subsection{Comparação de Desempenho}

No DVFS existe uma troca entre economia e desempenho, por isso também foram analisando os efeitos que a execução com a configuração do modelo proposto teve no tempo de execução das aplicações. As figuras \ref{fig:tempo_black}, \ref{fig:tempo_fluid}, \ref{fig:tempo_raytrace} e \ref{fig:tempo_swap} mostram os tempos de execução para P-state, Ondemand e modelo proposto.

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{Imagens/comp/relative_black_side_by_size_tempo.png}
	\caption{Comparação de tempo para o Blackscholes}
	\label{fig:tempo_black}
\end{figure}%

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{Imagens/comp/relative_fluid_side_by_size_tempo.png}
	\caption{Tempo FluidAnimate}
	\label{fig:tempo_fluid}
\end{figure}%

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{Imagens/comp/relative_swap_side_by_size_tempo.png}
	\caption{Comparação de tempo para o Swpations}
	\label{fig:tempo_swap}
\end{figure}%

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{Imagens/comp/relative_raytrace_side_by_size_tempo.png}
	\caption{Comparação de tempo para o Raytrace}
	\label{fig:tempo_raytrace}
\end{figure}%

\begin{table}[H]
	\label{tab:med_tempo}
	\begin{tabular}{c|c|c|c|c|}
		%\cline{2-5}
		& s             & \%            & s            & \%          \\ \hline
		\multicolumn{1}{c|}{blackscholes} & -12.82        & -7.90         & -36.75       & -22.66      \\ \hline
		\multicolumn{1}{c|}{swaptions}    & -41.38        & -24.96        & -41.75       & -25.18      \\ \hline
		\multicolumn{1}{c|}{raytrace}     & 41.05         & 3.78          & -161.67      & -14.90      \\ \hline
		\multicolumn{1}{c|}{fluidanimate} & 13.34         & 3.30          & -29.99       & -7.41       \\ \hline
		\multicolumn{1}{l|}{}              & \multicolumn{2}{c|}{Ondemand} & \multicolumn{2}{c|}{Intel} \\ %\cline{2-5} 
	\end{tabular}
	\centering
	\caption{Diferença total tempo}
\end{table}

Podemos ver que o desempenho foi pior para o modelo proposto, sendo esperado pois no geral foi obtido economia de energia. Também podemos confirmar que o FluidAnimate teve um maior tempo de execução no Ondemand mesmo executando a uma frequência próxima a da configuração proposta e isso levou a um maior consumo de energia como mostrado anteriormente.

\section{Energia utilizada no treinamento}

Como o objetivo é economizar energia e a etapa de treinamento requer um consumo considerável, foi estudada a redução do gasto reduzindo o número de amostras utilizadas no treinamento. As imagens \ref{fig:over_black}, \ref{fig:over_swap}, \ref{fig:over_raytrace} e \ref{fig:over_fluid} mostram o consumo de energia pelo número de amostras, junto com o EMP.

O numero de amostras foram escolhidas de forma que os pontos da borda são sempre selecionados, ou seja, para o maior e menor número de núcleos todas as frequências são utilizadas no treinamento, da mesma forma para maior e menor frequência, todas as núcleos são utilizadas, o restante dos pontos são escolhidos de forma aleatória. Essa estratégia se mostrou mais eficaz do que escolher todos os pontos de forma aleatória.

\begin{figure}[H]
	\centering
	\begin{subfigure}[t]{0.48\textwidth}
		\includegraphics[width=\linewidth]{Imagens/overhead/energy_Blackscholes.png}
		\caption{Energia}
	\end{subfigure}\hspace{0.04\textwidth}%
	\begin{subfigure}[t]{0.48\textwidth}
		\includegraphics[width=\linewidth]{Imagens/overhead/error_Blackscholes.png}
		\caption{Erro}
	\end{subfigure}%
	\caption{Energia utilizada no treinamento para o Blackscholes}
	\label{fig:over_black}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[t]{0.48\textwidth}
		\includegraphics[width=\linewidth]{Imagens/overhead/energy_Fluidanimate.png}
		\caption{Energia}
	\end{subfigure}\hspace{0.04\textwidth}%
	\begin{subfigure}[t]{0.48\textwidth}
		\includegraphics[width=\linewidth]{Imagens/overhead/error_Fluidanimate.png}
		\caption{Erro}
	\end{subfigure}%
	\caption{Energia utilizada no treinamento para o FluidAnimte}
	\label{fig:over_fluid}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[t]{0.48\textwidth}
		\includegraphics[width=\linewidth]{Imagens/overhead/energy_Swaptions.png}
		\caption{Energia}
	\end{subfigure}\hspace{0.04\textwidth}%
	\begin{subfigure}[t]{0.48\textwidth}
		\includegraphics[width=\linewidth]{Imagens/overhead/error_Swaptions.png}
		\caption{Erro}
	\end{subfigure}%
	\caption{Energia utilizada no treinamento para o Swaptions}
	\label{fig:over_swap}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[t]{0.48\textwidth}
		\includegraphics[width=\linewidth]{Imagens/overhead/energy_Raytrace.png}
		\caption{Energia}
	\end{subfigure}\hspace{0.04\textwidth}%
	\begin{subfigure}[t]{0.48\textwidth}
		\includegraphics[width=\linewidth]{Imagens/overhead/error_Raytrace.png}
		\caption{Erro}
	\end{subfigure}%
	\caption{Energia utilizada no treinamento para o Raytrace}
	\label{fig:over_raytrace}
\end{figure}

Para as aplicações Blackscholes, Swaptions e Raytrace, o erro começa a ser aceitável, próximo de 5\%, a partir de 400 amostras, e para o FluidAnimate que possui menos amostras, devido ao número de núcleos ser necessariamente potência de 2, teve erro aceitável com mais de 200 amostras. Com isso temos em média o consumo de 20 Mega Joules comparado com 50 Mega Joules utilizando todos os pontos. Apesar da redução de mais de 50\% o consumo de energia ainda é muito grande no treinamento. Levando em conta o consumo médio mostrado na tabela \ref{tab:dif_energia}, seria necessário rodar as aplicações mais de 4000 vezes para compensarmos a energia utilizada no treinamento. Uma possível solução para este problema seria utilizar um modelo matemático para o desempenho, com isso o número de amostras para ajustar o modelo cairia drasticamente, deixando ele bastante viável.

### 2021-03-23 thesis/intro/Capitulos/Revisao_bibliografica.tex
add tex files

% Capítulo 2
\chapter{Revisão Bibliografia} \label{cap:revisao_bibliografica}

O DVFS é a técnica mais comum empregada para obter economia de energia em sistemas com processadores modernos. Assim, a técnica tem sido extensivamente pesquisada com o objetivo de fornecer estratégias para selecionar a tensão e a frequência ideais para uma aplicação em arquitetura específica. Existem diversas teorias sobre a melhor forma de aplicá-la. No artigo \cite{Albers2014} foi provado que o problema de escalonamento de frequência com o estado ocioso para minimizar a energia é da classe NP-difícil. Embora seja complicado encontrar a melhor configuração possível, com algumas simplificações no modelo é possível encontrar bons resultados, sendo um problema  no qual não existe uma solução ótima.

Em \cite {Anghel2011}, os autores utilizaram dois algoritmos para escalonar a frequência dos processadores: um algoritmo inspirado no sistema imunológico humano para monitorar os estados de potência e desempenho do servidor e um algoritmo baseado em lógica fuzzy para alterar o estado de desempenho do servidor. Também em \cite{Cochran2011} foi introduzido um método de escalonamento para determinar os pontos de operação ideais do sistema para o número de núcleos e configurações de DVFS.

\cite{DaCosta2015} propôs uma abordagem que considera estados de atividade instantânea do sistema, nesse caso, a memória e a atividade de rede foram usadas para gerar uma configuração de gerenciamento de DVFS.

Registradores contadores de performance do RAPL também foram usados para executar o DVFS. Em \cite{Spiliopoulos2011}, os autores usaram um DVFS contínuo adaptável com base em um modelo de desempenho do processador. O modelo foi baseado na amostragem de contadores de desempenho em hardware para ver regularmente cargas de trabalho de desempenho/energia. Baseando-se nas predições as configurações apropriadas de voltagem e as frequências eram selecionadas.

\cite{Georgiou:2017} utilizaram um modelo de energia para uma arquitetura multissegmentada de vários núcleos e análise de recursos estáticos para avaliar a economia de energia e tempo de várias configurações de DVFS para o mesmo programa. Embora tenham sido capazes de identificar a melhor configuração sem a necessidade de executar o programa com cada configuração diferente e medir tempo e energia, a abordagem é bastante limitada, pois a análise estática não é dimensionada para arquiteturas e programas menos previsíveis.

Com base no que já foi pesquisado foi percebido que modelagem de potencia e desempenho de aplicações não são utilizados para o DVFS, que é a proposta deste trabalho. Além disso o foco do DVFS é redução do consumo de potência o que pode levar a um maior consumo de energia devido ao aumento no tempo total de execução. Outro problema é quando existe variações bruscas na carga do processador que causam uma variação muito grande na frequência que geralmente não leva a bons resultados.

### 2021-03-23 thesis/intro/Includes/Abstract.tex
add tex files

% Resumo em língua estrangeira (em inglês Abstract, em espanhol Resumen, em francês Résumé)
\begin{center}
	{\Large{\textbf{Energy-Optimal Configurations for high performance computer applications}}}
\end{center}

\vspace{1cm}

\begin{flushright}
	Author: Vitor Ramos Gomes da Silva\\
	Advisor: Prof. Dr. Samuel Xavier de Souza
\end{flushright}

\vspace{1cm}

\begin{center}
	\Large{\textsc{\textbf{Abstract}}}
\end{center}

\noindent This work proposes a methodology to minimize energy consumption
in high-performance computing by finding the ideal frequency and active cores parameters in the processor. Based on a model of the power draw of the system, developed based on the power consumption of the CMOS circuit, and an estimation of the execution time of the application using artificial intelligence is calculated the energy equation with frequency, number of active cores and input of the application as variables and from this equation is found the configuration that minimizes the total power consumption. The results of this method in four applications show that the proposed approach achieved savings of more than 30 \% when compared to the current energy saving scheme in Linux.

\noindent\textit{Keywords}: 
Energy Efficient Software, Energy Modeling, Frequency Scaling.

### 2021-03-23 thesis/intro/Includes/Agradecimentos.tex
add tex files

% Agradecimentos

\chapter*{Agradecimentos}
Aos meus pais, Cláudia Ramos Gomes da Silva e Paulo Sérgio Gomes da Silva, por tudo. A minha irmã, meus primos e amigos especialmente Andressa Stefany pelo apoio. Aos meus orientadores Samuel Xavier de Souza, que propôs a ideia para este trabalho e Alex Furtunato que co-orientou. A todos os professores do curso de Engenharia de Computação, que sempre estiveram dispostos a ajudar e contribuir para um melhor aprendizado. Ao NPAD e aos servidores Pedro Klisley e Daniel Holanda por também colaborarem na pesquisa. E a todos que direta ou indiretamente fizeram parte da minha formação, obrigado.

### 2021-03-23 thesis/intro/Includes/Capa.tex
add tex files

% Capa
% Proteção externa do trabalho e sobre a qual se imprimem as informações indispensáveis 
% à sua identificação.

% Especificação da capa
\begin{titlepage}
	\begin{center}

		% Cabeçalho (não deve ser modificado)
		% Contém o brasão da Universidade, o logotipo do Departamento, além dos dados
		% relacionados à vinculação do aluno (Universidade, Centro, Departamento e Curso)
		\begin{minipage}{2cm}
			\begin{center}
				\includegraphics[width=1.7cm, height=2.0cm]{Imagens/Brasao-UFRN.jpg}
			\end{center}
		\end{minipage}
		\begin{minipage}{11cm}
			\begin{center}
				\begin{espacosimples}
					{\small \textsc{Universidade Federal do Rio Grande do Norte}			\\
							  \textsc{Centro de Tecnologia}						\\
							  \textsc{Departamento de Engenharia de Computação e Automação}	\\
							  \textsc{Bacharelado em Engenharia de Computação}}
				\end{espacosimples}
			\end{center}
		\end{minipage}
		\begin{minipage}{2cm}
			\begin{center}
				\includegraphics[width=1.8cm, height=1.5cm]{Imagens/Logotipo-DCA.jpg}
			\end{center}
		\end{minipage}

		\vspace{6cm}

		% Título do trabalho
		{\setlength{\baselineskip}%
		{1.3\baselineskip}
		{\LARGE \textbf{Configurações otimizadas de energia para aplicações em computação de alto desempenho}}\par}

		\vspace{4cm}

		% Nome do aluno (autor)
		{\large \textbf{Vitor Ramos Gomes da Silva}}

		\vspace{7cm}

		% Local da instituição onde o trabalho deve ser apresentado e ano de entrega do mesmo
		Natal-RN\\ \today%03 de maio de 2018
	\end{center}
\end{titlepage}

### 2021-03-23 thesis/intro/Includes/Dedicatoria.tex
add tex files

% Dedicatória

\chapter*{}
\vspace*{\fill}
\begin{text}
\centering
\noindent
\textit{Aos meus pais e irmão que, com muito carinho e apoio, não mediram esforços para que eu chegasse até esta etapa tão importante em minha vida. Não posso esquecer dos meus amigos e colegas, pelo incentivo e pelo apoio constantes. E a todos aqueles que de alguma forma contribuíram no meu caminhar pelas pedras até esse momento.}
\end{text}
\vspace*{\fill}

% \vspace{15cm}
% \begin{flushright}
% Aos meus pais, que com muito carinho e apoio não mediram esforços para que eu chegasse até a esta etapa da minha vida.
% \end{flushright}

### 2021-03-23 thesis/intro/Includes/FolhaAprovacao.tex
add tex files

% Folha de aprovação
\begin{folhadeaprovacao}
	\setlength{\ABNTsignthickness}{0.4pt}
	\setlength{\ABNTsignwidth}{10cm}

	% Informações gerais acerca do trabalho 
	% (nome do autor, título, instituição à qual é submetido e natureza)
	\noindent 
	Monografia de Graduação sob o título \textit{Configurações otimizadas de energia para aplicações em computação de alto desempenho} apresentada por 
	Vitor Ramos Gomes da Silva e aceita pelo Departamento de Engenharia de Computação e Automação do
	Centro de Tecnologia da Universidade Federal do Rio Grande do Norte,
	sendo aprovada por todos os membros da banca examinadora abaixo especificada:

	% Membros da banca examinadora e respectivas filiações
	\assinatura
	{
		Prof. Dr. Samuel Xavier de Souza\\
		{\small Orientador(a)} 															\\ 
		{\footnotesize
			Departamento de computação e automação																	\\
		  	Universidade Federal do Rio Grande do Norte
		}
	}

	\assinatura
	{
		Prof. Dr. Alex Furtunato							
		\\
		{\small Co-orientador(a)}							
		\\ 
		{\footnotesize
			Departamento 																	\\
		  	O Instituto Federal de Educação, Ciência e Tecnologia do Rio Grande do Norte
		}
	}

	\assinatura
	{
		Luiz Felipe de Queiroz Silveira 						 \\ 
		{\footnotesize
			Departamento de computação e automação 																	 \\
		  	Universidade Federal do Rio Grande do Norte
		}
	}

% 	\assinatura
% 	{
% 		Titulação e nome do membro da banca examinadora 						 \\ 
% 		{\footnotesize
% 			Departamento 																	 \\
% 		  	Universidade
% 		}
% 	}

	\vfill

	\begin{center}
		Natal-RN, \today.
	\end{center}
\end{folhadeaprovacao}

### 2021-03-23 thesis/intro/Includes/FolhaRosto.tex
add tex files

% Folha de rosto
% Contém os elementos essenciais à identificação do trabalho.

% Título, nome do aluno e respectivo orientador e filiação
\titulo{\Large{Configurações otimizadas de energia para aplicações em computação de alto desempenho}}
\autor{Vitor Ramos Gomes da Silva}
\orientador[Orientador]{\par Prof. Dr. Samuel Xavier de Souza}
\coorientador[Co-Orientador]{\par Prof. Dr. Alex Furtunato}

\instituicao
{
	Universidade Federal do Rio Grande do Norte -- UFRN \par 
	Departamento de Engenharia de Computação e Automação -- DCA
}

% Natureza do trabalho (não deve ser modificada)
\comentario
{
Trabalho de Conclusão de Curso apresentado ao Departamento de Engenharia de Computação e Automação do Centro de Tecnologia da Universidade Federal do Rio Grande do Norte como parte dos requisitos necessários para obtenção do grau de Engenheiro de Computação com habilitação em Automação Industrial.
}

% Local e data
\local{Natal-RN}
\data{\today}

\folhaderosto

### 2021-03-23 thesis/intro/Includes/Resumo.tex
add tex files

% Resumo em língua vernácula
\begin{center}
	{\Large{\textbf{Configurações otimizadas de energia para aplicações em computação de alto desempenho}}}
\end{center}

\vspace{1cm}

\begin{flushright}
	Autor: Vitor Ramos Gomes da Silva\
	Orientador: Prof. Dr. Samuel Xavier de Souza
\end{flushright}

\vspace{1cm}

\begin{center}
	\Large{\textsc{\textbf{Resumo}}}
\end{center}

\noindent 
Este trabalho propõe uma metodologia para minimizar o consumo de energia 
em computação de alta performance, encontrando os parâmetros ideais de frequência e núcleos ativos no processador. A partir de uma modelagem da potência do sistema, desenvolvida com base no consumo de potência do circuito CMOS e uma estimativa do tempo de execução da aplicação utilizando inteligência artificial, é calculada a equação da energia em função da frequência, número de núcleos e entrada da aplicação. Com esta equação, é encontrada a configuração que minimiza o consumo total de energia. Foram feitos testes em quatro aplicações e os resultados desse método mostraram que essa abordagem chegou à economia de até 30\% quando comparado ao esquema de economia de energia existente atualmente no Linux.

\noindent\textit{Palavras-chave}: Software energeticamente eficiente, Modelagem do consumo de potência, Escalonamento de frequência

### 2021-03-23 thesis/intro/intro.tex
add tex files

\section{Motivation}
The energy efficiency of data centers has attained key importance in recent years, due to its high economic, environmental, and performance impact. 
Data center consumption was estimated in 2010 at between 1.1\% and 1.5\% of the worldwide electricity usage \cite{Dayarathna2016DataSurvey,Corcoran2017EmergingICT}, generating as much pollution as a nation like Argentina  \cite{Mathew2012Energy-awareNetworks}.
Power costs exceed in some cases the cost of purchasing hardware \cite{Rivoire2007ModelsOptimizations}. 
Nowadays, the leading petaflop supercomputers consume a range of 1–18 MW of electrical power, with 1.5 MW on average, which can be easily translated into millions of dollars per year in electricity bills \cite{Group2012HandbookSahni}.
Furthermore, the energy costs of powering a typical data center doubles every five years \cite{Buyya2013MasteringProgramming}.
Therefore, with such steep increase in electricity use and rising electricity costs, electricity bills have become a major expense for today’s data centers \cite{Poess2008EnergyResults}, \cite{Gao2013QualityCenters}. 
Due to these reasons, data center energy efficiency is now considered a chief concern for data center operators, often ahead of the traditional considerations of availability and security.

There are several approaches for green computing, from electrical materials to circuit design, systems integration, and software. These techniques may be different, but they share the same goal: substantially reduce overall system energy consumption without a corresponding negative impact on delivered performance. 
%From a systems-software perspective, the strategy of supplying minimal power necessary to accomplish the work can be accomplished by powering down system components when they are underutilized \cite{Mathew2012Energy-awareNetworks}. 
From a software systems perspective, the strategy of providing just enough power to get a job done can be accomplished by turning off system components when they are underutilized \cite{Mathew2012Energy-awareNetworks}.
%The processor and main memory are usually in the top three power draw components, which dominate energy consumption as shown in figure \ref{fig:powerbreakdown}. 
The processor and main memory are usually the components that dominate power consumption, as shown in Figure  \ref{fig:powerbreakdown}.
The processor can consume as much as 50\% of the total energy \cite{Fan2007, Barroso2007TheComputing, Malladi2012TowardsDRAM}. For that reason, modern processors incorporate several features for power management \cite{Rotem2012Power-managementBridge, Brown2005, Hackenberg2015}, such as Dynamic Power Management (DPM) and Dynamic Voltage and Frequency Scaling (DVFS). 
DPM encompasses a set of techniques for obtaining an energy-efficient computing by deactivating or reducing the performance of system components when they are idle or used partially \cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.
DVFS allows the frequency and voltage to be adjusted in run-time depending on current needs.

\begin{figure}[ht]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/power_breakdown/Xeon E5-2698 + DDR4 + SSD.png}
		\caption{This work}
		\label{fig:1}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/power_breakdown/Xeon E5507 + DDR3 + HDD.png}
		\caption{Study case in \cite{Malladi2012TowardsDRAM}}
		\label{fig:2}
	\end{subfigure}

	\hfill
	\caption{Power breakdown of a typical node of an HPC cluster at full use. The system used in this work (a) was built in 2016 and equipped with two Intel Xeon E5-2698, 128 GB of DDR4 memory and SSD as storage, while (b) the case study in \cite{Malladi2012TowardsDRAM} was built in 2012 and equipped with two Xeon E5507, 32GB of DDR3 memory and HDD as storage}
	\label{fig:powerbreakdown}
\end{figure}

DVFS is motivated by the well-known fact that frequency and power have a near-cubic relationship \cite{Dayarathna2016DataSurvey, Group2012HandbookSahni}. 
This implies that running the CPU at lower frequency causes a linear reduction in the performance while a near-cubic reduction in power and, consequently, it could lead to a near-square reduction in CPU energy. 
This can lead to expressive energy savings depending on the system and its infrastructure.
Although very promising, the system software has yet to determine when and what voltage and frequency to use when running applications.
Otherwise, not only will performance deteriorate, but in the worst case, energy consumption would also increase  \cite{Group2012HandbookSahni}.
%Otherwise, in the worst case, not only performance will degrade, but also energy consumption may increase \cite{Group2012HandbookSahni}. 
Indeed, reducing the frequency results in a longer execution time, which in turn increases the energy consumption of other system components such as memory and disks.
%This is because reducing the frequency results in longer run time which in turn increases the energy consumption of the other components in the system such as memory and disks. 
There is also an overhead of time and energy associated with a voltage and frequency switch.
%There are also overheads of time and energy associated with a voltage and frequency switch.
Thus, finding the most appropriate voltage and frequency to use at all times is not an easy task.
%Finding the most appropriate voltage and frequency to use at each time instant is not an easy task. 
Therefore, since its introduction in 1994, there has been a lot of research on DVFS algorithms.
%Since its introduction in 1994, there has been a huge amount of work on DVFS algorithms in the literature.

The DPM technique can achieve very large energy savings on systems where the static power is high or when the system remains inactive for a long period.
In that case, the problem is to determine when and which components to turn on/off.
With DPM, energy savings of 70\% have been reported ~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.
%In the case of DPM, the problem is to determine when and which components to turn on/off. This technique can archive very large energy savings on systems where the static power consumption is high or if the system stays idle for a long time. Power savings of 70\% have been reported~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.

At the same time as the stated techniques reduce the energy consumption of the system, they also lead to a complex trade-off concerning performance, which must be exploited to produce more energy-efficient algorithms.
%However, at the same time that these power-saving techniques reduce system energy, they also lead to a complex balance between energy savings and high performance, exposing trade-offs between performance and power consumption, which can be exploited to produce more energy-efficient algorithms.  %in the case of the CPU, this can be applied to each core. Here... 
Indeed, this work investigates whether the construction of a model of the energy consumption of an application can lead to significant energy savings. We propose an analytical energy model of an application in function of the two control variables present in most of HPC systems: CPU operating frequency and number of active cores. This model can serve as a base for new DVFS and DPM algorithms, as well as for the analysis of the contribution of each model parameter to the total energy consumption. The model equation includes the two parameters of the application and three parameters of the system. The application parameters incorporate characteristics of the percentage of parallelism and the size of the input. The system parameters include power-related and technology-dependent components that compose dynamic, static and leakage power.

The rest of this paper is divided as follows. In Section \ref{sec:relatedwork}, a general review of models is provided, showing the difference between each approach and its applications. In Section \ref{sec:models}, the models and its parameters are derived alongside its constraints. In section \ref{sec:experimentalvalidation}, the model is validated with the PARSEC benchmark applications. Forward, in Section \ref{sec:dvfs_optmin}, use cases of the model are presented, how it could be applied in DVFS and also analysis of the best saving energy policy depending on the system and the effect of the parallelism in the energy. Finally, the conclusion is make in Section \ref{sec:conclusion}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\section{Related Work} \label{sec:relatedwork}
%
%A model is a formal abstraction of a real system. Models for computer systems can be represented as equations, graphical models, rules, decision trees, sets of representative examples, neural networks, etc. The choice of representation affects the accuracy of the models, as well as their interpretability by people \cite{Hypothesis2012EncyclopediaLearning}. Accurate energy and power consumption models are very important for many energy efficiency schemes employed in computing equipment \cite{Rivoire2007ModelsOptimizations} and they can have multiple uses, including the design, forecasting, and optimization of data center systems. This work focuses on analytical models that could serve for energy optimization and subsequently analysis of important factors in the total energy draw.
%
%The desirable properties of a full-system energy consumption model include accuracy (accurate enough to allow the desired energy saving), speed (generate predictions quickly enough), generality and portability (should be suitable for as many systems as possible), inexpensiveness (should not requires expensive or intrusive infrastructure), and simplicity \cite{Rivoire2008AModels}. However, modeling the exact energy consumption behavior of an HPC system is not straightforward, either at the whole-system level or at the individual component. Patterns of energy consumption in data centers depend on multiple factors such as hardware specifications, workload, cooling requirements, types of applications, etc. Some of these factors cannot be measured easily. Furthermore, it is impractical to perform detailed measurements of the energy consumption of lower-level components without introducing overhead to the system.
%
%The definition of energy (E) is the total amount of work performed by a system over a time period (T) while power (P) is the rate at which the work is performed by the system. The relationship between these three quantities can be expressed as:
%
%\begin{equation}
%E = \int_{0}^{T}P(t)dt
%\label{qe:energy_definition_cont}
%\end{equation}
%
%%Or in the discrete format, as the sum of energy in time windows $t_i$.
%
%%\begin{equation}
%%    E = \sum_{i}t_iP_i
%%    \label{qe:energy_definition_discrete}
%%\end{equation}
%
%As a starting point, the models were classified with respect to his inputs, based on the survey \cite{Dayarathna2016DataSurvey}, where they analyzed more than 200 models pointing their characteristics and limitations, it was possible to divide the models into the categories where is more suitable to drive their purposes.
%According to a survey over 200 models~\cite{Dayarathna2016DataSurvey} that analyzed their characteristics and limitations, they can be divided into categories that are are more suitable to drive their purpose:
%\begin{itemize}
%	\item Temperature
%	\item System Utilization or Workload
%	\item Frequency
%	\item Other system states such as: cache miss, branch prediction, number of instructions executed, and more
%	\label{tab:input_type}
%\end{itemize}
%
%%maybe split into power and performance models.
%Energy models are often described as a combination of two models, the power model of the system and the performance model of the application.
%
%\subsection{Power models}
%Temperature-based models can provide a way of estimating power consumption without being very intrusive. With this kind of model it is possible to estimate energy with a single temperature sensor without the need to restart or introduce anything to the system. This is especially interesting for HPCs and data centers which can not be easily turned off. Another use of this model is in optimization of the refrigeration systems.%, but this will be not covered in this journal.
%
%A large part of the developed models are based on system states. These types of models normally take advantage of performance counters, which are provided by the CPU or the operating system and are capable of counting micro-architectural events, such as: instructions executed, cache-hits, miss-predicted branches, and much more. This kind of model is great for power estimation because they have information about several internal states of the computer. It is possible to achieve accurate predictions within 2\% of error \cite{Joseph2001Run-timeMicroprocessors} in average. However, performance counters are highly dependent on the system and architecture, making them not so easily portable. Also, there are limitations on how many counters can be read simultaneously in the hardware side. The CPU only allows a small number of counters (normally five) to be read at the same time. If more counters are requested, some events will be multiplexed, causing measurement loss. From the software side, measuring a large number of system counters can introduce a significant overhead.
%
%Additionally, hardware performance counters had shown other problems \cite{Weaver2008, Weaver2013a, Das2019SoK:Security, McGuire2009, Ramos2019AnCounters}. Events that should be exact and deterministic (such as the number of executed instructions) show run-to-run variations and overcounting on x86\_64 machines, even when running in strictly controlled environments. Although this does not seem to be a problem when performing energy estimation---since normally the models have configurable weights or parameters that can account for imprecision, they might not be fully reliable.
%
%%include utilization
%
%Frequency-based models serve as base for many power models because at a very low level every digital circuit (including modern processors) is composed of transistors. In these devices, there is a known relationship between power and frequency, which can be simplified by equation \ref{eq:power_simplified}. 
%\begin{equation}
%P = \alpha+\beta f^3,
%\label{eq:power_simplified}
%\end{equation}
%where $\alpha$ and $\beta$ are model parameters and $f$ is the operating frequency. This equation is presented in more detail in Section~\ref{sec:models}.
%
%These models are good for the optimization problems since it is in function of a variable that can be easily controlled: $f$.
%%, although it is highly dependent on the load as shown in figure \ref{fig:cpuload}.
%
%% \begin{figure}
%%     \centering
%%     \includegraphics[width=0.5\columnwidth]{cpu_load.png}
%%     \caption{Power with a fixed frequency for different CPU loads, the load means the amount of time that the CPU was busy in the given time windows}
%%     \label{fig:cpuload}
%% \end{figure}
%
%Hybrid models can also be found in the literature. They are models that use as input frequency and CPU utilization or system states and frequency. In \cite{Horyath2008Multi-modeClusters}, they combine frequency and CPU utilization to obtain an average error of 1\% such as: 
%\begin{equation}
%P = a_3fu+a_2f+a_1u+a_0,
%\label{eq:hybrid}
%\end{equation}
%where $u$ is CPU utilization and $a_0$, $a_1$, $a_2$, $a_3$ are model parameters.
%
%\subsection{Performance models}
%Another very common way to model the energy consumption of the system is with the workload. The workload is an abstract definition of the application. It represents the amount of work done in a given time and speed \ref{eq:workload_definition}. Workload can be defined as~\cite{Paolillo2018OptimisationParallelism, Group2012HandbookSahni, Kim2015RacingHeuristics}:
%\begin{equation}
%W = \int_{0}^{T_{\rm active}}s(t)dt = sT_{\rm active},
%\label{eq:workload_definition}
%\end{equation}
%where $T_{\rm active}$ is total active time, and $s$ is the speed of execution. This kind of model is often translated into utilization-based models so that it is possible to work it in real systems. 
%% The definition of utilization is the ration of the time that the system was active. One possible transition is defined bellow in equation \ref{eq:workload_utilization_tranlation}.
%Utilization can be defined as the ratio between the time that the system is active and the total time. Equation \ref{eq:workload_utilization_tranlation} defines workload in terms of utilization.
%\begin{equation}
%u = \frac{T_{\rm active}}{T_{\rm total}} = \frac{w/s}{T_{\rm total}},
%\label{eq:utilization_definition}
%\end{equation}
%\begin{equation}
%w = usT_{\rm total}.
%\label{eq:workload_utilization_tranlation}
%\end{equation}
%
%Models based on CPU utilization are the bases for DVFS algorithms. Even though the input is not controllable, it is very easy to measure system utilization with almost no overhead. It is also very portable. There is no need for training and it often works reasonably in any machine. A downside is that the theory behind the energy consumption optimization relies on the knowledge of the future workload \cite{Fu2018RaceMinimization, Group2012HandbookSahni}. This introduces the problem of how to estimate the workload.
%
%%This models can be forward subdivided into  schedules, ...
%
%\subsection{Online vs Offline models}

\section{Theoretical background} \label{sec:background}

A model is a formal representation of a real system. Computer system models can be represented in the form of equations, graphical models, rules, decision trees, representative collections of examples, or by neural networks, among others. The choice of representation affects the model's accuracy, as well as its interpretability by people~\cite{Hypothesis2012EncyclopediaLearning, Roy2019ForecastingNetwork, Zhu2019PredictingLearning}. Accurate energy and power consumption models are very important for many energy efficiency schemes employed in computing equipment \cite{Rivoire2007ModelsOptimizations}, and they can have multiple uses, including the design, forecasting, and optimization of data center systems. This work focuses on analytical models that could serve for energy optimization and subsequently, analysis of important factors in the total energy draw.

The desirable properties of a full-system model of energy consumption include accuracy (precision enough to allow the required energy saving), speed (fast enough to produce predictions), generality and portability (should be suitable for as many systems as possible), inexpensiveness (should not require costly or invasive infrastructure), and simplicity \cite{Rivoire2008AModels}. However, modeling the exact energy consumption behavior of an HPC system is not straightforward, either at the whole-system level or at the level of individual components. Patterns of energy consumption in data centers depend on multiple factors such as hardware specifications, workload, cooling requirements, or types of applications. Some of these factors cannot be measured easily. Furthermore, it is impractical to perform detailed measurements of the energy consumption of lower-level components without impacting the system with additional overload.

The concept of energy (E) is the total amount of work performed by a system over a period of time (T), while power (P) is the rate at which the system performs the work. The relation between these three amounts can be expressed as:

\begin{equation}
E = \int_{0}^{T}P(t)dt
\label{qe:energy_definition_cont}
\end{equation}

Several proposed models have already been classified with respect to its input parameters in \cite{Dayarathna2016DataSurvey}. In this survey, more than 200 models were analyzed according to their characteristics and limitations and classified into categories where the model is more suited to its objectives:
\begin{itemize}
	\item Temperature
	\item System Utilization or Workload
	\item Frequency
	\item Other system states such as cache miss, branch prediction, number of instructions executed, and more
	\label{tab:input_type}
\end{itemize}

Often, energy models are described as a combination of two main parts, the power model of the system and the performance model of the application.

\subsection{Power models}
Temperature-based models\cite{ZhangEnergyInfrastructure} can provide a way of estimating power consumption without being very intrusive. With this kind of model, it is possible to estimate energy with a single temperature sensor without the need to restart or introduce anything to the system. This is especially interesting for HPCs and data centers which, can not be easily turned off. Another use of this model is in the optimization of the refrigeration systems.

A large part of the developed models are based on parameters representing the state of the system. They normally take advantage of performance counters, which are provided by the CPU or the operating system and are capable of counting micro-architectural events, such as instructions executed, cache-hits, miss-predicted branches, and much more. 
This type of model is suitable for power estimation because it receives information about several internal states of the computer.

Frequency-based models serve as a base for many power models ~\cite{Sarwar1997, Butzen2007, Usman2013ANoC} because at a very low level every digital circuit (including modern processors) is composed of transistors. In these devices, there is a known relationship between power and frequency, which can be simplified by \cref{eq:power_simplified}: 
\begin{equation}
P = \alpha+\beta f^3
\label{eq:power_simplified}
\end{equation}
where $\alpha$ and $\beta$ are model parameters, and $f$ is the operating frequency (details of this equation are covered in Section~\ref{sec:models}). These types of models are suitable for optimization problems since they are a function of the operating frequency, which can be easily controlled.

\subsection{Performance models}
Another commonly used way to model the energy consumption of the system is the workload. The workload, which represents the amount of work done in a given time and speed, abstractly defines the application.

The workload ($W$) can be defined as~\cite{Paolillo2018OptimisationParallelism, Group2012HandbookSahni, Kim2015RacingHeuristics}:
\begin{equation}
W = \int_{0}^{\tau}s(t)dt = s\tau
\label{eq:workload_definition}
\end{equation}
where $\tau$ is total active time, and $s$ is the execution speed in instructions/second.
This equation is often taken up by utilization-based models, so that it could be applied to real systems.

Utilization can be defined as the ratio between the time that the system is active and the total time. Equation (\ref{eq:workload_utilization_tranlation}) defines workload in terms of CPU utilization ($u$):
\begin{equation}
u = \frac{\tau}{T} = \frac{W/s}{T},
\label{eq:utilization_definition}
\end{equation}
\begin{equation}
W = usT,
\label{eq:workload_utilization_tranlation}
\end{equation}
where $T$ the total execution time of the workload, and $s$ the execution speed.
Models based on CPU utilization are the base for DVFS algorithms. Even though the input is not controllable, it is straightforward  to measure system utilization with almost no overhead, and it is also very portable in terms of operating systems and architectures. 
%% Here is how to build the model or to optimize it, maybe too early to talk about AI without referencing existing approaches:
%There is no need for data and training, as with artificial intelligence algorithms, and adapts reasonably well to any machine. 
A downside is that optimization requires the knowledge of a wide variety of workloads which must be estimated somehow \cite{Fu2018RaceMinimization, Group2012HandbookSahni}.

% Markov model
Other approaches are also adopted to model energy, as for example in the work of Liu et al. \cite{Liu2016IntelligentSupplies} where the model uses a Markov chain to determine energy state transitions. This kind of approach needs heuristic algorithms to search for the best transitions and does not provide insights as an analytical equation does.

Although much work has been done in DVFS, the focus is still on the consumer electronics and laptop markets. 
For HPC, the notion of energy perception is relatively new~\cite{Beckman2005MakingSupercomputing}. 
Moreover, the operational characteristics of non-HPC and HPC systems are significantly different. 
First, the workload on non-HPC systems is very interactive with the end-user, but the workload on the HPC platform is not. 
Second, activities conducted on a non-HPC platform tend to share more machine resources.
In contrast, in HPC each job often runs with dedicated resources. 
Third, an HPC system usually is much larger than non-HPC systems, making it more  challenging to gather information, organize decisions, and execute global decisions.
Therefore, it is worthwhile to investigate whether a DVFS scheduling algorithm, which works well for conventional computing, remains effective for HPC.

\section{Related work} \label{sec:relatedwork}

Merkel et al.~\cite{Merkel2006BalancingSystems} developed an energy model for processors based on events. By counting a set of $n$ events and assuming that the processor consumes a fixed amount of energy $\alpha_i$ for each activity, they estimate the energy in a particular period as the number of occurrences $c_i$  multiplied by its correspondent weight $\alpha_i$:
\begin{equation}
E = \sum_{i=1}^{n}\alpha_ic_i.
\end{equation}

Another event-based model introduced by Roy et al. \cite{Roy2013AnAlgorithms}, described the computational energy consumed by a CPU for an algorithm $A$:
\begin{equation}
E(A) = P_{clk}T(A) + P_wW(A),
\end{equation}
where $P_{clk}$ is a processor clock leakage power, $T(A)$ is the total execution time, $W(A)$ is the total time taken by  non-I/O operations, and $P_w$ is used to capture the power consumption per operation performed by the CPU. $T(A)$ and $W(A)$ are estimated using performance features.

Models based on events are implemented using performance counters \cite{Fahad2019AComputing}, which have some drawbacks. They are highly dependent on the operating system and its architecture, making them not so easily portable. Additionally, there are limitations on the number of counters that can be read simultaneously. Furthermore, reading too many counters can add a non-negligible overhead. Thus, modern CPUs start to multiplex counters when more than a few are requested. There are also some well know problems regarding the precision of some events \cite{Weaver2008, Weaver2013a, Das2019SoK:Security, McGuire2009, Ramos2019AnCounters, Silva-De-Souza2020Containergy-aWorkloads}. Events that should be exact and deterministic (such as the number of executed instructions) show run-to-run variations and over-count on x86\_64 machines, even when running in strictly controlled environments. Our proposal is not based on performance counters and, therefore, is not vulnerable to those drawbacks.

Instruction-level energy model was also proposed in \cite{Characterization2013EnergyProcessor}, obtained from an energy per instruction (EPI) characterization made on Xeon Phi. Their model is expressed as:
\begin{equation}
E(f) = \frac{(p_1 - p_0)(c_1 - c_0)/f}{N}, 
\end{equation}
where $N$ is the total number of dynamic instructions, $p_0$ is the initial idle power, $p_1$ is the average dynamic power and ($c_1$-$c_0$) refers to the cumulative number of cycles the micro-benchmark performs. This model is suitable for estimating the energy after the application finishes executing, when it is possible to count the number of total cycles. However, it is challenging to use for optimization or forecasting since it does not have an application model to predict the number of cycles. Our model integrates the behavior of the application taking into account the execution time.

In \cite{Lewis2008Run-timeSystems} Lewis et al. described the overall system energy consumption using the following equation:
\begin{equation}
E = A_0(E_{proc} + E_{mem}) + A_1E_{em} + A_2E_{board} + A_3E_{hdd},    
\end{equation}
where, $A_0$, $A_1$, $A_2$, and $A_3$ are unknown constants that are calculated via linear regression analysis and those remain constant for a specific server architecture. This model, as the previous one, relies on knowledge of energy spent on each component, being  a suitable option for estimation after the application has already ran, but not for optimization of the run itself.

In another energy consumption model  based on system utilization, Mills et al. modeled the energy consumed by a compute node with CPU (single) executing at speed $\sigma$ as \cite{Mills2014EnergySystems},
\begin{equation}
E(\sigma,[t_1,t_2]) = \int_{t_1}^{t_2} \sigma^3 + \rho \sigma_max^3 dt,
\end{equation}

where $\rho$ stands for the overhead power  consumed regardless the processor speed, $t_1$ and $t_2$ are the initial and final execution time of the application. The overhead includes power consumption by all other system components such as memory, network, etc. For this reason, although the authors mentioned the energy consumption of a socket, their power model is generalized to the entire server.

Our work proposes a full-system energy model based on the CPU frequency and the number of cores.
The aim of the model is to understand and optimize the energy behavior of parallel applications in HPC systems according to application parameters such as the degree of parallelism and CPU parameters related to dynamic and static power. The proposed model differs from existing ones for including the frequency and number of cores in the same equation for estimating the energy for a specific application in a given configuration. This model can serve as a base for DVFS and DPM optimization problems that include frequency and number of active cores. It can also be used to analyze the contribution of each parameter (ex: level of parallelism) to energy consumption. The number of cores is an essential factor in HPC since applications are designed to run on multiple cores.

The proposed energy model is the product of an application-agnostic power model and an architecture-specific application performance model. The power model is based on the CMOS logic gates power draw as a function of the frequency~\cite{Sarwar1997, Butzen2007} augmented to include the number of cores. The performance model is based on Amdahl's law \cite{Amdahl1967ValidityCapabilities, Eyerman2010ModelingDesign, Shi2015ReevaluatingLaw}, which can be used to estimate runtime in multi-core systems. This model has been extended to include execution frequency and input size, characterizing the application on the target architecture.

\subsection{The contribution of this work}
Although a lot of work has been done in DVFS, the focus is still mainly on consumer electronics and laptop markets. For HPC, the notion of energy perception~\cite{Beckman2005MakingSupercomputing} is relatively new. Moreover, the operational characteristics of non-HPC and HPC systems are noticeably different. First, the workload on non-HPC systems is very interactive with the end-user, but the workload on the HPC platforms is not. Second, activities conducted on a non-HPC platform tend to share more resources of the machine. On the other hand, often in HPC, each job runs with dedicated resources. Third, an HPC system is normally much larger than non-HPC systems, which makes collecting information, organizing decisions, and executing global decisions more challenging. Therefore, it is worthwhile to investigate whether a DVFS scheduling algorithm, which works well for conventional computing, remains effective for HPC.

This work proposes a full-system energy model based on the CPU frequency and the number of cores. The goal of the model is to understand and optimize the energy behavior of parallel applications in HPC systems according to the application parameters such as the degree of parallelism and CPU parameters that relates to dynamic and static power. The proposed model differs from existing ones for including the frequency and number of cores in the same equation for estimating the energy for a specific application in a given configuration. This model can serve as a base for DVFS and DPM optimization problems that include both frequency and number of active cores. It can also be used to analyze the contribution of each parameter (ex: level of parallelism) to the energy consumption. The number of cores an important factor in HPC since applications are designed to run on multiple cores.

The proposed energy model is the product of an application-agnostic power model and an architecture-specific application performance model. The power model is based on the CMOS logic gate power draw in the function of the frequency~\cite{Sarwar1997, Butzen2007} augmented to include the number of cores. The performance model is based on Amdahl's law \cite{Amdahl1967ValidityCapabilities, Eyerman2010ModelingDesign, Shi2015ReevaluatingLaw}, which can be used to estimate execution time in multi-core systems. We have extended  this model to include the frequency and input size, characterizing the application on the target architecture.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\subsection{Goals of the research}
%
%The interest in developing this project is to make HPC increasingly viable, in addition to reducing maintenance costs and reducing the impact on the environment by saving energy.
%
%Apart from this, our research also has international collaborations with other high-performance computing centers such as the Federal University of Rio Grande do Norte, where the project was started, and others which I had collaborated with projects related to energy saving in the University of Bristol, and possible future collaborations with University of Lausanne on Switzerland.
%
%The main contribution of this thesis will be to produce better energy saving techniques for HPC based on dynamic models of applications’ performance, power/energy consumption analysis and extended DVFS techniques. As will be detailed in the next section, several objectives have been defined to address the multiple challenges of our proposal: consideration of the processor workload, instruction set and the input size of applications, more precise energy models based on additional execution parameters, alternative optimization strategies and automatization of the execution process.
%
%First experiments developed showed that this methodology has a potential and can be improved to develop software as an end product that can be used by servers and clouds to save energy. This search can also be easily expanded to encompass mobile devices such as smartphones or microprocessed devices like raspberry-pi and beaglebone. Contributions in the areas of power consumption modeling and algorithms for frequency and voltage control off the processor are the product of this research.
%
%\subsection{Work plan}
%
%Important points that will be addressed in this research:
%\begin{enumerate}
%	\item \textbf{Identify and analyzing important performance features for energy consumption}. Analyzing the application execution with multiple inputs on different configurations of active cores and frequencies, performance parameters and its relations can be chosen and associated to characterize the application. Since there is a large number of performance parameters, a study to select the principal components is necessary to reduce its complexity. 
%	\item \textbf{Splitting the application into optimizable execution steps}. The goal is to split the application execution into several stages and minimize the energy on each stage, for that, the performance model is going to be completely remodeled. Creating a complete profile of the application, including performance parameters, frequency, and a number of active cores. With this, we can Cluster the applications identifying common frequency scaling stages to optimize. So when the application is running we can collect these performance features and find the closest cluster that will already have the optimized configuration.
%	\item \textbf{Enhanced power models for HPC architectures: The power model is going to include additional parameters, such as the percentage of CPU utilization}. Further analysis of additional parameters, such as the influence of the instructions to be executed on the power consumed. Developing models that can reduce the overhead of training fitting and use fewer samples.
%	\item \textbf{Enhanced DVFS models for HPC architectures: Including new architectures with distributed memory and heterogeneous systems}. Also studied the viability of using mathematical and artificial intelligence models, their advantages and their trade-offs to produce the final model. With these tools develops we can extend the current DVFS algorithms and produce a Smart DVFS module for HPC.
%\end{enumerate}
%
%The new workflow that we want to implement can be illustrated on the figures \ref{fig:new_model} and \ref{fig:heterogeneous_systems}.
%
%\begin{figure}[h]
%	\centering
%	\includegraphics[width=\columnwidth]{intro/figures/new_model.png}
%	\caption{New model}
%	\label{fig:new_model}
%\end{figure}
%
%\begin{figure}[h]
%	\centering
%	\includegraphics[width=\columnwidth]{intro/figures/heterogeneous_systems.png}
%	\caption{Heterogeneous systems}
%	\label{fig:heterogeneous_systems}
%\end{figure}

### 2021-03-23 thesis/intro/main.tex
add tex files

\documentclass[a4paper]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}

\usepackage{float}
\usepackage{subfig}
\usepackage{booktabs}
\usepackage{tabu}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
%\usepackage{subcaption}
%\captionsetup{compatibility=false}

%\usepackage{apacite}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\title{Full name of the applicant: Aspirant [ASP]
SEMAPHORE Application ID: 34791278}
\author{Energy-Optimal Configurations for HPC Applications - Vitor Ramos Gomes da Silva}
\date{\today}

\begin{document}
\maketitle

\input{paper}

\bibliographystyle{plain}
\bibliography{refs}

\end{document}

### 2021-03-23 thesis/intro/paper.tex
add tex files

Keywords: High performance computing, Energetic efficient systems, Power Modeling, Performance Modeling

\section{Description of the research project}

One of the main concerns regarding the construction and use of high-performance computing (HPC) systems is energy consumption. Currently, medium-size systems can consume hundreds of Kilowatts while high-end systems can reach dozens of Megawatts as we shown in figure \ref{fig:pw_consumption_a}. Maintain these systems are very expensive and depending on the energy resources they use it can also cause a huge Environmental impact. Also The next generation of HPC systems, the Exascale computing, capable of reaching a billion billion calculations per second, will be strongly constrained by their energy consumption as shown by the trend of the figure \ref{fig:pw_consumption_b}. Investigating new ways to reduce the energy consumption on HPC is crucial.

\begin{figure}[h]
    \centering
    \subfloat[Power consumption top 500 systems]{{
    \includegraphics[width=0.45\textwidth]{figures/power_rank.png} 
    \label{fig:pw_consumption_a}
    }}%
    \qquad
    \subfloat[Power by computational power]{{
    \includegraphics[width=0.45\textwidth]{figures/power_tflops.png}
    \label{fig:pw_consumption_b}
    }}%
    \caption{Power consumption}%
    \label{fig:pw_consumption}%
\end{figure}

Analyzing the contribution to the server power consumption we can see from the Google research \cite{Fan2007, Barroso2007} and also on my research at Federal University of Rio Grande do Norte on figure \ref{fig:cpu_util} that the processor is responsible for about 45\% of the total server power draw at peak usage being the main component responsible for the server power draw. Reducing the power consumption of the processor is very important in order to reduce the overall energy consumption.

\begin{figure}[h]
    \centering
    \subfloat[Power consumption top 500 systems]{{
    \includegraphics[width=0.45\textwidth]{figures/cpu_util.png}
    \label{fig:cpu_util_a}
    }}%
    \qquad
    \subfloat[Power by computational power]{{
    \includegraphics[width=0.45\textwidth]{figures/power_share.png}
    \label{fig:cpu_util_b}
    }}%
    \caption{Power draw on HPC}%
    \label{fig:cpu_util}%
\end{figure}

Modern processors incorporate several features for power management such as independent processing cores that can be enabled/disabled by the operating system \cite{Rotem2012}, clock gating techniques for reducing the dynamic power dissipation of synchronous circuits \cite{Srinivasan2015} and Dynamic Voltage and Frequency Scaling (DVFS) \cite{Mittal2014}. Besides that, there is also architectural approaches to reduce energy consumption including the use of simple but more power efficient processors, such as the ARM architecture used on smartphones, and also specialized architectures such as GPUs and FPGAs.

\subsection{Goals of the research}

The interest in developing this project is to make HPC increasingly viable, in addition to reducing maintenance costs and reducing the impact on the environment by saving energy.

Apart from this, our research also has international collaborations with other high-performance computing centers such as the Federal University of Rio Grande do Norte, where the project was started, and others which I had collaborated with projects related to energy saving in the University of Bristol, and possible future collaborations with University of Lausanne on Switzerland.

The main contribution of this thesis will be to produce better energy saving techniques for HPC based on dynamic models of applications’ performance, power/energy consumption analysis and extended DVFS techniques. As will be detailed in the next section, several objectives have been defined to address the multiple challenges of our proposal: consideration of the processor workload, instruction set and the input size of applications, more precise energy models based on additional execution parameters, alternative optimization strategies and automatization of the execution process.

First experiments developed showed that this methodology has a potential and can be improved to develop software as an end product that can be used by servers and clouds to save energy. This search can also be easily expanded to encompass mobile devices such as smartphones or microprocessed devices like raspberry-pi and beaglebone. Contributions in the areas of power consumption modeling and algorithms for frequency and voltage control off the processor are the product of this research.

\subsection{State of art}

DVFS has been demonstrated to be a very effective technique for reducing the power consumption of processors \cite{Hackenberg2015, Dzhagaryan2014, Hahnel2012, Basmadjian2012, Travers2015, Miyoshi2002, Anghel2011, Pietri2014}. The technique tries to optimize power consumption by adjusting the frequency according to the current load of the processor. Generally, the frequency scales with the intensity of the load and the voltage scales to the minimum value that enables the selected frequency. Among other aspects, DVFS helps reducing energy consumption because it allows memory-bounded programs to be executed more efficiently \cite{Spiga2006}. Nonetheless, aspects such as load variability may compromise the effectiveness of DVFS. Another important aspect that is typically not taken into account in DVFS is the number of processing cores to be used by a parallel program.

The DVFS technique has been extensively researched with the aim of providing strategies for selecting the optimal voltage and frequency for a specific application and architecture. In \cite{Anghel2011}, the authors used two algorithms for scaling the frequency of the processors: a human-immune system inspired algorithm to monitor the server's power and performance states, and fuzzy logic based algorithm for changing the server's performance state. \cite{Cochran2011} introduced a scaling method for determining the system's optimal operation points for the number of threads and DVFS settings. In \cite{DaCosta2015}, an approach that considers instantaneous system activity states was proposed. In this case, the memory and network activity were used to generate a DVFS management setting.

%In \cite{Georgiou:2017}, the authors used an energy model for a multi-threaded, multi-core embedded architecture and static resources analysis to statically evaluate the energy and timing savings of various DVFS configurations for the same program. Although, they were able to identify the most optimal configuration without the need of measuring time and energy of the program executed on each configuration alternative, this approach is quite limited, as static analysis does not scale to less time-predictable architectures and programs.

The use of energy measurements on the DVFS is stimulated by the large number of performance counters including on modern architectures providing a lot of debugging information on real-time including energy and power consumption information. These counters are accessed through the operating system via several interfaces.
Power models use to estimate power consumption of the CPU are based on the logic gates \cite{Du2017, Butzen2007}.  A logic gate consumes power mainly when it's changing state, and the frequency with which it changes state will probably be, on average, proportional to the clock frequency. In addition to leak current due to imperfection on the manufacture, basic models are generally:

\begin{equation}
    P_{cpu}=P_{dynamic}+P_{leak}+P_{static}
\end{equation}
and
\begin{equation}
    P_{dynamic} \quad \alpha \quad CV^2
\end{equation}

Where V is the voltage and f the operating frequency.

\subsection{Research project}

\subsubsection{Preliminary Work}

This proposal is based on my previous work on this topic as part of a master thesis where a develop a methodology to find optimal parameters for energy-frequency relations and the number of active cores needed to run single-node HPC applications using an application-agnostic power model of the architecture combined with an architecture-aware performance model of the application.

The power model used is based on CMOS (Complementary Metal-Oxide-Semiconductors) modeling, which are the intrinsic components of CPU chips. According to the operating frequency~\cite{Sarwar1997}, it models the dynamic, static and leakage power of the circuit. Besides operating frequency, the power model also parametrizes the number of active sockets and the number of active cores per-socket on a multi-core architecture:

\begin{equation}
    P(f,p,n)=(af^3+bf)p+cs+d
    \label{eq_power_final}
\end{equation}

Where $f$ is the frequency, $p$ is the number of active cores, $s$ the number of sockets on the system and $a$, $b$, $c$ and $d$ are constants.

% To fit the power-model equation, the CPU was stressed up to 100\% and power information was acquired on regular intervals. The coefficients of (\ref{eq_power_final}), $a$, $b$, $c$ and $d$, were found by performing multi-linear regression. The figure below shows the resulting fit for the system.

% \begin{figure}[H]
%     \centering
%     \includegraphics{image6.png}
%     \caption{Caption}
%     \label{fig:my_label}
% \end{figure}

The Performance is modeled by characterizing the application on the target architecture. The idea is to predict the performance of the application at any given configuration. The model takes as inputs the operating frequency, the number of active cores and the input size. The modeling is done using a supervised learning method for regression called Support Vector Regression (SVM)~\cite{Ventura2009, Smola2004} with data collected from multiple runs for different numbers of active cores, frequencies, and input sizes.

To find optimal-energy configurations, we can minimizes the product of outcomes of the power and performance models. Our approach was validated on four PARSEC applications~\cite{Bienia2008} and compared to the \emph{Ondemand} governor, which is the default DVFS scheme for the Linux operating system. The figure \ref{fig:previous_model} illustrates the entire process.

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{figures/old_model.png}
    \caption{Previous model}
    \label{fig:previous_model}
\end{figure}

% The figure below shows the energy equation for one specific input on one of the applications tested:
% \begin{figure}[H]
%     \centering
%     \includegraphics{image5.png}
%     \caption{Caption}
%     \label{fig:image5}
% \end{figure}

%This experiments was performed at High Performance Computing Center at the UFRN (NPAD/UFRN) using compute nodes that consists of two Intel Xeon E5-2698 v3 processors with sixteen cores each and two hardware threads for each core. The maximum non-turbo frequency is 2.2GHz, and the total physical memory of the node is 128GB (8$\times$16GB). The operating system used is Linux CentOS 6.5, kernel 4.16.9 Turbo frequency and hardware multi-threading were disabled during all experiments.

%On the figure  \ref{fig:image5} we compare the proposed energy model with the real measured values.

% The steps need for this technique is:
% \begin{enumerate}
%     \item Run the power model fit once.
%     \item For each new application, run the performance model process.
%     \item Before the application start executing, use the power model and the performance model to compute the optimal energy configuration.
%     \item Set the frequency and disable cores according to the optimal values.
% \end{enumerate}

% All these steps can be performed without user intervention. The first step can be executed once the operating system is installed. The second step can be part of the installation process of the application, where the developer includes a configuration of how the application should be run and the operating system performs the process. The third and fourth are part of the operating system as a module.

The use of an application-agnostic power modeling of the target architecture helps to make the technique portable to other applications. That is, to estimate the energy-optimal frequency and number of active cores for a new application, only a performance characterization is needed. In comparison with methods that only use information of processor workload, like the Linux default governor Ondemand, our technique has the disadvantage of including additional steps of performance characterizations. However, our results show that we are able to find configurations using about 14$\times$ less energy when compared to the worst case of the Ondemand governor. When compared to the best case of this DVFS scheme, i.e. when the user guesses the optimal number of cores to be used, the proposed approach was able to find configurations using up to 23\% less energy to run the target application. The overall average energy saving reached 6\% for the proposed approach when compared to the best case.

% \begin{figure}[H]
%     \centering
%     \includegraphics{image4.png}
%     \caption{Caption}
%     \label{fig:image4}
% \end{figure}

%The figure \ref{fig:image4} show the relative energy consumption for one application comparing to the Ondemand governor to multiples inputs. The application run on Ondemand with different number of cores.
The success obtained with this approach is probably due to the fact that the prior knowledge of the application's performance on the target architecture can expose sufficiently relevant information, such as parallel speedups, which are more difficult to guess with runtime techniques based on DVFS.

\subsubsection{Work continuation and motivation}

Although DVFS helps reduce system consumption, there are still cases where it is not effective or could be better. When there are abrupt variations in processor workload,  or when reducing frequency does not lead to less energy consumption, due to the program need to run for longer. This could be solved with prior knowledge of the application and power consumption of the system that we propose.

The currently proposed method still has some limitations: it optimizes the execution for only one application at a time, the execution frequency must be set at the beginning of the program execution, and the input used by the program must be known. In addition, the configuration mechanisms described in the previous section require a strategic integration with the operating system in order to be able to execute them automatically.

The input size can be estimated, allowing to consider alternative strategies when used as an optimization parameter. Combined with the previous model, that estimation will allow to split applications into execution stages and to apply the best optimization strategy at each stage. However, that estimation implies the identification of useful measurable parameters, such as CPU utilization, Cache Hit/Miss or memory utilization.

There are still other alternatives to consider that deserve our attention. The previously proposed approach does not include frequency variations when running the application. Therefore, if the mathematical formulation can include this, it will be possible to find better results. Another parameter still not taken into account is that of the processor workload, which can bring a significant improvement to the power model. In addition to the inclusion of new parameters to the models, other models can still be proposed on mathematical formulation form or using, as in the case of SVR, learning methods. However, all this will require further analysis of the optimization alternatives that will appear in the benchmark tests as well as the study of the limits of each approach.

\subsection{Work plan}

Important points that will be addressed in this research:
\begin{enumerate}
    \item \textbf{Identify and analyzing important performance features for energy consumption}. Analyzing the application execution with multiple inputs on different configurations of active cores and frequencies, performance parameters and its relations can be chosen and associated to characterize the application. Since there is a large number of performance parameters, a study to select the principal components is necessary to reduce its complexity. 
    \item \textbf{Splitting the application into optimizable execution steps}. The goal is to split the application execution into several stages and minimize the energy on each stage, for that, the performance model is going to be completely remodeled. Creating a complete profile of the application, including performance parameters, frequency, and a number of active cores. With this, we can Cluster the applications identifying common frequency scaling stages to optimize. So when the application is running we can collect these performance features and find the closest cluster that will already have the optimized configuration.
    \item \textbf{Enhanced power models for HPC architectures: The power model is going to include additional parameters, such as the percentage of CPU utilization}. Further analysis of additional parameters, such as the influence of the instructions to be executed on the power consumed. Developing models that can reduce the overhead of training fitting and use fewer samples.
    \item \textbf{Enhanced DVFS models for HPC architectures: Including new architectures with distributed memory and heterogeneous systems}. Also studied the viability of using mathematical and artificial intelligence models, their advantages and their trade-offs to produce the final model. With these tools develops we can extend the current DVFS algorithms and produce a Smart DVFS module for HPC.
\end{enumerate}

The new workflow that we want to implement can be illustrated on the figures \ref{fig:new_model} and \ref{fig:heterogeneous_systems}.

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{figures/new_model.png}
    \caption{New model}
    \label{fig:new_model}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{figures/heterogeneous_systems.png}
    \caption{Heterogeneous systems}
    \label{fig:heterogeneous_systems}
\end{figure}

%\begin{figure}[h]
%    \centering
%    \includegraphics[width=\columnwidth]{figures/proposal.png}
%    \caption{Caption}
%    \label{fig:my_label}
%\end{figure}

% \newpage

% \section{Attachment}

% The figure below shows the resulting fit of the power model. To fit the power-model equation, the CPU was stressed up to 100\% and power information was acquired on regular intervals.

% \begin{figure}[H]
%     \centering
%     \includegraphics[width=10cm]{image6.png}
%     \caption{Power model fitting}
%     \label{fig:my_label}
% \end{figure}

% The energy model for one specific input for one of the applications tested is shown on the figure bellow
% \begin{figure}[H]
%     \centering
%     \includegraphics[width=10cm]{image5.png}
%     \caption{Energy model}
%     \label{fig:image5}
% \end{figure}

% The figure \ref{fig:image4} show the relative energy consumption for one application comparing to the Ondemand governor to multiples inputs. The application run on Ondemand with different number of cores.

% \begin{figure}[H]
%     \centering
%     \includegraphics{image4.png}
%     \caption{Comparison of the energy consumption of the proposed model to the Ondemand}
%     \label{fig:image4}
% \end{figure}

\section{Comments on changes made in the research project in case of resubmission(optional)}

Since the last submission, we have time to developed the first part of the project that was to collect performance counters a design a new application model that has a more precise definition of input size. The application architecture is shown on the figure \ref{fig:my_api}.

% \begin{figure}[h]
%     \centering
%     \includegraphics[height=7cm]{figures/architecture.png}
%     \caption{Tool developed to collect performance counters}
%     \label{fig:my_api}
% \end{figure}

\begin{figure}
    \subfloat[Architecture]{{
    \includegraphics[width=0.45\textwidth]{figures/architecture.png}
    \label{fig:my_api}
    }}%
    \qquad
    \subfloat[Clusters]{{
    \includegraphics[width=0.45\textwidth]{figures/graph_input_size2.png}
    \label{fig:sping_force_is}
    }}%
    \caption{Tool to cluster applications}
\end{figure}

We develop a complete workflow to measure and prepossess the performance counters, the steps include running multiple times, remove outlines, interpolation, and filtering, we can see an example application on the figure \ref{fig:workflow}.

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/workflow.png}
    \caption{Workflow}
    \label{fig:workflow}
\end{figure}

Finally we developed a new input size defined as:

\begin{equation}
I_{sz}=\frac{I_t}{I_m}*A
\end{equation}

Where $I_{t}$ is the total number of instructions, $I_m$ is the total number of memory instructions and A is the area of the original curve. This new input size can completely characterize the application execution and works as a fingerprint of the program so that we can proceed to the next step that is the clusterization.

On the figure, \ref{fig:sping_force_is} we display the clusters results that we obtain by comparing the input size curve of the PolyBench benchmark. An article describing the details of this first contribution was already submited to a conference and is waiting for approval.

\section{Activities report on the first year of doctorate}
ONLY for “1st grant - 2nd year” applicants

\section{Potential interdisciplinary approach of the research project (optional)}
This project involves Parallel and distributed computing, Operating system, Computer architecture, Programming, Mathematics, Physics, and Intelligent systems.

\section{Description of the work environment}
This research is developed in two HPC centers one at the host University, Faculte Polytechnique de Mons with a collaboration of the HPC center at Federal University of Rio Grande do Norte on Brazil (NPAD/UFRN). The environment is a supercomputer with multiple nodes, with 64 cores each, running a Linux system.

\section{Summary of the master’s thesis or equivalent}

This thesis proposes a methodology to minimize energy consumption in high-performance computing, finding the ideal frequency parameters and active cores in the processor. Based on a power modeling of the system, based on the power consumption of the CMOS circuit and an estimation of the execution time of the application using artificial intelligence, the energy equation is calculated as a function of frequency, number of cores and input of the application. With this equation, the configuration that minimizes the total power consumption is found. Tests were run on four applications and the results of this method showed that this approach achieved savings of up to 30\% when compared to the current energy-saving scheme on Linux.

\section{Additional comments (optional)}

\section{Ph.D. work calendar per month}

The above-mentioned objectives will be developed by following a structured work plan that will consist of the use of benchmarks, data collection and analysis, models proposal, experimental comparisons, and optimization steps.

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{figures/workplan.jpg}
    \caption{Work plan}
    \label{fig:workplan}
\end{figure}

### 2021-03-23 thesis/main.tex
add tex files

% Stanford University PhD thesis style -- modifications to the report style
% This is unofficial so you should always double check against the
% Registrar's office rules
% See http://library.stanford.edu/research/bibliography-management/latex-and-bibtex
% 
% Example of use below
% See the suthesis-2e.sty file for documentation
%
\documentclass{report}

\usepackage[breakable]{tcolorbox}
\usepackage{parskip} % Stop auto-indenting (to mimic markdown behaviour)

\usepackage{iftex}
\ifPDFTeX
\usepackage[T1]{fontenc}
\usepackage{mathpazo}
\else
\usepackage{fontspec}
\fi

% Basic figure setup, for now with no caption control since it's done
% automatically by Pandoc (which extracts ![](path) syntax from Markdown).
\usepackage{graphicx}
% Maintain compatibility with old templates. Remove in nbconvert 6.0
\let\Oldincludegraphics\includegraphics
% Ensure that by default, figures have no caption (until we provide a
% proper Figure object with a Caption API and a way to capture that
% in the conversion process - todo).
\usepackage{caption}
\DeclareCaptionFormat{nocaption}{}
\captionsetup{format=nocaption,aboveskip=0pt,belowskip=0pt}

\usepackage[Export]{adjustbox} % Used to constrain images to a maximum size
\adjustboxset{max size={0.9\linewidth}{0.9\paperheight}}
\usepackage{float}
\floatplacement{figure}{H} % forces figures to be placed at the correct location
\usepackage{xcolor} % Allow colors to be defined
\usepackage{enumerate} % Needed for markdown enumerations to work
\usepackage{geometry} % Used to adjust the document margins
\usepackage{amsmath} % Equations
\usepackage{amssymb} % Equations
\usepackage{textcomp} % defines textquotesingle
% Hack from http://tex.stackexchange.com/a/47451/13684:
\AtBeginDocument{%
	\def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
}
\usepackage{upquote} % Upright quotes for verbatim code
\usepackage{eurosym} % defines \euro
\usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
\usepackage{fancyvrb} % verbatim replacement that allows latex
\usepackage{grffile} % extends the file name processing of package graphics 
% to support a larger range
\makeatletter % fix for grffile with XeLaTeX
\def\Gread@@xetex#1{%
	\IfFileExists{"\Gin@base".bb}%
	{\Gread@eps{\Gin@base.bb}}%
	{\Gread@@xetex@aux#1}%
}
\makeatother

% The hyperref package gives us a pdf with properly built
% internal navigation ('pdf bookmarks' for the table of contents,
% internal cross-reference links, web links for URLs, etc.)
\usepackage{hyperref}
% The default LaTeX title has an obnoxious amount of whitespace. By default,
% titling removes some of it. It also provides customization options.
\usepackage{titling}
\usepackage{longtable} % longtable support required by pandoc >1.10
\usepackage{booktabs}  % table support for pandoc > 1.12.2
\usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
\usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
% normalem makes italics be italics, not underlines
\usepackage{mathrsfs}
\usepackage{outlines}

% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{nth}
\usepackage{todonotes}
\usepackage{url}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{float}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{listings}
%\usepackage[caption=false,font=footnotesize]{subfig}
%\usepackage[export]{adjustbox}

\newcommand*\rot{\rotatebox{90}}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
		T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\usepackage{suthesis-2e}
\dept{Computer Science}

\begin{document}
\title{Energy aware HPC}
\author{Vitor Ramos Gomes da Silva}
\principaladviser{Carlos Valderrama}
\coprincipaladviser{Pierre MANNEBACK}
\firstreader{Thierry DUTOIT}
\secondreader{Sidi MAHMOUDI}
\thirdreader{Samuel Xavier de Souza} %if needed

\beforepreface
\prefacesection{Preface}
Second Phd Committee
%\prefacesection{Acknowledgments}
%I would like to thank...
\afterpreface

\chapter{Introduction}
Energy consumption is a key to enable exascale High-performance Computing (HPC). Energy-optimized hardware and software combinations still could be inefficient if software operates poorly. 
Software operation relies on dynamic scaling of frequency and voltage (DVFS) as well as dynamic power management (DPM), but none have priority information on the application, which leads to inefficient software operation. This work proposes an analytical model that can be used to estimate energy-optimal software operation. 
This work proposes a new energy model that can be used as a basis to optimize DVFS and DPM, as well as to assess the contribution of parameters such as parallelism level and dynamic power to the energy consumption.
Since in HPC, applications can run on multiple cores and associated frequencies, both of which can be controlled for most systems, the novelty of the model includes such parameters as well as the workload, two application parameters and three system parameters.
\include{intro/intro}

%\chapter{Approch}
%\include{ch1/ch1}

\chapter{Models}
In this Section, the proposed power, performance and energy models are derived and validated
\include{models/models}
\include{ch5/ch5}

\chapter{Optmizers}
This chapter talks about the different types of optimizers.
\include{ch7/ch7}

\chapter{Framework}
This chapter introduces the framework developed, the fingerprint tool and pascal analyzer.
\include{fingerprint/fingerprint}
\include{pascalanalyzer/pascalanalyzer}

\chapter{Experiments}
This chapter will describe all the experiments that validate the models constraints.
\include{experiments/experiments}

%\appendix
%\chapter{A Long Proof}

\bibliographystyle{plain}
\bibliography{references}
\end{document}

### 2021-03-23 thesis/pascalanalyzer/DataStructure2.tex
add tex files

\documentclass[11pt]{article}

    \usepackage[breakable]{tcolorbox}
    \usepackage{parskip} % Stop auto-indenting (to mimic markdown behaviour)

    \usepackage{iftex}
    \ifPDFTeX
    	\usepackage[T1]{fontenc}
    	\usepackage{mathpazo}
    \else
    	\usepackage{fontspec}
    \fi

    % Basic figure setup, for now with no caption control since it's done
    % automatically by Pandoc (which extracts ![](path) syntax from Markdown).
    \usepackage{graphicx}
    % Maintain compatibility with old templates. Remove in nbconvert 6.0
    \let\Oldincludegraphics\includegraphics
    % Ensure that by default, figures have no caption (until we provide a
    % proper Figure object with a Caption API and a way to capture that
    % in the conversion process - todo).
    \usepackage{caption}
    \DeclareCaptionFormat{nocaption}{}
    \captionsetup{format=nocaption,aboveskip=0pt,belowskip=0pt}

    \usepackage[Export]{adjustbox} % Used to constrain images to a maximum size
    \adjustboxset{max size={0.9\linewidth}{0.9\paperheight}}
    \usepackage{float}
    \floatplacement{figure}{H} % forces figures to be placed at the correct location
    \usepackage{xcolor} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{textcomp} % defines textquotesingle
    % Hack from http://tex.stackexchange.com/a/47451/13684:
    \AtBeginDocument{%
        \def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
    }
    \usepackage{upquote} % Upright quotes for verbatim code
    \usepackage{eurosym} % defines \euro
    \usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics 
                         % to support a larger range
    \makeatletter % fix for grffile with XeLaTeX
    \def\Gread@@xetex#1{%
      \IfFileExists{"\Gin@base".bb}%
      {\Gread@eps{\Gin@base.bb}}%
      {\Gread@@xetex@aux#1}%
    }
    \makeatother

    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    % The default LaTeX title has an obnoxious amount of whitespace. By default,
    % titling removes some of it. It also provides customization options.
    \usepackage{titling}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2
    \usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
    \usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
                                % normalem makes italics be italics, not underlines
    \usepackage{mathrsfs}

    % Colors for the hyperref package
    \definecolor{urlcolor}{rgb}{0,.145,.698}
    \definecolor{linkcolor}{rgb}{.71,0.21,0.01}
    \definecolor{citecolor}{rgb}{.12,.54,.11}

    % ANSI colors
    \definecolor{ansi-black}{HTML}{3E424D}
    \definecolor{ansi-black-intense}{HTML}{282C36}
    \definecolor{ansi-red}{HTML}{E75C58}
    \definecolor{ansi-red-intense}{HTML}{B22B31}
    \definecolor{ansi-green}{HTML}{00A250}
    \definecolor{ansi-green-intense}{HTML}{007427}
    \definecolor{ansi-yellow}{HTML}{DDB62B}
    \definecolor{ansi-yellow-intense}{HTML}{B27D12}
    \definecolor{ansi-blue}{HTML}{208FFB}
    \definecolor{ansi-blue-intense}{HTML}{0065CA}
    \definecolor{ansi-magenta}{HTML}{D160C4}
    \definecolor{ansi-magenta-intense}{HTML}{A03196}
    \definecolor{ansi-cyan}{HTML}{60C6C8}
    \definecolor{ansi-cyan-intense}{HTML}{258F8F}
    \definecolor{ansi-white}{HTML}{C5C1B4}
    \definecolor{ansi-white-intense}{HTML}{A1A6B2}
    \definecolor{ansi-default-inverse-fg}{HTML}{FFFFFF}
    \definecolor{ansi-default-inverse-bg}{HTML}{000000}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Additional commands for more recent versions of Pandoc
    \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
    \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
    \newcommand{\ImportTok}[1]{{#1}}
    \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
    \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
    \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
    \newcommand{\BuiltInTok}[1]{{#1}}
    \newcommand{\ExtensionTok}[1]{{#1}}
    \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
    \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
    \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatibility definitions
    \def\gt{>}
    \def\lt{<}
    \let\Oldtex\TeX
    \let\Oldlatex\LaTeX
    \renewcommand{\TeX}{\textrm{\Oldtex}}
    \renewcommand{\LaTeX}{\textrm{\Oldlatex}}
    % Document parameters
    % Document title
    \title{DataStructure2}

% Pygments definitions
\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % For linebreaks inside Verbatim environment from package fancyvrb. 
    \makeatletter
        \newbox\Wrappedcontinuationbox 
        \newbox\Wrappedvisiblespacebox 
        \newcommand*\Wrappedvisiblespace {\textcolor{red}{\textvisiblespace}} 
        \newcommand*\Wrappedcontinuationsymbol {\textcolor{red}{\llap{\tiny$\m@th\hookrightarrow$}}} 
        \newcommand*\Wrappedcontinuationindent {3ex } 
        \newcommand*\Wrappedafterbreak {\kern\Wrappedcontinuationindent\copy\Wrappedcontinuationbox} 
        % Take advantage of the already applied Pygments mark-up to insert 
        % potential linebreaks for TeX processing. 
        %        {, <, #, %, $, ' and ": go to next line. 
        %        _, }, ^, &, >, - and ~: stay at end of broken line. 
        % Use of \textquotesingle for straight quote. 
        \newcommand*\Wrappedbreaksatspecials {% 
            \def\PYGZus{\discretionary{\char`\_}{\Wrappedafterbreak}{\char`\_}}% 
            \def\PYGZob{\discretionary{}{\Wrappedafterbreak\char`\{}{\char`\{}}% 
            \def\PYGZcb{\discretionary{\char`\}}{\Wrappedafterbreak}{\char`\}}}% 
            \def\PYGZca{\discretionary{\char`\^}{\Wrappedafterbreak}{\char`\^}}% 
            \def\PYGZam{\discretionary{\char`\&}{\Wrappedafterbreak}{\char`\&}}% 
            \def\PYGZlt{\discretionary{}{\Wrappedafterbreak\char`\<}{\char`\<}}% 
            \def\PYGZgt{\discretionary{\char`\>}{\Wrappedafterbreak}{\char`\>}}% 
            \def\PYGZsh{\discretionary{}{\Wrappedafterbreak\char`\#}{\char`\#}}% 
            \def\PYGZpc{\discretionary{}{\Wrappedafterbreak\char`\%}{\char`\%}}% 
            \def\PYGZdl{\discretionary{}{\Wrappedafterbreak\char`\$}{\char`\$}}% 
            \def\PYGZhy{\discretionary{\char`\-}{\Wrappedafterbreak}{\char`\-}}% 
            \def\PYGZsq{\discretionary{}{\Wrappedafterbreak\textquotesingle}{\textquotesingle}}% 
            \def\PYGZdq{\discretionary{}{\Wrappedafterbreak\char`\"}{\char`\"}}% 
            \def\PYGZti{\discretionary{\char`\~}{\Wrappedafterbreak}{\char`\~}}% 
        } 
        % Some characters . , ; ? ! / are not pygmentized. 
        % This macro makes them "active" and they will insert potential linebreaks 
        \newcommand*\Wrappedbreaksatpunct {% 
            \lccode`\~`\.\lowercase{\def~}{\discretionary{\hbox{\char`\.}}{\Wrappedafterbreak}{\hbox{\char`\.}}}% 
            \lccode`\~`\,\lowercase{\def~}{\discretionary{\hbox{\char`\,}}{\Wrappedafterbreak}{\hbox{\char`\,}}}% 
            \lccode`\~`\;\lowercase{\def~}{\discretionary{\hbox{\char`\;}}{\Wrappedafterbreak}{\hbox{\char`\;}}}% 
            \lccode`\~`\:\lowercase{\def~}{\discretionary{\hbox{\char`\:}}{\Wrappedafterbreak}{\hbox{\char`\:}}}% 
            \lccode`\~`\?\lowercase{\def~}{\discretionary{\hbox{\char`\?}}{\Wrappedafterbreak}{\hbox{\char`\?}}}% 
            \lccode`\~`\!\lowercase{\def~}{\discretionary{\hbox{\char`\!}}{\Wrappedafterbreak}{\hbox{\char`\!}}}% 
            \lccode`\~`\/\lowercase{\def~}{\discretionary{\hbox{\char`\/}}{\Wrappedafterbreak}{\hbox{\char`\/}}}% 
            \catcode`\.\active
            \catcode`\,\active 
            \catcode`\;\active
            \catcode`\:\active
            \catcode`\?\active
            \catcode`\!\active
            \catcode`\/\active 
            \lccode`\~`\~ 	
        }
    \makeatother

    \let\OriginalVerbatim=\Verbatim
    \makeatletter
    \renewcommand{\Verbatim}[1][1]{%
        %\parskip\z@skip
        \sbox\Wrappedcontinuationbox {\Wrappedcontinuationsymbol}%
        \sbox\Wrappedvisiblespacebox {\FV@SetupFont\Wrappedvisiblespace}%
        \def\FancyVerbFormatLine ##1{\hsize\linewidth
            \vtop{\raggedright\hyphenpenalty\z@\exhyphenpenalty\z@
                \doublehyphendemerits\z@\finalhyphendemerits\z@
                \strut ##1\strut}%
        }%
        % If the linebreak is at a space, the latter will be displayed as visible
        % space at end of first line, and a continuation symbol starts next line.
        % Stretch/shrink are however usually zero for typewriter font.
        \def\FV@Space {%
            \nobreak\hskip\z@ plus\fontdimen3\font minus\fontdimen4\font
            \discretionary{\copy\Wrappedvisiblespacebox}{\Wrappedafterbreak}
            {\kern\fontdimen2\font}%
        }%

        % Allow breaks at special characters using \PYG... macros.
        \Wrappedbreaksatspecials
        % Breaks at punctuation characters . , ; ? ! and / need catcode=\active 	
        \OriginalVerbatim[#1,codes*=\Wrappedbreaksatpunct]%
    }
    \makeatother

    % Exact colors from NB
    \definecolor{incolor}{HTML}{303F9F}
    \definecolor{outcolor}{HTML}{D84315}
    \definecolor{cellborder}{HTML}{CFCFCF}
    \definecolor{cellbackground}{HTML}{F7F7F7}

    % prompt
    \makeatletter
    \newcommand{\boxspacing}{\kern\kvtcb@left@rule\kern\kvtcb@boxsep}
    \makeatother
    \newcommand{\prompt}[4]{
        \ttfamily\llap{{\color{#2}[#3]:\hspace{3pt}#4}}\vspace{-\baselineskip}
    }

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy 
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=urlcolor,
      linkcolor=linkcolor,
      citecolor=citecolor,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\begin{document}

    \maketitle

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{from} \PY{n+nn}{collections} \PY{k+kn}{import} \PY{n}{defaultdict}
\PY{k+kn}{import} \PY{n+nn}{json}
\PY{k+kn}{import} \PY{n+nn}{pandas} \PY{k}{as} \PY{n+nn}{pd}
\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}

\PY{k}{def} \PY{n+nf}{multidict}\PY{p}{(}\PY{n}{n}\PY{p}{,} \PY{n}{dtype}\PY{o}{=}\PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{:}
    \PY{k}{if} \PY{n}{n} \PY{o}{==} \PY{l+m+mi}{0}\PY{p}{:} \PY{k}{return} \PY{n}{dtype}
    \PY{k}{return} \PY{n}{defaultdict}\PY{p}{(}\PY{k}{lambda}\PY{p}{:}\PY{n}{multidict}\PY{p}{(}\PY{n}{n}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,}\PY{n}{dtype}\PY{p}{)}\PY{p}{)}

\PY{n}{nested\PYZus{}dict} \PY{o}{=} \PY{k}{lambda}\PY{p}{:} \PY{n}{defaultdict}\PY{p}{(}\PY{n}{nested\PYZus{}dict}\PY{p}{)}

\PY{n}{freqs}\PY{o}{=} \PY{p}{[}\PY{l+m+mf}{1.2}\PY{p}{,}\PY{l+m+mf}{2.2}\PY{p}{]}
\PY{n}{cores}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{32}\PY{p}{]}
\PY{n}{inps}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{]}
\PY{n}{regions}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{]}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{measures} \PY{o}{=} \PY{n}{nested\PYZus{}dict}\PY{p}{(}\PY{p}{)}
\PY{c+c1}{\PYZsh{} measures= defaultdict(lambda:defaultdict(lambda:defaultdict(lambda:\PYZob{}\PYZcb{})))}

\PY{c+c1}{\PYZsh{} collect every n milliseconds + event}
\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
    \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
        \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
            \PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:}
                \PY{n}{measures}\PY{p}{[}\PY{n}{f}\PY{p}{]}\PY{p}{[}\PY{n}{c}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{]}\PY{p}{[}\PY{n}{e}\PY{p}{]}\PY{o}{=} \PY{p}{\PYZob{}}
                    \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{timestamp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{0.1}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{200}\PY{p}{\PYZcb{}}\PY{p}{,} \PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{timestamp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{0.2}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{220}\PY{p}{\PYZcb{}}\PY{p}{]}\PY{p}{,}
                    \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{total\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{10}\PY{p}{,}
                    \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{total\PYZus{}energy}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{10}
                    \PY{p}{\PYZcb{}}
                \PY{c+c1}{\PYZsh{} measures[e][f][c][i]= \PYZob{}}
                \PY{c+c1}{\PYZsh{}     \PYZdq{}sensor\PYZdq{}:[\PYZob{}\PYZdq{}timestamp\PYZdq{}:0.1,\PYZdq{}data\PYZdq{}:200\PYZcb{}, \PYZob{}\PYZdq{}timestamp\PYZdq{}:0.2,\PYZdq{}data\PYZdq{}:220\PYZcb{}],}
                \PY{c+c1}{\PYZsh{}     \PYZdq{}total\PYZus{}time\PYZdq{}:10,}
                \PY{c+c1}{\PYZsh{}     \PYZdq{}total\PYZus{}energy\PYZdq{}:10}
                \PY{c+c1}{\PYZsh{}     \PYZcb{}}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
                    \PY{n}{measures}\PY{p}{[}\PY{n}{f}\PY{p}{]}\PY{p}{[}\PY{n}{c}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{]}\PY{p}{[}\PY{n}{e}\PY{p}{]}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{o}{=} \PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{timestamp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{0.32}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{150}\PY{p}{\PYZcb{}}\PY{p}{\PYZcb{}}\PY{p}{,}\PY{p}{]}
                    \PY{c+c1}{\PYZsh{} measures[e][r][f][c][i]= [\PYZob{}\PYZdq{}start\PYZus{}time\PYZdq{}:0,\PYZdq{}stop\PYZus{}time\PYZdq{}:1,\PYZdq{}sensor\PYZdq{}:\PYZob{}\PYZdq{}timestamp\PYZdq{}:0.32,\PYZdq{}data\PYZdq{}:150\PYZcb{}\PYZcb{},]}

\PY{n+nb}{print}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{measures}\PY{p}{,} \PY{n}{indent}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
\{
    "1.2": \{
        "16": \{
            "2": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \}
        \},
        "32": \{
            "2": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \}
        \}
    \},
    "2.2": \{
        "16": \{
            "2": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \}
        \},
        "32": \{
            "2": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "sensor": [
                        \{
                            "timestamp": 0.1,
                            "data": 200
                        \},
                        \{
                            "timestamp": 0.2,
                            "data": 220
                        \}
                    ],
                    "total\_time": 10,
                    "total\_energy": 10,
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": \{
                                "timestamp": 0.32,
                                "data": 150
                            \}
                        \}
                    ]
                \}
            \}
        \}
    \}
\}

    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{measures} \PY{o}{=} \PY{n}{nested\PYZus{}dict}\PY{p}{(}\PY{p}{)}
\PY{c+c1}{\PYZsh{} measures= defaultdict(lambda:defaultdict(lambda:defaultdict(lambda:defaultdict(lambda:\PYZob{}\PYZcb{}))))}
\PY{c+c1}{\PYZsh{} collect every event}
\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
    \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
        \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
            \PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
                    \PY{n}{measures}\PY{p}{[}\PY{n}{f}\PY{p}{]}\PY{p}{[}\PY{n}{c}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{]}\PY{p}{[}\PY{n}{e}\PY{p}{]}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{o}{=} \PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{5}\PY{p}{\PYZcb{}}\PY{p}{,}\PY{p}{]}
                    \PY{c+c1}{\PYZsh{} measures[e][r][f][c][i]= [\PYZob{}\PYZdq{}start\PYZus{}time\PYZdq{}:0,\PYZdq{}stop\PYZus{}time\PYZdq{}:1,\PYZdq{}sensor\PYZdq{}:5\PYZcb{},]}
                    \PY{c+c1}{\PYZsh{} measures[\PYZdq{}\PYZpc{}s\PYZhy{}\PYZpc{}s\PYZhy{}\PYZpc{}s\PYZhy{}\PYZpc{}s\PYZhy{}\PYZpc{}s\PYZdq{}\PYZpc{}(f,c,i,e,r)]=[\PYZob{}\PYZdq{}start\PYZus{}time\PYZdq{}:0,\PYZdq{}stop\PYZus{}time\PYZdq{}:1,\PYZdq{}sensor\PYZdq{}:5\PYZcb{},]}

\PY{n+nb}{print}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{measures}\PY{p}{,}\PY{n}{indent}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
\{
    "1.2": \{
        "16": \{
            "2": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \}
        \},
        "32": \{
            "2": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \}
        \}
    \},
    "2.2": \{
        "16": \{
            "2": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \}
        \},
        "32": \{
            "2": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \},
            "5": \{
                "0": \{
                    "2": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ],
                    "4": [
                        \{
                            "start\_time": 0,
                            "stop\_time": 1,
                            "sensor": 5
                        \}
                    ]
                \}
            \}
        \}
    \}
\}

    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} difficulty to transform this into a xarray, dataframe, numpy}
\PY{c+c1}{\PYZsh{} performance of execution operations on each format}

\PY{n}{measures}\PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
    \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
        \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
            \PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
                    \PY{n}{measures}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{f}\PY{p}{,}\PY{n}{c}\PY{p}{,}\PY{n}{i}\PY{p}{,}\PY{n}{e}\PY{p}{,}\PY{n}{r}\PY{p}{)}\PY{p}{]}\PY{o}{=}\PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{5}\PY{p}{\PYZcb{}}\PY{p}{]}

\PY{n+nb}{print}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{measures}\PY{p}{,}\PY{n}{indent}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{)}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{p}{)}

\PY{k+kn}{import} \PY{n+nn}{pandas} \PY{k}{as} \PY{n+nn}{pd}
\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}
    \PY{n+nb}{map}\PY{p}{(}\PY{k}{lambda} \PY{n}{x}\PY{p}{:} \PY{n}{x}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{+}\PY{p}{[}\PY{n}{measures}\PY{p}{[}\PY{n}{x}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{]}\PY{p}{,}\PY{n}{measures}\PY{p}{)}\PY{p}{,}
    \PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}
\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}

\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{df}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
\{
    "1.2-16-2-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-16-2-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-16-5-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-16-5-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-32-2-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-32-2-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-32-5-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "1.2-32-5-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-16-2-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-16-2-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-16-5-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-16-5-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-32-2-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-32-2-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-32-5-0-2": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ],
    "2.2-32-5-0-4": [
        \{
            "start\_time": 0,
            "stop\_time": 1,
            "sensor": 5
        \}
    ]
\}

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
/home/vitor/anaconda3/lib/python3.7/site-packages/dask/config.py:168:
YAMLLoadWarning: calling yaml.load() without Loader={\ldots} is deprecated, as the
default Loader is unsafe. Please read https://msg.pyyaml.org/load for full
details.
  data = yaml.load(f.read()) or \{\}
/home/vitor/anaconda3/lib/python3.7/site-packages/distributed/config.py:20:
YAMLLoadWarning: calling yaml.load() without Loader={\ldots} is deprecated, as the
default Loader is unsafe. Please read https://msg.pyyaml.org/load for full
details.
  defaults = yaml.load(f)
    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
                             sensor
freq cores inputs execs reg
1.2  16    2      0     2         5
                        4         5
           5      0     2         5
                        4         5
     32    2      0     2         5
                        4         5
           5      0     2         5
                        4         5
2.2  16    2      0     2         5
                        4         5
           5      0     2         5
                        4         5
     32    2      0     2         5
                        4         5
           5      0     2         5
                        4         5
<xarray.Dataset>
Dimensions:  (cores: 2, execs: 1, freq: 2, inputs: 2, reg: 2)
Coordinates:
  * freq     (freq) object '1.2' '2.2'
  * cores    (cores) object '16' '32'
  * inputs   (inputs) object '2' '5'
  * execs    (execs) object '0'
  * reg      (reg) object '2' '4'
Data variables:
    sensor   (freq, cores, inputs, execs, reg) int64 5 5 5 5 5 5 {\ldots} 5 5 5 5 5 5
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} difficulty to transform this into a xarray, dataframe, numpy}
\PY{c+c1}{\PYZsh{} performance of execution operations on each format}

\PY{n}{measures}\PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
    \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
        \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
            \PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
                    \PY{n}{measures}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{f}\PY{p}{,}\PY{n}{c}\PY{p}{,}\PY{n}{i}\PY{p}{,}\PY{n}{e}\PY{p}{,}\PY{n}{r}\PY{p}{)}\PY{p}{]}\PY{o}{=}\PY{p}{[}
                        \PY{p}{[}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{]}\PY{p}{,}
                        \PY{p}{[}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{]}\PY{p}{,}
                        \PY{p}{[}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{]}
                        \PY{p}{]}

\PY{k+kn}{import} \PY{n+nn}{pandas} \PY{k}{as} \PY{n+nn}{pd}
\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}
    \PY{n+nb}{map}\PY{p}{(}\PY{k}{lambda} \PY{n}{x}\PY{p}{:} \PY{n}{x}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{+}\PY{n}{measures}\PY{p}{[}\PY{n}{x}\PY{p}{]}\PY{p}{,}\PY{n}{measures}\PY{p}{)}\PY{p}{,}
    \PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}
\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}

\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{df}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
                                 start       stop     sensor
freq cores inputs execs reg
1.2  16    2      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
           5      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
     32    2      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
           5      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
2.2  16    2      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
           5      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
     32    2      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
           5      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
<xarray.Dataset>
Dimensions:  (cores: 2, execs: 1, freq: 2, inputs: 2, reg: 2)
Coordinates:
  * freq     (freq) object '1.2' '2.2'
  * cores    (cores) object '16' '32'
  * inputs   (inputs) object '2' '5'
  * execs    (execs) object '0'
  * reg      (reg) object '2' '4'
Data variables:
    start    (freq, cores, inputs, execs, reg) object [0, 1, 3] {\ldots} [0, 1, 3]
    stop     (freq, cores, inputs, execs, reg) object [2, 3, 4] {\ldots} [2, 3, 4]
    sensor   (freq, cores, inputs, execs, reg) object [5, 4, 5] {\ldots} [5, 4, 5]
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{execs}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}
\PY{n}{measures}\PY{o}{=} \PY{p}{[}\PY{p}{]}
\PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n}{execs}\PY{p}{:}
    \PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
        \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
            \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}

                \PY{n}{aux}\PY{o}{=} \PY{p}{[}\PY{p}{]}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
                    \PY{c+c1}{\PYZsh{} start\PYZus{}time stop\PYZus{}time sensor}
                    \PY{n}{aux}\PY{o}{.}\PY{n}{append}\PY{p}{(}
                        \PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{region}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{n}{r}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}\PY{p}{\PYZcb{}}\PY{p}{)}
                    \PY{n}{aux}\PY{o}{.}\PY{n}{append}\PY{p}{(}
                        \PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{region}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{n}{r}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{4}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}\PY{p}{\PYZcb{}}\PY{p}{)}

                \PY{n}{measures}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{aux}\PY{p}{)}

\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}
\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{measures}\PY{p}{)}
\PY{n}{idxs}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{MultiIndex}\PY{o}{.}\PY{n}{from\PYZus{}product}\PY{p}{(}\PY{p}{[}\PY{n}{execs}\PY{p}{,}\PY{n}{freqs}\PY{p}{,}\PY{n}{cores}\PY{p}{,}\PY{n}{inps}\PY{p}{]}\PY{p}{,}\PY{n}{names}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
\PY{n}{df}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{n}{idxs}\PY{p}{)}

\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{df}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
                                                                         0  \textbackslash{}
freq cores inputs execs
1    1.2   16     2      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
           32     2      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
     2.2   16     2      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
           32     2      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}

                                                                         1  \textbackslash{}
freq cores inputs execs
1    1.2   16     2      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
           32     2      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
     2.2   16     2      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
           32     2      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}

                                                                         2  \textbackslash{}
freq cores inputs execs
1    1.2   16     2      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
           32     2      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
     2.2   16     2      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
           32     2      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
                  5      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}

                                                                         3
freq cores inputs execs
1    1.2   16     2      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
           32     2      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
     2.2   16     2      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
           32     2      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
                  5      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
<xarray.Dataset>
Dimensions:  (cores: 2, execs: 2, freq: 1, inputs: 2)
Coordinates:
  * freq     (freq) int64 1
  * cores    (cores) float64 1.2 2.2
  * inputs   (inputs) int64 16 32
  * execs    (execs) int64 2 5
Data variables:
    0        (freq, cores, inputs, execs) object \{'region': 2, 'start\_time': 0,
'stop\_time': 2, 'sensor': \{\}\} {\ldots} \{'region': 2, 'start\_time': 0, 'stop\_time': 2,
'sensor': \{\}\}
    1        (freq, cores, inputs, execs) object \{'region': 2, 'start\_time': 3,
'stop\_time': 4, 'sensor': \{\}\} {\ldots} \{'region': 2, 'start\_time': 3, 'stop\_time': 4,
'sensor': \{\}\}
    2        (freq, cores, inputs, execs) object \{'region': 4, 'start\_time': 0,
'stop\_time': 2, 'sensor': \{\}\} {\ldots} \{'region': 4, 'start\_time': 0, 'stop\_time': 2,
'sensor': \{\}\}
    3        (freq, cores, inputs, execs) object \{'region': 4, 'start\_time': 3,
'stop\_time': 4, 'sensor': \{\}\} {\ldots} \{'region': 4, 'start\_time': 3, 'stop\_time': 4,
'sensor': \{\}\}
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{execs}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}
\PY{n}{measures}\PY{o}{=} \PY{p}{[}\PY{p}{]}
\PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n}{execs}\PY{p}{:}
    \PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
        \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
            \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
                    \PY{n}{measures}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{[}\PY{n}{e}\PY{p}{,}\PY{n}{f}\PY{p}{,}\PY{n}{c}\PY{p}{,}\PY{n}{i}\PY{p}{,}\PY{n}{r}\PY{p}{,}\PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dados1}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{\PYZcb{}}\PY{p}{,}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dados2}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{\PYZcb{}}\PY{p}{]}\PY{p}{]}\PY{p}{)}

\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}
\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{measures}\PY{p}{,}\PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{exec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dados}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{exec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}

\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{df}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
                                                     dados
exec freq cores inputs reg
1    1.2  16    2      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
                5      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
          32    2      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
                5      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
     2.2  16    2      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
                5      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
          32    2      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
                5      2    [\{'dados1': 0\}, \{'dados2': 0\}]
                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
<xarray.Dataset>
Dimensions:  (cores: 2, exec: 1, freq: 2, inputs: 2, reg: 2)
Coordinates:
  * exec     (exec) int64 1
  * freq     (freq) float64 1.2 2.2
  * cores    (cores) int64 16 32
  * inputs   (inputs) int64 2 5
  * reg      (reg) int64 2 4
Data variables:
    dados    (exec, freq, cores, inputs, reg) object [\{'dados1': 0\}, \{'dados2':
0\}] {\ldots} [\{'dados1': 0\}, \{'dados2': 0\}]
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{execs}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{2}\PY{p}{]}
\PY{n}{measures}\PY{o}{=} \PY{p}{[}\PY{p}{]}
\PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n}{execs}\PY{p}{:}
    \PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
        \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
            \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}

                \PY{n}{aux}\PY{o}{=} \PY{n}{defaultdict}\PY{p}{(}\PY{k}{lambda}\PY{p}{:}\PY{p}{[}\PY{p}{[}\PY{p}{]}\PY{p}{,}\PY{p}{[}\PY{p}{]}\PY{p}{,}\PY{p}{[}\PY{p}{]}\PY{p}{]}\PY{p}{)}
                \PY{k}{for} \PY{n}{rx} \PY{o+ow}{in} \PY{p}{[}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{,}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{]}\PY{p}{:}
                    \PY{c+c1}{\PYZsh{} start\PYZus{}time stop\PYZus{}time sensor}
                    \PY{c+c1}{\PYZsh{} aux[rx].append(}
                    \PY{c+c1}{\PYZsh{}     \PYZob{}\PYZdq{}start\PYZus{}time\PYZdq{}:0,\PYZdq{}stop\PYZus{}time\PYZdq{}:2,\PYZdq{}sensor\PYZdq{}:\PYZob{}\PYZcb{}\PYZcb{})}
                    \PY{n}{aux}\PY{p}{[}\PY{n}{rx}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}
                    \PY{n}{aux}\PY{p}{[}\PY{n}{rx}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}
                    \PY{n}{aux}\PY{p}{[}\PY{n}{rx}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}\PY{p}{)}

                \PY{n}{reg\PYZus{}k}\PY{o}{=} \PY{n}{aux}\PY{o}{.}\PY{n}{keys}\PY{p}{(}\PY{p}{)}
                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{aux}\PY{p}{:}
                    \PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{)}
                    \PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
                    \PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{p}{)}
                    \PY{n}{measures}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{)}

\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}
\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{measures}\PY{p}{,}\PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
\PY{c+c1}{\PYZsh{} df[\PYZdq{}start\PYZdq{}]= df[\PYZdq{}start\PYZdq{}].apply(lambda x: np.array(x))}
\PY{c+c1}{\PYZsh{} df[\PYZdq{}stop\PYZdq{}]= df[\PYZdq{}stop\PYZdq{}].apply(lambda x: np.array(x))}
\PY{n}{idxs}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{MultiIndex}\PY{o}{.}\PY{n}{from\PYZus{}product}\PY{p}{(}\PY{p}{[}\PY{n}{execs}\PY{p}{,}\PY{n}{freqs}\PY{p}{,}\PY{n}{cores}\PY{p}{,}\PY{n}{inps}\PY{p}{,}\PY{n}{reg\PYZus{}k}\PY{p}{]}\PY{p}{,}\PY{n}{names}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{regs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
\PY{n}{df}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{n}{idxs}\PY{p}{)}

\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
<xarray.Dataset>
Dimensions:  (cores: 2, execs: 2, freq: 2, inputs: 2, regs: 2)
Coordinates:
  * execs    (execs) int64 1 2
  * freq     (freq) float64 1.2 2.2
  * cores    (cores) int64 16 32
  * inputs   (inputs) int64 2 5
  * regs     (regs) int64 2 4
Data variables:
    start    (execs, freq, cores, inputs, regs) object [0 0] [0 0] {\ldots} [0 0]
    stop     (execs, freq, cores, inputs, regs) object [1 1] [1 1] {\ldots} [1 1]
    sensor   (execs, freq, cores, inputs, regs) object [\{\} \{\}] {\ldots} [\{\} \{\}]
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{data}\PY{o}{=} \PY{p}{[}\PY{p}{]}
\PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n}{execs}\PY{p}{:}
    \PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
        \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
            \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
                \PY{k}{for} \PY{n}{x} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{random}\PY{o}{.}\PY{n}{randint}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{)}\PY{p}{)}\PY{p}{:}
                    \PY{n}{data}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{[}\PY{n}{e}\PY{p}{,}\PY{n}{f}\PY{p}{,}\PY{n}{c}\PY{p}{,}\PY{n}{i}\PY{p}{,}\PY{n}{x}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
\PY{n}{dfx}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{data}\PY{p}{,}\PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{exec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{core}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ent}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
\PY{n}{dfx}\PY{o}{=} \PY{n}{dfx}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{exec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{core}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ent}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}
\PY{n}{dfx}
\PY{c+c1}{\PYZsh{} xrx= dfx.to\PYZus{}xarray()}
\PY{c+c1}{\PYZsh{} xrx}
\end{Verbatim}
\end{tcolorbox}

            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
\prompt{Out}{outcolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
                        data
exec freq core inp ent
1    1.2  16   2   0       1
                   1       1
                   2       1
                   3       1
               5   0       1
                   1       1
                   2       1
          32   2   0       1
               5   0       1
     2.2  16   2   0       1
                   1       1
                   2       1
               5   0       1
                   1       1
          32   2   0       1
               5   0       1
                   1       1
2    1.2  16   2   0       1
                   1       1
               5   0       1
                   1       1
          32   2   0       1
                   1       1
                   2       1
               5   0       1
                   1       1
                   2       1
     2.2  16   2   0       1
                   1       1
                   2       1
                   3       1
               5   0       1
          32   2   0       1
               5   0       1
\end{Verbatim}
\end{tcolorbox}

    \hypertarget{benchmarks}{%
\section{Benchmarks}\label{benchmarks}}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{from} \PY{n+nn}{itertools} \PY{k+kn}{import} \PY{n}{product}
\PY{k+kn}{import} \PY{n+nn}{random}
\PY{k+kn}{import} \PY{n+nn}{string}
\PY{k+kn}{import} \PY{n+nn}{sys}
\PY{k+kn}{from} \PY{n+nn}{numpy} \PY{k+kn}{import} \PY{n}{prod}
\PY{k+kn}{import} \PY{n+nn}{pickle}
\PY{k+kn}{import} \PY{n+nn}{gzip}
\PY{k+kn}{import} \PY{n+nn}{os}
\PY{k+kn}{import} \PY{n+nn}{gc}

\PY{k}{def} \PY{n+nf}{si\PYZus{}format}\PY{p}{(}\PY{n}{n}\PY{p}{,} \PY{n}{precision}\PY{o}{=} \PY{l+m+mi}{2}\PY{p}{)}\PY{p}{:}
    \PY{n}{sufx}\PY{o}{=} \PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{M}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{1e6}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{K}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{1e3}\PY{p}{\PYZcb{}}
    \PY{n}{sufix}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}
    \PY{k}{for} \PY{n}{k}\PY{p}{,}\PY{n}{v} \PY{o+ow}{in} \PY{n}{sufx}\PY{o}{.}\PY{n}{items}\PY{p}{(}\PY{p}{)}\PY{p}{:}
        \PY{k}{if} \PY{n}{n} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{n}{v}\PY{p}{:}
            \PY{n}{n}\PY{o}{/}\PY{o}{=}\PY{n}{v}
            \PY{n}{sufix}\PY{o}{=} \PY{n}{k}
            \PY{k}{break}
    \PY{k}{return} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{:.}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{f\PYZcb{} }\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{b}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{n}\PY{p}{,}\PY{n}{precision}\PY{p}{,}\PY{n}{sufix}\PY{p}{)}

\PY{k}{def} \PY{n+nf}{get\PYZus{}size}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{n}{seen}\PY{o}{=}\PY{k+kc}{None}\PY{p}{)}\PY{p}{:}
    \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Recursively finds size of objects\PYZdq{}\PYZdq{}\PYZdq{}}
    \PY{n}{size} \PY{o}{=} \PY{n}{sys}\PY{o}{.}\PY{n}{getsizeof}\PY{p}{(}\PY{n}{obj}\PY{p}{)}
    \PY{k}{if} \PY{n}{seen} \PY{o+ow}{is} \PY{k+kc}{None}\PY{p}{:}
        \PY{n}{seen} \PY{o}{=} \PY{n+nb}{set}\PY{p}{(}\PY{p}{)}
    \PY{n}{obj\PYZus{}id} \PY{o}{=} \PY{n+nb}{id}\PY{p}{(}\PY{n}{obj}\PY{p}{)}
    \PY{k}{if} \PY{n}{obj\PYZus{}id} \PY{o+ow}{in} \PY{n}{seen}\PY{p}{:}
        \PY{k}{return} \PY{l+m+mi}{0}
    \PY{c+c1}{\PYZsh{} Important mark as seen *before* entering recursion to gracefully handle}
    \PY{c+c1}{\PYZsh{} self\PYZhy{}referential objects}
    \PY{n}{seen}\PY{o}{.}\PY{n}{add}\PY{p}{(}\PY{n}{obj\PYZus{}id}\PY{p}{)}
    \PY{k}{if} \PY{n+nb}{isinstance}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{n+nb}{dict}\PY{p}{)}\PY{p}{:}
        \PY{n}{size} \PY{o}{+}\PY{o}{=} \PY{n+nb}{sum}\PY{p}{(}\PY{p}{[}\PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{v}\PY{p}{,} \PY{n}{seen}\PY{p}{)} \PY{k}{for} \PY{n}{v} \PY{o+ow}{in} \PY{n}{obj}\PY{o}{.}\PY{n}{values}\PY{p}{(}\PY{p}{)}\PY{p}{]}\PY{p}{)}
        \PY{n}{size} \PY{o}{+}\PY{o}{=} \PY{n+nb}{sum}\PY{p}{(}\PY{p}{[}\PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{k}\PY{p}{,} \PY{n}{seen}\PY{p}{)} \PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{obj}\PY{o}{.}\PY{n}{keys}\PY{p}{(}\PY{p}{)}\PY{p}{]}\PY{p}{)}
    \PY{k}{elif} \PY{n+nb}{hasattr}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZus{}\PYZus{}dict\PYZus{}\PYZus{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}\PY{p}{:}
        \PY{n}{size} \PY{o}{+}\PY{o}{=} \PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{obj}\PY{o}{.}\PY{n+nv+vm}{\PYZus{}\PYZus{}dict\PYZus{}\PYZus{}}\PY{p}{,} \PY{n}{seen}\PY{p}{)}
    \PY{k}{elif} \PY{n+nb}{hasattr}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZus{}\PYZus{}iter\PYZus{}\PYZus{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)} \PY{o+ow}{and} \PY{o+ow}{not} \PY{n+nb}{isinstance}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{p}{(}\PY{n+nb}{str}\PY{p}{,} \PY{n+nb}{bytes}\PY{p}{,} \PY{n+nb}{bytearray}\PY{p}{)}\PY{p}{)}\PY{p}{:}
        \PY{n}{size} \PY{o}{+}\PY{o}{=} \PY{n+nb}{sum}\PY{p}{(}\PY{p}{[}\PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{seen}\PY{p}{)} \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{obj}\PY{p}{]}\PY{p}{)}
    \PY{k}{return} \PY{n}{size}

\PY{c+c1}{\PYZsh{} cores freq input exec reg=}
\PY{k}{def} \PY{n+nf}{bench}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{,} \PY{n}{nbytes}\PY{p}{)}\PY{p}{:}
    \PY{n}{vx}\PY{o}{=} \PY{p}{[}\PY{p}{]}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{Keys:}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{prod}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{)}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ Bytes:}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

\PY{c+c1}{\PYZsh{}     vx= [[ \PYZdq{}var\PYZob{}\PYZcb{}\PYZus{}\PYZob{}\PYZcb{}\PYZdq{}.format(i, j) for j in range(vs) ] for i, vs in enumerate(v\PYZus{}sizes)]}
    \PY{k}{for} \PY{n}{i}\PY{p}{,} \PY{n}{vs} \PY{o+ow}{in} \PY{n+nb}{enumerate}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{)}\PY{p}{:}
        \PY{n}{vx}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{[} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{var}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{\PYZus{}}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)} \PY{k}{for} \PY{n}{j} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{vs}\PY{p}{)} \PY{p}{]}\PY{p}{)}

    \PY{n}{multi\PYZus{}dict} \PY{o}{=} \PY{n}{nested\PYZus{}dict}\PY{p}{(}\PY{p}{)}
    \PY{n}{normal\PYZus{}dict} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{product}\PY{p}{(}\PY{o}{*}\PY{n}{vx}\PY{p}{)}\PY{p}{:}
        \PY{n}{data}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{random}\PY{o}{.}\PY{n}{choices}\PY{p}{(}\PY{n}{string}\PY{o}{.}\PY{n}{ascii\PYZus{}letters}\PY{p}{,}\PY{n}{k}\PY{o}{=}\PY{n+nb}{int}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{)}\PY{p}{)}

        \PY{n}{aux}\PY{o}{=} \PY{n}{multi\PYZus{}dict}
        \PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{idx}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{:}
            \PY{n}{aux} \PY{o}{=} \PY{n}{aux}\PY{p}{[}\PY{n}{k}\PY{p}{]}
        \PY{n}{aux}\PY{p}{[}\PY{n}{idx}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{]}\PY{o}{=} \PY{n}{data}

        \PY{n}{normal\PYZus{}dict}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{idx}\PY{p}{)}\PY{p}{]}\PY{o}{=} \PY{n}{data}

    \PY{n}{s1}\PY{o}{=} \PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{multi\PYZus{}dict}\PY{p}{)}
    \PY{n}{s2}\PY{o}{=} \PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{normal\PYZus{}dict}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{RAM       }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s1}\PY{p}{)}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s2}\PY{p}{)}\PY{p}{,} \PY{n}{s1}\PY{o}{/}\PY{n}{s2} \PY{p}{)}

    \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{w+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
        \PY{n}{tf}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{multi\PYZus{}dict}\PY{p}{)}\PY{p}{)}

    \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{w+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
        \PY{n}{tf}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{normal\PYZus{}dict}\PY{p}{)}\PY{p}{)}

    \PY{n}{s1}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{n}{s2}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{DISK json }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s1}\PY{p}{)}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s2}\PY{p}{)}\PY{p}{,} \PY{n}{s1}\PY{o}{/}\PY{n}{s2} \PY{p}{)}

    \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.pkl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{wb+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
        \PY{n}{pickle}\PY{o}{.}\PY{n}{dump}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{loads}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{multi\PYZus{}dict}\PY{p}{)}\PY{p}{)}\PY{p}{,} \PY{n}{tf}\PY{p}{,} \PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}

    \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.pkl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{wb+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
        \PY{n}{pickle}\PY{o}{.}\PY{n}{dump}\PY{p}{(}\PY{n}{normal\PYZus{}dict}\PY{p}{,} \PY{n}{tf}\PY{p}{,} \PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}

    \PY{n}{s1}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.pkl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{n}{s2}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.pkl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{DISK pkl  }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s1}\PY{p}{)}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s2}\PY{p}{)}\PY{p}{,} \PY{n}{s1}\PY{o}{/}\PY{n}{s2} \PY{p}{)}

    \PY{k}{with} \PY{n}{gzip}\PY{o}{.}\PY{n}{GzipFile}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.gz}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{wb}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
        \PY{n}{tf}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{multi\PYZus{}dict}\PY{p}{)}\PY{o}{.}\PY{n}{encode}\PY{p}{(}\PY{p}{)}\PY{p}{)}

    \PY{k}{with} \PY{n}{gzip}\PY{o}{.}\PY{n}{GzipFile}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.gz}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{wb}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
        \PY{n}{tf}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{normal\PYZus{}dict}\PY{p}{)}\PY{o}{.}\PY{n}{encode}\PY{p}{(}\PY{p}{)}\PY{p}{)}

    \PY{n}{s1}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.gz}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{n}{s2}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.gz}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{DISK gz  }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s1}\PY{p}{)}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s2}\PY{p}{)}\PY{p}{,} \PY{n}{s1}\PY{o}{/}\PY{n}{s2}\PY{p}{)}

    \PY{n}{gc}\PY{o}{.}\PY{n}{collect}\PY{p}{(}\PY{p}{)}

\PY{c+c1}{\PYZsh{}bench([2,3,2],10)}
\PY{c+c1}{\PYZsh{} bench([32,10,10,10,10],10)}
\PY{c+c1}{\PYZsh{} bench([32,10,10,10,10],1e3)}
\PY{c+c1}{\PYZsh{} bench([32,10,10,10,10],100e3)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e1}\PY{p}{)}
\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e2}\PY{p}{)}
\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e3}\PY{p}{)}
\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e4}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]

Keys: 24.00 Kb  Bytes: 10.00 b

RAM        2.57 Mb 4.73 Mb 0.542656059372869
DISK json  616.52 Kb 1.26 Mb 0.49046778042959427
DISK pkl   380.99 Kb 1.21 Mb 0.31507013577280013
DISK gz   221.87 Kb 270.49 Kb 0.8202447410255462

Keys: 24.00 Kb  Bytes: 100.00 b

RAM        4.73 Mb 6.89 Mb 0.6860778510924217
DISK json  2.78 Mb 3.42 Mb 0.8125601404741001
DISK pkl   2.54 Mb 3.37 Mb 0.7541981461470956
DISK gz   1.82 Mb 1.87 Mb 0.9686695423189734

Keys: 24.00 Kb  Bytes: 1.00 Kb

RAM        26.33 Mb 28.49 Mb 0.9240995518926407
DISK json  24.38 Mb 25.02 Mb 0.974398129272095
DISK pkl   24.22 Mb 25.04 Mb 0.9669301631449064
DISK gz   17.58 Mb 17.64 Mb 0.9962784816626885

Keys: 24.00 Kb  Bytes: 10.00 Kb

RAM        242.33 Mb 244.49 Mb 0.9911560500830847
DISK json  240.38 Mb 241.02 Mb 0.9973425857927034
DISK pkl   240.24 Mb 241.07 Mb 0.996564850797469
DISK gz   175.11 Mb 175.07 Mb 1.0001946831608381
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{32}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]

Keys: 320.00 Kb  Bytes: 1.00 b

RAM        13.38 Mb 37.27 Mb 0.3588984160511397
DISK json  5.23 Mb 13.98 Mb 0.3738659513590844
DISK pkl   1.49 Mb 12.72 Mb 0.1175192426066018
DISK gz   537.40 Kb 1.24 Mb 0.4322204295799909
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{import} \PY{n+nn}{time}

\PY{k}{def} \PY{n+nf}{bench2}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{,} \PY{n}{nbytes}\PY{p}{)}\PY{p}{:}
    \PY{n}{vx}\PY{o}{=} \PY{p}{[}\PY{p}{]}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{Keys:}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{prod}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{)}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ Bytes:}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
    \PY{k}{for} \PY{n}{i}\PY{p}{,} \PY{n}{vs} \PY{o+ow}{in} \PY{n+nb}{enumerate}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{)}\PY{p}{:}
        \PY{n}{vx}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{[} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{var}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{\PYZus{}}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)} \PY{k}{for} \PY{n}{j} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{vs}\PY{p}{)} \PY{p}{]}\PY{p}{)}

    \PY{n}{multi\PYZus{}dict} \PY{o}{=} \PY{n}{nested\PYZus{}dict}\PY{p}{(}\PY{p}{)}
    \PY{n}{normal\PYZus{}dict} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
    \PY{n}{all\PYZus{}comb}\PY{o}{=} \PY{n}{product}\PY{p}{(}\PY{o}{*}\PY{n}{vx}\PY{p}{)}
    \PY{n}{ti}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{all\PYZus{}comb}\PY{p}{:}
        \PY{n}{data}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{random}\PY{o}{.}\PY{n}{choices}\PY{p}{(}\PY{n}{string}\PY{o}{.}\PY{n}{ascii\PYZus{}letters}\PY{p}{,}\PY{n}{k}\PY{o}{=}\PY{n+nb}{int}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{)}\PY{p}{)}

        \PY{n}{aux}\PY{o}{=} \PY{n}{multi\PYZus{}dict}
        \PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{idx}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{:}
            \PY{n}{aux} \PY{o}{=} \PY{n}{aux}\PY{p}{[}\PY{n}{k}\PY{p}{]}
        \PY{n}{aux}\PY{p}{[}\PY{n}{idx}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{]}\PY{o}{=} \PY{n}{data}
    \PY{n}{tf}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{n}{s1}\PY{o}{=} \PY{n}{tf}\PY{o}{\PYZhy{}}\PY{n}{ti}

    \PY{n}{ti}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{all\PYZus{}comb}\PY{p}{:}
        \PY{n}{data}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{random}\PY{o}{.}\PY{n}{choices}\PY{p}{(}\PY{n}{string}\PY{o}{.}\PY{n}{ascii\PYZus{}letters}\PY{p}{,}\PY{n}{k}\PY{o}{=}\PY{n+nb}{int}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{)}\PY{p}{)}
        \PY{n}{normal\PYZus{}dict}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{idx}\PY{p}{)}\PY{p}{]}\PY{o}{=} \PY{n}{data}
    \PY{n}{tf}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{n}{s2}\PY{o}{=} \PY{n}{tf}\PY{o}{\PYZhy{}}\PY{n}{ti}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Creation time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{s1}\PY{p}{,} \PY{n}{s2}\PY{p}{)}

    \PY{c+c1}{\PYZsh{}multi\PYZus{}dict= json.loads(json.dumps(multi\PYZus{}dict))}
    \PY{n}{ti}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{all\PYZus{}comb}\PY{p}{:}
        \PY{n}{aux}\PY{o}{=} \PY{n}{multi\PYZus{}dict}
        \PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{idx}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{:}
            \PY{n}{aux} \PY{o}{=} \PY{n}{aux}\PY{p}{[}\PY{n}{k}\PY{p}{]}
        \PY{n}{z}\PY{o}{=} \PY{n}{aux}\PY{p}{[}\PY{n}{idx}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{]}
    \PY{n}{tf}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{n}{s1}\PY{o}{=} \PY{n}{tf}\PY{o}{\PYZhy{}}\PY{n}{ti}

    \PY{n}{ti}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{all\PYZus{}comb}\PY{p}{:}
        \PY{n}{z}\PY{o}{=} \PY{n}{normal\PYZus{}dict}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{idx}\PY{p}{)}\PY{p}{]}
    \PY{n}{tf}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
    \PY{n}{s2}\PY{o}{=} \PY{n}{tf}\PY{o}{\PYZhy{}}\PY{n}{ti}
    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Acess time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{s1}\PY{p}{,} \PY{n}{s2}\PY{p}{)}

    \PY{n}{gc}\PY{o}{.}\PY{n}{collect}\PY{p}{(}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e1}\PY{p}{)}
\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e2}\PY{p}{)}
\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e3}\PY{p}{)}
\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e4}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]

Keys: 24.00 Kb  Bytes: 10.00 b

Creation time 0.12965726852416992 2.384185791015625e-07
Acess time 4.76837158203125e-07 2.384185791015625e-07

Keys: 24.00 Kb  Bytes: 100.00 b

Creation time 0.6197633743286133 4.76837158203125e-07
Acess time 2.384185791015625e-07 4.76837158203125e-07

Keys: 24.00 Kb  Bytes: 1.00 Kb

Creation time 5.764535903930664 2.384185791015625e-07
Acess time 7.152557373046875e-07 2.384185791015625e-07

Keys: 24.00 Kb  Bytes: 10.00 Kb

Creation time 56.628950119018555 7.152557373046875e-07
Acess time 4.76837158203125e-07 0.0
    \end{Verbatim}

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{0}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{*}\PY{l+m+mi}{1000}\PY{p}{,}\PY{l+m+mf}{1e4}\PY{p}{)}
\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{*}\PY{l+m+mi}{1000}\PY{p}{,}\PY{l+m+mf}{1e4}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]

Keys: 1.00 b  Bytes: 10.00 Kb

RAM        314.94 Kb 19.23 Kb 16.38003848754356
DISK json  23.89 Kb 18.90 Kb 1.2643276710588982
DISK pkl   23.91 Kb 18.92 Kb 1.263903573694227
DISK gz   10.34 Kb 10.14 Kb 1.020122312093115

Keys: 1.00 b  Bytes: 10.00 Kb

Creation time 0.002998828887939453 4.76837158203125e-07
Acess time 7.152557373046875e-07 2.384185791015625e-07
    \end{Verbatim}

    % Add a bibliography block to the postdoc

\end{document}

### 2021-03-23 thesis/pascalanalyzer/arango.tex
add tex files

--- 2021-03-23 thesis/pascalanalyzer/DataStructure2.tex
@@
     % Document title
-    \title{DataStructure2}
+    \title{arango}
 
@@
     \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{k+kn}{from} \PY{n+nn}{collections} \PY{k+kn}{import} \PY{n}{defaultdict}
-\PY{k+kn}{import} \PY{n+nn}{json}
+\prompt{In}{incolor}{138}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{k+kn}{from} \PY{n+nn}{arango} \PY{k+kn}{import} \PY{n}{ArangoClient}
+\PY{k+kn}{import} \PY{n+nn}{os}\PY{o}{,} \PY{n+nn}{json}
+\PY{c+c1}{\PYZsh{} Initialize the ArangoDB client.}
+\PY{n}{client} \PY{o}{=} \PY{n}{ArangoClient}\PY{p}{(}\PY{n}{hosts}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{http://172.17.0.2:8529}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
+\PY{n}{sys\PYZus{}db} \PY{o}{=} \PY{n}{client}\PY{o}{.}\PY{n}{db}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}system}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{username}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{root}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{password}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{oN1WiCr2sm4m1H4g}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
+\PY{k}{if} \PY{o+ow}{not} \PY{n}{sys\PYZus{}db}\PY{o}{.}\PY{n}{has\PYZus{}database}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{energy}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
+    \PY{n}{sys\PYZus{}db}\PY{o}{.}\PY{n}{create\PYZus{}database}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{energy}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
+\PY{n}{energy\PYZus{}db}\PY{o}{=} \PY{n}{client}\PY{o}{.}\PY{n}{db}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{energy}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{username}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{root}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{password}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{oN1WiCr2sm4m1H4g}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
+\PY{k}{if} \PY{o+ow}{not} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{has\PYZus{}collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{run}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
+    \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{create\PYZus{}collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{run}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
+\PY{n}{run\PYZus{}col}\PY{o}{=} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{run}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
+\PY{k}{if} \PY{o+ow}{not} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{has\PYZus{}collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
+    \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{create\PYZus{}collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
+\PY{n}{model\PYZus{}col}\PY{o}{=} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
+\PY{k}{if} \PY{o+ow}{not} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{has\PYZus{}graph}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
+    \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{create\PYZus{}graph}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
+\PY{n}{md\PYZus{}graph}\PY{o}{=} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{graph}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
+\PY{k}{if} \PY{o+ow}{not} \PY{n}{md\PYZus{}graph}\PY{o}{.}\PY{n}{has\PYZus{}edge\PYZus{}definition}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
+    \PY{n}{md\PYZus{}graph}\PY{o}{.}\PY{n}{create\PYZus{}edge\PYZus{}definition}\PY{p}{(}
+        \PY{n}{edge\PYZus{}collection}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}
+        \PY{n}{from\PYZus{}vertex\PYZus{}collections}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{,}
+        \PY{n}{to\PYZus{}vertex\PYZus{}collections}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{run}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}
+    \PY{p}{)}
+\PY{n}{md\PYZus{}}\PY{o}{=} \PY{n}{md\PYZus{}graph}\PY{o}{.}\PY{n}{edge\PYZus{}collection}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{model\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
+\end{Verbatim}
+\end{tcolorbox}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{137}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{n}{sys\PYZus{}db}\PY{o}{.}\PY{n}{delete\PYZus{}database}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{energy}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
+\end{Verbatim}
+\end{tcolorbox}
+
+            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
+\prompt{Out}{outcolor}{137}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+True
+\end{Verbatim}
+\end{tcolorbox}
+
+    \hypertarget{data}{%
+\section{Data}\label{data}}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{139}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{n}{dpath}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/home/vitor/Documents/energy\PYZus{}scripts\PYZus{}superpc/results/}\PY{l+s+s2}{\PYZdq{}}
+\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{os}\PY{o}{.}\PY{n}{listdir}\PY{p}{(}\PY{n}{dpath}\PY{o}{+}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ondemand}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
+    \PY{n}{fname}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{dpath}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ondemand}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{n}{f}\PY{p}{)}
+    \PY{n}{data}\PY{o}{=} \PY{n}{json}\PY{o}{.}\PY{n}{load}\PY{p}{(}\PY{n+nb}{open}\PY{p}{(}\PY{n}{fname}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{r+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
+    \PY{n}{run\PYZus{}col}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{n}{data}\PY{p}{)}
+\end{Verbatim}
+\end{tcolorbox}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{140}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{n}{dpath}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/home/vitor/Documents/energy\PYZus{}scripts\PYZus{}superpc/results/}\PY{l+s+s2}{\PYZdq{}}
+\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{os}\PY{o}{.}\PY{n}{listdir}\PY{p}{(}\PY{n}{dpath}\PY{o}{+}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{performance}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
+    \PY{n}{fname}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{dpath}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{performance}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{n}{f}\PY{p}{)}
+    \PY{n}{data}\PY{o}{=} \PY{n}{json}\PY{o}{.}\PY{n}{load}\PY{p}{(}\PY{n+nb}{open}\PY{p}{(}\PY{n}{fname}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{r+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
+    \PY{n}{run\PYZus{}col}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{n}{data}\PY{p}{)}
+\end{Verbatim}
+\end{tcolorbox}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{141}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{n}{dpath}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/home/vitor/Documents/energy\PYZus{}data\PYZus{}measuments/parsec/ondemand}\PY{l+s+s2}{\PYZdq{}}
+\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{os}\PY{o}{.}\PY{n}{listdir}\PY{p}{(}\PY{n}{dpath}\PY{p}{)}\PY{p}{:}
+    \PY{k}{if} \PY{o+ow}{not} \PY{n}{f}\PY{o}{.}\PY{n}{endswith}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:} \PY{k}{continue}
+    \PY{n}{fname}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{dpath}\PY{p}{,}\PY{n}{f}\PY{p}{)}
+    \PY{n}{data}\PY{o}{=} \PY{n}{json}\PY{o}{.}\PY{n}{load}\PY{p}{(}\PY{n+nb}{open}\PY{p}{(}\PY{n}{fname}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{r+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
+    \PY{n}{run\PYZus{}col}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{n}{data}\PY{p}{)}
+\end{Verbatim}
+\end{tcolorbox}
+
+    \hypertarget{models}{%
+\section{Models}\label{models}}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{133}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{n}{dpath}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/home/vitor/Documents/energy\PYZus{}scripts\PYZus{}superpc/models/equation}\PY{l+s+s2}{\PYZdq{}}
+\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{os}\PY{o}{.}\PY{n}{listdir}\PY{p}{(}\PY{n}{dpath}\PY{p}{)}\PY{p}{:}
+    \PY{k}{if} \PY{o+ow}{not} \PY{n}{f}\PY{o}{.}\PY{n}{endswith}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:} \PY{k}{continue}
+    \PY{n}{fname}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{dpath}\PY{p}{,}\PY{n}{f}\PY{p}{)}
+    \PY{n}{data}\PY{o}{=} \PY{n}{json}\PY{o}{.}\PY{n}{load}\PY{p}{(}\PY{n+nb}{open}\PY{p}{(}\PY{n}{fname}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{r+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
+    \PY{n}{aux}\PY{o}{=} \PY{n}{model\PYZus{}col}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{n}{data}\PY{p}{)}
+    \PY{n}{md\PYZus{}}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{p}{\PYZob{}}
+        \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}from}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:} \PY{n}{aux}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}id}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{,}
+        \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}to}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:} \PY{n}{data}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{info}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{orig\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{,}
+        \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{vertex}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:} \PY{n}{aux}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}key}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}
+    \PY{p}{\PYZcb{}}\PY{p}{)}
+\end{Verbatim}
+\end{tcolorbox}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{134}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{n}{dpath}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/home/vitor/Documents/energy\PYZus{}scripts\PYZus{}superpc/models/svr}\PY{l+s+s2}{\PYZdq{}}
+\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{os}\PY{o}{.}\PY{n}{listdir}\PY{p}{(}\PY{n}{dpath}\PY{p}{)}\PY{p}{:}
+    \PY{k}{if} \PY{o+ow}{not} \PY{n}{f}\PY{o}{.}\PY{n}{endswith}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{:} \PY{k}{continue}
+    \PY{n}{fname}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{dpath}\PY{p}{,}\PY{n}{f}\PY{p}{)}
+    \PY{n}{data}\PY{o}{=} \PY{n}{json}\PY{o}{.}\PY{n}{load}\PY{p}{(}\PY{n+nb}{open}\PY{p}{(}\PY{n}{fname}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{r+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
+    \PY{n}{aux}\PY{o}{=} \PY{n}{model\PYZus{}col}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{n}{data}\PY{p}{)}
+    \PY{n}{md\PYZus{}}\PY{o}{.}\PY{n}{insert}\PY{p}{(}\PY{p}{\PYZob{}}
+        \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}from}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:} \PY{n}{aux}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}id}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{,}
+        \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}to}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:} \PY{n}{data}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{info}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{orig\PYZus{}data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{,}
+        \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{vertex}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:} \PY{n}{aux}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZus{}key}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}
+    \PY{p}{\PYZcb{}}\PY{p}{)}
+\end{Verbatim}
+\end{tcolorbox}
+
+    \hypertarget{advantages}{%
+\section{Advantages}\label{advantages}}
+
+\begin{itemize}
+\tightlist
+\item
+  No need to organize the files
+\item
+  Faster and easier queries
+\item
+  Shared data across programs and research
+\end{itemize}
+
+\hypertarget{disadvantages}{%
+\section{Disadvantages}\label{disadvantages}}
+
+\begin{itemize}
+\tightlist
+\item
+  Need of db server
+\item
+  Need of network
+\item
+  Non local server maybe slow ( need to cache )
+\end{itemize}
+
+    \hypertarget{examples}{%
+\section{Examples}\label{examples}}
+
+\hypertarget{group-data-of-the-same-run}{%
+\subsection{Group data of the same
+run}\label{group-data-of-the-same-run}}
+
+    \hypertarget{ondemand-groupby-argument}{%
+\subsection{Ondemand groupby argument}\label{ondemand-groupby-argument}}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{111}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{n}{ondemand} \PY{o}{=} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{aql}\PY{o}{.}\PY{n}{execute}\PY{p}{(}
+\PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}}
+\PY{l+s+sd}{for d in run}
+\PY{l+s+sd}{    filter d.config.governor == \PYZdq{}ondemand\PYZdq{}}
+\PY{l+s+sd}{    filter d.config.data\PYZus{}descriptor.keys == [\PYZdq{}cores\PYZdq{}, \PYZdq{}input\PYZdq{}, \PYZdq{}repetitions\PYZdq{}]}
+\PY{l+s+sd}{    collect args = d.config.arguments  into same\PYZus{}arg}
+\PY{l+s+sd}{    return }
+\PY{l+s+sd}{    \PYZob{}}
+\PY{l+s+sd}{        keys: same\PYZus{}arg[*].d.\PYZus{}key,}
+\PY{l+s+sd}{        pkgs: same\PYZus{}arg[*].d.config.pkg,}
+\PY{l+s+sd}{        args}
+\PY{l+s+sd}{    \PYZcb{}}
+\PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}}\PY{p}{)}
+\PY{k}{for} \PY{n}{d} \PY{o+ow}{in} \PY{n}{ondemand}\PY{p}{:}
+    \PY{n+nb}{print}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{pkgs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
+    \PY{n+nb}{print}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{args}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
+\end{Verbatim}
+\end{tcolorbox}
+
+    \begin{Verbatim}[commandchars=\\\{\}]
+['./canneal']
+['\_\_nt\_\_ 15000 2000 2500000.nets 128', '\_\_nt\_\_ 15000 2000 2500000.nets 256',
+'\_\_nt\_\_ 15000 2000 2500000.nets 384', '\_\_nt\_\_ 15000 2000 2500000.nets 512',
+'\_\_nt\_\_ 15000 2000 2500000.nets 640']
+['./canneal', './canneal', './canneal']
+['\_\_nt\_\_ 15000 2000 2500000.nets 3600', '\_\_nt\_\_ 15000 2000 2500000.nets 4200',
+'\_\_nt\_\_ 15000 2000 2500000.nets 4800', '\_\_nt\_\_ 15000 2000 2500000.nets 5400',
+'\_\_nt\_\_ 15000 2000 2500000.nets 6000']
+['./fluidanimate', './fluidanimate']
+['\_\_nt\_\_ 200 in\_500K.fluid out.fluid', '\_\_nt\_\_ 300 in\_500K.fluid out.fluid',
+'\_\_nt\_\_ 400 in\_500K.fluid out.fluid', '\_\_nt\_\_ 500 in\_500K.fluid out.fluid',
+'\_\_nt\_\_ 600 in\_500K.fluid out.fluid']
+['./xhpl', './xhpl', './xhpl']
+['\_\_nt\_\_ 7000', '\_\_nt\_\_ 8000', '\_\_nt\_\_ 9000', '\_\_nt\_\_ 10000', '\_\_nt\_\_ 11000']
+['./blackscholes', './blackscholes', './blackscholes']
+['\_\_nt\_\_ in\_5\_10.txt output.txt', '\_\_nt\_\_ in\_6\_10.txt output.txt', '\_\_nt\_\_
+in\_7\_10.txt output.txt', '\_\_nt\_\_ in\_8\_10.txt output.txt', '\_\_nt\_\_ in\_9\_10.txt
+output.txt']
+['./x264', './x264', './x264']
+['--quiet --qp 20 --partitions b8x8,i4x4 --ref 5 --direct auto --b-pyramid
+--weightb --mixed-refs --no-fast-pskip --me umh --subme 7 --analyse b8x8,i4x4
+--threads \_\_nt\_\_ -o eledream.264 eledream\_1920x1080\_306.y4m', '--quiet --qp 20
+--partitions b8x8,i4x4 --ref 5 --direct auto --b-pyramid --weightb --mixed-refs
+--no-fast-pskip --me umh --subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o
+eledream.264 eledream\_1920x1080\_357.y4m', '--quiet --qp 20 --partitions
+b8x8,i4x4 --ref 5 --direct auto --b-pyramid --weightb --mixed-refs --no-fast-
+pskip --me umh --subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
+eledream\_1920x1080\_408.y4m', '--quiet --qp 20 --partitions b8x8,i4x4 --ref 5
+--direct auto --b-pyramid --weightb --mixed-refs --no-fast-pskip --me umh
+--subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
+eledream\_1920x1080\_459.y4m', '--quiet --qp 20 --partitions b8x8,i4x4 --ref 5
+--direct auto --b-pyramid --weightb --mixed-refs --no-fast-pskip --me umh
+--subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
+eledream\_1920x1080\_512.y4m']
+['./dedup', './dedup', './dedup', './dedup']
+['-c -p -v -t \_\_nt\_\_ -i FC-6-x86\_64-disc1\_05.iso -o output.dat.ddp', '-c -p -v
+-t \_\_nt\_\_ -i FC-6-x86\_64-disc1\_06.iso -o output.dat.ddp', '-c -p -v -t \_\_nt\_\_ -i
+FC-6-x86\_64-disc1\_07.iso -o output.dat.ddp', '-c -p -v -t \_\_nt\_\_ -i
+FC-6-x86\_64-disc1\_08.iso -o output.dat.ddp', '-c -p -v -t \_\_nt\_\_ -i
+FC-6-x86\_64-disc1\_09.iso -o output.dat.ddp']
+['./swaptions', './swaptions', './swaptions', './swaptions']
+['-ns 32 -sm 2000000 -nt \_\_nt\_\_', '-ns 32 -sm 3000000 -nt \_\_nt\_\_', '-ns 32 -sm
+4000000 -nt \_\_nt\_\_', '-ns 32 -sm 5000000 -nt \_\_nt\_\_', '-ns 32 -sm 6000000 -nt
+\_\_nt\_\_']
+['./ferret', './ferret', './ferret']
+['corel\_5 lsh queries\_5 10 20 \_\_nt\_\_ output.txt', 'corel\_6 lsh queries\_6 10 20
+\_\_nt\_\_ output.txt', 'corel\_7 lsh queries\_7 10 20 \_\_nt\_\_ output.txt', 'corel\_8
+lsh queries\_8 10 20 \_\_nt\_\_ output.txt', 'corel\_9 lsh queries\_9 10 20 \_\_nt\_\_
+output.txt']
+['./vips', './vips', './vips']
+['im\_benchmark orion\_10800x10800.v output.v', 'im\_benchmark orion\_12600x12600.v
+output.v', 'im\_benchmark orion\_14400x14400.v output.v', 'im\_benchmark
+orion\_16200x16200.v output.v', 'im\_benchmark orion\_18000x18000.v output.v']
+['./openmc', './openmc', './openmc']
+['input1', 'input2', 'input3', 'input4', 'input5']
+['./bodytrack', './bodytrack', './bodytrack', './bodytrack']
+['sequenceB\_1043 4 204 1000 10 0 \_\_nt\_\_', 'sequenceB\_1043 4 408 1000 10 0
+\_\_nt\_\_', 'sequenceB\_1043 4 612 1000 10 0 \_\_nt\_\_', 'sequenceB\_1043 4 816 1000 10
+0 \_\_nt\_\_', 'sequenceB\_1043 4 1020 1000 10 0 \_\_nt\_\_']
+['./rtview', './rtview', './rtview']
+['thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 346 346',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 374 374',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 400 400',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 424 424',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 450 450']
+['./freqmine', './freqmine', './freqmine']
+['webdocs\_250k\_05.dat 11000', 'webdocs\_250k\_06.dat 11000', 'webdocs\_250k\_07.dat
+11000', 'webdocs\_250k\_08.dat 11000', 'webdocs\_250k\_09.dat 11000']
+    \end{Verbatim}
+
+    \hypertarget{performance-groupby-argument}{%
+\subsection{Performance groupby
+argument}\label{performance-groupby-argument}}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{112}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{n}{performance} \PY{o}{=} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{aql}\PY{o}{.}\PY{n}{execute}\PY{p}{(}
+\PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}}
+\PY{l+s+sd}{for d in run}
+\PY{l+s+sd}{    filter d.config.data\PYZus{}descriptor.extras.sensors}
+\PY{l+s+sd}{    filter d.config.data\PYZus{}descriptor.keys == [\PYZdq{}cores\PYZdq{}, \PYZdq{}frequency\PYZdq{}, \PYZdq{}input\PYZdq{}, \PYZdq{}repetitions\PYZdq{}]}
+\PY{l+s+sd}{    collect args = d.config.arguments  into same\PYZus{}arg}
+\PY{l+s+sd}{    return }
+\PY{l+s+sd}{    \PYZob{}}
+\PY{l+s+sd}{        keys: same\PYZus{}arg[*].d.\PYZus{}key,}
+\PY{l+s+sd}{        pkgs: same\PYZus{}arg[*].d.config.pkg,}
+\PY{l+s+sd}{        args}
+\PY{l+s+sd}{    \PYZcb{}}
+\PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}}\PY{p}{)}
+\PY{k}{for} \PY{n}{d} \PY{o+ow}{in} \PY{n}{performance}\PY{p}{:}
+    \PY{n+nb}{print}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{pkgs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
+    \PY{c+c1}{\PYZsh{}print(d[\PYZdq{}args\PYZdq{}])}
+\end{Verbatim}
+\end{tcolorbox}
+
+    \begin{Verbatim}[commandchars=\\\{\}]
+['completo\_canneal\_1', 'completo\_canneal\_3', 'completo\_canneal\_4',
+'completo\_canneal\_2']
+['completo\_xhpl\_2', 'completo\_xhpl']
+['completo\_black\_3', 'completo\_black\_2', 'completo\_black\_1']
+['completo\_black\_4']
+['completo\_black\_6', 'completo\_black\_5']
+['completo\_x264\_4']
+['completo\_x264\_2', 'completo\_x264\_1', 'completo\_x264\_3']
+['completo\_dedup\_2', 'completo\_dedup\_1', 'completo\_dedup\_4', 'completo\_dedup\_3']
+['completo\_swaptions\_3', 'completo\_swaptions\_2']
+['completo\_swaptions\_1']
+['completo\_fluid\_2']
+['completo\_fluid\_4', 'completo\_fluid\_3']
+['completo\_fluid\_1']
+['completo\_ferret\_1', 'completo\_ferret\_2', 'completo\_ferret\_3',
+'completo\_ferret\_4']
+['completo\_vips\_2', 'completo\_vips\_3', 'completo\_vips\_4', 'completo\_vips\_1']
+['completo\_openmc\_kernel\_novo']
+['completo\_rtview\_1']
+['completo\_bodytrack\_3']
+['completo\_bodytrack\_2', 'completo\_bodytrack\_1']
+['completo\_rtview\_2']
+['completo\_rtview\_3']
+['completo\_rtview\_4']
+['completo\_freqmine\_2', 'completo\_freqmine\_1']
+['completo\_freq']
+    \end{Verbatim}
+
+    \hypertarget{groupby-keys}{%
+\subsection{Groupby keys}\label{groupby-keys}}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{174}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{n}{ts}\PY{o}{=} \PY{l+m+mf}{0.9}
+\PY{n}{eq}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}\PYZdq{}}\PY{l+s+s2}{equation}\PY{l+s+se}{\PYZbs{}\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}
+\PY{n}{performance} \PY{o}{=} \PY{n}{energy\PYZus{}db}\PY{o}{.}\PY{n}{aql}\PY{o}{.}\PY{n}{execute}\PY{p}{(}
+\PY{l+s+sa}{f}\PY{l+s+s2}{\PYZdq{}\PYZdq{}\PYZdq{}}
+\PY{l+s+s2}{for d in model}
+\PY{l+s+s2}{    filter d.info.type == }\PY{l+s+si}{\PYZob{}eq\PYZcb{}}
+\PY{l+s+s2}{    filter d.train\PYZus{}sz == }\PY{l+s+si}{\PYZob{}ts\PYZcb{}}
+\PY{l+s+s2}{    collect args = d.config.arguments  into same\PYZus{}arg}
+\PY{l+s+s2}{    return }
+\PY{l+s+s2}{    }\PY{l+s+s2}{\PYZob{}\PYZob{}}
+\PY{l+s+s2}{        keys: same\PYZus{}arg[*].d.\PYZus{}key,}
+\PY{l+s+s2}{        pkgs: same\PYZus{}arg[*].d.config.pkg,}
+\PY{l+s+s2}{        maes: same\PYZus{}arg[*].d.info.mae,}
+\PY{l+s+s2}{        args}
+\PY{l+s+s2}{    \PYZcb{}\PYZcb{}}
+\PY{l+s+s2}{\PYZdq{}\PYZdq{}\PYZdq{}}\PY{p}{)}
+\PY{k}{for} \PY{n}{d} \PY{o+ow}{in} \PY{n}{performance}\PY{p}{:}
+    \PY{n}{idx\PYZus{}min}\PY{o}{=} \PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{maes}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{o}{.}\PY{n}{index}\PY{p}{(}\PY{n+nb}{min}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{maes}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}\PY{p}{)}
+    \PY{n+nb}{print}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{pkgs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{idx\PYZus{}min}\PY{p}{]}\PY{p}{)}
+    \PY{n+nb}{print}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{args}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
+    \PY{n+nb}{print}\PY{p}{(}\PY{n}{d}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{keys}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{idx\PYZus{}min}\PY{p}{]}\PY{p}{)}
+    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{*}\PY{l+m+mi}{100}\PY{p}{)}
+\end{Verbatim}
+\end{tcolorbox}
+
+    \begin{Verbatim}[commandchars=\\\{\}]
+completo\_canneal\_2
+['\_\_nt\_\_ 15000 2000 2500000.nets 128', '\_\_nt\_\_ 15000 2000 2500000.nets 256',
+'\_\_nt\_\_ 15000 2000 2500000.nets 384', '\_\_nt\_\_ 15000 2000 2500000.nets 512',
+'\_\_nt\_\_ 15000 2000 2500000.nets 640']
+64562
+--------------------------------------------------------------------------------
+--------------------
+completo\_xhpl
+['\_\_nt\_\_ 7000', '\_\_nt\_\_ 8000', '\_\_nt\_\_ 9000', '\_\_nt\_\_ 10000', '\_\_nt\_\_ 11000']
+64574
+--------------------------------------------------------------------------------
+--------------------
+completo\_black\_1
+['\_\_nt\_\_ in\_156K.txt out\_156K.txt', '\_\_nt\_\_ in\_312K.txt out\_312K.txt', '\_\_nt\_\_
+in\_625K.txt out\_625K.txt', '\_\_nt\_\_ in\_1M.txt out\_1M.txt', '\_\_nt\_\_ in\_2M.txt
+out\_2M.txt', '\_\_nt\_\_ in\_5M.txt out\_5M.txt', '\_\_nt\_\_ in\_10M.txt out\_10M.txt']
+64592
+--------------------------------------------------------------------------------
+--------------------
+completo\_black\_4
+['\_\_nt\_\_ in\_312K.txt out\_312K.txt', '\_\_nt\_\_ in\_625K.txt out\_625K.txt', '\_\_nt\_\_
+in\_1M.txt out\_1M.txt', '\_\_nt\_\_ in\_2M.txt out\_2M.txt', '\_\_nt\_\_ in\_5M.txt
+out\_5M.txt', '\_\_nt\_\_ in\_10M.txt out\_10M.txt']
+64603
+--------------------------------------------------------------------------------
+--------------------
+completo\_black\_5
+['\_\_nt\_\_ in\_5\_10.txt output.txt', '\_\_nt\_\_ in\_6\_10.txt output.txt', '\_\_nt\_\_
+in\_7\_10.txt output.txt', '\_\_nt\_\_ in\_8\_10.txt output.txt', '\_\_nt\_\_ in\_9\_10.txt
+output.txt']
+64617
+--------------------------------------------------------------------------------
+--------------------
+completo\_x264\_4
+['--quiet --qp 20 --partitions b8x8,i4x4 --ref 5 --direct auto --b-pyramid
+--weightb --mixed-refs --no-fast-pskip --me umh --subme 7 --analyse b8x8,i4x4
+--threads \_\_nt\_\_ -o eledream.264 eledream\_1920x1080\_306.y4m', '--quiet --qp 20
+--partitions b8x8,i4x4 --ref 5 --direct auto --b-pyramid --weightb --mixed-refs
+--no-fast-pskip --me umh --subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o
+eledream.264 eledream\_1920x1080\_357.y4m', '--quiet --qp 20 --partitions
+b8x8,i4x4 --ref 5 --direct auto --b-pyramid --weightb --mixed-refs --no-fast-
+pskip --me umh --subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
+eledream\_1920x1080\_408.y4m', '--quiet --qp 20 --partitions b8x8,i4x4 --ref 5
+--direct auto --b-pyramid --weightb --mixed-refs --no-fast-pskip --me umh
+--subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
+eledream\_1920x1080\_459.y4m', '--quiet --qp 20 --partitions b8x8,i4x4 --ref 5
+--direct auto --b-pyramid --weightb --mixed-refs --no-fast-pskip --me umh
+--subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
+eledream\_1920x1080\_512.y4m']
+64623
+--------------------------------------------------------------------------------
+--------------------
+completo\_x264\_1
+['--quiet --qp 20 --partitions b8x8,i4x4 --ref 5 --direct auto --b-pyramid
+--weightb --mixed-refs --no-fast-pskip --me umh --subme 7 --analyse b8x8,i4x4
+--threads \_\_nt\_\_ -o eledream.264 eledream\_1920x1080\_512.y4m', '--quiet --qp 20
+--partitions b8x8,i4x4 --ref 5 --direct auto --b-pyramid --weightb --mixed-refs
+--no-fast-pskip --me umh --subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o
+eledream.264 eledream\_1920x1080\_459.y4m', '--quiet --qp 20 --partitions
+b8x8,i4x4 --ref 5 --direct auto --b-pyramid --weightb --mixed-refs --no-fast-
+pskip --me umh --subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
+eledream\_1920x1080\_408.y4m', '--quiet --qp 20 --partitions b8x8,i4x4 --ref 5
+--direct auto --b-pyramid --weightb --mixed-refs --no-fast-pskip --me umh
+--subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
+eledream\_1920x1080\_357.y4m', '--quiet --qp 20 --partitions b8x8,i4x4 --ref 5
+--direct auto --b-pyramid --weightb --mixed-refs --no-fast-pskip --me umh
+--subme 7 --analyse b8x8,i4x4 --threads \_\_nt\_\_ -o eledream.264
+eledream\_1920x1080\_306.y4m']
+64636
+--------------------------------------------------------------------------------
+--------------------
+completo\_dedup\_3
+['-c -p -v -t \_\_nt\_\_ -i FC-6-x86\_64-disc1\_05.iso -o output.dat.ddp', '-c -p -v
+-t \_\_nt\_\_ -i FC-6-x86\_64-disc1\_06.iso -o output.dat.ddp', '-c -p -v -t \_\_nt\_\_ -i
+FC-6-x86\_64-disc1\_07.iso -o output.dat.ddp', '-c -p -v -t \_\_nt\_\_ -i
+FC-6-x86\_64-disc1\_08.iso -o output.dat.ddp', '-c -p -v -t \_\_nt\_\_ -i
+FC-6-x86\_64-disc1\_09.iso -o output.dat.ddp']
+64667
+--------------------------------------------------------------------------------
+--------------------
+completo\_swaptions\_2
+['-ns 32 -sm 2000000 -nt \_\_nt\_\_', '-ns 32 -sm 3000000 -nt \_\_nt\_\_', '-ns 32 -sm
+4000000 -nt \_\_nt\_\_', '-ns 32 -sm 5000000 -nt \_\_nt\_\_', '-ns 32 -sm 6000000 -nt
+\_\_nt\_\_']
+64680
+--------------------------------------------------------------------------------
+--------------------
+completo\_swaptions\_1
+['-ns 64 -sm 2000000 -nt \_\_nt\_\_', '-ns 64 -sm 3000000 -nt \_\_nt\_\_', '-ns 64 -sm
+4000000 -nt \_\_nt\_\_', '-ns 64 -sm 5000000 -nt \_\_nt\_\_', '-ns 64 -sm 6000000 -nt
+\_\_nt\_\_', '-ns 64 -sm 7000000 -nt \_\_nt\_\_']
+64686
+--------------------------------------------------------------------------------
+--------------------
+completo\_fluid\_2
+['1 125 in\_500K.fluid', '1 250 in\_500K.fluid', '1 500 in\_500K.fluid', '1 1000
+in\_500K.fluid', '1 2000 in\_500K.fluid', '1 4000 in\_500K.fluid']
+64697
+--------------------------------------------------------------------------------
+--------------------
+completo\_fluid\_4
+['1 200 in\_500K.fluid out.fluid', '1 300 in\_500K.fluid out.fluid', '1 400
+in\_500K.fluid out.fluid', '1 500 in\_500K.fluid out.fluid', '1 600 in\_500K.fluid
+out.fluid']
+64703
+--------------------------------------------------------------------------------
+--------------------
+completo\_fluid\_1
+['1 62 in\_500K.fluid', '1 125 in\_500K.fluid', '1 250 in\_500K.fluid', '1 500
+in\_500K.fluid', '1 1000 in\_500K.fluid', '1 2000 in\_500K.fluid', '1 4000
+in\_500K.fluid']
+64715
+--------------------------------------------------------------------------------
+--------------------
+completo\_ferret\_1
+['corel\_5 lsh queries\_5 10 20 \_\_nt\_\_ output.txt', 'corel\_6 lsh queries\_6 10 20
+\_\_nt\_\_ output.txt', 'corel\_7 lsh queries\_7 10 20 \_\_nt\_\_ output.txt', 'corel\_8
+lsh queries\_8 10 20 \_\_nt\_\_ output.txt', 'corel\_9 lsh queries\_9 10 20 \_\_nt\_\_
+output.txt']
+64721
+--------------------------------------------------------------------------------
+--------------------
+completo\_vips\_4
+['im\_benchmark orion\_10800x10800.v output.v', 'im\_benchmark orion\_12600x12600.v
+output.v', 'im\_benchmark orion\_14400x14400.v output.v', 'im\_benchmark
+orion\_16200x16200.v output.v', 'im\_benchmark orion\_18000x18000.v output.v']
+64757
+--------------------------------------------------------------------------------
+--------------------
+completo\_openmc\_kernel\_novo
+['input1', 'input2', 'input3', 'input4', 'input5']
+64769
+--------------------------------------------------------------------------------
+--------------------
+completo\_rtview\_1
+['nave.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 100 100', 'nave.obj
+-automove -nthreads \_\_nt\_\_ -frames 10000 -res 140 140', 'nave.obj -automove
+-nthreads \_\_nt\_\_ -frames 10000 -res 196 196', 'nave.obj -automove -nthreads
+\_\_nt\_\_ -frames 10000 -res 274 274', 'nave.obj -automove -nthreads \_\_nt\_\_ -frames
+10000 -res 384 384', 'nave.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 537
+537']
+64775
+--------------------------------------------------------------------------------
+--------------------
+completo\_bodytrack\_3
+['sequenceB\_1043 4 204 1000 10 0 \_\_nt\_\_', 'sequenceB\_1043 4 408 1000 10 0
+\_\_nt\_\_', 'sequenceB\_1043 4 612 1000 10 0 \_\_nt\_\_', 'sequenceB\_1043 4 816 1000 10
+0 \_\_nt\_\_', 'sequenceB\_1043 4 1020 1000 10 0 \_\_nt\_\_']
+64781
+--------------------------------------------------------------------------------
+--------------------
+completo\_bodytrack\_1
+['sequenceB\_1043 4 64 1000 10 0 \_\_nt\_\_', 'sequenceB\_1043 4 128 1000 10 0
+\_\_nt\_\_', 'sequenceB\_1043 4 256 1000 10 0 \_\_nt\_\_', 'sequenceB\_1043 4 512 1000 10
+0 \_\_nt\_\_', 'sequenceB\_1043 4 1024 1000 10 0 \_\_nt\_\_']
+64798
+--------------------------------------------------------------------------------
+--------------------
+completo\_rtview\_2
+['thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 100 100',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 140 140',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 196 196',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 274 274',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 384 384',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 537 537']
+64805
+--------------------------------------------------------------------------------
+--------------------
+completo\_rtview\_3
+['thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 100 100',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 200 200',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 300 300',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 400 400',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 500 500']
+64814
+--------------------------------------------------------------------------------
+--------------------
+completo\_rtview\_4
+['thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 346 346',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 374 374',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 400 400',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 424 424',
+'thai\_statue.obj -automove -nthreads \_\_nt\_\_ -frames 10000 -res 450 450']
+64828
+--------------------------------------------------------------------------------
+--------------------
+completo\_freqmine\_1
+['webdocs\_250k\_05.dat 11000', 'webdocs\_250k\_06.dat 11000', 'webdocs\_250k\_07.dat
+11000', 'webdocs\_250k\_08.dat 11000', 'webdocs\_250k\_09.dat 11000']
+64849
+--------------------------------------------------------------------------------
+--------------------
+completo\_freq
+['webdocs.dat 90000']
+64857
+--------------------------------------------------------------------------------
+--------------------
+    \end{Verbatim}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{13}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
 \PY{k+kn}{import} \PY{n+nn}{pandas} \PY{k}{as} \PY{n+nn}{pd}
 \PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}
-
-\PY{k}{def} \PY{n+nf}{multidict}\PY{p}{(}\PY{n}{n}\PY{p}{,} \PY{n}{dtype}\PY{o}{=}\PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{:}
-    \PY{k}{if} \PY{n}{n} \PY{o}{==} \PY{l+m+mi}{0}\PY{p}{:} \PY{k}{return} \PY{n}{dtype}
-    \PY{k}{return} \PY{n}{defaultdict}\PY{p}{(}\PY{k}{lambda}\PY{p}{:}\PY{n}{multidict}\PY{p}{(}\PY{n}{n}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,}\PY{n}{dtype}\PY{p}{)}\PY{p}{)}
-
-\PY{n}{nested\PYZus{}dict} \PY{o}{=} \PY{k}{lambda}\PY{p}{:} \PY{n}{defaultdict}\PY{p}{(}\PY{n}{nested\PYZus{}dict}\PY{p}{)}
-
-\PY{n}{freqs}\PY{o}{=} \PY{p}{[}\PY{l+m+mf}{1.2}\PY{p}{,}\PY{l+m+mf}{2.2}\PY{p}{]}
-\PY{n}{cores}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{32}\PY{p}{]}
-\PY{n}{inps}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{]}
-\PY{n}{regions}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{]}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{n}{measures} \PY{o}{=} \PY{n}{nested\PYZus{}dict}\PY{p}{(}\PY{p}{)}
-\PY{c+c1}{\PYZsh{} measures= defaultdict(lambda:defaultdict(lambda:defaultdict(lambda:\PYZob{}\PYZcb{})))}
-
-\PY{c+c1}{\PYZsh{} collect every n milliseconds + event}
-\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
-    \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
-        \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
-            \PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:}
-                \PY{n}{measures}\PY{p}{[}\PY{n}{f}\PY{p}{]}\PY{p}{[}\PY{n}{c}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{]}\PY{p}{[}\PY{n}{e}\PY{p}{]}\PY{o}{=} \PY{p}{\PYZob{}}
-                    \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{timestamp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{0.1}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{200}\PY{p}{\PYZcb{}}\PY{p}{,} \PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{timestamp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{0.2}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{220}\PY{p}{\PYZcb{}}\PY{p}{]}\PY{p}{,}
-                    \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{total\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{10}\PY{p}{,}
-                    \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{total\PYZus{}energy}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{10}
-                    \PY{p}{\PYZcb{}}
-                \PY{c+c1}{\PYZsh{} measures[e][f][c][i]= \PYZob{}}
-                \PY{c+c1}{\PYZsh{}     \PYZdq{}sensor\PYZdq{}:[\PYZob{}\PYZdq{}timestamp\PYZdq{}:0.1,\PYZdq{}data\PYZdq{}:200\PYZcb{}, \PYZob{}\PYZdq{}timestamp\PYZdq{}:0.2,\PYZdq{}data\PYZdq{}:220\PYZcb{}],}
-                \PY{c+c1}{\PYZsh{}     \PYZdq{}total\PYZus{}time\PYZdq{}:10,}
-                \PY{c+c1}{\PYZsh{}     \PYZdq{}total\PYZus{}energy\PYZdq{}:10}
-                \PY{c+c1}{\PYZsh{}     \PYZcb{}}
-                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
-                    \PY{n}{measures}\PY{p}{[}\PY{n}{f}\PY{p}{]}\PY{p}{[}\PY{n}{c}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{]}\PY{p}{[}\PY{n}{e}\PY{p}{]}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{o}{=} \PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{timestamp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{0.32}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{150}\PY{p}{\PYZcb{}}\PY{p}{\PYZcb{}}\PY{p}{,}\PY{p}{]}
-                    \PY{c+c1}{\PYZsh{} measures[e][r][f][c][i]= [\PYZob{}\PYZdq{}start\PYZus{}time\PYZdq{}:0,\PYZdq{}stop\PYZus{}time\PYZdq{}:1,\PYZdq{}sensor\PYZdq{}:\PYZob{}\PYZdq{}timestamp\PYZdq{}:0.32,\PYZdq{}data\PYZdq{}:150\PYZcb{}\PYZcb{},]}
-
-\PY{n+nb}{print}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{measures}\PY{p}{,} \PY{n}{indent}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{)}\PY{p}{)}
-\PY{n+nb}{print}\PY{p}{(}\PY{p}{)}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-\{
-    "1.2": \{
-        "16": \{
-            "2": \{
-                "0": \{
-                    "sensor": [
-                        \{
-                            "timestamp": 0.1,
-                            "data": 200
-                        \},
-                        \{
-                            "timestamp": 0.2,
-                            "data": 220
-                        \}
-                    ],
-                    "total\_time": 10,
-                    "total\_energy": 10,
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ]
-                \}
-            \},
-            "5": \{
-                "0": \{
-                    "sensor": [
-                        \{
-                            "timestamp": 0.1,
-                            "data": 200
-                        \},
-                        \{
-                            "timestamp": 0.2,
-                            "data": 220
-                        \}
-                    ],
-                    "total\_time": 10,
-                    "total\_energy": 10,
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ]
-                \}
-            \}
-        \},
-        "32": \{
-            "2": \{
-                "0": \{
-                    "sensor": [
-                        \{
-                            "timestamp": 0.1,
-                            "data": 200
-                        \},
-                        \{
-                            "timestamp": 0.2,
-                            "data": 220
-                        \}
-                    ],
-                    "total\_time": 10,
-                    "total\_energy": 10,
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ]
-                \}
-            \},
-            "5": \{
-                "0": \{
-                    "sensor": [
-                        \{
-                            "timestamp": 0.1,
-                            "data": 200
-                        \},
-                        \{
-                            "timestamp": 0.2,
-                            "data": 220
-                        \}
-                    ],
-                    "total\_time": 10,
-                    "total\_energy": 10,
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ]
-                \}
-            \}
-        \}
-    \},
-    "2.2": \{
-        "16": \{
-            "2": \{
-                "0": \{
-                    "sensor": [
-                        \{
-                            "timestamp": 0.1,
-                            "data": 200
-                        \},
-                        \{
-                            "timestamp": 0.2,
-                            "data": 220
-                        \}
-                    ],
-                    "total\_time": 10,
-                    "total\_energy": 10,
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ]
-                \}
-            \},
-            "5": \{
-                "0": \{
-                    "sensor": [
-                        \{
-                            "timestamp": 0.1,
-                            "data": 200
-                        \},
-                        \{
-                            "timestamp": 0.2,
-                            "data": 220
-                        \}
-                    ],
-                    "total\_time": 10,
-                    "total\_energy": 10,
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ]
-                \}
-            \}
-        \},
-        "32": \{
-            "2": \{
-                "0": \{
-                    "sensor": [
-                        \{
-                            "timestamp": 0.1,
-                            "data": 200
-                        \},
-                        \{
-                            "timestamp": 0.2,
-                            "data": 220
-                        \}
-                    ],
-                    "total\_time": 10,
-                    "total\_energy": 10,
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ]
-                \}
-            \},
-            "5": \{
-                "0": \{
-                    "sensor": [
-                        \{
-                            "timestamp": 0.1,
-                            "data": 200
-                        \},
-                        \{
-                            "timestamp": 0.2,
-                            "data": 220
-                        \}
-                    ],
-                    "total\_time": 10,
-                    "total\_energy": 10,
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": \{
-                                "timestamp": 0.32,
-                                "data": 150
-                            \}
-                        \}
-                    ]
-                \}
-            \}
-        \}
-    \}
-\}
-
-    \end{Verbatim}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{n}{measures} \PY{o}{=} \PY{n}{nested\PYZus{}dict}\PY{p}{(}\PY{p}{)}
-\PY{c+c1}{\PYZsh{} measures= defaultdict(lambda:defaultdict(lambda:defaultdict(lambda:defaultdict(lambda:\PYZob{}\PYZcb{}))))}
-\PY{c+c1}{\PYZsh{} collect every event}
-\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
-    \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
-        \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
-            \PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:}
-                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
-                    \PY{n}{measures}\PY{p}{[}\PY{n}{f}\PY{p}{]}\PY{p}{[}\PY{n}{c}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{]}\PY{p}{[}\PY{n}{e}\PY{p}{]}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{o}{=} \PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{5}\PY{p}{\PYZcb{}}\PY{p}{,}\PY{p}{]}
-                    \PY{c+c1}{\PYZsh{} measures[e][r][f][c][i]= [\PYZob{}\PYZdq{}start\PYZus{}time\PYZdq{}:0,\PYZdq{}stop\PYZus{}time\PYZdq{}:1,\PYZdq{}sensor\PYZdq{}:5\PYZcb{},]}
-                    \PY{c+c1}{\PYZsh{} measures[\PYZdq{}\PYZpc{}s\PYZhy{}\PYZpc{}s\PYZhy{}\PYZpc{}s\PYZhy{}\PYZpc{}s\PYZhy{}\PYZpc{}s\PYZdq{}\PYZpc{}(f,c,i,e,r)]=[\PYZob{}\PYZdq{}start\PYZus{}time\PYZdq{}:0,\PYZdq{}stop\PYZus{}time\PYZdq{}:1,\PYZdq{}sensor\PYZdq{}:5\PYZcb{},]}
-
-\PY{n+nb}{print}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{measures}\PY{p}{,}\PY{n}{indent}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{)}\PY{p}{)}
-\PY{n+nb}{print}\PY{p}{(}\PY{p}{)}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-\{
-    "1.2": \{
-        "16": \{
-            "2": \{
-                "0": \{
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ]
-                \}
-            \},
-            "5": \{
-                "0": \{
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ]
-                \}
-            \}
-        \},
-        "32": \{
-            "2": \{
-                "0": \{
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ]
-                \}
-            \},
-            "5": \{
-                "0": \{
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ]
-                \}
-            \}
-        \}
-    \},
-    "2.2": \{
-        "16": \{
-            "2": \{
-                "0": \{
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ]
-                \}
-            \},
-            "5": \{
-                "0": \{
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ]
-                \}
-            \}
-        \},
-        "32": \{
-            "2": \{
-                "0": \{
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ]
-                \}
-            \},
-            "5": \{
-                "0": \{
-                    "2": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ],
-                    "4": [
-                        \{
-                            "start\_time": 0,
-                            "stop\_time": 1,
-                            "sensor": 5
-                        \}
-                    ]
-                \}
-            \}
-        \}
-    \}
-\}
-
-    \end{Verbatim}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{c+c1}{\PYZsh{} difficulty to transform this into a xarray, dataframe, numpy}
-\PY{c+c1}{\PYZsh{} performance of execution operations on each format}
-
-\PY{n}{measures}\PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
-\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
-    \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
-        \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
-            \PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:}
-                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
-                    \PY{n}{measures}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{f}\PY{p}{,}\PY{n}{c}\PY{p}{,}\PY{n}{i}\PY{p}{,}\PY{n}{e}\PY{p}{,}\PY{n}{r}\PY{p}{)}\PY{p}{]}\PY{o}{=}\PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{5}\PY{p}{\PYZcb{}}\PY{p}{]}
-
-\PY{n+nb}{print}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{measures}\PY{p}{,}\PY{n}{indent}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{)}\PY{p}{)}
-\PY{n+nb}{print}\PY{p}{(}\PY{p}{)}
-
-\PY{k+kn}{import} \PY{n+nn}{pandas} \PY{k}{as} \PY{n+nn}{pd}
-\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}
-    \PY{n+nb}{map}\PY{p}{(}\PY{k}{lambda} \PY{n}{x}\PY{p}{:} \PY{n}{x}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{+}\PY{p}{[}\PY{n}{measures}\PY{p}{[}\PY{n}{x}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{]}\PY{p}{,}\PY{n}{measures}\PY{p}{)}\PY{p}{,}
-    \PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}
-\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
-
-\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
-\PY{n+nb}{print}\PY{p}{(}\PY{n}{df}\PY{p}{)}
-\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-\{
-    "1.2-16-2-0-2": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "1.2-16-2-0-4": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "1.2-16-5-0-2": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "1.2-16-5-0-4": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "1.2-32-2-0-2": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "1.2-32-2-0-4": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "1.2-32-5-0-2": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "1.2-32-5-0-4": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "2.2-16-2-0-2": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "2.2-16-2-0-4": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "2.2-16-5-0-2": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "2.2-16-5-0-4": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "2.2-32-2-0-2": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "2.2-32-2-0-4": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "2.2-32-5-0-2": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ],
-    "2.2-32-5-0-4": [
-        \{
-            "start\_time": 0,
-            "stop\_time": 1,
-            "sensor": 5
-        \}
-    ]
-\}
-
-    \end{Verbatim}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-/home/vitor/anaconda3/lib/python3.7/site-packages/dask/config.py:168:
-YAMLLoadWarning: calling yaml.load() without Loader={\ldots} is deprecated, as the
-default Loader is unsafe. Please read https://msg.pyyaml.org/load for full
-details.
-  data = yaml.load(f.read()) or \{\}
-/home/vitor/anaconda3/lib/python3.7/site-packages/distributed/config.py:20:
-YAMLLoadWarning: calling yaml.load() without Loader={\ldots} is deprecated, as the
-default Loader is unsafe. Please read https://msg.pyyaml.org/load for full
-details.
-  defaults = yaml.load(f)
-    \end{Verbatim}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-                             sensor
-freq cores inputs execs reg
-1.2  16    2      0     2         5
-                        4         5
-           5      0     2         5
-                        4         5
-     32    2      0     2         5
-                        4         5
-           5      0     2         5
-                        4         5
-2.2  16    2      0     2         5
-                        4         5
-           5      0     2         5
-                        4         5
-     32    2      0     2         5
-                        4         5
-           5      0     2         5
-                        4         5
-<xarray.Dataset>
-Dimensions:  (cores: 2, execs: 1, freq: 2, inputs: 2, reg: 2)
-Coordinates:
-  * freq     (freq) object '1.2' '2.2'
-  * cores    (cores) object '16' '32'
-  * inputs   (inputs) object '2' '5'
-  * execs    (execs) object '0'
-  * reg      (reg) object '2' '4'
-Data variables:
-    sensor   (freq, cores, inputs, execs, reg) int64 5 5 5 5 5 5 {\ldots} 5 5 5 5 5 5
-    \end{Verbatim}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{c+c1}{\PYZsh{} difficulty to transform this into a xarray, dataframe, numpy}
-\PY{c+c1}{\PYZsh{} performance of execution operations on each format}
-
-\PY{n}{measures}\PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
-\PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
-    \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
-        \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
-            \PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:}
-                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
-                    \PY{n}{measures}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{f}\PY{p}{,}\PY{n}{c}\PY{p}{,}\PY{n}{i}\PY{p}{,}\PY{n}{e}\PY{p}{,}\PY{n}{r}\PY{p}{)}\PY{p}{]}\PY{o}{=}\PY{p}{[}
-                        \PY{p}{[}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{]}\PY{p}{,}
-                        \PY{p}{[}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{]}\PY{p}{,}
-                        \PY{p}{[}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{]}
-                        \PY{p}{]}
-
-\PY{k+kn}{import} \PY{n+nn}{pandas} \PY{k}{as} \PY{n+nn}{pd}
-\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}
-    \PY{n+nb}{map}\PY{p}{(}\PY{k}{lambda} \PY{n}{x}\PY{p}{:} \PY{n}{x}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{+}\PY{n}{measures}\PY{p}{[}\PY{n}{x}\PY{p}{]}\PY{p}{,}\PY{n}{measures}\PY{p}{)}\PY{p}{,}
-    \PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}
-\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
-
-\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
-\PY{n+nb}{print}\PY{p}{(}\PY{n}{df}\PY{p}{)}
-\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-                                 start       stop     sensor
-freq cores inputs execs reg
-1.2  16    2      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-           5      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-     32    2      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-           5      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-2.2  16    2      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-           5      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-     32    2      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-           5      0     2    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-                        4    [0, 1, 3]  [2, 3, 4]  [5, 4, 5]
-<xarray.Dataset>
-Dimensions:  (cores: 2, execs: 1, freq: 2, inputs: 2, reg: 2)
-Coordinates:
-  * freq     (freq) object '1.2' '2.2'
-  * cores    (cores) object '16' '32'
-  * inputs   (inputs) object '2' '5'
-  * execs    (execs) object '0'
-  * reg      (reg) object '2' '4'
-Data variables:
-    start    (freq, cores, inputs, execs, reg) object [0, 1, 3] {\ldots} [0, 1, 3]
-    stop     (freq, cores, inputs, execs, reg) object [2, 3, 4] {\ldots} [2, 3, 4]
-    sensor   (freq, cores, inputs, execs, reg) object [5, 4, 5] {\ldots} [5, 4, 5]
-    \end{Verbatim}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{n}{execs}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}
-\PY{n}{measures}\PY{o}{=} \PY{p}{[}\PY{p}{]}
-\PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n}{execs}\PY{p}{:}
-    \PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
-        \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
-            \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
-
-                \PY{n}{aux}\PY{o}{=} \PY{p}{[}\PY{p}{]}
-                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
-                    \PY{c+c1}{\PYZsh{} start\PYZus{}time stop\PYZus{}time sensor}
-                    \PY{n}{aux}\PY{o}{.}\PY{n}{append}\PY{p}{(}
-                        \PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{region}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{n}{r}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}\PY{p}{\PYZcb{}}\PY{p}{)}
-                    \PY{n}{aux}\PY{o}{.}\PY{n}{append}\PY{p}{(}
-                        \PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{region}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{n}{r}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop\PYZus{}time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{4}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}\PY{p}{\PYZcb{}}\PY{p}{)}
-
-                \PY{n}{measures}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{aux}\PY{p}{)}
-
-\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}
-\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{measures}\PY{p}{)}
-\PY{n}{idxs}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{MultiIndex}\PY{o}{.}\PY{n}{from\PYZus{}product}\PY{p}{(}\PY{p}{[}\PY{n}{execs}\PY{p}{,}\PY{n}{freqs}\PY{p}{,}\PY{n}{cores}\PY{p}{,}\PY{n}{inps}\PY{p}{]}\PY{p}{,}\PY{n}{names}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
-\PY{n}{df}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{n}{idxs}\PY{p}{)}
-
-\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
-\PY{n+nb}{print}\PY{p}{(}\PY{n}{df}\PY{p}{)}
-\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-                                                                         0  \textbackslash{}
-freq cores inputs execs
-1    1.2   16     2      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-                  5      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-           32     2      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-                  5      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-     2.2   16     2      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-                  5      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-           32     2      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-                  5      \{'region': 2, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-
-                                                                         1  \textbackslash{}
-freq cores inputs execs
-1    1.2   16     2      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-                  5      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-           32     2      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-                  5      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-     2.2   16     2      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-                  5      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-           32     2      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-                  5      \{'region': 2, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-
-                                                                         2  \textbackslash{}
-freq cores inputs execs
-1    1.2   16     2      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-                  5      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-           32     2      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-                  5      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-     2.2   16     2      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-                  5      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-           32     2      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-                  5      \{'region': 4, 'start\_time': 0, 'stop\_time': 2,{\ldots}
-
-                                                                         3
-freq cores inputs execs
-1    1.2   16     2      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-                  5      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-           32     2      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-                  5      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-     2.2   16     2      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-                  5      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-           32     2      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-                  5      \{'region': 4, 'start\_time': 3, 'stop\_time': 4,{\ldots}
-<xarray.Dataset>
-Dimensions:  (cores: 2, execs: 2, freq: 1, inputs: 2)
-Coordinates:
-  * freq     (freq) int64 1
-  * cores    (cores) float64 1.2 2.2
-  * inputs   (inputs) int64 16 32
-  * execs    (execs) int64 2 5
-Data variables:
-    0        (freq, cores, inputs, execs) object \{'region': 2, 'start\_time': 0,
-'stop\_time': 2, 'sensor': \{\}\} {\ldots} \{'region': 2, 'start\_time': 0, 'stop\_time': 2,
-'sensor': \{\}\}
-    1        (freq, cores, inputs, execs) object \{'region': 2, 'start\_time': 3,
-'stop\_time': 4, 'sensor': \{\}\} {\ldots} \{'region': 2, 'start\_time': 3, 'stop\_time': 4,
-'sensor': \{\}\}
-    2        (freq, cores, inputs, execs) object \{'region': 4, 'start\_time': 0,
-'stop\_time': 2, 'sensor': \{\}\} {\ldots} \{'region': 4, 'start\_time': 0, 'stop\_time': 2,
-'sensor': \{\}\}
-    3        (freq, cores, inputs, execs) object \{'region': 4, 'start\_time': 3,
-'stop\_time': 4, 'sensor': \{\}\} {\ldots} \{'region': 4, 'start\_time': 3, 'stop\_time': 4,
-'sensor': \{\}\}
-    \end{Verbatim}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{n}{execs}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}
-\PY{n}{measures}\PY{o}{=} \PY{p}{[}\PY{p}{]}
-\PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n}{execs}\PY{p}{:}
-    \PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
-        \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
-            \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
-                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{regions}\PY{p}{:}
-                    \PY{n}{measures}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{[}\PY{n}{e}\PY{p}{,}\PY{n}{f}\PY{p}{,}\PY{n}{c}\PY{p}{,}\PY{n}{i}\PY{p}{,}\PY{n}{r}\PY{p}{,}\PY{p}{[}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dados1}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{\PYZcb{}}\PY{p}{,}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dados2}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mi}{0}\PY{p}{\PYZcb{}}\PY{p}{]}\PY{p}{]}\PY{p}{)}
-
-\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}
-\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{measures}\PY{p}{,}\PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{exec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dados}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{exec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{reg}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
-
-\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
-\PY{n+nb}{print}\PY{p}{(}\PY{n}{df}\PY{p}{)}
-\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-                                                     dados
-exec freq cores inputs reg
-1    1.2  16    2      2    [\{'dados1': 0\}, \{'dados2': 0\}]
-                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
-                5      2    [\{'dados1': 0\}, \{'dados2': 0\}]
-                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
-          32    2      2    [\{'dados1': 0\}, \{'dados2': 0\}]
-                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
-                5      2    [\{'dados1': 0\}, \{'dados2': 0\}]
-                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
-     2.2  16    2      2    [\{'dados1': 0\}, \{'dados2': 0\}]
-                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
-                5      2    [\{'dados1': 0\}, \{'dados2': 0\}]
-                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
-          32    2      2    [\{'dados1': 0\}, \{'dados2': 0\}]
-                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
-                5      2    [\{'dados1': 0\}, \{'dados2': 0\}]
-                       4    [\{'dados1': 0\}, \{'dados2': 0\}]
-<xarray.Dataset>
-Dimensions:  (cores: 2, exec: 1, freq: 2, inputs: 2, reg: 2)
-Coordinates:
-  * exec     (exec) int64 1
-  * freq     (freq) float64 1.2 2.2
-  * cores    (cores) int64 16 32
-  * inputs   (inputs) int64 2 5
-  * reg      (reg) int64 2 4
-Data variables:
-    dados    (exec, freq, cores, inputs, reg) object [\{'dados1': 0\}, \{'dados2':
-0\}] {\ldots} [\{'dados1': 0\}, \{'dados2': 0\}]
-    \end{Verbatim}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{n}{execs}\PY{o}{=} \PY{p}{[}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{2}\PY{p}{]}
-\PY{n}{measures}\PY{o}{=} \PY{p}{[}\PY{p}{]}
-\PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n}{execs}\PY{p}{:}
-    \PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
-        \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
-            \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
-
-                \PY{n}{aux}\PY{o}{=} \PY{n}{defaultdict}\PY{p}{(}\PY{k}{lambda}\PY{p}{:}\PY{p}{[}\PY{p}{[}\PY{p}{]}\PY{p}{,}\PY{p}{[}\PY{p}{]}\PY{p}{,}\PY{p}{[}\PY{p}{]}\PY{p}{]}\PY{p}{)}
-                \PY{k}{for} \PY{n}{rx} \PY{o+ow}{in} \PY{p}{[}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{,}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{]}\PY{p}{:}
-                    \PY{c+c1}{\PYZsh{} start\PYZus{}time stop\PYZus{}time sensor}
-                    \PY{c+c1}{\PYZsh{} aux[rx].append(}
-                    \PY{c+c1}{\PYZsh{}     \PYZob{}\PYZdq{}start\PYZus{}time\PYZdq{}:0,\PYZdq{}stop\PYZus{}time\PYZdq{}:2,\PYZdq{}sensor\PYZdq{}:\PYZob{}\PYZcb{}\PYZcb{})}
-                    \PY{n}{aux}\PY{p}{[}\PY{n}{rx}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}
-                    \PY{n}{aux}\PY{p}{[}\PY{n}{rx}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}
-                    \PY{n}{aux}\PY{p}{[}\PY{n}{rx}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}\PY{p}{)}
-
-                \PY{n}{reg\PYZus{}k}\PY{o}{=} \PY{n}{aux}\PY{o}{.}\PY{n}{keys}\PY{p}{(}\PY{p}{)}
-                \PY{k}{for} \PY{n}{r} \PY{o+ow}{in} \PY{n}{aux}\PY{p}{:}
-                    \PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{)}
-                    \PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
-                    \PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{p}{)}
-                    \PY{n}{measures}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{aux}\PY{p}{[}\PY{n}{r}\PY{p}{]}\PY{p}{)}
-
-\PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k}{as} \PY{n+nn}{np}
-\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{measures}\PY{p}{,}\PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{start}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{stop}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{sensor}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
-\PY{c+c1}{\PYZsh{} df[\PYZdq{}start\PYZdq{}]= df[\PYZdq{}start\PYZdq{}].apply(lambda x: np.array(x))}
-\PY{c+c1}{\PYZsh{} df[\PYZdq{}stop\PYZdq{}]= df[\PYZdq{}stop\PYZdq{}].apply(lambda x: np.array(x))}
-\PY{n}{idxs}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{MultiIndex}\PY{o}{.}\PY{n}{from\PYZus{}product}\PY{p}{(}\PY{p}{[}\PY{n}{execs}\PY{p}{,}\PY{n}{freqs}\PY{p}{,}\PY{n}{cores}\PY{p}{,}\PY{n}{inps}\PY{p}{,}\PY{n}{reg\PYZus{}k}\PY{p}{]}\PY{p}{,}\PY{n}{names}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{execs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{cores}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inputs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{regs}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
-\PY{n}{df}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{n}{idxs}\PY{p}{)}
-
-\PY{n}{xr}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}xarray}\PY{p}{(}\PY{p}{)}
-\PY{n+nb}{print}\PY{p}{(}\PY{n}{xr}\PY{p}{)}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-<xarray.Dataset>
-Dimensions:  (cores: 2, execs: 2, freq: 2, inputs: 2, regs: 2)
-Coordinates:
-  * execs    (execs) int64 1 2
-  * freq     (freq) float64 1.2 2.2
-  * cores    (cores) int64 16 32
-  * inputs   (inputs) int64 2 5
-  * regs     (regs) int64 2 4
-Data variables:
-    start    (execs, freq, cores, inputs, regs) object [0 0] [0 0] {\ldots} [0 0]
-    stop     (execs, freq, cores, inputs, regs) object [1 1] [1 1] {\ldots} [1 1]
-    sensor   (execs, freq, cores, inputs, regs) object [\{\} \{\}] {\ldots} [\{\} \{\}]
-    \end{Verbatim}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{n}{data}\PY{o}{=} \PY{p}{[}\PY{p}{]}
-\PY{k}{for} \PY{n}{e} \PY{o+ow}{in} \PY{n}{execs}\PY{p}{:}
-    \PY{k}{for} \PY{n}{f} \PY{o+ow}{in} \PY{n}{freqs}\PY{p}{:}
-        \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{cores}\PY{p}{:}
-            \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{inps}\PY{p}{:}
-                \PY{k}{for} \PY{n}{x} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{random}\PY{o}{.}\PY{n}{randint}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{)}\PY{p}{)}\PY{p}{:}
-                    \PY{n}{data}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{[}\PY{n}{e}\PY{p}{,}\PY{n}{f}\PY{p}{,}\PY{n}{c}\PY{p}{,}\PY{n}{i}\PY{p}{,}\PY{n}{x}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
-\PY{n}{dfx}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{data}\PY{p}{,}\PY{n}{columns}\PY{o}{=}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{exec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{core}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ent}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
-\PY{n}{dfx}\PY{o}{=} \PY{n}{dfx}\PY{o}{.}\PY{n}{set\PYZus{}index}\PY{p}{(}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{exec}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{freq}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{core}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{inp}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ent}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}
-\PY{n}{dfx}
-\PY{c+c1}{\PYZsh{} xrx= dfx.to\PYZus{}xarray()}
-\PY{c+c1}{\PYZsh{} xrx}
+\PY{k+kn}{from} \PY{n+nn}{sklearn}\PY{n+nn}{.}\PY{n+nn}{model\PYZus{}selection} \PY{k+kn}{import} \PY{n}{train\PYZus{}test\PYZus{}split}
+
+\PY{n}{df}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{random}\PY{o}{.}\PY{n}{rand}\PY{p}{(}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{)}\PY{p}{)}
+
+\PY{n}{X}\PY{o}{=} \PY{n}{df}\PY{p}{[}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{]}\PY{p}{]}
+\PY{n}{Y}\PY{o}{=} \PY{n}{df}\PY{p}{[}\PY{l+m+mi}{4}\PY{p}{]}
+
+\PY{n}{Xtrain}\PY{p}{,} \PY{n}{\PYZus{}}\PY{p}{,} \PY{n}{Ytrain}\PY{p}{,} \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{train\PYZus{}test\PYZus{}split}\PY{p}{(}\PY{n}{X}\PY{p}{,} \PY{n}{Y}\PY{p}{,} \PY{n}{train\PYZus{}size}\PY{o}{=}\PY{l+m+mi}{5}\PY{p}{)}
+\PY{n}{idx}\PY{o}{=} \PY{n}{Xtrain}\PY{o}{.}\PY{n}{index}
+\end{Verbatim}
+\end{tcolorbox}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{19}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{n}{df}\PY{o}{=}\PY{n}{df}\PY{o}{.}\PY{n}{sort\PYZus{}values}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}
+\PY{n}{df}
 \end{Verbatim}
@@
             \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
-\prompt{Out}{outcolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-                        data
-exec freq core inp ent
-1    1.2  16   2   0       1
-                   1       1
-                   2       1
-                   3       1
-               5   0       1
-                   1       1
-                   2       1
-          32   2   0       1
-               5   0       1
-     2.2  16   2   0       1
-                   1       1
-                   2       1
-               5   0       1
-                   1       1
-          32   2   0       1
-               5   0       1
-                   1       1
-2    1.2  16   2   0       1
-                   1       1
-               5   0       1
-                   1       1
-          32   2   0       1
-                   1       1
-                   2       1
-               5   0       1
-                   1       1
-                   2       1
-     2.2  16   2   0       1
-                   1       1
-                   2       1
-                   3       1
-               5   0       1
-          32   2   0       1
-               5   0       1
-\end{Verbatim}
-\end{tcolorbox}
-
-    \hypertarget{benchmarks}{%
-\section{Benchmarks}\label{benchmarks}}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{k+kn}{from} \PY{n+nn}{itertools} \PY{k+kn}{import} \PY{n}{product}
-\PY{k+kn}{import} \PY{n+nn}{random}
-\PY{k+kn}{import} \PY{n+nn}{string}
-\PY{k+kn}{import} \PY{n+nn}{sys}
-\PY{k+kn}{from} \PY{n+nn}{numpy} \PY{k+kn}{import} \PY{n}{prod}
-\PY{k+kn}{import} \PY{n+nn}{pickle}
-\PY{k+kn}{import} \PY{n+nn}{gzip}
-\PY{k+kn}{import} \PY{n+nn}{os}
-\PY{k+kn}{import} \PY{n+nn}{gc}
-
-\PY{k}{def} \PY{n+nf}{si\PYZus{}format}\PY{p}{(}\PY{n}{n}\PY{p}{,} \PY{n}{precision}\PY{o}{=} \PY{l+m+mi}{2}\PY{p}{)}\PY{p}{:}
-    \PY{n}{sufx}\PY{o}{=} \PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{M}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{1e6}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{K}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+m+mf}{1e3}\PY{p}{\PYZcb{}}
-    \PY{n}{sufix}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}
-    \PY{k}{for} \PY{n}{k}\PY{p}{,}\PY{n}{v} \PY{o+ow}{in} \PY{n}{sufx}\PY{o}{.}\PY{n}{items}\PY{p}{(}\PY{p}{)}\PY{p}{:}
-        \PY{k}{if} \PY{n}{n} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{n}{v}\PY{p}{:}
-            \PY{n}{n}\PY{o}{/}\PY{o}{=}\PY{n}{v}
-            \PY{n}{sufix}\PY{o}{=} \PY{n}{k}
-            \PY{k}{break}
-    \PY{k}{return} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{:.}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{f\PYZcb{} }\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{b}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{n}\PY{p}{,}\PY{n}{precision}\PY{p}{,}\PY{n}{sufix}\PY{p}{)}
-
-\PY{k}{def} \PY{n+nf}{get\PYZus{}size}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{n}{seen}\PY{o}{=}\PY{k+kc}{None}\PY{p}{)}\PY{p}{:}
-    \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Recursively finds size of objects\PYZdq{}\PYZdq{}\PYZdq{}}
-    \PY{n}{size} \PY{o}{=} \PY{n}{sys}\PY{o}{.}\PY{n}{getsizeof}\PY{p}{(}\PY{n}{obj}\PY{p}{)}
-    \PY{k}{if} \PY{n}{seen} \PY{o+ow}{is} \PY{k+kc}{None}\PY{p}{:}
-        \PY{n}{seen} \PY{o}{=} \PY{n+nb}{set}\PY{p}{(}\PY{p}{)}
-    \PY{n}{obj\PYZus{}id} \PY{o}{=} \PY{n+nb}{id}\PY{p}{(}\PY{n}{obj}\PY{p}{)}
-    \PY{k}{if} \PY{n}{obj\PYZus{}id} \PY{o+ow}{in} \PY{n}{seen}\PY{p}{:}
-        \PY{k}{return} \PY{l+m+mi}{0}
-    \PY{c+c1}{\PYZsh{} Important mark as seen *before* entering recursion to gracefully handle}
-    \PY{c+c1}{\PYZsh{} self\PYZhy{}referential objects}
-    \PY{n}{seen}\PY{o}{.}\PY{n}{add}\PY{p}{(}\PY{n}{obj\PYZus{}id}\PY{p}{)}
-    \PY{k}{if} \PY{n+nb}{isinstance}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{n+nb}{dict}\PY{p}{)}\PY{p}{:}
-        \PY{n}{size} \PY{o}{+}\PY{o}{=} \PY{n+nb}{sum}\PY{p}{(}\PY{p}{[}\PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{v}\PY{p}{,} \PY{n}{seen}\PY{p}{)} \PY{k}{for} \PY{n}{v} \PY{o+ow}{in} \PY{n}{obj}\PY{o}{.}\PY{n}{values}\PY{p}{(}\PY{p}{)}\PY{p}{]}\PY{p}{)}
-        \PY{n}{size} \PY{o}{+}\PY{o}{=} \PY{n+nb}{sum}\PY{p}{(}\PY{p}{[}\PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{k}\PY{p}{,} \PY{n}{seen}\PY{p}{)} \PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{obj}\PY{o}{.}\PY{n}{keys}\PY{p}{(}\PY{p}{)}\PY{p}{]}\PY{p}{)}
-    \PY{k}{elif} \PY{n+nb}{hasattr}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZus{}\PYZus{}dict\PYZus{}\PYZus{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}\PY{p}{:}
-        \PY{n}{size} \PY{o}{+}\PY{o}{=} \PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{obj}\PY{o}{.}\PY{n+nv+vm}{\PYZus{}\PYZus{}dict\PYZus{}\PYZus{}}\PY{p}{,} \PY{n}{seen}\PY{p}{)}
-    \PY{k}{elif} \PY{n+nb}{hasattr}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZus{}\PYZus{}iter\PYZus{}\PYZus{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)} \PY{o+ow}{and} \PY{o+ow}{not} \PY{n+nb}{isinstance}\PY{p}{(}\PY{n}{obj}\PY{p}{,} \PY{p}{(}\PY{n+nb}{str}\PY{p}{,} \PY{n+nb}{bytes}\PY{p}{,} \PY{n+nb}{bytearray}\PY{p}{)}\PY{p}{)}\PY{p}{:}
-        \PY{n}{size} \PY{o}{+}\PY{o}{=} \PY{n+nb}{sum}\PY{p}{(}\PY{p}{[}\PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{seen}\PY{p}{)} \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n}{obj}\PY{p}{]}\PY{p}{)}
-    \PY{k}{return} \PY{n}{size}
-
-\PY{c+c1}{\PYZsh{} cores freq input exec reg=}
-\PY{k}{def} \PY{n+nf}{bench}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{,} \PY{n}{nbytes}\PY{p}{)}\PY{p}{:}
-    \PY{n}{vx}\PY{o}{=} \PY{p}{[}\PY{p}{]}
-    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{Keys:}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{prod}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{)}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ Bytes:}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
-
-\PY{c+c1}{\PYZsh{}     vx= [[ \PYZdq{}var\PYZob{}\PYZcb{}\PYZus{}\PYZob{}\PYZcb{}\PYZdq{}.format(i, j) for j in range(vs) ] for i, vs in enumerate(v\PYZus{}sizes)]}
-    \PY{k}{for} \PY{n}{i}\PY{p}{,} \PY{n}{vs} \PY{o+ow}{in} \PY{n+nb}{enumerate}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{)}\PY{p}{:}
-        \PY{n}{vx}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{[} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{var}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{\PYZus{}}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)} \PY{k}{for} \PY{n}{j} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{vs}\PY{p}{)} \PY{p}{]}\PY{p}{)}
-
-    \PY{n}{multi\PYZus{}dict} \PY{o}{=} \PY{n}{nested\PYZus{}dict}\PY{p}{(}\PY{p}{)}
-    \PY{n}{normal\PYZus{}dict} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
-    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{product}\PY{p}{(}\PY{o}{*}\PY{n}{vx}\PY{p}{)}\PY{p}{:}
-        \PY{n}{data}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{random}\PY{o}{.}\PY{n}{choices}\PY{p}{(}\PY{n}{string}\PY{o}{.}\PY{n}{ascii\PYZus{}letters}\PY{p}{,}\PY{n}{k}\PY{o}{=}\PY{n+nb}{int}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{)}\PY{p}{)}
-
-        \PY{n}{aux}\PY{o}{=} \PY{n}{multi\PYZus{}dict}
-        \PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{idx}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{:}
-            \PY{n}{aux} \PY{o}{=} \PY{n}{aux}\PY{p}{[}\PY{n}{k}\PY{p}{]}
-        \PY{n}{aux}\PY{p}{[}\PY{n}{idx}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{]}\PY{o}{=} \PY{n}{data}
-
-        \PY{n}{normal\PYZus{}dict}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{idx}\PY{p}{)}\PY{p}{]}\PY{o}{=} \PY{n}{data}
-
-    \PY{n}{s1}\PY{o}{=} \PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{multi\PYZus{}dict}\PY{p}{)}
-    \PY{n}{s2}\PY{o}{=} \PY{n}{get\PYZus{}size}\PY{p}{(}\PY{n}{normal\PYZus{}dict}\PY{p}{)}
-    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{RAM       }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s1}\PY{p}{)}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s2}\PY{p}{)}\PY{p}{,} \PY{n}{s1}\PY{o}{/}\PY{n}{s2} \PY{p}{)}
-
-    \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{w+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
-        \PY{n}{tf}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{multi\PYZus{}dict}\PY{p}{)}\PY{p}{)}
-
-    \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{w+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
-        \PY{n}{tf}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{normal\PYZus{}dict}\PY{p}{)}\PY{p}{)}
-
-    \PY{n}{s1}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
-    \PY{n}{s2}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.json}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
-    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{DISK json }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s1}\PY{p}{)}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s2}\PY{p}{)}\PY{p}{,} \PY{n}{s1}\PY{o}{/}\PY{n}{s2} \PY{p}{)}
-
-    \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.pkl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{wb+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
-        \PY{n}{pickle}\PY{o}{.}\PY{n}{dump}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{loads}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{multi\PYZus{}dict}\PY{p}{)}\PY{p}{)}\PY{p}{,} \PY{n}{tf}\PY{p}{,} \PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}
-
-    \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.pkl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{wb+}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
-        \PY{n}{pickle}\PY{o}{.}\PY{n}{dump}\PY{p}{(}\PY{n}{normal\PYZus{}dict}\PY{p}{,} \PY{n}{tf}\PY{p}{,} \PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}
-
-    \PY{n}{s1}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.pkl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
-    \PY{n}{s2}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.pkl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
-    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{DISK pkl  }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s1}\PY{p}{)}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s2}\PY{p}{)}\PY{p}{,} \PY{n}{s1}\PY{o}{/}\PY{n}{s2} \PY{p}{)}
-
-    \PY{k}{with} \PY{n}{gzip}\PY{o}{.}\PY{n}{GzipFile}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.gz}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{wb}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
-        \PY{n}{tf}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{multi\PYZus{}dict}\PY{p}{)}\PY{o}{.}\PY{n}{encode}\PY{p}{(}\PY{p}{)}\PY{p}{)}
-
-    \PY{k}{with} \PY{n}{gzip}\PY{o}{.}\PY{n}{GzipFile}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.gz}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{wb}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)} \PY{k}{as} \PY{n}{tf}\PY{p}{:}
-        \PY{n}{tf}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{json}\PY{o}{.}\PY{n}{dumps}\PY{p}{(}\PY{n}{normal\PYZus{}dict}\PY{p}{)}\PY{o}{.}\PY{n}{encode}\PY{p}{(}\PY{p}{)}\PY{p}{)}
-
-    \PY{n}{s1}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f1.gz}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
-    \PY{n}{s2}\PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{getsize}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{/tmp/f2.gz}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
-    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{DISK gz  }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s1}\PY{p}{)}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{s2}\PY{p}{)}\PY{p}{,} \PY{n}{s1}\PY{o}{/}\PY{n}{s2}\PY{p}{)}
-
-    \PY{n}{gc}\PY{o}{.}\PY{n}{collect}\PY{p}{(}\PY{p}{)}
-
-\PY{c+c1}{\PYZsh{}bench([2,3,2],10)}
-\PY{c+c1}{\PYZsh{} bench([32,10,10,10,10],10)}
-\PY{c+c1}{\PYZsh{} bench([32,10,10,10,10],1e3)}
-\PY{c+c1}{\PYZsh{} bench([32,10,10,10,10],100e3)}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e1}\PY{p}{)}
-\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e2}\PY{p}{)}
-\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e3}\PY{p}{)}
-\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e4}\PY{p}{)}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-
-Keys: 24.00 Kb  Bytes: 10.00 b
-
-RAM        2.57 Mb 4.73 Mb 0.542656059372869
-DISK json  616.52 Kb 1.26 Mb 0.49046778042959427
-DISK pkl   380.99 Kb 1.21 Mb 0.31507013577280013
-DISK gz   221.87 Kb 270.49 Kb 0.8202447410255462
-
-Keys: 24.00 Kb  Bytes: 100.00 b
-
-RAM        4.73 Mb 6.89 Mb 0.6860778510924217
-DISK json  2.78 Mb 3.42 Mb 0.8125601404741001
-DISK pkl   2.54 Mb 3.37 Mb 0.7541981461470956
-DISK gz   1.82 Mb 1.87 Mb 0.9686695423189734
-
-Keys: 24.00 Kb  Bytes: 1.00 Kb
-
-RAM        26.33 Mb 28.49 Mb 0.9240995518926407
-DISK json  24.38 Mb 25.02 Mb 0.974398129272095
-DISK pkl   24.22 Mb 25.04 Mb 0.9669301631449064
-DISK gz   17.58 Mb 17.64 Mb 0.9962784816626885
-
-Keys: 24.00 Kb  Bytes: 10.00 Kb
-
-RAM        242.33 Mb 244.49 Mb 0.9911560500830847
-DISK json  240.38 Mb 241.02 Mb 0.9973425857927034
-DISK pkl   240.24 Mb 241.07 Mb 0.996564850797469
-DISK gz   175.11 Mb 175.07 Mb 1.0001946831608381
-    \end{Verbatim}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{32}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{)}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-
-Keys: 320.00 Kb  Bytes: 1.00 b
-
-RAM        13.38 Mb 37.27 Mb 0.3588984160511397
-DISK json  5.23 Mb 13.98 Mb 0.3738659513590844
-DISK pkl   1.49 Mb 12.72 Mb 0.1175192426066018
-DISK gz   537.40 Kb 1.24 Mb 0.4322204295799909
-    \end{Verbatim}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{k+kn}{import} \PY{n+nn}{time}
-
-\PY{k}{def} \PY{n+nf}{bench2}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{,} \PY{n}{nbytes}\PY{p}{)}\PY{p}{:}
-    \PY{n}{vx}\PY{o}{=} \PY{p}{[}\PY{p}{]}
-    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{Keys:}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{prod}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{)}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ Bytes:}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{si\PYZus{}format}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
-    \PY{k}{for} \PY{n}{i}\PY{p}{,} \PY{n}{vs} \PY{o+ow}{in} \PY{n+nb}{enumerate}\PY{p}{(}\PY{n}{v\PYZus{}sizes}\PY{p}{)}\PY{p}{:}
-        \PY{n}{vx}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{[} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{var}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{\PYZus{}}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)} \PY{k}{for} \PY{n}{j} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{vs}\PY{p}{)} \PY{p}{]}\PY{p}{)}
-
-    \PY{n}{multi\PYZus{}dict} \PY{o}{=} \PY{n}{nested\PYZus{}dict}\PY{p}{(}\PY{p}{)}
-    \PY{n}{normal\PYZus{}dict} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
-    \PY{n}{all\PYZus{}comb}\PY{o}{=} \PY{n}{product}\PY{p}{(}\PY{o}{*}\PY{n}{vx}\PY{p}{)}
-    \PY{n}{ti}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
-    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{all\PYZus{}comb}\PY{p}{:}
-        \PY{n}{data}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{random}\PY{o}{.}\PY{n}{choices}\PY{p}{(}\PY{n}{string}\PY{o}{.}\PY{n}{ascii\PYZus{}letters}\PY{p}{,}\PY{n}{k}\PY{o}{=}\PY{n+nb}{int}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{)}\PY{p}{)}
-
-        \PY{n}{aux}\PY{o}{=} \PY{n}{multi\PYZus{}dict}
-        \PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{idx}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{:}
-            \PY{n}{aux} \PY{o}{=} \PY{n}{aux}\PY{p}{[}\PY{n}{k}\PY{p}{]}
-        \PY{n}{aux}\PY{p}{[}\PY{n}{idx}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{]}\PY{o}{=} \PY{n}{data}
-    \PY{n}{tf}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
-    \PY{n}{s1}\PY{o}{=} \PY{n}{tf}\PY{o}{\PYZhy{}}\PY{n}{ti}
-
-    \PY{n}{ti}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
-    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{all\PYZus{}comb}\PY{p}{:}
-        \PY{n}{data}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{random}\PY{o}{.}\PY{n}{choices}\PY{p}{(}\PY{n}{string}\PY{o}{.}\PY{n}{ascii\PYZus{}letters}\PY{p}{,}\PY{n}{k}\PY{o}{=}\PY{n+nb}{int}\PY{p}{(}\PY{n}{nbytes}\PY{p}{)}\PY{p}{)}\PY{p}{)}
-        \PY{n}{normal\PYZus{}dict}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{idx}\PY{p}{)}\PY{p}{]}\PY{o}{=} \PY{n}{data}
-    \PY{n}{tf}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
-    \PY{n}{s2}\PY{o}{=} \PY{n}{tf}\PY{o}{\PYZhy{}}\PY{n}{ti}
-    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Creation time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{s1}\PY{p}{,} \PY{n}{s2}\PY{p}{)}
-
-    \PY{c+c1}{\PYZsh{}multi\PYZus{}dict= json.loads(json.dumps(multi\PYZus{}dict))}
-    \PY{n}{ti}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
-    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{all\PYZus{}comb}\PY{p}{:}
-        \PY{n}{aux}\PY{o}{=} \PY{n}{multi\PYZus{}dict}
-        \PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{idx}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{:}
-            \PY{n}{aux} \PY{o}{=} \PY{n}{aux}\PY{p}{[}\PY{n}{k}\PY{p}{]}
-        \PY{n}{z}\PY{o}{=} \PY{n}{aux}\PY{p}{[}\PY{n}{idx}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{]}
-    \PY{n}{tf}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
-    \PY{n}{s1}\PY{o}{=} \PY{n}{tf}\PY{o}{\PYZhy{}}\PY{n}{ti}
-
-    \PY{n}{ti}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
-    \PY{k}{for} \PY{n}{idx} \PY{o+ow}{in} \PY{n}{all\PYZus{}comb}\PY{p}{:}
-        \PY{n}{z}\PY{o}{=} \PY{n}{normal\PYZus{}dict}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZhy{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{idx}\PY{p}{)}\PY{p}{]}
-    \PY{n}{tf}\PY{o}{=} \PY{n}{time}\PY{o}{.}\PY{n}{time}\PY{p}{(}\PY{p}{)}
-    \PY{n}{s2}\PY{o}{=} \PY{n}{tf}\PY{o}{\PYZhy{}}\PY{n}{ti}
-    \PY{n+nb}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Acess time}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{s1}\PY{p}{,} \PY{n}{s2}\PY{p}{)}
-
-    \PY{n}{gc}\PY{o}{.}\PY{n}{collect}\PY{p}{(}\PY{p}{)}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e1}\PY{p}{)}
-\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e2}\PY{p}{)}
-\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e3}\PY{p}{)}
-\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{16}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{,}\PY{l+m+mi}{5}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{10}\PY{p}{]}\PY{p}{,}\PY{l+m+mf}{1e4}\PY{p}{)}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-
-Keys: 24.00 Kb  Bytes: 10.00 b
-
-Creation time 0.12965726852416992 2.384185791015625e-07
-Acess time 4.76837158203125e-07 2.384185791015625e-07
-
-Keys: 24.00 Kb  Bytes: 100.00 b
-
-Creation time 0.6197633743286133 4.76837158203125e-07
-Acess time 2.384185791015625e-07 4.76837158203125e-07
-
-Keys: 24.00 Kb  Bytes: 1.00 Kb
-
-Creation time 5.764535903930664 2.384185791015625e-07
-Acess time 7.152557373046875e-07 2.384185791015625e-07
-
-Keys: 24.00 Kb  Bytes: 10.00 Kb
-
-Creation time 56.628950119018555 7.152557373046875e-07
-Acess time 4.76837158203125e-07 0.0
-    \end{Verbatim}
-
-    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
-\prompt{In}{incolor}{0}{\boxspacing}
-\begin{Verbatim}[commandchars=\\\{\}]
-\PY{n}{bench}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{*}\PY{l+m+mi}{1000}\PY{p}{,}\PY{l+m+mf}{1e4}\PY{p}{)}
-\PY{n}{bench2}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{*}\PY{l+m+mi}{1000}\PY{p}{,}\PY{l+m+mf}{1e4}\PY{p}{)}
-\end{Verbatim}
-\end{tcolorbox}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-
-Keys: 1.00 b  Bytes: 10.00 Kb
-
-RAM        314.94 Kb 19.23 Kb 16.38003848754356
-DISK json  23.89 Kb 18.90 Kb 1.2643276710588982
-DISK pkl   23.91 Kb 18.92 Kb 1.263903573694227
-DISK gz   10.34 Kb 10.14 Kb 1.020122312093115
-
-Keys: 1.00 b  Bytes: 10.00 Kb
-
-Creation time 0.002998828887939453 4.76837158203125e-07
-Acess time 7.152557373046875e-07 2.384185791015625e-07
-    \end{Verbatim}
+\prompt{Out}{outcolor}{19}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+          0         1         2         3         4         5         6  \textbackslash{}
+9  0.777892  0.000535  0.239624  0.882285  0.951852  0.443976  0.966355
+8  0.932123  0.027919  0.050847  0.592801  0.867408  0.164765  0.748370
+3  0.675394  0.257285  0.523616  0.820933  0.657142  0.600372  0.220450
+4  0.887332  0.288416  0.690859  0.118682  0.694805  0.611857  0.046453
+5  0.800843  0.670505  0.759913  0.061861  0.054205  0.747787  0.822460
+1  0.037602  0.672393  0.133586  0.004869  0.203755  0.168597  0.928038
+2  0.702845  0.714748  0.298797  0.758865  0.256273  0.791469  0.854921
+0  0.182818  0.803162  0.356512  0.286891  0.948721  0.509564  0.684132
+7  0.475779  0.818054  0.397074  0.239534  0.630659  0.024413  0.467922
+6  0.539352  0.862582  0.722857  0.116203  0.177171  0.788025  0.039456
+
+          7         8         9
+9  0.035550  0.404241  0.547084
+8  0.857539  0.840110  0.636682
+3  0.132569  0.576291  0.461130
+4  0.279059  0.129637  0.241594
+5  0.924498  0.617560  0.317410
+1  0.621204  0.808938  0.369552
+2  0.004346  0.345329  0.451903
+0  0.576865  0.166287  0.171629
+7  0.421451  0.344078  0.591117
+6  0.796943  0.106906  0.065239
+\end{Verbatim}
+\end{tcolorbox}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{20}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{n}{f}\PY{o}{=} \PY{n}{df}\PY{o}{.}\PY{n}{to\PYZus{}json}\PY{p}{(}\PY{p}{)}
+\PY{n}{df2}\PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{read\PYZus{}json}\PY{p}{(}\PY{n}{f}\PY{p}{)}
+\PY{n}{df2}
+\end{Verbatim}
+\end{tcolorbox}
+
+            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
+\prompt{Out}{outcolor}{20}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+          0         1         2         3         4         5         6  \textbackslash{}
+9  0.777892  0.000535  0.239624  0.882285  0.951852  0.443976  0.966355
+8  0.932123  0.027919  0.050847  0.592801  0.867408  0.164765  0.748370
+3  0.675394  0.257285  0.523616  0.820933  0.657142  0.600372  0.220450
+4  0.887332  0.288416  0.690859  0.118682  0.694805  0.611857  0.046453
+5  0.800843  0.670505  0.759913  0.061861  0.054205  0.747787  0.822460
+1  0.037602  0.672393  0.133586  0.004869  0.203755  0.168597  0.928038
+2  0.702845  0.714748  0.298797  0.758865  0.256273  0.791469  0.854921
+0  0.182818  0.803162  0.356512  0.286891  0.948721  0.509564  0.684132
+7  0.475779  0.818054  0.397074  0.239534  0.630659  0.024413  0.467922
+6  0.539352  0.862582  0.722857  0.116203  0.177171  0.788025  0.039456
+
+          7         8         9
+9  0.035550  0.404241  0.547084
+8  0.857539  0.840110  0.636682
+3  0.132569  0.576291  0.461130
+4  0.279059  0.129637  0.241594
+5  0.924498  0.617560  0.317410
+1  0.621204  0.808938  0.369552
+2  0.004346  0.345329  0.451903
+0  0.576865  0.166287  0.171629
+7  0.421451  0.344078  0.591117
+6  0.796943  0.106906  0.065239
+\end{Verbatim}
+\end{tcolorbox}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{28}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{n}{x}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{def}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{AAA}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{fff}\PY{l+s+s2}{\PYZdq{}}\PY{p}{:}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{CCC}\PY{l+s+s2}{\PYZdq{}}\PY{p}{\PYZcb{}}
+\PY{n}{y}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{abc\PYZus{}def}\PY{l+s+s2}{\PYZdq{}}
+
+\PY{n}{title}\PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}
+\PY{k}{for} \PY{n}{k} \PY{o+ow}{in} \PY{n}{x}\PY{p}{:}
+    \PY{k}{if} \PY{n}{k} \PY{o+ow}{in} \PY{n}{y}\PY{p}{:}
+        \PY{n}{title}\PY{o}{=} \PY{n}{x}\PY{p}{[}\PY{n}{k}\PY{p}{]}
+\PY{n}{title}
+\end{Verbatim}
+\end{tcolorbox}
+
+            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
+\prompt{Out}{outcolor}{28}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+'AAA'
+\end{Verbatim}
+\end{tcolorbox}
+
+    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
+\prompt{In}{incolor}{44}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+\PY{l+m+mf}{46.615389}\PY{o}{\PYZhy{}}\PY{l+m+mf}{59.829612}
+\end{Verbatim}
+\end{tcolorbox}
+
+            \begin{tcolorbox}[breakable, size=fbox, boxrule=.5pt, pad at break*=1mm, opacityfill=0]
+\prompt{Out}{outcolor}{44}{\boxspacing}
+\begin{Verbatim}[commandchars=\\\{\}]
+-13.214222999999997
+\end{Verbatim}
+\end{tcolorbox}
 

### 2021-03-23 thesis/pascalanalyzer/main.tex
add tex files

%\documentclass[sigconf, anonymous]{acmart}
\documentclass[sigconf]{acmart}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{booktabs} % For formal tables
\usepackage{footnote}
\usepackage[lofdepth,lotdepth]{subfig}
\usepackage[ruled]{algorithm2e} % For algorithms
\usepackage{listings, xcolor}
\usepackage{todonotes}\setlength{\marginparwidth}{3.5cm}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
%\acmDOI{10.475/123_4}
\acmDOI{}

% ISBN
%\acmISBN{123-4567-24-567/08/06}
\acmISBN{}

%Conference
\acmConference[CGO'18]{International Symposium on Code Generation and Optimization}{February 2018}{Vienna, Austria} 
\acmYear{2018}
%\copyrightyear{2016}

%\acmPrice{15.00}

\begin{document}
\title{SPERF a scalability profiler for shared-memory parallel applications}
%\titlenote{Produces the permission block, and
%  copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
%  \texttt{acmart.pdf} document}

\author{Vitor~R.~G.~Silva}
%\authornote{Dr.~Trovato insisted his name be first.}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{Univ. Federal do Rio Grande do Norte}
  \streetaddress{Departamento de Engenharia de Computação e Automação}
%  \city{Natal - RN} 
%  \state{RN} 
%  \postcode{59078-900}
}
\email{ramos.vitor89@gmail.com}

\author{Marcio~O.~Jales}
%\authornote{Dr.~Trovato insisted his name be first.}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{Univ. Federal do Rio Grande do Norte}
  \streetaddress{Departamento de Engenharia de Computação e Automação}
%  \city{Natal - RN} 
%  \state{RN} 
%  \postcode{59078-900}
}
\email{marcio.dojcosta@gmail.com}

\author{Alex~F.~A.~Furtunato}
\affiliation{%
  \institution{IFRN}
  \streetaddress{Diretoria Acad\^{e}mica de Gest\~{a}o e Tecnologia da Informa\c{c}\~{a}o}
%  \city{Natal - RN} 
%  \state{RN} 
%  \postcode{59078-900}
}
\email{alex.furtunato@ifrn.edu.br}

\author{Luis~F.Q~Silveira}
\affiliation{%
  \institution{Univ. Federal do Rio Grande do Norte}
  \streetaddress{Departamento de Engenharia de Computação e Automação}
%  \city{Natal - RN} 
%  \state{RN} 
%  \postcode{59078-900}
}
\email{lfelipeqs@gmail.com}

\author{Samuel~Xavier-de-Souza}
\affiliation{%
  \institution{Univ. Federal do Rio Grande do Norte}
  \streetaddress{Departamento de Engenharia de Computação e Automação}
%  \city{Natal - RN} 
%  \state{RN} 
%  \postcode{59078-900}
}
\email{samuel@dca.ufrn.br}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{Vitor R. G. Silva et al.}

\begin{abstract}
SPERF is presented as a tool for performing automating measurements of efficiency for scalability analysis of parallel applications in shared memory systems. With a simple configuration timings measurements are taken when varying the number of threads and input sizes for a target application. The measurements are taken at relevant points of the code, such as loops and parallel region delimiters. This allows a hierarchical scalability analysis to easily identify specific regions of code that, when optimized, are the best candidates to improve the overall scalability of the program. The tool is written in C++ and supports the analysis of C\/C++ programs that have been parallelized with OpenMP or PThreads. A statistical analysis of the tool shows that it has a minimal level of intrusion since the target application's performance measured using the tool is basically the same as if it was measured without using it.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10010940.10011003.10011002</concept_id>
<concept_desc>Software and its engineering~Software performance</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Software performance}

\keywords{Parallel Computing, Muticore Architecture, Parsec Benchmark}

\maketitle

\section{Introduction}
\todo[inline,caption={REVISÃO: teste de nota}]{REVISÃO: linha 1.
\newline linha 2 ....}
Existing profiling tools focus on analyzing hotspots or bottlenecks using statistical measurements from a single run. Naturally, many applications may present different bottlenecks/hotspots for different input sizes. In the multi-core and many-core era, bottlenecks may also change when varying the number of cores used in the computation. Moreover, eliminating performance bottleneck from a single representative run is not sufficient to ensure that the program will scale its performance along with newer generations of multicore/manycore processors because scalability concepts involve analyzing the normalized performance across runs.

The fundamentals metrics to measure the parallel execution are Speedup and Efficiency. The Speedup, $S$, is related with measures of execution time of task on parallel application and is defined how:

\begin{equation}
  S = \frac{T_{S}}{T_{P}}
  \label{eq:speedup}
\end{equation}
where, $T_{S}$ is the measure of task execution time on single core mode, and $T_{P}$ is the measure time of a task running on multi-core mode. The Efficiency, $E$, of a parallel application is defined by the ratio of Speedup, $S$, by  the number of cores, $N$, used to running that application, measured on partial or entire code. Thus, the Efficiency is:

\begin{equation}
  E = \frac{S}{N}
  \label{eq:efficiency}
\end{equation}

Collecting efficiency information is the first step on identifying bottlenecks on software scalability. In the traditional approach, collecting the speedup information demands the declaration of variables that will be used to time the application, use of appropriate functions to take the time information, handle the resulting information and calculation and execute the application over and over again, this process demand some work.

In this paper we present a tool for automate this process along with her configurations and features on section \ref{sec:sperf}. Also we validate our tool showing that the intrusiveness of this tool is minimal on section \ref{sec:validation}. To demonstrate analysis that can be done with Sperf we show on section \ref{sec:results} the results on some well known benchmarks applications. And, for the last, the Sections \ref{sec:future_work} and \ref{sec:conclusions} discuss about our conclusion and listed some wanted features that can be attended by future works.

\section{Scalability Measurement}
\label{sec:scalability_measurement}
Scalability on parallel system is the capability of handle growing on the problem size. There are three types that we can classify an program on:
\begin{itemize}
\item Scalable if by increasing the problem size with a fixed number of cores the efficiency not decreases.
\item Weak scalable the efficiency when the problem size increases with the same rate of the cores have to be constant or grow.
\item Strong scalable keeping the problem size fixed and increasing the number of cores the efficiency does not decreases. 
\end{itemize}
The scalability is an important measurement on a parallel system because the most commons ways of parallelization is by data division where each core gets a slice of the input data to process and on this systems the length of the date is very important.

\subsection{Profiling for scalability measure}
\label{subsec:profiling}

There are several ways of profiling the application, and they can be classified by how the collect data into two classes, event base or statistical.

\begin{itemize}
\item Event base works by interrupt the program each time a certain event occurs like enter o leaving a function, system calls and others.
\item Statistical profiling operating by sampling the call stack of the target program on regular intervals.
\end{itemize}

On the profile based on events that is triggered by system calls or function call can cause performance changes due to the high intrusiveness and can have some side effects such as on memory caches or instruction decoding pipeline being able to causing inaccurate results. Statistical in the other hand can be less intrusive and also have less side effects but is not numerical accurate.

On the HPC (High-performance computing) community numerous researches have proposed plenty of tools that aim to profile applications like \cite{Kufrin2005}, \cite{Servat2016}, \cite{Morris2010},
\cite{Graham1982}, \cite{Shende2006}, \cite{Malony2014}
\cite{Servat2013}, \cite{Servat2013}, \cite{Adhianto2010a} \cite{Chittimalli2012}, \cite{Tiku1971}, \cite{Knupfer2012}, \cite{Southern2016},
\cite{Woo1995}, \cite{Bienia2008}, \cite{Bienia2010} \cite{Adhianto2010}, \cite{Srivastava}, \cite{Tallent}
\cite{Architecture2015}, \cite{Tiku1971a}, \cite{Tallent}
\cite{Adhianto2010}, \cite{Srivastava}
frameworks Score-p \citep{Knupfer2012}, ATOM \cite{Srivastava}. Fine-grained profiles Scalable fine-grained call path tracing \cite{Tallent} 
Gprof \cite{Graham1982}, Call path profiler \cite{Adhianto2010} not sure Splash \cite{Woo1995}

But these tools focus on profiling the target application with a fixed input and/or a fixed number of threads, in this way is not possible to see the overall scalability of the program. SPERF was created on this purpose running the target application multiple times with different inputs and number of threads. This way the developer can look at the scalability of each region on multiple core and data length.

%The HPCToolkit solution, for example, consists of a large project that provides a call path tracing with low-overhead and asynchronous statistical sampling to collect fine-grained context-sensitive data across multiple threads. To support tasks as performance tuning, runtime scheduling and automatic parallelization, a data-dependence profiler has been developed. Finally, a profiler tool suited to runtime managed environments and even to profile heterogeneous parallel code are among the solutions to collect sensitive metrics to support the parallel performance analysis.

\section{SPERF}
\label{sec:sperf}

SPERF join the best of the ways of profiling two being less intrusive making possible to analyze the real time of the application with precision.Was developed in C programming language inside a Linux system environment and compiled with GNU\/g++ standard 98 with the focus on see the overall scalability and it is a event based profiling, triggered by specific functions to do this its performance a instrumentation on the source code that insert function on the begin and in the end of parallel region. And these functions are used to collect data.

The instrumentation is done by adding two marks on the region of interest. To do this there is two functions available on the SPERF package on the sperfops.h file, sperf\_start(n) and sperf\_stop(n). Those two functions annotates the region where the time will be acquired, sperf\_start delimits the beginning of the area, while sperf\_stop marks the end. These functions use the id passed by argument to get the time of each thread spent on the region marked.They may be used as many times as wished by the user, in order to time multiples regions. Therefore, the user should pass to these functions, as argument, an integer value that enumerates the correspondent parallel regions. In other words, if one wishes to time "n" areas of code, the first area should be delimited by sprof\_start(1) and sprof\_stop(1), the second by sprof\_start(2) and sprof\_stop(2) and so on.

This step of instrumentation can be done automatically on programs that use OPENMP \cite{Architecture2015} where the tool will seek for OPENMP directives and will mark the begin and end of the each parallel region, but manual regions can also be added.

After that the user can compile the program as was compiled before there is no special library needed and profile the application by invoke SPERF and pass, as command line arguments, the target application path (sperf app). Thus. SPERF will read a configuration file that tell the number of threads and all arguments that program will run, them execute the program from all possible combinations.To set the number of threads in the target application SPERF uses an environment variable OMP\_NUM\_THREADS on OPENMP application or pass by argument to the program where the user configure previously what argument in his application is the number of threads.Then on each call to sperf\_stop will be send information about the time spent on the parallel region using pipes method of interprocess communication on C\/Linux and SPERF will annotate the time of each region marked and output to the user a file with these informations.

\subsection{Options available}

The tool have the following options that is passed by command line:
\begin{itemize}
\item -i -p instrument use to performance the automatic instrumentation on the source code the next argument is the path to the source code where there is going to instrument all .cpp and .c files by default also is possible the user tell what extensions will be instrumented using -e option the next argument is the extensions separated by ','
\item -o output file set the name of the output file by default the name is the same of the application.
\item -c configuration set the name of the configuration file if no option is passed the program will look for the configuration file with the same name of the application. (use to configure testes that will be performed )
\item -t specify the type of output file, the can be JSON, XML and CSV.
\end{itemize}

\subsection{Execution configuration}
To configure how the profiler will run, SPERF have a configuration file. This file consist of 6 variables.

There is two ways of setting the threads that target will run, one is listing all desired values using the variable list\_threads\_values={1,2,3,6,8} and the other is by configuring automated steps that can be arithmetic progression (constant increase) and geometric progression.To do this there three variable type\_of\_step set the type of progression power or constant, value\_of\_step sets the increasing rate and max\_number\_threads set the maximum value of thread.

Furthermore the user can set a list of arguments with the variable list\_of\_args={arg11 arg12 ..., arg21 arg22 ..., ...}

\begin{itemize}
\item number\_of\_tests - number of times that SPERF will run.
\item list\_threads\_values - set specific values from number of threads.
\item type\_of\_step (power or constant) - set the type of increasing can be either constant rate of power.
\item value\_of\_step - set the value of the step.
\item max\_number\_threads - set maximum value of thread .
\item list\_of\_args - list of the argument separated by ",".
\end{itemize}

How many times the "test" will run, the initial and final number of threads that the target will use, how will be the increment from the initial value and the final value and how many tests will be performed. Regarding to the increment, there are two manners: arithmetic progression (constant increase) and geometric progression. That means, if an initial and final value of 1 and 16 are set, respectively, with constant increment 2, the target application will execute using 1, 3, 5, 7 and so on until 16 threads (15 will be the last in this case). Besides, the user informs how many times the target will be tested. So, considering the previous case, providing a value of 50, the cycle of 1 to 15 threads with constant increment of 2 will repeat 50 times.

\subsection{Data visualization}
The default output from SPERF is a JSON (JavaScript Object Notation) file from each parallel region marked. This archive contains information about arguments, threads, the begin and end line of the region on the code, speedup, efficiency and time from execution. But it can also produce files in XML(eXtensible Markup Language) and CVS (Comma-separated values).

The format on the JSON has the tags, "Region" informing the initial and final line of the parallel region, "Filename" with the name of the source file that the regions was included, "Execution" where has information about all input data used and inside each the tag input data with information of watch arguments was used and all the runs with different number of threads.

%TODO : Change the tags for English. Why XML? Why not JSON?
\lstset{
    string=[s]{"}{"},
    stringstyle=\color{red},
    comment=[l]{:},
    commentstyle=\color{black},
}

\begin{lstlisting}
{
"Region":"initial line, final line",
"Filename":"program.cpp",
"Execution 1":{
	"InputData 1":{
		"Arguments":"arg1 arg2",
		"Run 1":{
			"Number of threads":"1",
			"Time":"t",
			"Speedup":"s",
			"Efficiency":"e"},
		...
		"Run m":{
			...
		}
	},
	...
	"InputData n":{
		...
	}
},
...
"Execution e":{
	...
}
}
\end{lstlisting}

The CSV output file is a simple table that show the efficiency for the combination of each argument with all threads.The sample CSV file.

% Table
\begin{table}[H]%
\caption{Sample CSV}
\label{tab:sampleCSV}
\begin{minipage}{\columnwidth}
\begin{center}
\begin{tabular}{lllll}
     & num\_threads & num\_threads & num\_threads & num\_threads \\
arg1 & efficiency & efficiency & efficiency & efficiency \\
arg2 & efficiency & efficiency & efficiency & efficiency \\
arg3 & efficiency & efficiency & efficiency & efficiency \\
arg4 & efficiency & efficiency & efficiency & efficiency \\
\end{tabular}
\end{center}
\bigskip\centering
\footnotesize\emph{Source:} Author
\emph{Note:} Efficiency for the entire application.
\end{minipage}
\end{table}%

\section{Validation}
\label{sec:validation}

The validation of our tool is the next step of the development process, as fundamental as the tool coding process itself. The tests were performed using two target applications, called, Blackscholes and Freqmine, executed on a server structured with 2 CPUs Intel Xeon Sixteen-Core E5-2698v3. The programs are part of the PARSEC Benchmark Suite \cite{Southern2016} e \cite{Bienia2010} (Princeton Repository for Shared- Memory Computers), which consists of a package with a variety of multi-thread programs, whether OPENMP or Pthreads. Its use is essential, once it represents real computational challenges and, thus, renowned benchmarks. More information about PARSEC may be found in [parsec info].

Despite the promising results at first sight, we have to find a manner to measure how intrusive is the tool, that means, examine how the action of SPERF disturbs the target application to find out, in the last instance, if it happens as to become impracticable.

The tests performed to show the intrusiveness is minimal was to proof that there is no significant distance between the times obtained with SPERF and without SPERF using the one way analysis of variance ANOVA F test \cite{Tiku1971}.This technique can be used to compare means of two or more samples (using the F distribution), testing the null hypothesis that samples in all groups are drawn from populations with the same mean values.

\subsection{BlackSchoels test}
The Blackscholes application is a partial differential equation that, in the way implemented on PARSEC, calculates prices for a portfolio of European options analytically. In our test to this paper, the application was compiled to execute using Pthreads. All the annotations were made on the Blackscholes.c source code where was manually instrumented to perform the tests adding the sperf\_start and sperf\_stop functions available on the sperof.h file on the region of interest. The results shown in the figure \ref{fig:black100b} was perform using the following configuration on SPERF and invoking $./sperf blackscholes$ and the figure \ref{fig:black2100b} using the Linux time program time [13].
\\
\\
$number\_of\_tests=96$\\
$max\_number\_threads=32$\\
$type\_of\_step=power$\\
$value\_of\_step=2$\\
$list\_of\_args=\{in\_312K.txt \quad outputFile.txt\}$

\begin{figure}[H]
	\includegraphics[width=7cm,height=6cm]{figures/black100b}
	\caption{With SPERF}
	\label{fig:black100b}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=7cm,height=6cm]{figures/black2100b}
  \caption{With Linux Time}
  \label{fig:black2100b}
\end{figure}

\subsection{Freqmine test}
Freqmine is an application that employs an array-based version of the FP-growth (Frequent Pattern-growth) method for Frequent Itemset Mining.This program provide multiples alternatives of parallelization and for this test the application was compiled with to execute with OpenMP what makes it possible to use the automatic instrumentation on this application. With the command, ./sperf -i -p /pathToFreqmineFolder, the code was marked on 7 different regions, after that was compiled normally and invoking $./sperf freqmine$ with the configuration on SPERF was obtained the results on the figure \ref{fig:freq100b} and the figure \ref{fig:freq2100b} with Linux time.
\\
\\
$number\_of\_tests=96$\\
$max\_number\_threads=32$\\
$type\_of\_step=power$\\
$value\_of\_step=2$\\
$list\_of\_args=\{kosarak\_990k.dat \quad 1200 \quad output.txt\}$

\begin{figure}[H]
	\includegraphics[width=7cm,height=6cm]{figures/freq100b}
	\caption{With SPERF}
	\label{fig:freq100b}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=7cm,height=6cm]{figures/freq2100b}
  \caption{With Linux Time}
  \label{fig:freq2100b}
\end{figure}

\subsection{Intrusiveness measurement}

The statistical result obtained from the test above from Blackscholes and Freqmine was shown in the tables \ref{tab:blackstatistics} and \ref{tab:freqstatistics}.Where we can compare the result mean and standard deviation from two groups one with SPERF and the other without.

\begin{table}[H]
\centering
\caption{Statistics for BlackSchoels}
\label{tab:blackstatistics}
\begin{tabular}{llll}
                & With Sperf & Without Sperf & Total    \\
N               & 96         & 96            & 192      \\
$\sum{X}$       & 100,19     & 103,0134      & 203,2034 \\
Mean            & 1,0436     & 1,0731        & 1,0584   \\
$\sum{X^2}$     & 105,5093   & 112,4857      & 217995   \\
Std.Dev.        & 0,0998     & 0,1431        & 124     
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Statistics for Freqmine}
\label{tab:freqstatistics}
\begin{tabular}{llll}
            & With Sperf & Without Sperf & Total    \\
N           & 96         & 96            & 192      \\
$\sum{X}$   & 202,8016   & 205,44        & 408,2416 \\
Mean        & 2,1125     & 2,14          & 2,1263   \\
$\sum{X^2}$ & 429,9198   & 440,428       & 870,3478 \\
Std.Dev.    & 0,1256     & 0,091         & 0,1102  
\end{tabular}
\end{table}

Given the summary statistics, the calculations of the hypothesis test are shown in tabular form in the tables \ref{tab:blackAnova} and \ref{tab:freqsAnova}.

\begin{table}[H]
\centering
\caption{Anova f test}
\label{tab:blackAnova}
\begin{tabular}{lllll}
Source             & SS     & df  & MS     &             \\
Between-treatments & 0,0415 & 1   & 0,0415 & F = 2,72671 \\
Within-treatments  & 2893   & 190 & 0,0152 &             \\
Total              & 2,9345 & 191 &        &            
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Anova f test}
\label{tab:freqsAnova}
\begin{tabular}{lllll}
Source             & SS     & df  & MS     &             \\
Between-treatments & 0,0363 & 1   & 0,0363 & F = 3.01548 \\
Within-treatments  & 2,2844 & 190 & 0,012  &             \\
Total              & 2,3206 & 191 &        &            
\end{tabular}
\end{table}

If we take a significance level of 5\% the critical value of F with 1 and 190 degrees of freedom on numerator and denominator respectively is 3.8909. Size booth values of F calculated is lower then the critical there is no significant statistical difference on the mean value according to ANOVA.

%\begin{figure}[H]
%	\includegraphics[width=7cm,height=4.5cm]{ftest}
%	\caption{F distribution}
%	\label{fig:Fdist}
%\end{figure}

\section{Results}
\label{sec:results}

To show how we can interpret the result from SPERF. Was made some analyzes on Freqmine and Blackschoes. The test performed this time was from 1 to 32 threads in a  geometric progression of ratio 2 also varying the data size. The input data used was provide by parsec as a native input and was appropriately divide into several parts such the data size was increase in the same ratio as the number of threads so we can see the scalability.

\subsection{Blackschoes}
The test was performed 4 time and the configuration to this test on Blackscholes was:
\\
\\
$number\_of\_tests=4$\\
$list\_threads\_values=\{\}$\\
$max\_number\_threads=32$\\
$type\_of\_step=power$\\
$value\_of\_step=2$\\
$list\_of\_args=\{$\\
$\texttt{number\_of\_threads} \quad in\_312K.txt \quad out\_312K.txt,$\\
$\texttt{number\_of\_threads} \quad in\_625K.txt \quad out\_625K.txt,$\\
$\texttt{number\_of\_threads} \quad in\_1M.txt \quad out\_1M.txt,$\\
$\texttt{number\_of\_threads} \quad in\_2M.txt \quad out\_2M.txt,$\\
$\texttt{number\_of\_threads} \quad in\_5M.txt \quad out\_5M.txt,$\\
$\texttt{number\_of\_threads} \quad in\_10M.txt \quad out\_10M.txt,$\\
$\}$

There was generated 2 files from Blacksholes one from the entire and other to the region marked size it only have one parallel region. Inside this files was the tables with the efficiency from each configuration. To the a more reliable result was get the median value of the executions and the result was plotted on a 3D graph to a better visualization as shown in the figure \ref{fig:blackRegions}.

\begin{figure}[H]
     \centering
     \subfloat[][Entire application]{\includegraphics[width=6cm,height=5cm]{blackFig1}\label{fig:blackFig1}}

     \subfloat[][Region 1]{\includegraphics[width=6cm,height=5cm]{blackFig2}\label{fig:blackFig2}}

     \caption{Regions of blackschoels}
     \label{fig:blackRegions}
\end{figure}

From this results we can see that the region that was parallelized has some part there is strong scalable on the argument 2 from 2 cores to 8.
And some parts the efficiency decay and start increases again. This observation was not possible with only the analyses of the entire application. Also we can see in general that is scalable because looking only the direction of the problem size the efficiency was roughly constant.

\subsection{Freqmine}
In the freqmine test size there two parameters that control the problem size was found a combination that represent the data size increases to be roughly the same ratio as the number of thread. The configuration used was:
\\
\\
$number\_of\_tests=4$ \\
$list\_threads\_values=\{\}$ \\
$max\_number\_threads=32$ \\
$type\_of\_step=power$ \\
$value\_of\_step=2$ \\
$list\_of\_args=\{$ \\
$kosarak\_30k.dat \quad 50 \quad kosarak\_30kout.dat,$\\
$kosarak\_61k.dat \quad 60 \quad kosarak\_61kout.dat,$\\
$kosarak\_123k.dat \quad 120 \quad kosarak\_123kout.dat,$\\
$kosarak\_247k.dat \quad 240 \quad kosarak\_247kout.dat,$\\
$kosarak\_495k.dat \quad 540 \quad kosarak\_495kout.dat,$\\
$kosarak\_990k.dat \quad 1400 \quad kosarak\_990kout.dat,$\\
$\}$

This test created eight files.Seven from region marked and one from the entire application. Also plotting the 3D graphs where we can see on figure \ref{fig:freqRegions}.

\begin{figure}[H]
     \centering
     \subfloat[][Entire application]{\includegraphics[width=4cm,height=3.5cm]{freqFig1}\label{fig:freqFig1}}
     \subfloat[][Region 1]{\includegraphics[width=4cm,height=3.5cm]{freqFig2}\label{fig:freqFig2}}

     \subfloat[][Region 2]{\includegraphics[width=4cm,height=3.5cm]{freqFig3}\label{fig:freqFig3}}
     \subfloat[][Region 3]{\includegraphics[width=4cm,height=3.5cm]{freqFig4}\label{fig:freqFig4}}
\end{figure}

\begin{figure}[H]
     \centering
     \ContinuedFloat
     \subfloat[][Region 4]{\includegraphics[width=4cm,height=3.5cm]{freqFig5}\label{fig:freqFig5}}
     \subfloat[][Region 5]{\includegraphics[width=4cm,height=3.5cm]{freqFig6}\label{fig:freqFig6}}

     \subfloat[][Region 6]{\includegraphics[width=4cm,height=3.5cm]{freqFig7}\label{fig:freqFig7}}
     \subfloat[][Region 7]{\includegraphics[width=4cm,height=3.5cm]{freqFig8}\label{fig:freqFig8}}

     \caption{Regions of freqmine}
     \label{fig:freqRegions}
\end{figure}

On this application we can see that the regions have complete different profiles. The regions 4 and 7 the efficiency decay drastically with increasing in the number of threads and there is no weak or strong parts. The regions 2, 3 and 5 responds better to the parallelization and we can see that by increase the problem size the efficiency grows significantly and there is a appearance of strong and weak parts. The regions 1 and 6 have some interesting results where we can see efficiency greater than 1 witch implies on super linear speedup. And the overall application have some decays on the efficiency but grew back and in general is scalable.

One of the roles of SPERF gets clearer when examining these figures. Such efficiency profiling may give to the user a deeper idea of how the target application behaves and improve the parallelization on the crucial parts.All of that work to achieve that results on the tables could be done manually, but with a higher cost to the user. The library functions make the code instrumentation simpler and the automatization factor presents and handle the output faster.

\section{Future Work}
\label{sec:future_work}

The tool is still in its first versions, so there are some features that are not supported yet, but, future works should improve these faults. Follow some new features suggestions:
\begin{itemize}
\item Support operating systems other than Linux. 
\item Automatic instrument code using Pthread due to the high granularity presents on this model.
\item Supporting multiples ways of passing program inputs
\item Instrumentation support through the compiler.
\item Support to distributed system MPI(Message Passing Interface).
\item Graphical Interface of better controls and visualizations features.
\end{itemize}

\section{Conclusions}
\label{sec:conclusions}

Scalability of the application is one of the most important ways to measure the limits of a parallel application, and we face a lack of tools that automatize this process. In this sense, this paper proposes SPERF as an alternative to accomplish this task. Considering the results obtained, we suggest that using our profiler may be very convenient.

SPERF makes really simple to analyze the scalability of the program with only a few options to configure and without making extensive modifications to source code it's possible to run the program with different data sizes and cores, and show the user in a presentable way the efficiency of each parallel region on the code.Furthermore, the level of intrusion measured shows that the influence of the instrumentation inside the user's code is minimal, thus confirming the convenience of the profiler proposed.

The future goals of the tool it is to support distributed systems (MPI) and developing a graphical user interface to a better interaction and data presentation.

% Bibliography

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}

### 2021-03-23 thesis/pascalanalyzer/pascalanalyzer.tex
add tex files

\section{Pascal Analyzer}

Run application collecting information while executing a program with
different configurations of cores, frequencies, inputs arguments.

\subsection{Architecture}
\begin{figure}
	\includegraphics{pascalanalyzer/arch2}
\end{figure}
\subsection{Modules description}
PascalRun: Provie a monitor class to collect information while executing  a program 
with different configurations of cores, frequencies, inputs arguments. The information collected can be via sensor specified by the user,  instrumented regions (source code and openmp binaries) and the entire  program execution time. The data is stored and processed with PascalData  module that can resume all on a DataArray (xArray module) with times, speedups or efficiency.

PascalData: Load saved data from pascalrun and can convert and process to varius formats.

PascalModel: Create AI and Analytical models from pascal data

Instrumenter: Instrument source code creating marks at parallel regions

libinst: Instrument binary code creating marks at parallel regions

cpufreq: resposible of CPU control

IPMI, RAPL, RAPL\_sample, ProcList: Sensors

\subsection{Features}

\subsubsection{Options}
\begin{outline}
	\1 list of cores
	\1 list of frequencies
	\1 list of inputs
	\1 list of sensors
		\2 ipmi
		\2 rapl (sysfs)
		\2 rapl (perf)
		\2 process cores monitor
	\1 number of repetitions
	\1 choice of CPU governor
	\1 idle time between task
	\1 enable/disable hyperthread
	\1 enable/disable cores
	\1 collect times from marks manually inserted by the user in the source code
	\1 collect times from parallel regions using binary instrumentation (openmp)
	\1 list of sensors
	\1 instrument source code
	\1 output file name
	\1 verbosity level
		\2 0 nothing
		\2 1 only display run configuration
		\2 2 display run configuration and sensors data
		\2 3 display everything
\end{outline}

\subsubsection{Execution}

Examples execution using command line
\begin{lstlisting}
pascalanalyzer ./myapp --bin --idle_time 5 --cores 1,4  
		--frequencies 3000000,2800000 --verbosity 3

pascalanalyzer ./myapp -c 1,32 -v 2 -o myoutput.json
\end{lstlisting}

Example execution python
\begin{lstlisting}
from pascalanalyzer.pascalrun import PascalRun
from pascalanalyzer.sensors.rapl import RAPL
rapl_sensor = RAPL()

app = PascalRun(
	application="myapp",
	repetitions=5,
	sensors=[rapl_sensor],
	idle_time=30,
	governor="userspace",
	disable_cores=True,
	disable_hyperthread=True,
)
app.run(
	cores=[1,16,32],
	frequencies=[3000000,2800000],
	inputs=[["arg1,"arg2"], ["arg3,"arg4"]]
)
app.save("output_file")
\end{lstlisting}

Example model creating python
\begin{lstlisting}
from pascalanalyzer import PascalData
from pascalanalyzer.pascalmodel import *

data = PascalData(filename)
df = data.energy()

opt = LeastSquaresOptmizer("model= lambda x, N, p, f:"
                           "x[0]*f**2*p+x[1]/(f*p)*N**x[7]+x[2]*f**2"
                           "+x[3]*p+x[4]/f*N**x[6]+x[5]", 8)

pascalmodel= create_model(df,inputs=["input","cores","frequency"],
                             output=["ipmi_energy"], model= opt,
                             config=data.config, train_sz=0.9)
\end{lstlisting}

\subsection{Data structure}

\subsubsection{Proposals}
\subsubsection{neasted dictonaries}
\begin{lstlisting}
{
	key1:
	{
		key2:
		{
			key3:
			{
				data1: val1,
				data2: val2
			},
			...
		},
		...
	},
	...
}
...
\end{lstlisting}

\subsubsection{index dictonary}
\begin{lstlisting}
{
"key1;key2;key3;...": [
	{
		data1: val1,
		data2: val2
	},
	...
],
}
...
\end{lstlisting}

\subsubsection{Benchmarking}

{\centering

Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 10.00 b

\begin{tabular}{llllr}
	\toprule
	{} & Storage &    Neasted &      Index &  Neasted/Index \\
	\midrule
	0 &     RAM &    2.57 Mb &    4.73 Mb &       0.542656 \\
	1 &   JSON  &  616.52 Kb &    1.26 Mb &       0.490468 \\
	2 &     PKL &  380.99 Kb &    1.21 Mb &       0.315070 \\
	3 &      GZ &  221.96 Kb &  270.36 Kb &       0.820989 \\
	\bottomrule
\end{tabular}

Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 100.00 b

\begin{tabular}{llllr}
	\toprule
	{} & Storage &  Neasted &    Index &  Neasted/Index \\
	\midrule
	0 &     RAM &  4.73 Mb &  6.89 Mb &       0.686078 \\
	1 &   JSON  &  2.78 Mb &  3.42 Mb &       0.812560 \\
	2 &     PKL &  2.54 Mb &  3.37 Mb &       0.754198 \\
	3 &      GZ &  1.82 Mb &  1.87 Mb &       0.968731 \\
	\bottomrule
\end{tabular}

Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 1.00 Kb

\begin{tabular}{llllr}
	\toprule
	{} & Storage &   Neasted &     Index &  Neasted/Index \\
	\midrule
	0 &     RAM &  26.33 Mb &  28.49 Mb &       0.924100 \\
	1 &   JSON  &  24.38 Mb &  25.02 Mb &       0.974398 \\
	2 &     PKL &  24.22 Mb &  25.04 Mb &       0.966930 \\
	3 &      GZ &  17.58 Mb &  17.64 Mb &       0.996294 \\
	\bottomrule
\end{tabular}

Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 10.00 Kb

\begin{tabular}{llllr}
	\toprule
	{} & Storage &    Neasted &      Index &  Neasted/Index \\
	\midrule
	0 &     RAM &  242.33 Mb &  244.49 Mb &       0.991156 \\
	1 &   JSON  &  240.38 Mb &  241.02 Mb &       0.997343 \\
	2 &     PKL &  240.24 Mb &  241.07 Mb &       0.996565 \\
	3 &      GZ &  175.11 Mb &  175.07 Mb &       1.000198 \\
	\bottomrule
\end{tabular}

Keys: [32,10,10,10,10] 320.00 Kb  Data bytes: 1.00 b

\begin{tabular}{llllr}
	\toprule
	{} & Storage &    Neasted &     Index &  Neasted/Index \\
	\midrule
	0 &     RAM &   13.38 Mb &  37.27 Mb &       0.358895 \\
	1 &   JSON  &    5.23 Mb &  13.98 Mb &       0.373866 \\
	2 &     PKL &    1.49 Mb &  12.70 Mb &       0.117686 \\
	3 &      GZ &  537.08 Kb &   1.24 Mb &       0.431880 \\
	\bottomrule
\end{tabular}

Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 10.00 b

\begin{tabular}{llrr}
	\toprule
	{} &      Time &   Neasted &  Index \\
	\midrule
	0 &  Creation &  0.073802 &    0.0 \\
	1 &     Acess &  0.000000 &    0.0 \\
	\bottomrule
\end{tabular}

Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 100.00 b

\begin{tabular}{llrr}
	\toprule
	{} &      Time &   Neasted &  Index \\
	\midrule
	0 &  Creation &  0.426866 &    0.0 \\
	1 &     Acess &  0.000000 &    0.0 \\
	\bottomrule
\end{tabular}

Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 1.00 Kb

\begin{tabular}{llrr}
	\toprule
	{} &      Time &   Neasted &  Index \\
	\midrule
	0 &  Creation &  4.162496 &    0.0 \\
	1 &     Acess &  0.000000 &    0.0 \\
	\bottomrule
\end{tabular}

Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 10.00 Kb

\begin{tabular}{llrr}
	\toprule
	{} &      Time &    Neasted &  Index \\
	\midrule
	0 &  Creation &  42.476255 &    0.0 \\
	1 &     Acess &   0.000000 &    0.0 \\
	\bottomrule
\end{tabular}

Keys: [1]*1000 1.00 Kb  Bytes: 10.00 Kb

\begin{tabular}{llllr}
	\toprule
	{} & Storage &    Neasted &     Index &  Neasted/Index \\
	\midrule
	0 &     RAM &  314.94 Kb &  19.23 Kb &      16.380038 \\
	1 &   JSON  &   23.89 Kb &  18.90 Kb &       1.264328 \\
	2 &     PKL &   23.91 Kb &  18.92 Kb &       1.263904 \\
	3 &      GZ &   10.34 Kb &  10.14 Kb &       1.019525 \\
	\bottomrule
\end{tabular}

Keys: [1]*1000 1.00 Kb  Data bytes: 10.00 Kb

\begin{tabular}{llrr}
	\toprule
	{} &      Time &   Neasted &  Index \\
	\midrule
	0 &  Creation &  0.003023 &    0.0 \\
	1 &     Acess &  0.000000 &    0.0 \\
	\bottomrule
\end{tabular}

}

From the tests, we conclude that index dictionary is a better option for speed and nested dictionary for space, although if the data size is significantly higher than the number of keys it doesn't save a lot of space. Because of that, we choose to work with index.

\subsubsection{Structure}
The data structure holds

\begin{lstlisting}
{
    "config": {
        "pkg": "./openmp_simple_bin",
        "execdate": "29-03-2020_13:07:07",
        "command": "run_test.py TestMonitor.test_valid_combinations",
        "hostname": "MSI",
        "kernel": "Linux-4.4.0-18362-Microsoft-x86_64-with-Ubuntu-18.04-bionic",
        "disable_hyperthread": true,
        "disable_cores": true,
        "idle_time": 0,
        "data_descriptor": {
            "values": [ ... ],
            "extras": {
                "regions": {
                "values": [ ... ]
            }
         },
        "keys": [ ... ]
    },
    "arguments": [ ... ],
    "governor": "userspace"
    },
    "data": {
        "4;1.2e-06;1;1": {
            "regions": {
                "0x1": [
                        [ ... ],
                        ...
                    ],
                ...
            },
            ...,
            "total_time": 0.3075871467590332
        },
        ...
    }
}
\end{lstlisting}

\subsection{Overhead}
\subsubsection{BlackSchoels test}

$number\_of\_tests=96$\\
$max\_number\_threads=32$\\
$type\_of\_step=power$\\
$value\_of\_step=2$\\
$list\_of\_args=\{in\_312K.txt \quad outputFile.txt\}$

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{pascalanalyzer/figures/black100b}
	\caption{With SPERF}
	\label{fig:black100b}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{pascalanalyzer/figures/black2100b}
	\caption{With Linux Time}
	\label{fig:black2100b}
	\end{subfigure}
\end{figure}

\subsubsection{Freqmine test}

$number\_of\_tests=96$\\
$max\_number\_threads=32$\\
$type\_of\_step=power$\\
$value\_of\_step=2$\\
$list\_of\_args=\{kosarak\_990k.dat \quad 1200 \quad output.txt\}$

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/freq100b}
		\caption{With SPERF}
		\label{fig:freq100b}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{pascalanalyzer/figures/freq2100b}
		\caption{With Linux Time}
		\label{fig:freq2100b}
	\end{subfigure}
\end{figure}

\subsubsection{Intrusiveness measurement}

The statistical result obtained from the test above from Blackscholes and Freqmine was shown in the tables \ref{tab:blackstatistics} and \ref{tab:freqstatistics}.Where we can compare the result mean and standard deviation from two groups one with SPERF and the other without.

\begin{table}[H]
	\centering
	\caption{Statistics for BlackSchoels}
	\label{tab:blackstatistics}
	\begin{tabular}{llll}
		& With Sperf & Without Sperf & Total    \\
		N               & 96         & 96            & 192      \\
		$\sum{X}$       & 100,19     & 103,0134      & 203,2034 \\
		Mean            & 1,0436     & 1,0731        & 1,0584   \\
		$\sum{X^2}$     & 105,5093   & 112,4857      & 217995   \\
		Std.Dev.        & 0,0998     & 0,1431        & 124     
	\end{tabular}
\end{table}

\begin{table}[H]
	\centering
	\caption{Statistics for Freqmine}
	\label{tab:freqstatistics}
	\begin{tabular}{llll}
		& With Sperf & Without Sperf & Total    \\
		N           & 96         & 96            & 192      \\
		$\sum{X}$   & 202,8016   & 205,44        & 408,2416 \\
		Mean        & 2,1125     & 2,14          & 2,1263   \\
		$\sum{X^2}$ & 429,9198   & 440,428       & 870,3478 \\
		Std.Dev.    & 0,1256     & 0,091         & 0,1102  
	\end{tabular}
\end{table}

Given the summary statistics, the calculations of the hypothesis test are shown in tabular form in the tables \ref{tab:blackAnova} and \ref{tab:freqsAnova}.

\begin{table}[H]
	\centering
	\caption{Anova f test}
	\label{tab:blackAnova}
	\begin{tabular}{lllll}
		Source             & SS     & df  & MS     &             \\
		Between-treatments & 0,0415 & 1   & 0,0415 & F = 2,72671 \\
		Within-treatments  & 2893   & 190 & 0,0152 &             \\
		Total              & 2,9345 & 191 &        &            
	\end{tabular}
\end{table}

\begin{table}[H]
	\centering
	\caption{Anova f test}
	\label{tab:freqsAnova}
	\begin{tabular}{lllll}
		Source             & SS     & df  & MS     &             \\
		Between-treatments & 0,0363 & 1   & 0,0363 & F = 3.01548 \\
		Within-treatments  & 2,2844 & 190 & 0,012  &             \\
		Total              & 2,3206 & 191 &        &            
	\end{tabular}
\end{table}

If we take a significance level of 5\% the critical value of F with 1 and 190 degrees of freedom on numerator and denominator respectively is 3.8909. Size booth values of F calculated is lower then the critical there is no significant statistical difference on the mean value according to ANOVA.

%\begin{figure}[H]
%	\includegraphics[width=7cm,height=4.5cm]{ftest}
%	\caption{F distribution}
%	\label{fig:Fdist}
%\end{figure}

\subsubsection{Sampling}

Energy estimation method for power samples
Time difference
\begin{equation}
	E= \sum_i^{N-1}(t_{i+1}-t_i)p_i+s_N(t_{tot}-t_N)
\end{equation}

Constant sample time
\begin{equation}
E= s_r\sum_i^{N}p_i=TP_{avg}
\end{equation}

Error due to sampling
\begin{itemize}
	\item Time difference $\approx$ 4\%
	\item Constant sample time 	$\approx$ 1\%
\end{itemize}

\subsection{Database}

\begin{outline}
	\1 Advantages
		\2 No need to organize the files
		\2 Faster and easier queries
		\2 Shared data across programs and research
	\1 Disadvantages
		\2 Need of db server
		\2 Need of network
		\2 Non local server maybe slow ( need to cache )
\end{outline}

### 2021-03-23 thesis/test/main.tex
add tex files

%=================================================================
\documentclass{ieeeaccess}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage[capitalise]{cleveref}
\usepackage{subcaption}
\usepackage{float}
%\usepackage{todonotes}

\usepackage{caption,setspace}
\captionsetup{font={sf,small,stretch=0.80},labelfont={bf,color=accessblue}}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}
\history{Date of publication xxxx 00, 0000, date of current version xxxx 00, 0000.}
\doi{10.1109/ACCESS.2017.DOI}

\title{
Analytical Energy Model Parametrized by Workload, Clock Frequency and Number of Active Cores for Share-Memory High-Performance Computing Applications
}

\author{
\uppercase{Vitor R. G. Silva}\authorrefmark{1},
\uppercase{Carlos Valderrama Sakuyama}\authorrefmark{1} \IEEEmembership{Senior Member, IEEE}, and \uppercase{Samuel Xavier-de-Souza}\authorrefmark{3} \IEEEmembership{Senior Member, IEEE}}
\address[1]{Department of Electronics and Microelectronics (SEMi), University of Mons, 7000 Mons, Belgium}
\address[2]{Department of Computer Engineering and Automation, Universidade Federal do Rio Grande do Norte, Natal, Brazil}
%\tfootnote{NPAD}

\markboth
{Author \headeretal: Preparation of Papers for IEEE TRANSACTIONS and JOURNALS}
{Author \headeretal: Preparation of Papers for IEEE TRANSACTIONS and JOURNALS}

\corresp{Vitor R. G. Silva: vitor.ramosgomesdasilva@umons.ac.be, Carlos Valderrama Sakuyama: carlos.valderrama@umons.ac.be, Samuel Xavier-de-Souza: samuel@dca.ufrn.br}

\begin{abstract}
Energy consumption is a key to enable exascale High-performance Computing (HPC). Energy-optimized hardware and software combinations still could be inefficient if software operates poorly. 
Software operation relies on dynamic scaling of frequency and voltage (DVFS) as well as dynamic power management (DPM), but none have privileged information on the application behavior, which may lead to energy-inefficient software operation. 
This work proposes an analytical model that can be used to estimate energy-optimal software configurations and provide knowledgeable hints to improve DVFS and DPM techniques for single-node HPC applications. 
Additionally, model parameters such as the level of parallelism and dynamic power provide insights into how the modeled application consumes energy, which can be useful for energy-efficient software development.
This novel analytical model takes the number of active cores, the operating frequencies, and the problem size as inputs to provide energy consumption estimation.
We present the modeling of 13 parallel applications employed to determine energy-optimal configurations for a number of different problem sizes. 
The results show that on average up to 70\% of energy could be saved and that if the right number of cores could be predicted, then only 14\% of the energy would be wasted.
We also compare the proposed model to standard machine-learning modeling with respect to training overhead and accuracy. The results show that our approach generates about 10 times less energy overhead for the same level of accuracy.
\end{abstract}

\begin{keywords}
Energy Model; Dynamic Frequency and Voltage Scaling; Dynamic Power Management; High Performance Computing
\end{keywords}

\titlepgskip=-15pt

\maketitle

\section{Introduction}
Data center's energy efficiency has attained crucial importance in recent years due to its high economic, environmental, and performance impact. 
Data center energy consumption was estimated to be between 1.1\% and 1.5\% of worldwide electricity usage in 2010 \cite{Dayarathna2016DataSurvey,Corcoran2017EmergingICT}, generating as much pollution as a nation like Argentina  \cite{Mathew2012Energy-awareNetworks}.
Power costs exceed in some cases the cost of purchasing hardware \cite{Rivoire2007ModelsOptimizations}. 
Nowadays, the leading petaflop supercomputers consume a range of 1–18 MW of electrical power, with 1.5 MW on average, which can be easily translated into millions of dollars per year in electricity bills \cite{Group2012HandbookSahni}.
Furthermore, the energy costs of powering a typical data center doubles every five years \cite{Buyya2013MasteringProgramming}.
Therefore, with such a steep increase in electricity use and rising electricity costs, electricity bills have become a significant expense for today's data centers \cite{Poess2008EnergyResults,Gao2013QualityCenters}. 
Due to these reasons, data center energy efficiency is now considered a chief concern for data center operators, often ahead of the traditional considerations of availability and security.

There are several approaches for green computing, from electrical materials to circuit design, systems integration, and software. These techniques may differ, but they share the same goal: substantially reduce overall system energy consumption without a corresponding negative impact on delivered performance. 
From a software systems perspective, the strategy of providing just enough power to get a job done can be accomplished by turning off system components when they are underutilized \cite{Mathew2012Energy-awareNetworks}.
The processor and main memory are usually the components that dominate power consumption, as shown in Figure  \ref{fig:powerbreakdown}.
The processor can consume as much as 50\% of the total energy~\cite{Fan2007, Barroso2007TheComputing, Malladi2012TowardsDRAM}. For that reason, modern processors incorporate several features for power management~\cite{Rotem2012Power-managementBridge, Brown2005, Hackenberg2015}, such as Dynamic Power Management (DPM) and Dynamic Voltage and Frequency Scaling (DVFS). 
DPM encompasses a set of techniques for obtaining energy-efficient computing by deactivating or reducing the system component's performance when they are idle or partially utilized~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.
DVFS allows the frequency and voltage to be adjusted in run-time depending on current needs.

\begin{figure}[htb!]
    \centering

    \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{power_breakdown/Xeon E5-2698 + DDR4 + SSD.png}
    \caption{This work.}
    \label{fig:1}
    \end{subfigure}
    %
    \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{power_breakdown/Xeon E5507 + DDR3 + HDD.png}
    \caption{Study case in \cite{Malladi2012TowardsDRAM}.}
    \label{fig:2}
    \end{subfigure}

    \hfill
    \caption{Power breakdown of a typical node of an HPC cluster at full use. The system used in this work (a) was built in 2016 and equipped with two Intel Xeon E5-2698, 128 GB of DDR4 memory and SSD as storage, while (b) the case study in \cite{Malladi2012TowardsDRAM} was built in 2012 and equipped with two Xeon E5507, 32GB of DDR3 memory and HDD as storage.}
    \label{fig:powerbreakdown}
\end{figure}

DVFS is motivated by the well-known fact that frequency and power have a near-cubic relationship \cite{Dayarathna2016DataSurvey, Group2012HandbookSahni}. 
This implies that running the CPU at lower frequency causes a linear reduction in the performance while a near-cubic reduction in power and, consequently, could lead to a near-square reduction in CPU energy. 
This can lead to dramatic energy savings depending on the system and its architecture.
Although very promising, the system software has yet to determine when and what voltage and frequency to use when running applications.
Otherwise, not only will performance deteriorate, but in the worst case, energy consumption would also increase  \cite{Group2012HandbookSahni}.
Indeed, reducing the frequency results in a longer execution time, which increases the energy consumption of other system components such as memory and disks.
There is also an overhead of time and energy associated with a voltage and frequency switch.
Thus, finding the most appropriate voltage and frequency to use  in all circumstances is not an easy task.
Therefore, since its introduction in 1994, there has been a tremendous amount of research on DVFS algorithms.

The DPM technique can achieve substantial energy savings on systems where the static power is high, or the system remains inactive for a long time.
In that case, the problem is to determine when and which components to turn on/off.
With DPM, energy savings of 70\% have been reported ~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.
%In the case of DPM, the problem is determining when and which components to turn on/off. This technique can achieve substantial energy savings on systems where the static power consumption is high or if the system stays idle for a long time. Power savings of 70\% have been reported~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.

At the same time as the stated techniques reduce the system's energy consumption, they also lead to a complex performance trade-off that must be exploited to produce energy-efficient algorithms.
%However, at the same time, that these power-saving techniques reduce system energy, they also lead to a complex balance between energy savings and high performance, exposing trade-offs between performance and power consumption, which can be exploited to produce more energy-efficient algorithms.  %in the case of the CPU, this can be applied to each core. Here... 
Indeed, this work investigates whether the construction of a model of the energy consumption of an application can lead to significant energy savings. We propose an analytical energy model of an application in function of the two control variables present in most HPC systems: CPU operating frequency and number of active cores. 
%%This model can serve as a base for new DVFS and DPM algorithms as well as for analyzing the contribution of individual model parameters to the total energy consumption.
This model, which characterizes an application, contributes to the objectives of DVFS and DPM techniques since it serves to the analysis of the contribution of each of the parameters to the total energy consumption.
Thus, the model includes two application-dependent parameters and three parameters relating to the system. The application parameters incorporate characteristics of the percentage of parallelism and the size of the input. The system parameters include power-related and technology-dependent components that compose dynamic, static, and leakage power.

The rest of this paper is organized as follows. In Sections \ref{sec:background} and \ref{sec:relatedwork}, a general review of existing models is provided, showing the difference between each approach and its applications. In Section \ref{sec:models}, the proposed model and its parameters are derived alongside its constraints. In Section \ref{sec:experimentalvalidation}, the model is validated with the PARSEC benchmark applications. Forward, in Section \ref{sec:dvfs_optmin}, use cases of the model are presented as well as how it could be applied in DVFS. Finally, the conclusion is made in Section \ref{sec:conclusion}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theoretical background} \label{sec:background}

A model is a formal representation of a real system. Computer system models can be represented in the form of equations, graphical models, rules, decision trees, representative collections of examples, or by neural networks, among others. The choice of representation affects the model's accuracy, as well as its interpretability by people~\cite{Hypothesis2012EncyclopediaLearning, Roy2019ForecastingNetwork, Zhu2019PredictingLearning}. Accurate energy and power consumption models are very important for many energy efficiency schemes employed in computing equipment \cite{Rivoire2007ModelsOptimizations}, and they can have multiple uses, including the design, forecasting, and optimization of data center systems. This work focuses on analytical models that could serve for energy optimization and subsequently, analysis of important factors in the total energy draw.

The desirable properties of a full-system model of energy consumption include accuracy (precision enough to allow the required energy saving), speed (fast enough to produce predictions), generality and portability (should be suitable for as many systems as possible), inexpensiveness (should not require costly or invasive infrastructure), and simplicity \cite{Rivoire2008AModels}. However, modeling the exact energy consumption behavior of an HPC system is not straightforward, either at the whole-system level or at the level of individual components. Patterns of energy consumption in data centers depend on multiple factors such as hardware specifications, workload, cooling requirements, or types of applications. Some of these factors cannot be measured easily. Furthermore, it is impractical to perform detailed measurements of the energy consumption of lower-level components without impacting the system with additional overload.

The concept of energy (E) is the total amount of work performed by a system over a period of time (T), while power (P) is the rate at which the system performs the work. The relation between these three amounts can be expressed as:

\begin{equation}
    E = \int_{0}^{T}P(t)dt
    \label{qe:energy_definition_cont}
\end{equation}

Several proposed models have already been classified with respect to its input parameters in \cite{Dayarathna2016DataSurvey}. In this survey, more than 200 models were analyzed according to their characteristics and limitations and classified into categories where the model is more suited to its objectives:
\begin{itemize}
    \item Temperature
    \item System Utilization or Workload
    \item Frequency
    \item Other system states such as cache miss, branch prediction, number of instructions executed, and more
    \label{tab:input_type}
\end{itemize}

Often, energy models are described as a combination of two main parts, the power model of the system and the performance model of the application.

\subsection{Power models}
Temperature-based models\cite{ZhangEnergyInfrastructure} can provide a way of estimating power consumption without being very intrusive. With this kind of model, it is possible to estimate energy with a single temperature sensor without the need to restart or introduce anything to the system. This is especially interesting for HPCs and data centers which, can not be easily turned off. Another use of this model is in the optimization of the refrigeration systems.

A large part of the developed models are based on parameters representing the state of the system. They normally take advantage of performance counters, which are provided by the CPU or the operating system and are capable of counting micro-architectural events, such as instructions executed, cache-hits, miss-predicted branches, and much more. 
This type of model is suitable for power estimation because it receives information about several internal states of the computer.

Frequency-based models serve as a base for many power models ~\cite{Sarwar1997, Butzen2007, Usman2013ANoC} because at a very low level every digital circuit (including modern processors) is composed of transistors. In these devices, there is a known relationship between power and frequency, which can be simplified by \cref{eq:power_simplified}: 
\begin{equation}
    P = \alpha+\beta f^3
    \label{eq:power_simplified}
\end{equation}
where $\alpha$ and $\beta$ are model parameters, and $f$ is the operating frequency (details of this equation are covered in Section~\ref{sec:models}). These types of models are suitable for optimization problems since they are a function of the operating frequency, which can be easily controlled.

\subsection{Performance models}
Another commonly used way to model the energy consumption of the system is the workload. The workload, which represents the amount of work done in a given time and speed, abstractly defines the application.

The workload ($W$) can be defined as~\cite{Paolillo2018OptimisationParallelism, Group2012HandbookSahni, Kim2015RacingHeuristics}:
\begin{equation}
    W = \int_{0}^{\tau}s(t)dt = s\tau
    \label{eq:workload_definition}
\end{equation}
where $\tau$ is total active time, and $s$ is the execution speed in instructions/second.
This equation is often taken up by utilization-based models, so that it could be applied to real systems.

Utilization can be defined as the ratio between the time that the system is active and the total time. Equation (\ref{eq:workload_utilization_tranlation}) defines workload in terms of CPU utilization ($u$):
\begin{equation}
    u = \frac{\tau}{T} = \frac{W/s}{T},
    \label{eq:utilization_definition}
\end{equation}
\begin{equation}
    W = usT,
    \label{eq:workload_utilization_tranlation}
\end{equation}
where $T$ the total execution time of the workload, and $s$ the execution speed.
Models based on CPU utilization are the base for DVFS algorithms. Even though the input is not controllable, it is straightforward  to measure system utilization with almost no overhead, and it is also very portable in terms of operating systems and architectures. 
%% Here is how to build the model or to optimize it, maybe too early to talk about AI without referencing existing approaches:
%There is no need for data and training, as with artificial intelligence algorithms, and adapts reasonably well to any machine. 
A downside is that optimization requires the knowledge of a wide variety of workloads which must be estimated somehow \cite{Fu2018RaceMinimization, Group2012HandbookSahni}.

% Markov model
Other approaches are also adopted to model energy, as for example in the work of Liu et al. \cite{Liu2016IntelligentSupplies} where the model uses a Markov chain to determine energy state transitions. This kind of approach needs heuristic algorithms to search for the best transitions and does not provide insights as an analytical equation does.

Although much work has been done in DVFS, the focus is still on the consumer electronics and laptop markets. 
For HPC, the notion of energy perception is relatively new~\cite{Beckman2005MakingSupercomputing}. 
Moreover, the operational characteristics of non-HPC and HPC systems are significantly different. 
First, the workload on non-HPC systems is very interactive with the end-user, but the workload on the HPC platform is not. 
Second, activities conducted on a non-HPC platform tend to share more machine resources.
In contrast, in HPC each job often runs with dedicated resources. 
Third, an HPC system usually is much larger than non-HPC systems, making it more  challenging to gather information, organize decisions, and execute global decisions.
Therefore, it is worthwhile to investigate whether a DVFS scheduling algorithm, which works well for conventional computing, remains effective for HPC.

\section{Related work} \label{sec:relatedwork}

Merkel et al.~\cite{Merkel2006BalancingSystems} developed an energy model for processors based on events. By counting a set of $n$ events and assuming that the processor consumes a fixed amount of energy $\alpha_i$ for each activity, they estimate the energy in a particular period as the number of occurrences $c_i$  multiplied by its correspondent weight $\alpha_i$:
\begin{equation}
    E = \sum_{i=1}^{n}\alpha_ic_i.
\end{equation}

Another event-based model introduced by Roy et al. \cite{Roy2013AnAlgorithms}, described the computational energy consumed by a CPU for an algorithm $A$:
\begin{equation}
    E(A) = P_{clk}T(A) + P_wW(A),
\end{equation}
where $P_{clk}$ is a processor clock leakage power, $T(A)$ is the total execution time, $W(A)$ is the total time taken by  non-I/O operations, and $P_w$ is used to capture the power consumption per operation performed by the CPU. $T(A)$ and $W(A)$ are estimated using performance features.

Models based on events are implemented using performance counters \cite{Fahad2019AComputing}, which have some drawbacks. They are highly dependent on the operating system and its architecture, making them not so easily portable. Additionally, there are limitations on the number of counters that can be read simultaneously. Furthermore, reading too many counters can add a non-negligible overhead. Thus, modern CPUs start to multiplex counters when more than a few are requested. There are also some well know problems regarding the precision of some events \cite{Weaver2008, Weaver2013a, Das2019SoK:Security, McGuire2009, Ramos2019AnCounters, Silva-De-Souza2020Containergy-aWorkloads}. Events that should be exact and deterministic (such as the number of executed instructions) show run-to-run variations and over-count on x86\_64 machines, even when running in strictly controlled environments. Our proposal is not based on performance counters and, therefore, is not vulnerable to those drawbacks.

Instruction-level energy model was also proposed in \cite{Characterization2013EnergyProcessor}, obtained from an energy per instruction (EPI) characterization made on Xeon Phi. Their model is expressed as:
\begin{equation}
    E(f) = \frac{(p_1 - p_0)(c_1 - c_0)/f}{N}, 
\end{equation}
where $N$ is the total number of dynamic instructions, $p_0$ is the initial idle power, $p_1$ is the average dynamic power and ($c_1$-$c_0$) refers to the cumulative number of cycles the micro-benchmark performs. This model is suitable for estimating the energy after the application finishes executing, when it is possible to count the number of total cycles. However, it is challenging to use for optimization or forecasting since it does not have an application model to predict the number of cycles. Our model integrates the behavior of the application taking into account the execution time.

In \cite{Lewis2008Run-timeSystems} Lewis et al. described the overall system energy consumption using the following equation:
\begin{equation}
E = A_0(E_{proc} + E_{mem}) + A_1E_{em} + A_2E_{board} + A_3E_{hdd},    
\end{equation}
where, $A_0$, $A_1$, $A_2$, and $A_3$ are unknown constants that are calculated via linear regression analysis and those remain constant for a specific server architecture. This model, as the previous one, relies on knowledge of energy spent on each component, being  a suitable option for estimation after the application has already ran, but not for optimization of the run itself.

In another energy consumption model  based on system utilization, Mills et al. modeled the energy consumed by a compute node with CPU (single) executing at speed $\sigma$ as \cite{Mills2014EnergySystems},
\begin{equation}
    E(\sigma,[t_1,t_2]) = \int_{t_1}^{t_2} \sigma^3 + \rho \sigma_max^3 dt,
\end{equation}

where $\rho$ stands for the overhead power  consumed regardless the processor speed, $t_1$ and $t_2$ are the initial and final execution time of the application. The overhead includes power consumption by all other system components such as memory, network, etc. For this reason, although the authors mentioned the energy consumption of a socket, their power model is generalized to the entire server.

Our work proposes a full-system energy model based on the CPU frequency and the number of cores.
The aim of the model is to understand and optimize the energy behavior of parallel applications in HPC systems according to application parameters such as the degree of parallelism and CPU parameters related to dynamic and static power. The proposed model differs from existing ones for including the frequency and number of cores in the same equation for estimating the energy for a specific application in a given configuration. This model can serve as a base for DVFS and DPM optimization problems that include frequency and number of active cores. It can also be used to analyze the contribution of each parameter (ex: level of parallelism) to energy consumption. The number of cores is an essential factor in HPC since applications are designed to run on multiple cores.

The proposed energy model is the product of an application-agnostic power model and an architecture-specific application performance model. The power model is based on the CMOS logic gates power draw as a function of the frequency~\cite{Sarwar1997, Butzen2007} augmented to include the number of cores. The performance model is based on Amdahl's law \cite{Amdahl1967ValidityCapabilities, Eyerman2010ModelingDesign, Shi2015ReevaluatingLaw}, which can be used to estimate runtime in multi-core systems. This model has been extended to include execution frequency and input size, characterizing the application on the target architecture.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Modeling energy with performance and power} \label{sec:models}
In this section, the three models proposed for power, performance and energy are described.

\subsection{Power Model} \label{sec:powermodel}
The complexity of the modern processor's circuits makes it very difficult to consider all the components and interconnections. A viable approach for modeling the CPU's power draw is to model their building components, mainly made out of logic gates. Thus, modeling the power consumption can be resumed to model the logic gates and multiplying this by the total number of gates, reducing the complexity of the modeling process.

The mature technology to manufacture logic gates is CMOS. Nowadays, it has been replaced by FINFET. In general, in these technologies, there are three main components of power dissipation \cite{Rauber2014, Goel2016, Du2017, Gonzalez1997},  namely, static power $P_{\rm static}$, dynamic power $P_{\rm dynamic}$, and leakage power $P_{\rm leak}$, that accumulated compose the total power draw.
% \begin{equation}
% P=P_{\rm static}+P_{\rm leak}+P_{\rm dynamic}.
% \label{eq:power_breakdown}
% \end{equation}

The dynamic power and leakage power behavior can be approximated by \cref{eq:power_dyn} and \cref{eq:power_leak}, respectively, as shown by Sarwar et al. and Butzen et al~\cite{Sarwar1997, Butzen2007}.
\begin{equation}
P_{dynamic}=CV^2f,
\label{eq:power_dyn}
\end{equation}
\begin{equation}
P_{leak} \propto V,
\label{eq:power_leak}
\end{equation}
where $C$ is the load capacitance, $V$ the voltage applied to the circuit and $f$ the switching frequency.

Another common approximation is to expect a linear relationship between the voltage and the applied frequency~\cite{Usman2013ANoC} such that:
\begin{equation}
f \propto V,
\label{eq:f_v}
\end{equation}
Thus, the proposed model for one processing core of a multi-core processor is derived by using \cref{eq:power_dyn}, \cref{eq:power_leak} and \cref{eq:f_v} to write \cref{eq:total_power}.
\begin{equation}
P(f)= c_1f^3+c_2f+c_3,
\label{eq:total_power}
\end{equation}
where $c_1$ $c_2$, and $c_3$ are the model's parameters associated with the dynamic, leakage and static power aspects, respectively. Including the number of active cores $p$, the proposed estimation of the power consumption of the whole processor becomes \cref{eq:power_final}
\begin{equation}
P(f,p)= p(c_1f^3+c_2f)+c_3,
\label{eq:power_final}
\end{equation}

\subsection{Performance Model} \label{sec:performancemodel}
To model the application execution time, we consider a program as a set of instructions executed on a mean frequency $f$ with $c_k$ instructions per cycle. The time $T_f$ that this program will take to complete at a given frequency is devised as follows:
\begin{equation}
T_f=\frac{I}{c_kf},
\label{eq:freqrel}
\end{equation}
where $I$ is the total number of instructions and $c_k$ the ratio of instructions per unit of time.

The next step is to include the number of cores in the equation. Amdahl's law \cite{Amdahl1967ValidityCapabilities}, described in \cref{eq:amdahl}, gives the theoretical background for that. It describes the speedup in latency of the execution of a task at a fixed workload.
\begin{equation}
S=\frac{T_s}{T_p}=\frac{1}{1-w+\frac{w}{p}},
\label{eq:amdahl}
\end{equation}
where $T_s$ is the serial time, $T_p$ the parallel time, $S$ is the theoretical speedup of the execution of the whole task, $w$ is the proportion of the execution time that benefits from improving system resources, and $p$ is the part of the task that benefits from improved system resources. Combining this with \cref{eq:freqrel}, the parallel time at frequency $f$ can be written as:
\begin{equation}
T_p=\frac{T_s}{S}=\frac{T_f}{\frac{1}{1-w+\frac{w}{p}}},
\label{eq:parallel_time}
\end{equation}

We can then write the equation of the program execution time as a function of frequency, number of cores and parallelism  as \cref{eq:performance} and subsequently derive \cref{eq:performance_2}:
\begin{equation}
T(f,p)=\frac{I}{ \frac{c_kf}{1-w+\frac{w}{p}} },
\label{eq:performance}
\end{equation}
\begin{equation}
T(f,p)=\frac{d_1(p-wp+w)}{fp},
\label{eq:performance_2}
\end{equation}
where $d_1$ is a constant.

Finally, to fully characterize the application, a parameter representing the application's workload, called input size $N$, is introduced, representing the number of basic operations need to complete a problem \cite{Kumar1994AnalyzingArchitectures}. In Oliveira et al. \cite{Oliveira2018ApplicationCores}, they showed that this parameter could generally be described as exponential. Therefore the proposed performance model is presented in \cref{eq:performance_final}. This resulting equation can describe the behavior of the execution time of a program for an input $N$, frequency $f$, and active cores $p$:
\begin{equation}
T(f,p,N)=\frac{d_1N^{d_2}(p-wp+w)}{fp},
\label{eq:performance_final}
\end{equation}
where $d_1$, $d_2$ and $w$ are constants that depend on the application.

\subsection{Energy Model} \label{sec:energymodel}
Combining the power model output described in~\cref{sec:powermodel} and the characterization of the application performance described in \cref{sec:performancemodel}, the total energy can be modeled as:
\begin{equation}
E(f,p,N)=P(f,p)\times{\rm T}(f,p,N),
\label{eq:en_combination}
\end{equation}
where $P(f,p)$ is the total power modeled by~\cref{eq:power_final}, ${T}(f,p,N)$ is the execution time estimated by the \cref{eq:performance_final}, $f$ is the frequency, $p$ is the number of active cores, and $N$ is the input size. The final equation can be written as:
\begin{equation}
    E(f,p,N)=\frac{d_1N^{d_2}(p-wp+w)(p(c_1f^3+c_2f)+c_3)}{fp}.
    \label{eq:en_final}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Experimental validation} \label{sec:experimentalvalidation}
In this section, the models presented in~\cref{sec:powermodel} and~\cref{sec:performancemodel} were validated with a benchmark specific for multi-core architectures. Additionally, in order to assess the modeling overhead and accuracy, our proposal was then compared to machine learning approaches. We compared against Support Vector Regression (SVR)~\cite{Smola2004}, Decision Tree \cite{Kitts2006RegressionLecture}, k-nearest neighbors \cite{Altman1992AnRegression}, Multilayer perceptron \cite{Murtagh1991MultilayerRegression}, and some new methods such as Gao et al. \cite{Gao2019DendriticPrediction}. However, SVR was chosen as the most representative because in our tests it performed best without an aggressive fine-tuning.

\subsection{Case-Study Applications} \label{sec:casestudyapplication}
The PARSEC parallel benchmark suite, version 3.0~\cite{Bienia2008}, OpenMC \cite{Romano2015OpenMC:Development} and LINPACK (HPL) \cite{Dongarra1988TheExplanation}, were chosen as case studies. The PARSEC benchmark focused on emerging workloads and was designed to represent the next-generation shared-memory programs for chip-multiprocessors. It covers an ample range of areas such as financial analysis, computer vision, engineering, enterprise storage, animation, similarity search, data mining, machine learning, and media processing. The OpenMC and the LINPACK are two classic HPC programs.

\subsection{Case-Study Architecture} \label{sec:casestudyarchitecture}
The experiments were executed in one computer node equipped with two Intel Xeon E5-2698 v3 processors with sixteen cores each and two hardware threads for each core. The overall view of the architecture is shown in Figure \ref{fig:architecture}. The maximum non-turbo frequency is 2.3GHz, and the total physical memory of the node is 128GB (8$\times$16GB). Turbo frequency and hardware multi-threading were disabled during all experiments. The operating system used was Linux CentOS 6.5, kernel 4.16.

\begin{figure}[htb!]
    \centering
    \includegraphics[width=\columnwidth]{architecture.png}
    \caption{Node architecture (the image was made with the lstop application).}
    \label{fig:architecture}
\end{figure}

The Linux kernel has many different policies for power management, depending on the driver. In the default driver, the acpi-cpufreq, the options are:
\begin{itemize}
    \item Powersave
    \item Performance
    \item Ondemand
    \item Conservative
    \item Userspace
\end{itemize}
In this work, the frequency control was performed using the Userspace governor, and the core control was accomplished by modifying the appropriate system files with the default CPU-hotplug driver.

The architecture is equipped with the Intelligent Platform Management Interface (IPMI), a set of interfaces allowing out-of-band management of computer systems and platform-status monitoring via the local network~\cite{Schwenkler2006IntelligentInterface}. It can monitor variables and resources such as the system's temperature, voltage, fans, and power supplies, with independent sensors attached to the hardware.

\subsection{Fitting the Models} \label{sec:fitting}
To find the parameters of the \cref{eq:en_final}, 10 uniformly random configurations of frequencies ($f$), cores ($p$) and inputs ($N$) were chosen from the range $1<=p<=32$, $1.2<=f<=2.2$ and $1<=N<=5$ respectively. The application was executed for each chosen configuration, and the measured  energy and time values were collected. For the input size if we assume that all CPU instructions are executed at approximately the same time, the number of basic operations will be directly correlated with the time. Thus, we can estimate the problem size by looking at the execution time, allowing us to divide a large problem size into several smaller ones knowing their relationship as done in the work of Oliveira ~\cite{Oliveira2018ApplicationCores}. The unity can also vary depending on the definition. In our case for simplicity, we assign numbers from 1 the smallest size to 10 the largest, increasing the problem linearly, that way its also possible to interpolate any input in between these values.

%The application's input was chosen to increase the workload linearly, from smallest to largest, so that most of the input space is covered~\cite{Oliveira2018ApplicationCores}.

For each configuration, samples of the power were collected using IPMI every 1 second. This sampling rate was chosen because the order of magnitude of the mean run time of the applications is minutes. Therefore, this rate provides enough samples to measure average power. Additionally, timestamps and the total run time were collected. The total energy spent on each configuration is estimated by first interpolation the power samples using a first order method and then integrating this function in the time.

From the sampled data, the parameters of the model can be estimated. This results in an optimization problem of finding the parameters that minimize the distance between estimated and the measured values. To solve this minimization problem, non-linear least squares have been applied.

The python library scikit-learning was used to build the SVR model~\cite{PedregosaFandVaroquauxGandGramfortAandMichel2011Scikit-learn:Python}. The SVR was trained using the same data used for parameters estimation of \cref{eq:en_final} with a grid search used to find the best kernel function and the best values for the hyper-parameters penalty for the wrong ($C$) and ($\gamma$). For this data, the best function was the Radial Base Function (RBF), and the hyper-parameters were $C=10^4$ and $\gamma=0.5$.

\subsection{Verifying Hypothesis}
\subsubsection{Frequency and voltage relation}

\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{hypothesis/freq_volt_rel.png}
    \caption{Frequency voltage relation}
    \label{fig:freq_volt_rel}
\end{figure}

\begin{table}[H]
\caption{Frequency voltage relation}
\centering
\begin{tabular}{|c|c|c|}
\hline
\begin{tabular}[c]{@{}c@{}}Freq. \\ Setted\end{tabular} & Volt. & \begin{tabular}[c]{@{}c@{}}Measured \\ Freq.\end{tabular} \\ \hline
2.2                                                     & 0.77  & 2.199                                                     \\ \hline
2.1                                                     & 0.76  & 2.099                                                     \\ \hline
2.0                                                     & 0.75  & 2.000                                                     \\ \hline
1.9                                                     & 0.74  & 1.899                                                     \\ \hline
1.8                                                     & 0.73  & 1.799                                                     \\ \hline
1.7                                                     & 0.72  & 1.699                                                     \\ \hline
1.6                                                     & 0.71  & 1.599                                                     \\ \hline
1.5                                                     & 0.70  & 1.500                                                     \\ \hline
1.4                                                     & 0.69  & 1.397                                                     \\ \hline
1.3                                                     & 0.68  & 1.297                                                     \\ \hline
1.2                                                     & 0.67  & 1.200                                                     \\ \hline
\end{tabular}
\end{table}

\subsubsection{Instructions executed}

\begin{table}[H]
\caption{Cores variation}
\begin{tabular}{|c|c|c|c|}
\hline
Application  & Instructions & Dev.     & Dev. (\%) \\ \hline
Vip          & 7.97e+11     & 7.16e+06 & 0.00\%    \\ \hline
Openmc       & 8.17e+07     & 1.65e+04 & 0.02\%    \\ \hline
Rtview       & 9.91e+12     & 1.55e+09 & 0.02\%    \\ \hline
X264         & 4.52e+11     & 5.81e+07 & 0.01\%    \\ \hline
Bodytrack    & 1.86e+12     & 3.95e+10 & 2.13\%    \\ \hline
Fluidanimate & 2.09e+12     & 8.44e+10 & 4.04\%    \\ \hline
Xhpl         & 1.14e+08     & 1.24e+05 & 0.11\%    \\ \hline
Blackschole  & 3.75e+12     & 1.40e+09 & 0.04\%    \\ \hline
Dedup        & 1.02e+11     & 5.74e+07 & 0.06\%    \\ \hline
Swapti       & 2.43e+12     & 8.87e+08 & 0.04\%    \\ \hline
Canneal      & 1.19e+11     & 4.46e+07 & 0.04\%    \\ \hline
Freqmine     & 1.27e+12     & 4.78e+08 & 0.04\%    \\ \hline
Ferret       & 4.76e+11     & 7.04e+07 & 0.01\%    \\ \hline
\end{tabular}
\end{table}

\begin{table}[H]
\caption{Frequency Variation}
\begin{tabular}{|c|c|c|c|}
\hline
Application  & Instructions & Dev.     & Dev. (\%) \\ \hline
Vip          & 7.97e+11     & 1.16e+06 & 0.00\%    \\ \hline
Openmc       & 8.17e+07     & 4.52e+03 & 0.01\%    \\ \hline
Rtview       & 9.91e+12     & 6.64e+05 & 0.00\%    \\ \hline
X264         & 4.52e+11     & 1.54e+05 & 0.00\%    \\ \hline
Bodytrack    & 1.84e+12     & 2.54e+05 & 0.00\%    \\ \hline
Fluidanimate & 2.38e+12     & 1.70e+09 & 0.07\%    \\ \hline
Xhpl         & 1.14e+08     & 5.95e+03 & 0.01\%    \\ \hline
Blackschole  & 3.75e+12     & 4.36e+05 & 0.00\%    \\ \hline
Dedup        & 1.02e+11     & 8.32e+07 & 0.08\%    \\ \hline
Swapti       & 2.43e+12     & 1.48e+05 & 0.00\%    \\ \hline
Canneal      & 1.19e+11     & 3.01e+05 & 0.00\%    \\ \hline
Freqmine     & 1.27e+12     & 3.70e+08 & 0.03\%    \\ \hline
Ferret       & 4.76e+11     & 5.63e+07 & 0.01\%    \\ \hline
\end{tabular}
\end{table}

\subsubsection{Input size and instructions}

\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{hypothesis/input_instructions/fp/Blackschole.png}
    \caption{Caption}
    \label{fig:my_label}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{hypothesis/input_instructions/fp/Canneal.png}
    \caption{Caption}
    \label{fig:my_label}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{hypothesis/input_instructions/input_time/Blackschole.png}
    \caption{Caption}
    \label{fig:my_label}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{hypothesis/input_instructions/input_time/Canneal.png}
    \caption{Caption}
    \label{fig:my_label}
\end{figure}

\subsection{Measured versus Modeled Energy}
\label{sec:measuredversusmodeledenergy}

To validate the proposed energy model, all possible configurations were tested by varying the cores in a range of $1<=p<=32$, the frequency in $1.2<=f<=2.2$ and the input in $1<=N<=5$, which varies from 400 to over 1000 configurations depending on the application as some applications have restrictions on the number of cores. The mean percentage error was computed as the difference between the estimated and measured values according to the following equation:
\begin{equation}
    MPE = \frac{1}{N} \sum_i^N \frac{|y_{\rm estimated}-y_{\rm measured}|}{y_{\rm measured}}.
    \label{eq:mpe}
\end{equation}

\subsubsection{Frequency x Cores}
Figures \ref{fig:en_eq_black}, \ref{fig:en_eq_canneal} plot the measured and modeled energy consumption for some of the applications modeled. They  show some of the possible shapes that the model can take while varying the number of active cores, operating frequency, and input size.
\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
    	\centerline{\includegraphics[width=\columnwidth]{energy/freq_cores/completo_black_5.png}}
        \caption{Blackscholes.}
    	\label{fig:en_eq_black}
    \end{subfigure}
    %
    \begin{subfigure}[b]{0.48\textwidth}
	\centerline{\includegraphics[width=\columnwidth]{energy/freq_cores/completo_canneal_1.png}}
    \caption{Canneal.}
	\label{fig:en_eq_canneal}
    \end{subfigure}
	\caption{Example fit for a specific input size: Blackscholes (a) and Canneal (b).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
	}
\end{figure}
\subsubsection{Frequency x Input}
Figures \ref{fig:en_eq_black}, \ref{fig:en_eq_canneal} plot the measured and modeled energy consumption for some of the applications modeled. They  show some of the possible shapes that the model can take while varying the number of active cores, operating frequency, and input size.
\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
    	\centerline{\includegraphics[width=\columnwidth]{energy/freq_inps/completo_black_5.png}}
        \caption{Blackscholes.}
    	\label{fig:en_eq_black}
    \end{subfigure}
    %
    \begin{subfigure}[b]{0.48\textwidth}
	\centerline{\includegraphics[width=\columnwidth]{energy/freq_inps/completo_canneal_1.png}}
    \caption{Canneal.}
	\label{fig:en_eq_canneal}
    \end{subfigure}
	\caption{Example fit for a specific input size: Blackscholes (a) and Canneal (b).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
	}
\end{figure}
\subsubsection{Cores x Input}
Figures \ref{fig:en_eq_black}, \ref{fig:en_eq_canneal} plot the measured and modeled energy consumption for some of the applications modeled. They  show some of the possible shapes that the model can take while varying the number of active cores, operating frequency, and input size.
\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
    	\centerline{\includegraphics[width=\columnwidth]{energy/cores_inps/completo_black_5.png}}
        \caption{Blackscholes.}
    	\label{fig:en_eq_black}
    \end{subfigure}
    %
    \begin{subfigure}[b]{0.48\textwidth}
	\centerline{\includegraphics[width=\columnwidth]{energy/cores_inps/completo_canneal_1.png}}
    \caption{Canneal.}
	\label{fig:en_eq_canneal}
    \end{subfigure}
	\caption{Example fit for a specific input size: Blackscholes (a) and Canneal (b).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
	}
\end{figure}

\subsubsection{Validation}
The validation results for each application trained with 10 random configurations are displayed in figure \ref{fig:mpe_svr_eq} and the raw MPE values in  table~\ref{tab:mpe_svr_eq}.
\begin{figure}[htb!]
	\includegraphics[width=\columnwidth]{mpe_svr_eq.png}
	\caption{Comparison of the Mean Percentage Error between the proposed model and SVR. "Model mean" and "SVR mean" are the average of all MPE values for all applications.
	}
	\label{fig:mpe_svr_eq}
\end{figure}
\begin{table}[htb!]
\centering
\begin{tabular}{|c|c|c|}
\hline
Application  & Model & SVR   \\ \hline
Ferret       & 5.25     & 12.49  \\ \hline
Raytrace     & 6.36     & 11.95  \\ \hline
Fluianimate  & 2.44     & 22.90  \\ \hline
x264         & 8.28     & 15.33  \\ \hline
Vips         & 7.54     & 10.80  \\ \hline
Swaptions    & 6.54     & 18.57  \\ \hline
Canneal      & 3.12     & 6.13   \\ \hline
Dedup        & 8.85     & 13.70  \\ \hline
Freqmine     & 2.44     & 3.24   \\ \hline
Blackscholes & 2.18     & 11.00  \\ \hline
HPL          & 7.47     & 12.75  \\ \hline
Bodytrack    & 16.98    & 34.12  \\ \hline
Openmc       & 11.15    & 24.34  \\ \hline
\end{tabular}
\caption{Comparison of the Mean Percentage Error between the proposed model and SVR: raw values.}
\label{tab:mpe_svr_eq}
\end{table}

Figure \ref{fig:mpe_svr_eq} shows that the proposed model always performed better, with a lower MPE, than SVR when we were limited to 10 training points. This result will be further explored in \cref{sec:overhead} where we compare with different training sizes.

\subsection{Overheads on training} \label{sec:overhead}
It is known that machine learning is data-driven, in that sense the SVR model obtained using only 10 configurations could be improved, but what about the analytical model? 
To answer that question, the proposed model and the SVR were also trained with a varying  number of configurations.
We then compared the MPE and the amount of energy spent to create each model. 
This accuracy-energy trade-off is crucial since the energy overhead when building models defeats the primary goal of saving power when running applications.

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
    	\centerline{\includegraphics[width=\columnwidth]{overhead/completo_ferret_4.png}}
        \caption{MPE for Ferret.}
    	\label{fig:overhead_ferret}
    \end{subfigure}
    %
    \begin{subfigure}[b]{0.45\textwidth}
    	\centerline{\includegraphics[width=\columnwidth]{overhead/completo_vips_4.png}}
        \caption{MPE for Vips.}
    	\label{fig:overhead_vips}
    \end{subfigure}
    %
    \begin{subfigure}[b]{0.45\textwidth}
    	\centerline{\includegraphics[width=\columnwidth]{overhead/completo_x264_4.png}}
        \caption{MPE for x264.}
    	\label{fig:overhead_x264}
    \end{subfigure}
    \caption{MPE case studies: Ferret (a), Vips (b) and x264 (c): model (almost constant) versus SVR (lowest MPE beyond train-size 35 on case (b)).}
    \label{fig:overheadFerretVips}
\end{figure}

Figure \ref{fig:overheadFerretVips} shows the comparisons of MPE and energy spent to create each model for two selected applications. According to the results, the analytical model is found to be very stable, not changing much as more data is added, while the SVR keeps reshaping to adapt to the data. 
The error of the analytical model is almost constant but that of  the SVR, initially very high, drops at some point.
\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.44\textwidth}
    	\centerline{\includegraphics[width=\columnwidth]{overhead/overall.png}}
        \caption{Average energy spent on all applications during model creation. The two curves are identical because the same data were used to adjust the SVR and the model.}
    	\label{fig:overall_overhead}
    \end{subfigure}
    %
    \quad
    \begin{subfigure}[b]{0.44\textwidth}
    	\centerline{\includegraphics[width=\columnwidth]{overhead/overall_mpe.png}}
        \caption{
        MPE all applications: SVR needs 10 times more data to have an MPE lower than the model.}
    	\label{fig:overall_MPE}
    \end{subfigure}
    \caption{Overall results for energy and MPE for each training size.}
    \label{fig:overall_train}
\end{figure}
Figure \ref{fig:overall_train} presents the overall results, with the mean energy overhead and MPE for all applications.
The meeting point of the MPE for the SVR and the proposed model can be extracted from  \ref{fig:overall_MPE}.
There it shows that around 90 configurations the SVR starts to have a smaller error. The cost of that is the linear increase in energy spent on training. The increase in energy, about 10 times more, can be observed in Figure \ref{fig:overall_overhead}.

\subsection{Analysis}

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{analisys/pareto_static_high.png}
    \caption{Pareto static energy}
    \label{fig:pareto_static_h}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{analisys/pareto_static_low.png}
    \caption{Pareto static energy}
    \label{fig:pareto_static_l}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{analisys/pareto_w_high.png}
    \caption{Pareto w energy}
    \label{fig:pareto_w_h}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{analisys/pareto_static_low.png}
    \caption{Pareto w energy}
    \label{fig:pareto_w_l}
\end{figure}

\subsection{DVFS and DPM optimization} \label{sec:dvfs_optmin}
The effectiveness of the proposed approach during optimization was evaluated with a simple algorithm that finds the optimal frequency and number of active cores from the equation proposed. The results were then compared to the Linux default choices for power management.

% With \cref{eq:en_final}, it is possible to calculate energy consumption estimates for each possible configuration since there is a finite range of possible values for the frequency and number of cores. It is also possible to apply constraints on the execution time, frequency, and number of active cores. Then, the configuration that minimizes energy consumption for a given input can be selected.

% Current HPC managers leave the user the choice of how many cores to use. On this basis, three situations were analyzed with relation to the number of cores:

% \begin{enumerate}
%     \item Worst choice: number of cores that maximize the total energy consumed;
%     \item Random choice: energy consumed for a random choice of number of cores;
%     \item Best choice: number of cores that minimize the total energy consumed (oracle).
% \end{enumerate}

% The default option for the Linux governor is Ondemand, and by default, it has no DPM control for the number of active cores. As Ondemand only does DVFS, for comparison, each application was executed with all available cores in the system, from 1 to 32.

% Figures \ref{fig:energy_worst_case}, \ref{fig:energy_mean_case}, and \ref{fig:energy_best_case}, show the energy savings with respect to Ondemand, i.e. $\frac{Ondemand-Model_{min}}{Ondemand}$ for the three cases described above. The savings and losses for each case are:
% \begin{enumerate}
%     \item Worst choice: save 69.88\% on average;
%     \item Random choice: save 12.04\% on average;
%     \item Best choice: lost 14.06\% on average.
% \end{enumerate}
% \begin{figure}[htb!]
%     \centering
%     \includegraphics[width=\columnwidth]{dvfs_cmp_max.png}
%     \caption{Energy savings comparisons between the proposed model and the Worst case.}
%     \label{fig:energy_worst_case}
% \end{figure}
% \begin{figure}[htb!]
%     \centering
%     \includegraphics[width=\columnwidth]{dvfs_cmp_mean.png}
%     \caption{Energy savings comparisons between the proposed model and the Random case.}
%     \label{fig:energy_mean_case}
% \end{figure}
% \begin{figure}[htb!]
%     \centering
%     \includegraphics[width=\columnwidth]{dvfs_cmp_32.png}
%     \caption{Energy savings comparisons between the proposed model and the Best case.}
%     \label{fig:energy_best_case}
% \end{figure}

% By default, operating systems do not implement DPM at the core level, and in HPC the user usually explicitly chooses the number of cores to run their job. To give a better idea of the impact on the energy consumption of DPM at the core level, we analyzed the choices of the number of cores over a period of one year in the HPC center of the UFRN. The result is plotted in Figure \ref{fig:cpu_requests}.
% \begin{figure}[htb!]
%     \centering
%     \includegraphics[width=\columnwidth]{cpu_request.png}
%     \caption{Number of CPU requests during one year in HPC cluster, sorted by the number of cores requested per job.}
%     \label{fig:cpu_requests}
% \end{figure}

% Observe that the most common choice of many regular users is a single core requested per job, matching the worst case choice for all application analysed in this work. The best choice was quite often 32 cores, which is the third most popular choice among users, but it is 72 times less frequent than 1 core. This leads us to envision how much energy could be saved and encourage us towards future work using the proposed model for DPM or more advanced optimization algorithms.

% In practice, this approach can be brought into production by allowing the resource manager to perform these changes for the user using pre-scripts and post-scripts for high energy consumption job submissions.

\section{Conclusion} \label{sec:conclusion}
In this work we have proposed an energy model based on the operating frequency and the number of cores for a shared memory system. This model serves as a reference for DVFS and DPM optimization problems.

Results from 3 different HPC benchmarks demonstrate the potential of the proposed model. While consuming  10 times less energy than a machine learning approach, such as SVR, to characterize applications, it can be used to provide knowledgeable hints to improve DVFS and DPM algorithms by enabling analysis of the contribution of each model parameter (e.g. level of parallelism) to the energy consumption. Indeed, as shown in Section \ref{sec:dvfs_optmin}, when no oracle is available to choose the frequency and the number of cores the application should use, the proposed model can save around 12\% of energy for a random choice and up to 70\% for the worse possible choice. Considering the job history of our own HPC center which shows the prevalence of worse choices made by users, the potential energy savings are very significant and encourages further research.

Future work will demonstrate all possible analyzes achievable with the proposed equation as more advanced DVFS models can be derived from it. For instance, identifying the different phases of a target program will allow finer changes in frequency and, perhaps, in the number of active cores, to further improve the results presented here.

% \appendices
\section*{Acknowledgment}
We would like to acknowledge the use of UFRN´s High-Performance Computing Center (NPAD/UFRN), for running the experiments of this work, and thank NPAD´s team for providing the necessary support to perform the energy measurements. 
% \footnote{NPAD is the name of the High Performance Computing Center (from Portuguese: Núcleo de Processamento de Alto Desempenho) of the Universidade Federal do Rio Grande do Norte (UFRN)}.

\bibliographystyle{IEEEtran}
\bibliography{references.bib}

%\bibliographystyle{IEEEtran}
%\begin{thebibliography}{00}
%\bibliography{references}
%\end{thebibliography}

\section*{Appendix}

The following abbreviations are used in this manuscript:\\

\noindent 
\begin{tabular}{@{}ll}
DVFS Dynamic Frequency and Voltage Scaling\\
DPM Dynamic Power Management\\
SVR Support Vector Regression\\
RBF Radial Base Function\\
HPC High Performance Computing\\
IPMI The Intelligent Platform Management Interface\\
RBF Radial Base Function\\
MPE Mean Percentage Error
\end{tabular}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./figures/biography_vitorramos}}]{Vitor R. G. Silva}
Bsc in Science and Technology in 2016 and Computer Engineer in 2018 by Federal University of Rio Grande do Norte. His research interests are high-performance computing, operating systems, cybersecurity and Artificial intelligence.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./figures/biography_carlos}}]{Carlos A. V. Sakuyama}
Professor Valderrama is, since 2004, Director of the Electronics and Microelectronics Department at the Polytechnic Faculty of the University of Mons in Belgium.  His Department is member of the New Media Art Technology and the Information Technology Institutes. He obtained the Ph.D. degree in Microelectronics at the Institute Nationale Polytechnique de Grenoble INPG (today Grenoble Institute of Technology) at the TIMA lab in France, in 1998, the M.Sc. diploma at the Federal University of Rio de Janeiro, Brazil, in 1993, and the electrical-electronics engineering diploma at the National University of Cordoba, Argentina, in 1989. He was invited professor in several universities, the Catholic University of Cordoba (Argentina), the Federal University of Pernambuco and the Federal University of Rio Grande do Norte (Brazil) and the University of Castilla La Mancha (Spain). He was responsible for the creation of the spinoff Nsilition (2009, funded by the Walloon Region). He has participated in more than 18 national and international research projects from the development of 4G chips, tracking devices and architectures for IoT, HPC and space.  He serves as technical reviewer and committee member of multiple journals and international conferences. His research activity is supported by more than 180 publications on international conferences, more than 10 books chapters, and more than 30 scientific journals. He is IEEE senior member since 2006.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./figures/samuel}}]{Samuel Xavier-de-Souza}
SAMUEL XAVIER-DE-SOUZA (Senior Member, IEEE) was born in Natal, Brazil. He received the Computer Engineering degree from the Universidade Federal do Rio Grande do Norte-UFRN, Brazil, in 2000, and the Ph.D. degree in electrical engineering from Katholieke Universiteit Leuven, Belgium, in 2007. He worked for IMEC, Belgium, as a Software/Hardware Engineer, and for the Flemish Supercomputing Center, Belgium, as a High-Performance Computing Consultant. In 2009, he joined the Department of Computer Engineering and Automation of UFRN, where he currently holds the position of an Associate Professor. He is also a Founder and the Director of UFRN’s High-Performance Computing Center-NPAD. In 2016, he became a Royal Society-Newton Advanced Fellow. His research interests include software energy, scalable and efficient parallel systems, parallel algorithms, parallel architectures, and their applications.
\end{IEEEbiography}

\EOD

\end{document}

### 2021-03-23 thesis/test/name.tex
add tex files

\documentclass{ieeeaccess}
\usepackage{subcaption}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\usepackage{caption,setspace}
\captionsetup{font={sf,small,stretch=0.80},labelfont={bf,color=accessblue}}

\begin{document}
\history{Date of publication xxxx 00, 0000, date of current version xxxx 00, 0000.}
\doi{10.1109/ACCESS.2017.DOI}

\title{Preparation of Papers for IEEE ACCESS}
\author{\uppercase{First A. Author}\authorrefmark{1}, \IEEEmembership{Fellow, IEEE},
\uppercase{Second B. Author\authorrefmark{2}, and Third C. Author,
Jr}.\authorrefmark{3},
\IEEEmembership{Member, IEEE}}
\address[1]{National Institute of Standards and 
Technology, Boulder, CO 80305 USA (e-mail: author@boulder.nist.gov)}
\address[2]{Department of Physics, Colorado State University, Fort Collins, 
CO 80523 USA (e-mail: author@lamar.colostate.edu)}
\address[3]{Electrical Engineering Department, University of Colorado, Boulder, CO 
80309 USA}
\tfootnote{This paragraph of the first footnote will contain support 
information, including sponsor and financial support acknowledgment. For 
example, ``This work was supported in part by the U.S. Department of 
Commerce under Grant BS123456.''}

\markboth
{Author \headeretal: Preparation of Papers for IEEE TRANSACTIONS and JOURNALS}
{Author \headeretal: Preparation of Papers for IEEE TRANSACTIONS and JOURNALS}

\corresp{Corresponding author: First A. Author (e-mail: author@ boulder.nist.gov).}

\begin{abstract}
These instructions give you guidelines for preparing papers for 
IEEE Access. Use this document as a template if you are 
using \LaTeX. Otherwise, use this document as an 
instruction set. The electronic file of your paper will be formatted further 
at IEEE. Paper titles should be written in uppercase and lowercase letters, 
not all uppercase. Avoid writing long formulas with subscripts in the title; 
short formulas that identify the elements are fine (e.g., "Nd--Fe--B"). Do 
not write ``(Invited)'' in the title. Full names of authors are preferred in 
the author field, but are not required. Put a space between authors' 
initials. The abstract must be a concise yet comprehensive reflection of 
what is in your article. In particular, the abstract must be self-contained, 
without abbreviations, footnotes, or references. It should be a microcosm of 
the full article. The abstract must be between 150--250 words. Be sure that 
you adhere to these limits; otherwise, you will need to edit your abstract 
accordingly. The abstract must be written as one paragraph, and should not 
contain displayed mathematical equations or tabular material. The abstract 
should include three or four different keywords or phrases, as this will 
help readers to find it. It is important to avoid over-repetition of such 
phrases as this can result in a page being rejected by search engines. 
Ensure that your abstract reads well and is grammatically correct.
\end{abstract}

\begin{keywords}
Enter key words or phrases in alphabetical 
order, separated by commas. For a list of suggested keywords, send a blank 
e-mail to keywords@ieee.org or visit \underline
{http://www.ieee.org/organizations/pubs/ani\_prod/keywrd98.txt}
\end{keywords}

\titlepgskip=-15pt

\maketitle

\section{Introduction}
\label{sec:introduction}
\PARstart{T}{his} document is a template for \LaTeX. If you are 
reading a paper or PDF version of this document, please download the 
electronic file, trans\_jour.tex, from the IEEE Web site at \underline
{http://ieeeauthorcenter.ieee.org/create-your-ieee-article/}\break\underline{use-authoring-tools-and-ieee-article-templates/ieee-article-}\break\underline{templates/} so you can use it to prepare your manuscript. If 
you would prefer to use LaTeX, download IEEE's LaTeX style and sample files 
from the same Web page. You can also explore using the Overleaf editor at 
\underline
{https://www.overleaf.com/blog/278-how-to-use-overleaf-}\break\underline{with-ieee-collabratec-your-quick-guide-to-getting-started}\break\underline{\#.xsVp6tpPkrKM9}

If your paper is intended for a conference, please contact your conference 
editor concerning acceptable word processor formats for your particular 
conference.

IEEE will do the final formatting of your paper. If your paper is intended 
for a conference, please observe the conference page limits.

\subsection{Abbreviations and Acronyms}
Define abbreviations and acronyms the first time they are used in the text, 
even after they have already been defined in the abstract. Abbreviations 
such as IEEE, SI, ac, and dc do not have to be defined. Abbreviations that 
incorporate periods should not have spaces: write ``C.N.R.S.,'' not ``C. N. 
R. S.'' Do not use abbreviations in the title unless they are unavoidable 
(for example, ``IEEE'' in the title of this article).

\subsection{Other Recommendations}
Use one space after periods and colons. Hyphenate complex modifiers: 
``zero-field-cooled magnetization.'' Avoid dangling participles, such as, 
``Using \eqref{eq}, the potential was calculated.'' [It is not clear who or what 
used \eqref{eq}.] Write instead, ``The potential was calculated by using \eqref{eq},'' or 
``Using \eqref{eq}, we calculated the potential.''

Use a zero before decimal points: ``0.25,'' not ``.25.'' Use 
``cm$^{3}$,'' not ``cc.'' Indicate sample dimensions as ``0.1 cm 
$\times $ 0.2 cm,'' not ``0.1 $\times $ 0.2 cm$^{2}$.'' The 
abbreviation for ``seconds'' is ``s,'' not ``sec.'' Use 
``Wb/m$^{2}$'' or ``webers per square meter,'' not 
``webers/m$^{2}$.'' When expressing a range of values, write ``7 to 
9'' or ``7--9,'' not ``7$\sim $9.''

A parenthetical statement at the end of a sentence is punctuated outside of 
the closing parenthesis (like this). (A parenthetical sentence is punctuated 
within the parentheses.) In American English, periods and commas are within 
quotation marks, like ``this period.'' Other punctuation is ``outside''! 
Avoid contractions; for example, write ``do not'' instead of ``don't.'' The 
serial comma is preferred: ``A, B, and C'' instead of ``A, B and C.''

If you wish, you may write in the first person singular or plural and use 
the active voice (``I observed that $\ldots$'' or ``We observed that $\ldots$'' 
instead of ``It was observed that $\ldots$''). Remember to check spelling. If 
your native language is not English, please get a native English-speaking 
colleague to carefully proofread your paper.

Try not to use too many typefaces in the same article. You're writing
scholarly papers, not ransom notes. Also please remember that MathJax
can't handle really weird typefaces.

\subsection{Equations}
Number equations consecutively with equation numbers in parentheses flush 
with the right margin, as in \eqref{eq}. To make your equations more 
compact, you may use the solidus (~/~), the exp function, or appropriate 
exponents. Use parentheses to avoid ambiguities in denominators. Punctuate 
equations when they are part of a sentence, as in
\begin{equation}E=mc^2.\label{eq}\end{equation}

Be sure that the symbols in your equation have been defined before the 
equation appears or immediately following. Italicize symbols ($T$ might refer 
to temperature, but T is the unit tesla). Refer to ``\eqref{eq},'' not ``Eq. \eqref{eq}'' 
or ``equation \eqref{eq},'' except at the beginning of a sentence: ``Equation \eqref{eq} 
is $\ldots$ .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\section{Units}
Use either SI (MKS) or CGS as primary units. (SI units are strongly 
encouraged.) English units may be used as secondary units (in parentheses). 
This applies to papers in data storage. For example, write ``15 
Gb/cm$^{2}$ (100 Gb/in$^{2})$.'' An exception is when 
English units are used as identifiers in trade, such as ``3\textonehalf-in 
disk drive.'' Avoid combining SI and CGS units, such as current in amperes 
and magnetic field in oersteds. This often leads to confusion because 
equations do not balance dimensionally. If you must use mixed units, clearly 
state the units for each quantity in an equation.

The SI unit for magnetic field strength $H$ is A/m. However, if you wish to use 
units of T, either refer to magnetic flux density $B$ or magnetic field 
strength symbolized as $\mu _{0}H$. Use the center dot to separate 
compound units, e.g., ``A$\cdot $m$^{2}$.''

\section{Some Common Mistakes}
The word ``data'' is plural, not singular. The subscript for the 
permeability of vacuum $\mu _{0}$ is zero, not a lowercase letter 
``o.'' The term for residual magnetization is ``remanence''; the adjective 
is ``remanent''; do not write ``remnance'' or ``remnant.'' Use the word 
``micrometer'' instead of ``micron.'' A graph within a graph is an 
``inset,'' not an ``insert.'' The word ``alternatively'' is preferred to the 
word ``alternately'' (unless you really mean something that alternates). Use 
the word ``whereas'' instead of ``while'' (unless you are referring to 
simultaneous events). Do not use the word ``essentially'' to mean 
``approximately'' or ``effectively.'' Do not use the word ``issue'' as a 
euphemism for ``problem.'' When compositions are not specified, separate 
chemical symbols by en-dashes; for example, ``NiMn'' indicates the 
intermetallic compound Ni$_{0.5}$Mn$_{0.5}$ whereas 
``Ni--Mn'' indicates an alloy of some composition 
Ni$_{x}$Mn$_{1-x}$.

\begin{figure}[htb!]
    \begin{subfigure}[b]{0.48\textwidth}
	\centerline{\includegraphics[width=\columnwidth]{energy/completo_dedup_4.png}}
    \caption{Dedup.}
	\label{fig:en_eq_dedup}
    \end{subfigure}
    %
    \begin{subfigure}[b]{0.48\textwidth}
    	\centerline{\includegraphics[width=\columnwidth]{energy/completo_rtview_4.png}}
        \caption{Raytrace.}
    	\label{fig:en_eq_rtview}
    \end{subfigure}
    %
    \begin{subfigure}[b]{0.48\textwidth}
    	\centerline{\includegraphics[width=\columnwidth]{energy/completo_swaptions_1.png}}
        \caption{Swaptions.}
    	\label{fig:en_eq_swaptions}
    \end{subfigure}
    \caption{Example fit for a specific input size: Dedup (a), Raytrace (b) and Swaptions (c). “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
    }
\end{figure}

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt){fig1.png}
{Magnetization as a function of applied field.
It is good practice to explain the significance of the figure in the caption.\label{fig1}}

Be aware of the different meanings of the homophones ``affect'' (usually a 
verb) and ``effect'' (usually a noun), ``complement'' and ``compliment,'' 
``discreet'' and ``discrete,'' ``principal'' (e.g., ``principal 
investigator'') and ``principle'' (e.g., ``principle of measurement''). Do 
not confuse ``imply'' and ``infer.''

Prefixes such as ``non,'' ``sub,'' ``micro,'' ``multi,'' and ``ultra'' are 
not independent words; they should be joined to the words they modify, 
usually without a hyphen. There is no period after the ``et'' in the Latin 
abbreviation ``\emph{et al.}'' (it is also italicized). The abbreviation ``i.e.,'' means 
``that is,'' and the abbreviation ``e.g.,'' means ``for example'' (these 
abbreviations are not italicized).

A general IEEE styleguide is available at \underline{http://www.ieee.org/}\break\underline{authortools}.

\section{Guidelines for Graphics Preparation and Submission}
\label{sec:guidelines}

\subsection{Types of Graphics}
The following list outlines the different types of graphics published in 
IEEE journals. They are categorized based on their construction, and use of 
color/shades of gray:

\subsubsection{Color/Grayscale figures}
{Figures that are meant to appear in color, or shades of black/gray. Such 
figures may include photographs, illustrations, multicolor graphs, and 
flowcharts.}

\subsubsection{Line Art figures}
{Figures that are composed of only black lines and shapes. These figures 
should have no shades or half-tones of gray, only black and white.}

\subsubsection{Author photos}
{Head and shoulders shots of authors that appear at the end of our papers. }

\subsubsection{Tables}
{Data charts which are typically black and white, but sometimes include 
color.}

\begin{table}
\caption{Units for Magnetic Properties}
\label{table}
\setlength{\tabcolsep}{3pt}
\begin{tabular}{|p{25pt}|p{75pt}|p{115pt}|}
\hline
Symbol& 
Quantity& 
Conversion from Gaussian and \par CGS EMU to SI $^{\mathrm{a}}$ \\
\hline
$\Phi $& 
magnetic flux& 
1 Mx $\to  10^{-8}$ Wb $= 10^{-8}$ V$\cdot $s \\
$B$& 
magnetic flux density, \par magnetic induction& 
1 G $\to  10^{-4}$ T $= 10^{-4}$ Wb/m$^{2}$ \\
$H$& 
magnetic field strength& 
1 Oe $\to  10^{3}/(4\pi )$ A/m \\
$m$& 
magnetic moment& 
1 erg/G $=$ 1 emu \par $\to 10^{-3}$ A$\cdot $m$^{2} = 10^{-3}$ J/T \\
$M$& 
magnetization& 
1 erg/(G$\cdot $cm$^{3}) =$ 1 emu/cm$^{3}$ \par $\to 10^{3}$ A/m \\
4$\pi M$& 
magnetization& 
1 G $\to  10^{3}/(4\pi )$ A/m \\
$\sigma $& 
specific magnetization& 
1 erg/(G$\cdot $g) $=$ 1 emu/g $\to $ 1 A$\cdot $m$^{2}$/kg \\
$j$& 
magnetic dipole \par moment& 
1 erg/G $=$ 1 emu \par $\to 4\pi \times  10^{-10}$ Wb$\cdot $m \\
$J$& 
magnetic polarization& 
1 erg/(G$\cdot $cm$^{3}) =$ 1 emu/cm$^{3}$ \par $\to 4\pi \times  10^{-4}$ T \\
$\chi , \kappa $& 
susceptibility& 
1 $\to  4\pi $ \\
$\chi_{\rho }$& 
mass susceptibility& 
1 cm$^{3}$/g $\to  4\pi \times  10^{-3}$ m$^{3}$/kg \\
$\mu $& 
permeability& 
1 $\to  4\pi \times  10^{-7}$ H/m \par $= 4\pi \times  10^{-7}$ Wb/(A$\cdot $m) \\
$\mu_{r}$& 
relative permeability& 
$\mu \to \mu_{r}$ \\
$w, W$& 
energy density& 
1 erg/cm$^{3} \to  10^{-1}$ J/m$^{3}$ \\
$N, D$& 
demagnetizing factor& 
1 $\to  1/(4\pi )$ \\
\hline
\multicolumn{3}{p{251pt}}{Vertical lines are optional in tables. Statements that serve as captions for 
the entire table do not need footnote letters. }\\
\multicolumn{3}{p{251pt}}{$^{\mathrm{a}}$Gaussian units are the same as cg emu for magnetostatics; Mx 
$=$ maxwell, G $=$ gauss, Oe $=$ oersted; Wb $=$ weber, V $=$ volt, s $=$ 
second, T $=$ tesla, m $=$ meter, A $=$ ampere, J $=$ joule, kg $=$ 
kilogram, H $=$ henry.}
\end{tabular}
\label{tab1}
\end{table}

\subsection{Multipart figures}
Figures compiled of more than one sub-figure presented side-by-side, or 
stacked. If a multipart figure is made up of multiple figure
types (one part is lineart, and another is grayscale or color) the figure 
should meet the stricter guidelines.

\subsection{File Formats For Graphics}\label{formats}
Format and save your graphics using a suitable graphics processing program 
that will allow you to create the images as PostScript (PS), Encapsulated 
PostScript (.EPS), Tagged Image File Format (.TIFF), Portable Document 
Format (.PDF), Portable Network Graphics (.PNG), or Metapost (.MPS), sizes them, and adjusts 
the resolution settings. When 
submitting your final paper, your graphics should all be submitted 
individually in one of these formats along with the manuscript.

\subsection{Sizing of Graphics}
Most charts, graphs, and tables are one column wide (3.5 inches/88 
millimeters/21 picas) or page wide (7.16 inches/181 millimeters/43 
picas). The maximum depth a graphic can be is 8.5 inches (216 millimeters/54
picas). When choosing the depth of a graphic, please allow space for a 
caption. Figures can be sized between column and page widths if the author 
chooses, however it is recommended that figures are not sized less than 
column width unless when necessary.

There is currently one publication with column measurements that do not 
coincide with those listed above. Proceedings of the IEEE has a column 
measurement of 3.25 inches (82.5 millimeters/19.5 picas).

The final printed size of author photographs is exactly
1 inch wide by 1.25 inches tall (25.4 millimeters$\,\times\,$31.75 millimeters/6 
picas$\,\times\,$7.5 picas). Author photos printed in editorials measure 1.59 inches 
wide by 2 inches tall (40 millimeters$\,\times\,$50 millimeters/9.5 picas$\,\times\,$12 
picas).

\subsection{Resolution }
The proper resolution of your figures will depend on the type of figure it 
is as defined in the ``Types of Figures'' section. Author photographs, 
color, and grayscale figures should be at least 300dpi. Line art, including 
tables should be a minimum of 600dpi.

\subsection{Vector Art}
In order to preserve the figures' integrity across multiple computer 
platforms, we accept files in the following formats: .EPS/.PDF/.PS. All 
fonts must be embedded or text converted to outlines in order to achieve the 
best-quality results.

\subsection{Color Space}
The term color space refers to the entire sum of colors that can be 
represented within the said medium. For our purposes, the three main color 
spaces are Grayscale, RGB (red/green/blue) and CMYK 
(cyan/magenta/yellow/black). RGB is generally used with on-screen graphics, 
whereas CMYK is used for printing purposes.

All color figures should be generated in RGB or CMYK color space. Grayscale 
images should be submitted in Grayscale color space. Line art may be 
provided in grayscale OR bitmap colorspace. Note that ``bitmap colorspace'' 
and ``bitmap file format'' are not the same thing. When bitmap color space 
is selected, .TIF/.TIFF/.PNG are the recommended file formats.

\subsection{Accepted Fonts Within Figures}
When preparing your graphics IEEE suggests that you use of one of the 
following Open Type fonts: Times New Roman, Helvetica, Arial, Cambria, and 
Symbol. If you are supplying EPS, PS, or PDF files all fonts must be 
embedded. Some fonts may only be native to your operating system; without 
the fonts embedded, parts of the graphic may be distorted or missing.

A safe option when finalizing your figures is to strip out the fonts before 
you save the files, creating ``outline'' type. This converts fonts to 
artwork what will appear uniformly on any screen.

\subsection{Using Labels Within Figures}

\subsubsection{Figure Axis labels }
Figure axis labels are often a source of confusion. Use words rather than 
symbols. As an example, write the quantity ``Magnetization,'' or 
``Magnetization M,'' not just ``M.'' Put units in parentheses. Do not label 
axes only with units. As in Fig. 1, for example, write ``Magnetization 
(A/m)'' or ``Magnetization (A$\cdot$m$^{-1}$),'' not just ``A/m.'' Do not label axes with a ratio of quantities and 
units. For example, write ``Temperature (K),'' not ``Temperature/K.''

Multipliers can be especially confusing. Write ``Magnetization (kA/m)'' or 
``Magnetization (10$^{3}$ A/m).'' Do not write ``Magnetization 
(A/m)$\,\times\,$1000'' because the reader would not know whether the top 
axis label in Fig. 1 meant 16000 A/m or 0.016 A/m. Figure labels should be 
legible, approximately 8 to 10 point type.

\subsubsection{Subfigure Labels in Multipart Figures and Tables}
Multipart figures should be combined and labeled before final submission. 
Labels should appear centered below each subfigure in 8 point Times New 
Roman font in the format of (a) (b) (c).

\subsection{File Naming}
Figures (line artwork or photographs) should be named starting with the 
first 5 letters of the author's last name. The next characters in the 
filename should be the number that represents the sequential 
location of this image in your article. For example, in author 
``Anderson's'' paper, the first three figures would be named ander1.tif, 
ander2.tif, and ander3.ps.

Tables should contain only the body of the table (not the caption) and 
should be named similarly to figures, except that `.t' is inserted 
in-between the author's name and the table number. For example, author 
Anderson's first three tables would be named ander.t1.tif, ander.t2.ps, 
ander.t3.eps.

Author photographs should be named using the first five characters of the 
pictured author's last name. For example, four author photographs for a 
paper may be named: oppen.ps, moshc.tif, chen.eps, and duran.pdf.

If two authors or more have the same last name, their first initial(s) can 
be substituted for the fifth, fourth, third$\ldots$ letters of their surname 
until the degree where there is differentiation. For example, two authors 
Michael and Monica Oppenheimer's photos would be named oppmi.tif, and 
oppmo.eps.

\subsection{Referencing a Figure or Table Within Your Paper}
When referencing your figures and tables within your paper, use the 
abbreviation ``Fig.'' even at the beginning of a sentence. Do not abbreviate 
``Table.'' Tables should be numbered with Roman Numerals.

\subsection{Checking Your Figures: The IEEE Graphics Analyzer}
The IEEE Graphics Analyzer enables authors to pre-screen their graphics for 
compliance with IEEE Access standards before submission. 
The online tool, located at
\underline{http://graphicsqc.ieee.org/}, allows authors to 
upload their graphics in order to check that each file is the correct file 
format, resolution, size and colorspace; that no fonts are missing or 
corrupt; that figures are not compiled in layers or have transparency, and 
that they are named according to the IEEE Access naming 
convention. At the end of this automated process, authors are provided with 
a detailed report on each graphic within the web applet, as well as by 
email.

For more information on using the Graphics Analyzer or any other graphics 
related topic, contact the IEEE Graphics Help Desk by e-mail at 
graphics@ieee.org.

\subsection{Submitting Your Graphics}
Because IEEE will do the final formatting of your paper,
you do not need to position figures and tables at the top and bottom of each 
column. In fact, all figures, figure captions, and tables can be placed at 
the end of your paper. In addition to, or even in lieu of submitting figures 
within your final manuscript, figures should be submitted individually, 
separate from the manuscript in one of the file formats listed above in 
Section \ref{formats}. Place figure captions below the figures; place table titles 
above the tables. Please do not include captions as part of the figures, or 
put them in ``text boxes'' linked to the figures. Also, do not place borders 
around the outside of your figures.

\subsection{Color Processing/Printing in IEEE Journals}
All IEEE Transactions, Journals, and Letters allow an author to publish 
color figures on IEEE Xplore\textregistered\ at no charge, and automatically 
convert them to grayscale for print versions. In most journals, figures and 
tables may alternatively be printed in color if an author chooses to do so. 
Please note that this service comes at an extra expense to the author. If 
you intend to have print color graphics, include a note with your final 
paper indicating which figures or tables you would like to be handled that 
way, and stating that you are willing to pay the additional fee.

\section{Conclusion}
A conclusion section is not required. Although a conclusion may review the 
main points of the paper, do not replicate the abstract as the conclusion. A 
conclusion might elaborate on the importance of the work or suggest 
applications and extensions.

\appendices

Appendixes, if needed, appear before the acknowledgment.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in American English is 
without an ``e'' after the ``g.'' Use the singular heading even if you have 
many acknowledgments. Avoid expressions such as ``One of us (S.B.A.) would 
like to thank $\ldots$ .'' Instead, write ``F. A. Author thanks $\ldots$ .'' In most 
cases, sponsor and financial support acknowledgments are placed in the 
unnumbered footnote on the first page, not here.

\section*{References and Footnotes}

\subsection{References}
References need not be cited in text. When they are, they appear on the 
line, in square brackets, inside the punctuation. Multiple references are 
each numbered with separate brackets. When citing a section in a book, 
please give the relevant page numbers. In text, refer simply to the 
reference number. Do not use ``Ref.'' or ``reference'' except at the 
beginning of a sentence: ``Reference \cite{b3} shows $\ldots$ .'' Please do not use 
automatic endnotes in \emph{Word}, rather, type the reference list at the end of the 
paper using the ``References'' style.

Reference numbers are set flush left and form a column of their own, hanging 
out beyond the body of the reference. The reference numbers are on the line, 
enclosed in square brackets. In all references, the given name of the author 
or editor is abbreviated to the initial only and precedes the last name. Use 
them all; use \emph{et al.} only if names are not given. Use commas around Jr., 
Sr., and III in names. Abbreviate conference titles. When citing IEEE 
transactions, provide the issue number, page range, volume number, year, 
and/or month if available. When referencing a patent, provide the day and 
the month of issue, or application. References may not include all 
information; please obtain and include relevant information. Do not combine 
references. There must be only one reference with each number. If there is a 
URL included with the print reference, it can be included at the end of the 
reference.

Other than books, capitalize only the first word in a paper title, except 
for proper nouns and element symbols. For papers published in translation 
journals, please give the English citation first, followed by the original 
foreign-language citation See the end of this document for formats and 
examples of common references. For a complete discussion of references and 
their formats, see the IEEE style manual at
\underline{http://www.ieee.org/authortools}.

\subsection{Footnotes}
Number footnotes separately in superscript numbers.\footnote{It is recommended that footnotes be avoided (except for 
the unnumbered footnote with the receipt date on the first page). Instead, 
try to integrate the footnote information into the text.} Place the actual 
footnote at the bottom of the column in which it is cited; do not put 
footnotes in the reference list (endnotes). Use letters for table footnotes 
(see Table \ref{table}).

\section{Submitting Your Paper for Review}

\subsection{Final Stage}
When you submit your final version (after your paper has been accepted), 
print it in two-column format, including figures and tables. You must also 
send your final manuscript on a disk, via e-mail, or through a Web 
manuscript submission system as directed by the society contact. You may use 
\emph{Zip} for large files, or compress files using \emph{Compress, Pkzip, Stuffit,} or \emph{Gzip.}

Also, send a sheet of paper or PDF with complete contact information for all 
authors. Include full mailing addresses, telephone numbers, fax numbers, and 
e-mail addresses. This information will be used to send each author a 
complimentary copy of the journal in which the paper appears. In addition, 
designate one author as the ``corresponding author.'' This is the author to 
whom proofs of the paper will be sent. Proofs are sent to the corresponding 
author only.

\subsection{Review Stage Using ScholarOne\textregistered\ Manuscripts}
Contributions to the Transactions, Journals, and Letters may be submitted 
electronically on IEEE's on-line manuscript submission and peer-review 
system, ScholarOne\textregistered\ Manuscripts. You can get a listing of the 
publications that participate in ScholarOne at 
\underline{http://www.ieee.org/publications\_standards/publications/}\break\underline{authors/authors\_submission.html}.
First check if you have an existing account. If there is none, please create 
a new account. After logging in, go to your Author Center and click ``Submit 
First Draft of a New Manuscript.''

Along with other information, you will be asked to select the subject from a 
pull-down list. Depending on the journal, there are various steps to the 
submission process; you must complete all steps for a complete submission. 
At the end of each step you must click ``Save and Continue''; just uploading 
the paper is not sufficient. After the last step, you should see a 
confirmation that the submission is complete. You should also receive an 
e-mail confirmation. For inquiries regarding the submission of your paper on 
ScholarOne Manuscripts, please contact oprs-support@ieee.org or call +1 732 
465 5861.

ScholarOne Manuscripts will accept files for review in various formats. 
Please check the guidelines of the specific journal for which you plan to 
submit.

You will be asked to file an electronic copyright form immediately upon 
completing the submission process (authors are responsible for obtaining any 
security clearances). Failure to submit the electronic copyright could 
result in publishing delays later. You will also have the opportunity to 
designate your article as ``open access'' if you agree to pay the IEEE open 
access fee.

\subsection{Final Stage Using ScholarOne Manuscripts}
Upon acceptance, you will receive an email with specific instructions 
regarding the submission of your final files. To avoid any delays in 
publication, please be sure to follow these instructions. Most journals 
require that final submissions be uploaded through ScholarOne Manuscripts, 
although some may still accept final submissions via email. Final 
submissions should include source files of your accepted manuscript, high 
quality graphic files, and a formatted pdf file. If you have any questions 
regarding the final submission process, please contact the administrative 
contact for the journal.

In addition to this, upload a file with complete contact information for all 
authors. Include full mailing addresses, telephone numbers, fax numbers, and 
e-mail addresses. Designate the author who submitted the manuscript on 
ScholarOne Manuscripts as the ``corresponding author.'' This is the only 
author to whom proofs of the paper will be sent.

\subsection{Copyright Form}
Authors must submit an electronic IEEE Copyright Form (eCF) upon submitting 
their final manuscript files. You can access the eCF system through your 
manuscript submission system or through the Author Gateway. You are 
responsible for obtaining any necessary approvals and/or security 
clearances. For additional information on intellectual property rights, 
visit the IEEE Intellectual Property Rights department web page at 
\underline{http://www.ieee.org/publications\_standards/publications/}\break\underline{rights/index.html}.

\section{IEEE Publishing Policy}
The general IEEE policy requires that authors should only submit original 
work that has neither appeared elsewhere for publication, nor is under 
review for another refereed publication. The submitting author must disclose 
all prior publication(s) and current submissions when submitting a 
manuscript. Do not publish ``preliminary'' data or results. The submitting 
author is responsible for obtaining agreement of all coauthors and any 
consent required from employers or sponsors before submitting an article. 
The IEEE Access Department strongly discourages courtesy 
authorship; it is the obligation of the authors to cite only relevant prior 
work.

The IEEE Access Department does not publish conference 
records or proceedings, but can publish articles related to conferences that 
have undergone rigorous peer review. Minimally, two reviews are required for 
every article submitted for peer review.

\section{Publication Principles}
The two types of contents of that are published are; 1) peer-reviewed and 2) 
archival. The Access Department publishes scholarly 
articles of archival value as well as tutorial expositions and critical 
reviews of classical subjects and topics of current interest.

Authors should consider the following points:

\begin{enumerate}
\item Technical papers submitted for publication must advance the state of knowledge and must cite relevant prior work. 
\item The length of a submitted paper should be commensurate with the importance, or appropriate to the complexity, of the work. For example, an obvious extension of previously published work might not be appropriate for publication or might be adequately treated in just a few pages.
\item Authors must convince both peer reviewers and the editors of the scientific and technical merit of a paper; the standards of proof are higher when extraordinary or unexpected results are reported. 
\item Because replication is required for scientific progress, papers submitted for publication must provide sufficient information to allow readers to perform similar experiments or calculations and 
use the reported results. Although not everything need be disclosed, a paper 
must contain new, useable, and fully described information. For example, a 
specimen's chemical composition need not be reported if the main purpose of 
a paper is to introduce a new measurement technique. Authors should expect 
to be challenged by reviewers if the results are not supported by adequate 
data and critical details.
\item Papers that describe ongoing work or announce the latest technical achievement, which are suitable for presentation at a professional conference, may not be appropriate for publication.
\end{enumerate}

\section{Reference Examples}

\begin{itemize}

\item \emph{Basic format for books:}\\
J. K. Author, ``Title of chapter in the book,'' in \emph{Title of His Published Book, x}th ed. City of Publisher, (only U.S. State), Country: Abbrev. of Publisher, year, ch. $x$, sec. $x$, pp. \emph{xxx--xxx.}\\
See \cite{b1,b2}.

\item \emph{Basic format for periodicals:}\\
J. K. Author, ``Name of paper,'' \emph{Abbrev. Title of Periodical}, vol. \emph{x, no}. $x, $pp\emph{. xxx--xxx, }Abbrev. Month, year, DOI. 10.1109.\emph{XXX}.123456.\\
See \cite{b3}--\cite{b5}.

\item \emph{Basic format for reports:}\\
J. K. Author, ``Title of report,'' Abbrev. Name of Co., City of Co., Abbrev. State, Country, Rep. \emph{xxx}, year.\\
See \cite{b6,b7}.

\item \emph{Basic format for handbooks:}\\
\emph{Name of Manual/Handbook, x} ed., Abbrev. Name of Co., City of Co., Abbrev. State, Country, year, pp. \emph{xxx--xxx.}\\
See \cite{b8,b9}.

\item \emph{Basic format for books (when available online):}\\
J. K. Author, ``Title of chapter in the book,'' in \emph{Title of
Published Book}, $x$th ed. City of Publisher, State, Country: Abbrev.
of Publisher, year, ch. $x$, sec. $x$, pp. \emph{xxx--xxx}. [Online].
Available: \underline{http://www.web.com}\\
See \cite{b10}--\cite{b13}.

\item \emph{Basic format for journals (when available online):}\\
J. K. Author, ``Name of paper,'' \emph{Abbrev. Title of Periodical}, vol. $x$, no. $x$, pp. \emph{xxx--xxx}, Abbrev. Month, year. Accessed on: Month, Day, year, DOI: 10.1109.\emph{XXX}.123456, [Online].\\
See \cite{b14}--\cite{b16}.

\item \emph{Basic format for papers presented at conferences (when available online): }\\
J.K. Author. (year, month). Title. presented at abbrev. conference title. [Type of Medium]. Available: site/path/file\\
See \cite{b17}.

\item \emph{Basic format for reports and handbooks (when available online):}\\
J. K. Author. ``Title of report,'' Company. City, State, Country. Rep. no., (optional: vol./issue), Date. [Online] Available: site/path/file\\
See \cite{b18,b19}.

\item \emph{Basic format for patents (when available online):}\\
Name of the invention, by inventor's name. (year, month day). Patent Number [Type of medium]. Available: site/path/file\\
See \cite{b21}.

\item \emph{Basic format}\emph{for conference proceedings (published):}\\
J. K. Author, ``Title of paper,'' in \emph{Abbreviated Name of Conf.}, City of Conf., Abbrev. State (if given), Country, year, pp. \emph{xxxxxx.}\\
See \cite{b22}.

\item \emph{Example for papers presented at conferences (unpublished):}\\
See \cite{b23}.

\item \emph{Basic format for patents}$:$\\
J. K. Author, ``Title of patent,'' U.S. Patent \emph{x xxx xxx}, Abbrev. Month, day, year.\\
See \cite{b24}.

\item \emph{Basic format for theses (M.S.) and dissertations (Ph.D.):}
\begin{enumerate}
\item J. K. Author, ``Title of thesis,'' M.S. thesis, Abbrev. Dept., Abbrev. Univ., City of Univ., Abbrev. State, year.
\item J. K. Author, ``Title of dissertation,'' Ph.D. dissertation, Abbrev. Dept., Abbrev. Univ., City of Univ., Abbrev. State, year.
\end{enumerate}
See \cite{b25,b26}.

\item \emph{Basic format for the most common types of unpublished references:}
\begin{enumerate}
\item J. K. Author, private communication, Abbrev. Month, year.
\item J. K. Author, ``Title of paper,'' unpublished.
\item J. K. Author, ``Title of paper,'' to be published.
\end{enumerate}
See \cite{b27}--\cite{b29}.

\item \emph{Basic formats for standards:}
\begin{enumerate}
\item \emph{Title of Standard}, Standard number, date.
\item \emph{Title of Standard}, Standard number, Corporate author, location, date.
\end{enumerate}
See \cite{b30,b31}.

\item \emph{Article number in~reference examples:}\\
See \cite{b32,b33}.

\item \emph{Example when using et al.:}\\
See \cite{b34}.

\end{itemize}

\begin{thebibliography}{00}

\bibitem{b1} G. O. Young, ``Synthetic structure of industrial plastics,'' in \emph{Plastics,} 2\textsuperscript{nd} ed., vol. 3, J. Peters, Ed. New York, NY, USA: McGraw-Hill, 1964, pp. 15--64.

\bibitem{b2} W.-K. Chen, \emph{Linear Networks and Systems.} Belmont, CA, USA: Wadsworth, 1993, pp. 123--135.

\bibitem{b3} J. U. Duncombe, ``Infrared navigation---Part I: An assessment of feasibility,'' \emph{IEEE Trans. Electron Devices}, vol. ED-11, no. 1, pp. 34--39, Jan. 1959, 10.1109/TED.2016.2628402.

\bibitem{b4} E. P. Wigner, ``Theory of traveling-wave optical laser,'' \emph{Phys. Rev}., vol. 134, pp. A635--A646, Dec. 1965.

\bibitem{b5} E. H. Miller, ``A note on reflector arrays,'' \emph{IEEE Trans. Antennas Propagat}., to be published.

\bibitem{b6} E. E. Reber, R. L. Michell, and C. J. Carter, ``Oxygen absorption in the earth's atmosphere,'' Aerospace Corp., Los Angeles, CA, USA, Tech. Rep. TR-0200 (4230-46)-3, Nov. 1988.

\bibitem{b7} J. H. Davis and J. R. Cogdell, ``Calibration program for the 16-foot antenna,'' Elect. Eng. Res. Lab., Univ. Texas, Austin, TX, USA, Tech. Memo. NGL-006-69-3, Nov. 15, 1987.

\bibitem{b8} \emph{Transmission Systems for Communications}, 3\textsuperscript{rd} ed., Western Electric Co., Winston-Salem, NC, USA, 1985, pp. 44--60.

\bibitem{b9} \emph{Motorola Semiconductor Data Manual}, Motorola Semiconductor Products Inc., Phoenix, AZ, USA, 1989.

\bibitem{b10} G. O. Young, ``Synthetic structure of industrial
plastics,'' in Plastics, vol. 3, Polymers of Hexadromicon, J. Peters,
Ed., 2\textsuperscript{nd} ed. New York, NY, USA: McGraw-Hill, 1964, pp. 15-64.
[Online]. Available:
\underline{http://www.bookref.com}.

\bibitem{b11} \emph{The Founders' Constitution}, Philip B. Kurland
and Ralph Lerner, eds., Chicago, IL, USA: Univ. Chicago Press, 1987.
[Online]. Available: \underline{http://press-pubs.uchicago.edu/founders/}

\bibitem{b12} The Terahertz Wave eBook. ZOmega Terahertz Corp., 2014.
[Online]. Available:
\underline{http://dl.z-thz.com/eBook/zomega\_ebook\_pdf\_1206\_sr.pdf}. Accessed on: May 19, 2014.

\bibitem{b13} Philip B. Kurland and Ralph Lerner, eds., \emph{The
Founders' Constitution.} Chicago, IL, USA: Univ. of Chicago Press,
1987, Accessed on: Feb. 28, 2010, [Online] Available:
\underline{http://press-pubs.uchicago.edu/founders/}

\bibitem{b14} J. S. Turner, ``New directions in communications,'' \emph{IEEE J. Sel. Areas Commun}., vol. 13, no. 1, pp. 11-23, Jan. 1995.

\bibitem{b15} W. P. Risk, G. S. Kino, and H. J. Shaw, ``Fiber-optic frequency shifter using a surface acoustic wave incident at an oblique angle,'' \emph{Opt. Lett.}, vol. 11, no. 2, pp. 115--117, Feb. 1986.

\bibitem{b16} P. Kopyt \emph{et al., ``}Electric properties of graphene-based conductive layers from DC up to terahertz range,'' \emph{IEEE THz Sci. Technol.,} to be published. DOI: 10.1109/TTHZ.2016.2544142.

\bibitem{b17} PROCESS Corporation, Boston, MA, USA. Intranets:
Internet technologies deployed behind the firewall for corporate
productivity. Presented at INET96 Annual Meeting. [Online].
Available: \underline{http://home.process.com/Intranets/wp2.htp}

\bibitem{b18} R. J. Hijmans and J. van Etten, ``Raster: Geographic analysis and modeling with raster data,'' R Package Version 2.0-12, Jan. 12, 2012. [Online]. Available: \underline {http://CRAN.R-project.org/package=raster}

\bibitem{b19} Teralyzer. Lytera UG, Kirchhain, Germany [Online].
Available:
\underline{http://www.lytera.de/Terahertz\_THz\_Spectroscopy.php?id=home}, Accessed on: Jun. 5, 2014

\bibitem{b20} U.S. House. 102\textsuperscript{nd} Congress, 1\textsuperscript{st} Session. (1991, Jan. 11). \emph{H. Con. Res. 1, Sense of the Congress on Approval of}  \emph{Military Action}. [Online]. Available: LEXIS Library: GENFED File: BILLS

\bibitem{b21} Musical toothbrush with mirror, by L.M.R. Brooks. (1992, May 19). Patent D 326 189 [Online]. Available: NEXIS Library: LEXPAT File: DES

\bibitem{b22} D. B. Payne and J. R. Stern, ``Wavelength-switched pas- sively coupled single-mode optical network,'' in \emph{Proc. IOOC-ECOC,} Boston, MA, USA, 1985, pp. 585--590.

\bibitem{b23} D. Ebehard and E. Voges, ``Digital single sideband detection for interferometric sensors,'' presented at the \emph{2\textsuperscript{nd} Int. Conf. Optical Fiber Sensors,} Stuttgart, Germany, Jan. 2-5, 1984.

\bibitem{b24} G. Brandli and M. Dick, ``Alternating current fed power supply,'' U.S. Patent 4 084 217, Nov. 4, 1978.

\bibitem{b25} J. O. Williams, ``Narrow-band analyzer,'' Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, USA, 1993.

\bibitem{b26} N. Kawasaki, ``Parametric study of thermal and chemical nonequilibrium nozzle flow,'' M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.

\bibitem{b27} A. Harrison, private communication, May 1995.

\bibitem{b28} B. Smith, ``An approach to graphs of linear forms,'' unpublished.

\bibitem{b29} A. Brahms, ``Representation error for real numbers in binary computer arithmetic,'' IEEE Computer Group Repository, Paper R-67-85.

\bibitem{b30} IEEE Criteria for Class IE Electric Systems, IEEE Standard 308, 1969.

\bibitem{b31} Letter Symbols for Quantities, ANSI Standard Y10.5-1968.

\bibitem{b32} R. Fardel, M. Nagel, F. Nuesch, T. Lippert, and A. Wokaun, ``Fabrication of organic light emitting diode pixels by laser-assisted forward transfer,'' \emph{Appl. Phys. Lett.}, vol. 91, no. 6, Aug. 2007, Art. no. 061103.~

\bibitem{b33} J. Zhang and N. Tansu, ``Optical gain and laser characteristics of InGaN quantum wells on ternary InGaN substrates,'' \emph{IEEE Photon. J.}, vol. 5, no. 2, Apr. 2013, Art. no. 2600111

\bibitem{b34} S. Azodolmolky~\emph{et al.}, Experimental demonstration of an impairment aware network planning and operation tool for transparent/translucent optical networks,''~\emph{J. Lightw. Technol.}, vol. 29, no. 4, pp. 439--448, Sep. 2011.

\end{thebibliography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a1.png}}]{First A. Author} (M'76--SM'81--F'87) and all authors may include 
biographies. Biographies are often not included in conference-related
papers. This author became a Member (M) of IEEE in 1976, a Senior
Member (SM) in 1981, and a Fellow (F) in 1987. The first paragraph may
contain a place and/or date of birth (list place, then date). Next,
the author's educational background is listed. The degrees should be
listed with type of degree in what field, which institution, city,
state, and country, and year the degree was earned. The author's major
field of study should be lower-cased.

The second paragraph uses the pronoun of the person (he or she) and not the 
author's last name. It lists military and work experience, including summer 
and fellowship jobs. Job titles are capitalized. The current job must have a 
location; previous positions may be listed 
without one. Information concerning previous publications may be included. 
Try not to list more than three books or published articles. The format for 
listing publishers of a book within the biography is: title of book 
(publisher name, year) similar to a reference. Current and previous research 
interests end the paragraph. The third paragraph begins with the author's 
title and last name (e.g., Dr.\ Smith, Prof.\ Jones, Mr.\ Kajor, Ms.\ Hunter). 
List any memberships in professional societies other than the IEEE. Finally, 
list any awards and work for IEEE committees and publications. If a 
photograph is provided, it should be of good quality, and 
professional-looking. Following are two examples of an author's biography.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a2.png}}]{Second B. Author} was born in Greenwich Village, New York, NY, USA in 
1977. He received the B.S. and M.S. degrees in aerospace engineering from 
the University of Virginia, Charlottesville, in 2001 and the Ph.D. degree in 
mechanical engineering from Drexel University, Philadelphia, PA, in 2008.

From 2001 to 2004, he was a Research Assistant with the Princeton Plasma 
Physics Laboratory. Since 2009, he has been an Assistant Professor with the 
Mechanical Engineering Department, Texas A{\&}M University, College Station. 
He is the author of three books, more than 150 articles, and more than 70 
inventions. His research interests include high-pressure and high-density 
nonthermal plasma discharge processes and applications, microscale plasma 
discharges, discharges in liquids, spectroscopic diagnostics, plasma 
propulsion, and innovation plasma applications. He is an Associate Editor of 
the journal \emph{Earth, Moon, Planets}, and holds two patents.

Dr. Author was a recipient of the International Association of Geomagnetism 
and Aeronomy Young Scientist Award for Excellence in 2008, and the IEEE 
Electromagnetic Compatibility Society Best Symposium Paper Award in 2011. 
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a3.png}}]{Third C. Author, Jr.} (M'87) received the B.S. degree in mechanical 
engineering from National Chung Cheng University, Chiayi, Taiwan, in 2004 
and the M.S. degree in mechanical engineering from National Tsing Hua 
University, Hsinchu, Taiwan, in 2006. He is currently pursuing the Ph.D. 
degree in mechanical engineering at Texas A{\&}M University, College 
Station, TX, USA.

From 2008 to 2009, he was a Research Assistant with the Institute of 
Physics, Academia Sinica, Tapei, Taiwan. His research interest includes the 
development of surface processing and biological/medical treatment 
techniques using nonthermal atmospheric pressure plasmas, fundamental study 
of plasma sources, and fabrication of micro- or nanostructured surfaces.

Mr. Author's awards and honors include the Frew Fellowship (Australian 
Academy of Science), the I. I. Rabi Prize (APS), the European Frequency and 
Time Forum Award, the Carl Zeiss Research Award, the William F. Meggers 
Award and the Adolph Lomb Medal (OSA).
\end{IEEEbiography}

\EOD

\end{document}

### 2021-03-25 thesis/experiments/experiments.tex
first version

@@
 \centering
-\includegraphics[height= 7.5cm]{intro/Imagens/c_states}
+\includegraphics[height= 7.5cm]{intro/figures/c_states}
 \caption {C states}
@@
 \centering
-\includegraphics[height= 7.5cm]{intro/Imagens/p_states}
+\includegraphics[height= 7.5cm]{intro/figures/p_states}
 \caption {States P}
@@
 \centering
-\includegraphics [height = 7.5cm] {intro/Imagens/IPMI-Block-Diagram.png}
+\includegraphics [height = 7.5cm] {intro/figures/IPMI-Block-Diagram.png}
 \caption {IMPI diagram} {Image taken from \protect \url {https://pt.wikipedia.org/wiki/Intelligent_Platform_Management_Interface}, the main components of IPMI and how they communicate are shown}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/Blackschole.png}
+	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/blackscholes.png}
 	\caption{Caption}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/Canneal.png}
+	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/canneal.png}
 	\caption{Caption}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/Blackschole.png}
+	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/blackscholes.png}
 	\caption{Caption}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/Canneal.png}
+	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/canneal.png}
 	\caption{Caption}
@@
 \begin{figure}[htb!]
-	\includegraphics[width=\columnwidth]{models/figures/mpe_svr_eq.png}
+	\includegraphics[width=\columnwidth]{models/figures/mae_svr_eq.png}
 	\caption{Comparison of the Mean Percentage Error between the proposed model and SVR. "Model mean" and "SVR mean" are the average of all MPE values for all applications.
@@
 	\begin{subfigure}[b]{0.44\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/overall.png}}
+		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/overall_energy_10pts.png}}
 		\caption{Average energy spent on all applications during model creation. The two curves are identical because the same data were used to adjust the SVR and the model.}
@@
 	\begin{subfigure}[b]{0.44\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/overall_mpe.png}}
+		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/overall_mpe_10pts.png}}
 		\caption{

### 2021-03-25 thesis/model_analysis/model_analysis.tex
first version

\section{Model Analysis}

\subsection{Pareto points}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_high.png}
	\caption{Pareto static energy}
	\label{fig:pareto_static_h}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_low.png}
	\caption{Pareto static energy}
	\label{fig:pareto_static_l}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_w_high.png}
	\caption{Pareto w energy}
	\label{fig:pareto_w_h}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_low.png}
	\caption{Pareto w energy}
	\label{fig:pareto_w_l}
\end{figure}

\subsection{Optimization under constraints}

\subsection{Gradient and Countorns}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn0.png}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn0_3d.png}
	\end{subfigure}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn3.png}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn3_3d.png}
	\end{subfigure}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/pleak0.png}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/pleak0_3d.png}
	\end{subfigure}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/pleak10.png}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/pleak10_3d.png}
	\end{subfigure}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic0.png}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic0_3d.png}
	\end{subfigure}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic3000.png}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic3000_3d.png}
	\end{subfigure}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/w0.png}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/w0_3d.png}
	\end{subfigure}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/w1.png}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{models/figures/analisys/w1_3d.png}
	\end{subfigure}
\end{figure}

### 2021-03-25 thesis/fingerprint/fingerprint.tex
rename figures

@@
     \centering
-    \includegraphics[width=\textwidth]{fingerprint/figures/PERF_COUNT_HW_INSTRUCTIONS.png}
+    \includegraphics[width=\textwidth]{fingerprint/figures/workflow.png}
     \caption{Multiple executions of the Polybench 2mm program with different input sizes}
@@
     \centering
-    \includegraphics[width=\textwidth]{fingerprint/figures/PERF_COUNT_HW_INSTRUCTIONS_1.png}
+    \includegraphics[width=\textwidth]{fingerprint/figures/workflow_1.png}
     \caption{Single instance representation}
@@
     \centering
-    \includegraphics[width=\textwidth]{fingerprint/figures/PERF_COUNT_HW_INSTRUCTIONS_2.png}
+    \includegraphics[width=\textwidth]{fingerprint/figures/workflow_2.png}
     \caption{Interpolation}
@@
     \centering
-    \includegraphics[width=\textwidth]{fingerprint/figures/PERF_COUNT_HW_INSTRUCTIONS_3.png}
+    \includegraphics[width=\textwidth]{fingerprint/figures/workflow_3.png}
     \caption{Filtering}

### 2021-04-27 thesis/phases/phases.tex
phases section

\section{The impact of best-choice execution time slots partition of HPC applications on energy-saving strategies}

The energy efficiency of HPC can be achieved through dynamic workload planning. Dynamic scaling of frequency and voltage DVFS is the main energy efficiency technique to save power. The scaling algorithms can operate dynamically depending on the CPU workload. However, the required IT resources and workload vary throughout the applications' execution. The division of this workload directly impacts the resulting performance and the complexity of dynamic scaling algorithms. This work analyses the impact on the energy efficiency of different dynamic workload partitioning techniques, based on time division of allocated resources and DVFS. We evaluate the trade-off between the best time intervals partition and the overhead impact of dynamic scheduling techniques. We propose the optimal partition of application executions in time intervals providing the most energy efficient execution parameters, workload, frequency and number of cores, for each time interval.

\subsection{Introduction}
Talk about the importance of energy efficiency.

\subsection{Related work}
Some related work

\subsection{Methodology}

\subsubsection{Optimal energy configuration}
Splitting a program into phases, or time intervals with a selected set of execution parameters, has an impact on energy consumption. To find the ideal phase division, we need to compute the optimal energy configuration for a given number and duration of phases. For that, many different approaches could be applied, each having different trade-offs.

Using models ref [x], or a static analysis of the application, allows obtaining a very fast but less precise estimate. On the other hand, using the application and evaluating its behavior in each possible context would be a more reliable but too expensive method. Indeed, since the number of combinations grows exponentially, this is a very time consuming approach. Instead, our analysis approach will provide both, speed and precision. We will simply run the target application and collect performance data from every possible execution context. By context, we can consider, for instance, the frequency and/or number of cores. This data will then be used to analyse the behavior of phases in all possible combinations. For this, we assume that the number of instructions of the executed program is the same regardless of frequency and number of threads.

\begin{equation}
    I=c_{k1}f_1T_1=c_{k2}f_2T_2
\end{equation}

where $I$ is the number of instructions, $c_{k1}$ the rate of instructions per cycle, $f$ the frequency of the processor and $T$ the execution time. All the required parameters can be collected by simply running the application with the available frequencies.

From this equation we can split the execution of the program into several different phases $i$ as follows:
% \begin{equation}
%     I=xI+(1-x)I=xc_{k1}f_1+(1-x)c_{k2}f_2
% \end{equation}
\begin{equation}
    I=\sum_{i}{a_iI} %=\sum_i{a_ic_{ki}f_i}
\end{equation}

with $a_i$ as the percentages of instructions executed with frequency $f_i$, we can compute any combination of phases. For convenience, we write this as a function of the time:
% \begin{equation}
% T=\frac{I}{c_{k1}f_1}
% \end{equation}
\begin{equation}
I=Tc_{k}f
\end{equation}
\begin{equation}
    I=\sum_{i}{a_iI}=\sum_i{a_iTc_{ki}f_i}
\end{equation}
% \begin{equation}
% I=xI+(1-x)I=xTc_{k1}f_1+(1-x)Tc_{k2}f_2
% \end{equation}

the application being divided into phases of different duration, $a_i$  becomes the duration of the phase as a percentage of the total execution time. This general equation does not limit the number and duration of phases, and several phases can have the same frequency.

In the event that we decide to include the number of cores as a context parameter, the equation becomes more complex. Indeed, the number of cores can completely change the application's behavior but, depending on the type of application, it's still possible to find similarities. Many HPC applications use data parallelism, and the Amdahl's law can model most of it. For this kind of applications, we can estimate that:
\begin{equation}
T=\frac{T_s}{S}
\end{equation}
\begin{equation}
T_s=\frac{I}{c_kf}
\end{equation}
\begin{equation}
S=\frac{1}{1-w+\frac{w}{p}}
\end{equation}
\begin{equation}
T=\frac{I(1-w+\frac{w}{p})}{c_kf}=\frac{I}{c_kf}-\frac{Iw}{c_kf}+\frac{Iw}{c_kfp}
\end{equation}

Looking at a fixed frequency, we arrive at a very similar structure as before:
\begin{equation}
T=c_1I+\frac{c_2I}{p}
\end{equation}

As previously, we can use $I$ as the number of reference instructions, and compute $a$ as the percentage of single and multi-core instructions, or express it as a percentage of the total time T:
\begin{equation}
aT=c_1aI+\frac{c_2aI}{p}
\end{equation}

The extent of the hypotheses will later be validated, however, by taking advantage of both principles, we can now, based on real measurements, estimate any phase division and possible context combination.

% Here a need to show that the time is proportional to the number of executed instructions.
% WHY ?
% I am trying to say here that we can split the program using a percentage of the time because the time is proportional to the number of executed instructions. If we complete x% of the time, that will be x% of instructions. This implies that we can resume and continue the program from any point ensuring that we executed all instructions. If we have a previous execution for each configuration, we can compute any combination with minimal error.

\subsubsection{Data Collection}
Data, power samples, can be gathered by running the application in several possible context configurations. In a system with 32 cores that can run at 13 frequencies, we can have 416 (32*13) possible configurations. Power samples can be taken from dedicated sensors in the IPMI system [REF]. To avoid overload, a sample rate of 0.5 seconds is more than sufficient. Additionally, the behavior of an application can be affected by the workload, so it is also interesting to run applications with different inputs and assess their impact on the collected data [REF victor alex paper].

% We can consider different input sizes as an new application...

\subsubsection{The simulator}
To analyse the impact of configurations and identify optimal time-intervals, phases, we propose an integrator, which computes the energy on an interval for a given configuration, and a reference, a brute force test of all possible combinations. The integrator has order zero, and computers the energy between two timestamps, as shown in the figure \ref{fig:zero_order}, where the red and yellow parts area are the energy value.

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{phases/figures/zero_order.png}
    \caption{Power vs time intervals for a given application. The energy associated area is colored.}
    \label{fig:zero_order}
\end{figure}

Power and energy can be computed for a given time interval based on power samples. The pseudo-algorithm for this computation is described bellow:

\begin{lstlisting}
timeslice_energy(info, start, stop)
{
   power = info.power;
   time = info.time;

   t1 = start_time + (elapsed_time) * start;
   t2 = start_time + (elapsed_time) * stop;
   en = 0;

   // first time that is greater than t1
   for(i=0; i<nsamples; i++)
       if(time[i] >= t1)
           break;
   v1 = (i == 0)?0:i-1;
   // first time that is greater than t2
   for(i=v1; i<nsamples; i++)
       if(time[i] >= t2)
           break;
   v2 = (i == 0)?0:i-1;

   // interpolate and integrates the power samples
   if(v1 == v2)
       return (t2-t1)*power[v1];

   en = (time[v1+1]-t1)*power[v1];
   en += (t2-time[v2])*power[v2];
   // integrate samples in between
   for(i=v1+1; i<v2; i++)
       en += (time[i+1]-time[i])*power[i];
   return en;
}
\end{lstlisting}

The reference can simply be obtained by applying brute-force, which cycles through all configurations for all divisions and compute the minimal energy. Other more effective, but less accurate, methods can be applied.

\begin{lstlisting}
bruteforce(data, phases)
{
   min_en = -1;
   total_en = 0;
   for (i = 0; i < nphases; i++)
   {
       min_en = -1;
       for (j = 0; j<data.nconfigs; j++)
       {
           en = timeslice_energy(data.info[j], 
                        phases[i], phases[i + 1]);
           if (min_en == -1 || en < min_en)
           {
               min_en = en;
               min_conf = data.config[j];
           }
       }
       total_en += min_en;
   }
   return total_en;
}
\end{lstlisting}

\subsubsection{Optimizing phases}
The main goal of this section is to identify the optimal number and duration of time intervals. Indeed, several short phases can have an impact on switching overhead as well as few long phases on the overall energy.

Using the objective function, brute force can identify the optimal energy on each phase and any minimization method can provide the best phase division. By computing in a range of a number of phases, we can find the one that brings the lowest energy value. A genetic algorithm GA can be used to analyse a range of, for instance, 3 to 100 phases. A GA for a population of 10K individuals with a 0.1 elitism will ensure that 10\% of the best individuals will survive and 0.1 mutations will avoid local minimums within a maximum of 300 steps. A reproduction function will combine the half-phases of one individual with another, and the mutation will randomly change a division.

\subsection{Results}

% Normalization
% Show that number of instructions is the same...

In order to validate the approach, HPC benchmarks where chosen, 13 applications from PARSEC 3.0, HPC, and Openmc.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{fingerprint/figures/energy_per_phase.png}
    \caption{Relative energy vs number of phases using 13 applications from PARSEC 3.0, HPC, and Openmc benchmarks. Energy compared to the single-phase optimal configuration.}
    \label{fig:relative_energy}
\end{figure}

This evaluation shows very interesting results, as shown in the figure \ref{fig:relative_energy}. Indeed, for all cases, an optimal single-phase configuration gives the highest energy draw. On the contrary, it is not necessary to split the application into more than 35 phases. On average, the energy consumption stood still after 35 phases.

The results attend our expectation that the optimal with n divisions will always be better or equal than n-1 divisions. This can be explained by the fact that after we find the ideal frequency and core curves, increasing the number of phases will not change the control signals. Therefore the energy will remain the same, and there is no reason to increase the number of phases beyond that actually.

It is important to note that the additional energy savings that theoretically may be achieved when running with a greater number of phases will have a negative impact on performance and context switching. Moreover, the maximal energy saving compared to the single-phase case was just 2.6\%. This may be due to the type of applications, where a fine-grain fast control will not save more energy. On the contrary, this switching will cause to consume a lot more energy, and this is likely to happen with online DVFS algorithms, very sensitive to fast workload variations [REF].

We evaluated separately the impact on energy when varying the number of cores. Energy-cores figures doesn't change fundamentally varying the number of phases. Indeed, bellow we can see some examples with optimal energy cores selections for 60 and 70 phases (figure \ref{fig:cores_control}). Note the very small differences that will not change significantly the energy figures.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_3_cores_signals_cmp.png}
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_3_cores_signals_cmp.png}
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_cores_signals_cmp.png}  
\caption{Energy and number of cores with 60 and 70 phases. (a) bla bla (b) bla bla (c) blabla.}
    \label{fig:cores_control}
\end{figure}

We observed a similar behavior just considering frequency control energy optimisation (figure \ref{fig:freq_control}).

\begin{figure}[h]
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_3_freq_signals_cmp.png}
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_3_freq_signals_cmp.png}
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_freq_signals_cmp.png}
\caption{Energy and frequency with 60 and 70 phases. (a) bla bla (b) bla bla (c) blabla.}
    \label{fig:freq_control}
\end{figure}

The images below show the phases for all applications:
\begin{figure}[h]
\includegraphics[width=\columnwidth]{phases/figures/phase_division_35.png}
\end{figure}
In most applications, we can see a pattern of 3 phases wich for this applications are generally, and setup phase where it loads data from the disk, a computation phase where the data is processed, and a finalization phase where data is written back to the disk or display to the user.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{phases/figures/phase_division_3.png}
\end{figure}
Plotting for 3 phases, we can observe that better.

\subsection{Conclusion and future work}
The main conclusion of this work was that in HPC enviroment we don't need a lot of phases to optimize the energy consumption given that for one phase we can compute the optimal energy.

The final idea is to use the information of the best phase to find a performance metric that can provide us this division, in future work we intend to compute this metrics using combination of hardware performance counters enabling optimal phase division in real time. This can improve current DVFS algorithms giving them an workload metric that can improve the energy savings.

### 2021-07-08 thesis/intro/intro.tex
update thesis text

@@
 Power costs exceed in some cases the cost of purchasing hardware \cite{Rivoire2007ModelsOptimizations}. 
-Nowadays, the leading petaflop supercomputers consume a range of 1–18 MW of electrical power, with 1.5 MW on average, which can be easily translated into millions of dollars per year in electricity bills \cite{Group2012HandbookSahni}.
+Nowadays, the leading petaflop supercomputers consume a range of 1-18 MW of electrical power, with 1.5 MW on average, which can be easily translated into millions of dollars per year in electricity bills \cite{Group2012HandbookSahni}.
 Furthermore, the energy costs of powering a typical data center doubles every five years \cite{Buyya2013MasteringProgramming}.
-Therefore, with such steep increase in electricity use and rising electricity costs, electricity bills have become a major expense for today’s data centers \cite{Poess2008EnergyResults}, \cite{Gao2013QualityCenters}.
+Therefore, with such steep increase in electricity use and rising electricity costs, electricity bills have become a major expense for today's data centers \cite{Poess2008EnergyResults}, \cite{Gao2013QualityCenters}.
 Due to these reasons, data center energy efficiency is now considered a chief concern for data center operators, often ahead of the traditional considerations of availability and security.
@@
 
-The rest of this paper is divided as follows.
+%The rest of this paper is divided as follows.
 In Section \ref{sec:relatedwork}, a general review of models is provided, showing the difference between each approach and its applications.

### 2021-07-08 thesis/main.tex
first version
add phases in main
update thesis text

@@
 \usepackage{listings}
+
 %\usepackage[caption=false,font=footnotesize]{subfig}
 %\usepackage[export]{adjustbox}
+
+\lstset { %
+	language=C++,
+	backgroundcolor=\color{black!5}, % set backgroundcolor
+	basicstyle=\footnotesize,% basic font setting
+}
 
@@
 
-\usepackage{suthesis-2e}
-\dept{Computer Science}
+%\usepackage{suthesis-2e}
+%\dept{Computer Science}
+
+\usepackage[fpms]{umons-coverpage}
+\umonsAuthor{Vitor Ramos Gomes \textsc{SILVA}}
+\umonsTitle{Energy aware high-performance computing}
+\umonsDocumentType{A DISSERTATION SUBMITTED TO THE ELECTRONICS AND MICROELECTRONICS DEPARTMENT OF MONS UNIVERSITY IN PARTIAL FULFILLMENT OF THE REQUIREMENTS FOR THE DEGREE OF DOCTOR IN ENGINEERING SCIENCES AND TECHNOLOGY}
+%\umonsSupervisor{Supervisor\\ Carlos Alberto \textsc{VALDERRAMA}\\
+%				 Co-supervisor\\ Pierre \textsc{MANEBACK}}
+\umonsSupervisor{
+ 		Thesis committee:\\
+ 		Supervisor: Prof. Carlos Alberto \textsc{VALDERRAMA}\\
+ 		Co-supervisor: Prof. Pierre \textsc{MANEBACK}\\
+ 		Prof. Thierry \textsc{DUTOIT}\\
+ 		Prof. Sidi \textsc{MAHMOUDI}\\
+ 		Prof. Samuel \textsc{XAVIER-DE-SOUZA}
+}
+\umonsDate{April 2021}
 
 \begin{document}
-\title{Energy aware HPC}
-\author{Vitor Ramos Gomes da Silva}
-\principaladviser{Carlos Valderrama}
-\coprincipaladviser{Pierre MANNEBACK}
-\firstreader{Thierry DUTOIT}
-\secondreader{Sidi MAHMOUDI}
-\thirdreader{Samuel Xavier de Souza} %if needed
+%\title{Energy aware HPC}
+%\author{Vitor Ramos Gomes da Silva}
+%\principaladviser{Carlos Valderrama}
+%\coprincipaladviser{Pierre MANNEBACK}
+%\firstreader{Thierry DUTOIT}
+%\secondreader{Sidi MAHMOUDI}
+%\thirdreader{Samuel Xavier de Souza} %if needed
+% 
+%\beforepreface
+%\prefacesection{Preface}
+%%Second Phd Committee
+%\prefacesection{Acknowledgments}
+%%I would like to thank...
+%\afterpreface
 
-\beforepreface
-\prefacesection{Preface}
-Second Phd Committee
-%\prefacesection{Acknowledgments}
-%I would like to thank...
-\afterpreface
+\umonsCoverPage
 
-\chapter{Introduction}
+%\section*{Acknowledgments}
+%\vspace{1cm}
+%\begin{center}
+%	...
+%\end{center}
+%\newpage
+
+\section*{Abstract}
 Energy consumption is a key to enable exascale High-performance Computing (HPC). Energy-optimized hardware and software combinations still could be inefficient if software operates poorly. 
@@
 Since in HPC, applications can run on multiple cores and associated frequencies, both of which can be controlled for most systems, the novelty of the model includes such parameters as well as the workload, two application parameters and three system parameters.
+
+\vspace{2cm}
+\textbf{Keywords :} Energy Model; Dynamic Frequency and Voltage Scaling; Dynamic Power Management; High Performance Computing
+\newpage
+
+\tableofcontents
+\newpage
+\listoffigures
+\newpage
+\listoftables
+\newpage
+
+\chapter{Introduction}
+
 \include{intro/intro}
@@
 \include{models/models}
-\include{ch5/ch5}
+\include{model_analysis/model_analysis}
 
@@
 This chapter talks about the different types of optimizers.
-\include{ch7/ch7}
+\include{optmizers/optmizers}
+\include{phases/phases}
 

### 2021-07-08 thesis/phases/phases.tex
add phases in main
update thesis text

@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
 \includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_3_cores_signals_cmp.png}
 \includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_3_cores_signals_cmp.png}
-\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_cores_signals_cmp.png}  
-\caption{Energy and number of cores with 60 and 70 phases. (a) bla bla (b) bla bla (c) blabla.}
+%\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_cores_signals_cmp.png}  
+\caption{Energy and number of cores with 37 and 99 phases.}
     \label{fig:cores_control}
@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
 \includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_3_freq_signals_cmp.png}
 \includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_3_freq_signals_cmp.png}
-\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_freq_signals_cmp.png}
-\caption{Energy and frequency with 60 and 70 phases. (a) bla bla (b) bla bla (c) blabla.}
+%\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_freq_signals_cmp.png}
+\caption{Energy and frequency with 37 and 99 phases.}
     \label{fig:freq_control}

### 2022-04-05 thesis/main.tex
umons template

--- 2021-03-23 thesis/main.tex
@@
-% Stanford University PhD thesis style -- modifications to the report style
-% This is unofficial so you should always double check against the
-% Registrar's office rules
-% See http://library.stanford.edu/research/bibliography-management/latex-and-bibtex
-% 
-% Example of use below
-% See the suthesis-2e.sty file for documentation
-%
-\documentclass{report}
+\documentclass[
+papersize=a4,
+pagelayout=default,
+fontname=latinmodern,
+fontsize=11pt,
+twoside,
+final,
+faculty=fpms
+]{umons-Thesis}
 
-\usepackage[breakable]{tcolorbox}
-\usepackage{parskip} % Stop auto-indenting (to mimic markdown behaviour)
+% Styles
+\headstyles{umons}
+\pagestyle{umons}
 
-\usepackage{iftex}
-\ifPDFTeX
-\usepackage[T1]{fontenc}
-\usepackage{mathpazo}
-\else
-\usepackage{fontspec}
-\fi
+\usepackage[main=english]{babel}
+\usepackage[autostyle=true]{csquotes}
+\usepackage[babel=true]{microtype}
+\usepackage{float}
+\usepackage{graphicx}
+\usepackage{subcaption}
+\usepackage{booktabs}
+\usepackage{mathtools}
+\usepackage{siunitx}
+\sisetup{%
+	binary-units = true,%	loads units for binary data (\bit, \byte)
+}
+\usepackage[%
+hidelinks,%				hides links (removing color and border)
+pdfusetitle,%			uses the content of \author{}... for metadata
+pdfdisplaydoctitle,%	displays document title instead of file name in title bar
+]{hyperref}
+\usepackage{cleveref}
+\usepackage{outlines}
+\usepackage{listings}
 
-% Basic figure setup, for now with no caption control since it's done
-% automatically by Pandoc (which extracts ![](path) syntax from Markdown).
-\usepackage{graphicx}
-% Maintain compatibility with old templates. Remove in nbconvert 6.0
-\let\Oldincludegraphics\includegraphics
-% Ensure that by default, figures have no caption (until we provide a
-% proper Figure object with a Caption API and a way to capture that
-% in the conversion process - todo).
-\usepackage{caption}
-\DeclareCaptionFormat{nocaption}{}
-\captionsetup{format=nocaption,aboveskip=0pt,belowskip=0pt}
-
-\usepackage[Export]{adjustbox} % Used to constrain images to a maximum size
-\adjustboxset{max size={0.9\linewidth}{0.9\paperheight}}
-\usepackage{float}
-\floatplacement{figure}{H} % forces figures to be placed at the correct location
-\usepackage{xcolor} % Allow colors to be defined
-\usepackage{enumerate} % Needed for markdown enumerations to work
-\usepackage{geometry} % Used to adjust the document margins
-\usepackage{amsmath} % Equations
-\usepackage{amssymb} % Equations
-\usepackage{textcomp} % defines textquotesingle
-% Hack from http://tex.stackexchange.com/a/47451/13684:
-\AtBeginDocument{%
-	\def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
+\lstset { %
+	language=C++,
+	backgroundcolor=\color{black!5}, % set backgroundcolor
+	basicstyle=\footnotesize,% basic font setting
 }
-\usepackage{upquote} % Upright quotes for verbatim code
-\usepackage{eurosym} % defines \euro
-\usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
-\usepackage{fancyvrb} % verbatim replacement that allows latex
-\usepackage{grffile} % extends the file name processing of package graphics 
-% to support a larger range
-\makeatletter % fix for grffile with XeLaTeX
-\def\Gread@@xetex#1{%
-	\IfFileExists{"\Gin@base".bb}%
-	{\Gread@eps{\Gin@base.bb}}%
-	{\Gread@@xetex@aux#1}%
-}
-\makeatother
-
-% The hyperref package gives us a pdf with properly built
-% internal navigation ('pdf bookmarks' for the table of contents,
-% internal cross-reference links, web links for URLs, etc.)
-\usepackage{hyperref}
-% The default LaTeX title has an obnoxious amount of whitespace. By default,
-% titling removes some of it. It also provides customization options.
-\usepackage{titling}
-\usepackage{longtable} % longtable support required by pandoc >1.10
-\usepackage{booktabs}  % table support for pandoc > 1.12.2
-\usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
-\usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
-% normalem makes italics be italics, not underlines
-\usepackage{mathrsfs}
-\usepackage{outlines}
-
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{nth}
-\usepackage{todonotes}
-\usepackage{url}
-\usepackage{xcolor}
-\usepackage{textcomp}
-\usepackage{float}
-\usepackage{subcaption}
-\usepackage{cleveref}
-\usepackage{listings}
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%\usepackage[export]{adjustbox}
 
 \newcommand*\rot{\rotatebox{90}}
-
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
 		T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\DeclareOldFontCommand{\rm}{\normalfont\rmfamily}{\mathrm}
 
-\usepackage{suthesis-2e}
-\dept{Computer Science}
+%%        %%
+% DOCUMENT %
+% ======== %
+
+% Metadata
+% --------
+
+\author{Vitor \textsc{Ramos Gomes da Silva}}
+
+\date{\today}
+
+\title{Title}
+%\title{My very long two-line PhD thesis title}
+%\title{My supra long PhD thesis title just to verify that nothing is broken because of a long title}
+
+\committeeMembers{%
+	Prof. Carlos Alberto \textsc{VALDERRAMA}, Supervisor\\
+	Prof. Pierre \textsc{MANEBACK}, Co-supervisor\\
+	Prof. Thierry \textsc{DUTOIT}, Chair\\
+	Prof. Sidi \textsc{MAHMOUDI}, Chair\\
+	Prof. Samuel \textsc{XAVIER-DE-SOUZA}, Chair 
+}
+
+% Text
+% ----
 
 \begin{document}
-\title{Energy aware HPC}
-\author{Vitor Ramos Gomes da Silva}
-\principaladviser{Carlos Valderrama}
-\coprincipaladviser{Pierre MANNEBACK}
-\firstreader{Thierry DUTOIT}
-\secondreader{Sidi MAHMOUDI}
-\thirdreader{Samuel Xavier de Souza} %if needed
 
-\beforepreface
-\prefacesection{Preface}
-Second Phd Committee
-%\prefacesection{Acknowledgments}
-%I would like to thank...
-\afterpreface
+	\umonsThesisTitlePage
 
-\chapter{Introduction}
+	\frontmatter
+
+	\begin{umonsThesisAbstract}
 		Energy consumption is a key to enable exascale High-performance Computing (HPC). Energy-optimized hardware and software combinations still could be inefficient if software operates poorly. 
@@
 		Since in HPC, applications can run on multiple cores and associated frequencies, both of which can be controlled for most systems, the novelty of the model includes such parameters as well as the workload, two application parameters and three system parameters.
+	\end{umonsThesisAbstract}
+
+	% \vspace{2cm}
+	% \textbf{Keywords :} Energy Model; Dynamic Frequency and Voltage Scaling; Dynamic Power Management; High Performance Computing
+
+	\tableofcontents*
+	\listoffigures*
+	\listoftables*
+
+	\mainmatter
+
+	\chapter{Introduction}
+
 	\include{intro/intro}
 
-%\chapter{Approch}
-%\include{ch1/ch1}
-
-\chapter{Models}
+	\chapter{Proposed Models}
 	In this Section, the proposed power, performance and energy models are derived and validated
 	\include{models/models}
-\include{ch5/ch5}
-
-\chapter{Optmizers}
-This chapter talks about the different types of optimizers.
-\include{ch7/ch7}
+	\include{phases/phases}
+	\include{optmizers/optmizers}
 
@@
 
-\chapter{Experiments}
+	\chapter{Results (model validation)}
 	This chapter will describe all the experiments that validate the models constraints.
@@
 
-%\appendix
-%\chapter{A Long Proof}
+	\chapter{Conclusions}
 
@@
 	\bibliography{references}
+
+	\backmatter
+
 \end{document}

### 2022-04-05 thesis/experiments/experiments.tex
add journals and restructured

--- 2021-03-23 thesis/experiments/experiments.tex
@@
-\section{Frequency control} \label {sec: controle_freq}
+\section{Frequency control} \label{sec:frequency_control}
 
@@
 After an idle time, the processor begins to activate C states, starting with C0 where it is still fully active, then moving to C1, where some features are disabled, up to Cn, where all possible features are disabled.
-The ACPI standard establishes the functionalities that can be disabled between level C1 and level C3, as seen in Table \ref {tab: states_c}.
+The ACPI standard establishes the functionalities that can be disabled between level C1 and level C3, as seen in Table \ref {tab:states_c}.
 The other levels are specific to each manufacturer.
 
-\begin {table} [H]
+\begin{table}[H]
 \centering
-\begin {tabular} {| l | l | l |}
+\begin{tabular}{| l | l | l |}
 \hline
@@
 C3 & Sleep & Disable cache coherence \\ \hline
-\end {tabular}
-\caption {C states}
-\label {tab: states_c}
-\end {table}
+\end{tabular}
+\caption{C states}
+\label{tab:states_c}
+\end{table}
 
@@
 In states P, there is a trade-off between performance and energy savings.
-Figure \ref {fig: p_state} best illustrates the change of states, in which we can see which parts of the circuits are deactivated in states C, the latency to return to the active state, power consumption and also shows the relationship of states P with often.
+\cref{fig:p_state} best illustrates the change of states, in which we can see which parts of the circuits are deactivated in states C, the latency to return to the active state, power consumption and also shows the relationship of states P with often.
 
@@
 \centering
-\includegraphics[height= 7.5cm]{intro/Imagens/c_states}
-\caption {C states}
-\end {subfigure}%
+\includegraphics[width=\columnwidth]{intro/figures/c_states}
+\caption{C states}
+\end{subfigure}%
 ~
@@
 \centering
-\includegraphics[height= 7.5cm]{intro/Imagens/p_states}
-\caption {States P}
-\end {subfigure}
+\includegraphics[width=\columnwidth]{intro/figures/p_states}
+\caption{States P}
+\end{subfigure}
 \caption {Illustration of states C and P} {Altered image from \protect \url {https://www.thomas-krenn.com/en/wiki/Processor_P-states_and_C-states}}
-\label {fig: p_state}
-\end {figure}
+\label{fig:p_state}
+\end{figure}
 
@@
 
-On Linux there are several frequency management options \cite {Brown2005}.
+On Linux there are several frequency management options \cite {Brown2005ACPILinux}.
 The main ones are acpi-cpufreq, Intel P-state, AMD powernow.
@@
 
-\begin {itemize}
+\begin{itemize}
 \item Performance - configured as often as possible
@@
 \item Conservative - similar to Ondemand but more smoothly, the frequency increase is continuous instead of jumping.
-\end {itemize}
+\end{itemize}
 
-\section {Power consumption monitoring} \label {sec: monitoring}
-The Running Average Power Limit (RAPL) and Intelligent Platform Management Interface (IPMI) interfaces were used to measure the power consumed. described in \cite{November2013}.
+\section{Power consumption monitoring} \label{sec:power_consumption_monitoring}
+The Running Average Power Limit (RAPL) and Intelligent Platform Management Interface (IPMI) interfaces were used to measure the power consumed. described in \cite{IPMI2013ConfigurationGuide}.
 
-\subsection {IPMI}
-IPMI \cite {November2013} is a set of specifications for autonomous subsystems that provides processor, firmware and operating system independent management and monitoring.
+\subsection{IPMI}
+IPMI \cite{IPMI2013ConfigurationGuide} is a set of specifications for autonomous subsystems that provides processor, firmware and operating system independent management and monitoring.
 The use of IPMI allows system administrators to avoid having to travel to the server location, which is often far away, to perform their tasks.
@@
 
-\begin {figure} [H]
+\begin{figure}[H]
 \centering
-\includegraphics [height = 7.5cm] {intro/Imagens/IPMI-Block-Diagram.png}
-\caption {IMPI diagram} {Image taken from \protect \url {https://pt.wikipedia.org/wiki/Intelligent_Platform_Management_Interface}, the main components of IPMI and how they communicate are shown}
-\label {fig: IPMI}
-\end {figure}
+\includegraphics[height = 7.5cm] {intro/figures/IPMI-Block-Diagram.png}
+\caption{IMPI diagram} {Image taken from \protect \url {https://pt.wikipedia.org/wiki/Intelligent_Platform_Management_Interface}, the main components of IPMI and how they communicate are shown}
+\label{fig:IPMI}
+\end{figure}
 
@@
 
-\subsection {RAPL}
+\subsection{RAPL}
 
-Modern Intel microprocessors, based on the SandyBridge architecture, include the RAPL \cite {Rotem2012, Hahnel2012, Hackenberg2015} interface designed to limit the use of energy on a chip while ensuring maximum performance.
+Modern Intel microprocessors, based on the SandyBridge architecture, include the RAPL \cite {Rotem2012Power-managementBridge, Hahnel2012RAPL, Hackenberg2015AnProcessor} interface designed to limit the use of energy on a chip while ensuring maximum performance.
 This interface supports energy measurement capabilities through an integrated circuit that estimates energy use based on a model driven by architectural event counters for all components.
@@
 
-\section{Verifying models hypothesis}
-
-\subsection{Frequency voltage ratio}
-
-\begin{table}[H]
-	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|}
-		\hline
-		Freq (GHz) & 2.2   & 2.1   & 2.0   & 1.9   & 1.8   & 1.7   & 1.6   & 1.5   & 1.4   & 1.3   & 1.2   \\ \hline
-		V          & 0.77  & 0.76  & 0.75  & 0.74  & 0.73  & 0.72  & 0.71  & 0.70  & 0.69  & 0.68  & 0.67  \\ \hline
-		aperf      & 2.199 & 2.099 & 2.000 & 1.899 & 1.799 & 1.699 & 1.599 & 1.500 & 1.397 & 1.297 & 1.200 \\ \hline
-	\end{tabular}
-\end{table}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\columnwidth]{experiments/figures/freq_volt_rel.png}
-	\caption{Frequency voltage relation}
-	\label{fig:freq_volt_rel}
-\end{figure}
-
-\subsection{Instructions executed}
-
-\begin{table}[H]
-	\caption{Cores variation}
-	\begin{tabular}{|c|c|c|c|}
-		\hline
-		Application  & Instructions & Dev.     & Dev. (\%) \\ \hline
-		Vip          & 7.97e+11     & 7.16e+06 & 0.00\%    \\ \hline
-		Openmc       & 8.17e+07     & 1.65e+04 & 0.02\%    \\ \hline
-		Rtview       & 9.91e+12     & 1.55e+09 & 0.02\%    \\ \hline
-		X264         & 4.52e+11     & 5.81e+07 & 0.01\%    \\ \hline
-		Bodytrack    & 1.86e+12     & 3.95e+10 & 2.13\%    \\ \hline
-		Fluidanimate & 2.09e+12     & 8.44e+10 & 4.04\%    \\ \hline
-		Xhpl         & 1.14e+08     & 1.24e+05 & 0.11\%    \\ \hline
-		Blackschole  & 3.75e+12     & 1.40e+09 & 0.04\%    \\ \hline
-		Dedup        & 1.02e+11     & 5.74e+07 & 0.06\%    \\ \hline
-		Swapti       & 2.43e+12     & 8.87e+08 & 0.04\%    \\ \hline
-		Canneal      & 1.19e+11     & 4.46e+07 & 0.04\%    \\ \hline
-		Freqmine     & 1.27e+12     & 4.78e+08 & 0.04\%    \\ \hline
-		Ferret       & 4.76e+11     & 7.04e+07 & 0.01\%    \\ \hline
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-	\caption{Frequency Variation}
-	\begin{tabular}{|c|c|c|c|}
-		\hline
-		Application  & Instructions & Dev.     & Dev. (\%) \\ \hline
-		Vip          & 7.97e+11     & 1.16e+06 & 0.00\%    \\ \hline
-		Openmc       & 8.17e+07     & 4.52e+03 & 0.01\%    \\ \hline
-		Rtview       & 9.91e+12     & 6.64e+05 & 0.00\%    \\ \hline
-		X264         & 4.52e+11     & 1.54e+05 & 0.00\%    \\ \hline
-		Bodytrack    & 1.84e+12     & 2.54e+05 & 0.00\%    \\ \hline
-		Fluidanimate & 2.38e+12     & 1.70e+09 & 0.07\%    \\ \hline
-		Xhpl         & 1.14e+08     & 5.95e+03 & 0.01\%    \\ \hline
-		Blackschole  & 3.75e+12     & 4.36e+05 & 0.00\%    \\ \hline
-		Dedup        & 1.02e+11     & 8.32e+07 & 0.08\%    \\ \hline
-		Swapti       & 2.43e+12     & 1.48e+05 & 0.00\%    \\ \hline
-		Canneal      & 1.19e+11     & 3.01e+05 & 0.00\%    \\ \hline
-		Freqmine     & 1.27e+12     & 3.70e+08 & 0.03\%    \\ \hline
-		Ferret       & 4.76e+11     & 5.63e+07 & 0.01\%    \\ \hline
-	\end{tabular}
-\end{table}
-
-\subsection{Input size and instructions}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/Blackschole.png}
-	\caption{Caption}
-	\label{fig:my_label}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/Canneal.png}
-	\caption{Caption}
-	\label{fig:my_label}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/Blackschole.png}
-	\caption{Caption}
-	\label{fig:my_label}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/Canneal.png}
-	\caption{Caption}
-	\label{fig:my_label}
-\end{figure}
-
-\section{Model experimental validation} \label{sec:experimentalvalidation}
-In this section, the models presented in~\cref{sec:powermodel} and~\cref{sec:performancemodel} were validated with a benchmark specific for multi-core architectures.
-Additionally, in order to assess the modeling overhead and accuracy, our proposal was then compared to machine learning approaches.
-We compared against Support Vector Regression (SVR)~\cite{Smola2004}, Decision Tree \cite{Kitts2006RegressionLecture}, k-nearest neighbors \cite{Altman1992AnRegression}, Multilayer perceptron \cite{Murtagh1991MultilayerRegression}, and some new methods such as Gao et al.
-\cite{Gao2019DendriticPrediction}.
-However, SVR was chosen as the most representative because in our tests it performed best without an aggressive fine-tuning.
-
-\subsection{Case-Study Applications} \label{sec:casestudyapplication}
-The PARSEC parallel benchmark suite, version 3.0~\cite{Bienia2008}, OpenMC \cite{Romano2015OpenMC:Development} and LINPACK (HPL) \cite{Dongarra1988TheExplanation}, were chosen as case studies.
-The PARSEC benchmark focused on emerging workloads and was designed to represent the next-generation shared-memory programs for chip-multiprocessors.
-It covers an ample range of areas such as financial analysis, computer vision, engineering, enterprise storage, animation, similarity search, data mining, machine learning, and media processing.
-The OpenMC and the LINPACK are two classic HPC programs.
-
-\subsection{Case-Study Architecture} \label{sec:casestudyarchitecture}
+\section{Case-Study Architecture} \label{sec:casestudyarchitecture}
 The experiments were executed in one computer node equipped with two Intel Xeon E5-2698 v3 processors with sixteen cores each and two hardware threads for each core. 
-The overall view of the architecture is shown in Figure \ref{fig:architecture}.
-The maximum non-turbo frequency is 2.3GHz, and the total physical memory of the node is 128GB (8$\times$16GB).
+The overall view of the architecture is shown in \cref{fig:architecture}.
+The maximum non-turbo frequency was 2.3 GHz, and the total physical memory of the node was 128 GB (8 $\times$ 16 GB).
 Turbo frequency and hardware multi-threading were disabled during all experiments.
@@
 
-\begin{figure}[htb!]
+\begin{figure}[H]
 	\centering
@@
 
-The Linux kernel has many different policies for power management, depending on the driver. In the default driver, the acpi-cpufreq, the options are:
+The Linux kernel has many different policies for power management, depending on the driver.
+In the default driver, the acpi-cpufreq, the options are Powersave, Performance, Ondemand, Conservative, and Userspace.
+Each governor has a policy on how the frequency is selected.
+In this investigation, the frequency control was performed using the Userspace governor, which allows the user or any userspace program to set the CPU to a specific frequency.
+The core control was accomplished by modifying the appropriate system files with the default CPU-hotplug driver.
+
+The architecture was equipped with the intelligent platform management interface (IPMI), a set of interfaces allowing out-of-band management of computer systems and platform-status monitoring via the local network~\cite{Schwenkler2006IntelligentInterface}.
+It can monitor variables and resources, such as the system's temperature, voltage, fans, and power supplies, with independent sensors attached to the hardware.
+
+\section{Case-Study Applications} \label{sec:casestudyapplication}
+The applications blackscholes, bodytrack, canneal, dedup, fluidanimate, freqmine, raytrace, swaptions, vips and x264 from the PARSEC \url{https://parsec.cs.princeton.edu/download.htm} (accessed on 20 February  2020) parallel benchmark suite, version 3.0~\cite{Bienia2008TheSuite}, OpenMC \cite{Romano2015OpenMC:Development} and LINPACK (HPL) \cite{Dongarra1988TheExplanation}, were chosen as case studies.
+The PARSEC benchmark focused on emerging workloads and was designed to represent the next-generation shared-memory programs for chip-multiprocessors.
+It covers an ample range of areas, such as financial analysis, computer vision, engineering, enterprise storage, animation, similarity search, data mining, machine learning, and media processing.
+The OpenMC and the LINPACK are two classic HPC programs.
+
+%The other two experiments use applications Raytrace~\cite{Bienia2008} and OpenMC~\cite{Romano2015OpenMC:Development}.
+Raytrace is an application addressed on rendering photorealistic scenes.
+It is available on the PARSEC Benchmark suite, version 2.0.
+The PARSEC benchmark, focused on emerging workloads, represents the next-generation shared-memory programs for chip multiprocessors.
+It covers a wide range of areas such as financial analysis, computer vision, engineering, enterprise storage, animation, similarity search, data mining, machine learning, and media~processing.
+%
+%\textls[-20]{{OpenMC is an application that implements the Monte-Carlo method to simulate the transport of neutrons and photons. It is a classic program aimed at high-performance computing.}}
+
+\section{Validation the framework}\label{sec:validation_the_framework}
+In this Section, we discuss the measurements and simulation results for three distinct cases, including two real applications.
+We also demonstrate how external libraries and standalone visualization tools can be used to render the collected data and investigate performance aspects such as the program's scalability capacity.
+
+\subsection{Configuration} \label{subsec:configuration_pascal}
+
+We used three experiments to assess the tool and demonstrate its ability to support analysis aimed at observing parallel scalability.
+The first includes a runtime imbalance between processing units in a specific parallel region.
+In this case, the objective is to present how the analyzer helps users observe the impact of the inefficiency of a code part on the whole program.
+The code for this experiment is presented in \cref{lst:codes_regionscomparison}.
+It consists of two simple parallel regions with the same functionality that divide the iterations of a loop between the available threads.
+The difference between the regions lies in the strategies used to manipulate the sum variable, used in the example to store the values of the calculations performed in each thread.
+We assume that the {\tt anyop()} function, in this case, invariably has the same runtime in all function calls.
+
+The command presented on \cref{lst:pascal_usage_ex} was used as base to perform experiments.
+For the experiment of \cref{lst:codes_regionscomparison}, we use just the parameters {\tt -c}, {\tt -r}, {\tt -t}, {\tt -a} and {\tt -o}, including the 64 value to {\tt -c} option.
+
+\lstset{style=ccodestyle, frame=tb}
+\begin{lstlisting}[label={lst:pascal_usage_ex}, language=bash, caption={Command line showing how experiments were run through a terminal.}]
+	analyzer 
+	application
+	-c 1,2,4,8,...,32 # threads/cores
+	-r 3 # number of repetitions
+	-t auto # automated instrumentation
+	-a 1 # aggregation mode
+	--ipmi ip user psswd # energy sensor
+	--idtm 5 # idle time between runs
+	--dhpt # disable hyper-thread
+	--dcrs # disable cores
+	--ipts ... # application specific inputs
+	-o application.json # output file
+\end{lstlisting}
+
+\lstset{style=ccodestyle, frame=tb}
+\begin{lstlisting}[label={lst:codes_regionscomparison}, language=C, caption={Sample code used to visualize the impact of regions on program scalability.}]
+	int main(int argc, char **argv) {
+		unsigned long sum = 0;
+		int ops = atoi(argv[1]);
+
+		#pragma omp parallel for schedule(static) reduction(+: sum)
+		for (int i = 0; i < ops; i++)
+		sum += anyop();
+
+		#pragma omp parallel for schedule(static)
+		for (int i = 0; i < ops; i++) {
+			#pragma omp critical
+			sum += anyop();
+		}
+	}
+\end{lstlisting}
+
+The command return a .json file with all the information necessary for our analysis.
+We can quickly generate tables with the collected data from this file, thus supporting observation and analysis of scalability, energy trends, and model fitting.
+The code described in \cref{lst:regtab} and \cref{tab:regtable_dedup} are samples of how users can easily read and view collected data from the control terminal in a tabular way.
+
+\lstset{style=pythonStyle, frame=tb}
+\begin{lstlisting}[label={lst:regtab}, language=python, captionpos=b, caption={Example of using the Python API to load analyzer files.}]
+	from analyzer import Data
+
+	data = Data("application.json")
+	data.energy(regions=True)
+	# data.speedup()
+	# data.efficiency()
+	# data.regions()
+\end{lstlisting}
+
+%The result can be seen in tabular form, as shown in Table.~\ref{tab:regtable_dedup}.
+Using these tables, we can quickly generate figures showing the application's behavior from various points of view.
+In Figures \ref{fig:speedup_raytrace} and \ref{fig:speedup_openmc} we observe the trend of the speedup with the number of cores and input.
+\begin{table}[H]
+	\caption{Dataframe generated automatically from collected samples using the Python API.}
+	\label{tab:regtable_dedup}
+
+	\resizebox{\columnwidth}{!}{%
+		\begin{tabular}{ccccccc}
+			\toprule
+			\textbf{Repetition} & \textbf{Input} & \textbf{Cores} & \textbf{Regions} & \textbf{Start Time (s)} & \textbf{Stop Time (s)} & \multicolumn{1}{c}{\textbf{ipmi Energy (J)}} \\ \midrule
+			1 & 1 & 1 & 1 & 0.00 & 66.31 & 13,156.19 \\
+			1 & 1 & 1 & 2 & 0.00 & 66.27 & 13,148.87 \\
+			... & ... & ... & ... & ... & ... & ... \\
+			1 & 5 & 30 & 4 & 0.00 & 8.46 & 1903.82 \\
+			1 & 5 & 30 & 5 & 0.02 & 58.16 & 13,067.01 \\ \bottomrule
+	\end{tabular}}
+\end{table}
+
+\subsection{Pascal Analyzer Validation}\label{sec:results_pascal}
+
+The analyzer does not display graphics and other visual elements natively.
+However, the simple use of external libraries allows generating graphics and visualizing points of the program's behavior that you want to observe.
+In addition, specialized visualization tools can also be used to view the results.
+PaScal Viewer \cite{Silva2018}, for example, natively interprets the analyzer's output files, complementing its functionality and creating an integrated and appropriate environment for the program's parallel scalability analysis.
+
+To analyze the experiment presented in \cref{lst:codes_regionscomparison}, we use PaScal Viewer to observe in a hierarchical way how the different OpenMP clauses impacted the region's efficiencies.
+The efficiency pinpoints how a program can take advantage of increasing processing elements on a parallel system.
+It is defined as the ratio of speedup to the number of processing units.
+In this sense, PaScal Viewer displays an efficiency diagram for each analysis region, as presented in \cref{fig:pv_regionscomparison}.
+Comparing these diagrams, it is possible to see how the critical clause damages the scalability.
+In addition, it is also possible to relate that this second region affects the efficiency of the whole program due to the code characteristics.
+
+If the reduction clause replaces {\tt \#pragma~omp~crtical}, the second region and the entire program become more efficient, as shown in \cref{fig:pv_regionscomparison_2}.
+The PaScal Viewer diagrams mentioned in this work, the x-axis (i1--i7) corresponds to different data inputs and the y-axis (2--64) to the number of threads used in the processing.
+Figures~\ref{fig:pv_regionscomparison} and \ref{fig:pv_regionscomparison_2} were rendered using a tool feature that smoothes the color transition of diagrams.
+This feature uses interpolation to create a visual element where the color transition is less pronounced.
+The diagram axes only show the initial and final values with this option.
+The user can visualize diagrams with only discrete values or even compare the two presentation modes side by side.
+\cref{fig:visualizationmodes} shows the difference between discrete and smoothed modes considering the whole program and the simulation scenario that uses the clause {\tt \#pragma~omp~crtical}.
+\begin{figure}[H]
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_r1.pdf}
+		\caption{\centering}
+		\label{fig:pv_regionscomparison_a}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_r2.pdf}
+		\caption{\centering}
+		\label{fig:pv_regionscomparison_b}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_whole.pdf}
+		\caption{\centering}
+		\label{fig:pv_regionscomparison_c}
+	\end{subfigure}
+	\caption{Efficiency diagrams and impact of inner regions on program scalability.
+(x-axis = inputs; y-axis = number of threads).
+(\textbf{a}) First region diagram.
+(\textbf{b}) Second region diagram.
+(\textbf{c}) Whole program~diagram.}
+	\label{fig:pv_regionscomparison}
+\end{figure}
+\unskip
+\begin{figure}[H]
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_1.pdf}
+		\caption{\centering}
+		\label{fig:pv_regionscomparison_a_2}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_2.pdf}
+		\caption{\centering}
+		\label{fig:pv_regionscomparison_b_2}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_3.pdf}
+		\caption{\centering}
+		\label{fig:pv_regionscomparison_c_2}
+	\end{subfigure}
+	\caption{Efficiency diagrams after removing {\tt \#pragma~omp~crtical} clause.
+(x-axis = inputs; y-axis = number of threads).
+(\textbf{a}) First region diagram.
+(\textbf{b}) Second region diagram.
+(\textbf{c}) Whole program~diagram.}
+	\label{fig:pv_regionscomparison_2}
+\end{figure}
+\unskip
+\begin{figure}[H]
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/diagram_discretevalues.pdf}
+		\caption{\centering}
+		\label{fig:discretevalues}
+	\end{subfigure}
+	%
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/diagram_smoothedvalues.pdf}
+		\caption{\centering}
+		\label{fig:smoothedvalues}
+	\end{subfigure}
+
+	\caption{Visualization modes of diagrams provided by PaScal Viewer. (\textbf{a}) PaScal Viewer diagram on discrete mode. (\textbf{b}) PaScal Viewer diagram on smoothed mode.}
+	\label{fig:visualizationmodes}
+\end{figure}
+
+%The analyzer allows that the result can be seen in tabular form, as shown in Table.~\ref{tab:regtable_dedup}.
+Using these tables, we can quickly generate figures that show the application's behavior from various points of view.
+In Figures \ref{fig:speedup_raytrace} and \ref{fig:speedup_openmc} we can observe the trend of the speedup with the number of cores and input.
+
+%Figures \ref{fig:raytrace_en} and \ref{fig:openmc_en} present the energy variation with the number of cores.
+
+\textls[-19]{Other analysis objectives not supported by PaScal Viewer, such as visualization of the speedup curve or energy consumption, can be supported by traditional plots.
+Figures~\ref{fig:speedup} and \ref{fig:energy}} present the charts rendered for analysis by Raytrace and OpenMC applications.
+In these figures, it is possible to observe that the program efficiency varies according to the increase in the number of threads and the execution of inputs with higher processing loads.
+From the diagrams in \cref{fig:efficiency_all}, it is possible to observe that the applications present different behaviors concerning their efficiency variations and their scalability capabilities.
+OpenMC maintains its efficiency almost constant.
+This pattern represents strong scalability and indicates that the program can maintain its efficiency level when it uses a larger number of threads and processing larger inputs.
+On the other hand, \cref{fig:efficiency_all}a demonstrates that Raytrace achieves higher efficiency values when processing larger inputs.
+However, it is also possible to observe that increasing the number of processing units fixing the input size will not improve or hold the efficiency value.
+\begin{figure}[H]
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/speedup_completo_rtview_2 (1).pdf}
+		\caption{\centering}
+		\label{fig:speedup_raytrace}
+	\end{subfigure}
+	%
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/speedup_completo_openmc_kernel_novo (1).pdf}
+		\caption{\centering}
+		\label{fig:speedup_openmc}
+	\end{subfigure}
+	\caption{\textls[-30]{Speedup of the applications for several input sizes. (\textbf{a}) Raytrace speedup. (\textbf{b}) OpenMC speedup. }}
+	\label{fig:speedup}
+\end{figure}
+\unskip
+\begin{figure}[H]
+	\begin{subfigure}[b]{0.46\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/energy_completo_rtview_2 (1).pdf}
+		\caption{\centering}
+		\label{fig:raytrace_en}
+	\end{subfigure}
+	%
+	\begin{subfigure}[b]{0.46\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/energy_completo_openmc_kernel_novo (1).pdf}
+		\caption{\centering}
+		\label{fig:openmc_en}
+	\end{subfigure}
+
+	\caption{Energy consumption identified in the execution of applications while varying the number of cores for several input sizes.
+		%MDPI: Please add commas to numbers to indicate thousand on the Oy axis of the figures, e.g. 20,000 30,000 40,000.
+		(\textbf{a}) Raytrace energy consumption. (\textbf{b}) OpenMC energy consumption.}
+	\label{fig:energy}
+\end{figure}
+\unskip
+%Furthermore, using an external graphical interface module, we can load the .json file and intuitively visualize the program's scalability showing the efficiency.
+The efficiency pinpoint how efficiently a program can take advantage of increasing processing elements on a parallel system.
+It is defined as the ratio of speedup to the number of processors.
+%For example, Figure \ref{fig:efficiency_all} shows examples of graphics generated for the same applications as above.
+In these figures, it is possible to observe that the program efficiency varies according to the increase in the number of threads and the execution of inputs with higher processing loads.
+From the Figure \ref{fig:efficiency_raytrace} we can see that as we increase the input size, the efficiency increases while slowly decreasing when the number of cores grows, indicating a tendency for a scalable program.
+In the second Figure \ref{fig:efficiency_openmc} the efficiency is practically constant while varying the number of cores and the input indicating a strong scalable program.
+\begin{figure}[H]
+	\begin{subfigure}[b]{0.46\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_raytrace.png}
+		\caption{\centering}
+		\label{fig:efficiency_raytrace}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.46\textwidth}
+		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_openmc.png}
+		\caption{\centering}
+		\label{fig:efficiency_openmc}
+	\end{subfigure}
+
+	\caption{Efficiency diagram varying the input size and the number of cores.
+The color bar indicates the efficiency value in percentage.
+(\textbf{a}) Raytrace efficiency diagram.
+(\textbf{b}) OpenMC efficiency diagram.}
+	\label{fig:efficiency_all}
+\end{figure}
+
+The input values shown in Figures~\ref{fig:speedup}--\ref{fig:efficiency_all} indicate different data sets for processing.
+For example, the ``input 2'' represents a load that will require sequential processing with runtime twice as long as ``input 1''.
+Likewise, the ``input 3'' represents a load that will require sequential processing with runtime twice as long as ``input 2'', and so on.
+
+Even if the analyzer does not display graphics natively, analysis aimed at observing scalability and energy consumption of applications depends on precise measurements, reinforcing the advantage of the analyzer proposed in this work.
+
+% \begin{figure}[H]
+% \centering
+% \includegraphics[width=\linewidth]{pascalanalyzer/figures/results/efficiency_raytrace.png}
+% \caption{Efficiency for Raytrace}
+% \label{fig:efficiency_raytrace}
+% \end{figure}
+
+% \begin{figure}[H]
+% \centering
+% \includegraphics[width=\linewidth]{pascalanalyzer/figures/results/efficiency_raytrace_detailed.png}
+% \caption{Efficiency for Raytrace}
+% \label{fig:efficiency_raytrace_detailed}
+% \end{figure}
+
+% \begin{figure}[H]
+% \centering
+% \includegraphics[width=\linewidth]{pascalanalyzer/figures/results/efficiency_openmc.png}
+% \caption{Efficiency for OpenMC}
+% \label{fig:efficiency_openmc}
+% \end{figure}
+
+% \begin{figure}[H]
+% \centering
+% \includegraphics[width=\linewidth]{pascalanalyzer/figures/results/efficiency_openmc_detailed.png}
+% \caption{Efficiency for OpenMC}
+% \label{fig:efficiency_openmc_detailed}
+% \end{figure}
+\subsection{Fingerprint Tool validation}
+
+In this subsection, we first show the comparison between our tool and others already established, as well as the clustering process.
+
+\subsubsection{ACCURACY COMPARISON}
+
+% \begin{table*}[h]
+% \centering
+% \caption{Average}
+% \begin{tabular}{|c|c|c|c|c|c|c|}
+% \hline
+% Counter                                 & Pined values & Linux API   & PAPI      & PAPI Python & Perf tool   & MyPerf      \\ \hline
+% INSTRUCTIONS\_RETIRED                  & 226990030    & 227000691   & 227000620 & 225901249   & 227000572   & 227000650   \\ \hline
+% BRANCH\_INSTRUCTIONS\_RETIRED          & 9240000      & 9250617     & 9250566   & 9239617     & 9250552     & 9250501     \\ \hline
+% BR\_INST\_RETIRED:CONDITIONAL          & 8220000      & 8220000     & 8220000   & 8209717     & 8220000     & 8220000     \\ \hline
+% MEM\_UOP\_RETIRED:ANY\_LOADS           &              & 2484182672  &           &             & 2484383940  & 2484155029  \\ \hline
+% MEM\_UOP\_RETIRED:ANY\_STORES          &              & 189962002   &           &             & 189961539   & 189961321   \\ \hline
+% UOPS\_RETIRED:ANY                      &              & 12291082129 &           &             & 12290811038 & 12290901997 \\ \hline
+% PARTIAL\_RAT\_STALLS:MUL\_SINGLE\_UOP  &              & 600878      &           &             & 600151      & 600330      \\ \hline
+% ARITH:FPU\_DIV                         &              & 5801446     &           &             & 5801000     & 5800977     \\ \hline
+% FP\_COMP\_OPS\_EXE:X87                 &              & 48785528    &           &             & 48784834    & 48786021    \\ \hline
+% INST\_RETIRED:X87                      &              & 17200008    &           &             & 17200008    & 17200007    \\ \hline
+% FP\_COMP\_OPS\_EXE:SSE\_SCALAR\_DOUBLE &              & 5401694     &           &             & 5401842     & 5401679     \\ \hline
+% \end{tabular}
+% \label{tab:mean}
+% \end{table*}
+
+% \begin{table*}[h]
+% \centering
+% \caption{Standard deviation}
+% \begin{tabular}{|c|c|c|c|c|c|}
+% \hline
+% Counter                                    & Linux API & PAPI & PAPI Python & Perf tool & MyPerf \\ \hline
+% INSTRUCTIONS\_RETIRED                  & 396       & 133  & 337763      & 110       & 175    \\ \hline
+% BRANCH\_INSTRUCTIONS\_RETIRED          & 297       & 208  & 8485        & 379       & 91     \\ \hline
+% BR\_INST\_RETIRED:CONDITIONAL          & 0         & 0    & 3383        & 0         & 0      \\ \hline
+% MEM\_UOP\_RETIRED:ANY\_LOADS           & 37399     &      &             & 39217     & 38953  \\ \hline
+% MEM\_UOP\_RETIRED:ANY\_STORES          & 1513      &      &             & 1035      & 687    \\ \hline
+% UOPS\_RETIRED:ANY                      & 345246    &      &             & 335832    & 333298 \\ \hline
+% PARTIAL\_RAT\_STALLS:MUL\_SINGLE\_UOP  & 1222      &      &             & 252       & 521    \\ \hline
+% ARITH:FPU\_DIV                         & 1760      &      &             & 1621      & 1544   \\ \hline
+% FP\_COMP\_OPS\_EXE:X87                 & 1283      &      &             & 1920      & 3311   \\ \hline
+% INST\_RETIRED:X87                      & 4         &      &             & 4         & 3      \\ \hline
+% FP\_COMP\_OPS\_EXE:SSE\_SCALAR\_DOUBLE & 1547      &      &             & 3259      & 2097   \\ \hline
+% \end{tabular}
+% \label{tab:std}
+% \end{table*}
+
+\begin{table}[H]
+	\centering
+	\caption{Comparison}
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{llllllllll}
+			\hline
+			\multicolumn{6}{l|}{\textbf{Average*$10^{-6}$}} & \multicolumn{4}{l}{\textbf{Standard deviation}}\\ \hline
+			\textbf{Counters} & \begin{tabular}{l}\textbf{Pined} \\ \textbf{values}\end{tabular} & \begin{tabular}{l}\textbf{Linux} \\ \textbf{API}\end{tabular} & \textbf{PAPI} & \begin{tabular}{l}\textbf{PAPI} \\ \textbf{Python}\end{tabular} & \textbf{MyPerf}  & \begin{tabular}{|l}\textbf{Linux} \\ \textbf{API}\end{tabular} & \textbf{PAPI} & \begin{tabular}{l}\textbf{PAPI} \\ \textbf{Python}\end{tabular} & \textbf{MyPerf} \\ \hline
+			INSTRUCTIONS\_RETIRED                  & 226.99       & 227       & 227  & 225.9       & 227     & 396       & 133  & 337763      & 175    \\
+			BRANCH\_INSTRUCTIONS\_RETIRED          & 9.24         & 9.25      & 9.25 & 9.24        & 9.25    & 297       & 208  & 8485        & 91     \\
+			BR\_INST\_RETIRED:CONDITIONAL          & 8.22         & 8.22      & 8.22 & 8.21        & 8.22    & 0         & 0    & 3383        & 0      \\
+			MEM\_UOP\_RETIRED:ANY\_LOADS           &              & 2484.18   &      &             & 2484.16 & 37399     &      &             & 38953  \\
+			MEM\_UOP\_RETIRED:ANY\_STORES          &              & 189.96    &      &             & 189.96  & 1513      &      &             & 687    \\
+			UOPS\_RETIRED:ANY                      &              & 12291.08  &      &             & 12290.9 & 345246    &      &             & 333298 \\
+			PARTIAL\_RAT\_STALLS:MUL\_SINGLE\_UOP  &              & 0.6       &      &             & 0.6     & 1222      &      &             & 521    \\
+			ARITH:FPU\_DIV                         &              & 5.8       &      &             & 5.8     & 1760      &      &             & 1544   \\
+			FP\_COMP\_OPS\_EXE:X87                 &              & 48.79     &      &             & 48.79   & 1283      &      &             & 3311   \\
+			INST\_RETIRED:X87                      &              & 17.2      &      &             & 17.2    & 4         &      &             & 3      \\
+			FP\_COMP\_OPS\_EXE:SSE\_SCALAR\_DOUBLE &              & 5.4       &      &             & 5.4     & 1547      &      &             & 2097   \\ \hline
+		\end{tabular}
+	}
+	\label{tab:counters}
+\end{table}
+
+To validate the tool, we compared the results of the counters obtained with different APIs. 
+We used the hand-crafted assembly benchmark from \cite{Weaver2013Non-determinismImplementations}, designed to test determinism and accuracy of PMUs.
+We compared the values obtained from the Linux API, PAPI on C and PAPI on Python. 
+The events used for this comparison were instructions retired, branch instructions, memory read, memory load, and arithmetic operations.
+We ran the benchmark 30 times and calculated the mean and standard deviation as shown in table \ref{tab:counters}.
+Some events could not be measured using PAPI because the tool does not accept raw events and there are no equivalent events.
+
+Since the benchmark was hand-crafted with assembly, we know exactly the value for some counter events. 
+For this reason, the number of instructions, branch instructions, and conditional branch are pin. However, some other events are architecture-specific and there is no pined value. 
+In the latter case, we can still compare to the Linux API, which should be closest to the reality.
+
+The differences using the Linux low-level API, PAPI, and our tool are negligible (the average percentage distance is less than 0.01\% in all the cases). 
+As expected, PAPI on Python had the largest difference (with an average distance of 0.25\%) mainly due to an unsynchronized start that resulted in the loss of some instructions at the beginning of the execution.
+This can be an important problem if the application contains a small number of instructions.
+
+The standard deviation of the 30 executions shows that our tool has the smallest variation on a run-to-run on most events. On the contrary, PAPI on Python shows a big variation compared to the others.
+
+\subsubsection{CLUSTERING}
+
+To cluster applications first we need to have a way to compare two programs, for that we define a new variable that tries to compute a fingerprint to the program, this variable has to look similar when we execute the same program with different conditions and inputs.
+Thinking in the simplest model of a program as a Turing machine as everything can be done with a tap of memory and set of rules we empirically define the variable input size as described on the equation \ref{eq:input_size}.
+On real computers this is not too far from reality, most of the computations and input and output operations somehow pass through the memory, so analyzing the relationship of the total number of instructions and memory instructions can give us a good fingerprint.
+
+\begin{equation}
+	I_{sz} = \frac{I}{I_{m}} \\
+	\label{eq:input_size}
+\end{equation}
+
+Where $I_{sz}$ we called input size, $I$ the number of instructions executed, $I_m$ number of memory instructions execute.
+We observe that this variable demonstrate to have the proprieties that we are looking for, produce similar results to the same program with different input size, and environment.
+This can be used to identify programs but also to see the similarities between different applications.
+
+To compute the distance between two programs we use the Canberra metric \cite{Jurman2009CanberraLists}, described on the equation \ref{eq:canberra}.
+
+\begin{equation}
+	d(p,q)= \sum_{i=1}^{n}{\frac{|p_i-q_i|}{|p_i|+|q_i|}}
+	\label{eq:canberra}
+\end{equation}
+
+Where $p$ and $q$ are n-dimensional vectors.
+
+We compute the input size for all 30 applications of the Polybench with 3 different inputs.
+Running each program 15 times, with a sampling rate of 0.01 seconds collecting the total number of instruction, number of loads and write instructions, number of floating point operations, besides some software counters.
+After applying the post-processing to the data collected we compute the input size and perform a Hierarchical clustering using the linkage method of Ward \cite{Murtagh2011WardsAlgorithm}, that minimizes the total within-cluster variance.
+The results of the clustering can be seen on the figure \ref{fig:sping_force_is} and on the dendrogram on figure \ref{fig:dendograma_input_size}.
+
+%The exact HPCs was:
+%PERF_COUNT_HW_INSTRUCTIONS,
+%MEM_UOPS_RETIRED:ALL_LOADS,
+%MEM_UOPS_RETIRED:ALL_STORES,
+%FP_ARITH_INST_RETIRED:SCALAR
+
+\begin{figure*}[h]
+	\centering
+	\includegraphics[width=\textwidth]{fingerprint/figures/dendograma_input_size.png}
+	\caption{Dendrogram}
+	\label{fig:dendograma_input_size}
+\end{figure*}
+
+From this dendrogram, we can have an idea of how close two applications are.
+We choose the number of clusters that maximize the number of hits of the same program with different input sizes on the same cluster, in this case, was 5 clusters.
+
+It was observed that as input size grows the program behavior tends to a specific curve, but for small input sizes some have variation, so in some cases, the same program has been classified in more than one cluster.
+This can also happen if specific parts of the program are triggered with specific inputs, in which case it will also belong to more than one cluster.
+
+In order to have an overall classification of each program, we can pick up the frequency in which each appeared in the clusters and classified it in the cluster in which it appeared more often.
+In this case, the clusters are:
+
 \begin{itemize}
-	\item Powersave
-	\item Performance
-	\item Ondemand
-	\item Conservative
-	\item Userspace
+	\item Cluster 1: 2mm, 3mm, cholesky, correlation, covariance, floyd-warshall, gemm, gramschmidt, lu, ludcmp, nussinov, symm
+
+	\item Cluster 2: deriche, doitgen, syrk
+
+	\item Cluster 3: adi, fdtd-2d, jacobi-2d, syr2k
+
+	\item Cluster 4: atax, bicg, durbin, gemver, gesummv, mvt, trisolv, trmm
+
+	\item Cluster 5: heat-3d, seidel-2d
 \end{itemize}
-In this work, the frequency control was performed using the Userspace governor, and the core control was accomplished by modifying the appropriate system files with the default CPU-hotplug driver.
 
-The architecture is equipped with the Intelligent Platform Management Interface (IPMI), a set of interfaces allowing out-of-band management of computer systems and platform-status monitoring via the local network~\cite{Schwenkler2006IntelligentInterface}.
-It can monitor variables and resources such as the system's temperature, voltage, fans, and power supplies, with independent sensors attached to the hardware.
+On the figure, \ref{fig:sping_force_is} we display the clusters using the spring force algorithm where each program with a specific input is a node and the edge weight is the Canberra distance.
+To a better visualization, the name of the inputs was replaced by numbers, the EXTRALARGE is 3, LARGE 2 and MEDIUM 1.
 
-\subsection{Fitting the Models} \label{sec:fitting}
-To find the parameters of the \cref{eq:en_final}, 10 uniformly random configurations of frequencies ($f$), cores ($p$) and inputs ($N$) were chosen from the range $1<=p<=32$, $1.2<=f<=2.2$ and $1<=N<=5$ respectively.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{fingerprint/figures/graph_input_size.png}
+	\caption{Spring force of Canberra distance}
+	\label{fig:sping_force_is}
+\end{figure}
+
+The figure above shows a different way to observe in a reduced space the distance between clusters and applications and how they are organized.
+From this graph, we see that clustering it's well partitioned and we can clearly separate each cluster.
+It is also interesting to note that the applications of the cluster formed by the circle symbol are more separated, which may indicate that they were classified in this way because they did not fit into any other cluster.
+
+% \begin{figure*}[t]
+%     \begin{subfigure}{\textwidth}
+%         \includegraphics[width=\textwidth]{fingerprint/figures/dendograma_input_size.png}
+%     \end{subfigure}
+%     \\
+%     \begin{subfigure}{\textwidth}
+%         \includegraphics[width=\textwidth]{fingerprint/figures/dendograma_floating.png}
+%     \end{subfigure}
+%     %and so on
+% \end{figure*}
+
+To have an idea of how the behavior of the variable input size for each clusters is, it was plotted for the applications of clusters 1 and 2 shown on the figures \ref{fig:c1_input_size_0}, \ref{fig:c1_input_size_1}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{fingerprint/figures/cluster_input_0.png}
+	\caption{Input size - Cluster 1}
+	\label{fig:c1_input_size_0}
+\end{figure}
+
+From these figures, we can have an idea of what behavior was classified as the same class.
+On cluster 1 on the figure \ref{fig:c1_input_size_0} all applications showed the same behavior with approximately the same amplitude.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{fingerprint/figures/cluster_input_1.png}
+	\caption{Input size - Cluster 2}
+	\label{fig:c1_input_size_1}
+\end{figure}
+
+On figure \ref{fig:c1_input_size_1} we observe that curves that have similar shape regardless of scale on vertical and horizontal axis have also been classified as the same clusters.
+This is the wanted comportment for the classification because we are only interested in the overall shape of the curve.
+
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_input_4.png}
+%     \caption{Input size - Cluster 3}
+%     \label{fig:c1_input_size_2}
+% \end{figure}
+
+% To conclude this study we also clustering applications by its floating point behavior.
+For that we only use the counter correspondent to floating point operations.
+The figure \ref{fig:dendogram_fp} shows the dendrogram and the \ref{fig:sping_force_fp} the spring force graph plot.
+
+% \begin{figure*}[h]
+%     \includegraphics[width=\textwidth]{fingerprint/figuresdendograma_floating.png}
+%     \caption{Dendrogram}
+%     \label{fig:dendogram_fp}
+% \end{figure*}
+
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=\textwidth]{fingerprint/figures/graph_floating2.png}
+%     \caption{Spring force of Canberra distance}
+%     \label{fig:sping_force_fp}
+% \end{figure}
+
+% The figures \ref{fig:cluster_fp1},\ref{fig:cluster_fp2},\ref{fig:cluster_fp3} show the behavior on the same cluster for the float point operations.
+For this classification the number of clusters was 8.
+
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_fp_0.png}
+%     \caption{Floating point - Cluster 1}
+%     \label{fig:cluster_fp1}
+% \end{figure}
+
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_fp_4.png}
+%     \caption{Floating point - Cluster 2}
+%     \label{fig:cluster_fp2}
+% \end{figure}
+
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_fp_5.png}
+%     \caption{Floating point - Cluster 3}
+%     \label{fig:cluster_fp3}
+% \end{figure}
+
+% Here we can also observed again that applications with the same shape were classified as the same clusters.
+
+\section{Validation the model} \label{sec:experimentalvalidation}
+In this section, the models presented in~Sections \ref{sec:power_model} and~\ref{sec:performance_model} were validated with a benchmark specific for multi-core architectures.
+Additionally, in order to assess the modeling overhead and accuracy, our proposal was then compared to machine learning approaches.
+We compared against support vector regression (SVR)~\cite{Smola2004ARegression}, decision tree \cite{Kitts2006RegressionLecture}, k-nearest neighbors \cite{Altman1992AnRegression}, multilayer perceptron \cite{Murtagh1991MultilayerRegression}, and some new methods, such as Gao et al.
+\cite{Gao2019DendriticPrediction}.
+However, SVR was chosen as the most representative because it performed best in our tests without aggressive fine-tuning, as shown in \cref{fig:ml_models}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\columnwidth]{experiments/figures/ml_models.pdf}
+	\caption{Average of the mean squared error for all applications of our study case Section \ref{sec:casestudyapplication}.}
+	\label{fig:ml_models}
+\end{figure}
+
+\subsection{Verifying Hypothesis}
+
+In this section, we validate whether the assumptions of our model are valid for the system used.
+
+\subsubsection{Frequency and Voltage Relation}
+One of the assumptions was that the frequency and the voltage have a linear relationship, as indicated by ~\cref{eq:f_v}.
+To verify that, we build an experiment that sets the frequency to a specific value while sampling the voltage using the APERF and MPERF registers that provide feedback on the current CPU frequency.
+The average result of the sampling voltages is shown in  \cref{fig:freq_volt_rel}, where we can observe a near-perfect linear relation.
+This is because manufacturers implement this curve in the processors, using tables that relate ranges of frequencies to voltages so that they can precisely define any curve that will better suit their design.
+
+\begin{figure}[H]
+	\centering
+	\captionsetup[subfigure]{justification=centering}
+	\includegraphics[width=\columnwidth]{experiments/figures/freq_volt_rel.pdf}
+	\caption{Frequency voltage relation.}
+	\label{fig:freq_volt_rel}
+\end{figure}
+
+\subsubsection{Input Size and Instructions}
+We ran the applications with different inputs assuming  linear growth in the amount of work for one input to the other when building our model.
+However, measuring and controlling the amount of work would require much instrumentation and tuning to find an input corresponding to a certain amount of work.
+Therefore, to build our models, we use the time to reference the amount of work, assuming that the work is proportional to the executing time.
+\cref{fig:time_input} corresponds to the verification of this supposition.
+
+\cref{tab:corr_in_time} shows that the assumption was reasonable since the average correlation was 0.96 for all applications, indicating that growth in the number of instructions  will follow the time.
+This was the case for all applications that we ran in our benchmark and should hold for any data parallelism type of application.
+\begin{figure}[H]
+	\centering
+	\captionsetup[subfigure]{justification=centering}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/blackscholes.pdf}
+		\caption{Blackscholes.}
+		\label{fig:time_input_blackshoels}
+	\end{subfigure}
+	%
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/canneal.pdf}
+		\caption{Canneal.}
+		\label{fig:time_input_canneal}
+	\end{subfigure}
+
+	\hfill
+	\caption{Relation between time and instructions for each input size. }
+	\label{fig:time_input}
+\end{figure}
+\vspace{-6pt}
+
+\begin{table}[H]
+	\centering
+	\caption{Correlation of time and instructions for all applications.}
+	\begin{tabular}{ll}
+		\multicolumn{1}{l|}{\textbf{Application}} & \textbf{Correlation} \\ \hline
+		Blackscholes                                               & 0.99                                  \\
+		Bodytrack                                                  & 0.99                                  \\
+		Canneal                                                    & 0.99                                  \\
+		Dedup                                                      & 0.99                                  \\
+		Ferret                                                     & 0.96                                  \\
+		Fluidanimate                                               & 0.99                                  \\
+		Freqmineq                                                  & 0.99                                  \\
+		Openmc                                                     & 0.94                                  \\
+		Raytrace                                                   & 0.99                                  \\
+		Swaptions                                                  & 0.99                                  \\
+		Vips                                                       & 0.98                                  \\
+		x264                                                       & 0.99                                  \\
+		HPL                                                        & 0.79                                  \\ \hline
+	\end{tabular}
+	\label{tab:corr_in_time}
+\end{table}
+
+The next assumption was that the application's behavior was the same when varying the workload.
+This condition is necessary for using the model with an unknown input size because, if the behavior is the same, we can interpolate the known inputs.
+One way to verify this is to measure the rate of instructions per second normalized by the frequency, as shown in \cref{fig:fingerpritns}.
+\begin{figure}[H]
+	\centering
+	\captionsetup[subfigure]{justification=centering}
+
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/blackscholes.pdf}
+		\caption{Blacksholes.}
+		\label{fig:fp_baclscholes}
+	\end{subfigure}
+	%
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/canneal.pdf}
+		\caption{Canneal.}
+		\label{fig:fp_canneal}
+	\end{subfigure}
+
+	\hfill
+	\caption{Rate of instructions per second varying the input size normalized by the frequency.}
+	\label{fig:fingerpritns}
+\end{figure}
+
+\cref{fig:fingerpritns} shows that the applications have roughly the same curve when normalized; this also happens for all other applications in our benchmark.
+
+The final assumption is that the workload should also not vary depending on the number of cores or frequency.
+To verify, we measure the total number of executed instructions while varying the cores from 1 to 32.
+\cref{tab:cores_variation} shows the results.
+
+\begin{table}[H]
+	\centering
+	\caption{Variation of the number of instructions when changing the number of cores for the same input.}
+	%\begin{tabular}{c|c|c|c}
+	\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}cccc}
+		\toprule
+		& \textbf{Average} & \textbf{Standard}     & \textbf{Standard}\\
+		\multirow{-2}{*}{\textbf{Application}}& \textbf{Number of Instructions} & \textbf{Deviation}     & \textbf{Deviation} \textbf{(\%) }\\ \midrule
+		Vip          & $7.97 \times 10^{11}$     & $7.16 \times 10^{6}$  & 0.00 \\ %Please use Scientific notation.
+For example, “1.65e+04” should be changed to “1.65× 104”.
+Please confirm `%` whether to remove.
+		Openmc       & $8.17 \times 10^{7}$     & $ 1.65 \times 10^{4}$ & 0.02    \\ 
+		Rtview       & $9.91 \times 10^{12}$     & $ 1.55 \times 10^{9}$ & 0.02    \\ 
+		X264         & $4.52 \times 10^{11}$     & $ 5.81 \times 10^{7}$ & 0.01    \\ 
+		Bodytrack    & $1.86 \times 10^{12}$     & $ 3.95 \times 10^{10}$ & 2.13    \\ 
+		Fluidanimate & $2.09 \times 10^{12}$     & $ 8.44 \times 10^{10}$ & 4.04    \\ 
+		HPL          & $1.14 \times 10^{8}$      & $ 1.24 \times 10^{5}$ & 0.11    \\ 
+		Blackschole  & $3.75 \times 10^{12}$     & $ 1.40 \times 10^{9}$ & 0.04     \\ 
+		Dedup        & $1.02 \times 10^{11}$     & $ 5.74 \times 10^{7}$ & 0.06     \\ 
+		Swapti       & $2.43 \times 10^{12}$     & $ 8.87 \times 10^{8}$ & 0.04     \\ 
+		Canneal      & $1.19 \times 10^{11}$     & $ 4.46 \times 10^{7}$ & 0.04     \\ 
+		Freqmine     & $1.27 \times 10^{12}$     & $ 4.78 \times 10^{8}$ & 0.04     \\ 
+		Ferret       & $4.76 \times 10^{11}$     & $ 7.04 \times 10^{7}$ & 0.01     \\ \bottomrule
+	\end{tabular*}
+	\label{tab:cores_variation}
+\end{table}
+% Corrected
+
+\cref{tab:cores_variation} shows the standard deviation and what that corresponds to in terms of the total number of instructions as a percentage.
+
+The same test was performed for the frequency, varying from 1.2 to 2.2 GHz with 100~MHz steps. The results are shown in \cref{tab:freq_variation}.
+
+These results show that all the assumptions were reasonable, and we can safely move to the validation of the model's prediction.
+\begin{table}[H]
+	\caption{Variation of the number of instructions when changing the frequency for the same input.}
+	\centering
+	%\begin{tabular}{c|c|c|c}{\hsize}{@{\extracolsep{\fill}}cccc}
+	\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}cccc}
+		\toprule
+		& \textbf{Average} & \textbf{Standard}     & \textbf{Standard}\\
+		\multirow{-2}{*}{\textbf{Application}}& \textbf{Number of Instructions} & \textbf{Deviation}     & \textbf{Deviation} \textbf{(\%)} \\ \midrule
+		Vip          & $7.97 \times 10^{11}$      & $ 1.16 \times 10^{6}$  & 0.00      \\ 
+		Openmc       & $8.17 \times 10^{7}$       & $ 4.52 \times 10^{3}$  & 0.01      \\ 
+		Rtview       & $9.91 \times 10^{12}$      & $ 6.64 \times 10^{5}$  & 0.00      \\ 
+		X264         & $4.52 \times 10^{11}$      & $ 1.54 \times 10^{5}$  & 0.00      \\ 
+		Bodytrack    & $1.84 \times 10^{12}$      & $ 2.54 \times 10^{5}$  & 0.00      \\ 
+		Fluidanimate & $2.38 \times 10^{12}$      & $ 1.70 \times 10^{9}$  & 0.07      \\ 
+		HPL          & $1.14 \times 10^{8}$       & $ 5.95 \times 10^{3}$  & 0.01      \\ 
+		Blackschole  & $3.75 \times 10^{12}$      & $ 4.36 \times 10^{5}$  & 0.00      \\ 
+		Dedup        & $1.02 \times 10^{11}$      & $ 8.32 \times 10^{7}$  & 0.08      \\ 
+		Swapti       & $2.43 \times 10^{12}$      & $ 1.48 \times 10^{5}$  & 0.00      \\ 
+		Canneal      & $1.19 \times 10^{11}$      & $ 3.01 \times 10^{5}$  & 0.00      \\ 
+		Freqmine     & $1.27 \times 10^{12}$      & $ 3.70 \times 10^{8}$  & 0.03      \\ 
+		Ferret       & $4.76 \times 10^{11}$      & $ 5.63 \times 10^{7}$  & 0.01      \\\bottomrule
+	\end{tabular*}
+	\label{tab:freq_variation}
+\end{table}
+
+\subsection{Fitting the Models} \label{subsec:fitting}
+To find the parameters of  \cref{eq:en_final}, 10 uniformly random configurations of frequencies ($f$), cores ($p$) and inputs ($N$) were chosen from the range $1<=p<=32$, $1.2<=f<=2.2$ and $1<=N<=5$, respectively.
 The application was executed for each chosen configuration, and the measured energy and time values were collected.
-For the input size if we assume that all CPU instructions are executed at approximately the same time, the number of basic operations will be directly correlated with the time.
-Thus, we can estimate the problem size by looking at the execution time, allowing us to divide a large problem size into several smaller ones knowing their relationship as done in the work of Oliveira ~\cite{Oliveira2018ApplicationCores}.
+For the input size, if we assume that all CPU instructions take approximately the same time to execute, the number of basic operations will be directly correlated with the time.
+Thus, we can estimate the input size by looking at the execution time, allowing us to divide a large input size into several smaller ones, knowing their relationship, as performed in the work of Oliveira ~\cite{Oliveira2018ApplicationCharacterization}.
 The unity can also vary depending on the definition.
-In our case for simplicity, we assign numbers from 1 the smallest size to 10 the largest, increasing the problem linearly, that way its also possible to interpolate any input in between these values.
-
-%The application's input was chosen to increase the workload linearly, from smallest to largest, so that most of the input space is covered~\cite{Oliveira2018ApplicationCores}.
+For simplicity, we assign numbers from 1 to 10, increasing the problem linearly, so it is also possible to interpolate any input in between these values.
 
 For each configuration, samples of the power were collected using IPMI every 1 second.
-This sampling rate was chosen because the order of magnitude of the mean run time of the applications is minutes.
+This sampling rate was chosen based on the magnitude of the mean run time of the applications, which is in the order of minutes.
 Therefore, this rate provides enough samples to measure average power.
 Additionally, timestamps and the total run time were collected.
-The total energy spent on each configuration is estimated by first interpolation the power samples using a first order method and then integrating this function in the time.
+The total energy spent on each configuration is estimated by first interpolating the power samples using the first-order method and then integrating this function in the time.
 
-From the sampled data, the parameters of the model can be estimated.
-This results in an optimization problem of finding the parameters that minimize the distance between estimated and the measured values.
-To solve this minimization problem, non-linear least squares have been applied.
+The model's parameters are calculated by solving an optimization problem of finding the values that minimize the squared error of the prediction to the measured values using the non-linear least-squares method.
 
-The python library scikit-learning was used to build the SVR model~\cite{PedregosaFandVaroquauxGandGramfortAandMichel2011Scikit-learn:Python}.
-The SVR was trained using the same data used for parameters estimation of \cref{eq:en_final} with a grid search used to find the best kernel function and the best values for the hyper-parameters penalty for the wrong ($C$) and ($\gamma$).
-For this data, the best function was the Radial Base Function (RBF), and the hyper-parameters were $C=10^4$ and $\gamma=0.5$.
+The Python library Scikit-Learn was used to build the SVR model~\cite{Pedregosa2011Scikit-learn:Python}.
+The SVR was trained using the same data used for parameter estimation of Equation (\ref{eq:en_final}) %\cref{eq:en_final}
+with a grid search used to find the best kernel function and the best values for the hyper-parameters penalty for the wrong ($C$) and ($\gamma$).
+For this data, the best function was the radial base function (RBF), and the hyper-parameters were $C=10^4$ and $\gamma=0.5$.
 
 \subsection{Measured versus Modeled Energy}
-\label{sec:measuredversusmodeledenergy}
+\label{subsec:measuredversusmodeledenergy}
 
-To validate the proposed energy model, all possible configurations were tested by varying the cores in a range of $1<=p<=32$, the frequency in $1.2<=f<=2.2$ and the input in $1<=N<=5$, which varies from 400 to over 1000 configurations depending on the application as some applications have restrictions on the number of cores.
-The mean percentage error was computed as the difference between the estimated and measured values according to the following equation:
+To validate the model, we ran all possible configurations in the tested machine, varying the cores in a range of $1<=p<=32$, the frequency in $1.2<=f<=2.2$, and the input in $1<=N<=5$.
+The total number of configurations varies from 400 to over 1000 depending on the application, as some applications have restrictions on the number of cores that they can run.
+Once the data was collected, we computed the mean percentage error (MPE) according to the following equation:
 \begin{equation}
@@
 
-\subsubsection{Frequency x Cores}
-Figures \ref{fig:en_eq_black}, \ref{fig:en_eq_canneal} plot the measured and modeled energy consumption for some of the applications modeled.
-They  show some of the possible shapes that the model can take while varying the number of active cores, operating frequency, and input size.
+\subsubsection{Frequency $\times$ Cores}
+\cref{fig:en_eq_freq_cores_fc} plots the measured and modeled energy consumption for some of the applications modeled.
+In addition,  some of the possible shapes that the model can take while varying the number of active cores, and operating frequency, are shown.%please confirm that the intended meaning has been retained.
+% yes its the same
+
 \begin{figure}[H]
 	\centering
-	\begin{subfigure}[b]{0.48\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_cores/completo_black_5.png}}
-		\caption{Blackscholes.}
-		\label{fig:en_eq_black}
+	\captionsetup[subfigure]{justification=centering}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_cores/completo_black_5.pdf}}
+		\caption{}
+		\label{fig:en_eq_black_fc}
 	\end{subfigure}
 	%
-	\begin{subfigure}[b]{0.48\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_cores/completo_canneal_1.png}}
-		\caption{Canneal.}
-		\label{fig:en_eq_canneal}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_cores/completo_canneal_1.pdf}}
+		\caption{}
+		\label{fig:en_eq_canneal_fc}
 	\end{subfigure}
-	\caption{Example fit for a specific input size: Blackscholes (a) and Canneal (b).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
+
+	\caption{Example fit for a specific input size: Blackscholes (\textbf{a}) and Canneal (\textbf{b}).
+“measured values” are the sensor data, and “minimum energy” is the minimum energy model prediction.
 	}
+	\label{fig:en_eq_freq_cores_fc}
 \end{figure}
-\subsubsection{Frequency x Input}
-Figures \ref{fig:en_eq_black}, \ref{fig:en_eq_canneal} plot the measured and modeled energy consumption for some of the applications modeled.
-They  show some of the possible shapes that the model can take while varying the number of active cores, operating frequency, and input size.
+\subsubsection{Frequency $\times$ Input}
+\cref{fig:en_eq_freq_inp_fi} plots the measured and modeled energy consumption for some of the applications modeled.
+The diagrams show some of the possible shapes that the model can take while varying the operating frequency, and input size.
 \begin{figure}[H]
 	\centering
-	\begin{subfigure}[b]{0.48\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_inps/completo_black_5.png}}
-		\caption{Blackscholes.}
-		\label{fig:en_eq_black}
+	\captionsetup[subfigure]{justification=centering}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_inps/completo_black_5.pdf}}
+		\caption{}
+		\label{fig:en_eq_black_fi}
 	\end{subfigure}
 	%
-	\begin{subfigure}[b]{0.48\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_inps/completo_canneal_1.png}}
-		\caption{Canneal.}
-		\label{fig:en_eq_canneal}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_inps/completo_canneal_1.pdf}}
+		\caption{}
+		\label{fig:en_eq_canneal_fi}
 	\end{subfigure}
-	\caption{Example fit for a specific input size: Blackscholes (a) and Canneal (b).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
+
+	\caption{Example fit for a specific input size: Blackscholes (\textbf{a}) and Canneal (\textbf{b}).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
 	}
+	\label{fig:en_eq_freq_inp_fi}
 \end{figure}
-\subsubsection{Cores x Input}
-Figures \ref{fig:en_eq_black}, \ref{fig:en_eq_canneal} plot the measured and modeled energy consumption for some of the applications modeled.
-They  show some of the possible shapes that the model can take while varying the number of active cores, operating frequency, and input size.
+\subsubsection{Cores $\times$ Input}
+\cref{fig:en_eq_core_inp_ci} plots the measured and modeled energy consumption for some of the applications modeled.
+The diagrams  show some of the possible shapes that the model can take while varying the number of active cores, and input size.
 \begin{figure}[H]
 	\centering
-	\begin{subfigure}[b]{0.48\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/cores_inps/completo_black_5.png}}
-		\caption{Blackscholes.}
-		\label{fig:en_eq_black}
+	\captionsetup[subfigure]{justification=centering}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/cores_inps/completo_black_5.pdf}}
+		\caption{}
+		\label{fig:en_eq_black_ci}
 	\end{subfigure}
 	%
-	\begin{subfigure}[b]{0.48\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/cores_inps/completo_canneal_1.png}}
-		\caption{Canneal.}
-		\label{fig:en_eq_canneal}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/cores_inps/completo_canneal_1.pdf}}
+		\caption{}
+		\label{fig:en_eq_canneal_ci}
 	\end{subfigure}
-	\caption{Example fit for a specific input size: Blackscholes (a) and Canneal (b).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
+
+	\caption{Example fit for a specific input size: Blackscholes (\textbf{a}) and Canneal (\textbf{b}).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
 	}
+	\label{fig:en_eq_core_inp_ci}
 \end{figure}
 
-\subsubsection{Validation}
-The validation results for each application trained with 10 random configurations are displayed in figure \ref{fig:mpe_svr_eq} and the raw MPE values in  table~\ref{tab:mpe_svr_eq}.
-\begin{figure}[htb!]
-	\includegraphics[width=\columnwidth]{models/figures/mpe_svr_eq.png}
-	\caption{Comparison of the Mean Percentage Error between the proposed model and SVR. "Model mean" and "SVR mean" are the average of all MPE values for all applications.
+\subsubsection{Comparison}
+The average results for each application were calculated using a model trained with only 10 configurations, and the comparison is displayed \cref{fig:mpe_svr_eq}.
+%We calculated the raw MPE values shown in ~\cref{tab:mpe_svr_eq}.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=.8\columnwidth]{models/figures/mpe_svr_eq.pdf}
+	\caption{Comparison of the mean percentage error between the proposed model and SVR. ``Model mean'' and ``SVR mean'' are the average of all MPE values for all applications.
 	}
@@
 \end{figure}
-\begin{table}[htb!]
+
+\cref{fig:mpe_svr_eq} shows that the proposed model always performed better, with a lower MPE than SVR, when we were limited to 10 training points.
+This result is further explored in the next Section \ref{subsec:overhead}, where we undertake a comparison with different training sizes.
+The exact values are shown in \cref{tab:mpe_svr_eq}.
+
+\begin{table}[H]
 \centering
-	\begin{tabular}{|c|c|c|}
+\begin{tabular}{c|c|c}
 \hline
 Application  & Model & SVR   \\ \hline
-		Ferret       & 5.25     & 12.49  \\ \hline
-		Raytrace     & 6.36     & 11.95  \\ \hline
-		Fluianimate  & 2.44     & 22.90  \\ \hline
-		x264         & 8.28     & 15.33  \\ \hline
-		Vips         & 7.54     & 10.80  \\ \hline
-		Swaptions    & 6.54     & 18.57  \\ \hline
-		Canneal      & 3.12     & 6.13   \\ \hline
-		Dedup        & 8.85     & 13.70  \\ \hline
-		Freqmine     & 2.44     & 3.24   \\ \hline
-		Blackscholes & 2.18     & 11.00  \\ \hline
-		HPL          & 7.47     & 12.75  \\ \hline
-		Bodytrack    & 16.98    & 34.12  \\ \hline
-		Openmc       & 11.15    & 24.34  \\ \hline
+Ferret       & 5.25     & 12.49  \\ 
+Raytrace     & 6.36     & 11.95  \\
+Fluianimate  & 2.44     & 22.90  \\
+x264         & 8.28     & 15.33  \\
+Vips         & 7.54     & 10.80  \\
+Swaptions    & 6.54     & 18.57  \\
+Canneal      & 3.12     & 6.13   \\
+Dedup        & 8.85     & 13.70  \\
+Freqmine     & 2.44     & 3.24   \\
+Blackscholes & 2.18     & 11.00  \\
+HPL          & 7.47     & 12.75  \\
+Bodytrack    & 16.98    & 34.12  \\
+Openmc       & 11.15    & 24.34  \\
 \end{tabular}
@@
 
-Figure \ref{fig:mpe_svr_eq} shows that the proposed model always performed better, with a lower MPE, than SVR when we were limited to 10 training points.
-This result will be further explored in \cref{sec:overhead} where we compare with different training sizes.
-
-\subsection{Overheads on training} \label{sec:overhead}
-It is known that machine learning is data-driven, in that sense the SVR model obtained using only 10 configurations could be improved, but what about the analytical model? 
+\subsection{Overheads on Training} \label{subsec:overhead}
+It is known that machine learning is data-driven; in that sense, the SVR model obtained using only 10 configurations could be improved, but what about the analytical model? 
 To answer that question, the proposed model and the SVR were also trained with a varying number of configurations.
 We then compared the MPE and the amount of energy spent to create each model. 
-This accuracy-energy trade-off is crucial since the energy overhead when building models defeats the primary goal of saving power when running applications.
+This accuracy-energy trade-off is crucial since building models' energy overhead defeats the primary goal of saving power when running applications.
+\enlargethispage{.5cm}
 
@@
 	\centering
+	\captionsetup[subfigure]{justification=centering}
 	\begin{subfigure}[b]{0.45\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_ferret_4.png}}
+		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_ferret_4.pdf}}
 		\caption{MPE for Ferret.}
@@
 	\begin{subfigure}[b]{0.45\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_vips_4.png}}
-		\caption{MPE for Vips.}
-		\label{fig:overhead_vips}
+		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_black_6.pdf}}
+		\caption{MPE for Blackscholes.}
+		\label{fig:overhead_black}
+	\end{subfigure}
+	%
+	\par\bigskip
+	\begin{subfigure}[b]{0.45\textwidth}
+		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_x264_4.pdf}}
+		\caption{MPE for x264.}
+		\label{fig:overhead_x264}
 	\end{subfigure}
@@
 	\begin{subfigure}[b]{0.45\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_x264_4.png}}
-		\caption{MPE for x264.}
-		\label{fig:overhead_x264}
+		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_dedup_4.pdf}}
+		\caption{MPE for Dedup.}
+		\label{fig:overhead_dedup}
 	\end{subfigure}
-	\caption{MPE case studies: Ferret (a), Vips (b) and x264 (c): model (almost constant) versus SVR (lowest MPE beyond train-size 35 on case (b)).}
-	\label{fig:overheadFerretVips}
+
+	\caption{MPE of the case studies versus training size, comparing how many training points is necessary to reach an acceptable result.}
+	\label{fig:overheadapps}
 \end{figure}
 
-Figure \ref{fig:overheadFerretVips} shows the comparisons of MPE and energy spent to create each model for two selected applications.
-According to the results, the analytical model is found to be very stable, not changing much as more data is added, while the SVR keeps reshaping to adapt to the data.
-The error of the analytical model is almost constant but that of  the SVR, initially very high, drops at some point.
+\cref{fig:overheadapps} shows the comparisons of MPE and energy spent to create each model for two selected applications.
+According to the results, the analytical model is very stable, not changing much as more data is added, while the SVR keeps reshaping to adapt to the data.
+The error of the analytical model is almost constant but that of the SVR, initially very high, drops as more data is used in the training process.
 \begin{figure}[H]
 	\centering
-	\begin{subfigure}[b]{0.44\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/overall.png}}
-		\caption{Average energy spent on all applications during model creation. The two curves are identical because the same data were used to adjust the SVR and the model.}
+	\captionsetup[subfigure]{justification=centering}
+	\begin{subfigure}[t]{0.45\textwidth}
+		\includegraphics[width=\columnwidth]{models/figures/overhead/overall_energy_10pts.pdf}
+		\caption{}
 		\label{fig:overall_overhead}
@@
 	%
-	\quad
-	\begin{subfigure}[b]{0.44\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/overall_mpe.png}}
+	%\par\bigskip
+	\begin{subfigure}[t]{0.45\textwidth}
+		\includegraphics[width=\columnwidth]{models/figures/overhead/overall_mpe_10pts.pdf}
 		\caption{
-			MPE all applications: SVR needs 10 times more data to have an MPE lower than the model.}
+		}
 		\label{fig:overall_MPE}
 	\end{subfigure}
-	\caption{Overall results for energy and MPE for each training size.}
+	\caption{Overall results for energy and MPE for each training size.
+(\textbf{a}) Average energy spent on all applications during model creation.
+The two curves are identical because the same data were used to adjust the SVR and the model.
+(\textbf{b}) MPE of all applications: SVR needs 10 times more data to have an MPE lower than the proposed model.}
 	\label{fig:overall_train}
 \end{figure}
-Figure \ref{fig:overall_train} presents the overall results, with the mean energy overhead and MPE for all applications.
-The meeting point of the MPE for the SVR and the proposed model can be extracted from  \ref{fig:overall_MPE}.
-There it shows that around 90 configurations the SVR starts to have a smaller error.
+
+\cref{fig:overall_train} presents the overall results, with the mean energy overhead and MPE for all applications.
+The meeting point of the MPE for the SVR and the proposed model can be extracted from \cref{fig:overall_train}b.
+It shows that, in around 90 configurations, the SVR starts to have a smaller error.
 The cost of that is the linear increase in energy spent on training.
-The increase in energy, about 10 times more, can be observed in Figure \ref{fig:overall_overhead}.
+The increase in energy, about 10 times more, can be observed in \cref{fig:overall_train}a.
 
-\section{PMCs}
+\subsection{Analysis}
+One of the most significant advantages of using an analytical model is the understanding of the problem that an equation provides, making many different kinds of analysis possible that are otherwise impossible with a machine learning model.
+In this section, we discuss one of the possible analyses.
+In the following figures, we try to understand the contribution of each parameter of the equation to the total energy consumption.
 
-\section{Energy per instruction}
+For this analysis, we took the model of one of the applications and, varying one parameter of the equation, we display the energy versus performance (time) for all configurations.
+After that, we computed the Pareto frontier, a set of all Pareto efficient allocations, i.e., all the configurations where resources cannot be reallocated to make one individual better off without making at least one individual worse off.
+This gives us all the configurations where we have an optimal trade-off of performance and energy to choose from.
 
-\begin{lstlisting}
-xor rcx, rcx
-mov rax, 1
-mov rdx, 0
+\cref{fig:pareto_static} shows the Pareto frontier for several values for the static power parameter ($c_3$ in \cref{eq:en_final}) with configurations of frequency ranging from 1.2 to 5 GHz and cores from 1 to 64, so that we can also have an idea of what is the tendency when we increase the frequency and number of cores.
 
-loop:
-	targ_inst(*arg)
-	targ_inst(*arg)
-	targ_inst(*arg)
-	targ_inst(*arg)
-	targ_inst(*arg)
-	targ_inst(*arg)
-	targ_inst(*arg)
-	targ_inst(*arg)
-	targ_inst(*arg)
+\begin{figure}[H]
+	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_high.pdf}
+	\caption{Pareto frontier for several values of static power parameter.
+The arrows with blue heads indicate the maximum energy, while the arrows with a red head the minimal energy for each corresponding curve.}% In four-digit numbers, no comma or spaces to indicate thousand, whatever the numbers are in the table or body text, e.g., 4500; More than five-digit numbers, should have comma, e.g.
+59,400; 899,302.
+Please revise the figure.
+	\label{fig:pareto_static}
+\end{figure}
+%Fixed
 
-add rcx, 1
-cmp rcx, 9999999
-jne loop
-\end{lstlisting}
+From this figure, we can see that when increasing the value of the static power parameter, the total energy consumption increases as expected.
+We can also observe that the values that minimize the total energy consumption tend to be high frequency and multiple cores.
+This is one of the consequences of increasing the static power factor.
+As the dynamic factor proportionally decreases, its variables tend to have less impact on total consumption, enabling configurations with high frequency and several cores.
+This also enables chip-level optimization for choosing components that change the ratio between static and dynamic power.
 
-Estimating the energy of the instruction from this benchmark
+\cref{fig:pareto_w} shows the Pareto frontier in the same ranges described before but for the parameter corresponding to the level of parallelism of the application ($w$ in  \cref{eq:en_final}).
 
-$E_{total}=9999999(\frac{10}{13}inst+\frac{3}{13}loop)$
+\begin{figure}[H]
+	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_w_high.pdf}
+	\caption{Pareto frontier for several values of static power parameter.
+The arrows with blue heads indicate the maximum energy, while the arrows with a red head, the minimal, for each corresponding curve.}% In four-digit numbers, no comma or spaces to indicate thousand, whatever the numbers are in the table or body text, e.g., 4500; More than five-digit numbers, should have comma, e.g.
+59,400; 899,302.
+Please revise the figure.
+	\label{fig:pareto_w}
+\end{figure}
+%Fixed
 
-$E_{total}=7692306*inst+constant$
+In  \cref{fig:pareto_w}, we observe that, as the parallelism level increases the total energy decreases.
+The number of cores tends to be higher with a higher level of parallelism as expected, and the frequency shows an inverse relation.
 
-\subsection{Generic}
+\subsection{DVFS and DPM Optimization} \label{subsec:dvfs_optmin}
+The effectiveness of the proposed approach during optimization was evaluated with a simple algorithm that finds the optimal frequency and number of active cores from the proposed equation.
+The results were then compared to the Linux default choices for power management.
 
-\begin{figure}
+With \cref{eq:en_final}, it is possible to calculate energy consumption estimates for each possible configuration since there is a finite range of possible values for the frequency and number of cores.
+It is also possible to apply constraints on the execution time, frequency, and the number of active cores.
+Then, the configuration that minimizes energy consumption for a given input can be selected.
+The complete workflow is shown in \cref{fig:optim_workflow}.
+We can see that any optimization problem can be structured with our model and the system's constraints.
+In the following examples, the optimization problem that we build is to minimize the energy equation given the constraints of possible frequencies and the number of cores that our system can run.
+The algorithm selected to minimize was the newton-CG~\cite{Royer2020AOptimization}.
+
+Current HPC managers leave to the user the choice of how many cores to use. On this basis, three situations were analyzed in relation to the number of cores:
+
+\begin{enumerate}
+	\item Worst choice: number of cores that maximize the total energy consumed;
+	\item Random choice: energy consumed for a random choice of the number of cores;
+	\item Best choice: number of cores that minimize the total energy consumed (oracle).
+\end{enumerate}
+\begin{figure}[H]
 	\centering
-	\includegraphics[width=\textwidth]{experiments/figures/inst_en_args_generic.png}
-	\caption{Energy per instruction argument}
-	\label{fig:experiment_en1}
+	\includegraphics[width=0.8\columnwidth]{models/figures/DVFS optim.pdf}
+	\caption{Optimization workflow showing how DVFS and DPM optimization could be implemented from ou model.}
+	\label{fig:optim_workflow}
 \end{figure}
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{experiments/figures/inst_mean_en_generic.png}
-	\caption{Mean energy per instruction over all arguments}
-	\label{fig:experiment_en2}
-\end{figure}
+The default option for the Linux governor is Ondemand, and, by default, it has no DPM control for the number of active cores.
+As Ondemand only performs DVFS, \textls[-15]{for comparison, each application was executed with all available cores in the system, from 1 to 32.}
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{experiments/figures/inst_std_en_generic.png}
-	\caption{Standard deviation energy per instruction over all arguments}
-	\label{fig:experiment_en3}
-\end{figure}
+\textls[-15]{Figures \ref{fig:energy_worst_case}--\ref{fig:energy_best_case}, show the energy savings with respect to Ondemand, i.e., $\frac{Ondemand-Model_{min}}{Ondemand}$} for the three cases described above.
+The savings and losses for each case are:
 
-\subsection{SSE}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{experiments/figures/inst_en_args_sse.png}
-	\caption{Energy per instruction argument (sse)}
-	\label{fig:experiment_en4}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{experiments/figures/inst_mean_en_sse.png}
-	\caption{Mean energy per instruction over all arguments (sse)}
-	\label{fig:experiment_en5}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{experiments/figures/inst_std_en_sse.png}
-	\caption{Standard deviation energy per instruction over all arguments (sse)}
-	\label{fig:experiment_en6}
-\end{figure}
-
-\section{SSE and Generic}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{experiments/figures/inst_en_cmp_sse_generic.png}
-	\caption{Comparing SSE and generic instructions energy consumption}
-	\label{fig:experiment_en7}
-\end{figure}
-
-\section{Power}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{experiments/figures/inst_pw_generic.png}
-	\caption{Instructions power draw}
-	\label{fig:experiment_pw1}
-\end{figure}
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{experiments/figures/inst_pw_args.png}
-	\caption{Instructions power draw by argument}
-	\label{fig:experiment_pw2}
-\end{figure}
-
-\section{Tracing application methods}
-
-\begin{outline}
-	\1 Linux ptrace
-		\2 Breakpoint on each instruction	
-	\1 Intel pin
-		\2 JIT
-		\2 Not open source
-	\1 qemu user
-		\2 TCG accelerator (transform target to host instruction) similar to JIT
-		\2 Partial emulation
-\end{outline}
-
-We choose the less intrusive and open-source alternative qemu user, with an small modificaion we are able to trace an application instructions. Generating an table as follows:
-
-\begin{table}[H]
-	\centering
-	\begin{tabular}{|c|c|c|}
-		\hline
-		opcode         & mnemonic & args                          \\ \hline
-		48 89 e7       & movq     & \%rsp, \%rdi\textbackslash{}n \\ \hline
-		48 89 e7       & movq     & \%rsp, \%rdi\textbackslash{}n \\ \hline
-		e8 08 0e 00 00 & callq    & 0x4000a24ea0\textbackslash{}n \\ \hline
-		55             & pushq    & \%rbp\textbackslash{}n        \\ \hline
-		48 89 e5       & movq     & \%rsp, \%rbp\textbackslash{}n \\ \hline
-		41 57          & pushq    & \%r15\textbackslash{}n        \\ \hline
-		41 56          & pushq    & \%r14\textbackslash{}n        \\ \hline
-		41 55          & pushq    & \%r13\textbackslash{}n        \\ \hline
-		& \vdots & \\ \hline
-	\end{tabular}
-\end{table}
-
-\section{CPU load}
+\begin{enumerate}
+	\item Worst choice: save 69.88\% on average;
+	\item Random choice: save 12.04\% on average;
+	\item Best choice: lost 14.06\% on average.
+\end{enumerate}
 
@@
 	\centering
-	\begin{subfigure}[b]{0.45\textwidth}
+	\includegraphics[width=\columnwidth]{models/figures/dvfs_cmp_max.pdf}
+	\caption{Energy savings comparisons between the proposed model and the Worst case.}
+	\label{fig:energy_worst_case}
+\end{figure}
+
+\begin{figure}[H]
 	\centering
-		\includegraphics[width=1\columnwidth]{experiments/figures/pw_freq_load.png}
-		\caption{Instructions power draw varing CPU load}
-		\label{fig:experiment_pw_load}
-	\end{subfigure}
+	\includegraphics[width=\columnwidth]{models/figures/dvfs_cmp_mean.pdf}
+	\caption{Energy savings comparisons between the proposed model and the Random case.}
+	\label{fig:energy_mean_case}
+\end{figure}
+
+By default, operating systems do not implement DPM at the core level, and, in HPC, the user usually explicitly chooses the number of cores to run their job.
+To give a better idea of the impact on the energy consumption of DPM at the core level, we analyzed the choices of the number of cores over a period of one year in the HPC center at UFRN.
+The result is plotted in~\cref{fig:cpu_requests}.
+
+It is of note that the most common choice of many regular users is a single core requested per job, matching the worst-case choice for all applications analyzed in this investigation.
+The best choice was quite often 32 cores, which is the third most popular choice among users, but it is 72 times less frequent than 1 core.
+This led us to envision how much energy could be saved and encouraged us towards future research using the proposed model for DPM or more advanced optimization algorithms.
+
+In practice, this approach can be implemented by allowing the resource manager to perform these changes for the user using pre-scripts and post-scripts for high energy consumption job submissions.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\columnwidth]{models/figures/dvfs_cmp_32.pdf}
+	\caption{Energy savings comparisons between the proposed model and the Best case.}
+	\label{fig:energy_best_case}
+\end{figure}
+\vspace{-12pt}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\columnwidth]{experiments/figures/cpu_requestes.pdf}
+	\caption{Number of CPU requests during one year in HPC cluster, sorted by the number of cores requested per job.}
+	\label{fig:cpu_requests}
+\end{figure}
+
+%\section{PMCs}
 %
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centering
-		\includegraphics[width=1\columnwidth]{experiments/figures/pw_load.png}
-		\caption{Instructions power draw varing CPU load}
-		\label{fig:experiment_pw_load}
-	\end{subfigure}
-\end{figure}
+%\section{Energy per instruction}
+%
+%\begin{lstlisting}
+%xor rcx, rcx
+%mov rax, 1
+%mov rdx, 0
+%
+%loop:
+%	targ_inst(*arg)
+%	targ_inst(*arg)
+%	targ_inst(*arg)
+%	targ_inst(*arg)
+%	targ_inst(*arg)
+%	targ_inst(*arg)
+%	targ_inst(*arg)
+%	targ_inst(*arg)
+%	targ_inst(*arg)
+%
+%add rcx, 1
+%cmp rcx, 9999999
+%jne loop
+%\end{lstlisting}
+%
+%Estimating the energy of the instruction from this benchmark
+%
+%$E_{total}=9999999(\frac{10}{13}inst+\frac{3}{13}loop)$
+%
+%$E_{total}=7692306*inst+constant$
+%
+%\subsection{Generic}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\textwidth]{experiments/figures/inst_en_args_generic.png}
+%	\caption{Energy per instruction argument}
+%	\label{fig:experiment_en1}
+%\end{figure}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\textwidth]{experiments/figures/inst_mean_en_generic.png}
+%	\caption{Mean energy per instruction over all arguments}
+%	\label{fig:experiment_en2}
+%\end{figure}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\textwidth]{experiments/figures/inst_std_en_generic.png}
+%	\caption{Standard deviation energy per instruction over all arguments}
+%	\label{fig:experiment_en3}
+%\end{figure}
+%
+%\subsection{SSE}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\textwidth]{experiments/figures/inst_en_args_sse.png}
+%	\caption{Energy per instruction argument (sse)}
+%	\label{fig:experiment_en4}
+%\end{figure}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\textwidth]{experiments/figures/inst_mean_en_sse.png}
+%	\caption{Mean energy per instruction over all arguments (sse)}
+%	\label{fig:experiment_en5}
+%\end{figure}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\textwidth]{experiments/figures/inst_std_en_sse.png}
+%	\caption{Standard deviation energy per instruction over all arguments (sse)}
+%	\label{fig:experiment_en6}
+%\end{figure}
+%
+%\subsection{SSE and Generic}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\textwidth]{experiments/figures/inst_en_cmp_sse_generic.png}
+%	\caption{Comparing SSE and generic instructions energy consumption}
+%	\label{fig:experiment_en7}
+%\end{figure}
+%
+%\subsection{Power}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\textwidth]{experiments/figures/inst_pw_generic.png}
+%	\caption{Instructions power draw}
+%	\label{fig:experiment_pw1}
+%\end{figure}
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\textwidth]{experiments/figures/inst_pw_args.png}
+%	\caption{Instructions power draw by argument}
+%	\label{fig:experiment_pw2}
+%\end{figure}
+%
+%\section{Tracing application methods}
+%
+%\begin{outline}
+%	\1 Linux ptrace
+%		\2 Breakpoint on each instruction	
+%	\1 Intel pin
+%		\2 JIT
+%		\2 Not open source
+%	\1 qemu user
+%		\2 TCG accelerator (transform target to host instruction) similar to JIT
+%		\2 Partial emulation
+%\end{outline}
+%
+%We choose the less intrusive and open-source alternative qemu user, with an small modificaion we are able to trace an application instructions. Generating an table as follows:
+%
+%
+%\begin{table}[H]
+%	\centering
+%	\begin{tabular}{|c|c|c|}
+%		\hline
+%		opcode         & mnemonic & args                          \\ \hline
+%		48 89 e7       & movq     & \%rsp, \%rdi\textbackslash{}n \\ \hline
+%		48 89 e7       & movq     & \%rsp, \%rdi\textbackslash{}n \\ \hline
+%		e8 08 0e 00 00 & callq    & 0x4000a24ea0\textbackslash{}n \\ \hline
+%		55             & pushq    & \%rbp\textbackslash{}n        \\ \hline
+%		48 89 e5       & movq     & \%rsp, \%rbp\textbackslash{}n \\ \hline
+%		41 57          & pushq    & \%r15\textbackslash{}n        \\ \hline
+%		41 56          & pushq    & \%r14\textbackslash{}n        \\ \hline
+%		41 55          & pushq    & \%r13\textbackslash{}n        \\ \hline
+%		& \vdots & \\ \hline
+%	\end{tabular}
+%\end{table}
+%
+%
+%\section{CPU load}
+%
+%\begin{figure}[H]
+%	\centering
+%	\begin{subfigure}[b]{0.45\textwidth}
+%		\centering
+%		\includegraphics[width=1\columnwidth]{experiments/figures/pw_freq_load.png}
+%		\caption{Instructions power draw varing CPU load}
+%		\label{fig:experiment_pw_load}
+%	\end{subfigure}
+%	%
+%	\begin{subfigure}[b]{0.45\textwidth}
+%		\centering
+%		\includegraphics[width=1\columnwidth]{experiments/figures/pw_load.png}
+%		\caption{Instructions power draw varing CPU load}
+%		\label{fig:experiment_pw_load}
+%	\end{subfigure}
+%\end{figure}
+%
+%
+%\section{Analysis}
+%
+%\subsection{Pareto points}
+
+%
+%\begin{figure}
+
+%	\centering
+
+%	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_high.png}
+
+%	\caption{Pareto static energy}
+
+%	\label{fig:pareto_static_h}
+
+%\end{figure}
+
+%
+%
+%\begin{figure}
+
+%	\centering
+
+%	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_low.png}
+
+%	\caption{Pareto static energy}
+
+%	\label{fig:pareto_static_l}
+
+%\end{figure}
+
+%
+%
+%\begin{figure}
+
+%	\centering
+
+%	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_w_high.png}
+
+%	\caption{Pareto w energy}
+
+%	\label{fig:pareto_w_h}
+
+%\end{figure}
+
+%
+%\begin{figure}
+
+%	\centering
+
+%	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_low.png}
+
+%	\caption{Pareto w energy}
+
+%	\label{fig:pareto_w_l}
+
+%\end{figure}
+
+%
+%\subsection{Optimization under constraints}
+
+%
+%\subsection{Gradient and Countorns}
+
+%
+%\begin{figure}[H]
+
+%	\centering
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn0.png}
+
+%	\end{subfigure}
+
+%	%
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn0_3d.png}
+
+%	\end{subfigure}
+
+%\end{figure}
+
+%
+%\begin{figure}[H]
+
+%	\centering
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn3.png}
+
+%	\end{subfigure}
+
+%	%
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn3_3d.png}
+
+%	\end{subfigure}
+
+%\end{figure}
+
+%
+%\begin{figure}[H]
+
+%	\centering
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/pleak0.png}
+
+%	\end{subfigure}
+
+%	%
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/pleak0_3d.png}
+
+%	\end{subfigure}
+
+%\end{figure}
+
+%
+%\begin{figure}[H]
+
+%	\centering
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/pleak10.png}
+
+%	\end{subfigure}
+
+%	%
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/pleak10_3d.png}
+
+%	\end{subfigure}
+
+%\end{figure}
+
+%
+%\begin{figure}[H]
+
+%	\centering
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic0.png}
+
+%	\end{subfigure}
+
+%	%
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic0_3d.png}
+
+%	\end{subfigure}
+
+%\end{figure}
+
+%
+%\begin{figure}[H]
+
+%	\centering
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic3000.png}
+
+%	\end{subfigure}
+
+%	%
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic3000_3d.png}
+
+%	\end{subfigure}
+
+%\end{figure}
+
+%
+%\begin{figure}[H]
+
+%	\centering
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/w0.png}
+
+%	\end{subfigure}
+
+%	%
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/w0_3d.png}
+
+%	\end{subfigure}
+
+%\end{figure}
+
+%
+%\begin{figure}[H]
+
+%	\centering
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/w1.png}
+
+%	\end{subfigure}
+
+%	%
+
+%	\begin{subfigure}[b]{0.45\textwidth}
+
+%		\includegraphics[width=\textwidth]{models/figures/analisys/w1_3d.png}
+
+%	\end{subfigure}
+
+%\end{figure}

### 2022-04-05 thesis/fingerprint/fingerprint.tex
update thesis text
add journals and restructured

@@
-%\author{ \parbox{3 in}{\centering Huibert Kwakernaak*
-%         \thanks{*Use the $\backslash$thanks command to put information here}\\
-%         Faculty of Electrical Engineering, Mathematics and Computer Science\\
-%         University of Twente\\
-%         7500 AE Enschede, The Netherlands\\
-%         {\tt\small h.kwakernaak@autsubmit.com}}
-%         \hspace*{ 0.5 in}
-%         \parbox{3 in}{ \centering Pradeep Misra**
-%         \thanks{**The footnote marks may be inserted manually}\\
-%        Department of Electrical Engineering \\
-%         Wright State University\\
-%         Dayton, OH 45435, USA\\
-%         {\tt\small pmisra@cs.wright.edu}}
-%}
-
-\section{An Accurate Tool for Modeling, Fingerprinting, Comparison, and Clustering of Parallel Applications Based on Performance Counters}
+\section{Fingerprint tool}
 The analysis of application performance is essential to better exploit its potential on High-Performance Computing (HPC) architectures.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{INTRODUCTION}
+\subsection{Introduction}
 
@@
 Performance counters were initially introduced for debugging, but since then they have provided a lot of useful information about running applications without slowing down the execution.
-They have been used in several other areas, such as software profiling \cite{Melo2010, Kufrin2005, Knupfer2011}, CPU power modeling \cite{Zamani2012ASystems}, dynamic frequency and voltage scaling, vulnerability research and malware defense \cite{Demme2013OnCounters}.
+They have been used in several other areas, such as software profiling \cite{Melo2010Perf, Kufrin2005Perfsuite, Knupfer2011Scorep}, CPU power modeling \cite{Zamani2012ASystems}, dynamic frequency and voltage scaling, vulnerability research and malware defense \cite{Demme2013OnCounters}.
 
@@
 Although many tools have been developed using performance counters, programmable interfaces capable of providing good accuracy are still lacking, especially for high-level programming languages.
-Indeed, apart from PAPI \cite{Weaver2013, Mucci1999} and Perfmon \cite{EranianPerfmon2Interface, Eranian2005TheSpecification} there are only a few APIs allowing access to these counters, and many others are poorly documented, unstable, or designer for a specific purpose.
+Indeed, apart from PAPI \cite{Weaver2013Non-determinismImplementations, Mucci1999PAPI} and Perfmon \cite{Eranian2008Perfmon2} there are only a few APIs allowing access to these counters, and many others are poorly documented, unstable, or designer for a specific purpose.
 
@@
 Therefore, besides gathering information, post-processing modules are also needed. 
-Such modules will overcome the lack of precision of counters on some architectures, as indicated in \cite{Weaver2008,Weaver2013a,Das2019SoK:Security}. 
+Such modules will overcome the lack of precision of counters on some architectures, as indicated in \cite{Weaver2008CanTrusted,Weaver2013Non-determinismImplementations,Das2019SoK:Security}. 
 Events that must be precise and deterministic (such as retired instructions) show a variation on run-to-run and over-count on x86\_64 machines, even in strictly controlled environments.
@@
 
-The definition of reference parameters, such as input size of programs, and performance measures uniformization, allow us to cluster benchmarks such as PolyBench \cite{PolyBench/C3.2}, a collection of numerical computations with static control flow extracted from various application domains, with interesting results.
+The definition of reference parameters, such as input size of programs, and performance measures uniformization, allow us to cluster benchmarks such as PolyBench \cite{Gonzalez2021PolyBench}, a collection of numerical computations with static control flow extracted from various application domains, with interesting results.
 In addition to contributing to the standardization of kernel execution and monitoring, this clustering has identified applications with similar and comparable behavior in terms of input size, data transfer and access, resources used and computation; which facilitates the creation of test sets for a given environment, according to specific measurement parameters.
@@
 
-\subsection{RELATED WORK}
+\subsection{Related Work}
 
 There are only a few APIs allowing access to performance counters.
-PAPI \cite{Weaver2013, Mucci1999}, one of the most used libraries for accessing hardware performance counters, was originally developed to provide portable access to the counters found on a diverse collection of modern microprocessors.
+PAPI \cite{Weaver2013Non-determinismImplementations, Mucci1999PAPI}, one of the most used libraries for accessing hardware performance counters, was originally developed to provide portable access to the counters found on a diverse collection of modern microprocessors.
 Rather than learning and writing a new performance infrastructure every time it is ported to a new machine.
@@
 
-Perfctr \cite{MikaelPettersson.TheInterface.} supports per-kernel-thread and system-wide monitoring for most major processor architectures.
+Perfctr \cite{Nikolaev2011Perfctr} supports per-kernel-thread and system-wide monitoring for most major processor architectures.
 It is distributed as a stand-alone kernel patch.
@@
 
-The Intel VTUNE \cite{IntelCorp.TheAnalyzer.} performance analyzer comes with its own kernel interface, implemented by an open-source driver.
+The Intel VTUNE \cite{Intel2021Vtune} performance analyzer comes with its own kernel interface, implemented by an open-source driver.
 The interface supports system-wide monitoring only and is very specific to the needs of the tool.
 
-The problem with the approach of a tool and its own kernel interface is dangerous because, as mentioned on \cite{EranianPerfmon2Interface}, there is clearly code duplication, but more importantly, there is no coordination between the various interfaces that may coexist sharing access to the same PMU resource.
-To solve this problem, Perfmon2 \cite{Eranian2005TheSpecification} offers a standard interface that all tools can use.
+The problem with the approach of a tool and its own kernel interface is dangerous because, as mentioned on \cite{Eranian2008Perfmon2}, there is clearly code duplication, but more importantly, there is no coordination between the various interfaces that may coexist sharing access to the same PMU resource.
+To solve this problem, Perfmon2 \cite{Eranian2008Perfmon2} offers a standard interface that all tools can use.
 Unfortunately, it has not been widely adopted, just supported by a few architectures like the IA64.
@@
 
-\subsection{READING PERFORMANCE COUNTERS}
+\subsection{Reading Performance Counters}
 
@@
 Ideal hardware performance counters provide exact deterministic results.
-Real-world PMU implementations do not always live up to this ideal \cite{Weaver2008, Weaver2013a, Das2019SoK:Security, McGuire2009}.
+Real-world PMU implementations do not always live up to this ideal \cite{Weaver2008CanTrusted, Weaver2013Non-determinismImplementations, Das2019SoK:Security, McGuire2009AnalysisKernel}.
 Events that should be exact and deterministic (such as the number of executed instructions) show run-to-run variations and over counts on x86 64 machines, even when running in strictly controlled environments.
@@
 
-\subsection{IMPLEMENTATION}
+\subsection{Implementation}
 
@@
 The Workload and libpfm4 module are developed on C/C++ and interfaced with python using Python C API with SWIG (a software development tool to connects programs written in C/C++ with a variety of high-level programming languages).
-The libpfm4 module, developed by \cite{Eranian2008}, is used as an auxiliary library.
+The libpfm4 module, developed by \cite{Eranian2008Perfmon2}, is used as an auxiliary library.
 
@@
 \end{figure}
-
-\subsection{RESULTS}
-
-In this subsection, we first show the comparison between our tool and others already established, as well as the clustering process.
-
-\subsubsection{ACCURACY COMPARISON}
-
-% \begin{table*}[h]
-% \centering
-% \caption{Average}
-% \begin{tabular}{|c|c|c|c|c|c|c|}
-% \hline
-% Counter                                 & Pined values & Linux API   & PAPI      & PAPI Python & Perf tool   & MyPerf      \\ \hline
-% INSTRUCTIONS\_RETIRED                  & 226990030    & 227000691   & 227000620 & 225901249   & 227000572   & 227000650   \\ \hline
-% BRANCH\_INSTRUCTIONS\_RETIRED          & 9240000      & 9250617     & 9250566   & 9239617     & 9250552     & 9250501     \\ \hline
-% BR\_INST\_RETIRED:CONDITIONAL          & 8220000      & 8220000     & 8220000   & 8209717     & 8220000     & 8220000     \\ \hline
-% MEM\_UOP\_RETIRED:ANY\_LOADS           &              & 2484182672  &           &             & 2484383940  & 2484155029  \\ \hline
-% MEM\_UOP\_RETIRED:ANY\_STORES          &              & 189962002   &           &             & 189961539   & 189961321   \\ \hline
-% UOPS\_RETIRED:ANY                      &              & 12291082129 &           &             & 12290811038 & 12290901997 \\ \hline
-% PARTIAL\_RAT\_STALLS:MUL\_SINGLE\_UOP  &              & 600878      &           &             & 600151      & 600330      \\ \hline
-% ARITH:FPU\_DIV                         &              & 5801446     &           &             & 5801000     & 5800977     \\ \hline
-% FP\_COMP\_OPS\_EXE:X87                 &              & 48785528    &           &             & 48784834    & 48786021    \\ \hline
-% INST\_RETIRED:X87                      &              & 17200008    &           &             & 17200008    & 17200007    \\ \hline
-% FP\_COMP\_OPS\_EXE:SSE\_SCALAR\_DOUBLE &              & 5401694     &           &             & 5401842     & 5401679     \\ \hline
-% \end{tabular}
-% \label{tab:mean}
-% \end{table*}
-
-% \begin{table*}[h]
-% \centering
-% \caption{Standard deviation}
-% \begin{tabular}{|c|c|c|c|c|c|}
-% \hline
-% Counter                                    & Linux API & PAPI & PAPI Python & Perf tool & MyPerf \\ \hline
-% INSTRUCTIONS\_RETIRED                  & 396       & 133  & 337763      & 110       & 175    \\ \hline
-% BRANCH\_INSTRUCTIONS\_RETIRED          & 297       & 208  & 8485        & 379       & 91     \\ \hline
-% BR\_INST\_RETIRED:CONDITIONAL          & 0         & 0    & 3383        & 0         & 0      \\ \hline
-% MEM\_UOP\_RETIRED:ANY\_LOADS           & 37399     &      &             & 39217     & 38953  \\ \hline
-% MEM\_UOP\_RETIRED:ANY\_STORES          & 1513      &      &             & 1035      & 687    \\ \hline
-% UOPS\_RETIRED:ANY                      & 345246    &      &             & 335832    & 333298 \\ \hline
-% PARTIAL\_RAT\_STALLS:MUL\_SINGLE\_UOP  & 1222      &      &             & 252       & 521    \\ \hline
-% ARITH:FPU\_DIV                         & 1760      &      &             & 1621      & 1544   \\ \hline
-% FP\_COMP\_OPS\_EXE:X87                 & 1283      &      &             & 1920      & 3311   \\ \hline
-% INST\_RETIRED:X87                      & 4         &      &             & 4         & 3      \\ \hline
-% FP\_COMP\_OPS\_EXE:SSE\_SCALAR\_DOUBLE & 1547      &      &             & 3259      & 2097   \\ \hline
-% \end{tabular}
-% \label{tab:std}
-% \end{table*}
-
-\begin{table*}[h]
-\centering
-\caption{Comparison}
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-\hline
-\multicolumn{6}{|c|}{Average*$10^{-6}$} & \multicolumn{4}{c|}{Standard deviation}\\ \hline
-Counters & \begin{tabular}{c}Pined\\values\end{tabular} & \begin{tabular}{c}Linux\\API\end{tabular} & PAPI & \begin{tabular}{c}PAPI\\Python\end{tabular} & MyPerf  & \begin{tabular}{c}Linux\\API\end{tabular} & PAPI & \begin{tabular}{c}PAPI\\Python\end{tabular} & MyPerf \\ \hline
-INSTRUCTIONS\_RETIRED                  & 226.99       & 227       & 227  & 225.9       & 227     & 396       & 133  & 337763      & 175    \\ \hline
-BRANCH\_INSTRUCTIONS\_RETIRED          & 9.24         & 9.25      & 9.25 & 9.24        & 9.25    & 297       & 208  & 8485        & 91     \\ \hline
-BR\_INST\_RETIRED:CONDITIONAL          & 8.22         & 8.22      & 8.22 & 8.21        & 8.22    & 0         & 0    & 3383        & 0      \\ \hline
-MEM\_UOP\_RETIRED:ANY\_LOADS           &              & 2484.18   &      &             & 2484.16 & 37399     &      &             & 38953  \\ \hline
-MEM\_UOP\_RETIRED:ANY\_STORES          &              & 189.96    &      &             & 189.96  & 1513      &      &             & 687    \\ \hline
-UOPS\_RETIRED:ANY                      &              & 12291.08  &      &             & 12290.9 & 345246    &      &             & 333298 \\ \hline
-PARTIAL\_RAT\_STALLS:MUL\_SINGLE\_UOP  &              & 0.6       &      &             & 0.6     & 1222      &      &             & 521    \\ \hline
-ARITH:FPU\_DIV                         &              & 5.8       &      &             & 5.8     & 1760      &      &             & 1544   \\ \hline
-FP\_COMP\_OPS\_EXE:X87                 &              & 48.79     &      &             & 48.79   & 1283      &      &             & 3311   \\ \hline
-INST\_RETIRED:X87                      &              & 17.2      &      &             & 17.2    & 4         &      &             & 3      \\ \hline
-FP\_COMP\_OPS\_EXE:SSE\_SCALAR\_DOUBLE &              & 5.4       &      &             & 5.4     & 1547      &      &             & 2097   \\ \hline
-\end{tabular}
-}
-\label{tab:counters}
-\end{table*}
-
-To validate the tool, we compared the results of the counters obtained with different APIs. 
-We used the hand-crafted assembly benchmark from \cite{Weaver2013a}, designed to test determinism and accuracy of PMUs.
-We compared the values obtained from the Linux API, PAPI on C and PAPI on Python. 
-The events used for this comparison were instructions retired, branch instructions, memory read, memory load, and arithmetic operations.
-We ran the benchmark 30 times and calculated the mean and standard deviation as shown in table \ref{tab:counters}.
-Some events could not be measured using PAPI because the tool does not accept raw events and there are no equivalent events.
-
-Since the benchmark was hand-crafted with assembly, we know exactly the value for some counter events. 
-For this reason, the number of instructions, branch instructions, and conditional branch are pin. However, some other events are architecture-specific and there is no pined value. 
-In the latter case, we can still compare to the Linux API, which should be closest to the reality.
-
-The differences using the Linux low-level API, PAPI, and our tool are negligible (the average percentage distance is less than 0.01\% in all the cases). 
-As expected, PAPI on Python had the largest difference (with an average distance of 0.25\%) mainly due to an unsynchronized start that resulted in the loss of some instructions at the beginning of the execution.
-This can be an important problem if the application contains a small number of instructions.
-
-The standard deviation of the 30 executions shows that our tool has the smallest variation on a run-to-run on most events. On the contrary, PAPI on Python shows a big variation compared to the others.
-
-\subsubsection{CLUSTERING}
-
-To cluster applications first we need to have a way to compare two programs, for that we define a new variable that tries to compute a fingerprint to the program, this variable has to look similar when we execute the same program with different conditions and inputs.
-Thinking in the simplest model of a program as a Turing machine as everything can be done with a tap of memory and set of rules we empirically define the variable input size as described on the equation \ref{eq:input_size}.
-On real computers this is not too far from reality, most of the computations and input and output operations somehow pass through the memory, so analyzing the relationship of the total number of instructions and memory instructions can give us a good fingerprint.
-
-\begin{equation}
-    I_{sz} = \frac{I}{I_{m}} \\
-    \label{eq:input_size}
-\end{equation}
-
-Where $I_{sz}$ we called input size, $I$ the number of instructions executed, $I_m$ number of memory instructions execute.
-We observe that this variable demonstrate to have the proprieties that we are looking for, produce similar results to the same program with different input size, and environment.
-This can be used to identify programs but also to see the similarities between different applications.
-
-To compute the distance between two programs we use the Canberra metric \cite{Jurman2009CanberraLists}, described on the equation \ref{eq:canberra}.
-
-\begin{equation}
-    d(p,q)= \sum_{i=1}^{n}{\frac{|p_i-q_i|}{|p_i|+|q_i|}}
-    \label{eq:canberra}
-\end{equation}
-
-Where $p$ and $q$ are n-dimensional vectors.
-
-We compute the input size for all 30 applications of the Polybench with 3 different inputs.
-Running each program 15 times, with a sampling rate of 0.01 seconds collecting the total number of instruction, number of loads and write instructions, number of floating point operations, besides some software counters.
-After applying the post-processing to the data collected we compute the input size and perform a Hierarchical clustering using the linkage method of Ward \cite{Murtagh2011WardsAlgorithm}, that minimizes the total within-cluster variance.
-The results of the clustering can be seen on the figure \ref{fig:sping_force_is} and on the dendrogram on figure \ref{fig:dendograma_input_size}.
-
-%The exact HPCs was:
-%PERF_COUNT_HW_INSTRUCTIONS,
-%MEM_UOPS_RETIRED:ALL_LOADS,
-%MEM_UOPS_RETIRED:ALL_STORES,
-%FP_ARITH_INST_RETIRED:SCALAR
-
-\begin{figure*}[h]
-    \centering
-    \includegraphics[width=\textwidth]{fingerprint/figures/dendograma_input_size.png}
-    \caption{Dendrogram}
-    \label{fig:dendograma_input_size}
-\end{figure*}
-
-From this dendrogram, we can have an idea of how close two applications are.
-We choose the number of clusters that maximize the number of hits of the same program with different input sizes on the same cluster, in this case, was 5 clusters.
-
-It was observed that as input size grows the program behavior tends to a specific curve, but for small input sizes some have variation, so in some cases, the same program has been classified in more than one cluster.
-This can also happen if specific parts of the program are triggered with specific inputs, in which case it will also belong to more than one cluster.
-
-In order to have an overall classification of each program, we can pick up the frequency in which each appeared in the clusters and classified it in the cluster in which it appeared more often.
-In this case, the clusters are:
-
-\begin{itemize}
-    \item Cluster 1: 2mm, 3mm, cholesky, correlation, covariance, floyd-warshall, gemm, gramschmidt, lu, ludcmp, nussinov, symm
-
-    \item Cluster 2: deriche, doitgen, syrk
-
-    \item Cluster 3: adi, fdtd-2d, jacobi-2d, syr2k
-
-    \item Cluster 4: atax, bicg, durbin, gemver, gesummv, mvt, trisolv, trmm
-
-    \item Cluster 5: heat-3d, seidel-2d
-\end{itemize}
-
-On the figure, \ref{fig:sping_force_is} we display the clusters using the spring force algorithm where each program with a specific input is a node and the edge weight is the Canberra distance.
-To a better visualization, the name of the inputs was replaced by numbers, the EXTRALARGE is 3, LARGE 2 and MEDIUM 1.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{fingerprint/figures/graph_input_size.png}
-    \caption{Spring force of Canberra distance}
-    \label{fig:sping_force_is}
-\end{figure}
-
-The figure above shows a different way to observe in a reduced space the distance between clusters and applications and how they are organized.
-From this graph, we see that clustering it's well partitioned and we can clearly separate each cluster.
-It is also interesting to note that the applications of the cluster formed by the circle symbol are more separated, which may indicate that they were classified in this way because they did not fit into any other cluster.
-
-% \begin{figure*}[t]
-%     \begin{subfigure}{\textwidth}
-%         \includegraphics[width=\textwidth]{fingerprint/figures/dendograma_input_size.png}
-%     \end{subfigure}
-%     \\
-%     \begin{subfigure}{\textwidth}
-%         \includegraphics[width=\textwidth]{fingerprint/figures/dendograma_floating.png}
-%     \end{subfigure}
-%     %and so on
-% \end{figure*}
-
-To have an idea of how the behavior of the variable input size for each clusters is, it was plotted for the applications of clusters 1 and 2 shown on the figures \ref{fig:c1_input_size_0}, \ref{fig:c1_input_size_1}.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{fingerprint/figures/cluster_input_0.png}
-    \caption{Input size - Cluster 1}
-    \label{fig:c1_input_size_0}
-\end{figure}
-
-From these figures, we can have an idea of what behavior was classified as the same class.
-On cluster 1 on the figure \ref{fig:c1_input_size_0} all applications showed the same behavior with approximately the same amplitude.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{fingerprint/figures/cluster_input_1.png}
-    \caption{Input size - Cluster 2}
-    \label{fig:c1_input_size_1}
-\end{figure}
-
-On figure \ref{fig:c1_input_size_1} we observe that curves that have similar shape regardless of scale on vertical and horizontal axis have also been classified as the same clusters.
-This is the wanted comportment for the classification because we are only interested in the overall shape of the curve.
-
-% \begin{figure}[H]
-%     \centering
-%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_input_4.png}
-%     \caption{Input size - Cluster 3}
-%     \label{fig:c1_input_size_2}
-% \end{figure}
-
-% To conclude this study we also clustering applications by its floating point behavior.
-For that we only use the counter correspondent to floating point operations.
-The figure \ref{fig:dendogram_fp} shows the dendrogram and the \ref{fig:sping_force_fp} the spring force graph plot.
-
-% \begin{figure*}[h]
-%     \includegraphics[width=\textwidth]{fingerprint/figuresdendograma_floating.png}
-%     \caption{Dendrogram}
-%     \label{fig:dendogram_fp}
-% \end{figure*}
-
-% \begin{figure}[H]
-%     \centering
-%     \includegraphics[width=\textwidth]{fingerprint/figures/graph_floating2.png}
-%     \caption{Spring force of Canberra distance}
-%     \label{fig:sping_force_fp}
-% \end{figure}
-
-% The figures \ref{fig:cluster_fp1},\ref{fig:cluster_fp2},\ref{fig:cluster_fp3} show the behavior on the same cluster for the float point operations.
-For this classification the number of clusters was 8.
-
-% \begin{figure}[H]
-%     \centering
-%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_fp_0.png}
-%     \caption{Floating point - Cluster 1}
-%     \label{fig:cluster_fp1}
-% \end{figure}
-
-% \begin{figure}[H]
-%     \centering
-%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_fp_4.png}
-%     \caption{Floating point - Cluster 2}
-%     \label{fig:cluster_fp2}
-% \end{figure}
-
-% \begin{figure}[H]
-%     \centering
-%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_fp_5.png}
-%     \caption{Floating point - Cluster 3}
-%     \label{fig:cluster_fp3}
-% \end{figure}
-
-% Here we can also observed again that applications with the same shape were classified as the same clusters.
-
-\subsection{CONCLUSIONS}
-
-%\cite{Processors2012, Gregg2017}
-From the results, we observe that the API present overhead similar or lower to other low-level APIs, with the advantage of being in high abstraction and simplified configuration with a few lines of code, is possible to configure and gather counter data.
-
-The tool developed provide also provided a way to fingerprint programs and compute similarities between different programs or the same program with different inputs.
-This can be useful to reduce applications spaces for benchmarks as was done in Polybench clustering but also to analyze the behavior of a parameter providing insights to the programmer to find a possible bottleneck.
-
-We also provide a precise definition to input size that made possible fingerprint the programs, but it also can help programmers of benchmark applications to better create inputs with more precise growth of a particular parameter.
-
-%\subsection{FUTURE WORK}
-%
-%We pretend to use this tool to create a data set of applications behavior and automatically classify any program as belonging to a cluster or a set of clusters.
-The idea is to have a set of clusters that can describe most applications in this way we can know specific behaviors of the applications.
-This can be applied to various areas such as voltage scaling and frequency of the processor, once knowing the behavior of a particular program we can have an idea of what is the best strategy to control the frequency in order to save energy or increase performance.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{APPENDIX}
-
-% \begin{figure*}[h]
-%     \centering
-%     \includegraphics[width=\textwidth]{dendograma_input_size.png}
-%     \caption{Dendogram input size}
-%     \label{fig:dendograma_input_size}
-% \end{figure*}
-
-% \begin{figure*}[h]
-%     \centering
-%     \includegraphics[width=\textwidth]{dendograma_floating.png}
-%     \caption{Dendogram floating}
-%     \label{fig:dendograma_floating}
-% \end{figure*}
-
-This tool and all the analysis are available on github: \url{github.com/VitorRamos/performance_features}
-
-% \begin{lstlisting}[language=bash]
-% pip install performance-features
-% \end{lstlisting}
-
-%\subsection{ACKNOWLEDGMENT}
-%
-%This research was developed in the super-computing centers of the University of MONS in collaboration with the center of the Federal University of Rio Grande do Norte.

### 2022-04-05 thesis/intro/intro.tex
add journals and restructured

--- 2021-03-23 thesis/intro/intro.tex
@@
-\section{Motivation}
-The energy efficiency of data centers has attained key importance in recent years, due to its high economic, environmental, and performance impact. 
-Data center consumption was estimated in 2010 at between 1.1\% and 1.5\% of the worldwide electricity usage \cite{Dayarathna2016DataSurvey,Corcoran2017EmergingICT}, generating as much pollution as a nation like Argentina  \cite{Mathew2012Energy-awareNetworks}.
-Power costs exceed in some cases the cost of purchasing hardware \cite{Rivoire2007ModelsOptimizations}. 
-Nowadays, the leading petaflop supercomputers consume a range of 1–18 MW of electrical power, with 1.5 MW on average, which can be easily translated into millions of dollars per year in electricity bills \cite{Group2012HandbookSahni}.
-Furthermore, the energy costs of powering a typical data center doubles every five years \cite{Buyya2013MasteringProgramming}.
-Therefore, with such steep increase in electricity use and rising electricity costs, electricity bills have become a major expense for today’s data centers \cite{Poess2008EnergyResults}, \cite{Gao2013QualityCenters}.
-Due to these reasons, data center energy efficiency is now considered a chief concern for data center operators, often ahead of the traditional considerations of availability and security.
+\section{Motivation} \label{sec:motivation}
+Data center energy efficiency has become of crucial importance in recent years due to its high economic, environmental, and performance impact.
+For example, the leading petaflop supercomputers consume a range of 1–18 MW of electrical power, with 1.5 MW on average, which can be easily translated into millions of dollars per year in electricity bills \cite{Group2012HandbookSahni}.
+Data center energy consumption was estimated to be between 1.1\% and 1.5\% of worldwide electricity usage in 2010 \cite{Dayarathna2016DataSurvey,Corcoran2017EmergingICT}, generating as much pollution as a nation such as Argentina  \cite{Mathew2012Energy-awareNetworks}.
+In some cases, the power costs exceed the cost of purchasing hardware~\cite{Rivoire2007ModelsOptimizations}.
+Furthermore, the energy costs of powering a typical data center doubles every five years~\cite{Buyya2013Introduction}.
+Therefore, with such a steep increase in power use, electricity bills have become a significant expense for today's data centers \cite{Poess2008EnergyCenters,Gao2013QualityCenters}. 
+For these reasons, data center energy efficiency is now considered a primary concern for data center operators, often ahead of the traditional considerations of availability and security.
 
 There are several approaches for green computing, from electrical materials to circuit design, systems integration, and software.
-These techniques may be different, but they share the same goal: substantially reduce overall system energy consumption without a corresponding negative impact on delivered performance.
-%From a systems-software perspective, the strategy of supplying minimal power necessary to accomplish the work can be accomplished by powering down system components when they are underutilized \cite{Mathew2012Energy-awareNetworks}.
-From a software systems perspective, the strategy of providing just enough power to get a job done can be accomplished by turning off system components when they are underutilized \cite{Mathew2012Energy-awareNetworks}.
-%The processor and main memory are usually in the top three power draw components, which dominate energy consumption as shown in figure \ref{fig:powerbreakdown}. 
-The processor and main memory are usually the components that dominate power consumption, as shown in Figure  \ref{fig:powerbreakdown}.
-The processor can consume as much as 50\% of the total energy \cite{Fan2007, Barroso2007TheComputing, Malladi2012TowardsDRAM}.
-For that reason, modern processors incorporate several features for power management \cite{Rotem2012Power-managementBridge, Brown2005, Hackenberg2015}, such as Dynamic Power Management (DPM) and Dynamic Voltage and Frequency Scaling (DVFS).
-DPM encompasses a set of techniques for obtaining an energy-efficient computing by deactivating or reducing the performance of system components when they are idle or used partially \cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.
+These techniques may differ, but they share the same goal---to substantially reduce overall system energy consumption without a corresponding negative impact on delivered performance.
+The processor and main memory are  the components that usually dominate power consumption, as shown in \cref{fig:powerbreakdown}.
+The processor can consume as much as 50\% of the total energy~\cite{Fan2007PowerComputer, Barroso2007TheComputing, Malladi2012TowardsDRAM}.
+For that reason, modern processors incorporate several features for power management~\cite{Rotem2012Power-managementBridge, Brown2005ACPILinux, Hackenberg2015AnProcessor, Intel20200thLake}, such as dynamic power management (DPM) and dynamic voltage and frequency scaling (DVFS).
+DPM encompasses a set of techniques for obtaining energy-efficient computing by deactivating or reducing the system components' performance when they are idle or partially utilized~\mbox{\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}}.
 DVFS allows the frequency and voltage to be adjusted in run-time depending on current needs.
 
-\begin{figure}[ht]
+\begin{figure}[H]
 	\centering
+	\captionsetup[subfigure]{justification=centering}
+
 	\begin{subfigure}[b]{0.45\textwidth}
 		\includegraphics[width=\textwidth]{models/figures/power_breakdown/Xeon E5-2698 + DDR4 + SSD.png}
-		\caption{This work}
-		\label{fig:1}
+		\caption{}
+		\label{fig:powerbreakdown_a}
 	\end{subfigure}
@@
 		\includegraphics[width=\textwidth]{models/figures/power_breakdown/Xeon E5507 + DDR3 + HDD.png}
-		\caption{Study case in \cite{Malladi2012TowardsDRAM}}
-		\label{fig:2}
+		\caption{}
+		\label{fig:powerbreakdown_b}
 	\end{subfigure}
@@
 	\caption{Power breakdown of a typical node of an HPC cluster at full use.
-The system used in this work (a) was built in 2016 and equipped with two Intel Xeon E5-2698, 128 GB of DDR4 memory and SSD as storage, while (b) the case study in \cite{Malladi2012TowardsDRAM} was built in 2012 and equipped with two Xeon E5507, 32GB of DDR3 memory and HDD as storage}
+The system used in this study (\textbf{a}) was built in 2016 and equipped with two Intel Xeon E5-2698, 128 GB of DDR4 memory and SSD as storage, while (\textbf{b}) the case study in \cite{Malladi2012TowardsDRAM} was built in 2012 and equipped with two Xeon E5507, 32GB of DDR3 memory and HDD as storage.}
 	\label{fig:powerbreakdown}
@@
 
-DVFS is motivated by the well-known fact that frequency and power have a near-cubic relationship \cite{Dayarathna2016DataSurvey, Group2012HandbookSahni}. 
-This implies that running the CPU at lower frequency causes a linear reduction in the performance while a near-cubic reduction in power and, consequently, it could lead to a near-square reduction in CPU energy.
-This can lead to expressive energy savings depending on the system and its infrastructure.
+DVFS is motivated by the well-known fact that frequency and power have a near-cubic relationship \cite{Dayarathna2016DataSurvey, Group2012HandbookSahni}; this implies that running the CPU at a lower frequency causes a linear reduction in performance and a near-cubic reduction in power, which could lead to a near-square reduction in CPU energy.
+Because of this, it is possible to achieve dramatic energy savings just with frequency control, depending on the system and its architecture.
 Although very promising, the system software has yet to determine when and what voltage and frequency to use when running applications.
-Otherwise, not only will performance deteriorate, but in the worst case, energy consumption would also increase  \cite{Group2012HandbookSahni}.
-%Otherwise, in the worst case, not only performance will degrade, but also energy consumption may increase \cite{Group2012HandbookSahni}. 
-Indeed, reducing the frequency results in a longer execution time, which in turn increases the energy consumption of other system components such as memory and disks.
-%This is because reducing the frequency results in longer run time which in turn increases the energy consumption of the other components in the system such as memory and disks. 
-There is also an overhead of time and energy associated with a voltage and frequency switch.
-%There are also overheads of time and energy associated with a voltage and frequency switch.
-Thus, finding the most appropriate voltage and frequency to use at all times is not an easy task.
-%Finding the most appropriate voltage and frequency to use at each time instant is not an easy task. 
-Therefore, since its introduction in 1994, there has been a lot of research on DVFS algorithms.
-%Since its introduction in 1994, there has been a huge amount of work on DVFS algorithms in the literature.
+Otherwise, not only will performance deteriorate, but, in the worst case, energy consumption would also increase  \cite{Group2012HandbookSahni}.
+Indeed, reducing the frequency results in a longer execution time, which increases the energy consumption of other system components, such as memory and disks.
+There is also an overhead of time and energy associated with a voltage and frequency switch that needs to be considered.
+Thus, finding the most appropriate voltage and frequency to use in all circumstances is not easy.
+Therefore, since its introduction in 1994 \cite{Group2012HandbookSahni}, there has been a tremendous amount of research on DVFS algorithms.
 
-The DPM technique can achieve very large energy savings on systems where the static power is high or when the system remains inactive for a long period.
+The DPM technique can achieve substantial energy savings on systems where the static power is high, or the system remains inactive for a long time.
 In that case, the problem is to determine when and which components to turn on/off.
-With DPM, energy savings of 70\% have been reported ~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.
-%In the case of DPM, the problem is to determine when and which components to turn on/off.
-This technique can archive very large energy savings on systems where the static power consumption is high or if the system stays idle for a long time.
-Power savings of 70\% have been reported~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.
+With DPM, energy savings of 70\% have been reported~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.
 
-At the same time as the stated techniques reduce the energy consumption of the system, they also lead to a complex trade-off concerning performance, which must be exploited to produce more energy-efficient algorithms.
-%However, at the same time that these power-saving techniques reduce system energy, they also lead to a complex balance between energy savings and high performance, exposing trade-offs between performance and power consumption, which can be exploited to produce more energy-efficient algorithms.
-%in the case of the CPU, this can be applied to each core.
-Here...
-Indeed, this work investigates whether the construction of a model of the energy consumption of an application can lead to significant energy savings.
-We propose an analytical energy model of an application in function of the two control variables present in most of HPC systems: CPU operating frequency and number of active cores.
-This model can serve as a base for new DVFS and DPM algorithms, as well as for the analysis of the contribution of each model parameter to the total energy consumption.
-The model equation includes the two parameters of the application and three parameters of the system.
-The application parameters incorporate characteristics of the percentage of parallelism and the size of the input.
-The system parameters include power-related and technology-dependent components that compose dynamic, static and leakage power.
+However, at the same time, while these power-saving techniques reduce system energy, they can compromise  performance leading  to a complex trade-off that needs to be carefully exploited to produce more energy-efficient algorithms.
+Indeed, this study investigates whether the construction of an energy consumption model of an application can lead to significant energy savings.
 
-The rest of this paper is divided as follows.
-In Section \ref{sec:relatedwork}, a general review of models is provided, showing the difference between each approach and its applications.
-In Section \ref{sec:models}, the models and its parameters are derived alongside its constraints.
-In section \ref{sec:experimentalvalidation}, the model is validated with the PARSEC benchmark applications.
-Forward, in Section \ref{sec:dvfs_optmin}, use cases of the model are presented, how it could be applied in DVFS and also analysis of the best saving energy policy depending on the system and the effect of the parallelism in the energy.
-Finally, the conclusion is make in Section \ref{sec:conclusion}.
+We propose an analytical energy model for a given application in the function of the two control variables present in most HPC systems: CPU operating frequency and number of active cores.
+The model is composed of three application-dependent parameters and three parameters relating to the architecture of the system.
+The application parameters incorporate characteristics of the percentage of parallelism and the input size.
+The system architecture parameters include power-related and technology-dependent components, such as dynamic, static, and leakage power.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Theoretical Background} \label{sec:theoretical_background}
 
-%\section{Related Work} \label{sec:relatedwork}
-%
-%A model is a formal abstraction of a real system.
-Models for computer systems can be represented as equations, graphical models, rules, decision trees, sets of representative examples, neural networks, etc.
-The choice of representation affects the accuracy of the models, as well as their interpretability by people \cite{Hypothesis2012EncyclopediaLearning}.
-Accurate energy and power consumption models are very important for many energy efficiency schemes employed in computing equipment \cite{Rivoire2007ModelsOptimizations} and they can have multiple uses, including the design, forecasting, and optimization of data center systems.
-This work focuses on analytical models that could serve for energy optimization and subsequently analysis of important factors in the total energy draw.
-%
-%The desirable properties of a full-system energy consumption model include accuracy (accurate enough to allow the desired energy saving), speed (generate predictions quickly enough), generality and portability (should be suitable for as many systems as possible), inexpensiveness (should not requires expensive or intrusive infrastructure), and simplicity \cite{Rivoire2008AModels}.
-However, modeling the exact energy consumption behavior of an HPC system is not straightforward, either at the whole-system level or at the individual component.
-Patterns of energy consumption in data centers depend on multiple factors such as hardware specifications, workload, cooling requirements, types of applications, etc.
+A model is a formal representation of a natural system.
+The representation of computer system models includes equations, graphical models, rules, decision trees, representative collections of examples, and neural networks.
+The choice of representation affects the model's accuracy, as well as its interpretability by people~\cite{Hypothesis2012EncyclopediaLearning, Roy2019ForecastingNetwork, Zhu2019PredictingLearning}.
+Accurate energy and power consumption models are essential for many energy efficiency schemes employed in computing equipment \cite{Rivoire2007ModelsOptimizations}, and they can have multiple uses, including the design, forecasting, and optimization of data center systems.
+This study focuses on analytical models that could aid energy optimization and analyses of crucial factors in the total energy draw.
+
+The desirable properties of a full-system model of energy consumption include accuracy, speed, generality and portability, inexpensiveness, and simplicity \cite{Rivoire2008AModels}.
+However, modeling an HPC system's exact energy consumption behavior is not straightforward, either at the whole-system level or at the level of individual components.
+Data centers' energy consumption patterns depend on multiple factors, such as hardware specifications, workload, cooling requirements, or the type of the applications.
 Some of these factors cannot be measured easily.
-Furthermore, it is impractical to perform detailed measurements of the energy consumption of lower-level components without introducing overhead to the system.
-%
-%The definition of energy (E) is the total amount of work performed by a system over a time period (T) while power (P) is the rate at which the work is performed by the system.
-The relationship between these three quantities can be expressed as:
-%
-%\begin{equation}
-%E = \int_{0}^{T}P(t)dt
-%\label{qe:energy_definition_cont}
-%\end{equation}
-%
-%%Or in the discrete format, as the sum of energy in time windows $t_i$.
-%
-%%\begin{equation}
-%%    E = \sum_{i}t_iP_i
-%%    \label{qe:energy_definition_discrete}
-%%\end{equation}
-%
-%As a starting point, the models were classified with respect to his inputs, based on the survey \cite{Dayarathna2016DataSurvey}, where they analyzed more than 200 models pointing their characteristics and limitations, it was possible to divide the models into the categories where is more suitable to drive their purposes.
-%According to a survey over 200 models~\cite{Dayarathna2016DataSurvey} that analyzed their characteristics and limitations, they can be divided into categories that are are more suitable to drive their purpose:
-%\begin{itemize}
-%	\item Temperature
-%	\item System Utilization or Workload
-%	\item Frequency
-%	\item Other system states such as: cache miss, branch prediction, number of instructions executed, and more
-%	\label{tab:input_type}
-%\end{itemize}
-%
-%%maybe split into power and performance models.
-%Energy models are often described as a combination of two models, the power model of the system and the performance model of the application.
-%
-%\subsection{Power models}
-%Temperature-based models can provide a way of estimating power consumption without being very intrusive.
-With this kind of model it is possible to estimate energy with a single temperature sensor without the need to restart or introduce anything to the system.
-This is especially interesting for HPCs and data centers which can not be easily turned off.
-Another use of this model is in optimization of the refrigeration systems.%, but this will be not covered in this journal.
-%
-%A large part of the developed models are based on system states.
-These types of models normally take advantage of performance counters, which are provided by the CPU or the operating system and are capable of counting micro-architectural events, such as: instructions executed, cache-hits, miss-predicted branches, and much more.
-This kind of model is great for power estimation because they have information about several internal states of the computer.
-It is possible to achieve accurate predictions within 2\% of error \cite{Joseph2001Run-timeMicroprocessors} in average.
-However, performance counters are highly dependent on the system and architecture, making them not so easily portable.
-Also, there are limitations on how many counters can be read simultaneously in the hardware side.
-The CPU only allows a small number of counters (normally five) to be read at the same time.
-If more counters are requested, some events will be multiplexed, causing measurement loss.
-From the software side, measuring a large number of system counters can introduce a significant overhead.
-%
-%Additionally, hardware performance counters had shown other problems \cite{Weaver2008, Weaver2013a, Das2019SoK:Security, McGuire2009, Ramos2019AnCounters}.
-Events that should be exact and deterministic (such as the number of executed instructions) show run-to-run variations and overcounting on x86\_64 machines, even when running in strictly controlled environments.
-Although this does not seem to be a problem when performing energy estimation---since normally the models have configurable weights or parameters that can account for imprecision, they might not be fully reliable.
-%
-%%include utilization
-%
-%Frequency-based models serve as base for many power models because at a very low level every digital circuit (including modern processors) is composed of transistors.
-In these devices, there is a known relationship between power and frequency, which can be simplified by equation \ref{eq:power_simplified}.
-%\begin{equation}
-%P = \alpha+\beta f^3,
-%\label{eq:power_simplified}
-%\end{equation}
-%where $\alpha$ and $\beta$ are model parameters and $f$ is the operating frequency. This equation is presented in more detail in Section~\ref{sec:models}.
-%
-%These models are good for the optimization problems since it is in function of a variable that can be easily controlled: $f$.
-%%, although it is highly dependent on the load as shown in figure \ref{fig:cpuload}.
-%
-%% \begin{figure}
-%%     \centering
-%%     \includegraphics[width=0.5\columnwidth]{cpu_load.png}
-%%     \caption{Power with a fixed frequency for different CPU loads, the load means the amount of time that the CPU was busy in the given time windows}
-%%     \label{fig:cpuload}
-%% \end{figure}
-%
-%Hybrid models can also be found in the literature.
-They are models that use as input frequency and CPU utilization or system states and frequency.
-In \cite{Horyath2008Multi-modeClusters}, they combine frequency and CPU utilization to obtain an average error of 1\% such as:
-%\begin{equation}
-%P = a_3fu+a_2f+a_1u+a_0,
-%\label{eq:hybrid}
-%\end{equation}
-%where $u$ is CPU utilization and $a_0$, $a_1$, $a_2$, $a_3$ are model parameters.
-%
-%\subsection{Performance models}
-%Another very common way to model the energy consumption of the system is with the workload.
-The workload is an abstract definition of the application.
-It represents the amount of work done in a given time and speed \ref{eq:workload_definition}.
-Workload can be defined as~\cite{Paolillo2018OptimisationParallelism, Group2012HandbookSahni, Kim2015RacingHeuristics}:
-%\begin{equation}
-%W = \int_{0}^{T_{\rm active}}s(t)dt = sT_{\rm active},
-%\label{eq:workload_definition}
-%\end{equation}
-%where $T_{\rm active}$ is total active time, and $s$ is the speed of execution.
-This kind of model is often translated into utilization-based models so that it is possible to work it in real systems.
-%% The definition of utilization is the ration of the time that the system was active. One possible transition is defined bellow in equation \ref{eq:workload_utilization_tranlation}.
-%Utilization can be defined as the ratio between the time that the system is active and the total time. Equation \ref{eq:workload_utilization_tranlation} defines workload in terms of utilization.
-%\begin{equation}
-%u = \frac{T_{\rm active}}{T_{\rm total}} = \frac{w/s}{T_{\rm total}},
-%\label{eq:utilization_definition}
-%\end{equation}
-%\begin{equation}
-%w = usT_{\rm total}.
-%\label{eq:workload_utilization_tranlation}
-%\end{equation}
-%
-%Models based on CPU utilization are the bases for DVFS algorithms.
-Even though the input is not controllable, it is very easy to measure system utilization with almost no overhead.
-It is also very portable.
-There is no need for training and it often works reasonably in any machine.
-A downside is that the theory behind the energy consumption optimization relies on the knowledge of the future workload \cite{Fu2018RaceMinimization, Group2012HandbookSahni}.
-This introduces the problem of how to estimate the workload.
-%
-%%This models can be forward subdivided into  schedules, ...
-%
-%\subsection{Online vs Offline models}
+Furthermore, it is impractical to perform detailed measurements of the energy consumption of lower-level components without additional overhead.
 
-\section{Theoretical background} \label{sec:background}
-
-A model is a formal representation of a real system.
-Computer system models can be represented in the form of equations, graphical models, rules, decision trees, representative collections of examples, or by neural networks, among others.
-The choice of representation affects the model's accuracy, as well as its interpretability by people~\cite{Hypothesis2012EncyclopediaLearning, Roy2019ForecastingNetwork, Zhu2019PredictingLearning}.
-Accurate energy and power consumption models are very important for many energy efficiency schemes employed in computing equipment \cite{Rivoire2007ModelsOptimizations}, and they can have multiple uses, including the design, forecasting, and optimization of data center systems.
-This work focuses on analytical models that could serve for energy optimization and subsequently, analysis of important factors in the total energy draw.
-
-The desirable properties of a full-system model of energy consumption include accuracy (precision enough to allow the required energy saving), speed (fast enough to produce predictions), generality and portability (should be suitable for as many systems as possible), inexpensiveness (should not require costly or invasive infrastructure), and simplicity \cite{Rivoire2008AModels}.
-However, modeling the exact energy consumption behavior of an HPC system is not straightforward, either at the whole-system level or at the level of individual components.
-Patterns of energy consumption in data centers depend on multiple factors such as hardware specifications, workload, cooling requirements, or types of applications.
-Some of these factors cannot be measured easily.
-Furthermore, it is impractical to perform detailed measurements of the energy consumption of lower-level components without impacting the system with additional overload.
-
-The concept of energy (E) is the total amount of work performed by a system over a period of time (T), while power (P) is the rate at which the system performs the work.
-The relation between these three amounts can be expressed as:
-
-\begin{equation}
-E = \int_{0}^{T}P(t)dt
-\label{qe:energy_definition_cont}
-\end{equation}
-
-Several proposed models have already been classified with respect to its input parameters in \cite{Dayarathna2016DataSurvey}.
-In this survey, more than 200 models were analyzed according to their characteristics and limitations and classified into categories where the model is more suited to its objectives:
+Several proposed models have already been classified concerning their input parameters, as shown by Dayarathna et al.~\cite{Dayarathna2016DataSurvey}, who analyzed more than 200 models according to their characteristics and limitations and classified them into categories where the model is more suited to its objectives:
 \begin{itemize}
-	\item Temperature
-	\item System Utilization or Workload
+	%\item Temperature
+	\item System utilization or workload
 	\item Frequency
-	\item Other system states such as cache miss, branch prediction, number of instructions executed, and more
+	\item Other system states, such as cache miss, branch prediction, number of instructions executed, and more
 	\label{tab:input_type}
@@
 Often, energy models are described as a combination of two main parts, the power model of the system and the performance model of the application.
+This is because the concept of energy (E) is the total amount of work performed by a system over a period of time (T), while power (P) is the rate at which the system performs the work.
+The relation between these three amounts can be expressed as:
+\begin{equation}
+	E = \int_{0}^{T}P(t)dt.
+	\label{eq:energy_definition_cont}
+\end{equation}
 
-\subsection{Power models}
-Temperature-based models\cite{ZhangEnergyInfrastructure} can provide a way of estimating power consumption without being very intrusive.
-With this kind of model, it is possible to estimate energy with a single temperature sensor without the need to restart or introduce anything to the system.
-This is especially interesting for HPCs and data centers which, can not be easily turned off.
-Another use of this model is in the optimization of the refrigeration systems.
+\subsection{Power Models}
 
-A large part of the developed models are based on parameters representing the state of the system.
-They normally take advantage of performance counters, which are provided by the CPU or the operating system and are capable of counting micro-architectural events, such as instructions executed, cache-hits, miss-predicted branches, and much more.
-This type of model is suitable for power estimation because it receives information about several internal states of the computer.
+The modeling of system parameters is becoming popular nowadays with the advantage of performance counters provided by the CPU or the operating system.
+These counters can measure micro-architectural events, such as instructions executed, cache hits, miss-predicted branches, and more; thus, providing a base for many different estimations of power usage.
+This makes this type of model very suitable for power estimation because it can use information about several internal states of the computer.
 
-Frequency-based models serve as a base for many power models ~\cite{Sarwar1997, Butzen2007, Usman2013ANoC} because at a very low level every digital circuit (including modern processors) is composed of transistors.
-In these devices, there is a known relationship between power and frequency, which can be simplified by \cref{eq:power_simplified}:
+Frequency-based models are the most common kind of model.
+They serve as a base for many power models~\cite{Sarwar1997CmosCalculation, Butzen2007LeakageGates, Usman2013ANoC}.
+These models utilize the fact that every digital circuit (including modern processors) is composed of transistors.
+Thus, modeling one transistor's interaction and scaling this to the chip can give a reasonable estimate of the entire system's energy.
+One of the most common frequency-based %please confirm that the intended meaning has been retained.
+model approximations is defined as follows: 
 \begin{equation}
-P = \alpha+\beta f^3
+	P = \alpha+\beta f^3,
 	\label{eq:power_simplified}
 \end{equation}
-where $\alpha$ and $\beta$ are model parameters, and $f$ is the operating frequency (details of this equation are covered in Section~\ref{sec:models}).
-These types of models are suitable for optimization problems since they are a function of the operating frequency, which can be easily controlled.
+where $\alpha$ and $\beta$ are model parameters, and $f$ is the operating frequency (details of this equation are covered in (\cref{chapter:models}).
+This type of model is suitable for optimization problems since these are a function of the operating frequency, which can be easily controlled.
 
-\subsection{Performance models}
-Another commonly used way to model the energy consumption of the system is the workload.
-The workload, which represents the amount of work done in a given time and speed, abstractly defines the application.
+% We confirm that the meaning has been retained
 
-The workload ($W$) can be defined as~\cite{Paolillo2018OptimisationParallelism, Group2012HandbookSahni, Kim2015RacingHeuristics}:
+\subsection{Performance Models}
+The most common way to model the application performance is using the workload.
+The workload is an abstract representation of the amount of work done for a given time and speed.
+The workload ($W$) can be defined in many different ways.
+One common way, used in many studies, such as Paolillo et al.
+~\cite{Paolillo2018OptimisationParallelism}, Francis et al.
+\cite{ Group2012HandbookSahni}, and Kim et al.
+\cite{Kim2015RacingHeuristics}, is the following:
 \begin{equation}
-W = \int_{0}^{\tau}s(t)dt = s\tau
+	W = \int_{0}^{\tau}s(t)dt = s\tau,
 	\label{eq:workload_definition}
@@
 where $\tau$ is total active time, and $s$ is the execution speed in instructions/second.
-This equation is often taken up by utilization-based models, so that it could be applied to real systems.
 
-Utilization can be defined as the ratio between the time that the system is active and the total time.
-Equation (\ref{eq:workload_utilization_tranlation}) defines workload in terms of CPU utilization ($u$):
+Utilization models \cite{Fu2018RaceMinimization, Group2012HandbookSahni} are also found in the literature, defined as the ratio between the time that the system is active and the total time (idle and active).
+These models are present in many DVFS algorithms present in Linux.
+They can be viewed as a good alternative to the workload since it is impossible to measure workload in real-time.
+\cref{eq:utilization_definition} defines workload in terms of CPU utilization ($u$):
 \begin{equation}
@@
 \end{equation}
-\begin{equation}
-W = usT,
-\label{eq:workload_utilization_tranlation}
-\end{equation}
-where $T$ the total execution time of the workload, and $s$ the execution speed.
-Models based on CPU utilization are the base for DVFS algorithms.
-Even though the input is not controllable, it is straightforward  to measure system utilization with almost no overhead, and it is also very portable in terms of operating systems and architectures.
-%% Here is how to build the model or to optimize it, maybe too early to talk about AI without referencing existing approaches:
-%There is no need for data and training, as with artificial intelligence algorithms, and adapts reasonably well to any machine. 
-A downside is that optimization requires the knowledge of a wide variety of workloads which must be estimated somehow \cite{Fu2018RaceMinimization, Group2012HandbookSahni}.
+where $T$ is the total execution time (idle and active), and $\tau$ is the active time, meaning when the processor was executing instructions.
+Models based on CPU utilization are the basis for DVFS algorithms.
+Even though this is not a controllable parameter, it is straightforward to measure system utilization with almost no overhead, and it is also very portable in terms of operating systems and architectures.
 
-% Markov model
-Other approaches are also adopted to model energy, as for example in the work of Liu et al.
-\cite{Liu2016IntelligentSupplies} where the model uses a Markov chain to determine energy state transitions.
-This kind of approach needs heuristic algorithms to search for the best transitions and does not provide insights as an analytical equation does.
-
-Although much work has been done in DVFS, the focus is still on the consumer electronics and laptop markets. 
-For HPC, the notion of energy perception is relatively new~\cite{Beckman2005MakingSupercomputing}. 
-Moreover, the operational characteristics of non-HPC and HPC systems are significantly different. 
-First, the workload on non-HPC systems is very interactive with the end-user, but the workload on the HPC platform is not. 
-Second, activities conducted on a non-HPC platform tend to share more machine resources.
-In contrast, in HPC each job often runs with dedicated resources. 
-Third, an HPC system usually is much larger than non-HPC systems, making it more  challenging to gather information, organize decisions, and execute global decisions.
-Therefore, it is worthwhile to investigate whether a DVFS scheduling algorithm, which works well for conventional computing, remains effective for HPC.
-
-\section{Related work} \label{sec:relatedwork}
+\section{Related work} \label{sec:related_work}
 
 Merkel et al.~\cite{Merkel2006BalancingSystems} developed an energy model for processors based on events.
-By counting a set of $n$ events and assuming that the processor consumes a fixed amount of energy $\alpha_i$ for each activity, they estimate the energy in a particular period as the number of occurrences $c_i$  multiplied by its correspondent weight $\alpha_i$:
+Their model assumes a fixed energy consumption $\alpha_i$ for each activity, and by counting the number of occurrences $c_i$ of every activity, they estimate the total energy as:
 \begin{equation}
 	E = \sum_{i=1}^{n}\alpha_ic_i.
+	\label{equation:rw_merkel}
 \end{equation}
 
-Another event-based model introduced by Roy et al. \cite{Roy2013AnAlgorithms}, described the computational energy consumed by a CPU for an algorithm $A$:
+Another event-based model, introduced by Roy et al.
+\cite{Roy2013AnAlgorithms}, described the computational energy consumed by a CPU for an algorithm $A$ as the Equation (\ref{equation:rw_event_based})%\cref{equation:rw_event_based}:
 \begin{equation}
 	E(A) = P_{clk}T(A) + P_wW(A),
+	\label{equation:rw_event_based}
 \end{equation}
@@
 
-Models based on events are implemented using performance counters \cite{Fahad2019AComputing}, which have some drawbacks.
-They are highly dependent on the operating system and its architecture, making them not so easily portable.
-Additionally, there are limitations on the number of counters that can be read simultaneously.
-Furthermore, reading too many counters can add a non-negligible overhead.
-Thus, modern CPUs start to multiplex counters when more than a few are requested.
-There are also some well know problems regarding the precision of some events \cite{Weaver2008, Weaver2013a, Das2019SoK:Security, McGuire2009, Ramos2019AnCounters, Silva-De-Souza2020Containergy-aWorkloads}.
-Events that should be exact and deterministic (such as the number of executed instructions) show run-to-run variations and over-count on x86\_64 machines, even when running in strictly controlled environments.
-Our proposal is not based on performance counters and, therefore, is not vulnerable to those drawbacks.
+Models based on events present some drawbacks, they are highly dependent on the operating system and its architecture, making them problematic to port for other platforms.
+There are also limitations regarding the number of simultaneous events that can coexist without adding a non-negligible overhead.
+Additionally, there are cases where events need multiplexing, for example, when using more hardware events than %please confirm that the intended meaning has been retained.
+the CPU can provide.
+There are also some well know problems regarding the precision of some events, as shown in many studies~\cite{Weaver2008CanTrusted, Weaver2013Non-determinismImplementations, Das2019SoK:Security, McGuire2009AnalysisKernel, Ramos2019AnCounters, Silva-de-Souza2020ContainergyAWorkloads}.
+Some events that should be exact and deterministic (such as the number of executed instructions) show run-to-run variations and over-count on various architectures, even when running in strictly controlled environments.
+Because of that, our proposed model is not dependent on events and, therefore, not vulnerable to those drawbacks.
+% We confirm that the meaning is the same
 
-Instruction-level energy model was also proposed in \cite{Characterization2013EnergyProcessor}, obtained from an energy per instruction (EPI) characterization made on Xeon Phi.
+An instruction-level energy model was also proposed in \cite{Shao2013EnergyProcessor} by Yakun et.
+al.
+Where they proposed an energy per instruction (EPI) characterization made on Xeon Phi.
 Their model is expressed as:
@@
 	E(f) = \frac{(p_1 - p_0)(c_1 - c_0)/f}{N}, 
+	\label{eq:rw_instruction_level}
 \end{equation}
-where $N$ is the total number of dynamic instructions, $p_0$ is the initial idle power, $p_1$ is the average dynamic power and ($c_1$-$c_0$) refers to the cumulative number of cycles the micro-benchmark performs.
-This model is suitable for estimating the energy after the application finishes executing, when it is possible to count the number of total cycles.
-However, it is challenging to use for optimization or forecasting since it does not have an application model to predict the number of cycles.
-Our model integrates the behavior of the application taking into account the execution time.
+where $N$ is the total number of dynamic instructions, $p_0$ is the initial idle power, $p_1$ is the average dynamic power, and ($c_1$ $-$ $c_0$) refers to the cumulative number of cycles the micro-benchmark performs.
+This model is suitable for estimating the energy after the application finishes executing when it is possible to count the total cycles.
+However, it is challenging to use for optimization or forecasting since it does not have an application model to predict the cycles.
+Our model integrates the behavior of the application, taking into account the execution time.
 
-In \cite{Lewis2008Run-timeSystems} Lewis et al. described the overall system energy consumption using the following equation:
+Lewis et al.~\cite{Lewis2008Run-timeSystems} described the overall system energy consumption using the following equation:
 \begin{equation}
 	E = A_0(E_{proc} + E_{mem}) + A_1E_{em} + A_2E_{board} + A_3E_{hdd},
+	\label{eq:lewvis}
 \end{equation}
 where, $A_0$, $A_1$, $A_2$, and $A_3$ are unknown constants that are calculated via linear regression analysis and those remain constant for a specific server architecture.
-This model, as the previous one, relies on knowledge of energy spent on each component, being  a suitable option for estimation after the application has already ran, but not for optimization of the run itself.
+This model, as the previous one, relies on knowledge of energy spent on each component, being  a suitable option for estimation after the application has already run, but not for optimization of the run itself, which is the aim of our model.
 
 In another energy consumption model  based on system utilization, Mills et al.
-modeled the energy consumed by a compute node with CPU (single) executing at speed $\sigma$ as \cite{Mills2014EnergySystems},
+\cite{Mills2014EnergySystems} modeled the energy consumed by a compute node with CPU (single) executing at speed $\sigma$ as Equation (\ref{eq:mills})%\cref{eq:mills},
 \begin{equation}
 	E(\sigma,[t_1,t_2]) = \int_{t_1}^{t_2} \sigma^3 + \rho \sigma_max^3 dt,
+	\label{eq:mills}
 \end{equation}
+where $\rho$ stands for the overhead power consumed regardless of the processor speed, $t_1$ and $t_2$ are the application's initial and final execution times.
+The overhead includes power consumption by all other system components, such as memory, network, and more.
+For this reason, although the authors mentioned the energy consumption of a socket, their power model is generalized to the entire server.
+This model lacks a closed-form, i.e., it depends on the definition of $\alpha(t)$ to be complete.
+Our model has a closed-form which facilitates analyses.
 
-where $\rho$ stands for the overhead power  consumed regardless the processor speed, $t_1$ and $t_2$ are the initial and final execution time of the application.
-The overhead includes power consumption by all other system components such as memory, network, etc.
-For this reason, although the authors mentioned the energy consumption of a socket, their power model is generalized to the entire server.
+\subsection{The contribution of this work}
 
-Our work proposes a full-system energy model based on the CPU frequency and the number of cores.
-The aim of the model is to understand and optimize the energy behavior of parallel applications in HPC systems according to application parameters such as the degree of parallelism and CPU parameters related to dynamic and static power.
-The proposed model differs from existing ones for including the frequency and number of cores in the same equation for estimating the energy for a specific application in a given configuration.
-This model can serve as a base for DVFS and DPM optimization problems that include frequency and number of active cores.
+Although much work has been done on DVFS, the focus is still on the consumer electronics and laptop markets. 
+For HPC, the notion of energy perception is relatively new~\cite{Feng2003MakingSupercomputing}. 
+Moreover, the operational characteristics of non-HPC and HPC systems are significantly different. 
+First, the workload on non-HPC systems is very interactive with the end-user, but the workload on the HPC platform is not.
+Second, activities conducted on a non-HPC platform tend to share more machine resources.
+In contrast, in HPC, each job often runs with dedicated resources. 
+Third, an HPC system is usually much larger than a non-HPC system, making it more challenging to gather information, organize, and execute global decisions.
+Therefore, it is worthwhile to investigate whether a DVFS scheduling algorithm, which works well for conventional computing, remains effective for HPC.
+
+Our paper proposes a full-system energy model based on the CPU frequency and the number of cores.
+The model aims to understand and optimize the energy behavior of parallel applications in HPC systems according to application parameters, such as the degree of parallelism and CPU parameters related to dynamic and static power.
+The proposed model differs from existing ones, including the frequency and number of cores in the same equation for estimating the energy for a specific application in a given configuration.
+This model can serve as a base for considering DVFS and DPM optimization problems, including frequency and active cores.
 It can also be used to analyze the contribution of each parameter (ex: level of parallelism) to energy consumption.
-The number of cores is an essential factor in HPC since applications are designed to run on multiple cores.
+Furthermore, the number of cores is essential in HPC since applications are designed to run on multiple cores.
 
 The proposed energy model is the product of an application-agnostic power model and an architecture-specific application performance model.
-The power model is based on the CMOS logic gates power draw as a function of the frequency~\cite{Sarwar1997, Butzen2007} augmented to include the number of cores.
-The performance model is based on Amdahl's law \cite{Amdahl1967ValidityCapabilities, Eyerman2010ModelingDesign, Shi2015ReevaluatingLaw}, which can be used to estimate runtime in multi-core systems.
-This model has been extended to include execution frequency and input size, characterizing the application on the target architecture.
+The power model is based on the CMOS logic gates power draw as a function of the frequency~\cite{Sarwar1997CmosCalculation, Butzen2007LeakageGates} augmented to include the number of cores.
+The performance model is based on Amdahl's law \cite{Amdahl1967ValidityCapabilities, Eyerman2010ModelingDesign, Gustafson1988ReevaluatingLaw}, which can be used to estimate runtime in multi-core systems.
+In addition, this model has been extended to include execution frequency and input size, characterizing the application on the target architecture.
 
-\subsection{The contribution of this work}
-Although a lot of work has been done in DVFS, the focus is still mainly on consumer electronics and laptop markets.
-For HPC, the notion of energy perception~\cite{Beckman2005MakingSupercomputing} is relatively new.
-Moreover, the operational characteristics of non-HPC and HPC systems are noticeably different.
-First, the workload on non-HPC systems is very interactive with the end-user, but the workload on the HPC platforms is not.
-Second, activities conducted on a non-HPC platform tend to share more resources of the machine.
-On the other hand, often in HPC, each job runs with dedicated resources.
-Third, an HPC system is normally much larger than non-HPC systems, which makes collecting information, organizing decisions, and executing global decisions more challenging.
-Therefore, it is worthwhile to investigate whether a DVFS scheduling algorithm, which works well for conventional computing, remains effective for HPC.
+\cref{tab:related_work} summarizes the models comparing the system dependencies and the controllable variables.
+\begin{table}[H]
+	\footnotesize
+	\caption{Related work summary.\label{tab1}}
+	\setlength{\tabcolsep}{2.3mm}\begin{tabular}{cccc}
+		\toprule
+		%\begin{adjustwidth}{-\extralength}{0 cm}
 
-This work proposes a full-system energy model based on the CPU frequency and the number of cores.
-The goal of the model is to understand and optimize the energy behavior of parallel applications in HPC systems according to the application parameters such as the degree of parallelism and CPU parameters that relates to dynamic and static power.
-The proposed model differs from existing ones for including the frequency and number of cores in the same equation for estimating the energy for a specific application in a given configuration.
-This model can serve as a base for DVFS and DPM optimization problems that include both frequency and number of active cores.
-It can also be used to analyze the contribution of each parameter (ex: level of parallelism) to the energy consumption.
-The number of cores an important factor in HPC since applications are designed to run on multiple cores.
+		\textbf{Model}  & \textbf{System Dependency}       & \textbf{Variable}                     & \textbf{Controllable Variables}         \\ \midrule
+		Merkel et al. \cite{Merkel2006BalancingSystems} & performance counters    & number of activities & - \\
+		Roy et al. \cite{Roy2013AnAlgorithms}&performance counters &io operations, total time    & - \\
+		Yakun et al. \cite{Shao2013EnergyProcessor}&number of instructions &frequency&frequency \\
+		Lewis et al. \cite{Lewis2008Run-timeSystems}&energy of subcomponents&energy of subcomponents & - \\
+		Mills et al. \cite{Mills2014EnergySystems}&power of subcomponents&total time, frequency & frequency \\ 
+		Our model & - & frequency, cores, input size & frequency, cores, input size\\ \bottomrule
+	\end{tabular}
+	%\caption{Related work summary}
+	\label{tab:related_work}
+	%\bottomrule
+\end{table}
 
-The proposed energy model is the product of an application-agnostic power model and an architecture-specific application performance model.
-The power model is based on the CMOS logic gate power draw in the function of the frequency~\cite{Sarwar1997, Butzen2007} augmented to include the number of cores.
-The performance model is based on Amdahl's law \cite{Amdahl1967ValidityCapabilities, Eyerman2010ModelingDesign, Shi2015ReevaluatingLaw}, which can be used to estimate execution time in multi-core systems.
-We have extended  this model to include the frequency and input size, characterizing the application on the target architecture.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\subsection{Goals of the research}
+%The main contributions of the proposed model are:
 %
-%The interest in developing this project is to make HPC increasingly viable, in addition to reducing maintenance costs and reducing the impact on the environment by saving energy.
-%
-%Apart from this, our research also has international collaborations with other high-performance computing centers such as the Federal University of Rio Grande do Norte, where the project was started, and others which I had collaborated with projects related to energy saving in the University of Bristol, and possible future collaborations with University of Lausanne on Switzerland.
-%
-%The main contribution of this thesis will be to produce better energy saving techniques for HPC based on dynamic models of applications’ performance, power/energy consumption analysis and extended DVFS techniques.
-As will be detailed in the next section, several objectives have been defined to address the multiple challenges of our proposal: consideration of the processor workload, instruction set and the input size of applications, more precise energy models based on additional execution parameters, alternative optimization strategies and automatization of the execution process.
-%
-%First experiments developed showed that this methodology has a potential and can be improved to develop software as an end product that can be used by servers and clouds to save energy.
-This search can also be easily expanded to encompass mobile devices such as smartphones or microprocessed devices like raspberry-pi and beaglebone.
-Contributions in the areas of power consumption modeling and algorithms for frequency and voltage control off the processor are the product of this research.
-%
-%\subsection{Work plan}
-%
-%Important points that will be addressed in this research:
-%\begin{enumerate}
-%	\item \textbf{Identify and analyzing important performance features for energy consumption}.
-Analyzing the application execution with multiple inputs on different configurations of active cores and frequencies, performance parameters and its relations can be chosen and associated to characterize the application.
-Since there is a large number of performance parameters, a study to select the principal components is necessary to reduce its complexity.
-%	\item \textbf{Splitting the application into optimizable execution steps}.
-The goal is to split the application execution into several stages and minimize the energy on each stage, for that, the performance model is going to be completely remodeled.
-Creating a complete profile of the application, including performance parameters, frequency, and a number of active cores.
-With this, we can Cluster the applications identifying common frequency scaling stages to optimize.
-So when the application is running we can collect these performance features and find the closest cluster that will already have the optimized configuration.
-%	\item \textbf{Enhanced power models for HPC architectures: The power model is going to include additional parameters, such as the percentage of CPU utilization}.
-Further analysis of additional parameters, such as the influence of the instructions to be executed on the power consumed.
-Developing models that can reduce the overhead of training fitting and use fewer samples.
-%	\item \textbf{Enhanced DVFS models for HPC architectures: Including new architectures with distributed memory and heterogeneous systems}.
-Also studied the viability of using mathematical and artificial intelligence models, their advantages and their trade-offs to produce the final model.
-With these tools develops we can extend the current DVFS algorithms and produce a Smart DVFS module for HPC.
-%\end{enumerate}
-%
-%The new workflow that we want to implement can be illustrated on the figures \ref{fig:new_model} and \ref{fig:heterogeneous_systems}.
-%
-%\begin{figure}[h]
-%	\centering
-%	\includegraphics[width=\columnwidth]{intro/figures/new_model.png}
-%	\caption{New model}
-%	\label{fig:new_model}
-%\end{figure}
-%
-%\begin{figure}[h]
-%	\centering
-%	\includegraphics[width=\columnwidth]{intro/figures/heterogeneous_systems.png}
-%	\caption{Heterogeneous systems}
-%	\label{fig:heterogeneous_systems}
-%\end{figure}
+%\begin{itemize}
+%	\item Simple model: faster to fit and compute, good for DVFS and DPM optimization.
+%	\item Parameters with logical meaning: helps to understand the contribution of each specific term.
+%	\item Analytical analysis: several analyses can be derived from the equation.
+%	\item Controllable variables: the equation is in the function of parameters that we can control directly.
+%\end{itemize}

### 2022-04-05 thesis/main.tex
add journals and restructured

@@
 final,
-faculty=fpms
+faculty=fpms,
 ]{umons-Thesis}
@@
 \pagestyle{umons}
+%\setcounter{secnumdepth}{4}
 
+\usepackage{breakcites}
 \usepackage[main=english]{babel}
@@
 \usepackage{siunitx}
+\usepackage{multirow}
+
 \sisetup{%
@@
 ]{hyperref}
-\usepackage{cleveref}
+\usepackage[capitalise]{cleveref}
 \usepackage{outlines}
 \usepackage{listings}
+
+\definecolor{light-gray}{gray}{1}
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{gray}{1}
+\definecolor{commentgreen}{RGB}{2,112,10}
+
+\colorlet{punct}{red!60!black}
+\definecolor{background}{HTML}{EEEEEE}
+\definecolor{delim}{RGB}{20,105,176}
+\colorlet{numb}{magenta!60!black}
+
+\lstdefinestyle{jsonstyle}{
+	commentstyle=\color{commentgreen},
+	backgroundcolor=\color{light-gray},
+	basicstyle=\ttfamily\small,
+	breakatwhitespace=false, 
+	breaklines=true, 
+	captionpos=b, 
+	keepspaces=true, 
+	numbers=left, 
+	numbersep=5pt, 
+	showspaces=false, 
+	showstringspaces=false,
+	showtabs=false, 
+	tabsize=1,
+	literate={\ \ }{{\ }}1
+}
+
+\lstdefinelanguage{json}{
+	basicstyle=\ttfamily\small\normalfont,
+	numbers=left,
+	numberstyle=\scriptsize,
+	stepnumber=1,
+	numbersep=4pt,
+	showstringspaces=false,
+	breaklines=true,
+	frame=lines,
+	tabsize=1,
+	literate={\ \ }{{\ }}1,
+	backgroundcolor=\color{background},
+	morecomment=[f][\color{red}][0]{*},
+	literate=
+	{:}{{{\color{punct}{:}}}}{1}
+	{,}{{{\color{punct}{,}}}}{1}
+	{\{}{{{\color{delim}{\{}}}}{1}
+	{\}}{{{\color{delim}{\}}}}}{1}
+	{[}{{{\color{delim}{[}}}}{1}
+	{]}{{{\color{delim}{]}}}}{1},
+}
+
+\lstdefinestyle{ccodestyle}{
+	language=C,
+	commentstyle=\color{commentgreen},
+	backgroundcolor=\color{light-gray},
+	basicstyle=\ttfamily\small,
+	breakatwhitespace=false, 
+	breaklines=true, 
+	captionpos=b, 
+	keepspaces=true, 
+	numbers=left, 
+	numbersep=5pt, 
+	showspaces=false, 
+	showstringspaces=false,
+	showtabs=false, 
+	tabsize=2,
+	xleftmargin=0.5cm,
+	emph={int,char,double,float,unsigned,return},
+	emphstyle={\color{blue}}
+}
+
+\lstdefinestyle{pythonStyle}{
+	%backgroundcolor=\color{backcolour},
+	backgroundcolor=\color{light-gray},
+	basicstyle=\ttfamily\small,
+	commentstyle=\color{codegreen}\itshape,
+	keywordstyle=\color{magenta}\bfseries,
+	%numberstyle=\tiny\color{codegray},
+	stringstyle=\color{codepurple},
+	belowcaptionskip=\medskipamount,
+	breakatwhitespace=false, 
+	breaklines=true, 
+	captionpos=t, 
+	keepspaces=true,
+	frame=top,frame=bottom,
+	framexbottommargin=5pt,
+	%framextopmargin=5pt,
+	xleftmargin=0.5cm,
+	numbers=left, 
+	numbersep=5pt, 
+	showspaces=false, 
+	showstringspaces=false,
+	showtabs=false, 
+	tabsize=2
+}
 
@@
 \newcommand*\rot{\rotatebox{90}}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+\def\BibTeX{{\rmfamily B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
 		T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
 \DeclareOldFontCommand{\rm}{\normalfont\rmfamily}{\mathrm}
+\DeclareOldFontCommand{\sf}{\normalfont\sffamily}{\mathsf}
+\DeclareOldFontCommand{\tt}{\normalfont\ttfamily}{\mathtt}
+\DeclareOldFontCommand{\bf}{\normalfont\bfseries}{\mathbf}
+\DeclareOldFontCommand{\it}{\normalfont\itshape}{\mathit}
+\DeclareOldFontCommand{\sl}{\normalfont\slshape}{\@nomath\sl}
+\DeclareOldFontCommand{\sc}{\normalfont\scshape}{\@nomath\sc}
 
+\crefname{lstlisting}{listing}{listings}
 %%        %%
@@
 	\chapter{Introduction}
-
+	This chapter will...
 	\include{intro/intro}
 
-	\chapter{Proposed Models}
-	In this Section, the proposed power, performance and energy models are derived and validated
+	\chapter{Proposed Models} \label{chapter:models}
+	In this chapter, the proposed power, performance and energy models are derived and validated.
+	Energy consumption is crucial in high-performance computing (HPC), especially to enable the next exascale generation.
+Hence, modern systems implement various hardware and software features for power management.
+Nonetheless, due to numerous different implementations, we can always push the limits of software to achieve the most efficient use of our hardware.
+To be
+	energy efficient, the software relies on dynamic voltage and frequency scaling (DVFS), as well as dynamic power management (DPM).
+Yet, none have privileged information on the hardware architecture and application behavior, which may lead to energy-inefficient software operation.
+	This study  proposes analytical modeling for architecture and application behavior that can be used to estimate energy-optimal software configurations and provide knowledgeable hints to improve DVFS and DPM techniques for single-node HPC applications.
+	Additionally, model parameters, such as the level of parallelism and dynamic power, provide insights into how the modeled application consumes energy, which can be helpful for energy-efficient software development and operation.
+	This novel analytical model takes the number of active cores, the operating frequencies, and the input size as inputs to provide energy consumption estimation.
+	We present the modeling of 13 parallel applications employed to determine energy-optimal configurations for several different input sizes.
+	The results show that up to 70\% of energy could be saved in the best scenario compared to the default Linux choice and 14\% on average.
+	We also compare the proposed model with standard machine-learning modeling concerning training overhead and accuracy.
+The results show that our approach generates about 10 times less energy overhead for the same level of accuracy.
 	\include{models/models}
 	\include{phases/phases}
-	\include{optmizers/optmizers}
+	%\include{optmizers/optmizers}
 
@@
 	This chapter introduces the framework developed, the fingerprint tool and pascal analyzer.
+
+	\include{pascalanalyzer/pascalanalyzer}
 	\include{fingerprint/fingerprint}
-	\include{pascalanalyzer/pascalanalyzer}
 
-	\chapter{Results (model validation)}
+	\chapter{Experiments and Results}
 	This chapter will describe all the experiments that validate the models constraints.
@@
 	\chapter{Conclusions}
+	This chapter will...
+	\include{conclusion}
 
-	\bibliographystyle{plain}
-	\bibliography{references}
+	%\bibliographystyle{plain}
+	%\bibliography{references.bib}
+	\include{references.tex}
 

### 2022-04-05 thesis/optmizers/optmizers.tex
add journals and restructured

--- 2021-03-23 thesis/ch7/ch7.tex
@@
 
-The complexity of this problem also varies, depending on the choice of power function it can result in linear programming \cite{Kim2015RacingHeuristics}, quadratic programming \cite{Horyath2008Multi-modeClusters}, until NP-hard problems \cite{Fu2018RaceMinimization}.
-
-\subsection{DVFS optimization} \label{sec:dvfs_optmin}
-The effectiveness of the proposed approach for optimization was evaluated with a simple algorithm that finds the optimal frequency and number of active cores from the equation proposed.
-The results were then compared to the Linux default choices for power management.
-
-With the equation developed \ref{eq:en_final}, it is possible to calculate energy consumption estimations for every possible configuration since there is a finite interval of possible values for the frequency and cores.
-Then, the configuration that minimizes energy consumption for a given input can be selected.
-It is also possible to apply constraints on the execution time, frequency, and the number of active cores.
-
-Current HPC managers leave to the user the choice of how many cores to use. Because of that three situations were analyzed with relation to the number of cores:
-
-\begin{enumerate}
-	\item Worst choice: number of cores that maximize the total energy consumed
-	\item Random choice: energy consumed for a random choice of number of cores
-	\item Best choice: number of cores that minimize the total energy consumed (oracle)
-\end{enumerate}
-
-The default option for the Linux governor is the Ondemand, and by default, it does not have a DPM control for the number of active cores.
-Since Ondemand only does DVFS, to allow the comparison, each application was executed with all number of cores available in the system 1 to 32.
-
-Figures \ref{fig:energy_worst_case}, \ref{fig:energy_mean_case} and \ref{fig:energy_best_case}, shows the energy savings with respect to Ondemand in the following way $\frac{Ondemand-Model_{min}}{Ondemand}$ with all three cases described above.
-The list below shows the savings and losses for each case:
-
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width=\columnwidth,height=8.5cm]{models/figures/dvfs_cmp_max.png}
-	\caption{Energy saving comparison with the proposed model and for the worst case}
-	\label{fig:energy_worst_case}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\columnwidth]{models/figures/dvfs_cmp_mean.png}
-	\caption{Energy saving comparison with the proposed model and for the random case}
-	\label{fig:energy_mean_case}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\columnwidth]{models/figures/dvfs_cmp_32.png}
-	\caption{Energy saving comparison with the proposed model and for the best case}
-	\label{fig:energy_best_case}
-\end{figure}
-
-The list below shows the savings and losses for each case:
-
-\begin{enumerate}
-	\item Worst choice: save 69.88\% on average
-	\item Random choice: save 12.04\% on average
-	\item Best choice: lost 14.06\% on average
-\end{enumerate}
-
-%This can be explained by the linear speedup characteristic from this application. Thus, it is to be expected that on constant efficiency, the better solution will be with more cores.
-
-%The other results on the other applications show that the proposed approach, even in cases on the bests configurations for DVFS, demonstrate that can obtain energy saving of up 18\%, comparing to the Ondemand governor.
-
-%\begin{figure}[H]
-%    \centering
-%    \includegraphics[width=\columnwidth]{dvfs/completo_black_ondemand_nolimit_2_10.png}
-%    \caption{Caption}
-%    \label{fig:my_label}
-%\end{figure}
-%
-%In general, for the case-study applications and case-study architecture, the optimal-energy configurations tend to be the ones using the highest frequency, which characterizes a race-to-idle rather than a pace-to-idle optimal behavior~\cite{Kim2015RacingHeuristics}.
-This can be explained by the large static power observed in the considered architecture, evidenced by the large $c_3$ parameter in (\ref{eq_power_final}) that was fitted in (\ref{eq:fittedpower}).
-With a large static power, using a pace-to-idle strategy, i.e.
-the use of frequencies lower than the maximum, is expected to be effective only if the sum of the leakage and the dynamic power parcels is larger than the static power parcel.
-%Based on the fitted power model, this would never happen, i.e. the sum of leakage and dynamic power is always less than the static power,
-%\begin{equation*}
-%p(0.29f^3+0.97f)+9.18s < 198.59, 
-%\end{equation*}
-%even if we use the maximum number of cores, $p=32$ and $s=2$, and the maximum frequency, $f=2.2$.
-%Nevertheless, race-to-idle was not always the best strategy because energy scales with the execution time, which in turn scales inversely with the number of active cores and the operating frequency, and because power scales linearly with the number of cores, but exponentially with the frequency.
-%
-%The optimal number of active cores depends on the parallel scalability of the application.
-The more scalable the application, the more cores it requires to minimize energy.
-A scalable application can increasingly exchange the speedup of more cores with lower frequencies in order to spend less energy.
-This is because of the linear relationship between power and number of cores and the exponential relationship between power and frequency.
-%
-%In all cases, the method proposed here outperformed the worst case of the Ondemand governor.
-On average, the difference in energy consumption was about 790\%, being 1298\% the maximum difference and 59\% the minimum.
-In general, the energy consumption of the DFVS scheme was larger for smaller numbers of cores.
-Nonetheless, it was not always the case that the best number of cores for this scheme was the maximum, i.e.
-32 cores.
-Possibly, for architectures with larger number of cores, choosing the exact number the minimizes energy consumption would be less evident.
-
-Since the number of cores in HPC is requested by the user, the number of CPU requests during one year in UFRN's cluster was analyzed. The result is plotted in the figure \ref{fig:cpu_requests}.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\columnwidth]{models/figures/cpu_request.png}
-	\caption{Number of CPUs requested by the users during 2016-2017}
-	\label{fig:cpu_requests}
-\end{figure}
-
-The worst choice was always with a single core and is the most common choice of many regular users.
-The best choice was quite often 32 cores, which is the third most popular choice of the users but it is $72\times$ less than 1 core.
-This can point a direction of how much energy can be saved by using this technique for DPM or more advanced optimization algorithms that can be derived from our model.
-
-In practice, this approach can be brought into production by allowing the resource manager to perform these changes for the user using pre-scripts and post-scripts for job submissions with energy consumption requirements.
-This approach is simple to be implemented and can work in almost any Linux system.
-
-\section{Conclusion} \label{sec:conclusion}
-This work proposes an energy model based on the frequency and number of cores for a full shared memory system.
-This model can serve as a base for DVFS and DPM optimization problems that include both frequency and active cores, as well for analysis of the contribution of each parameter (ex: parallelism level) to the energy consumption.
-
-Results from 3 HPC benchmarks running in one cluster demonstrate the potential of the proposed novel model.
-While consuming less energy than traditional machine learning approaches, it can serve as bases from DVFS and DPM algorithm as shown in the \ref{sec:dvfs_optmin} in an average case saving about 12\% up to 69\%.
-The previous knowledge of the application's performance can expose sufficiently relevant information, such as parallel speedups, that is harder to guess in run-time techniques based on DVFS.
-
-A weakness of the proposed model is the need for information about the input size of the application that can be complex to derive.
-A possible solution would be to precisely define what is input size, given an definition in function of common variable for all applications like throughput for example.
-Future work will demonstrate all the possible analysis that is possible to archive with the equation as long as more advanced DVFS models that can be derived using the equation.
-For instance identification of different phases of the target program and thus, it will enable more fine-grained changes of the frequency and, perhaps, the number of active cores, to further improve the results presented here.
-
-Another important aspect that is typically not taken into account is the number of processing cores to be used by a parallel program.
-This choice is left to the user, which often is not trivial as shown in this paper.
-
-%%%% moved to the conclusions.
-% A weak of this approach is that the model of performance and power needed to be trained with previously known input sizes.
-This weakness can be treated with a future solution that can incorporate a prediction of input size during execution of the application.
+The complexity of this problem also varies, depending on the choice of power function it can result in linear programming \cite{Kim2015RacingHeuristics}, quadratic programming \cite{Horyath2008Multi-mode}, until NP-hard problems \cite{Fu2018RaceMinimization}.

### 2022-04-05 thesis/pascalanalyzer/pascalanalyzer.tex
first version
add journals and restructured

@@
-\section{Pascal Analyzer}
+\section{Pascal Analyzer} \label{sec:pascal_analyzer}
+High-performance computing systems have become increasingly dynamic, complex, and unpredictable.
+To help build software that uses full-system capabilities, performance measurement and analysis tools exploit extensive execution analysis focusing on single-run results.
+Despite being effective in identifying performance hotspots and bottlenecks, these tools are not sufficiently suitable to evaluate the overall scalability trends of parallel applications.
+Either they lack the support for combining data from multiple runs or collect excessive data, causing unnecessary overhead.
+In this work, we present a tool for automatically measuring and comparing several executions of a parallel application according to various scenarios characterized by the input arrangements, the number of threads, number of cores, and frequencies.
+%In this work, we present a tool for measuring and comparing several executions of a parallel application automatically with different input arrangements, number of threads, cores, and frequencies to understand its scalability by observing how it uses the available computational resources across configurations.
+Unlike other existing performance analysis tools, the proposed work covers some gaps in specialized features necessary to better understand computational resources scalability trends across configurations.
+%The proposed work covers some gaps and unspecialized features of existing performance analysis tools. 
+%The tool features automatic instrumentation and direct mapping of parallel regions, accuracy-preserving data reductions, and ease of use to compare multiple configurations for improving analysis and productivity.
+In order to improve scalability analysis and productivity over the vast spectrum of possible configurations, the proposed tool features automatic instrumentation, direct mapping of parallel regions, accuracy-preserving data reductions, and ease of use.
+%In addition, it aims at parallel scalability, instead focus on single-run performance, aggregating a minimal level of intrusion (<1\% on average), which is fundamental for accurately understanding the behavior and scalability trends of the parallel application.
+As it aims at accurately understanding scalability trends of parallel applications, detailed single-run performance analyses show minimal intrusion (less than 1\% overhead).
 
-Run application collecting information while executing a program with
-different configurations of cores, frequencies, inputs arguments.
+\subsection{Introduction} \label{sec:introduction_pascal}
 
-\subsection{Architecture}
-\begin{figure}
-	\includegraphics{pascalanalyzer/arch2}
+Developing parallel programs capable of exploiting the computational power of high-performance systems is as well-known as it is challenging~\cite{Huck2007, Islam2019, Weber2019}. 
+During the development process, the code optimization step is a fundamental part of the software construction strategy and is supported by performance measurement and analysis tools~\cite{Bergel2019, Weber2019, Huck2005, Geimer2010, Shende2006, Adhianto2010, Miller1995, Galobardes2015, Pillet2007, Islam2019}.
+The main objective of these tools is to help developers understand the execution characteristics, allowing the identification of bottlenecks and behavioral phenomena that compromise the program’s efficiency, thus guiding possible improvements~\cite{Brink2020, Huck2007}.
+
+Nowadays, due to the complexity of parallel systems, correctly identifying and locating performance and scalability bottlenecks depends on the developer's ability to compare several measurements in different execution configurations~\cite{Bergel2019, Silva2018}.
+This investigative process tends to be tedious and complex, requiring in-depth knowledge of the problem domain, parallel systems, and measurement and analysis tools.
+%Mastering the analysis tools becomes necessary mainly because they differ from each other in terms of measurement strategy, the collected metrics, and the many points and focus of observation. 
+Mastering the analysis tools is necessary because they differ from each other in terms of measurement strategy, metrics collected, and the many points and focus of observation.
+Even though some solutions propose organizing and combining metrics from different sources, these distinct characteristics can compel developers to exploit more than one tool depending on their analysis~interests.
+
+%Despite their differences, every existing tool has its own contributions in helping to understand the inner working of a program, and some can present a vast and varied set of metrics with a high degree of detail.
+Every existing tool has its own contributions in helping to understand the inner working of a program. Some can present a vast and varied set of metrics with a high degree of detail. 
+However, developers may have difficulty using these tools when their interest is in analyzing parallel scalability. 
+%The main challenge arises from the focus on investigating a single run, because scalability analysis requires contrasting runs of different configurations. 
+The main challenge arises from the emphasis on studying a single run because scalability analysis requires contrasting runs of different configurations.
+%In this sense, it tends to be more significant to measure and compare only a few data across executions of multiple configurations than to collect a vast amount of finer-grain data of single-configuration runs.
+In terms of productivity, measuring and comparing only a few data between runs of multiple configurations tends to be more significant to scalability analysis than collecting a vast amount of finer-grain data of single-configuration runs.
+Furthermore, due to the number of measurements, there is a natural overhead associated with the execution of the analysis tool itself.
+Some works indicate this overhead can reach 40\% of the program runtime, which influences and damages the observation of its efficiency variation~\cite{Eriksson2016}.
+As such, a measurement and analysis approach that uses only the data needed to understand the application's efficiency variation can be advantageous.
+First, because measuring and collecting only the necessary data limits the degree of intrusion and, second, because this strategy makes the analysis tool simpler and easier to use.
+%Finally, the tool's availability can also be a drawback as some apply to specific compilers and execution frameworks.
+Indeed, the framework version is a limitation for some traditional tools like HPCToolkit~\cite{Adhianto2010} and TAU (Tuning and Analysis Utilities)
+%MDPI: Is this an abbreviation? If affirmative, please add the explanation. Please carefully check and confirm -- INCLUDED ON THE NEXT PARAGRAPH
+~\cite{Shende2006}.
+
+From this context, we present a measurement tool that focuses on analyzing the parallel scalability of programs.
+It primarily provides runtime and power consumption measures in an approach that favors inspection of the overall behavior of the program before starting a more in-depth investigation of points that naturally require more detail, time and effort.
+The work explores and combines strategies already used in other solutions, such as tracing, post-mortem analysis, and automatic instrumentation based on shared libraries.
+Additionally, it includes an instrumentation model that automatically relates parallel code regions to analysis regions, using a measurement strategy based on the intrinsic and primary behavior of the parallel frameworks.
+Due to this approach, the tool allows analyzing programs developed for shared-memory environments regardless of the compiler or framework versions.
+The framework version, which in some cases also requires specific compiler versions, is a limitation for some traditional tools like HPCToolkit~\cite{Adhianto2010} and TAU (Tuning and Analysis Utilities)~\cite{Shende2006}.
+
+The proposed tool uses a tracing-based measurement technique to identify and measure the boundaries of the analysis regions.
+The tracing approach supports a hierarchical region analysis model allowing users to see how the efficiency of inner parts can impact the program's overall scalability.
+In addition, the tool uses an aggregation feature to reduce the volume of data produced while preserving accurate analysis capability with minimal overhead.
+The tool also includes usability features that reduce the developer's efforts during the measurement and analysis process.
+These features allow one to automate the process of running and merging data collected from multiple configurations, bypassing manual efforts and avoiding the drawbacks of non-computer-oriented analysis.
+%The tool does not display visuals elements natively.
+Still, its interface allows the collected data to be exported and viewed from the control terminal in tabular form.
+Data exported via the interface are organized into data frames that can be interpreted by other tools or even rendered directly from graphic libraries.
+The collected data, available from the terminal in tabular form, can be exported as data blocks that can be interpreted by other tools or even rendered by graphics libraries to facilitate the visualization and data analysis.
+
+%The proposed tool is an alternative for realizing parallel scalability analysis more efficiently than single-run-centric performance measurement and analysis tools. 
+We offer an alternative tool for realizing parallel scalability analysis more efficiently than single-run-centric performance measurement and analysis tools.
+In addition to that, our framework can also be used to promote energy-aware software; methods that rely on accurate and efficient energy collection could benefit from energy consumption data for the entire program and specific regions of the application \cite{10.1007/978-3-319-58667-0_22,10.1007/978-3-319-58667-0_21,7016382,10.1145/3321551}.
+To summarize, the main contributions of this work are:
+
+\begin{itemize}
+	\item A lightweight automated process of measuring and comparing runs with different configurations.
+	\item Transparent linking of parallel code sections to analysis regions.
+	\item Hierarchical views and degrees of efficiency for the inner parts of a program.
+	\item The imposed overhead is low and minimally intrusive, mainly because of the measurement approach and the number of metrics collected.
+	\item It provides energy consumption measures for the different configuration parameters. 
+\end{itemize}
+
+The following sections describe how the tool is built and used for performance analysis.
+Section \ref{sec:related_work_pascal} presents the related works.
+Section \ref{sec:design_and_features_pascal} describes the tool architecture and goals; it explains the main features, usage, and collected data.
+The experimental results are presented in Section \ref{sec:validation_the_framework}.
+Finally, the contributions are summarized in Section \ref{sec:conclusions_pascal}, with an outlook of future works.
+
+\subsection{Related Works} \label{sec:related_work_pascal}
+
+Performance analysis aimed at code optimization is an essential activity for developing parallel applications with scalable performance.
+For this purpose, performance analysis tools, like Caliper~\cite{Boehme2016}, HPCToolkit~\cite{Adhianto2010}, Scalasca~\cite{Geimer2010}, TAU~\cite{Shende2006}, Vampir~\cite{Weber2019}, and VTune~\cite{VTune}, are useful because they help developers identify where the application uses the available computing resources inefficiently through collecting detailed data of its execution.
+Such tools differ from each other mainly in terms of data measurement and analysis strategy: tracing or profiling, post-mortem or real-time; individual or comprehensive observation focus; and for providing or not providing visual elements that facilitate the judgment of collected data.
+Other solutions, like the DASHING framework~\cite{Islam2019} and the HATCHET library~\cite{Brink2020}, allow inspection of performance data from multiple sources.
+In this case, the goal of DASHING and HATCHET is to provide a more robust data set to support the analysis process.
+
+In the performance analysis process, observing the parallel program's execution as a sequence of events representing significant activities is essential to understand its behavior~\cite{Pantazopoulos1997}.
+Events are basic units of the analysis process, and the way they are observed influences strategies for collecting performance data.
+Profiling-based analysis tools collect information from events that occurs in the program execution and commonly operate by statistical sampling using interruptions.
+These interruptions can be caused, for example, by periodic breaks or hardware events.
+Interruptions allow checking the system state and the information collected depends on the focus of observation.
+Profiling-based tools use statistical techniques to describe the program behavior in terms of aggregate performance metrics.
+They usually ignore the chronology of events, but are known to be useful for identifying, for example, load imbalance, high communication time, or excessive routine calls.
+Paradyn~\cite{Miller1995}, and Periscope~\cite{Gerndt2010} are tools that adopt this strategy.
+
+On the other hand, tracing-based analysis tools collect performance data from events that occur when the program takes over a particular state.
+Tracing can provide valuable information about the time dimension of a program, allowing users to check when and where transitions of routines, communications, and specific events occur.
+This measurement strategy tends to be more invasive and intrusive, generating a larger dataset than the profiling-based approach.
+Vampir~\cite{Weber2019} and Paraver~\cite{Labarta2005} are examples of this group.
+Some tools like HPCToolkit~\cite{Adhianto2010}, Scalasca~\cite{Geimer2010}, TAU~\cite{Shende2006} and VTune~\cite{VTune} support both profiling and~tracing.
+
+%In this work we propose tracing to measure the program and identify when parts of its code were executed. 
+In this work we propose tracing to identify when parts of a code were executed. 
+%To mitigate some of the disadvantages usually related to tracing, the tool implements features such as automatic instrumentation mode and aggregation resource, in addition to adopting an analysis based only on time.
+To mitigate some of the disadvantages related to tracing, in addition to adopting a time-based analysis, the tool implements features such as automatic instrumentation mode and resources aggregation.
+%The automatic instrumentation mode makes it possible for users to analyze the program's execution without modifying the compiled code. 
+Our automatic instrumentation mode allows users to analyze program execution without modifying the source or the executable.
+Although many other analysis tools use the automatic instrumentation mode, this work presents a different way of mapping and measuring the code parts. 
+%In our proposal, parallel regions are the main focus of analysis, and there is no limitation on the version of OpenMP used in the development, as in tools such as HPCToolkit~\cite{Adhianto2010} and TAU~\cite{Shende2006}.
+In our proposal, parallel regions are the main focus of analysis. 
+%The strategy is efficient in identifying the scalability trend of parts or the whole parallel program in execution.
+The strategy is effective enough to identify the scaling trends of parts or of the entire running parallel program.
+%Furthermore, collecting only data related to runtime reduces the associated overhead with the measurement process. 
+Furthermore, collecting only specific runtime data reduces the overhead associated with the measurement process.
+As with other similar tools, such as HPCToolkit~\cite{Adhianto2010} and TAU~\cite{Shende2006}, there is no limitation on the version of OpenMP~used.
+
+Runtime overhead is an attribute associated with the set of additional instructions that are executed to collect the program measurements.
+The time to perform this ``extra code'' varies in line with the tool and is directly related, in quantity and degree of detail, to the aspects it measures.
+Limiting this overhead is crucial because it can compromise the understanding of program behavior, and divert optimization efforts to less effective points. 
+According to some comparative and practical studies in this regard, tests on traditional tools have shown a runtime overhead ranging from 2\% to 40\%~\cite{Eriksson2016}.
+%Fundamental to the proposal of this work, the overhead presented by the proposed tool differs according to the program under analysis but can be considered negligible (<1\%) for analyzing scalability trends of parallel applications.
+Although different depending on the program analyzed, the overhead resulting from our instrumentation strategy can be considered negligible (less than 1\%) and optimal for analyzing trends in parallel applications.
+
+Several performance analysis tools employ a post-mortem approach. 
+%In this kind of approach, the tools collect performance metrics during the program execution, but the data interpretation is carried out only after its execution. 
+In this case, the tool performs performance measurements and data collection while the program is running, and then the collected data will be interpreted.
+HPCToolkit~\cite{Adhianto2010}, Scalasca~\cite{Geimer2010}, TAU~\cite{Shende2006} and VTune~\cite{VTune} are examples of post-mortem performance analysis tools. 
+%Post-mortem performance analysis tools may require the storage of large volumes of performance data, but they are more suitable to provide an overall view of the execution. 
+Depending on the type of analysis, these tools may require storing large amounts of data, but they are best suited to provide an overall view of the execution.
+%On the other hand, real-time analysis tools enable performance analysis during the execution of the program. 
+In contrast, run-time analysis tools perform both operations at run time.
+This approach allows detecting waiting states and communication inefficiencies accurately. 
+%However, an online analysis environment usually requires the coordinated usage of extra tools, which increases the analysis infrastructure complexity. 
+Periscope~\cite{Gerndt2010} and Paradyn~\cite{Miller1995} are examples of tools in this group.
+However, run-time analysis generally requires the coordinated action of tools, which increases the structural complexity.
+%The need for synchronous initialization, communication between all resources used in the analysis, and the additional load on the network are disadvantages of the real-time approach. 
+The need for synchronous initialization and communication between analysis resources as well as their impact on the runtime environment are the drawbacks of this approach.
+%In addition, the analysis that depends on data collected at the end of the program's execution, as in the case of scalability analysis, are potentially damaged by the infrastructure overhead and do not, in practice, benefit from characteristics of real-time tools.
+In addition, the analysis, particularly that of scalability, which relies on the data collected at the end of the execution, is potentially degraded by the infrastructure overhead and therefore does not benefit from the characteristics of real-time tools.
+
+\textls[-15]{Energy management solutions can also provide similar features for example GEOPM~\cite{10.1007/978-3-319-58667-0_21},} allows the measurement of time and energy of specific regions.
+The main advantage that we could not find in any other framework is that our tool can provide specific data of energy consumption of parallel regions in a completely automated way.
+For example, in GEOPM, it is necessary to instrument the source code to obtain the same result.
+
+%To investigate the scalability trend, it is more important to identify the program's general behavior than its state at a given moment of execution. 
+%Although the current performance analysis tools are helpful and allow to identify numerous aspects regarding the program's behavior, none of them presents characteristics directly focused on observing the scalability trend of parallel code.
+An indispensable feature for this purpose includes automating the program's measurement process in different configurations because a more accurate understanding of scalability requires measuring and comparing multiple code runs.
+%Focused on more practical and objective analysis, we propose an alternative tool to observing a parallel program's scalability trend.
+It collects only information that allows deriving the overall program runtime, or parts of it, according to the user's interest, and includes relevant functionalities to an efficient measurement and analysis process.
+Among the supported features it provides: the automation of program runs based on different parameters; the generation of data sets that contain metrics with different execution parameters; and the reduction of the generated data volume on disk.
+
+%Although the current performance analysis tools are helpful and allow to identify numerous aspects regarding the program status, to investigate scalability trends, it is more important to identify the behavior of parallel code rather than its state at a given moment of the execution, therefore the post-mortem approach is more suitable.
+By focusing on objective analysis, we offer an alternative tool to observe the scalability trend of a parallel program.
+Thus, we collect just the information allowing us to infer the overall behavior of the program, or, according to the interest of the user, specifically chosen parts of it, including features relevant to an effective measurement and analysis process.
+To this end, it is essential to also automate the processes to obtain comparative measurements of several runs. 
+Supported features are the automation of program executions according to various configuration parameters, including a user-specified granularity, the generation of datasets containing comparative measurements of runtime parameters, and efficient optimization of the amount of produced data.
+%In this work, the tool mitigates the disadvantage related to the volume of data set by aggregating the collected data.
+Regarding the amount of data collected, our approach uses data aggregation, which facilitates storage and further analysis.
+
+\subsection{Design and Features} \label{sec:design_and_features_pascal}
+
+This section presents the architecture, components, and interconnections used in the design of the proposed tool.
+It also includes a discussion about main features, instrumentation modes, and output file structure that contains the data measured and collected by the proposed tool.
+
+\subsubsection{Architecture} \label{subsec:architecture_pascal}
+The tool we propose to measure, combine and compare multiple runs of a parallel application implements two main concepts to achieve this goal: actuators and sensors.
+
+Actuators characterize the parameters we intend to observe.
+They are variables representing elements external to the program and which, when changed, can influence aspects such as the performance and efficiency of the running application.~Therefore, analyzing the result of actuators' variation is essential to understand how these elements impact program behavior, especially scalability and power consumption.
+By default, the tool implements actuators controlling the number of active cores and threads, the program input, and the CPU operating frequency.
+
+The sensor is the concept we use to represent the elements addressed to measure and monitor the variation of actuators.
+Both sensors and actuators are plugged into the tool core as modules.
+Currently, the tool implements three types of sensors:
+
+\begin{itemize}
+	\item Begin/end: collects data at launch and end of each program run.
+	\item Time sample: periodically collects data.
+	\item Events-based: collects data when a specific event happens.
+\end{itemize}
+
+Currently, there are sensors to measure time, energy, and performance counters.
+The default sensor is a \textit{begin/end} type used to collect the execution time of the whole application.
+To measure energy consumption, we developed sampling sensors capable of retrieving data from RAPL (Running Average Power Limit) and IPMI (Intelligent Platform Management Interface), which are interfaces that provide power information from the CPU
+%MDPI: Please add the explabnation % added in abbreviations
+and the entire system. Apart from that, there are \textit{time sample} and \textit{begin/end} sensors to gather performance counters data.
+
+Measurements that support scalability analysis are taken from event-based sensors.
+For this, the tool includes markers that trigger events to automatically identify the boundaries of parallel regions defined by developers in codes that use POSIX Threads or OpenMP.
+%MDPI: Are these abbreviations? If affirmative, please add the explanations. Please carefully check and confirm % added in abbreviations
+This feature is essential for relating parallel code sections to analysis regions and measuring execution times directly from binary code.
+The triggers for this measurement mode are implemented in a wrapper library and therefore already available in the system.
+The tool also provides manual marks that can be inserted into the source code to monitor specific parts of the program.
+Furthermore, any system event can be set as a trigger.
+\cref{fig:pascal_architecture} describes the integration of each part of the proposed software.
+The idea is that actuators and sensors are modular parts that can easily be added or removed.
+The tool's core is responsible for all operations, launching the application, and data gathering.
+
+\begin{figure}[H]	
+	%\includegraphics[width=15 cm]{Definitions/logo-mdpi}
+	\includegraphics[scale=0.7]{pascalanalyzer/figures/designfeatures/pascal_philosophy.pdf}
+	\caption{Architecture showing the interconnections of the central parts of the tool.
+Either binary (using the wrapper library) or an instrumented source code, the target application can be launched on the target platform by the tool core following the configuration parameters chosen by the user while deploying actuators/sensors.
+The resulting data collection is stored in Json file format for post analysis and visualization (GUI).
+\label{fig:pascal_architecture}}
 \end{figure}
-\subsection{Modules description}
-PascalRun: Provie a monitor class to collect information while executing  a program 
-with different configurations of cores, frequencies, inputs arguments.
-The information collected can be via sensor specified by the user,  instrumented regions (source code and openmp binaries) and the entire  program execution time.
-The data is stored and processed with PascalData  module that can resume all on a DataArray (xArray module) with times, speedups or efficiency.
 
-PascalData: Load saved data from pascalrun and can convert and process to varius formats.
+\subsubsection{Instrumentation and Intrusiveness} \label{subsec:intrusiveness_pascal}
 
-PascalModel: Create AI and Analytical models from pascal data
+The instrumentation module is one of the most crucial.
+In addition to automating the instrumentation, it determines the overload and level of intrusiveness of the instrumentation process.
+This module is designed to execute the fewest instructions possible in the most optimized manner.
+Currently, there is instrumentation support for C/C++ languages through shared libraries that work with either automatic or manual instrumentation.
+Manual instrumentation is preferred when it is necessary to examine the application program in specific code sections, regardless of the type of content.
 
-Instrumenter: Instrument source code creating marks at parallel regions
+The manual instrumentation mode is implemented in three routines: one for initialization, another to mark the beginning of the region of interest ({\tt analyzer\_start}), and finally, a routine to mark the end of that region ({\tt analyzer\_stop}).
+The initialization routine is called when loading the library to create the necessary data structures and set up the data exchange communication.
+The routines {\tt analyzer\_start} and {\tt analyzer\_stop} identify threads in a region and store timestamps.
+These routines are implemented in such a way that only one thread at a time writes in a designated position of a two-dimensional array, thus ensuring thread safety and eliminating the necessity of locks.
 
-libinst: Instrument binary code creating marks at parallel regions
+The automatic instrumentation includes a routine allowing us to intercept the creation of threads via the {\tt LD\_PRELOAD} environment variable.
+This routine overwrites parts of an existing native library.
+In this manner, the functions responsible for thread spawning, such as {\tt pthread\_create} in the pthread library, {\tt GOMP\_parallel} (GCC implementation), {\tt \_\_kmpc\_fork\_call} (Clang implementation) in the OpenMP framework, and similar functions for other compilers, are intercepted to identify the parallel regions automatically.
+This approach is less intrusive than other tools, such as using a debugger interface with breakpoints or performing binary code instrumentation.
 
-cpufreq: resposible of CPU control
+A key point of performance analysis tools, particularly important when analyzing real-time program execution, is that instrumentation must have a negligible impact on program behavior and execution time.
+Indeed, as mentioned in Section \ref{subsec:architecture_pascal}, we support three types of sensors, each of which has a different degree of intrusion.
+There is hardly any sensors intrusion at the start and then at the end of the program execution since it is simply the invocation of both data collection routines.
 
-IPMI, RAPL, RAPL\_sample, ProcList: Sensors
+With sampling sensors, the degree of intrusion depends on the sampling rate and the total execution time of the application under study.
+Therefore, the intrusion needs to be assessed on a case-by-case basis with this type of sensor.
 
-\subsection{Features}
+%It is fixed concerning the number of instructions and diverges with the number of threads.
+For example, suppose the user intends to analyze a parallelized region that is executed by four processing units.
+In that case, the execution time overhead is approximately equal to $4 \times T_{measurement}$, where $T_{measurement}$ corresponds to the time taken to perform a measurement, comprising the identification of events within the limits of the region involved in the measurement.
 
-\subsubsection{Options}
-\begin{outline}
-	\1 list of cores
-	\1 list of frequencies
-	\1 list of inputs
-	\1 list of sensors
-		\2 ipmi
-		\2 rapl (sysfs)
-		\2 rapl (perf)
-		\2 process cores monitor
-	\1 number of repetitions
-	\1 choice of CPU governor
-	\1 idle time between task
-	\1 enable/disable hyperthread
-	\1 enable/disable cores
-	\1 collect times from marks manually inserted by the user in the source code
-	\1 collect times from parallel regions using binary instrumentation (openmp)
-	\1 list of sensors
-	\1 instrument source code
-	\1 output file name
-	\1 verbosity level
-		\2 0 nothing
-		\2 1 only display run configuration
-		\2 2 display run configuration and sensors data
-		\2 3 display everything
-\end{outline}
+%The overhead for this example is the same when measuring two parallel regions with two cores and one parallel region with four cores.
+The \cref{lst:overheadsample1} and~\ref{lst:overheadsample2} present code snippets that correspond to the two scenarios mentioned and which, when analyzed using the tool, are therefore impacted by the same overhead.
 
-\subsubsection{Execution}
+The instrumentation overhead does not depend on the number of instructions or the runtime required to process the set of instructions to be analyzed.
+However, it varies according to the number of processing units used and measurements taken.
+To estimate the order of magnitude of the overhead ($T_{measurement}$), %MDPI: Is the italics necessary?
+Please carefully check and confirm.
+% yes
+we measured recurring calls to the proposed sampling functions delimiting the regions in a simple benchmark code.
+This experiment was carried out in the same target architecture described in the experimental results of Section \ref{sec:casestudyarchitecture}, and the code structure is presented on Listing~~\ref{lst:overheadcode}.
 
-Examples execution using command line
-\begin{lstlisting}
-pascalanalyzer ./myapp --bin --idle_time 5 --cores 1,4  
-		--frequencies 3000000,2800000 --verbosity 3
+\lstset{style=ccodestyle, frame=tb}
 
-pascalanalyzer ./myapp -c 1,32 -v 2 -o myoutput.json
+\begin{lstlisting}[label={lst:overheadcode}, language=C, caption={Code used to measure the overhead of instrumentation functions ({\tt analyzer\_start} and {\tt analyzer\_stop}) defined as $T_{measurement}$.}]
+int main(int argc, char** argv) {
+	get_time(begin);
+	#pragma omp parallel for
+	for (c=0; c<n_iterations; c++) {
+		analyzer_start(1);
+		usleep(1e4); // to simulate a simple operation
+		analyzer_stop(1);
+	} 
+	get_time(end);
+	time = begin-end;
+}
 \end{lstlisting}
 
-Example execution python
-\begin{lstlisting}
-from pascalanalyzer.pascalrun import PascalRun
-from pascalanalyzer.sensors.rapl import RAPL
-rapl_sensor = RAPL()
+The time required to execute $N$ %MDPI: Is the italic necessary? Please carefully check and confirm % yes
+calls to the {\tt analyzer\_start} and {\tt analyzer\_stop} functions was obtained by the routine {\tt gettimeofday}, from the {\tt sys/time.h} library.
+The algorithm ran with 1 %MDPI: We used the scientific notations, please confirm.
+% ok
+$\times~10^{4}$, 2 $\times~10^{4}$, 3 $\times~10^{4}$, 4 $\times~10^{4}$, 5 $\times~10^{4}$, and 1 $\times~10^{5}$ calls to the pair of functions, each test repetited ten times, and the mean, median, and variance values computed.
+\cref{fig:overhead} shows these results.
 
-app = PascalRun(
-	application="myapp",
-	repetitions=5,
-	sensors=[rapl_sensor],
-	idle_time=30,
-	governor="userspace",
-	disable_cores=True,
-	disable_hyperthread=True,
-)
-app.run(
-	cores=[1,16,32],
-	frequencies=[3000000,2800000],
-	inputs=[["arg1,"arg2"], ["arg3,"arg4"]]
-)
-app.save("output_file")
+\begin{figure}[H]
+\centering
+\captionsetup{justification=centering}
+\begin{subfigure}[b]{0.45\textwidth}	
+	\includegraphics[width=\textwidth]{pascalanalyzer/figures/designfeatures/instrumentation_overhead_summary.pdf}
+	\caption{\centering}
+	\label{fig:overhead_1}
+\end{subfigure}
+%
+\begin{subfigure}[b]{0.46\textwidth}
+	\includegraphics[width=\textwidth]{pascalanalyzer/figures/designfeatures/instrumentation_overhead_boxplot.pdf}
+	\caption{\centering}
+	\label{fig:overhead_2}
+\end{subfigure}
+\caption{Measuring variance of the time to single instrumentation, 
+	%MDPI: 1. For this figure, please use the scientific notations.
+	%          2. Please add commas to numbers in the Ox axis of the figures, e.g.: 20,000 30,000 40,000 etc. - CHANGED
+	i.e., a call to {\tt analyzer\_start} and {\tt analyzer\_stop} while varying the number of measurements.
+(\textbf{a}) Time consumed from sampling a region one time in seconds.
+%MDPI: We moved the explanations of subfigures here, please confirm.
+	(\textbf{b}) Box plot showing the statistics of the sampling time of a single region.}
+\label{fig:overhead}
+\end{figure}
+
+In Figure \ref{fig:overhead}a, we can see the mean, median, and variance of the time for a single call to our instrumentation function while varying the number of calls/measurements.
+The Figure \ref{fig:overhead}a complements these results showing the variance in each execution.
+%From these results, we can approximate the average value of $T_{measurement}$ to be $4e^{-6}$ seconds in this system.
+In other words, for each one million measurements, an overhead of approximately four seconds is added to the total program execution time.
+
+\cref{tab:overhead} shows the results from the same experiment above comparing the time with and without the analyzer.
+We can observe that proportional impact (overhead) was constant while the number of interactions increased.
+\cref{tab:overhead} also presents the data referring to the simulation using the TAU profiling tool.
+For this simulation, we replaced the analyzer directives with the time measurement directives ({\tt TAU\_PROFILE\_START} and {\tt TAU\_PROFILE\_STOP}) of TAU, which allowed us to approximate the measurement and analysis conditions.
+
+\begin{table}[H]
+\caption{Instrumentation overhead estimation varying the number of samples collected for TAU and~Analyzer.}
+\label{tab:overhead}
+%\centering
+\newcolumntype{C}{>{\centering\arraybackslash}X}
+\begin{tabularx}{\textwidth}{CCCCCC}\toprule
+	\multirow{2}{*}{\vspace{-6pt}\textbf{Iterations}} & \multicolumn{3}{c}{\textbf{Time (s)}} & \multicolumn{2}{c}{\textbf{Overhead (\%)}} \\ \cmidrule{2-6}
+	& \textbf{Real Time} & \textbf{TAU} & \textbf{Analyzer} & \textbf{TAU} & \textbf{Analyzer} \\ \midrule
+	10,000	& 100.933	& 100.992	& 101.053	& 0.058	& 0.118 \\
+	20,000	& 201.863	& 201.980	& 202.094	& 0.057	& 0.114 \\
+	30,000	& 302.797	& 302.977	& 303.142	& 0.059	& 0.114 \\
+	40,000	& 403.738	& 403.978	& 404.176	& 0.059	& 0.108 \\
+	50,000	& 504.675	& 504.959	& 505.212	& 0.056	& 0.106 \\
+	100,000	& 1009.359	& 1009.927	& 1010.432	& 0.056	& 0.106 \\ \midrule
+\end{tabularx}
+\end{table}
+
+\textls[-20]{From the results, it is possible to see that the tool proposed in this work has a higher overhead than that presented by the TAU, but the analysis capabilities are distinct.
+TAU adds the individual runtime of each thread to define the execution time of a parallel region.
+This strategy does not consider the simultaneous action of threads in processing instructions and can count the same period of time more than once, damaging a precise measurement of the parallel region.
+The analyzer works around this problem because the intersection periods are counted just once.
+In this sense, although the TAU has lower overhead, a scalability analysis depends on a more accurate measurement like the one proposed in this work.}
+\newpage
+Varying the number of threads impacts the cost of instrumentation.
+As shown in \mbox{\cref{tab:overhead2}}, there is a trend for the percentage of overhead to increase concerning the application's runtime.
+However, it is also possible to see that this increase is negligible, considering the exponential increase in the number of threads.
+The increase in the processing load, on the other hand, has a beneficial impact on the relationship between execution time and the percentage of overhead.
+This behavior occurs because the runtime will increase with the greater need for processing, and the overhead will only change if the number of threads is also changed.
+
+%The longer the execution time elapsed in each measurement, the smaller is the proportional impact of $T_{measurement}$ on the analysis as shown in Table .
+%(9.181-9.142)/10000 = 0.0039s/Iteration
+%(91.81-91.42)/100000 = 0.0000039s/Iteration
+%100*(9.181-9.142)/9.181 = 0.4247 %over
+%100*(91.81-91.42)/91.81 = 0.4247 %over
+%The instrumentation overhead generated when measuring a region is located in the time of the outermost region. 
+%This means that the execution time indicated by the tool for a region is only influenced by the instrumentation when there are more internal analysis regions. 
+%In practice, the implementation strategy used in this tool offers the user the possibility of verifying more precisely the performance of certain parts of the code.
+
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+%\begin{table*}
+%\centering
+%\caption{Overhead estimation varying the number of samples collected}
+%\label{tab:overhead}
+%\begin{tabular}{rrrrrrr}
+%\midrule
+%\multirow{2}{*}{Iterations} & \multicolumn{2}{c}{Time (s)} & \multicolumn{3}{c}{Time %estimation of single instrumentation (s)} & \multirow{2}{*}{Overhead (\%)} \\ %\cmidrule{2-6}
+% & without analyzer & with analyzer & median %& average & variance & \\ \midrule
+%10000 & 9.14 & 9.18 & 4.05e-06 & %4.03e-06 & 2.40e-14 & 0.44 \\
+%20000 & 18.28 & 18.36 & 4.03e-06 & %4.04e-06 & 1.24e-14 & 0.44 \\
+%30000 & 27.43 & 27.54 & 3.88e-06 %& 3.88e-06 & 1.35e-14 & 0.42 \\
+%40000 & 36.57 & 36.72 & 3.89e-06 %& 3.90e-06 & 2.22e-14 & 0.43 \\
+%50000 & 45.71 & 45.91 & 3.94e-06 %& 3.93e-06 & 2.52e-14 & 0.43 \\
+%100000 & 91.42 & 91.81 & 3.93e-06 %& 3.90e-06 & 2.10e-14 & 0.43 \\ \midrule
+%\end{tabular}
+%\end{table*}
+
+\begin{table}[H]
+\caption{Instrumentation overhead while varying the number of threads with the number of samples fixed to one million.}
+\label{tab:overhead2}
+%\centering
+\newcolumntype{C}{>{\centering\arraybackslash}X}
+\begin{tabularx}{\textwidth}{CCCC}\toprule
+	\multirow{2}{*}{\vspace{-6pt}\textbf{Threads}} & \multicolumn{2}{c}{\textbf{Time (s)}} & \multirow{2}{*}{\vspace{-6pt}\textbf{Overhead (\%)}} \\ \cmidrule{2-3}
+	& \textbf{Direct} & \textbf{with Analyzer} \\ \midrule
+	1 & 1009.360 & 1010.430 & 0.107 \\
+	2	& 504.733	 & 505.374	 & 0.127 \\
+	4	& 252.399	 & 252.742	 & 0.135 \\
+	8	& 126.215	 & 126.390	 & 0.138 \\
+	16	& 63.109	 & 63.199 & 0.143 \\ \bottomrule
+\end{tabularx}
+\end{table}
+
+For pluggable sensors, overhead is generally not a concern as they run on a separate thread and have very low CPU usage, which presents only a few scenarios where they can cause interference.
+
+\textls[-30]{One of these scenarios is where the application needs all the machine's resources at the same time that we have sensors that can respond faster than the processing speed at a given sample rate.
+In this case, the overhead would be directly associated with the sample rate and how the system handles concurrency.
+In general, this is rarer to happen in HPC (High-performance computing) since it would require an application with a perfect linear scaling.}
+
+Another possible scenario is I/O overhead, where some network, disk, or memory resource becomes unavailable due to high sensor usage.
+However, this scenario is even rarer as sensors seldom produce data that quickly and, in all the performed tests, this was not close to being an issue.
+
+\subsubsection{Features and Usage} \label{subsec:features_and_usage_pascal}
+
+The analyzer is a simple and easy-to-use tool that allows the user to understand the general behavior of a program before investing in efforts that require in-depth and sometimes longer-term analysis.
+However, to meet the objectives that best match the needs and resources of the user, the tool provides functionalities to parameterize the measurement process.
+Among the main ones are:
+\begin{enumerate}
+	\item Automated execution and deployment of parameterized code, actuators, and sensors
+	%Automated code execution with different actuators and sensors saving time of a manual process.
+	\item Automated low-overhead binary code instrumentation.
+	\item No need of the source code or to recompile, unless the user desires it. 
+	%Automated instrumentation in the binary with low overhead eliminating the need to access source code or recompile.
+	\item Flexible user-defined observation regions: analysis of contexts within and beyond parallel regions.
+	%The flexibility to identify specific observation regions: which allows the analysis of contexts beyond the blocks of instructions defined as parallel.
+	\item Automatic instrumentation of parallel and remaining intermediate serial regions.
+	%The ability to compute intermediate regions and automatically derive sensor values in certain parts of the program between user-defined regions.
+In scalability analysis, this becomes very useful since parallel regions can be instrumented automatically and the remaining regions can already be classified as serial, allowing the identification of all parts of the program.
+	\item Aggregation, by region, of the collected metrics, which significantly reduces the amount of data stored on disk and RAM usage; useful in cases where it would be impossible to store information about all sensor events, such as the instrumentation of a for loop with billions of iterations.
+In this case it is possible to choose how to group the data either by mean, median, minimum or maximum values.
+	\item Hierarchical regions: The proposed analyzer also facilitates the identification of aligned regions, enabling the identification of the calling hierarchy of inner regions and block analysis.
+\end{enumerate}
+\newpage
+The tool can be used in the command line or via its API. The API provides calls to integrate sensors and actuators as shown in the Listing \ref{lst:pascal_api}.
+%MDPI: Please provide the explanation % added in abbreviations
+
+\lstset{style=pythonStyle,frame=tb}
+\begin{lstlisting}[label={lst:pascal_api}, language=python, caption={Python script showing some API features provided by the tool, on how a custom run can be configured.}, captionpos=b]
+from analyzer.run import Run
+from analyzer.actuators import CPUFrequencies
+from analyzer.actuators import CPUCores
+from analyzer.sensors import RAPL
+from analyzer.sensors import fingerprint
+from itertools import product
+
+lsensors = [
+RAPL(), 
+fingerprint(counters=["INSTRUCTIONS"])
+]
+cores = [
+CPUCores(c) for c in range(1,32)
+]
+freqs = [
+CPUFrequencies(2800000), 
+CPUFrequencies(2600000)
+]
+
+# all combinations
+configs = list(product(cores,freqs))
+app = Run(application="a.out",
+repetitions=10, 
+instr_auto=True, 
+sensors=lsensors)
+app.run(configs)
+app.savedata("out.json")
 \end{lstlisting}
 
-Example model creating python
-\begin{lstlisting}
-from pascalanalyzer import PascalData
-from pascalanalyzer.pascalmodel import *
+\subsubsection{Exported Data Structure} \label{subsec:datastructure_pascal}
 
-data = PascalData(filename)
-df = data.energy()
+The data collected by the tool is exported to a .json file and stored on disk.
+The data structure is divided into two large groups.
+One group contains information about the configuration parameters driving deployment and execution.
+The other includes the performance measurements.
+The data is grouped in keys representing the processed input, the number of processing elements used in the execution, and the simulation ID.
+Since the tool can perform the same configuration many times, a simulation ID distinguishes runs that use the same input and number of processing elements.
 
-opt = LeastSquaresOptmizer("model= lambda x, N, p, f:"
-                           "x[0]*f**2*p+x[1]/(f*p)*N**x[7]+x[2]*f**2"
-                           "+x[3]*p+x[4]/f*N**x[6]+x[5]", 8)
+\cref{lst:datafile} presents an example file exported by the analyzer.
+There we can see the main parts of the exported structure, with the header where essential information about the system is written followed by the description of the collected data, where the list of actuators is presented as well as the specific type of information that was collected from the sensors.
+Finally the samples are presented separated by actuator configurations, and sensor type.
 
-pascalmodel= create_model(df,inputs=["input","cores","frequency"],
-                             output=["ipmi_energy"], model= opt,
-                             config=data.config, train_sz=0.9)
-\end{lstlisting}
+The keys and values in the .json file do not represent information that is easy for the user to identify and understand.
+However, it can be easily interpreted by a script or visualization software such as the PaScal Viewer~\cite{Silva2018}.
+The proposed tool provides a visualization module, responsible for interpreting and presenting data in an organized and user-friendly manner.
 
-\subsection{Data structure}
-
-\subsubsection{Proposals}
-\subsubsection{neasted dictonaries}
-\begin{lstlisting}
-{
-	key1:
-	{
-		key2:
-		{
-			key3:
-			{
-				data1: val1,
-				data2: val2
-			},
-			...
-		},
-		...
-	},
-	...
-}
-...
-\end{lstlisting}
-
-\subsubsection{index dictonary}
-\begin{lstlisting}
-{
-"key1;key2;key3;...": [
-	{
-		data1: val1,
-		data2: val2
-	},
-	...
-],
-}
-...
-\end{lstlisting}
-
-\subsubsection{Benchmarking}
-
-{\centering
-
-Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 10.00 b
-
-\begin{tabular}{llllr}
-	\toprule
-	{} & Storage &    Neasted &      Index &  Neasted/Index \\
-	\midrule
-	0 &     RAM &    2.57 Mb &    4.73 Mb &       0.542656 \\
-	1 &   JSON  &  616.52 Kb &    1.26 Mb &       0.490468 \\
-	2 &     PKL &  380.99 Kb &    1.21 Mb &       0.315070 \\
-	3 &      GZ &  221.96 Kb &  270.36 Kb &       0.820989 \\
-	\bottomrule
-\end{tabular}
-
-Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 100.00 b
-
-\begin{tabular}{llllr}
-	\toprule
-	{} & Storage &  Neasted &    Index &  Neasted/Index \\
-	\midrule
-	0 &     RAM &  4.73 Mb &  6.89 Mb &       0.686078 \\
-	1 &   JSON  &  2.78 Mb &  3.42 Mb &       0.812560 \\
-	2 &     PKL &  2.54 Mb &  3.37 Mb &       0.754198 \\
-	3 &      GZ &  1.82 Mb &  1.87 Mb &       0.968731 \\
-	\bottomrule
-\end{tabular}
-
-Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 1.00 Kb
-
-\begin{tabular}{llllr}
-	\toprule
-	{} & Storage &   Neasted &     Index &  Neasted/Index \\
-	\midrule
-	0 &     RAM &  26.33 Mb &  28.49 Mb &       0.924100 \\
-	1 &   JSON  &  24.38 Mb &  25.02 Mb &       0.974398 \\
-	2 &     PKL &  24.22 Mb &  25.04 Mb &       0.966930 \\
-	3 &      GZ &  17.58 Mb &  17.64 Mb &       0.996294 \\
-	\bottomrule
-\end{tabular}
-
-Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 10.00 Kb
-
-\begin{tabular}{llllr}
-	\toprule
-	{} & Storage &    Neasted &      Index &  Neasted/Index \\
-	\midrule
-	0 &     RAM &  242.33 Mb &  244.49 Mb &       0.991156 \\
-	1 &   JSON  &  240.38 Mb &  241.02 Mb &       0.997343 \\
-	2 &     PKL &  240.24 Mb &  241.07 Mb &       0.996565 \\
-	3 &      GZ &  175.11 Mb &  175.07 Mb &       1.000198 \\
-	\bottomrule
-\end{tabular}
-
-Keys: [32,10,10,10,10] 320.00 Kb  Data bytes: 1.00 b
-
-\begin{tabular}{llllr}
-	\toprule
-	{} & Storage &    Neasted &     Index &  Neasted/Index \\
-	\midrule
-	0 &     RAM &   13.38 Mb &  37.27 Mb &       0.358895 \\
-	1 &   JSON  &    5.23 Mb &  13.98 Mb &       0.373866 \\
-	2 &     PKL &    1.49 Mb &  12.70 Mb &       0.117686 \\
-	3 &      GZ &  537.08 Kb &   1.24 Mb &       0.431880 \\
-	\bottomrule
-\end{tabular}
-
-Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 10.00 b
-
-\begin{tabular}{llrr}
-	\toprule
-	{} &      Time &   Neasted &  Index \\
-	\midrule
-	0 &  Creation &  0.073802 &    0.0 \\
-	1 &     Acess &  0.000000 &    0.0 \\
-	\bottomrule
-\end{tabular}
-
-Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 100.00 b
-
-\begin{tabular}{llrr}
-	\toprule
-	{} &      Time &   Neasted &  Index \\
-	\midrule
-	0 &  Creation &  0.426866 &    0.0 \\
-	1 &     Acess &  0.000000 &    0.0 \\
-	\bottomrule
-\end{tabular}
-
-Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 1.00 Kb
-
-\begin{tabular}{llrr}
-	\toprule
-	{} &      Time &   Neasted &  Index \\
-	\midrule
-	0 &  Creation &  4.162496 &    0.0 \\
-	1 &     Acess &  0.000000 &    0.0 \\
-	\bottomrule
-\end{tabular}
-
-Keys: [16,10,5,3,10] 24.00 Kb  Data bytes: 10.00 Kb
-
-\begin{tabular}{llrr}
-	\toprule
-	{} &      Time &    Neasted &  Index \\
-	\midrule
-	0 &  Creation &  42.476255 &    0.0 \\
-	1 &     Acess &   0.000000 &    0.0 \\
-	\bottomrule
-\end{tabular}
-
-Keys: [1]*1000 1.00 Kb  Bytes: 10.00 Kb
-
-\begin{tabular}{llllr}
-	\toprule
-	{} & Storage &    Neasted &     Index &  Neasted/Index \\
-	\midrule
-	0 &     RAM &  314.94 Kb &  19.23 Kb &      16.380038 \\
-	1 &   JSON  &   23.89 Kb &  18.90 Kb &       1.264328 \\
-	2 &     PKL &   23.91 Kb &  18.92 Kb &       1.263904 \\
-	3 &      GZ &   10.34 Kb &  10.14 Kb &       1.019525 \\
-	\bottomrule
-\end{tabular}
-
-Keys: [1]*1000 1.00 Kb  Data bytes: 10.00 Kb
-
-\begin{tabular}{llrr}
-	\toprule
-	{} &      Time &   Neasted &  Index \\
-	\midrule
-	0 &  Creation &  0.003023 &    0.0 \\
-	1 &     Acess &  0.000000 &    0.0 \\
-	\bottomrule
-\end{tabular}
-
-}
-
-From the tests, we conclude that index dictionary is a better option for speed and nested dictionary for space, although if the data size is significantly higher than the number of keys it doesn't save a lot of space.
-Because of that, we choose to work with index.
-
-\subsubsection{Structure}
-The data structure holds
-
-\begin{lstlisting}
+\lstset{style=jsonstyle, frame=tb}
+\begin{lstlisting}[label={lst:datafile}, language=json, caption={Sample of exported data file showing the internal structure and organization of the data.}]
 {
 	"config": {
-        "pkg": "./openmp_simple_bin",
-        "execdate": "29-03-2020_13:07:07",
-        "command": "run_test.py TestMonitor.test_valid_combinations",
-        "hostname": "MSI",
-        "kernel": "Linux-4.4.0-18362-Microsoft-x86_64-with-Ubuntu-18.04-bionic",
-        "disable_hyperthread": true,
-        "disable_cores": true,
-        "idle_time": 0,
+		****************** Header ******************
+		"pkg": "./application", 
+		"execdate": "00-00-0000_00:00:00",
+		"kernel": "Linux-4.4",
+		"command": "analyzer ...",
+		"hostname": "host",
+		...
+		******** Data collected description ********
 		"data_descriptor": {
-            "values": [ ... ],
+			******** Mandatory collected values ********
+			"values": [ 
+			"start_time", "stop_time", ... 
+			],
+			************** Actuators list **************
+			"keys": [ "A1", "A2" ...],
+			********* Sensor data description **********
 			"extras": {
-                "regions": {
-                "values": [ ... ]
+				"regions": { ... }
+				"sensors": { ... }
 			}
 		},
-        "keys": [ ... ]
+		"arguments": ["input_1", ...]
 	},
-    "arguments": [ ... ],
-    "governor": "userspace"
-    },
+	************** Data collected **************
 	"data": {
-        "4;1.2e-06;1;1": {
-            "regions": {
-                "0x1": [
-                        [ ... ],
-                        ...
-                    ],
-                ...
-            },
-            ...,
-            "total_time": 0.3075871467590332
-        },
-        ...
+		***** Actuator values separated by ";" *****
+		"X;Y;Z": {
+			********* Data collected by sensor *********
+			"regions": { ... },
+			"sensors": { ... },
+			"start_time": ti,
+			"stop_time": tf
+		}
 	}
@@
 \end{lstlisting}
-
-\subsection{Overhead}
-\subsubsection{BlackSchoels test}
-
-$number\_of\_tests=96$\\
-$max\_number\_threads=32$\\
-$type\_of\_step=power$\\
-$value\_of\_step=2$\\
-$list\_of\_args=\{in\_312K.txt \quad outputFile.txt\}$
-
-\begin{figure}[H]
-	\centering
-	\begin{subfigure}[b]{0.45\textwidth}
-	\includegraphics[width=\textwidth]{pascalanalyzer/figures/black100b}
-	\caption{With SPERF}
-	\label{fig:black100b}
-	\end{subfigure}
-	%
-	\begin{subfigure}[b]{0.45\textwidth}
-	\includegraphics[width=\textwidth]{pascalanalyzer/figures/black2100b}
-	\caption{With Linux Time}
-	\label{fig:black2100b}
-	\end{subfigure}
-\end{figure}
-
-\subsubsection{Freqmine test}
-
-$number\_of\_tests=96$\\
-$max\_number\_threads=32$\\
-$type\_of\_step=power$\\
-$value\_of\_step=2$\\
-$list\_of\_args=\{kosarak\_990k.dat \quad 1200 \quad output.txt\}$
-
-\begin{figure}[H]
-	\centering
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/freq100b}
-		\caption{With SPERF}
-		\label{fig:freq100b}
-	\end{subfigure}
-	%
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/freq2100b}
-		\caption{With Linux Time}
-		\label{fig:freq2100b}
-	\end{subfigure}
-\end{figure}
-
-\subsubsection{Intrusiveness measurement}
-
-The statistical result obtained from the test above from Blackscholes and Freqmine was shown in the tables \ref{tab:blackstatistics} and \ref{tab:freqstatistics}.Where we can compare the result mean and standard deviation from two groups one with SPERF and the other without.
-
-\begin{table}[H]
-	\centering
-	\caption{Statistics for BlackSchoels}
-	\label{tab:blackstatistics}
-	\begin{tabular}{llll}
-		& With Sperf & Without Sperf & Total    \\
-		N               & 96         & 96            & 192      \\
-		$\sum{X}$       & 100,19     & 103,0134      & 203,2034 \\
-		Mean            & 1,0436     & 1,0731        & 1,0584   \\
-		$\sum{X^2}$     & 105,5093   & 112,4857      & 217995   \\
-		Std.Dev.        & 0,0998     & 0,1431        & 124     
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-	\centering
-	\caption{Statistics for Freqmine}
-	\label{tab:freqstatistics}
-	\begin{tabular}{llll}
-		& With Sperf & Without Sperf & Total    \\
-		N           & 96         & 96            & 192      \\
-		$\sum{X}$   & 202,8016   & 205,44        & 408,2416 \\
-		Mean        & 2,1125     & 2,14          & 2,1263   \\
-		$\sum{X^2}$ & 429,9198   & 440,428       & 870,3478 \\
-		Std.Dev.    & 0,1256     & 0,091         & 0,1102  
-	\end{tabular}
-\end{table}
-
-Given the summary statistics, the calculations of the hypothesis test are shown in tabular form in the tables \ref{tab:blackAnova} and \ref{tab:freqsAnova}.
-
-\begin{table}[H]
-	\centering
-	\caption{Anova f test}
-	\label{tab:blackAnova}
-	\begin{tabular}{lllll}
-		Source             & SS     & df  & MS     &             \\
-		Between-treatments & 0,0415 & 1   & 0,0415 & F = 2,72671 \\
-		Within-treatments  & 2893   & 190 & 0,0152 &             \\
-		Total              & 2,9345 & 191 &        &            
-	\end{tabular}
-\end{table}
-
-\begin{table}[H]
-	\centering
-	\caption{Anova f test}
-	\label{tab:freqsAnova}
-	\begin{tabular}{lllll}
-		Source             & SS     & df  & MS     &             \\
-		Between-treatments & 0,0363 & 1   & 0,0363 & F = 3.01548 \\
-		Within-treatments  & 2,2844 & 190 & 0,012  &             \\
-		Total              & 2,3206 & 191 &        &            
-	\end{tabular}
-\end{table}
-
-If we take a significance level of 5\% the critical value of F with 1 and 190 degrees of freedom on numerator and denominator respectively is 3.8909.
-Size booth values of F calculated is lower then the critical there is no significant statistical difference on the mean value according to ANOVA.
-
-%\begin{figure}[H]
-%	\includegraphics[width=7cm,height=4.5cm]{ftest}
-%	\caption{F distribution}
-%	\label{fig:Fdist}
-%\end{figure}
-
-\subsubsection{Sampling}
-
-Energy estimation method for power samples
-Time difference
-\begin{equation}
-	E= \sum_i^{N-1}(t_{i+1}-t_i)p_i+s_N(t_{tot}-t_N)
-\end{equation}
-
-Constant sample time
-\begin{equation}
-E= s_r\sum_i^{N}p_i=TP_{avg}
-\end{equation}
-
-Error due to sampling
-\begin{itemize}
-	\item Time difference $\approx$ 4\%
-	\item Constant sample time 	$\approx$ 1\%
-\end{itemize}
-
-\subsection{Database}
-
-\begin{outline}
-	\1 Advantages
-		\2 No need to organize the files
-		\2 Faster and easier queries
-		\2 Shared data across programs and research
-	\1 Disadvantages
-		\2 Need of db server
-		\2 Need of network
-		\2 Non local server maybe slow ( need to cache )
-\end{outline}

### 2022-04-14 thesis/conclusion.tex
add missing tex files

\section{Conclusion} \label{sec:conclusion}
This paper proposes an energy model based on the operating frequency and the number of cores for a shared memory system. This model serves as a reference for DVFS and DPM optimization problems.

Results from three different HPC benchmarks demonstrate the potential of the proposed model while consuming 10 times less energy than a machine learning approach, such as SVR, to characterize applications. Moreover, it can provide knowledge-based hints to improve DVFS and DPM algorithms by enabling analysis of the contribution of each model parameter (e.g., level of parallelism) to the energy consumption. Indeed, as shown in Section \ref{subsec:dvfs_optmin}, when no oracle is available to choose the frequency and the number of cores the application should use, the proposed model can save around 12\% of energy for a random choice and up to 70\% for the worse possible choice. Considering the job history of our own HPC center, which shows the prevalence of worse possible choices made by users, the potential energy savings are very significant and encourage further research.

Although the model is promising, it still has some limitations. The main one is related to the input size, which needs to be estimated to create the application model and optimize the application. Another limitation concerns the power model, which does not consider the load variation, so our model ends up using an average of the energy consumption, which is enough to obtain good results but limits its implementation in real-time optimization.
Future research is intended to solve both problems, first adapting the model to use the ratio of executed instructions as input size, something which is more tangible and easy to measure in modern systems without much overhead, and adding  new parameters to the power model to account for the load. This would allow us to develop more advanced DVFS models that could identify different phases of a target program with more subtle changes in frequency, and, perhaps, in the number of active cores to  further improve the results presented here.

%Future work will demonstrate all possible analyzes achievable with the proposed equation as more advanced DVFS models can be derived from it. For instance, identifying the different phases of a target program will allow more subtle changes in frequency and, perhaps, in the number of active cores to improve further the results presented here.

This work proposes an energy model based on the frequency and number of cores for a full shared memory system. This model can serve as a base for DVFS and DPM optimization problems that include both frequency and active cores, as well for analysis of the contribution of each parameter (ex: parallelism level) to the energy consumption.

Results from 3 HPC benchmarks running in one cluster demonstrate the potential of the proposed novel model. While consuming less energy than traditional machine learning approaches, it can serve as bases from DVFS and DPM algorithm as shown in the \ref{subsec:dvfs_optmin} in an average case saving about 12\% up to 69\%. The previous knowledge of the application's performance can expose sufficiently relevant information, such as parallel speedups, that is harder to guess in run-time techniques based on DVFS.

A weakness of the proposed model is the need for information about the input size of the application that can be complex to derive. A possible solution would be to precisely define what is input size, given an definition in function of common variable for all applications like throughput for example. Future work will demonstrate all the possible analysis that is possible to archive with the equation as long as more advanced DVFS models that can be derived using the equation. For instance identification of different phases of the target program and thus, it will enable more fine-grained changes of the frequency and, perhaps, the number of active cores, to further improve the results presented here.

Another important aspect that is typically not taken into account is the number of processing cores to be used by a parallel program. This choice is left to the user, which often is not trivial as shown in this paper.

\subsection{Conclusions and Future Works Pascal}\label{sec:conclusions_pascal}

This work introduces a practical and easy-to-use tool for measuring and analyzing the efficiency of parallel applications. The proposed tool focuses on observing scalability and energy consumption and implements features that enables analysis at hierarchical levels of the program's inner parts. It also simplifies the comparison of the application runs in different configurations, helping developers target software optimization efforts.

The tool has a low level of intrusion added to the program's performance measurement under analysis, which is a fundamental aspect of understanding the program's behavior and scalability capacity. 
Although it does not offer graphic elements to visualize the collected metrics, the tool itself makes it possible to examine the data from the command terminal. It also allows this data to be analyzed by an interface provided in Python or through .json files. 
%Other tools, such as PaScal Viewer, can compensate for the absence of these graphic resources. Indeed, PaScal Viewer is a web-based visualization tool focused on analyzing the scalability trend of parallel applications, which has been integrated with PaScal Analyzer in a complementary way.

\textls[-20]{In future work, we intend to evolve this tool to include the ability to predict speedup and efficiency from a few samples using state-of-the-art prediction models present in the \mbox{literature \cite{Alex2020WhenParallelSpeedups, Vitor2022AnalyticalEnergyModel}}. The idea is to present the general behavior of the program and its scalability trend and reduce the execution time necessary to compose a comprehensive analysis. Furthermore, we intend to include features that allow the observation of parallel applications in distributed environments that use the Message Passing Interface (MPI) standard.}

\subsection{Conclusion fingerprint}

%\cite{Processors2012, Gregg2017}
From the results, we observe that the API present overhead similar or lower to other low-level APIs, with the advantage of being in high abstraction and simplified configuration with a few lines of code, is possible to configure and gather counter data.

The tool developed provide also provided a way to fingerprint programs and compute similarities between different programs or the same program with different inputs. This can be useful to reduce applications spaces for benchmarks as was done in Polybench clustering but also to analyze the behavior of a parameter providing insights to the programmer to find a possible bottleneck.

We also provide a precise definition to input size that made possible fingerprint the programs, but it also can help programmers of benchmark applications to better create inputs with more precise growth of a particular parameter.

%\subsection{FUTURE WORK}
%
%We pretend to use this tool to create a data set of applications behavior and automatically classify any program as belonging to a cluster or a set of clusters. The idea is to have a set of clusters that can describe most applications in this way we can know specific behaviors of the applications. This can be applied to various areas such as voltage scaling and frequency of the processor, once knowing the behavior of a particular program we can have an idea of what is the best strategy to control the frequency in order to save energy or increase performance.

### 2022-04-14 thesis/references.tex
add missing tex files

\begin{thebibliography}{999}

	\bibitem[Ishfag(2012)]{Group2012HandbookSahni}
	Ishfag, A.; Sanjay, R. \textit{Handbook of Energy-Aware and Green Computing}; Chapman \& Hall/CRC: London, England, 2012; Volume~1, pp. 702--713.

	\bibitem[Dayarathna(2016)]{Dayarathna2016DataSurvey}
	Dayarathna, M.; Wen, Y.; Fan, R. Data Center Energy Consumption Modeling: A Survey. {\em IEEE Commun. Surv. Tutor.} {\bf 2016}, {\em 18}, 732--794. [\href{http://doi.org/10.1109/COMST.2015.2481183}{CrossRef}]

	\bibitem[Corcoran(2017)]{Corcoran2017EmergingICT}
	Corcoran, P.; Andrae, A. \emph{Emerging Trends in Electricity Consumption for Consumer ICT}; National University of Ireland: Galway, Ireland, 2013; pp. 1--56.

	\bibitem[Mathew(2012)]{Mathew2012Energy-awareNetworks}
	Mathew, V.; Sitaraman, R. K.; Shenoy, P. Energy-aware load balancing in content delivery networks. {In Proceedings of the 2012 Proceedings IEEE INFOCOM, Orlando, FL, USA, 25--30 March 2012}; pp. 954--962.

	\bibitem[Rivoire(2007)]{Rivoire2007ModelsOptimizations}
	Rivoire, S.; Shah, M.A.; Ranganathan, P.; Kozyrakis, C.; Meza, J. Models and Metrics to Enable Energy-Efficiency Optimizations. {\em Computer} {\bf 2007}, {\em 40}, 39--48. [\href{http://dx.doi.org/10.1109/MC.2007.436}{CrossRef}]

	\bibitem[Buyya(2013)]{Buyya2013Introduction}
	Buyya, R.; Vecchiola, C.; Selvi, S.T. \textit{Mastering Cloud Computing}; Morgan Kaufmann Publishers Inc.: San Francisco, CA, USA, 2013; pp. 3--27

	\bibitem[Poess(2008)]{Poess2008EnergyCenters}
	Poess, M.; Nambiar, R.O. Energy cost, the key challenge of today's data centers. {\em Proc. VLDB Endow.} {\bf 2008}, {\em 1}, 1229--1240. [\href{http://dx.doi.org/10.14778/1454159.1454162}{CrossRef}]

	\bibitem[Gao(2013)]{Gao2013QualityCenters}
	Gao, Y.; Guan, H.; Qi, Z.; Wang, B.; Liu, L. Quality of service aware power management for virtualized data centers. {\em J. Syst. Archit.} {\bf 2013}, {\em 59}, 245--259. [\href{http://dx.doi.org/10.1016/j.sysarc.2013.03.007}{CrossRef}]

	\bibitem[Fan(2007)]{Fan2007PowerComputer}
	Fan, X.; Weber, W.D.; Barroso, L.A. Power provisioning for a warehouse-sized computer. {\em ACM SIGARCH Comput. Archit. News} {\bf 2007}, {\em 35}, 13--23. [\href{http://dx.doi.org/10.1145/1273440.1250665}{CrossRef}]

	\bibitem[Barroso(2007)]{Barroso2007TheComputing}
	Barroso, L.A.; H{\"{o}}lzle, U. The Case for Energy-Proportional Computing. {\em Computer} {\bf 2007}, {\em 40}, 33--37. [\href{http://dx.doi.org/10.1109/MC.2007.443}{CrossRef}]

	\bibitem[Malladi(2012)]{Malladi2012TowardsDRAM}
	Malladi, K.T.; Nothaft, F.A.; Periyathambi, K.; Lee, B.C.; Kozyrakis, C.; Horowitz, M. Towards energy-proportional datacenter memory with mobile DRAM. {In  Proceedings of the 2012 39th Annual International Symposium on Computer Architecture (ISCA), Portland, OR, USA, 9--13 June 2012}; pp. 37--48.

	\bibitem[Rotem(2012)]{Rotem2012Power-managementBridge}
	Rotem, E.; Naveh, A.; Ananthakrishnan, A.; Weissmann, E.; Rajwan, D. Power-Management Architecture of the Intel Microarchitecture Code-Named Sandy Bridge. {\em IEEE Micro} {\bf 2012}, {\em 32} 20--27. [\href{http://dx.doi.org/10.1109/MM.2012.12}{CrossRef}]

	\bibitem[Brown(2005)]{Brown2005ACPILinux}
	Brown, L.; Moore, R.; Li, D.S.; Yu, L.; Keshavamurthy, A.; Pallipadi, V. ACPI in Linux. {\em Symposium } {\bf 2005}, {\em 51}, 1--51.

	\bibitem[Hackenberg(2015)]{Hackenberg2015AnProcessor}
	Hackenberg, D.; Schone, R.; Ilsche, T.; Molka, D.; Schuchart, J.; Geyer, R. An Energy Efficiency Feature Survey of the Intel Haswell Processor. {In  Proceedings of the  2015 IEEE International Parallel and Distributed Processing Symposium Workshop, Hyderabad, India, 25--29 May 2015}; pp. 896--904.

	\bibitem[{Intel}(2020)]{Intel20200thLake}
	Intel. \textit{12th Generation Intel {\textregistered} Core™ Processors}; Intel: Santa Clara, CA, USA, 2020; pp. 420--430

	\bibitem[Shuja(2012)]{Shuja2012Energy-efficientCenters}
	Shuja, J.; Madani, S.A.; Bilal, K.; Hayat, K.; Khan, S.U.; Sarwar, S. Energy-efficient data centers. {\em Computing} {\bf 2012}, {\em 94}, 973--994. [\href{http://dx.doi.org/10.1007/s00607-012-0211-2}{CrossRef}]

	\bibitem[Benini(2000)]{Benini2000AManagement}
	Benini, L.; Bogliolo, A.; De~Micheli, G. A survey of design techniques for system-level dynamic power management. {\em IEEE Trans. Very Large Scale Integr. (VLSI) Syst.} {\bf 2000}, {\em 8}, 299--316. [\href{http://dx.doi.org/10.1109/92.845896}{CrossRef}]

	\bibitem[Merkel(2006)]{Merkel2006BalancingSystems}
	Merkel, A.; Bellosa, F. Balancing power consumption in multiprocessor systems. {\em ACM SIGOPS/EuroSys Eur. Conf. Comput. Syst.} {\bf 2006}, \emph{40}, 403--4014.

	\bibitem[Roy(2013)]{Roy2013AnAlgorithms}
	Roy, S.; Rudra, A.; Verma, A. An energy complexity model for algorithms. In Proceedings of the 4th conference on Innovations in Theoretical Computer Science, New York, NY, USA, 9--12 January 2013.

	\bibitem[Weaver(2008)]{Weaver2008CanTrusted}
	Weaver, V.M.; McKee, S.A. Can hardware performance counters be trusted? {In Proceedings of the 2008 IEEE International Symposium on Workload Characterization, Seattle, WA, USA, 14--16 September 2008}; pp. 141--150.

	\bibitem[Weaver(2013)]{Weaver2013Non-determinismImplementations}
	Weaver, V.M.; Terpstra, D.; Moore, S. Non-determinism and overcount on modern hardware performance counter implementations. {In Proceedings of the 2013 IEEE International Symposium on Performance Analysis of Systems and Software (ISPASS), Austin, TX, USA, 21--23 April 2013}; pp. 215--224.

	\bibitem[Das(2019)]{Das2019SoK:Security}
	Das, S.; Werner, J.; Antonakakis, M.; Polychronakis, M.; Monrose, F. SoK: The Challenges, Pitfalls, and Perils of Using Hardware Performance Counters for Security. {In Proceedings of the 2019 IEEE Symposium on Security and Privacy (SP), San Francisco, CA, USA, 19--23 May 2019}; pp. 20--38.

	\bibitem[Guire(2009)]{McGuire2009AnalysisKernel}
	Mc~Guire, N.; Okech, P.; Schiesser, G. Analysis of Inherent Randomness of the Linux Kernel. {In Proceedings of the Eleventh RealTime Linux Workshop, Dresden, Germany, 28–30 September 2009}.

	\bibitem[Ramos(2019)]{Ramos2019AnCounters}
	Ramos, V.; Valderrama, C.; Xavier~de Souza, S.; Manneback, P. An Accurate Tool for Modeling, Fingerprinting, Comparison, and Clustering of Parallel Applications Based on Performance Counters. {In Proceedings of the  IEEE International Parallel and Distributed Processing, Rio de Janeiro, Brazil, 20--24 May 2019}; pp. 797--804.

	\bibitem[Silva-de Souza(2020)]{Silva-de-Souza2020ContainergyAWorkloads}
	Silva-de Souza, W.; Iranfar, A.; Br{\'{a}}ulio, A.; Zapater, M.; Xavier-de Souza, S.; Olcoz, K.; Atienza, D. Containergy—A Container-Based Energy and Performance Profiling Tool for Next Generation Workloads. {\em Energies} {\bf 2020}, {\em 13}, ~2162. [\href{http://dx.doi.org/10.3390/en13092162}{CrossRef}]

	\bibitem[Shao and Brooks(2013)]{Shao2013EnergyProcessor}
	Shao, Y.S.; Brooks, D. Energy characterization and instruction-level energy model of Intel's Xeon Phi processor. In Proceedings of the  International Symposium on Low Power Electronics and Design (ISLPED), Beijing, China, 4--6 September 2013; pp. 389--394.

	\bibitem[Lewis(2008)]{Lewis2008Run-timeSystems}
	Lewis, A.; Ghosh, S.; Tzeng, N.F. Run-time energy consumption estimation based on workload in server systems. {In Proceedings of the 2008 Conference on Power Aware Computing and Systems, San Diego, CA, USA, 8--10 December 2008}; pp. 3--4.

	\bibitem[Mills(2014)]{Mills2014EnergySystems}
	Mills, B.; Znati, T.; Melhem, R.; Ferreira, K.B.; Grant, R.E. Energy Consumption of Resilience Mechanisms in Large Scale Systems. In Proceedings of the 2014 22nd Euromicro International Conference on Parallel, Distributed, and Network-Based Processing, Turin, Italy, 12--14 February 2014;  pp. 528--535.

	\bibitem[Feng(2003)]{Feng2003MakingSupercomputing}
	Feng, W.c. Making a Case for Efficient Supercomputing. {\em Queue} {\bf 2003}, {\em 1}, 54--64. [\href{http://dx.doi.org/10.1145/957717.957772}{CrossRef}]

	\bibitem[Sarwar(1997)]{Sarwar1997CmosCalculation}
	Sarwar, A. Cmos power consumption and cpd calculation. In {\em Proceeding: Design Considerations for Logic Products};  Texas Instruments: Dallas, TX, USA, {1997}.

	\bibitem[Butzen and Ribas(2007)]{Butzen2007LeakageGates}
	Butzen, P.; Ribas, R. {\em Leakage Current in Sub-Micrometer CMOS Gates}; {Universidade Federal do Rio Grande do Sul}: Porto Alegre, Brazil, { 2007}; pp. 1--30.

	\bibitem[Amdahl(1967)]{Amdahl1967ValidityCapabilities}
	Amdahl, G.M. {Validity of the single processor approach to achieving large scale computing capabilities}. In Proceedings of the Spring Joint Computer Conference on---AFIPS '67 (Spring), New York, NY, USA, 18--20 April 1967.

	\bibitem[Eyerman(2010)]{Eyerman2010ModelingDesign}
	Eyerman, S.; Eeckhout, L. {Modeling critical sections in Amdahl's law and its implications for multicore design}. In Proceedings of the 37th Annual International Symposium on Computer Architecture---ISCA '10, New York, NY, USA, 19--23 June 2010.

	\bibitem[Gustafson(1988)]{Gustafson1988ReevaluatingLaw}
	Gustafson, J.L. Reevaluating Amdahl's law. {\em Commun. ACM} {\bf 1988}, {\em 31}, 532--533. [\href{http://dx.doi.org/10.1145/42411.42415}{CrossRef}]

	\bibitem[Seel(2012)]{Hypothesis2012EncyclopediaLearning}
	Seel, N.M. \emph{Encyclopedia of the Sciences of Learning}; {Springer:  {Berlin/Heidelberg, Germany,} %newly added information, please confirm
	} {1988}; pp. 223--242.

	\bibitem[Roy(2019)]{Roy2019ForecastingNetwork}
	Roy, P.; Mahapatra, G.S.; Dey, K.N. Forecasting of software reliability using neighborhood fuzzy particle swarm optimization based novel neural network. {\em IEEE/CAA J. Autom. Sin.} {\bf 2019}, {\em 6}, 1365--1383. [\href{http://dx.doi.org/10.1109/JAS.2019.1911753}{CrossRef}]

	\bibitem[Zhu(2019)]{Zhu2019PredictingLearning}
	Zhu, W.; Liu, X.; Xu, M.; Wu, H. Predicting the results of RNA molecular specific hybridization using machine learning. {\em IEEE/CAA J. Autom. Sin.} {\bf 2019}, {\em 6}, 1384--1396. [\href{http://dx.doi.org/10.1109/JAS.2019.1911756}{CrossRef}]

	\bibitem[Rivoire(2008)]{Rivoire2008AModels}
	Rivoire, S.; Ranganathan, P.; Kozyrakis, C. A comparison of high-level full-system power models. {In Proceedings of the 2008 Conference on Power Aware Computing and Systems, San Diego, CA, USA, 8--10 December 2008}; pp. 1--5.

	\bibitem[Usman(2013)]{Usman2013ANoC}
	Usman, S.; Khan, S.U.; Khan, S. A comparative study of voltage/frequency scaling in NoC. {In Proceedings of the  IEEE International Conference on Electro-Information Technology, Rapid City, SD, USA, 9--11 May 2013}; pp. 1--5.

	\bibitem[Paolillo(2018)]{Paolillo2018OptimisationParallelism}
	Paolillo, A. Optimisation of Performance Metrics of Embedded Hard Real-Time
	Systems using Software/Hardware Parallelism, Ph.D. Thesis, Université libre de Bruxelles, Brussels, Belgium, 2018.

	\bibitem[Kim(2015)]{Kim2015RacingHeuristics}
	Kim, D.H.; Imes, C.; Hoffmann, H. Racing and Pacing to Idle: Theoretical and Empirical Analysis of
	Energy Optimization Heuristics. {In Proceedings of the  2015 IEEE 3rd International Conference on Cyber-Physical Systems, Networks, and Applications, Hong Kong, China, 19--21 August 2015}; pp. 78--85.

	\bibitem[Fu(2018)]{Fu2018RaceMinimization}
	Fu, C.; Chau, V.; Li, M.; Xue, C.J. Race to idle or not: Balancing the memory sleep time with DVS for energy minimization. {\em J. Comb. Optim.} {\bf 2018}, {\em 35}, 860--894. [\href{http://dx.doi.org/10.1007/s10878-017-0229-7}{CrossRef}]

	\bibitem[Rauber(2014)]{Rauber2014EnergyScaling}
	Rauber, T.; R{\"{u}}nger, G.; Schwind, M.; Xu, H.; Melzner, S. Energy measurement, modeling, and prediction for processors with frequency scaling. {\em J. Supercomput.} {\bf 2014}, {\em 70}, 1451--1476. [\href{http://dx.doi.org/10.1007/s11227-014-1236-4}{CrossRef}]

	\bibitem[Goel(2016)]{Goel2016AProcessors}
	Goel, B.; McKee, S.A. A Methodology for Modeling Dynamic and Static Power Consumption for Multicore Processors. {In Proceedings of the IEEE International Parallel and Distributed Processing Symposium, Chicago, IL, USA, 23--27 May 2016}; pp. 273--282.

	\bibitem[Du(2017)]{Du2017ModelingSystems}
	Du, Z.; Ge, R.; Lee, V.W.; Vuduc, R.; Bader, D.A.; He, L. Modeling the Power Variability of Core Speed Scaling on Homogeneous Multicore Systems. {\em Sci. Program.} {\bf 2017}, {\em 2017}, 1--13. [\href{http://dx.doi.org/10.1155/2017/8686971}{CrossRef}]

	\bibitem[Gonzalez(1997)]{Gonzalez1997SupplyCMOS}
	Gonzalez, R.; Gordon, B.; Horowitz, M. Supply and threshold voltage scaling for low power CMOS. {\em IEEE J. Solid-State Circuits} {\bf 1997}, {\em 32}, 1210--1216. [\href{http://dx.doi.org/10.1109/4.604077}{CrossRef}]

	\bibitem[Silva(2019)]{Silva2019Energy-OptimalApplications}
	Silva, V.R.G.; Furtunato, A.F.A.; Georgiou, K.; Sakuyama, C.A.V.; Eder, K.; Xavier-de Souza, S. Energy-Optimal Configurations for Single-Node HPC Applications. {In Proceedings of the  2019 International Conference on High Performance Computing \& Simulation (HPCS), Dublin, Ireland, 15--19 July 2019}; pp. 448--454.

	\bibitem[Kumar(1994)]{Kumar1994AnalyzingArchitectures}
	Kumar, V.; Gupta, A. Analyzing Scalability of Parallel Algorithms and Architectures. {\em J. Parallel  Distrib. Comput.} {\bf 1994}, {\em 22}, 379--391. [\href{http://dx.doi.org/10.1006/jpdc.1994.1099}{CrossRef}]

	\bibitem[Oliveira(2018)]{Oliveira2018ApplicationCharacterization}
	Oliveira, V.H.F.; Furtunato, A.F.A.; Silveira, L.F.; Georgiou, K.; Eder, K.; Xavier-de Souza, S. Application Speedup Characterization. {In Proceedings of the ACM/SPEC International Conference on Performance Engineering, Berlin, Germany, 9--13 April 2018}; pp. 43--44.

	\bibitem[Smola(2004)]{Smola2004ARegression}
	Smola, A.J.; Sch{\"{o}}lkopf, B. A tutorial on support vector regression. {\em Stat. Comput.} {\bf 2004}, {\em 14}, 199--222. [\href{http://dx.doi.org/10.1023/B:STCO.0000035301.49549.88}{CrossRef}]

	\bibitem[Kitts(2006)]{Kitts2006RegressionLecture}
	Kitts, B. Regression Trees Lecture. {\em Data Min.} {\bf 2006}, 6--7.

	\bibitem[Altman(1992)]{Altman1992AnRegression}
	Altman, N.S. An Introduction to Kernel and Nearest-Neighbor Nonparametric Regression. {\em  Am. Stat.} {\bf 1992}, {\em 46}, 175--185.

	\bibitem[Murtagh(1991)]{Murtagh1991MultilayerRegression}
	Murtagh, F. Multilayer perceptrons for classification and regression. {\em Neurocomputing} {\bf 1991}, {\em 2}, 183--197. [\href{http://dx.doi.org/10.1016/0925-2312(91)90023-5}{CrossRef}]

	\bibitem[Gao(2019)]{Gao2019DendriticPrediction}
	Gao, S.; Zhou, M.; Wang, Y.; Cheng, J.; Yachi, H.; Wang, J. Dendritic Neuron Model With Effective Learning Algorithms for Classification, Approximation, and Prediction. {\em IEEE Trans. Neural Netw. Learn. Syst.} {\bf 2019}, {\em 30}, 601--614. [\href{http://dx.doi.org/10.1109/TNNLS.2018.2846646}{CrossRef}]

	\bibitem[Schwenkler(2006)]{Schwenkler2006IntelligentInterface}
	Schwenkler, T.; Deutschland, S. Intelligent Platform Management Interface. In \emph{Sicheres Netzwerkmanagement}; Springer: Berlin/Heidelberg, Germany, 2006; pp. 169--207.

	\bibitem[Bienia(2008)]{Bienia2008TheSuite}
	Bienia, C.; Kumar, S.; Singh, J.P.; Li, K. The PARSEC benchmark suite. In Proceedings of the 17th international conference on Parallel architectures and compilation techniques---PACT '08, New York, NY, USA, 25--29 October 2008.

	\bibitem[Romano(2015)]{Romano2015OpenMC:Development}
	Romano, P.K.; Horelik, N.E.; Herman, B.R.; Nelson, A.G.; Forget, B.; Smith, K. OpenMC: A state-of-the-art Monte Carlo code for research and development. {\em Ann. Nucl. Energy} {\bf 2015}, {\em 82}, 90--97. [\href{http://dx.doi.org/10.1016/j.anucene.2014.07.048}{CrossRef}]

	\bibitem[Dongarra(1988)]{Dongarra1988TheExplanation}
	Dongarra, J.J. The LINPACK Benchmark: An explanation. In Proceedings of the 1st International Conference on Supercomputing, Athens, Greece, 8--12 June 1987.

	\bibitem[Pedregosa(2011)]{Pedregosa2011Scikit-learn:Python}
	Pedregosa, F; Varoquaux, G.; Gramfort, A.; Michel, V.; Thirion, B.; Grisel, O.; Blondel, M.;  Prettenhofer, P.; Weiss, R.; Dubourg, V.; et al. Scikit-learn: Machine Learning in {\{}P{\}}ython. {\em J. Mach. Learn. Res.} {\bf 2011}, {\em 12}, 2825--2830.

	\bibitem[Royer(2020)]{Royer2020AOptimization}
	Royer, C.W.; O’Neill, M.; Wright, S.J. A Newton-CG algorithm with complexity guarantees for smooth
	unconstrained optimization. {\em Math. Programm.} {\bf 2020}, {\em 180}, 451--488.

	\bibitem[Horyath(2008)]{Horyath2008Multi-mode}
	Tibor Horyath and Kevin Skadron.
	\newblock {Multi-mode energy management for multi-tier server clusters}.
	\newblock {\em Parallel Architectures and Compilation Techniques - Conference Proceedings, PACT}, 270--279, 2008.

	\bibitem[Demme(2013)]{Demme2013OnCounters}
	John Demme, Matthew Maycock, Jared Schmitz, Adrian Tang, Adam Waksman, Simha
	Sethumadhavan, and Salvatore Stolfo.
	\newblock {On the feasibility of online malware detection with performance
	counters}.
	\newblock {\em ACM SIGARCH Computer Architecture News}, 41(3):559, 2013.

	\bibitem[Eranian(2008)]{Eranian2008Perfmon2}
	Stephane Eranian.
	\newblock {Perfmon2: a standard performance monitoring interface for Linux}.
	\newblock {\em Slides, perfmon2 overview}, 2008.

	\bibitem[Hahnel(2012)]{Hahnel2012RAPL}
	Marcus H{\"{a}}hnel, Bj{\"{o}}rn D{\"{o}}bel, Marcus V{\"{o}}lp, and Hermann
	H{\"{a}}rtig.
	\newblock {Measuring energy consumption for short code paths using RAPL}.
	\newblock {\em ACM SIGMETRICS Performance Evaluation Review}, 40(3):13, 2012.

	\bibitem[Zamani(2012)]{Zamani2012ASystems}
	Reza Zamani and Ahmad Afsahi.
	\newblock {A study of hardware performance monitoring counter selection in
	power modeling of computing systems}.
	\newblock {\em 2012 International Green Computing Conference, IGCC 2012}, 2012.

	\bibitem[IPMI(2013)]{IPMI2013ConfigurationGuide}
	Published November.
	\newblock {IPMI Configuration User Guide}.
	\newblock 2017(November), 2013.

	\bibitem[Murtagh(2011)]{Murtagh2011WardsAlgorithm}
	Fionn Murtagh and Pierre Legendre.
	\newblock {Ward's Hierarchical Clustering Method: Clustering Criterion and
	Agglomerative Algorithm}.
	\newblock (June):1--20, 2011.

	\bibitem[Mucci(1999)]{Mucci1999PAPI}
	PJ~Mucci, Shirley Browne, Christine Deane, and George Ho.
	\newblock {PAPI: A portable interface to hardware performance counters}.
	\newblock {\em Proceedings of the department of defense HPCMP users group
	conference}, 32:7–10, 1999.

	\bibitem[Luo(2005)]{Luo2005PropertiesDifferentiators}
	Jianwen Luo, Kui Ying, Ping He, and Jing Bai.
	\newblock {Properties of Savitzky-Golay digital differentiators}.
	\newblock {\em Digital Signal Processing: A Review Journal}, 15(2):122--136,
	2005.

	\bibitem[Kufrin(2005)]{Kufrin2005Perfsuite}
	Rick Kufrin.
	\newblock {Perfsuite: An accessible, open source performance analysis
	environment for linux}.
	\newblock {\em Dans Presented at The 6th International Conference on Linux
	Clusters: The HPC Revolution}, 151(April):5, 2005.

	\bibitem[Knupfer(2011)]{Knupfer2011Scorep}
	Andreas Kn{\"{u}}pfer and Christian R{\"{o}}ssel.
	\newblock {Score-P – A Joint Performance Measurement Run-Time Infrastructure
	for}.
	\newblock 1--12, 2011.

	\bibitem[Jurman(2009)]{Jurman2009CanberraLists}
	Giuseppe Jurman, Samantha Riccadonna, Roberto Visintainer, and Cesare
	Furlanello.
	\newblock {Canberra distance on ranked lists}.
	\newblock {\em Proceedings, Advances in Ranking-NIPS 09 Workshop}, pages
	22--27, 2009.

	\bibitem[Hang(2017)]{Hang2017CubicApplications}
	Houjun Hang, Xing Yao, Qingqing Li, and Michel Artiles.
	\newblock {Cubic B-Spline Curves with Shape Parameter and Their Applications}.
	\newblock {\em Mathematical Problems in Engineering}, 2017:1--8, 2017.

	\bibitem[Intel(2013)]{Intel2013IntelGuide}
	{Intel}.
	\newblock {Intel{\textregistered} 64 and IA-32 Architectures Software
	Developer’s Manual, Volume 3 (3A, 3B {\&} 3C): System Programming Guide}.
	\newblock 3(253665):1--1386, 2013.

	%%%%%%%%%%%%%%

	\bibitem[Huck(2007)]{Huck2007}
	Huck, K.; Malony, A.; Shende, S.; Morris, A.
	\newblock Scalable, Automated Performance Analysis with TAU and PerfExplorer.
	\newblock In Proceedings of the PARCO, Aachen, Germany, 4--7 September 
	2007; Volume~15, pp. 629--636.

	\bibitem[Islam(2019)]{Islam2019}
	Islam, T.; Ayala, A.; Jensen, Q.; Ibrahim, K.
	\newblock Toward a Programmable Analysis and Visualization Framework for
	Interactive Performance Analytics.
	\newblock In Proceedings of the IEEE/ACM International Workshop on Programming and Performance
	Visualization Tools (ProTools), Denver, CO, USA, 17 November 2019; pp. 70--77.
	\newblock [\href{http://doi.org/10.1109/ProTools49597.2019.00015}{CrossRef}]

	\bibitem[Weber(2019)]{Weber2019}
	\textls[-20]{Weber, M.; Ziegenbalg, J.; Wesarg, B.
		\newblock Online Performance Analysis with the Vampir Tool Set. In {\em Tools
			for High Performance Computing 2017, Proceedings of the 11th International Workshop on Parallel Tools for High Performance Computing, Dresden, Germany, 11--12 September, 2017
		}; Springer International Publishing: Cham, Switzerland,
		2019; pp. 129--143.
		\newblock
	} [\href{http://dx.doi.org/10.1007/978-3-030-11987-4_8}{CrossRef}]

	\bibitem[Bergel(2019)]{Bergel2019}
	\textls[-30]{Bergel, A.; Bhatele, A.; Boehme, D.; Gralka, P.; Griffin, K.; Hermanns, M.A.;
		Okanović, D.; Pearce, O.; Vierjahn, T. \emph{Visual Analytics Challenges in
			Analyzing Calling Context Trees}; Springer: Cham, Switzerland, 2019; pp. 233--249.
		\newblock
	} [\href{http://dx.doi.org/10.1007/978-3-030-17872-7_14}{CrossRef}]

	\bibitem[Malony(2005)]{Huck2005}
	Huck, K.; Malony, A.
	\newblock PerfExplorer: A Performance Data Mining Framework For Large-Scale
	Parallel Computing.
	\newblock In Proceedings of the SC '05: Proceedings of the 2005 ACM/IEEE Conference on
	Supercomputing, Seattle, WA, USA, 12--18 November 2005; p.~41.
	\newblock [\href{http://dx.doi.org/10.1109/SC.2005.55}{CrossRef}]

	\bibitem[Geimer(2010)]{Geimer2010}
	Geimer, M.; Wolf, F.; Wylie, B.; {\'A}brah{\'a}m, E.; Becker, D.; Mohr, B.
	\newblock The Scalasca performance toolset architecture.
	\newblock {\em Concurr. Comput. Pract. Exp.} {\bf
		2010}, {\em 22}, 702--719. [\href{http://dx.doi.org/10.1002/cpe.1556}{CrossRef}]

	\bibitem[Malony(2006)]{Shende2006}
	Shende, S.S.; Malony, A.D.
	\newblock The Tau Parallel Performance System.
	\newblock {\em Int. J. High Perform. Comput. Appl.} {\bf 2006}, {\em 20},~287--311.
	\newblock [\href{http://dx.doi.org/10.1177/1094342006064482}{CrossRef}]

	\bibitem[Adhianto(2010)]{Adhianto2010}
	Adhianto, L.; Banerjee, S.; Fagan, M.; Krentel, M.; Marin, G.; Mellor-Crummey,
	J.; Tallent, N.R.
	\newblock HPCTOOLKIT: Tools for performance analysis of optimized parallel
	programs.
	\newblock {\em Concurr. Comput. Pract. Exp.} {\bf 2010},
	{\em 22},~685--701.
	\newblock [\href{http://dx.doi.org/10.1002/cpe.1553}{CrossRef}]

	\bibitem[Miller(1995)]{Miller1995}
	Miller, B.; Callaghan, M.; Cargille, J.; Hollingsworth, J.; Irvin, R.;
	Karavanic, K.; Kunchithapadam, K.; Newhall, T.
	\newblock The Paradyn parallel performance measurement tool.
	\newblock {\em Computer} {\bf 1995}, {\em 28},~37--46.
	\newblock [\href{http://dx.doi.org/10.1109/2.471178}{CrossRef}]

	\bibitem[Galobardes(2015)]{Galobardes2015}
	Galobardes, E.C.
	\newblock \emph{Automatic Tuning of HPC Applications. The Periscope Tuning Framework};
	\newblock Shaker: Herzogenrath, Germany, 2015.

	\bibitem[Pillet(2007)]{Pillet2007}
	Pillet, V.; Labarta, J.; Cortes, T.; Girona, S.
	\newblock PARAVER: A Tool to Visualize and Analyze Parallel Code. In Proceedings of the WoTUG-18: Transputer and Occam Developments, Manchester, UK, 9--13 April 2007.

	\bibitem[Brink(2020)]{Brink2020}
	Brink, S.; Lumsden, I.; Scully-Allison, C.; Williams, K.; Pearce, O.; Gamblin,
	T.; Taufer, M.; Isaacs, K.E.; Bhatele, A.
	\newblock Usability and Performance Improvements in Hatchet.
	\newblock In Proceedings of the IEEE/ACM International Workshop on HPC User Support Tools
	(HUST) and Workshop on Programming and Performance Visualization Tools
	(ProTools), Atlanta, GA, USA, 18 November 2020; pp. 49--58.
	\newblock [\href{http://dx.doi.org/10.1109/HUSTProtools51951.2020.00013}{CrossRef}]

	\bibitem[Silva(2018)]{Silva2018}
	Silva, A.B.N.; Cunha, D.A.M.; Silva, V.R.G.; Furtunato, A.F.A.; Souza, S.X.-d.-S.
	\newblock PaScal Viewer: A Tool for the Visualization of Parallel Scalability
	Trends.
	\newblock In Proceedings of the ESPT/VPA@SC, Dallas, TX, USA, 11--16 November 2018.

	\bibitem[Eriksson(2016)]{Eriksson2016}
	Eriksson, J.; Ojeda-may, P.; Ponweiser, T.; Steinreiter, T.
	\newblock \emph{Profiling and Tracing Tools for Performance Analysis of Large Scale Applications};
	\newblock PRACE---Partnership for Advanced Computing in Europe: Brussels, Belgium, 2016; 
	pp. 1--30.
	\newblock Available online: \url{https://prace-ri.eu/wp-content/uploads/WP237.pdf} (accessed on 12 January 2020).

	\bibitem[Roberts(2017)]{10.1007/978-3-319-58667-0_22}
	Roberts, S.I.; Wright, S.A.; Fahmy, S.A.; Jarvis, S.A.
	\newblock Metrics for Energy-Aware Software Optimisation.
	\newblock In \emph{High Performance Computing, Proceedings of the 32nd International Conference, ISC High Performance 2017, Frankfurt, Germany, 18--22 June 2017}; Kunkel, J.M., Yokota, R., Balaji, P.,
	Keyes, D., Eds.; Springer International Publishing: Cham, Switzerland, 2017; pp.
	413--430.

	\bibitem[Eastep(2017)]{10.1007/978-3-319-58667-0_21}
	Eastep, J.; Sylvester, S.; Cantalupo, C.; Geltz, B.; Ardanaz, F.; Al-Rawi, A.;
	Livingston, K.; Keceli, F.; Maiterth, M.; Jana, S.
	\newblock Global Extensible Open Power Manager: A Vehicle for HPC Community
	Collaboration on Co-Designed Energy Management Solutions.
	\newblock In \emph{High Performance Computing, Proceedings of the 32nd International Conference, ISC High Performance 2017, Frankfurt, Germany, 18--22 June 2017}; Kunkel, J.M., Yokota, R., Balaji, P.,
	Keyes, D., Eds.; Springer International Publishing: Cham, Switzerland, 2017; pp.
	394--412.

	\bibitem[Hackenberg(2014)]{7016382}
	Hackenberg, D.; Ilsche, T.; Schuchart, J.; Schöne, R.; Nagel, W.E.; Simon, M.;
	Georgiou, Y.
	\newblock HDEEM: High Definition Energy Efficiency Monitoring.
	\newblock In Proceedings of the Energy Efficient Supercomputing Workshop, New Orleans, LA, USA, 16 November 2014; pp. 1--10.
	\newblock [\href{http://dx.doi.org/10.1109/E2SC.2014.13}{CrossRef}]

	\bibitem[Roberts(2019)]{10.1145/3321551}
	Roberts, S.I.; Wright, S.A.; Fahmy, S.A.; Jarvis, S.A.
	\newblock The Power-Optimised Software Envelope.
	\newblock {\em ACM Trans. Archit. Code Optim.} {\bf 2019}, {\em 16}, 1--27.
	\newblock [\href{http://dx.doi.org/10.1145/3321551}{CrossRef}]

	\bibitem[Boehme(2016)]{Boehme2016}
	Boehme, D.; Gamblin, T.; Beckingsale, D.; Bremer, P.T.; Gimenez, A.; LeGendre,
	M.; Pearce, O.; Schulz, M.
	\newblock Caliper: Performance Introspection for HPC Software Stacks.
	\newblock In Proceedings of the SC '16: Proceedings of the International Conference for High
	Performance Computing, Networking, Storage and Analysis, Salt Lake City, UT, USA, 13--18 November 2016; pp. 550--560.
	\newblock [\href{http://dx.doi.org/10.1109/SC.2016.46}{CrossRef}]

	\bibitem[Intel(2021)]{Intel2021Vtune}
	Corporation, I.
	\newblock \text{Intel VTune}.
	\newblock Available online: \url{https://software.intel.com/vtune} (accessed on 15 February 2020).

	\bibitem[Houstis(1997)]{Pantazopoulos1997}
	Pantazopoulos, K.N.; Houstis, E.
	\newblock {Performance Analysis and Visualization Tools for Parallel Computing}.
	\newblock 1997.
	\newblock Available online: \url{https://docs.lib.purdue.edu/cstech/1346} (accessed on 20 May 2020).
	%MDPI: Please add the publisher and location.

	\bibitem[Ott(2010)]{Gerndt2010}
	Gerndt, M.; Ott, M.
	\newblock Automatic Performance Analysis with Periscope.
	\newblock {\em Concurr. Comput. Pract. Exp.} {\bf
		2010}, {\em 22},~736--748. [\href{http://dx.doi.org/10.1002/cpe.1551}{CrossRef}]

	\bibitem[Labarta(2005)]{Labarta2005}
	Labarta, J.; Gimenez, J.; Martínez, E.; González, P.; Servat, H.; Llort, G.;
	Aguilar, X.
	\newblock Scalability of Tracing and Visualization Tools. In Proceedings of the {International Conference ParCo}, Prague, Czech Republic, 10--13 September 2005;
	\newblock pp. 869--876.

	\bibitem[Bienia(2008)]{Bienia2008}
	Bienia, C.; Kumar, S.; Singh, J.P.; Li, K.
	\newblock {The PARSEC benchmark suite: Characterization and architectural
		implications}.
	\newblock {In Proceedings of the International Conference on Parallel
		Architectures and Compilation Techniques}, Toronto, ON, Canada, {25--29 October 2008}; pp. 72--81.
	\newblock [\href{http://dx.doi.org/10.1145/1454115.1454128}{CrossRef}]

	\bibitem[Alex(2020)]{Alex2020WhenParallelSpeedups}
	\newblock Furtunato, A.F.A.; Georgiou, K.; Eder, K.; Xavier-De-Souza, S.
	\newblock When Parallel Speedups Hit the Memory Wall.
	\newblock {\em IEEE Access} {\bf 2020}, {\em  8},  79225--79238. [\href{http://dx.doi.org/10.1109/ACCESS.2020.2990418}{CrossRef}]

	\bibitem[Vitor(2020)]{Vitor2022AnalyticalEnergyModel}
	\newblock Silva, V.R.G.; Valderrama, C.; Manneback, P.; Xavier-de-Souza, S.
	\newblock Analytical Energy Model Parametrized by Workload, Clock Frequency and Number of Active Cores for Share-Memory High-Performance Computing Applications.
	\newblock {\em  Energies} {\bf 2022}, {\em  15},  1213. [\href{http://dx.doi.org/10.3390/en15031213}{CrossRef}]

	%%%%%%

	\bibitem[Melo(2010)]{Melo2010Perf}
	A.~C. de~Melo, ``{The New Linux 'perf' Tools},'' {\em Linux Kongress}, 2010.

	\bibitem[Gonzalez(2021)]{Gonzalez2021PolyBench}Abella-Gonzalez, M., Carollo-Fernandez, P., Pouchet, L., Rastello, F. \& Rodrıiguez, G. PolyBench/Python: Benchmarking Python Environments with Polyhedral Optimizations. {\em Proceedings Of The 30th ACM SIGPLAN International Conference On Compiler Construction}. pp. 59-70 (2021)

	\bibitem[Nikolaev(2011)]{Nikolaev2011Perfctr}Nikolaev, R. \& Back, G. Perfctr-Xen: A Framework for Performance Counter Virtualization. {\em Proceedings Of The 7th ACM SIGPLAN/SIGOPS International Conference On Virtual Execution Environments}. pp. 15-26 (2011)

\end{thebibliography}

### 2022-04-27 thesis/experiments/experiments.tex
carlos revision 1

@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_r1.pdf}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:pv_regionscomparison_a}
@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_r2.pdf}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:pv_regionscomparison_b}
@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_whole.pdf}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:pv_regionscomparison_c}
@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_1.pdf}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:pv_regionscomparison_a_2}
@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_2.pdf}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:pv_regionscomparison_b_2}
@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_3.pdf}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:pv_regionscomparison_c_2}
@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/diagram_discretevalues.pdf}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:discretevalues}
@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/diagram_smoothedvalues.pdf}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:smoothedvalues}
@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/speedup_completo_rtview_2 (1).pdf}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:speedup_raytrace}
@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/speedup_completo_openmc_kernel_novo (1).pdf}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:speedup_openmc}
@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/energy_completo_rtview_2 (1).pdf}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:raytrace_en}
@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/energy_completo_openmc_kernel_novo (1).pdf}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:openmc_en}
@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_raytrace.png}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:efficiency_raytrace}
@@
 		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_openmc.png}
-		\caption{\centering}
+%		\caption{\centering}
 		\label{fig:efficiency_openmc}
@@
 % Here we can also observed again that applications with the same shape were classified as the same clusters.
-
-\section{Validation the model} \label{sec:experimentalvalidation}
-In this section, the models presented in~Sections \ref{sec:power_model} and~\ref{sec:performance_model} were validated with a benchmark specific for multi-core architectures.
-Additionally, in order to assess the modeling overhead and accuracy, our proposal was then compared to machine learning approaches.
-We compared against support vector regression (SVR)~\cite{Smola2004ARegression}, decision tree \cite{Kitts2006RegressionLecture}, k-nearest neighbors \cite{Altman1992AnRegression}, multilayer perceptron \cite{Murtagh1991MultilayerRegression}, and some new methods, such as Gao et al.
-\cite{Gao2019DendriticPrediction}.
-However, SVR was chosen as the most representative because it performed best in our tests without aggressive fine-tuning, as shown in \cref{fig:ml_models}.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\columnwidth]{experiments/figures/ml_models.pdf}
-	\caption{Average of the mean squared error for all applications of our study case Section \ref{sec:casestudyapplication}.}
-	\label{fig:ml_models}
-\end{figure}
-
-\subsection{Verifying Hypothesis}
-
-In this section, we validate whether the assumptions of our model are valid for the system used.
-
-\subsubsection{Frequency and Voltage Relation}
-One of the assumptions was that the frequency and the voltage have a linear relationship, as indicated by ~\cref{eq:f_v}.
-To verify that, we build an experiment that sets the frequency to a specific value while sampling the voltage using the APERF and MPERF registers that provide feedback on the current CPU frequency.
-The average result of the sampling voltages is shown in  \cref{fig:freq_volt_rel}, where we can observe a near-perfect linear relation.
-This is because manufacturers implement this curve in the processors, using tables that relate ranges of frequencies to voltages so that they can precisely define any curve that will better suit their design.
-
-\begin{figure}[H]
-	\centering
-	\captionsetup[subfigure]{justification=centering}
-	\includegraphics[width=\columnwidth]{experiments/figures/freq_volt_rel.pdf}
-	\caption{Frequency voltage relation.}
-	\label{fig:freq_volt_rel}
-\end{figure}
-
-\subsubsection{Input Size and Instructions}
-We ran the applications with different inputs assuming  linear growth in the amount of work for one input to the other when building our model.
-However, measuring and controlling the amount of work would require much instrumentation and tuning to find an input corresponding to a certain amount of work.
-Therefore, to build our models, we use the time to reference the amount of work, assuming that the work is proportional to the executing time.
-\cref{fig:time_input} corresponds to the verification of this supposition.
-
-\cref{tab:corr_in_time} shows that the assumption was reasonable since the average correlation was 0.96 for all applications, indicating that growth in the number of instructions  will follow the time.
-This was the case for all applications that we ran in our benchmark and should hold for any data parallelism type of application.
-\begin{figure}[H]
-	\centering
-	\captionsetup[subfigure]{justification=centering}
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/blackscholes.pdf}
-		\caption{Blackscholes.}
-		\label{fig:time_input_blackshoels}
-	\end{subfigure}
-	%
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/canneal.pdf}
-		\caption{Canneal.}
-		\label{fig:time_input_canneal}
-	\end{subfigure}
-
-	\hfill
-	\caption{Relation between time and instructions for each input size. }
-	\label{fig:time_input}
-\end{figure}
-\vspace{-6pt}
-
-\begin{table}[H]
-	\centering
-	\caption{Correlation of time and instructions for all applications.}
-	\begin{tabular}{ll}
-		\multicolumn{1}{l|}{\textbf{Application}} & \textbf{Correlation} \\ \hline
-		Blackscholes                                               & 0.99                                  \\
-		Bodytrack                                                  & 0.99                                  \\
-		Canneal                                                    & 0.99                                  \\
-		Dedup                                                      & 0.99                                  \\
-		Ferret                                                     & 0.96                                  \\
-		Fluidanimate                                               & 0.99                                  \\
-		Freqmineq                                                  & 0.99                                  \\
-		Openmc                                                     & 0.94                                  \\
-		Raytrace                                                   & 0.99                                  \\
-		Swaptions                                                  & 0.99                                  \\
-		Vips                                                       & 0.98                                  \\
-		x264                                                       & 0.99                                  \\
-		HPL                                                        & 0.79                                  \\ \hline
-	\end{tabular}
-	\label{tab:corr_in_time}
-\end{table}
-
-The next assumption was that the application's behavior was the same when varying the workload.
-This condition is necessary for using the model with an unknown input size because, if the behavior is the same, we can interpolate the known inputs.
-One way to verify this is to measure the rate of instructions per second normalized by the frequency, as shown in \cref{fig:fingerpritns}.
-\begin{figure}[H]
-	\centering
-	\captionsetup[subfigure]{justification=centering}
-
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/blackscholes.pdf}
-		\caption{Blacksholes.}
-		\label{fig:fp_baclscholes}
-	\end{subfigure}
-	%
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/canneal.pdf}
-		\caption{Canneal.}
-		\label{fig:fp_canneal}
-	\end{subfigure}
-
-	\hfill
-	\caption{Rate of instructions per second varying the input size normalized by the frequency.}
-	\label{fig:fingerpritns}
-\end{figure}
-
-\cref{fig:fingerpritns} shows that the applications have roughly the same curve when normalized; this also happens for all other applications in our benchmark.
-
-The final assumption is that the workload should also not vary depending on the number of cores or frequency.
-To verify, we measure the total number of executed instructions while varying the cores from 1 to 32.
-\cref{tab:cores_variation} shows the results.
-
-\begin{table}[H]
-	\centering
-	\caption{Variation of the number of instructions when changing the number of cores for the same input.}
-	%\begin{tabular}{c|c|c|c}
-	\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}cccc}
-		\toprule
-		& \textbf{Average} & \textbf{Standard}     & \textbf{Standard}\\
-		\multirow{-2}{*}{\textbf{Application}}& \textbf{Number of Instructions} & \textbf{Deviation}     & \textbf{Deviation} \textbf{(\%) }\\ \midrule
-		Vip          & $7.97 \times 10^{11}$     & $7.16 \times 10^{6}$  & 0.00 \\ %Please use Scientific notation.
-For example, “1.65e+04” should be changed to “1.65× 104”.
-Please confirm `%` whether to remove.
-		Openmc       & $8.17 \times 10^{7}$     & $ 1.65 \times 10^{4}$ & 0.02    \\ 
-		Rtview       & $9.91 \times 10^{12}$     & $ 1.55 \times 10^{9}$ & 0.02    \\ 
-		X264         & $4.52 \times 10^{11}$     & $ 5.81 \times 10^{7}$ & 0.01    \\ 
-		Bodytrack    & $1.86 \times 10^{12}$     & $ 3.95 \times 10^{10}$ & 2.13    \\ 
-		Fluidanimate & $2.09 \times 10^{12}$     & $ 8.44 \times 10^{10}$ & 4.04    \\ 
-		HPL          & $1.14 \times 10^{8}$      & $ 1.24 \times 10^{5}$ & 0.11    \\ 
-		Blackschole  & $3.75 \times 10^{12}$     & $ 1.40 \times 10^{9}$ & 0.04     \\ 
-		Dedup        & $1.02 \times 10^{11}$     & $ 5.74 \times 10^{7}$ & 0.06     \\ 
-		Swapti       & $2.43 \times 10^{12}$     & $ 8.87 \times 10^{8}$ & 0.04     \\ 
-		Canneal      & $1.19 \times 10^{11}$     & $ 4.46 \times 10^{7}$ & 0.04     \\ 
-		Freqmine     & $1.27 \times 10^{12}$     & $ 4.78 \times 10^{8}$ & 0.04     \\ 
-		Ferret       & $4.76 \times 10^{11}$     & $ 7.04 \times 10^{7}$ & 0.01     \\ \bottomrule
-	\end{tabular*}
-	\label{tab:cores_variation}
-\end{table}
-% Corrected
-
-\cref{tab:cores_variation} shows the standard deviation and what that corresponds to in terms of the total number of instructions as a percentage.
-
-The same test was performed for the frequency, varying from 1.2 to 2.2 GHz with 100~MHz steps. The results are shown in \cref{tab:freq_variation}.
-
-These results show that all the assumptions were reasonable, and we can safely move to the validation of the model's prediction.
-\begin{table}[H]
-	\caption{Variation of the number of instructions when changing the frequency for the same input.}
-	\centering
-	%\begin{tabular}{c|c|c|c}{\hsize}{@{\extracolsep{\fill}}cccc}
-	\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}cccc}
-		\toprule
-		& \textbf{Average} & \textbf{Standard}     & \textbf{Standard}\\
-		\multirow{-2}{*}{\textbf{Application}}& \textbf{Number of Instructions} & \textbf{Deviation}     & \textbf{Deviation} \textbf{(\%)} \\ \midrule
-		Vip          & $7.97 \times 10^{11}$      & $ 1.16 \times 10^{6}$  & 0.00      \\ 
-		Openmc       & $8.17 \times 10^{7}$       & $ 4.52 \times 10^{3}$  & 0.01      \\ 
-		Rtview       & $9.91 \times 10^{12}$      & $ 6.64 \times 10^{5}$  & 0.00      \\ 
-		X264         & $4.52 \times 10^{11}$      & $ 1.54 \times 10^{5}$  & 0.00      \\ 
-		Bodytrack    & $1.84 \times 10^{12}$      & $ 2.54 \times 10^{5}$  & 0.00      \\ 
-		Fluidanimate & $2.38 \times 10^{12}$      & $ 1.70 \times 10^{9}$  & 0.07      \\ 
-		HPL          & $1.14 \times 10^{8}$       & $ 5.95 \times 10^{3}$  & 0.01      \\ 
-		Blackschole  & $3.75 \times 10^{12}$      & $ 4.36 \times 10^{5}$  & 0.00      \\ 
-		Dedup        & $1.02 \times 10^{11}$      & $ 8.32 \times 10^{7}$  & 0.08      \\ 
-		Swapti       & $2.43 \times 10^{12}$      & $ 1.48 \times 10^{5}$  & 0.00      \\ 
-		Canneal      & $1.19 \times 10^{11}$      & $ 3.01 \times 10^{5}$  & 0.00      \\ 
-		Freqmine     & $1.27 \times 10^{12}$      & $ 3.70 \times 10^{8}$  & 0.03      \\ 
-		Ferret       & $4.76 \times 10^{11}$      & $ 5.63 \times 10^{7}$  & 0.01      \\\bottomrule
-	\end{tabular*}
-	\label{tab:freq_variation}
-\end{table}
-
-\subsection{Fitting the Models} \label{subsec:fitting}
-To find the parameters of  \cref{eq:en_final}, 10 uniformly random configurations of frequencies ($f$), cores ($p$) and inputs ($N$) were chosen from the range $1<=p<=32$, $1.2<=f<=2.2$ and $1<=N<=5$, respectively.
-The application was executed for each chosen configuration, and the measured energy and time values were collected.
-For the input size, if we assume that all CPU instructions take approximately the same time to execute, the number of basic operations will be directly correlated with the time.
-Thus, we can estimate the input size by looking at the execution time, allowing us to divide a large input size into several smaller ones, knowing their relationship, as performed in the work of Oliveira ~\cite{Oliveira2018ApplicationCharacterization}.
-The unity can also vary depending on the definition.
-For simplicity, we assign numbers from 1 to 10, increasing the problem linearly, so it is also possible to interpolate any input in between these values.
-
-For each configuration, samples of the power were collected using IPMI every 1 second.
-This sampling rate was chosen based on the magnitude of the mean run time of the applications, which is in the order of minutes.
-Therefore, this rate provides enough samples to measure average power.
-Additionally, timestamps and the total run time were collected.
-The total energy spent on each configuration is estimated by first interpolating the power samples using the first-order method and then integrating this function in the time.
-
-The model's parameters are calculated by solving an optimization problem of finding the values that minimize the squared error of the prediction to the measured values using the non-linear least-squares method.
-
-The Python library Scikit-Learn was used to build the SVR model~\cite{Pedregosa2011Scikit-learn:Python}.
-The SVR was trained using the same data used for parameter estimation of Equation (\ref{eq:en_final}) %\cref{eq:en_final}
-with a grid search used to find the best kernel function and the best values for the hyper-parameters penalty for the wrong ($C$) and ($\gamma$).
-For this data, the best function was the radial base function (RBF), and the hyper-parameters were $C=10^4$ and $\gamma=0.5$.
-
-\subsection{Measured versus Modeled Energy}
-\label{subsec:measuredversusmodeledenergy}
-
-To validate the model, we ran all possible configurations in the tested machine, varying the cores in a range of $1<=p<=32$, the frequency in $1.2<=f<=2.2$, and the input in $1<=N<=5$.
-The total number of configurations varies from 400 to over 1000 depending on the application, as some applications have restrictions on the number of cores that they can run.
-Once the data was collected, we computed the mean percentage error (MPE) according to the following equation:
-\begin{equation}
-	MPE = \frac{1}{N} \sum_i^N \frac{|y_{\rm estimated}-y_{\rm measured}|}{y_{\rm measured}}.
-	\label{eq:mpe}
-\end{equation}
-
-\subsubsection{Frequency $\times$ Cores}
-\cref{fig:en_eq_freq_cores_fc} plots the measured and modeled energy consumption for some of the applications modeled.
-In addition,  some of the possible shapes that the model can take while varying the number of active cores, and operating frequency, are shown.%please confirm that the intended meaning has been retained.
-% yes its the same
-
-\begin{figure}[H]
-	\centering
-	\captionsetup[subfigure]{justification=centering}
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_cores/completo_black_5.pdf}}
-		\caption{}
-		\label{fig:en_eq_black_fc}
-	\end{subfigure}
-	%
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_cores/completo_canneal_1.pdf}}
-		\caption{}
-		\label{fig:en_eq_canneal_fc}
-	\end{subfigure}
-
-	\caption{Example fit for a specific input size: Blackscholes (\textbf{a}) and Canneal (\textbf{b}).
-“measured values” are the sensor data, and “minimum energy” is the minimum energy model prediction.
-	}
-	\label{fig:en_eq_freq_cores_fc}
-\end{figure}
-\subsubsection{Frequency $\times$ Input}
-\cref{fig:en_eq_freq_inp_fi} plots the measured and modeled energy consumption for some of the applications modeled.
-The diagrams show some of the possible shapes that the model can take while varying the operating frequency, and input size.
-\begin{figure}[H]
-	\centering
-	\captionsetup[subfigure]{justification=centering}
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_inps/completo_black_5.pdf}}
-		\caption{}
-		\label{fig:en_eq_black_fi}
-	\end{subfigure}
-	%
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_inps/completo_canneal_1.pdf}}
-		\caption{}
-		\label{fig:en_eq_canneal_fi}
-	\end{subfigure}
-
-	\caption{Example fit for a specific input size: Blackscholes (\textbf{a}) and Canneal (\textbf{b}).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
-	}
-	\label{fig:en_eq_freq_inp_fi}
-\end{figure}
-\subsubsection{Cores $\times$ Input}
-\cref{fig:en_eq_core_inp_ci} plots the measured and modeled energy consumption for some of the applications modeled.
-The diagrams  show some of the possible shapes that the model can take while varying the number of active cores, and input size.
-\begin{figure}[H]
-	\centering
-	\captionsetup[subfigure]{justification=centering}
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/cores_inps/completo_black_5.pdf}}
-		\caption{}
-		\label{fig:en_eq_black_ci}
-	\end{subfigure}
-	%
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/cores_inps/completo_canneal_1.pdf}}
-		\caption{}
-		\label{fig:en_eq_canneal_ci}
-	\end{subfigure}
-
-	\caption{Example fit for a specific input size: Blackscholes (\textbf{a}) and Canneal (\textbf{b}).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
-	}
-	\label{fig:en_eq_core_inp_ci}
-\end{figure}
-
-\subsubsection{Comparison}
-The average results for each application were calculated using a model trained with only 10 configurations, and the comparison is displayed \cref{fig:mpe_svr_eq}.
-%We calculated the raw MPE values shown in ~\cref{tab:mpe_svr_eq}.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=.8\columnwidth]{models/figures/mpe_svr_eq.pdf}
-	\caption{Comparison of the mean percentage error between the proposed model and SVR. ``Model mean'' and ``SVR mean'' are the average of all MPE values for all applications.
-	}
-	\label{fig:mpe_svr_eq}
-\end{figure}
-
-\cref{fig:mpe_svr_eq} shows that the proposed model always performed better, with a lower MPE than SVR, when we were limited to 10 training points.
-This result is further explored in the next Section \ref{subsec:overhead}, where we undertake a comparison with different training sizes.
-The exact values are shown in \cref{tab:mpe_svr_eq}.
-
-\begin{table}[H]
-\centering
-\begin{tabular}{c|c|c}
-\hline
-Application  & Model & SVR   \\ \hline
-Ferret       & 5.25     & 12.49  \\ 
-Raytrace     & 6.36     & 11.95  \\
-Fluianimate  & 2.44     & 22.90  \\
-x264         & 8.28     & 15.33  \\
-Vips         & 7.54     & 10.80  \\
-Swaptions    & 6.54     & 18.57  \\
-Canneal      & 3.12     & 6.13   \\
-Dedup        & 8.85     & 13.70  \\
-Freqmine     & 2.44     & 3.24   \\
-Blackscholes & 2.18     & 11.00  \\
-HPL          & 7.47     & 12.75  \\
-Bodytrack    & 16.98    & 34.12  \\
-Openmc       & 11.15    & 24.34  \\
-\end{tabular}
-\caption{Comparison of the Mean Percentage Error between the proposed model and SVR: raw values.}
-\label{tab:mpe_svr_eq}
-\end{table}
-
-\subsection{Overheads on Training} \label{subsec:overhead}
-It is known that machine learning is data-driven; in that sense, the SVR model obtained using only 10 configurations could be improved, but what about the analytical model? 
-To answer that question, the proposed model and the SVR were also trained with a varying number of configurations.
-We then compared the MPE and the amount of energy spent to create each model. 
-This accuracy-energy trade-off is crucial since building models' energy overhead defeats the primary goal of saving power when running applications.
-\enlargethispage{.5cm}
-
-\begin{figure}[H]
-	\centering
-	\captionsetup[subfigure]{justification=centering}
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_ferret_4.pdf}}
-		\caption{MPE for Ferret.}
-		\label{fig:overhead_ferret}
-	\end{subfigure}
-	%
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_black_6.pdf}}
-		\caption{MPE for Blackscholes.}
-		\label{fig:overhead_black}
-	\end{subfigure}
-	%
-	\par\bigskip
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_x264_4.pdf}}
-		\caption{MPE for x264.}
-		\label{fig:overhead_x264}
-	\end{subfigure}
-	%
-	\begin{subfigure}[b]{0.45\textwidth}
-		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_dedup_4.pdf}}
-		\caption{MPE for Dedup.}
-		\label{fig:overhead_dedup}
-	\end{subfigure}
-
-	\caption{MPE of the case studies versus training size, comparing how many training points is necessary to reach an acceptable result.}
-	\label{fig:overheadapps}
-\end{figure}
-
-\cref{fig:overheadapps} shows the comparisons of MPE and energy spent to create each model for two selected applications.
-According to the results, the analytical model is very stable, not changing much as more data is added, while the SVR keeps reshaping to adapt to the data.
-The error of the analytical model is almost constant but that of the SVR, initially very high, drops as more data is used in the training process.
-\begin{figure}[H]
-	\centering
-	\captionsetup[subfigure]{justification=centering}
-	\begin{subfigure}[t]{0.45\textwidth}
-		\includegraphics[width=\columnwidth]{models/figures/overhead/overall_energy_10pts.pdf}
-		\caption{}
-		\label{fig:overall_overhead}
-	\end{subfigure}
-	%
-	%\par\bigskip
-	\begin{subfigure}[t]{0.45\textwidth}
-		\includegraphics[width=\columnwidth]{models/figures/overhead/overall_mpe_10pts.pdf}
-		\caption{
-		}
-		\label{fig:overall_MPE}
-	\end{subfigure}
-	\caption{Overall results for energy and MPE for each training size.
-(\textbf{a}) Average energy spent on all applications during model creation.
-The two curves are identical because the same data were used to adjust the SVR and the model.
-(\textbf{b}) MPE of all applications: SVR needs 10 times more data to have an MPE lower than the proposed model.}
-	\label{fig:overall_train}
-\end{figure}
-
-\cref{fig:overall_train} presents the overall results, with the mean energy overhead and MPE for all applications.
-The meeting point of the MPE for the SVR and the proposed model can be extracted from \cref{fig:overall_train}b.
-It shows that, in around 90 configurations, the SVR starts to have a smaller error.
-The cost of that is the linear increase in energy spent on training.
-The increase in energy, about 10 times more, can be observed in \cref{fig:overall_train}a.
-
-\subsection{Analysis}
-One of the most significant advantages of using an analytical model is the understanding of the problem that an equation provides, making many different kinds of analysis possible that are otherwise impossible with a machine learning model.
-In this section, we discuss one of the possible analyses.
-In the following figures, we try to understand the contribution of each parameter of the equation to the total energy consumption.
-
-For this analysis, we took the model of one of the applications and, varying one parameter of the equation, we display the energy versus performance (time) for all configurations.
-After that, we computed the Pareto frontier, a set of all Pareto efficient allocations, i.e., all the configurations where resources cannot be reallocated to make one individual better off without making at least one individual worse off.
-This gives us all the configurations where we have an optimal trade-off of performance and energy to choose from.
-
-\cref{fig:pareto_static} shows the Pareto frontier for several values for the static power parameter ($c_3$ in \cref{eq:en_final}) with configurations of frequency ranging from 1.2 to 5 GHz and cores from 1 to 64, so that we can also have an idea of what is the tendency when we increase the frequency and number of cores.
-
-\begin{figure}[H]
-	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_high.pdf}
-	\caption{Pareto frontier for several values of static power parameter.
-The arrows with blue heads indicate the maximum energy, while the arrows with a red head the minimal energy for each corresponding curve.}% In four-digit numbers, no comma or spaces to indicate thousand, whatever the numbers are in the table or body text, e.g., 4500; More than five-digit numbers, should have comma, e.g.
-59,400; 899,302.
-Please revise the figure.
-	\label{fig:pareto_static}
-\end{figure}
-%Fixed
-
-From this figure, we can see that when increasing the value of the static power parameter, the total energy consumption increases as expected.
-We can also observe that the values that minimize the total energy consumption tend to be high frequency and multiple cores.
-This is one of the consequences of increasing the static power factor.
-As the dynamic factor proportionally decreases, its variables tend to have less impact on total consumption, enabling configurations with high frequency and several cores.
-This also enables chip-level optimization for choosing components that change the ratio between static and dynamic power.
-
-\cref{fig:pareto_w} shows the Pareto frontier in the same ranges described before but for the parameter corresponding to the level of parallelism of the application ($w$ in  \cref{eq:en_final}).
-
-\begin{figure}[H]
-	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_w_high.pdf}
-	\caption{Pareto frontier for several values of static power parameter.
-The arrows with blue heads indicate the maximum energy, while the arrows with a red head, the minimal, for each corresponding curve.}% In four-digit numbers, no comma or spaces to indicate thousand, whatever the numbers are in the table or body text, e.g., 4500; More than five-digit numbers, should have comma, e.g.
-59,400; 899,302.
-Please revise the figure.
-	\label{fig:pareto_w}
-\end{figure}
-%Fixed
-
-In  \cref{fig:pareto_w}, we observe that, as the parallelism level increases the total energy decreases.
-The number of cores tends to be higher with a higher level of parallelism as expected, and the frequency shows an inverse relation.
-
-\subsection{DVFS and DPM Optimization} \label{subsec:dvfs_optmin}
-The effectiveness of the proposed approach during optimization was evaluated with a simple algorithm that finds the optimal frequency and number of active cores from the proposed equation.
-The results were then compared to the Linux default choices for power management.
-
-With \cref{eq:en_final}, it is possible to calculate energy consumption estimates for each possible configuration since there is a finite range of possible values for the frequency and number of cores.
-It is also possible to apply constraints on the execution time, frequency, and the number of active cores.
-Then, the configuration that minimizes energy consumption for a given input can be selected.
-The complete workflow is shown in \cref{fig:optim_workflow}.
-We can see that any optimization problem can be structured with our model and the system's constraints.
-In the following examples, the optimization problem that we build is to minimize the energy equation given the constraints of possible frequencies and the number of cores that our system can run.
-The algorithm selected to minimize was the newton-CG~\cite{Royer2020AOptimization}.
-
-Current HPC managers leave to the user the choice of how many cores to use. On this basis, three situations were analyzed in relation to the number of cores:
-
-\begin{enumerate}
-	\item Worst choice: number of cores that maximize the total energy consumed;
-	\item Random choice: energy consumed for a random choice of the number of cores;
-	\item Best choice: number of cores that minimize the total energy consumed (oracle).
-\end{enumerate}
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{models/figures/DVFS optim.pdf}
-	\caption{Optimization workflow showing how DVFS and DPM optimization could be implemented from ou model.}
-	\label{fig:optim_workflow}
-\end{figure}
-
-The default option for the Linux governor is Ondemand, and, by default, it has no DPM control for the number of active cores.
-As Ondemand only performs DVFS, \textls[-15]{for comparison, each application was executed with all available cores in the system, from 1 to 32.}
-
-\textls[-15]{Figures \ref{fig:energy_worst_case}--\ref{fig:energy_best_case}, show the energy savings with respect to Ondemand, i.e., $\frac{Ondemand-Model_{min}}{Ondemand}$} for the three cases described above.
-The savings and losses for each case are:
-
-\begin{enumerate}
-	\item Worst choice: save 69.88\% on average;
-	\item Random choice: save 12.04\% on average;
-	\item Best choice: lost 14.06\% on average.
-\end{enumerate}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\columnwidth]{models/figures/dvfs_cmp_max.pdf}
-	\caption{Energy savings comparisons between the proposed model and the Worst case.}
-	\label{fig:energy_worst_case}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\columnwidth]{models/figures/dvfs_cmp_mean.pdf}
-	\caption{Energy savings comparisons between the proposed model and the Random case.}
-	\label{fig:energy_mean_case}
-\end{figure}
-
-By default, operating systems do not implement DPM at the core level, and, in HPC, the user usually explicitly chooses the number of cores to run their job.
-To give a better idea of the impact on the energy consumption of DPM at the core level, we analyzed the choices of the number of cores over a period of one year in the HPC center at UFRN.
-The result is plotted in~\cref{fig:cpu_requests}.
-
-It is of note that the most common choice of many regular users is a single core requested per job, matching the worst-case choice for all applications analyzed in this investigation.
-The best choice was quite often 32 cores, which is the third most popular choice among users, but it is 72 times less frequent than 1 core.
-This led us to envision how much energy could be saved and encouraged us towards future research using the proposed model for DPM or more advanced optimization algorithms.
-
-In practice, this approach can be implemented by allowing the resource manager to perform these changes for the user using pre-scripts and post-scripts for high energy consumption job submissions.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\columnwidth]{models/figures/dvfs_cmp_32.pdf}
-	\caption{Energy savings comparisons between the proposed model and the Best case.}
-	\label{fig:energy_best_case}
-\end{figure}
-\vspace{-12pt}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\columnwidth]{experiments/figures/cpu_requestes.pdf}
-	\caption{Number of CPU requests during one year in HPC cluster, sorted by the number of cores requested per job.}
-	\label{fig:cpu_requests}
-\end{figure}
-
-%\section{PMCs}
-%
-%\section{Energy per instruction}
-%
-%\begin{lstlisting}
-%xor rcx, rcx
-%mov rax, 1
-%mov rdx, 0
-%
-%loop:
-%	targ_inst(*arg)
-%	targ_inst(*arg)
-%	targ_inst(*arg)
-%	targ_inst(*arg)
-%	targ_inst(*arg)
-%	targ_inst(*arg)
-%	targ_inst(*arg)
-%	targ_inst(*arg)
-%	targ_inst(*arg)
-%
-%add rcx, 1
-%cmp rcx, 9999999
-%jne loop
-%\end{lstlisting}
-%
-%Estimating the energy of the instruction from this benchmark
-%
-%$E_{total}=9999999(\frac{10}{13}inst+\frac{3}{13}loop)$
-%
-%$E_{total}=7692306*inst+constant$
-%
-%\subsection{Generic}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\textwidth]{experiments/figures/inst_en_args_generic.png}
-%	\caption{Energy per instruction argument}
-%	\label{fig:experiment_en1}
-%\end{figure}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\textwidth]{experiments/figures/inst_mean_en_generic.png}
-%	\caption{Mean energy per instruction over all arguments}
-%	\label{fig:experiment_en2}
-%\end{figure}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\textwidth]{experiments/figures/inst_std_en_generic.png}
-%	\caption{Standard deviation energy per instruction over all arguments}
-%	\label{fig:experiment_en3}
-%\end{figure}
-%
-%\subsection{SSE}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\textwidth]{experiments/figures/inst_en_args_sse.png}
-%	\caption{Energy per instruction argument (sse)}
-%	\label{fig:experiment_en4}
-%\end{figure}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\textwidth]{experiments/figures/inst_mean_en_sse.png}
-%	\caption{Mean energy per instruction over all arguments (sse)}
-%	\label{fig:experiment_en5}
-%\end{figure}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\textwidth]{experiments/figures/inst_std_en_sse.png}
-%	\caption{Standard deviation energy per instruction over all arguments (sse)}
-%	\label{fig:experiment_en6}
-%\end{figure}
-%
-%\subsection{SSE and Generic}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\textwidth]{experiments/figures/inst_en_cmp_sse_generic.png}
-%	\caption{Comparing SSE and generic instructions energy consumption}
-%	\label{fig:experiment_en7}
-%\end{figure}
-%
-%\subsection{Power}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\textwidth]{experiments/figures/inst_pw_generic.png}
-%	\caption{Instructions power draw}
-%	\label{fig:experiment_pw1}
-%\end{figure}
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\textwidth]{experiments/figures/inst_pw_args.png}
-%	\caption{Instructions power draw by argument}
-%	\label{fig:experiment_pw2}
-%\end{figure}
-%
-%\section{Tracing application methods}
-%
-%\begin{outline}
-%	\1 Linux ptrace
-%		\2 Breakpoint on each instruction	
-%	\1 Intel pin
-%		\2 JIT
-%		\2 Not open source
-%	\1 qemu user
-%		\2 TCG accelerator (transform target to host instruction) similar to JIT
-%		\2 Partial emulation
-%\end{outline}
-%
-%We choose the less intrusive and open-source alternative qemu user, with an small modificaion we are able to trace an application instructions. Generating an table as follows:
-%
-%
-%\begin{table}[H]
-%	\centering
-%	\begin{tabular}{|c|c|c|}
-%		\hline
-%		opcode         & mnemonic & args                          \\ \hline
-%		48 89 e7       & movq     & \%rsp, \%rdi\textbackslash{}n \\ \hline
-%		48 89 e7       & movq     & \%rsp, \%rdi\textbackslash{}n \\ \hline
-%		e8 08 0e 00 00 & callq    & 0x4000a24ea0\textbackslash{}n \\ \hline
-%		55             & pushq    & \%rbp\textbackslash{}n        \\ \hline
-%		48 89 e5       & movq     & \%rsp, \%rbp\textbackslash{}n \\ \hline
-%		41 57          & pushq    & \%r15\textbackslash{}n        \\ \hline
-%		41 56          & pushq    & \%r14\textbackslash{}n        \\ \hline
-%		41 55          & pushq    & \%r13\textbackslash{}n        \\ \hline
-%		& \vdots & \\ \hline
-%	\end{tabular}
-%\end{table}
-%
-%
-%\section{CPU load}
-%
-%\begin{figure}[H]
-%	\centering
-%	\begin{subfigure}[b]{0.45\textwidth}
-%		\centering
-%		\includegraphics[width=1\columnwidth]{experiments/figures/pw_freq_load.png}
-%		\caption{Instructions power draw varing CPU load}
-%		\label{fig:experiment_pw_load}
-%	\end{subfigure}
-%	%
-%	\begin{subfigure}[b]{0.45\textwidth}
-%		\centering
-%		\includegraphics[width=1\columnwidth]{experiments/figures/pw_load.png}
-%		\caption{Instructions power draw varing CPU load}
-%		\label{fig:experiment_pw_load}
-%	\end{subfigure}
-%\end{figure}
-%
-%
-%\section{Analysis}
-%
-%\subsection{Pareto points}
-
-%
-%\begin{figure}
-
-%	\centering
-
-%	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_high.png}
-
-%	\caption{Pareto static energy}
-
-%	\label{fig:pareto_static_h}
-
-%\end{figure}
-
-%
-%
-%\begin{figure}
-
-%	\centering
-
-%	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_low.png}
-
-%	\caption{Pareto static energy}
-
-%	\label{fig:pareto_static_l}
-
-%\end{figure}
-
-%
-%
-%\begin{figure}
-
-%	\centering
-
-%	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_w_high.png}
-
-%	\caption{Pareto w energy}
-
-%	\label{fig:pareto_w_h}
-
-%\end{figure}
-
-%
-%\begin{figure}
-
-%	\centering
-
-%	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_low.png}
-
-%	\caption{Pareto w energy}
-
-%	\label{fig:pareto_w_l}
-
-%\end{figure}
-
-%
-%\subsection{Optimization under constraints}
-
-%
-%\subsection{Gradient and Countorns}
-
-%
-%\begin{figure}[H]
-
-%	\centering
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn0.png}
-
-%	\end{subfigure}
-
-%	%
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn0_3d.png}
-
-%	\end{subfigure}
-
-%\end{figure}
-
-%
-%\begin{figure}[H]
-
-%	\centering
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn3.png}
-
-%	\end{subfigure}
-
-%	%
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn3_3d.png}
-
-%	\end{subfigure}
-
-%\end{figure}
-
-%
-%\begin{figure}[H]
-
-%	\centering
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/pleak0.png}
-
-%	\end{subfigure}
-
-%	%
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/pleak0_3d.png}
-
-%	\end{subfigure}
-
-%\end{figure}
-
-%
-%\begin{figure}[H]
-
-%	\centering
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/pleak10.png}
-
-%	\end{subfigure}
-
-%	%
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/pleak10_3d.png}
-
-%	\end{subfigure}
-
-%\end{figure}
-
-%
-%\begin{figure}[H]
-
-%	\centering
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic0.png}
-
-%	\end{subfigure}
-
-%	%
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic0_3d.png}
-
-%	\end{subfigure}
-
-%\end{figure}
-
-%
-%\begin{figure}[H]
-
-%	\centering
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic3000.png}
-
-%	\end{subfigure}
-
-%	%
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic3000_3d.png}
-
-%	\end{subfigure}
-
-%\end{figure}
-
-%
-%\begin{figure}[H]
-
-%	\centering
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/w0.png}
-
-%	\end{subfigure}
-
-%	%
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/w0_3d.png}
-
-%	\end{subfigure}
-
-%\end{figure}
-
-%
-%\begin{figure}[H]
-
-%	\centering
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/w1.png}
-
-%	\end{subfigure}
-
-%	%
-
-%	\begin{subfigure}[b]{0.45\textwidth}
-
-%		\includegraphics[width=\textwidth]{models/figures/analisys/w1_3d.png}
-
-%	\end{subfigure}
-
-%\end{figure}

### 2022-04-27 thesis/intro/intro.tex
carlos revision 1

@@
 
-\section{Theoretical Background} \label{sec:theoretical_background}
+\section{Objectives}
 
-A model is a formal representation of a natural system.
-The representation of computer system models includes equations, graphical models, rules, decision trees, representative collections of examples, and neural networks.
-The choice of representation affects the model's accuracy, as well as its interpretability by people~\cite{Hypothesis2012EncyclopediaLearning, Roy2019ForecastingNetwork, Zhu2019PredictingLearning}.
-Accurate energy and power consumption models are essential for many energy efficiency schemes employed in computing equipment \cite{Rivoire2007ModelsOptimizations}, and they can have multiple uses, including the design, forecasting, and optimization of data center systems.
-This study focuses on analytical models that could aid energy optimization and analyses of crucial factors in the total energy draw.
-
-The desirable properties of a full-system model of energy consumption include accuracy, speed, generality and portability, inexpensiveness, and simplicity \cite{Rivoire2008AModels}.
-However, modeling an HPC system's exact energy consumption behavior is not straightforward, either at the whole-system level or at the level of individual components.
-Data centers' energy consumption patterns depend on multiple factors, such as hardware specifications, workload, cooling requirements, or the type of the applications.
-Some of these factors cannot be measured easily.
-Furthermore, it is impractical to perform detailed measurements of the energy consumption of lower-level components without additional overhead.
-
-Several proposed models have already been classified concerning their input parameters, as shown by Dayarathna et al.~\cite{Dayarathna2016DataSurvey}, who analyzed more than 200 models according to their characteristics and limitations and classified them into categories where the model is more suited to its objectives:
-\begin{itemize}
-	%\item Temperature
-	\item System utilization or workload
-	\item Frequency
-	\item Other system states, such as cache miss, branch prediction, number of instructions executed, and more
-	\label{tab:input_type}
-\end{itemize}
-
-Often, energy models are described as a combination of two main parts, the power model of the system and the performance model of the application.
-This is because the concept of energy (E) is the total amount of work performed by a system over a period of time (T), while power (P) is the rate at which the system performs the work.
-The relation between these three amounts can be expressed as:
-\begin{equation}
-	E = \int_{0}^{T}P(t)dt.
-	\label{eq:energy_definition_cont}
-\end{equation}
-
-\subsection{Power Models}
-
-The modeling of system parameters is becoming popular nowadays with the advantage of performance counters provided by the CPU or the operating system.
-These counters can measure micro-architectural events, such as instructions executed, cache hits, miss-predicted branches, and more; thus, providing a base for many different estimations of power usage.
-This makes this type of model very suitable for power estimation because it can use information about several internal states of the computer.
-
-Frequency-based models are the most common kind of model.
-They serve as a base for many power models~\cite{Sarwar1997CmosCalculation, Butzen2007LeakageGates, Usman2013ANoC}.
-These models utilize the fact that every digital circuit (including modern processors) is composed of transistors.
-Thus, modeling one transistor's interaction and scaling this to the chip can give a reasonable estimate of the entire system's energy.
-One of the most common frequency-based %please confirm that the intended meaning has been retained.
-model approximations is defined as follows: 
-\begin{equation}
-	P = \alpha+\beta f^3,
-	\label{eq:power_simplified}
-\end{equation}
-where $\alpha$ and $\beta$ are model parameters, and $f$ is the operating frequency (details of this equation are covered in (\cref{chapter:models}).
-This type of model is suitable for optimization problems since these are a function of the operating frequency, which can be easily controlled.
-
-% We confirm that the meaning has been retained
-
-\subsection{Performance Models}
-The most common way to model the application performance is using the workload.
-The workload is an abstract representation of the amount of work done for a given time and speed.
-The workload ($W$) can be defined in many different ways.
-One common way, used in many studies, such as Paolillo et al.
-~\cite{Paolillo2018OptimisationParallelism}, Francis et al.
-\cite{ Group2012HandbookSahni}, and Kim et al.
-\cite{Kim2015RacingHeuristics}, is the following:
-\begin{equation}
-	W = \int_{0}^{\tau}s(t)dt = s\tau,
-	\label{eq:workload_definition}
-\end{equation}
-where $\tau$ is total active time, and $s$ is the execution speed in instructions/second.
-
-Utilization models \cite{Fu2018RaceMinimization, Group2012HandbookSahni} are also found in the literature, defined as the ratio between the time that the system is active and the total time (idle and active).
-These models are present in many DVFS algorithms present in Linux.
-They can be viewed as a good alternative to the workload since it is impossible to measure workload in real-time.
-\cref{eq:utilization_definition} defines workload in terms of CPU utilization ($u$):
-\begin{equation}
-	u = \frac{\tau}{T} = \frac{W/s}{T},
-	\label{eq:utilization_definition}
-\end{equation}
-where $T$ is the total execution time (idle and active), and $\tau$ is the active time, meaning when the processor was executing instructions.
-Models based on CPU utilization are the basis for DVFS algorithms.
-Even though this is not a controllable parameter, it is straightforward to measure system utilization with almost no overhead, and it is also very portable in terms of operating systems and architectures.
-
-\section{Related work} \label{sec:related_work}
-
-Merkel et al.~\cite{Merkel2006BalancingSystems} developed an energy model for processors based on events.
-Their model assumes a fixed energy consumption $\alpha_i$ for each activity, and by counting the number of occurrences $c_i$ of every activity, they estimate the total energy as:
-\begin{equation}
-	E = \sum_{i=1}^{n}\alpha_ic_i.
-	\label{equation:rw_merkel}
-\end{equation}
-
-Another event-based model, introduced by Roy et al.
-\cite{Roy2013AnAlgorithms}, described the computational energy consumed by a CPU for an algorithm $A$ as the Equation (\ref{equation:rw_event_based})%\cref{equation:rw_event_based}:
-\begin{equation}
-	E(A) = P_{clk}T(A) + P_wW(A),
-	\label{equation:rw_event_based}
-\end{equation}
-where $P_{clk}$ is a processor clock leakage power, $T(A)$ is the total execution time, $W(A)$ is the total time taken by  non-I/O operations, and $P_w$ is used to capture the power consumption per operation performed by the CPU.
-$T(A)$ and $W(A)$ are estimated using performance features.
-
-Models based on events present some drawbacks, they are highly dependent on the operating system and its architecture, making them problematic to port for other platforms.
-There are also limitations regarding the number of simultaneous events that can coexist without adding a non-negligible overhead.
-Additionally, there are cases where events need multiplexing, for example, when using more hardware events than %please confirm that the intended meaning has been retained.
-the CPU can provide.
-There are also some well know problems regarding the precision of some events, as shown in many studies~\cite{Weaver2008CanTrusted, Weaver2013Non-determinismImplementations, Das2019SoK:Security, McGuire2009AnalysisKernel, Ramos2019AnCounters, Silva-de-Souza2020ContainergyAWorkloads}.
-Some events that should be exact and deterministic (such as the number of executed instructions) show run-to-run variations and over-count on various architectures, even when running in strictly controlled environments.
-Because of that, our proposed model is not dependent on events and, therefore, not vulnerable to those drawbacks.
-% We confirm that the meaning is the same
-
-An instruction-level energy model was also proposed in \cite{Shao2013EnergyProcessor} by Yakun et.
-al.
-Where they proposed an energy per instruction (EPI) characterization made on Xeon Phi.
-Their model is expressed as:
-\begin{equation}
-	E(f) = \frac{(p_1 - p_0)(c_1 - c_0)/f}{N}, 
-	\label{eq:rw_instruction_level}
-\end{equation}
-where $N$ is the total number of dynamic instructions, $p_0$ is the initial idle power, $p_1$ is the average dynamic power, and ($c_1$ $-$ $c_0$) refers to the cumulative number of cycles the micro-benchmark performs.
-This model is suitable for estimating the energy after the application finishes executing when it is possible to count the total cycles.
-However, it is challenging to use for optimization or forecasting since it does not have an application model to predict the cycles.
-Our model integrates the behavior of the application, taking into account the execution time.
-
-Lewis et al.~\cite{Lewis2008Run-timeSystems} described the overall system energy consumption using the following equation:
-\begin{equation}
-	E = A_0(E_{proc} + E_{mem}) + A_1E_{em} + A_2E_{board} + A_3E_{hdd},
-	\label{eq:lewvis}
-\end{equation}
-where, $A_0$, $A_1$, $A_2$, and $A_3$ are unknown constants that are calculated via linear regression analysis and those remain constant for a specific server architecture.
-This model, as the previous one, relies on knowledge of energy spent on each component, being  a suitable option for estimation after the application has already run, but not for optimization of the run itself, which is the aim of our model.
-
-In another energy consumption model  based on system utilization, Mills et al.
-\cite{Mills2014EnergySystems} modeled the energy consumed by a compute node with CPU (single) executing at speed $\sigma$ as Equation (\ref{eq:mills})%\cref{eq:mills},
-\begin{equation}
-	E(\sigma,[t_1,t_2]) = \int_{t_1}^{t_2} \sigma^3 + \rho \sigma_max^3 dt,
-	\label{eq:mills}
-\end{equation}
-where $\rho$ stands for the overhead power consumed regardless of the processor speed, $t_1$ and $t_2$ are the application's initial and final execution times.
-The overhead includes power consumption by all other system components, such as memory, network, and more.
-For this reason, although the authors mentioned the energy consumption of a socket, their power model is generalized to the entire server.
-This model lacks a closed-form, i.e., it depends on the definition of $\alpha(t)$ to be complete.
-Our model has a closed-form which facilitates analyses.
-
-\subsection{The contribution of this work}
+\section{Contributions}
 
@@
 
-%The main contributions of the proposed model are:
-%
-%\begin{itemize}
-%	\item Simple model: faster to fit and compute, good for DVFS and DPM optimization.
-%	\item Parameters with logical meaning: helps to understand the contribution of each specific term.
-%	\item Analytical analysis: several analyses can be derived from the equation.
-%	\item Controllable variables: the equation is in the function of parameters that we can control directly.
-%\end{itemize}
+The main contributions of the proposed model are:
+
+% Contribuições por capitulo...
+\begin{itemize}
+	\item Simple model: faster to fit and compute, good for DVFS and DPM optimization.
+	\item Parameters with logical meaning: helps to understand the contribution of each specific term.
+	\item Analytical analysis: several analyses can be derived from the equation.
+	\item Controllable variables: the equation is in the function of parameters that we can control directly.
+\end{itemize}
+
+\section{Organization}

### 2022-04-27 thesis/optmization.tex
carlos revision 1

\section{Validation the model} \label{sec:experimentalvalidation}
In this section, the models presented in~Sections \ref{sec:power_model} and~\ref{sec:performance_model} were validated with a benchmark specific for multi-core architectures. Additionally, in order to assess the modeling overhead and accuracy, our proposal was then compared to machine learning approaches. We compared against support vector regression (SVR)~\cite{Smola2004ARegression}, decision tree \cite{Kitts2006RegressionLecture}, k-nearest neighbors \cite{Altman1992AnRegression}, multilayer perceptron \cite{Murtagh1991MultilayerRegression}, and some new methods, such as Gao et al. \cite{Gao2019DendriticPrediction}. However, SVR was chosen as the most representative because it performed best in our tests without aggressive fine-tuning, as shown in \cref{fig:ml_models}.

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{experiments/figures/ml_models.pdf}
	\caption{Average of the mean squared error for all applications of our study case Section \ref{sec:casestudyapplication}.}
	\label{fig:ml_models}
\end{figure}

\subsection{Verifying Hypothesis}

In this section, we validate whether the assumptions of our model are valid for the system used.

\subsection{Frequency and Voltage Relation}
One of the assumptions was that the frequency and the voltage have a linear relationship, as indicated by ~\cref{eq:f_v}. To verify that, we build an experiment that sets the frequency to a specific value while sampling the voltage using the APERF and MPERF registers that provide feedback on the current CPU frequency. The average result of the sampling voltages is shown in  \cref{fig:freq_volt_rel}, where we can observe a near-perfect linear relation. This is because manufacturers implement this curve in the processors, using tables that relate ranges of frequencies to voltages so that they can precisely define any curve that will better suit their design.

\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}
	\includegraphics[width=\columnwidth]{experiments/figures/freq_volt_rel.pdf}
	\caption{Frequency voltage relation.}
	\label{fig:freq_volt_rel}
\end{figure}

\subsection{Input Size and Instructions}
We ran the applications with different inputs assuming  linear growth in the amount of work for one input to the other when building our model. However, measuring and controlling the amount of work would require much instrumentation and tuning to find an input corresponding to a certain amount of work. Therefore, to build our models, we use the time to reference the amount of work, assuming that the work is proportional to the executing time. \cref{fig:time_input} corresponds to the verification of this supposition.

\cref{tab:corr_in_time} shows that the assumption was reasonable since the average correlation was 0.96 for all applications, indicating that growth in the number of instructions  will follow the time. This was the case for all applications that we ran in our benchmark and should hold for any data parallelism type of application.
\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/blackscholes.pdf}
		\caption{Blackscholes.}
		\label{fig:time_input_blackshoels}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/input_time/canneal.pdf}
		\caption{Canneal.}
		\label{fig:time_input_canneal}
	\end{subfigure}

	\hfill
	\caption{Relation between time and instructions for each input size. }
	\label{fig:time_input}
\end{figure}
\vspace{-6pt}

\begin{table}[H]
	\centering
	\caption{Correlation of time and instructions for all applications.}
	\begin{tabular}{ll}
		\multicolumn{1}{l|}{\textbf{Application}} & \textbf{Correlation} \\ \hline
		Blackscholes                                               & 0.99                                  \\
		Bodytrack                                                  & 0.99                                  \\
		Canneal                                                    & 0.99                                  \\
		Dedup                                                      & 0.99                                  \\
		Ferret                                                     & 0.96                                  \\
		Fluidanimate                                               & 0.99                                  \\
		Freqmineq                                                  & 0.99                                  \\
		Openmc                                                     & 0.94                                  \\
		Raytrace                                                   & 0.99                                  \\
		Swaptions                                                  & 0.99                                  \\
		Vips                                                       & 0.98                                  \\
		x264                                                       & 0.99                                  \\
		HPL                                                        & 0.79                                  \\ \hline
	\end{tabular}
	\label{tab:corr_in_time}
\end{table}

The next assumption was that the application's behavior was the same when varying the workload. This condition is necessary for using the model with an unknown input size because, if the behavior is the same, we can interpolate the known inputs. One way to verify this is to measure the rate of instructions per second normalized by the frequency, as shown in \cref{fig:fingerpritns}.
\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}

	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/blackscholes.pdf}
		\caption{Blacksholes.}
		\label{fig:fp_baclscholes}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\columnwidth]{models/figures/hypothesis/input_instructions/fp/canneal.pdf}
		\caption{Canneal.}
		\label{fig:fp_canneal}
	\end{subfigure}

	\hfill
	\caption{Rate of instructions per second varying the input size normalized by the frequency.}
	\label{fig:fingerpritns}
\end{figure}

\cref{fig:fingerpritns} shows that the applications have roughly the same curve when normalized; this also happens for all other applications in our benchmark.

The final assumption is that the workload should also not vary depending on the number of cores or frequency. To verify, we measure the total number of executed instructions while varying the cores from 1 to 32. \cref{tab:cores_variation} shows the results.

\begin{table}[H]
	\centering
	\caption{Variation of the number of instructions when changing the number of cores for the same input.}
	%\begin{tabular}{c|c|c|c}
	\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}cccc}
		\toprule
		& \textbf{Average} & \textbf{Standard}     & \textbf{Standard}\\
		\multirow{-2}{*}{\textbf{Application}}& \textbf{Number of Instructions} & \textbf{Deviation}     & \textbf{Deviation} \textbf{(\%) }\\ \midrule
		Vip          & $7.97 \times 10^{11}$     & $7.16 \times 10^{6}$  & 0.00 \\ %Please use Scientific notation. For example, “1.65e+04” should be changed to “1.65× 104”. Please confirm `%` whether to remove.
		Openmc       & $8.17 \times 10^{7}$     & $ 1.65 \times 10^{4}$ & 0.02    \\ 
		Rtview       & $9.91 \times 10^{12}$     & $ 1.55 \times 10^{9}$ & 0.02    \\ 
		X264         & $4.52 \times 10^{11}$     & $ 5.81 \times 10^{7}$ & 0.01    \\ 
		Bodytrack    & $1.86 \times 10^{12}$     & $ 3.95 \times 10^{10}$ & 2.13    \\ 
		Fluidanimate & $2.09 \times 10^{12}$     & $ 8.44 \times 10^{10}$ & 4.04    \\ 
		HPL          & $1.14 \times 10^{8}$      & $ 1.24 \times 10^{5}$ & 0.11    \\ 
		Blackschole  & $3.75 \times 10^{12}$     & $ 1.40 \times 10^{9}$ & 0.04     \\ 
		Dedup        & $1.02 \times 10^{11}$     & $ 5.74 \times 10^{7}$ & 0.06     \\ 
		Swapti       & $2.43 \times 10^{12}$     & $ 8.87 \times 10^{8}$ & 0.04     \\ 
		Canneal      & $1.19 \times 10^{11}$     & $ 4.46 \times 10^{7}$ & 0.04     \\ 
		Freqmine     & $1.27 \times 10^{12}$     & $ 4.78 \times 10^{8}$ & 0.04     \\ 
		Ferret       & $4.76 \times 10^{11}$     & $ 7.04 \times 10^{7}$ & 0.01     \\ \bottomrule
	\end{tabular*}
	\label{tab:cores_variation}
\end{table}
% Corrected

\cref{tab:cores_variation} shows the standard deviation and what that corresponds to in terms of the total number of instructions as a percentage.

The same test was performed for the frequency, varying from 1.2 to 2.2 GHz with 100~MHz steps. The results are shown in \cref{tab:freq_variation}.

These results show that all the assumptions were reasonable, and we can safely move to the validation of the model's prediction.
\begin{table}[H]
	\caption{Variation of the number of instructions when changing the frequency for the same input.}
	\centering
	%\begin{tabular}{c|c|c|c}{\hsize}{@{\extracolsep{\fill}}cccc}
	\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}cccc}
		\toprule
		& \textbf{Average} & \textbf{Standard}     & \textbf{Standard}\\
		\multirow{-2}{*}{\textbf{Application}}& \textbf{Number of Instructions} & \textbf{Deviation}     & \textbf{Deviation} \textbf{(\%)} \\ \midrule
		Vip          & $7.97 \times 10^{11}$      & $ 1.16 \times 10^{6}$  & 0.00      \\ 
		Openmc       & $8.17 \times 10^{7}$       & $ 4.52 \times 10^{3}$  & 0.01      \\ 
		Rtview       & $9.91 \times 10^{12}$      & $ 6.64 \times 10^{5}$  & 0.00      \\ 
		X264         & $4.52 \times 10^{11}$      & $ 1.54 \times 10^{5}$  & 0.00      \\ 
		Bodytrack    & $1.84 \times 10^{12}$      & $ 2.54 \times 10^{5}$  & 0.00      \\ 
		Fluidanimate & $2.38 \times 10^{12}$      & $ 1.70 \times 10^{9}$  & 0.07      \\ 
		HPL          & $1.14 \times 10^{8}$       & $ 5.95 \times 10^{3}$  & 0.01      \\ 
		Blackschole  & $3.75 \times 10^{12}$      & $ 4.36 \times 10^{5}$  & 0.00      \\ 
		Dedup        & $1.02 \times 10^{11}$      & $ 8.32 \times 10^{7}$  & 0.08      \\ 
		Swapti       & $2.43 \times 10^{12}$      & $ 1.48 \times 10^{5}$  & 0.00      \\ 
		Canneal      & $1.19 \times 10^{11}$      & $ 3.01 \times 10^{5}$  & 0.00      \\ 
		Freqmine     & $1.27 \times 10^{12}$      & $ 3.70 \times 10^{8}$  & 0.03      \\ 
		Ferret       & $4.76 \times 10^{11}$      & $ 5.63 \times 10^{7}$  & 0.01      \\\bottomrule
	\end{tabular*}
	\label{tab:freq_variation}
\end{table}

\section{Fitting the Models} \label{subsec:fitting}
To find the parameters of  \cref{eq:en_final}, 10 uniformly random configurations of frequencies ($f$), cores ($p$) and inputs ($N$) were chosen from the range $1<=p<=32$, $1.2<=f<=2.2$ and $1<=N<=5$, respectively. The application was executed for each chosen configuration, and the measured energy and time values were collected. For the input size, if we assume that all CPU instructions take approximately the same time to execute, the number of basic operations will be directly correlated with the time. Thus, we can estimate the input size by looking at the execution time, allowing us to divide a large input size into several smaller ones, knowing their relationship, as performed in the work of Oliveira ~\cite{Oliveira2018ApplicationCharacterization}. The unity can also vary depending on the definition. For simplicity, we assign numbers from 1 to 10, increasing the problem linearly, so it is also possible to interpolate any input in between these values.

For each configuration, samples of the power were collected using IPMI every 1 second. This sampling rate was chosen based on the magnitude of the mean run time of the applications, which is in the order of minutes. Therefore, this rate provides enough samples to measure average power. Additionally, timestamps and the total run time were collected. The total energy spent on each configuration is estimated by first interpolating the power samples using the first-order method and then integrating this function in the time.

The model's parameters are calculated by solving an optimization problem of finding the values that minimize the squared error of the prediction to the measured values using the non-linear least-squares method.

The Python library Scikit-Learn was used to build the SVR model~\cite{Pedregosa2011Scikit-learn:Python}. The SVR was trained using the same data used for parameter estimation of Equation (\ref{eq:en_final}) %\cref{eq:en_final}
with a grid search used to find the best kernel function and the best values for the hyper-parameters penalty for the wrong ($C$) and ($\gamma$). For this data, the best function was the radial base function (RBF), and the hyper-parameters were $C=10^4$ and $\gamma=0.5$.

\section{Measured versus Modeled Energy}
\label{subsec:measuredversusmodeledenergy}

To validate the model, we ran all possible configurations in the tested machine, varying the cores in a range of $1<=p<=32$, the frequency in $1.2<=f<=2.2$, and the input in $1<=N<=5$. The total number of configurations varies from 400 to over 1000 depending on the application, as some applications have restrictions on the number of cores that they can run. Once the data was collected, we computed the mean percentage error (MPE) according to the following equation:
\begin{equation}
	MPE = \frac{1}{N} \sum_i^N \frac{|y_{\rm estimated}-y_{\rm measured}|}{y_{\rm measured}}.
	\label{eq:mpe}
\end{equation}

\subsection{Frequency $\times$ Cores}
\cref{fig:en_eq_freq_cores_fc} plots the measured and modeled energy consumption for some of the applications modeled. In addition,  some of the possible shapes that the model can take while varying the number of active cores, and operating frequency, are shown.%please confirm that the intended meaning has been retained.
% yes its the same

\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_cores/completo_black_5.pdf}}
		\caption{}
		\label{fig:en_eq_black_fc}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_cores/completo_canneal_1.pdf}}
		\caption{}
		\label{fig:en_eq_canneal_fc}
	\end{subfigure}

	\caption{Example fit for a specific input size: Blackscholes (\textbf{a}) and Canneal (\textbf{b}).  “measured values” are the sensor data, and “minimum energy” is the minimum energy model prediction.
	}
	\label{fig:en_eq_freq_cores_fc}
\end{figure}
\subsection{Frequency $\times$ Input}
\cref{fig:en_eq_freq_inp_fi} plots the measured and modeled energy consumption for some of the applications modeled. The diagrams show some of the possible shapes that the model can take while varying the operating frequency, and input size.
\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_inps/completo_black_5.pdf}}
		\caption{}
		\label{fig:en_eq_black_fi}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/freq_inps/completo_canneal_1.pdf}}
		\caption{}
		\label{fig:en_eq_canneal_fi}
	\end{subfigure}

	\caption{Example fit for a specific input size: Blackscholes (\textbf{a}) and Canneal (\textbf{b}).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
	}
	\label{fig:en_eq_freq_inp_fi}
\end{figure}
\subsection{Cores $\times$ Input}
\cref{fig:en_eq_core_inp_ci} plots the measured and modeled energy consumption for some of the applications modeled. The diagrams  show some of the possible shapes that the model can take while varying the number of active cores, and input size.
\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/cores_inps/completo_black_5.pdf}}
		\caption{}
		\label{fig:en_eq_black_ci}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/energy/cores_inps/completo_canneal_1.pdf}}
		\caption{}
		\label{fig:en_eq_canneal_ci}
	\end{subfigure}

	\caption{Example fit for a specific input size: Blackscholes (\textbf{a}) and Canneal (\textbf{b}).  “measured values” are the sensor data and “minimum energy” is the minimum energy model prediction.
	}
	\label{fig:en_eq_core_inp_ci}
\end{figure}

\subsection{Comparison}
The average results for each application were calculated using a model trained with only 10 configurations, and the comparison is displayed \cref{fig:mpe_svr_eq}. %We calculated the raw MPE values shown in ~\cref{tab:mpe_svr_eq}.
\begin{figure}[H]
	\centering
	\includegraphics[width=.8\columnwidth]{models/figures/mpe_svr_eq.pdf}
	\caption{Comparison of the mean percentage error between the proposed model and SVR. ``Model mean'' and ``SVR mean'' are the average of all MPE values for all applications.
	}
	\label{fig:mpe_svr_eq}
\end{figure}

\cref{fig:mpe_svr_eq} shows that the proposed model always performed better, with a lower MPE than SVR, when we were limited to 10 training points. This result is further explored in the next Section \ref{subsec:overhead}, where we undertake a comparison with different training sizes. The exact values are shown in \cref{tab:mpe_svr_eq}.

\begin{table}[H]
\centering
\begin{tabular}{c|c|c}
\hline
Application  & Model & SVR   \\ \hline
Ferret       & 5.25     & 12.49  \\ 
Raytrace     & 6.36     & 11.95  \\
Fluianimate  & 2.44     & 22.90  \\
x264         & 8.28     & 15.33  \\
Vips         & 7.54     & 10.80  \\
Swaptions    & 6.54     & 18.57  \\
Canneal      & 3.12     & 6.13   \\
Dedup        & 8.85     & 13.70  \\
Freqmine     & 2.44     & 3.24   \\
Blackscholes & 2.18     & 11.00  \\
HPL          & 7.47     & 12.75  \\
Bodytrack    & 16.98    & 34.12  \\
Openmc       & 11.15    & 24.34  \\
\end{tabular}
\caption{Comparison of the Mean Percentage Error between the proposed model and SVR: raw values.}
\label{tab:mpe_svr_eq}
\end{table}

\section{Overheads on Training} \label{subsec:overhead}
It is known that machine learning is data-driven; in that sense, the SVR model obtained using only 10 configurations could be improved, but what about the analytical model? 
To answer that question, the proposed model and the SVR were also trained with a varying number of configurations.
We then compared the MPE and the amount of energy spent to create each model. 
This accuracy-energy trade-off is crucial since building models' energy overhead defeats the primary goal of saving power when running applications.
\enlargethispage{.5cm}

\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_ferret_4.pdf}}
		\caption{MPE for Ferret.}
		\label{fig:overhead_ferret}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_black_6.pdf}}
		\caption{MPE for Blackscholes.}
		\label{fig:overhead_black}
	\end{subfigure}
	%
	\par\bigskip
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_x264_4.pdf}}
		\caption{MPE for x264.}
		\label{fig:overhead_x264}
	\end{subfigure}
	%
	\begin{subfigure}[b]{0.45\textwidth}
		\centerline{\includegraphics[width=\columnwidth]{models/figures/overhead/completo_dedup_4.pdf}}
		\caption{MPE for Dedup.}
		\label{fig:overhead_dedup}
	\end{subfigure}

	\caption{MPE of the case studies versus training size, comparing how many training points is necessary to reach an acceptable result.}
	\label{fig:overheadapps}
\end{figure}

\cref{fig:overheadapps} shows the comparisons of MPE and energy spent to create each model for two selected applications. According to the results, the analytical model is very stable, not changing much as more data is added, while the SVR keeps reshaping to adapt to the data. 
The error of the analytical model is almost constant but that of the SVR, initially very high, drops as more data is used in the training process.
\begin{figure}[H]
	\centering
	\captionsetup[subfigure]{justification=centering}
	\begin{subfigure}[t]{0.45\textwidth}
		\includegraphics[width=\columnwidth]{models/figures/overhead/overall_energy_10pts.pdf}
		\caption{}
		\label{fig:overall_overhead}
	\end{subfigure}
	%
	%\par\bigskip
	\begin{subfigure}[t]{0.45\textwidth}
		\includegraphics[width=\columnwidth]{models/figures/overhead/overall_mpe_10pts.pdf}
		\caption{
		}
		\label{fig:overall_MPE}
	\end{subfigure}
	\caption{Overall results for energy and MPE for each training size. (\textbf{a}) Average energy spent on all applications during model creation. The two curves are identical because the same data were used to adjust the SVR and the model. (\textbf{b}) MPE of all applications: SVR needs 10 times more data to have an MPE lower than the proposed model.}
	\label{fig:overall_train}
\end{figure}

\cref{fig:overall_train} presents the overall results, with the mean energy overhead and MPE for all applications.
The meeting point of the MPE for the SVR and the proposed model can be extracted from \cref{fig:overall_train}b.
It shows that, in around 90 configurations, the SVR starts to have a smaller error. The cost of that is the linear increase in energy spent on training. The increase in energy, about 10 times more, can be observed in \cref{fig:overall_train}a.

\section{Analysis}
One of the most significant advantages of using an analytical model is the understanding of the problem that an equation provides, making many different kinds of analysis possible that are otherwise impossible with a machine learning model. In this section, we discuss one of the possible analyses. In the following figures, we try to understand the contribution of each parameter of the equation to the total energy consumption.

For this analysis, we took the model of one of the applications and, varying one parameter of the equation, we display the energy versus performance (time) for all configurations. After that, we computed the Pareto frontier, a set of all Pareto efficient allocations, i.e., all the configurations where resources cannot be reallocated to make one individual better off without making at least one individual worse off. This gives us all the configurations where we have an optimal trade-off of performance and energy to choose from.

\cref{fig:pareto_static} shows the Pareto frontier for several values for the static power parameter ($c_3$ in \cref{eq:en_final}) with configurations of frequency ranging from 1.2 to 5 GHz and cores from 1 to 64, so that we can also have an idea of what is the tendency when we increase the frequency and number of cores.

\begin{figure}[H]
	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_high.pdf}
	\caption{Pareto frontier for several values of static power parameter. The arrows with blue heads indicate the maximum energy, while the arrows with a red head the minimal energy for each corresponding curve.}% In four-digit numbers, no comma or spaces to indicate thousand, whatever the numbers are in the table or body text, e.g., 4500; More than five-digit numbers, should have comma, e.g. 59,400; 899,302. Please revise the figure.
	\label{fig:pareto_static}
\end{figure}
%Fixed

From this figure, we can see that when increasing the value of the static power parameter, the total energy consumption increases as expected. We can also observe that the values that minimize the total energy consumption tend to be high frequency and multiple cores. This is one of the consequences of increasing the static power factor. As the dynamic factor proportionally decreases, its variables tend to have less impact on total consumption, enabling configurations with high frequency and several cores. This also enables chip-level optimization for choosing components that change the ratio between static and dynamic power.

\cref{fig:pareto_w} shows the Pareto frontier in the same ranges described before but for the parameter corresponding to the level of parallelism of the application ($w$ in  \cref{eq:en_final}).

\begin{figure}[H]
	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_w_high.pdf}
	\caption{Pareto frontier for several values of static power parameter. The arrows with blue heads indicate the maximum energy, while the arrows with a red head, the minimal, for each corresponding curve.}% In four-digit numbers, no comma or spaces to indicate thousand, whatever the numbers are in the table or body text, e.g., 4500; More than five-digit numbers, should have comma, e.g. 59,400; 899,302. Please revise the figure.
	\label{fig:pareto_w}
\end{figure}
%Fixed

In  \cref{fig:pareto_w}, we observe that, as the parallelism level increases the total energy decreases. The number of cores tends to be higher with a higher level of parallelism as expected, and the frequency shows an inverse relation.

\section{DVFS and DPM Optimization} \label{subsec:dvfs_optmin}
The effectiveness of the proposed approach during optimization was evaluated with a simple algorithm that finds the optimal frequency and number of active cores from the proposed equation. The results were then compared to the Linux default choices for power management.

With \cref{eq:en_final}, it is possible to calculate energy consumption estimates for each possible configuration since there is a finite range of possible values for the frequency and number of cores. It is also possible to apply constraints on the execution time, frequency, and the number of active cores. Then, the configuration that minimizes energy consumption for a given input can be selected. The complete workflow is shown in \cref{fig:optim_workflow}. We can see that any optimization problem can be structured with our model and the system's constraints. In the following examples, the optimization problem that we build is to minimize the energy equation given the constraints of possible frequencies and the number of cores that our system can run. The algorithm selected to minimize was the newton-CG~\cite{Royer2020AOptimization}.

Current HPC managers leave to the user the choice of how many cores to use. On this basis, three situations were analyzed in relation to the number of cores:

\begin{enumerate}
	\item Worst choice: number of cores that maximize the total energy consumed;
	\item Random choice: energy consumed for a random choice of the number of cores;
	\item Best choice: number of cores that minimize the total energy consumed (oracle).
\end{enumerate}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\columnwidth]{models/figures/DVFS optim.pdf}
	\caption{Optimization workflow showing how DVFS and DPM optimization could be implemented from ou model.}
	\label{fig:optim_workflow}
\end{figure}

The default option for the Linux governor is Ondemand, and, by default, it has no DPM control for the number of active cores. As Ondemand only performs DVFS, \textls[-15]{for comparison, each application was executed with all available cores in the system, from 1 to 32.}

\textls[-15]{Figures \ref{fig:energy_worst_case}--\ref{fig:energy_best_case}, show the energy savings with respect to Ondemand, i.e., $\frac{Ondemand-Model_{min}}{Ondemand}$} for the three cases described above. The savings and losses for each case are:

\begin{enumerate}
	\item Worst choice: save 69.88\% on average;
	\item Random choice: save 12.04\% on average;
	\item Best choice: lost 14.06\% on average.
\end{enumerate}

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{models/figures/dvfs_cmp_max.pdf}
	\caption{Energy savings comparisons between the proposed model and the Worst case.}
	\label{fig:energy_worst_case}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{models/figures/dvfs_cmp_mean.pdf}
	\caption{Energy savings comparisons between the proposed model and the Random case.}
	\label{fig:energy_mean_case}
\end{figure}

By default, operating systems do not implement DPM at the core level, and, in HPC, the user usually explicitly chooses the number of cores to run their job. To give a better idea of the impact on the energy consumption of DPM at the core level, we analyzed the choices of the number of cores over a period of one year in the HPC center at UFRN. The result is plotted in~\cref{fig:cpu_requests}.

It is of note that the most common choice of many regular users is a single core requested per job, matching the worst-case choice for all applications analyzed in this investigation. The best choice was quite often 32 cores, which is the third most popular choice among users, but it is 72 times less frequent than 1 core. This led us to envision how much energy could be saved and encouraged us towards future research using the proposed model for DPM or more advanced optimization algorithms.

In practice, this approach can be implemented by allowing the resource manager to perform these changes for the user using pre-scripts and post-scripts for high energy consumption job submissions.
\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{models/figures/dvfs_cmp_32.pdf}
	\caption{Energy savings comparisons between the proposed model and the Best case.}
	\label{fig:energy_best_case}
\end{figure}
\vspace{-12pt}

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{experiments/figures/cpu_requestes.pdf}
	\caption{Number of CPU requests during one year in HPC cluster, sorted by the number of cores requested per job.}
	\label{fig:cpu_requests}
\end{figure}

%\section{PMCs}
%
%\section{Energy per instruction}
%
%\begin{lstlisting}
%xor rcx, rcx
%mov rax, 1
%mov rdx, 0
%
%loop:
%	targ_inst(*arg)
%	targ_inst(*arg)
%	targ_inst(*arg)
%	targ_inst(*arg)
%	targ_inst(*arg)
%	targ_inst(*arg)
%	targ_inst(*arg)
%	targ_inst(*arg)
%	targ_inst(*arg)
%
%add rcx, 1
%cmp rcx, 9999999
%jne loop
%\end{lstlisting}
%
%Estimating the energy of the instruction from this benchmark
%
%$E_{total}=9999999(\frac{10}{13}inst+\frac{3}{13}loop)$
%
%$E_{total}=7692306*inst+constant$
%
%\subsection{Generic}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=\textwidth]{experiments/figures/inst_en_args_generic.png}
%	\caption{Energy per instruction argument}
%	\label{fig:experiment_en1}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=\textwidth]{experiments/figures/inst_mean_en_generic.png}
%	\caption{Mean energy per instruction over all arguments}
%	\label{fig:experiment_en2}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=\textwidth]{experiments/figures/inst_std_en_generic.png}
%	\caption{Standard deviation energy per instruction over all arguments}
%	\label{fig:experiment_en3}
%\end{figure}
%
%\subsection{SSE}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=\textwidth]{experiments/figures/inst_en_args_sse.png}
%	\caption{Energy per instruction argument (sse)}
%	\label{fig:experiment_en4}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=\textwidth]{experiments/figures/inst_mean_en_sse.png}
%	\caption{Mean energy per instruction over all arguments (sse)}
%	\label{fig:experiment_en5}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=\textwidth]{experiments/figures/inst_std_en_sse.png}
%	\caption{Standard deviation energy per instruction over all arguments (sse)}
%	\label{fig:experiment_en6}
%\end{figure}
%
%\subsection{SSE and Generic}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=\textwidth]{experiments/figures/inst_en_cmp_sse_generic.png}
%	\caption{Comparing SSE and generic instructions energy consumption}
%	\label{fig:experiment_en7}
%\end{figure}
%
%\subsection{Power}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=\textwidth]{experiments/figures/inst_pw_generic.png}
%	\caption{Instructions power draw}
%	\label{fig:experiment_pw1}
%\end{figure}
%\begin{figure}
%	\centering
%	\includegraphics[width=\textwidth]{experiments/figures/inst_pw_args.png}
%	\caption{Instructions power draw by argument}
%	\label{fig:experiment_pw2}
%\end{figure}
%
%\section{Tracing application methods}
%
%\begin{outline}
%	\1 Linux ptrace
%		\2 Breakpoint on each instruction	
%	\1 Intel pin
%		\2 JIT
%		\2 Not open source
%	\1 qemu user
%		\2 TCG accelerator (transform target to host instruction) similar to JIT
%		\2 Partial emulation
%\end{outline}
%
%We choose the less intrusive and open-source alternative qemu user, with an small modificaion we are able to trace an application instructions. Generating an table as follows:
%
%
%\begin{table}[H]
%	\centering
%	\begin{tabular}{|c|c|c|}
%		\hline
%		opcode         & mnemonic & args                          \\ \hline
%		48 89 e7       & movq     & \%rsp, \%rdi\textbackslash{}n \\ \hline
%		48 89 e7       & movq     & \%rsp, \%rdi\textbackslash{}n \\ \hline
%		e8 08 0e 00 00 & callq    & 0x4000a24ea0\textbackslash{}n \\ \hline
%		55             & pushq    & \%rbp\textbackslash{}n        \\ \hline
%		48 89 e5       & movq     & \%rsp, \%rbp\textbackslash{}n \\ \hline
%		41 57          & pushq    & \%r15\textbackslash{}n        \\ \hline
%		41 56          & pushq    & \%r14\textbackslash{}n        \\ \hline
%		41 55          & pushq    & \%r13\textbackslash{}n        \\ \hline
%		& \vdots & \\ \hline
%	\end{tabular}
%\end{table}
%
%
%\section{CPU load}
%
%\begin{figure}[H]
%	\centering
%	\begin{subfigure}[b]{0.45\textwidth}
%		\centering
%		\includegraphics[width=1\columnwidth]{experiments/figures/pw_freq_load.png}
%		\caption{Instructions power draw varing CPU load}
%		\label{fig:experiment_pw_load}
%	\end{subfigure}
%	%
%	\begin{subfigure}[b]{0.45\textwidth}
%		\centering
%		\includegraphics[width=1\columnwidth]{experiments/figures/pw_load.png}
%		\caption{Instructions power draw varing CPU load}
%		\label{fig:experiment_pw_load}
%	\end{subfigure}
%\end{figure}
%
%
%\section{Analysis}
%
%\subsection{Pareto points}

%
%\begin{figure}

%	\centering

%	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_high.png}

%	\caption{Pareto static energy}

%	\label{fig:pareto_static_h}

%\end{figure}

%
%
%\begin{figure}

%	\centering

%	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_low.png}

%	\caption{Pareto static energy}

%	\label{fig:pareto_static_l}

%\end{figure}

%
%
%\begin{figure}

%	\centering

%	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_w_high.png}

%	\caption{Pareto w energy}

%	\label{fig:pareto_w_h}

%\end{figure}

%
%\begin{figure}

%	\centering

%	\includegraphics[width=\columnwidth]{models/figures/analisys/pareto_static_low.png}

%	\caption{Pareto w energy}

%	\label{fig:pareto_w_l}

%\end{figure}

%
%\subsection{Optimization under constraints}

%
%\subsection{Gradient and Countorns}

%
%\begin{figure}[H]

%	\centering

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn0.png}

%	\end{subfigure}

%	%

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn0_3d.png}

%	\end{subfigure}

%\end{figure}

%
%\begin{figure}[H]

%	\centering

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn3.png}

%	\end{subfigure}

%	%

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/pdyn3_3d.png}

%	\end{subfigure}

%\end{figure}

%
%\begin{figure}[H]

%	\centering

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/pleak0.png}

%	\end{subfigure}

%	%

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/pleak0_3d.png}

%	\end{subfigure}

%\end{figure}

%
%\begin{figure}[H]

%	\centering

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/pleak10.png}

%	\end{subfigure}

%	%

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/pleak10_3d.png}

%	\end{subfigure}

%\end{figure}

%
%\begin{figure}[H]

%	\centering

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic0.png}

%	\end{subfigure}

%	%

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic0_3d.png}

%	\end{subfigure}

%\end{figure}

%
%\begin{figure}[H]

%	\centering

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic3000.png}

%	\end{subfigure}

%	%

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/pstatic3000_3d.png}

%	\end{subfigure}

%\end{figure}

%
%\begin{figure}[H]

%	\centering

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/w0.png}

%	\end{subfigure}

%	%

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/w0_3d.png}

%	\end{subfigure}

%\end{figure}

%
%\begin{figure}[H]

%	\centering

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/w1.png}

%	\end{subfigure}

%	%

%	\begin{subfigure}[b]{0.45\textwidth}

%		\includegraphics[width=\textwidth]{models/figures/analisys/w1_3d.png}

%	\end{subfigure}

%\end{figure}

### 2022-04-27 thesis/phases/phases.tex
add journals and restructured
carlos revision 1

@@
-\section{The impact of best-choice execution time slots partition of HPC applications on energy-saving strategies}
+\section{Model with phases}
 
@@
 
-\subsection{Introduction}
+\section{Introduction}
 Talk about the importance of energy efficiency.
 
-\subsection{Related work}
+\section{Related work}
 Some related work
 
-\subsection{Methodology}
+\section{Methodology}
 
-\subsubsection{Optimal energy configuration}
+\section{Optimal energy configuration}
 Splitting a program into phases, or time intervals with a selected set of execution parameters, has an impact on energy consumption.
@@
 
-\subsubsection{Data Collection}
+\section{Data Collection}
 Data, power samples, can be gathered by running the application in several possible context configurations.
@@
 
-\subsubsection{The simulator}
+\section{The simulator}
 To analyse the impact of configurations and identify optimal time-intervals, phases, we propose an integrator, which computes the energy on an interval for a given configuration, and a reference, a brute force test of all possible combinations.
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
     \centering
@@
 
-\subsubsection{Optimizing phases}
+\section{Optimizing phases}
 The main goal of this section is to identify the optimal number and duration of time intervals.
@@
 
-\subsection{Results}
+\section{Results}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \includegraphics[width=\columnwidth]{fingerprint/figures/energy_per_phase.png}
@@
 The images below show the phases for all applications:
-\begin{figure}[h]
+\begin{figure}[H]
 \includegraphics[width=\columnwidth]{phases/figures/phase_division_35.png}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \includegraphics[width=\columnwidth]{phases/figures/phase_division_3.png}
@@
 
-\subsection{Conclusion and future work}
+\section{Conclusion and future work}
 The main conclusion of this work was that in HPC enviroment we don't need a lot of phases to optimize the energy consumption given that for one phase we can compute the optimal energy.

### 2022-04-27 thesis/conclusion.tex
change subsection to section and labels

@@
-\section{Conclusion} \label{sec:conclusion}
+\section{Conclusion Models} \label{sec:conclusion_models}
 This paper proposes an energy model based on the operating frequency and the number of cores for a shared memory system. This model serves as a reference for DVFS and DPM optimization problems.
@@
 
-\subsection{Conclusions and Future Works Pascal}\label{sec:conclusions_pascal}
+\subsection{Conclusions and Future Works Pascal} \label{sec:conclusions_pascal}
 
@@
 
-\subsection{Conclusion fingerprint}
+\subsection{Conclusion fingerprint tool}
 
@@
 This can be applied to various areas such as voltage scaling and frequency of the processor, once knowing the behavior of a particular program we can have an idea of what is the best strategy to control the frequency in order to save energy or increase performance.
+
+\section{Conclusion and future work: Phases}
+The main conclusion of this work was that in HPC enviroment we don't need a lot of phases to optimize the energy consumption given that for one phase we can compute the optimal energy.
+
+The final idea is to use the information of the best phase to find a performance metric that can provide us this division, in future work we intend to compute this metrics using combination of hardware performance counters enabling optimal phase division in real time.
+This can improve current DVFS algorithms giving them an workload metric that can improve the energy savings.

### 2022-04-27 thesis/experiments/experiments.tex
change subsection to section and labels

@@
 
-\section{Validation the framework}\label{sec:validation_the_framework}
+\section{Pascal framework validation} \label{sec:pascal_framework_validation}
 In this Section, we discuss the measurements and simulation results for three distinct cases, including two real applications.
 We also demonstrate how external libraries and standalone visualization tools can be used to render the collected data and investigate performance aspects such as the program's scalability capacity.
-
-\subsection{Configuration} \label{subsec:configuration_pascal}
 
@@
 
-\subsection{Pascal Analyzer Validation}\label{sec:results_pascal}
+\subsection{Case of study}\label{subsec:pfv_case_of_study}
 
@@
 % \end{figure}
-\subsection{Fingerprint Tool validation}
+\section{Fingerprint Tool validation} \label{sec:fingerprint_tool_validation}
 
@@
 
-\subsubsection{ACCURACY COMPARISON}
+%\subsection{Accuracy comparison} \label{subsec:ftv_accuracy_comparison}
 
@@
 
-\subsubsection{CLUSTERING}
+\subsubsection{Case of study} \label{subsec:ftv_case_of_study}
 

### 2022-04-27 thesis/fingerprint/fingerprint.tex
change subsection to section and labels

@@
-\section{Fingerprint tool}
+\section{Profiling with performance counters} \label{sec:profiling_with_performance_counters}
 The analysis of application performance is essential to better exploit its potential on High-Performance Computing (HPC) architectures.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Introduction}
+\section{Collecting performance counters} \label{sec:collecting_performance_countersl}
 
@@
 
-\subsection{Related Work}
+\section{State of the art performance counters APIs} \label{sec:state_of_the_art_performance_counters_APIs}
 
@@
 
-\subsection{Reading Performance Counters}
+\section{Accounting for errors and limitations} \label{sec:accounting_for_errors_and_limitations}
 
@@
 
-\subsection{Implementation}
+\section{Implementing a tool to read performance counters precisely} \label{sec:implementing_a_tool_to_read_performance_counters_precisely}
 
@@
 
-\subsubsection{ARCHITECTURE}
+\section{PMC tool architecture} \label{sec:PMC_tool_architecture}
 
@@
 
-\subsubsection{POST-PROCESSING}
-\label{sec:posprocessing}
+\section{Post-processing} \label{sec:posprocessing}
 % With this API we create clusterize the applications of Polybench applications.

### 2022-04-27 thesis/optmization.tex
change subsection to section and labels

@@
 
-\subsection{Verifying Hypothesis}
+\subsection{Verifying Hypothesis} \label{subsec:ev_verifying_hypothesis}
 
@@
 
-\subsection{Frequency and Voltage Relation}
+\subsection{Frequency and Voltage Relation} \label{subsec:ev_frequency_and_voltage Relation}
 One of the assumptions was that the frequency and the voltage have a linear relationship, as indicated by ~\cref{eq:f_v}.
@@
 
-\subsection{Input Size and Instructions}
+\subsection{Input Size and Instructions} \label{subsec:ev_input_size_and_instructions}
 We ran the applications with different inputs assuming  linear growth in the amount of work for one input to the other when building our model.
@@
 
-\section{Fitting the Models} \label{subsec:fitting}
+\section{Fitting the Models} \label{subsec:fitting_the_models}
 To find the parameters of  \cref{eq:en_final}, 10 uniformly random configurations of frequencies ($f$), cores ($p$) and inputs ($N$) were chosen from the range $1<=p<=32$, $1.2<=f<=2.2$ and $1<=N<=5$, respectively.
@@
 
-\section{Measured versus Modeled Energy}
-\label{subsec:measuredversusmodeledenergy}
+\section{Measured versus Modeled Energy} \label{sec:measured_versus_modeled_energy}
 
@@
 
-\subsection{Frequency $\times$ Cores}
+\subsection{Frequency $\times$ Cores} \label{subsec:mvme_frequency_x_cores}
 \cref{fig:en_eq_freq_cores_fc} plots the measured and modeled energy consumption for some of the applications modeled.
@@
 \end{figure}
-\subsection{Frequency $\times$ Input}
+
+\subsection{Frequency $\times$ Input} \label{subsec:mvme_frequency_x_input}
 \cref{fig:en_eq_freq_inp_fi} plots the measured and modeled energy consumption for some of the applications modeled.
@@
 \end{figure}
-\subsection{Cores $\times$ Input}
+
+\subsection{Cores $\times$ Input} \label{subsec:mvme_cores_x_input}
 \cref{fig:en_eq_core_inp_ci} plots the measured and modeled energy consumption for some of the applications modeled.
@@
 
-\subsection{Comparison}
+\section{Comparison} \label{sec:comparison}
 The average results for each application were calculated using a model trained with only 10 configurations, and the comparison is displayed \cref{fig:mpe_svr_eq}.
@@
 
-\section{Overheads on Training} \label{subsec:overhead}
+\section{Overheads on Training} \label{subsec:overhead_on_training}
 It is known that machine learning is data-driven; in that sense, the SVR model obtained using only 10 configurations could be improved, but what about the analytical model? 
@@
 
-\section{Analysis}
+\section{Analysis} \label{sec:analysis}
 One of the most significant advantages of using an analytical model is the understanding of the problem that an equation provides, making many different kinds of analysis possible that are otherwise impossible with a machine learning model.
@@
 
-\section{DVFS and DPM Optimization} \label{subsec:dvfs_optmin}
+\section{DVFS and DPM Optimization} \label{sec:dvfs_dpm_optminzation}
 The effectiveness of the proposed approach during optimization was evaluated with a simple algorithm that finds the optimal frequency and number of active cores from the proposed equation.

### 2022-04-27 thesis/pascalanalyzer/pascalanalyzer.tex
carlos revision 1
change subsection to section and labels

@@
 
-\subsection{Introduction} \label{sec:introduction_pascal}
+\section{Introduction to profiling tools} \label{sec:introduction_to_profiling_tools}
 
@@
 
-\subsection{Related Works} \label{sec:related_work_pascal}
+\section{State of the art: profiling and tracing tools} \label{sec:state_of_the_art_profiling_and_tracing_tools}
 
@@
 
-\subsection{Design and Features} \label{sec:design_and_features_pascal}
-
-This section presents the architecture, components, and interconnections used in the design of the proposed tool.
+%\section{Design and Features} \label{sec:design_and_features}
+%
+%This section presents the architecture, components, and interconnections used in the design of the proposed tool.
 It also includes a discussion about main features, instrumentation modes, and output file structure that contains the data measured and collected by the proposed tool.
 
-\subsubsection{Architecture} \label{subsec:architecture_pascal}
+\section{Pascal architecture} \label{sec:pascal_architecture}
 The tool we propose to measure, combine and compare multiple runs of a parallel application implements two main concepts to achieve this goal: actuators and sensors.
@@
 
-\subsubsection{Instrumentation and Intrusiveness} \label{subsec:intrusiveness_pascal}
+\section{Instrumentation and Intrusiveness} \label{sec:instrumentation_and_intrusiveness}
 
@@
 	\includegraphics[width=\textwidth]{pascalanalyzer/figures/designfeatures/instrumentation_overhead_summary.pdf}
-	\caption{\centering}
+%	\caption{\centering}
 	\label{fig:overhead_1}
@@
 	\includegraphics[width=\textwidth]{pascalanalyzer/figures/designfeatures/instrumentation_overhead_boxplot.pdf}
-	\caption{\centering}
+%	\caption{\centering}
 	\label{fig:overhead_2}
@@
 
-\subsubsection{Features and Usage} \label{subsec:features_and_usage_pascal}
+\section{Features and Usage} \label{sec:features_and_usage}
 
@@
 
-\subsubsection{Exported Data Structure} \label{subsec:datastructure_pascal}
+\section{Exported Data Structure} \label{sec:exported_data_structure}
 

### 2022-04-27 thesis/phases/phases.tex
change subsection to section and labels

@@
-\section{Model with phases}
+\section{Model with phases} \label{sec:model_with_phases}
 
@@
 
-\section{Introduction}
+\section{Introduction to models on time} \label{sec:introduction_to_models_on_time}
 Talk about the importance of energy efficiency.
 
-\section{Related work}
+\section{Related work: Phases} \label{sec:related_work_phases}
 Some related work
 
-\section{Methodology}
-
-\section{Optimal energy configuration}
+\section{Optimal energy configuration} \label{sec:Optimal_energy_configuration}
 Splitting a program into phases, or time intervals with a selected set of execution parameters, has an impact on energy consumption.
@@
 
-\section{Data Collection}
+\section{Data Collection} \label{sec:data_collection}
 Data, power samples, can be gathered by running the application in several possible context configurations.
@@
 
-\section{The simulator}
+\section{The simulator} \label{sec:the_simulator}
 To analyse the impact of configurations and identify optimal time-intervals, phases, we propose an integrator, which computes the energy on an interval for a given configuration, and a reference, a brute force test of all possible combinations.
@@
 
-\section{Optimizing phases}
+\section{Optimizing phases} \label{sec:optimizing_phases}
 The main goal of this section is to identify the optimal number and duration of time intervals.
@@
 
-\section{Results}
+\section{Results: Phases experiments} \label{sec:results_phases_experiments}
 
@@
 Plotting for 3 phases, we can observe that better.
-
-\section{Conclusion and future work}
-The main conclusion of this work was that in HPC enviroment we don't need a lot of phases to optimize the energy consumption given that for one phase we can compute the optimal energy.
-
-The final idea is to use the information of the best phase to find a performance metric that can provide us this division, in future work we intend to compute this metrics using combination of hardware performance counters enabling optimal phase division in real time.
-This can improve current DVFS algorithms giving them an workload metric that can improve the energy savings.

### 2022-05-03 thesis/conclusion.tex
fix refs

@@
 Moreover, it can provide knowledge-based hints to improve DVFS and DPM algorithms by enabling analysis of the contribution of each model parameter (e.g., level of parallelism) to the energy consumption.
-Indeed, as shown in Section \ref{subsec:dvfs_optmin}, when no oracle is available to choose the frequency and the number of cores the application should use, the proposed model can save around 12\% of energy for a random choice and up to 70\% for the worse possible choice.
+Indeed, as shown in Section \ref{sec:model_validation}, when no oracle is available to choose the frequency and the number of cores the application should use, the proposed model can save around 12\% of energy for a random choice and up to 70\% for the worse possible choice.
 Considering the job history of our own HPC center, which shows the prevalence of worse possible choices made by users, the potential energy savings are very significant and encourage further research.
@@
 Results from 3 HPC benchmarks running in one cluster demonstrate the potential of the proposed novel model.
-While consuming less energy than traditional machine learning approaches, it can serve as bases from DVFS and DPM algorithm as shown in the \ref{subsec:dvfs_optmin} in an average case saving about 12\% up to 69\%.
+While consuming less energy than traditional machine learning approaches, it can serve as bases from DVFS and DPM algorithm as shown in the \ref{sec:model_validation} in an average case saving about 12\% up to 69\%.
 The previous knowledge of the application's performance can expose sufficiently relevant information, such as parallel speedups, that is harder to guess in run-time techniques based on DVFS.

### 2022-05-03 thesis/optmization.tex
fix refs

@@
-\section{Validation the model} \label{sec:experimentalvalidation}
+\section{Model Validation} \label{sec:model_validation}
 In this section, the models presented in~Sections \ref{sec:power_model} and~\ref{sec:performance_model} were validated with a benchmark specific for multi-core architectures.
@@
 \cref{fig:mpe_svr_eq} shows that the proposed model always performed better, with a lower MPE than SVR, when we were limited to 10 training points.
-This result is further explored in the next Section \ref{subsec:overhead}, where we undertake a comparison with different training sizes.
+This result is further explored in the next Section \ref{sec:overhead_on_training}, where we undertake a comparison with different training sizes.
 The exact values are shown in \cref{tab:mpe_svr_eq}.

### 2022-05-03 thesis/pascalanalyzer/pascalanalyzer.tex
fix refs

@@
 The following sections describe how the tool is built and used for performance analysis.
-Section \ref{sec:related_work_pascal} presents the related works.
-Section \ref{sec:design_and_features_pascal} describes the tool architecture and goals; it explains the main features, usage, and collected data.
-The experimental results are presented in Section \ref{sec:validation_the_framework}.
+Section \ref{sec:state_of_the_art_profiling_and_tracing_tools} presents the related works.
+Section \ref{sec:pascal_architecture} describes the tool architecture and goals; it explains the main features, usage, and collected data.
+The experimental results are presented in Section \ref{sec:pascal_framework_validation}.
 Finally, the contributions are summarized in Section \ref{sec:conclusions_pascal}, with an outlook of future works.
@@
 Performance analysis aimed at code optimization is an essential activity for developing parallel applications with scalable performance.
-For this purpose, performance analysis tools, like Caliper~\cite{Boehme2016}, HPCToolkit~\cite{Adhianto2010}, Scalasca~\cite{Geimer2010}, TAU~\cite{Shende2006}, Vampir~\cite{Weber2019}, and VTune~\cite{VTune}, are useful because they help developers identify where the application uses the available computing resources inefficiently through collecting detailed data of its execution.
+For this purpose, performance analysis tools, like Caliper~\cite{Boehme2016}, HPCToolkit~\cite{Adhianto2010}, Scalasca~\cite{Geimer2010}, TAU~\cite{Shende2006}, Vampir~\cite{Weber2019}, and VTune~\cite{Intel2021Vtune}, are useful because they help developers identify where the application uses the available computing resources inefficiently through collecting detailed data of its execution.
 Such tools differ from each other mainly in terms of data measurement and analysis strategy: tracing or profiling, post-mortem or real-time; individual or comprehensive observation focus; and for providing or not providing visual elements that facilitate the judgment of collected data.
@@
 Vampir~\cite{Weber2019} and Paraver~\cite{Labarta2005} are examples of this group.
-Some tools like HPCToolkit~\cite{Adhianto2010}, Scalasca~\cite{Geimer2010}, TAU~\cite{Shende2006} and VTune~\cite{VTune} support both profiling and~tracing.
+Some tools like HPCToolkit~\cite{Adhianto2010}, Scalasca~\cite{Geimer2010}, TAU~\cite{Shende2006} and VTune~\cite{Intel2021Vtune} support both profiling and~tracing.
 
@@
 In this case, the tool performs performance measurements and data collection while the program is running, and then the collected data will be interpreted.
-HPCToolkit~\cite{Adhianto2010}, Scalasca~\cite{Geimer2010}, TAU~\cite{Shende2006} and VTune~\cite{VTune} are examples of post-mortem performance analysis tools. 
+HPCToolkit~\cite{Adhianto2010}, Scalasca~\cite{Geimer2010}, TAU~\cite{Shende2006} and VTune~\cite{Intel2021Vtune} are examples of post-mortem performance analysis tools. 
 %Post-mortem performance analysis tools may require the storage of large volumes of performance data, but they are more suitable to provide an overall view of the execution. 
@@
 A key point of performance analysis tools, particularly important when analyzing real-time program execution, is that instrumentation must have a negligible impact on program behavior and execution time.
-Indeed, as mentioned in Section \ref{subsec:architecture_pascal}, we support three types of sensors, each of which has a different degree of intrusion.
+Indeed, as mentioned in Section \ref{sec:pascal_architecture}, we support three types of sensors, each of which has a different degree of intrusion.
 There is hardly any sensors intrusion at the start and then at the end of the program execution since it is simply the invocation of both data collection routines.

### 2022-05-03 thesis/framework/validation.tex
change structure

--- 2022-04-27 thesis/experiments/experiments.tex
@@
-\section{Frequency control} \label{sec:frequency_control}
-
-In modern systems, control over the processor frequency can be done by hardware with independent circuits as well as by software.
-For that, there is the Advanced Configuration and Power Interface (ACPI) an open standard adopted by operating systems to configure hardware components related to power management.
-
-In the ACPI two important states for the DVFS are defined that optimize the energy consumption.
-They are "C", which is activated when the processor is not executing any instructions, and "P", which is activated while the processor is operating.
-These states have several levels and at each level the frequency and voltage are changed.
-
-State P starts at level P0, where frequency and voltage are the maximum possible, then P1, where both decrease, until reaching the last state, Pn, where frequency and voltage are the lowest possible.
-The change of state depends on the level of utilization of the processor.
-To stay in each state, the level of processor usage must be within specific limits.
-After exceeding these limits for a certain time, the state will change to the next state corresponding to that new level of processor usage.
-The number of possible states depends on each manufacturer.
-
-After an idle time, the processor begins to activate C states, starting with C0 where it is still fully active, then moving to C1, where some features are disabled, up to Cn, where all possible features are disabled.
-The ACPI standard establishes the functionalities that can be disabled between level C1 and level C3, as seen in Table \ref {tab:states_c}.
-The other levels are specific to each manufacturer.
-
-\begin{table}[H]
-	\centering
-	\begin{tabular}{| l | l | l |}
-		\hline
-		Mode & Name & Functionality \\ \hline
-		C0 & operating state & Active processor \\ \hline
-		C1 & Halt & Stop executing instructions \\ \hline
-		C2 & Stop-Clock & Disable the internal clock \\ \hline
-		C3 & Sleep & Disable cache coherence \\ \hline
-	\end{tabular}
-	\caption{C states}
-	\label{tab:states_c}
-\end{table}
-
-In state C, the higher the level the greater the energy savings, but returning to the fully functional level is more difficult.
-In states P, there is a trade-off between performance and energy savings.
-\cref{fig:p_state} best illustrates the change of states, in which we can see which parts of the circuits are deactivated in states C, the latency to return to the active state, power consumption and also shows the relationship of states P with often.
-
-\begin {figure} [H]
-\centering
-\begin{subfigure}[t]{0.5\textwidth}
-	\centering
-	\includegraphics[width=\columnwidth]{intro/figures/c_states}
-	\caption{C states}
-\end{subfigure}%
-~
-\begin{subfigure}[t]{0.5\textwidth}
-	\centering
-	\includegraphics[width=\columnwidth]{intro/figures/p_states}
-	\caption{States P}
-\end{subfigure}
-\caption {Illustration of states C and P} {Altered image from \protect \url {https://www.thomas-krenn.com/en/wiki/Processor_P-states_and_C-states}}
-\label{fig:p_state}
-\end{figure}
-
-% Are states C and P orthogonal, do they operate independently?
-
-For this management, the operating system provides in the user space a way to control the frequency. This work used an operating system that has Linux as its core.
-
-Linux is compatible with several modern architectures and is widely used in servers, smartphones and supercomputers.
-It was based on the UNIX system which has the philosophy of treating everything on the system as a file, including settings and input and output devices, such as keyboard, mouse and hard drive.
-Another important feature is that it is modular and parts of the system can be loaded or removed during execution.
-
-On Linux there are several frequency management options \cite {Brown2005ACPILinux}.
-The main ones are acpi-cpufreq, Intel P-state, AMD powernow.
-In this work, acpi-cpufreq is used, which is standard and allows direct control of frequency through system files.
-Acpi-cpufreq is a Linux module that uses implemented policies that dynamically decide the frequency to be used.
-Some of these policies are:
-
-\begin{itemize}
-\item Performance - configured as often as possible
-\item Powersave - configured as low as possible
-\item Userspace - the user chooses the frequency to be used
-\item Ondemand - controls the frequency depending on the processor load. When the load increases the frequency also increases accordingly.
-\item Conservative - similar to Ondemand but more smoothly, the frequency increase is continuous instead of jumping.
-\end{itemize}
-
-\section{Power consumption monitoring} \label{sec:power_consumption_monitoring}
-The Running Average Power Limit (RAPL) and Intelligent Platform Management Interface (IPMI) interfaces were used to measure the power consumed. described in \cite{IPMI2013ConfigurationGuide}.
-
-\subsection{IPMI}
-IPMI \cite{IPMI2013ConfigurationGuide} is a set of specifications for autonomous subsystems that provides processor, firmware and operating system independent management and monitoring.
-The use of IPMI allows system administrators to avoid having to travel to the server location, which is often far away, to perform their tasks.
-Also, servers are located in places with low temperature and with a lot of noise due to the ventilation system, and one should avoid spending too much time in these places.
-With remote management, it is possible to turn the system on and off, remotely access the (BIOS) and reinstall the system in case of any serious failure.
-
-\begin{figure}[H]
-\centering
-\includegraphics[height = 7.5cm] {intro/figures/IPMI-Block-Diagram.png}
-\caption{IMPI diagram} {Image taken from \protect \url {https://pt.wikipedia.org/wiki/Intelligent_Platform_Management_Interface}, the main components of IPMI and how they communicate are shown}
-\label{fig:IPMI}
-\end{figure}
-
-Access to the IPMI network can be done using the HTTP protocol or a tool made available by the manufacturer (ipmitool), which also performs access via the network.
-It is also used to monitor the status of the platform with a set of sensors coupled with system temperatures, voltages, fans and power supplies.
-
-\subsection{RAPL}
-
-Modern Intel microprocessors, based on the SandyBridge architecture, include the RAPL \cite {Rotem2012Power-managementBridge, Hahnel2012RAPL, Hackenberg2015AnProcessor} interface designed to limit the use of energy on a chip while ensuring maximum performance.
-This interface supports energy measurement capabilities through an integrated circuit that estimates energy use based on a model driven by architectural event counters for all components.
-It also provides temperature readings and current leak models.
-Estimates are made available in model-specific registers (MSR), updated in milliseconds.
-The energy estimates offered by RAPL were validated by Intel, which showed excellent results.
-
-\section{Case-Study Architecture} \label{sec:casestudyarchitecture}
-The experiments were executed in one computer node equipped with two Intel Xeon E5-2698 v3 processors with sixteen cores each and two hardware threads for each core. 
-The overall view of the architecture is shown in \cref{fig:architecture}.
-The maximum non-turbo frequency was 2.3 GHz, and the total physical memory of the node was 128 GB (8 $\times$ 16 GB).
-Turbo frequency and hardware multi-threading were disabled during all experiments.
-The operating system used was Linux CentOS 6.5, kernel 4.16.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\columnwidth]{models/figures/architecture.png}
-\caption{Node architecture (the image was made with the lstop application).}
-\label{fig:architecture}
-\end{figure}
-
-The Linux kernel has many different policies for power management, depending on the driver.
-In the default driver, the acpi-cpufreq, the options are Powersave, Performance, Ondemand, Conservative, and Userspace.
-Each governor has a policy on how the frequency is selected.
-In this investigation, the frequency control was performed using the Userspace governor, which allows the user or any userspace program to set the CPU to a specific frequency.
-The core control was accomplished by modifying the appropriate system files with the default CPU-hotplug driver.
-
-The architecture was equipped with the intelligent platform management interface (IPMI), a set of interfaces allowing out-of-band management of computer systems and platform-status monitoring via the local network~\cite{Schwenkler2006IntelligentInterface}.
-It can monitor variables and resources, such as the system's temperature, voltage, fans, and power supplies, with independent sensors attached to the hardware.
-
-\section{Case-Study Applications} \label{sec:casestudyapplication}
-The applications blackscholes, bodytrack, canneal, dedup, fluidanimate, freqmine, raytrace, swaptions, vips and x264 from the PARSEC \url{https://parsec.cs.princeton.edu/download.htm} (accessed on 20 February  2020) parallel benchmark suite, version 3.0~\cite{Bienia2008TheSuite}, OpenMC \cite{Romano2015OpenMC:Development} and LINPACK (HPL) \cite{Dongarra1988TheExplanation}, were chosen as case studies.
-The PARSEC benchmark focused on emerging workloads and was designed to represent the next-generation shared-memory programs for chip-multiprocessors.
-It covers an ample range of areas, such as financial analysis, computer vision, engineering, enterprise storage, animation, similarity search, data mining, machine learning, and media processing.
-The OpenMC and the LINPACK are two classic HPC programs.
-
-%The other two experiments use applications Raytrace~\cite{Bienia2008} and OpenMC~\cite{Romano2015OpenMC:Development}.
-Raytrace is an application addressed on rendering photorealistic scenes.
-It is available on the PARSEC Benchmark suite, version 2.0.
-The PARSEC benchmark, focused on emerging workloads, represents the next-generation shared-memory programs for chip multiprocessors.
-It covers a wide range of areas such as financial analysis, computer vision, engineering, enterprise storage, animation, similarity search, data mining, machine learning, and media~processing.
-%
-%\textls[-20]{{OpenMC is an application that implements the Monte-Carlo method to simulate the transport of neutrons and photons. It is a classic program aimed at high-performance computing.}}
-
 \section{Pascal framework validation} \label{sec:pascal_framework_validation}

### 2022-05-03 thesis/experiments/setup.tex
fix section titles

--- 2022-04-27 thesis/experiments/experiments.tex
@@
 
-\section{Case-Study Architecture} \label{sec:casestudyarchitecture}
+\section{Case-Study architecture} \label{sec:casestudyarchitecture}
 The experiments were executed in one computer node equipped with two Intel Xeon E5-2698 v3 processors with sixteen cores each and two hardware threads for each core. 
@@
 
-\section{Case-Study Applications} \label{sec:casestudyapplication}
+\section{Case-Study applications} \label{sec:casestudyapplication}
 The applications blackscholes, bodytrack, canneal, dedup, fluidanimate, freqmine, raytrace, swaptions, vips and x264 from the PARSEC \url{https://parsec.cs.princeton.edu/download.htm} (accessed on 20 February  2020) parallel benchmark suite, version 3.0~\cite{Bienia2008TheSuite}, OpenMC \cite{Romano2015OpenMC:Development} and LINPACK (HPL) \cite{Dongarra1988TheExplanation}, were chosen as case studies.
@@
 %\textls[-20]{{OpenMC is an application that implements the Monte-Carlo method to simulate the transport of neutrons and photons. It is a classic program aimed at high-performance computing.}}
-
-\section{Validation the framework}\label{sec:validation_the_framework}
-In this Section, we discuss the measurements and simulation results for three distinct cases, including two real applications.
-We also demonstrate how external libraries and standalone visualization tools can be used to render the collected data and investigate performance aspects such as the program's scalability capacity.
-
-\subsection{Configuration} \label{subsec:configuration_pascal}
-
-We used three experiments to assess the tool and demonstrate its ability to support analysis aimed at observing parallel scalability.
-The first includes a runtime imbalance between processing units in a specific parallel region.
-In this case, the objective is to present how the analyzer helps users observe the impact of the inefficiency of a code part on the whole program.
-The code for this experiment is presented in \cref{lst:codes_regionscomparison}.
-It consists of two simple parallel regions with the same functionality that divide the iterations of a loop between the available threads.
-The difference between the regions lies in the strategies used to manipulate the sum variable, used in the example to store the values of the calculations performed in each thread.
-We assume that the {\tt anyop()} function, in this case, invariably has the same runtime in all function calls.
-
-The command presented on \cref{lst:pascal_usage_ex} was used as base to perform experiments.
-For the experiment of \cref{lst:codes_regionscomparison}, we use just the parameters {\tt -c}, {\tt -r}, {\tt -t}, {\tt -a} and {\tt -o}, including the 64 value to {\tt -c} option.
-
-\lstset{style=ccodestyle, frame=tb}
-\begin{lstlisting}[label={lst:pascal_usage_ex}, language=bash, caption={Command line showing how experiments were run through a terminal.}]
-	analyzer 
-	application
-	-c 1,2,4,8,...,32 # threads/cores
-	-r 3 # number of repetitions
-	-t auto # automated instrumentation
-	-a 1 # aggregation mode
-	--ipmi ip user psswd # energy sensor
-	--idtm 5 # idle time between runs
-	--dhpt # disable hyper-thread
-	--dcrs # disable cores
-	--ipts ... # application specific inputs
-	-o application.json # output file
-\end{lstlisting}
-
-\lstset{style=ccodestyle, frame=tb}
-\begin{lstlisting}[label={lst:codes_regionscomparison}, language=C, caption={Sample code used to visualize the impact of regions on program scalability.}]
-	int main(int argc, char **argv) {
-		unsigned long sum = 0;
-		int ops = atoi(argv[1]);
-
-		#pragma omp parallel for schedule(static) reduction(+: sum)
-		for (int i = 0; i < ops; i++)
-		sum += anyop();
-
-		#pragma omp parallel for schedule(static)
-		for (int i = 0; i < ops; i++) {
-			#pragma omp critical
-			sum += anyop();
-		}
-	}
-\end{lstlisting}
-
-The command return a .json file with all the information necessary for our analysis.
-We can quickly generate tables with the collected data from this file, thus supporting observation and analysis of scalability, energy trends, and model fitting.
-The code described in \cref{lst:regtab} and \cref{tab:regtable_dedup} are samples of how users can easily read and view collected data from the control terminal in a tabular way.
-
-\lstset{style=pythonStyle, frame=tb}
-\begin{lstlisting}[label={lst:regtab}, language=python, captionpos=b, caption={Example of using the Python API to load analyzer files.}]
-	from analyzer import Data
-
-	data = Data("application.json")
-	data.energy(regions=True)
-	# data.speedup()
-	# data.efficiency()
-	# data.regions()
-\end{lstlisting}
-
-%The result can be seen in tabular form, as shown in Table.~\ref{tab:regtable_dedup}.
-Using these tables, we can quickly generate figures showing the application's behavior from various points of view.
-In Figures \ref{fig:speedup_raytrace} and \ref{fig:speedup_openmc} we observe the trend of the speedup with the number of cores and input.
-\begin{table}[H]
-	\caption{Dataframe generated automatically from collected samples using the Python API.}
-	\label{tab:regtable_dedup}
-
-	\resizebox{\columnwidth}{!}{%
-		\begin{tabular}{ccccccc}
-			\toprule
-			\textbf{Repetition} & \textbf{Input} & \textbf{Cores} & \textbf{Regions} & \textbf{Start Time (s)} & \textbf{Stop Time (s)} & \multicolumn{1}{c}{\textbf{ipmi Energy (J)}} \\ \midrule
-			1 & 1 & 1 & 1 & 0.00 & 66.31 & 13,156.19 \\
-			1 & 1 & 1 & 2 & 0.00 & 66.27 & 13,148.87 \\
-			... & ... & ... & ... & ... & ... & ... \\
-			1 & 5 & 30 & 4 & 0.00 & 8.46 & 1903.82 \\
-			1 & 5 & 30 & 5 & 0.02 & 58.16 & 13,067.01 \\ \bottomrule
-	\end{tabular}}
-\end{table}
-
-\subsection{Pascal Analyzer Validation}\label{sec:results_pascal}
-
-The analyzer does not display graphics and other visual elements natively.
-However, the simple use of external libraries allows generating graphics and visualizing points of the program's behavior that you want to observe.
-In addition, specialized visualization tools can also be used to view the results.
-PaScal Viewer \cite{Silva2018}, for example, natively interprets the analyzer's output files, complementing its functionality and creating an integrated and appropriate environment for the program's parallel scalability analysis.
-
-To analyze the experiment presented in \cref{lst:codes_regionscomparison}, we use PaScal Viewer to observe in a hierarchical way how the different OpenMP clauses impacted the region's efficiencies.
-The efficiency pinpoints how a program can take advantage of increasing processing elements on a parallel system.
-It is defined as the ratio of speedup to the number of processing units.
-In this sense, PaScal Viewer displays an efficiency diagram for each analysis region, as presented in \cref{fig:pv_regionscomparison}.
-Comparing these diagrams, it is possible to see how the critical clause damages the scalability.
-In addition, it is also possible to relate that this second region affects the efficiency of the whole program due to the code characteristics.
-
-If the reduction clause replaces {\tt \#pragma~omp~crtical}, the second region and the entire program become more efficient, as shown in \cref{fig:pv_regionscomparison_2}.
-The PaScal Viewer diagrams mentioned in this work, the x-axis (i1--i7) corresponds to different data inputs and the y-axis (2--64) to the number of threads used in the processing.
-Figures~\ref{fig:pv_regionscomparison} and \ref{fig:pv_regionscomparison_2} were rendered using a tool feature that smoothes the color transition of diagrams.
-This feature uses interpolation to create a visual element where the color transition is less pronounced.
-The diagram axes only show the initial and final values with this option.
-The user can visualize diagrams with only discrete values or even compare the two presentation modes side by side.
-\cref{fig:visualizationmodes} shows the difference between discrete and smoothed modes considering the whole program and the simulation scenario that uses the clause {\tt \#pragma~omp~crtical}.
-\begin{figure}[H]
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_r1.pdf}
-%		\caption{\centering}
-		\label{fig:pv_regionscomparison_a}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_r2.pdf}
-%		\caption{\centering}
-		\label{fig:pv_regionscomparison_b}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/regionscomparison_whole.pdf}
-%		\caption{\centering}
-		\label{fig:pv_regionscomparison_c}
-	\end{subfigure}
-	\caption{Efficiency diagrams and impact of inner regions on program scalability.
-(x-axis = inputs; y-axis = number of threads).
-(\textbf{a}) First region diagram.
-(\textbf{b}) Second region diagram.
-(\textbf{c}) Whole program~diagram.}
-	\label{fig:pv_regionscomparison}
-\end{figure}
-\unskip
-\begin{figure}[H]
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_1.pdf}
-%		\caption{\centering}
-		\label{fig:pv_regionscomparison_a_2}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_2.pdf}
-%		\caption{\centering}
-		\label{fig:pv_regionscomparison_b_2}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_rg_3.pdf}
-%		\caption{\centering}
-		\label{fig:pv_regionscomparison_c_2}
-	\end{subfigure}
-	\caption{Efficiency diagrams after removing {\tt \#pragma~omp~crtical} clause.
-(x-axis = inputs; y-axis = number of threads).
-(\textbf{a}) First region diagram.
-(\textbf{b}) Second region diagram.
-(\textbf{c}) Whole program~diagram.}
-	\label{fig:pv_regionscomparison_2}
-\end{figure}
-\unskip
-\begin{figure}[H]
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/diagram_discretevalues.pdf}
-%		\caption{\centering}
-		\label{fig:discretevalues}
-	\end{subfigure}
-	%
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/diagram_smoothedvalues.pdf}
-%		\caption{\centering}
-		\label{fig:smoothedvalues}
-	\end{subfigure}
-
-	\caption{Visualization modes of diagrams provided by PaScal Viewer. (\textbf{a}) PaScal Viewer diagram on discrete mode. (\textbf{b}) PaScal Viewer diagram on smoothed mode.}
-	\label{fig:visualizationmodes}
-\end{figure}
-
-%The analyzer allows that the result can be seen in tabular form, as shown in Table.~\ref{tab:regtable_dedup}.
-Using these tables, we can quickly generate figures that show the application's behavior from various points of view.
-In Figures \ref{fig:speedup_raytrace} and \ref{fig:speedup_openmc} we can observe the trend of the speedup with the number of cores and input.
-
-%Figures \ref{fig:raytrace_en} and \ref{fig:openmc_en} present the energy variation with the number of cores.
-
-\textls[-19]{Other analysis objectives not supported by PaScal Viewer, such as visualization of the speedup curve or energy consumption, can be supported by traditional plots.
-Figures~\ref{fig:speedup} and \ref{fig:energy}} present the charts rendered for analysis by Raytrace and OpenMC applications.
-In these figures, it is possible to observe that the program efficiency varies according to the increase in the number of threads and the execution of inputs with higher processing loads.
-From the diagrams in \cref{fig:efficiency_all}, it is possible to observe that the applications present different behaviors concerning their efficiency variations and their scalability capabilities.
-OpenMC maintains its efficiency almost constant.
-This pattern represents strong scalability and indicates that the program can maintain its efficiency level when it uses a larger number of threads and processing larger inputs.
-On the other hand, \cref{fig:efficiency_all}a demonstrates that Raytrace achieves higher efficiency values when processing larger inputs.
-However, it is also possible to observe that increasing the number of processing units fixing the input size will not improve or hold the efficiency value.
-\begin{figure}[H]
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/speedup_completo_rtview_2 (1).pdf}
-%		\caption{\centering}
-		\label{fig:speedup_raytrace}
-	\end{subfigure}
-	%
-	\begin{subfigure}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/speedup_completo_openmc_kernel_novo (1).pdf}
-%		\caption{\centering}
-		\label{fig:speedup_openmc}
-	\end{subfigure}
-	\caption{\textls[-30]{Speedup of the applications for several input sizes. (\textbf{a}) Raytrace speedup. (\textbf{b}) OpenMC speedup. }}
-	\label{fig:speedup}
-\end{figure}
-\unskip
-\begin{figure}[H]
-	\begin{subfigure}[b]{0.46\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/energy_completo_rtview_2 (1).pdf}
-%		\caption{\centering}
-		\label{fig:raytrace_en}
-	\end{subfigure}
-	%
-	\begin{subfigure}[b]{0.46\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/energy_completo_openmc_kernel_novo (1).pdf}
-%		\caption{\centering}
-		\label{fig:openmc_en}
-	\end{subfigure}
-
-	\caption{Energy consumption identified in the execution of applications while varying the number of cores for several input sizes.
-		%MDPI: Please add commas to numbers to indicate thousand on the Oy axis of the figures, e.g. 20,000 30,000 40,000.
-		(\textbf{a}) Raytrace energy consumption. (\textbf{b}) OpenMC energy consumption.}
-	\label{fig:energy}
-\end{figure}
-\unskip
-%Furthermore, using an external graphical interface module, we can load the .json file and intuitively visualize the program's scalability showing the efficiency.
-The efficiency pinpoint how efficiently a program can take advantage of increasing processing elements on a parallel system.
-It is defined as the ratio of speedup to the number of processors.
-%For example, Figure \ref{fig:efficiency_all} shows examples of graphics generated for the same applications as above.
-In these figures, it is possible to observe that the program efficiency varies according to the increase in the number of threads and the execution of inputs with higher processing loads.
-From the Figure \ref{fig:efficiency_raytrace} we can see that as we increase the input size, the efficiency increases while slowly decreasing when the number of cores grows, indicating a tendency for a scalable program.
-In the second Figure \ref{fig:efficiency_openmc} the efficiency is practically constant while varying the number of cores and the input indicating a strong scalable program.
-\begin{figure}[H]
-	\begin{subfigure}[b]{0.46\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_raytrace.png}
-%		\caption{\centering}
-		\label{fig:efficiency_raytrace}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.46\textwidth}
-		\includegraphics[width=\textwidth]{pascalanalyzer/figures/results/efficiency_openmc.png}
-%		\caption{\centering}
-		\label{fig:efficiency_openmc}
-	\end{subfigure}
-
-	\caption{Efficiency diagram varying the input size and the number of cores.
-The color bar indicates the efficiency value in percentage.
-(\textbf{a}) Raytrace efficiency diagram.
-(\textbf{b}) OpenMC efficiency diagram.}
-	\label{fig:efficiency_all}
-\end{figure}
-
-The input values shown in Figures~\ref{fig:speedup}--\ref{fig:efficiency_all} indicate different data sets for processing.
-For example, the ``input 2'' represents a load that will require sequential processing with runtime twice as long as ``input 1''.
-Likewise, the ``input 3'' represents a load that will require sequential processing with runtime twice as long as ``input 2'', and so on.
-
-Even if the analyzer does not display graphics natively, analysis aimed at observing scalability and energy consumption of applications depends on precise measurements, reinforcing the advantage of the analyzer proposed in this work.
-
-% \begin{figure}[H]
-% \centering
-% \includegraphics[width=\linewidth]{pascalanalyzer/figures/results/efficiency_raytrace.png}
-% \caption{Efficiency for Raytrace}
-% \label{fig:efficiency_raytrace}
-% \end{figure}
-
-% \begin{figure}[H]
-% \centering
-% \includegraphics[width=\linewidth]{pascalanalyzer/figures/results/efficiency_raytrace_detailed.png}
-% \caption{Efficiency for Raytrace}
-% \label{fig:efficiency_raytrace_detailed}
-% \end{figure}
-
-% \begin{figure}[H]
-% \centering
-% \includegraphics[width=\linewidth]{pascalanalyzer/figures/results/efficiency_openmc.png}
-% \caption{Efficiency for OpenMC}
-% \label{fig:efficiency_openmc}
-% \end{figure}
-
-% \begin{figure}[H]
-% \centering
-% \includegraphics[width=\linewidth]{pascalanalyzer/figures/results/efficiency_openmc_detailed.png}
-% \caption{Efficiency for OpenMC}
-% \label{fig:efficiency_openmc_detailed}
-% \end{figure}
-\subsection{Fingerprint Tool validation}
-
-In this subsection, we first show the comparison between our tool and others already established, as well as the clustering process.
-
-\subsubsection{ACCURACY COMPARISON}
-
-% \begin{table*}[h]
-% \centering
-% \caption{Average}
-% \begin{tabular}{|c|c|c|c|c|c|c|}
-% \hline
-% Counter                                 & Pined values & Linux API   & PAPI      & PAPI Python & Perf tool   & MyPerf      \\ \hline
-% INSTRUCTIONS\_RETIRED                  & 226990030    & 227000691   & 227000620 & 225901249   & 227000572   & 227000650   \\ \hline
-% BRANCH\_INSTRUCTIONS\_RETIRED          & 9240000      & 9250617     & 9250566   & 9239617     & 9250552     & 9250501     \\ \hline
-% BR\_INST\_RETIRED:CONDITIONAL          & 8220000      & 8220000     & 8220000   & 8209717     & 8220000     & 8220000     \\ \hline
-% MEM\_UOP\_RETIRED:ANY\_LOADS           &              & 2484182672  &           &             & 2484383940  & 2484155029  \\ \hline
-% MEM\_UOP\_RETIRED:ANY\_STORES          &              & 189962002   &           &             & 189961539   & 189961321   \\ \hline
-% UOPS\_RETIRED:ANY                      &              & 12291082129 &           &             & 12290811038 & 12290901997 \\ \hline
-% PARTIAL\_RAT\_STALLS:MUL\_SINGLE\_UOP  &              & 600878      &           &             & 600151      & 600330      \\ \hline
-% ARITH:FPU\_DIV                         &              & 5801446     &           &             & 5801000     & 5800977     \\ \hline
-% FP\_COMP\_OPS\_EXE:X87                 &              & 48785528    &           &             & 48784834    & 48786021    \\ \hline
-% INST\_RETIRED:X87                      &              & 17200008    &           &             & 17200008    & 17200007    \\ \hline
-% FP\_COMP\_OPS\_EXE:SSE\_SCALAR\_DOUBLE &              & 5401694     &           &             & 5401842     & 5401679     \\ \hline
-% \end{tabular}
-% \label{tab:mean}
-% \end{table*}
-
-% \begin{table*}[h]
-% \centering
-% \caption{Standard deviation}
-% \begin{tabular}{|c|c|c|c|c|c|}
-% \hline
-% Counter                                    & Linux API & PAPI & PAPI Python & Perf tool & MyPerf \\ \hline
-% INSTRUCTIONS\_RETIRED                  & 396       & 133  & 337763      & 110       & 175    \\ \hline
-% BRANCH\_INSTRUCTIONS\_RETIRED          & 297       & 208  & 8485        & 379       & 91     \\ \hline
-% BR\_INST\_RETIRED:CONDITIONAL          & 0         & 0    & 3383        & 0         & 0      \\ \hline
-% MEM\_UOP\_RETIRED:ANY\_LOADS           & 37399     &      &             & 39217     & 38953  \\ \hline
-% MEM\_UOP\_RETIRED:ANY\_STORES          & 1513      &      &             & 1035      & 687    \\ \hline
-% UOPS\_RETIRED:ANY                      & 345246    &      &             & 335832    & 333298 \\ \hline
-% PARTIAL\_RAT\_STALLS:MUL\_SINGLE\_UOP  & 1222      &      &             & 252       & 521    \\ \hline
-% ARITH:FPU\_DIV                         & 1760      &      &             & 1621      & 1544   \\ \hline
-% FP\_COMP\_OPS\_EXE:X87                 & 1283      &      &             & 1920      & 3311   \\ \hline
-% INST\_RETIRED:X87                      & 4         &      &             & 4         & 3      \\ \hline
-% FP\_COMP\_OPS\_EXE:SSE\_SCALAR\_DOUBLE & 1547      &      &             & 3259      & 2097   \\ \hline
-% \end{tabular}
-% \label{tab:std}
-% \end{table*}
-
-\begin{table}[H]
-	\centering
-	\caption{Comparison}
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{llllllllll}
-			\hline
-			\multicolumn{6}{l|}{\textbf{Average*$10^{-6}$}} & \multicolumn{4}{l}{\textbf{Standard deviation}}\\ \hline
-			\textbf{Counters} & \begin{tabular}{l}\textbf{Pined} \\ \textbf{values}\end{tabular} & \begin{tabular}{l}\textbf{Linux} \\ \textbf{API}\end{tabular} & \textbf{PAPI} & \begin{tabular}{l}\textbf{PAPI} \\ \textbf{Python}\end{tabular} & \textbf{MyPerf}  & \begin{tabular}{|l}\textbf{Linux} \\ \textbf{API}\end{tabular} & \textbf{PAPI} & \begin{tabular}{l}\textbf{PAPI} \\ \textbf{Python}\end{tabular} & \textbf{MyPerf} \\ \hline
-			INSTRUCTIONS\_RETIRED                  & 226.99       & 227       & 227  & 225.9       & 227     & 396       & 133  & 337763      & 175    \\
-			BRANCH\_INSTRUCTIONS\_RETIRED          & 9.24         & 9.25      & 9.25 & 9.24        & 9.25    & 297       & 208  & 8485        & 91     \\
-			BR\_INST\_RETIRED:CONDITIONAL          & 8.22         & 8.22      & 8.22 & 8.21        & 8.22    & 0         & 0    & 3383        & 0      \\
-			MEM\_UOP\_RETIRED:ANY\_LOADS           &              & 2484.18   &      &             & 2484.16 & 37399     &      &             & 38953  \\
-			MEM\_UOP\_RETIRED:ANY\_STORES          &              & 189.96    &      &             & 189.96  & 1513      &      &             & 687    \\
-			UOPS\_RETIRED:ANY                      &              & 12291.08  &      &             & 12290.9 & 345246    &      &             & 333298 \\
-			PARTIAL\_RAT\_STALLS:MUL\_SINGLE\_UOP  &              & 0.6       &      &             & 0.6     & 1222      &      &             & 521    \\
-			ARITH:FPU\_DIV                         &              & 5.8       &      &             & 5.8     & 1760      &      &             & 1544   \\
-			FP\_COMP\_OPS\_EXE:X87                 &              & 48.79     &      &             & 48.79   & 1283      &      &             & 3311   \\
-			INST\_RETIRED:X87                      &              & 17.2      &      &             & 17.2    & 4         &      &             & 3      \\
-			FP\_COMP\_OPS\_EXE:SSE\_SCALAR\_DOUBLE &              & 5.4       &      &             & 5.4     & 1547      &      &             & 2097   \\ \hline
-		\end{tabular}
-	}
-	\label{tab:counters}
-\end{table}
-
-To validate the tool, we compared the results of the counters obtained with different APIs. 
-We used the hand-crafted assembly benchmark from \cite{Weaver2013Non-determinismImplementations}, designed to test determinism and accuracy of PMUs.
-We compared the values obtained from the Linux API, PAPI on C and PAPI on Python. 
-The events used for this comparison were instructions retired, branch instructions, memory read, memory load, and arithmetic operations.
-We ran the benchmark 30 times and calculated the mean and standard deviation as shown in table \ref{tab:counters}.
-Some events could not be measured using PAPI because the tool does not accept raw events and there are no equivalent events.
-
-Since the benchmark was hand-crafted with assembly, we know exactly the value for some counter events. 
-For this reason, the number of instructions, branch instructions, and conditional branch are pin. However, some other events are architecture-specific and there is no pined value. 
-In the latter case, we can still compare to the Linux API, which should be closest to the reality.
-
-The differences using the Linux low-level API, PAPI, and our tool are negligible (the average percentage distance is less than 0.01\% in all the cases). 
-As expected, PAPI on Python had the largest difference (with an average distance of 0.25\%) mainly due to an unsynchronized start that resulted in the loss of some instructions at the beginning of the execution.
-This can be an important problem if the application contains a small number of instructions.
-
-The standard deviation of the 30 executions shows that our tool has the smallest variation on a run-to-run on most events. On the contrary, PAPI on Python shows a big variation compared to the others.
-
-\subsubsection{CLUSTERING}
-
-To cluster applications first we need to have a way to compare two programs, for that we define a new variable that tries to compute a fingerprint to the program, this variable has to look similar when we execute the same program with different conditions and inputs.
-Thinking in the simplest model of a program as a Turing machine as everything can be done with a tap of memory and set of rules we empirically define the variable input size as described on the equation \ref{eq:input_size}.
-On real computers this is not too far from reality, most of the computations and input and output operations somehow pass through the memory, so analyzing the relationship of the total number of instructions and memory instructions can give us a good fingerprint.
-
-\begin{equation}
-	I_{sz} = \frac{I}{I_{m}} \\
-	\label{eq:input_size}
-\end{equation}
-
-Where $I_{sz}$ we called input size, $I$ the number of instructions executed, $I_m$ number of memory instructions execute.
-We observe that this variable demonstrate to have the proprieties that we are looking for, produce similar results to the same program with different input size, and environment.
-This can be used to identify programs but also to see the similarities between different applications.
-
-To compute the distance between two programs we use the Canberra metric \cite{Jurman2009CanberraLists}, described on the equation \ref{eq:canberra}.
-
-\begin{equation}
-	d(p,q)= \sum_{i=1}^{n}{\frac{|p_i-q_i|}{|p_i|+|q_i|}}
-	\label{eq:canberra}
-\end{equation}
-
-Where $p$ and $q$ are n-dimensional vectors.
-
-We compute the input size for all 30 applications of the Polybench with 3 different inputs.
-Running each program 15 times, with a sampling rate of 0.01 seconds collecting the total number of instruction, number of loads and write instructions, number of floating point operations, besides some software counters.
-After applying the post-processing to the data collected we compute the input size and perform a Hierarchical clustering using the linkage method of Ward \cite{Murtagh2011WardsAlgorithm}, that minimizes the total within-cluster variance.
-The results of the clustering can be seen on the figure \ref{fig:sping_force_is} and on the dendrogram on figure \ref{fig:dendograma_input_size}.
-
-%The exact HPCs was:
-%PERF_COUNT_HW_INSTRUCTIONS,
-%MEM_UOPS_RETIRED:ALL_LOADS,
-%MEM_UOPS_RETIRED:ALL_STORES,
-%FP_ARITH_INST_RETIRED:SCALAR
-
-\begin{figure*}[h]
-	\centering
-	\includegraphics[width=\textwidth]{fingerprint/figures/dendograma_input_size.png}
-	\caption{Dendrogram}
-	\label{fig:dendograma_input_size}
-\end{figure*}
-
-From this dendrogram, we can have an idea of how close two applications are.
-We choose the number of clusters that maximize the number of hits of the same program with different input sizes on the same cluster, in this case, was 5 clusters.
-
-It was observed that as input size grows the program behavior tends to a specific curve, but for small input sizes some have variation, so in some cases, the same program has been classified in more than one cluster.
-This can also happen if specific parts of the program are triggered with specific inputs, in which case it will also belong to more than one cluster.
-
-In order to have an overall classification of each program, we can pick up the frequency in which each appeared in the clusters and classified it in the cluster in which it appeared more often.
-In this case, the clusters are:
-
-\begin{itemize}
-	\item Cluster 1: 2mm, 3mm, cholesky, correlation, covariance, floyd-warshall, gemm, gramschmidt, lu, ludcmp, nussinov, symm
-
-	\item Cluster 2: deriche, doitgen, syrk
-
-	\item Cluster 3: adi, fdtd-2d, jacobi-2d, syr2k
-
-	\item Cluster 4: atax, bicg, durbin, gemver, gesummv, mvt, trisolv, trmm
-
-	\item Cluster 5: heat-3d, seidel-2d
-\end{itemize}
-
-On the figure, \ref{fig:sping_force_is} we display the clusters using the spring force algorithm where each program with a specific input is a node and the edge weight is the Canberra distance.
-To a better visualization, the name of the inputs was replaced by numbers, the EXTRALARGE is 3, LARGE 2 and MEDIUM 1.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\textwidth]{fingerprint/figures/graph_input_size.png}
-	\caption{Spring force of Canberra distance}
-	\label{fig:sping_force_is}
-\end{figure}
-
-The figure above shows a different way to observe in a reduced space the distance between clusters and applications and how they are organized.
-From this graph, we see that clustering it's well partitioned and we can clearly separate each cluster.
-It is also interesting to note that the applications of the cluster formed by the circle symbol are more separated, which may indicate that they were classified in this way because they did not fit into any other cluster.
-
-% \begin{figure*}[t]
-%     \begin{subfigure}{\textwidth}
-%         \includegraphics[width=\textwidth]{fingerprint/figures/dendograma_input_size.png}
-%     \end{subfigure}
-%     \\
-%     \begin{subfigure}{\textwidth}
-%         \includegraphics[width=\textwidth]{fingerprint/figures/dendograma_floating.png}
-%     \end{subfigure}
-%     %and so on
-% \end{figure*}
-
-To have an idea of how the behavior of the variable input size for each clusters is, it was plotted for the applications of clusters 1 and 2 shown on the figures \ref{fig:c1_input_size_0}, \ref{fig:c1_input_size_1}.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\textwidth]{fingerprint/figures/cluster_input_0.png}
-	\caption{Input size - Cluster 1}
-	\label{fig:c1_input_size_0}
-\end{figure}
-
-From these figures, we can have an idea of what behavior was classified as the same class.
-On cluster 1 on the figure \ref{fig:c1_input_size_0} all applications showed the same behavior with approximately the same amplitude.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\textwidth]{fingerprint/figures/cluster_input_1.png}
-	\caption{Input size - Cluster 2}
-	\label{fig:c1_input_size_1}
-\end{figure}
-
-On figure \ref{fig:c1_input_size_1} we observe that curves that have similar shape regardless of scale on vertical and horizontal axis have also been classified as the same clusters.
-This is the wanted comportment for the classification because we are only interested in the overall shape of the curve.
-
-% \begin{figure}[H]
-%     \centering
-%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_input_4.png}
-%     \caption{Input size - Cluster 3}
-%     \label{fig:c1_input_size_2}
-% \end{figure}
-
-% To conclude this study we also clustering applications by its floating point behavior.
-For that we only use the counter correspondent to floating point operations.
-The figure \ref{fig:dendogram_fp} shows the dendrogram and the \ref{fig:sping_force_fp} the spring force graph plot.
-
-% \begin{figure*}[h]
-%     \includegraphics[width=\textwidth]{fingerprint/figuresdendograma_floating.png}
-%     \caption{Dendrogram}
-%     \label{fig:dendogram_fp}
-% \end{figure*}
-
-% \begin{figure}[H]
-%     \centering
-%     \includegraphics[width=\textwidth]{fingerprint/figures/graph_floating2.png}
-%     \caption{Spring force of Canberra distance}
-%     \label{fig:sping_force_fp}
-% \end{figure}
-
-% The figures \ref{fig:cluster_fp1},\ref{fig:cluster_fp2},\ref{fig:cluster_fp3} show the behavior on the same cluster for the float point operations.
-For this classification the number of clusters was 8.
-
-% \begin{figure}[H]
-%     \centering
-%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_fp_0.png}
-%     \caption{Floating point - Cluster 1}
-%     \label{fig:cluster_fp1}
-% \end{figure}
-
-% \begin{figure}[H]
-%     \centering
-%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_fp_4.png}
-%     \caption{Floating point - Cluster 2}
-%     \label{fig:cluster_fp2}
-% \end{figure}
-
-% \begin{figure}[H]
-%     \centering
-%     \includegraphics[width=\textwidth]{fingerprint/figures/cluster_fp_5.png}
-%     \caption{Floating point - Cluster 3}
-%     \label{fig:cluster_fp3}
-% \end{figure}
-
-% Here we can also observed again that applications with the same shape were classified as the same clusters.

### 2022-05-03 thesis/main.tex
carlos revision 1
change subsection to section and labels
fix section titles

@@
 
-	\chapter{Proposed Models} \label{chapter:models}
+	\chapter{Application Models} \label{chapter:models}
 	In this chapter, the proposed power, performance and energy models are derived and validated.
@@
 	\include{models/models}
+	%\include{optmizers/optmizers}
+
+	\chapter{Energy optimization using the models}
+	\include{experiments/setup}
+	\include{optmization/optmization}
+
+	\chapter{Extending the models with phases}
 	\include{phases/phases}
-	%\include{optmizers/optmizers}
 
@@
 
-	\chapter{Experiments and Results}
 	This chapter will describe all the experiments that validate the models constraints.
-	\include{experiments/experiments}
+	\include{framework/validation}
 

### 2022-05-03 thesis/optmization/optmization.tex
fix section titles

--- 2022-05-03 thesis/optmization.tex
@@
-\section{Model Validation} \label{sec:model_validation}
+\section{Model validation} \label{sec:model_validation}
 In this section, the models presented in~Sections \ref{sec:power_model} and~\ref{sec:performance_model} were validated with a benchmark specific for multi-core architectures.
@@
 
-\section{Fitting the Models} \label{subsec:fitting_the_models}
+\section{Fitting the models} \label{subsec:fitting_the_models}
 To find the parameters of  \cref{eq:en_final}, 10 uniformly random configurations of frequencies ($f$), cores ($p$) and inputs ($N$) were chosen from the range $1<=p<=32$, $1.2<=f<=2.2$ and $1<=N<=5$, respectively.
@@
 
-\section{Measured versus Modeled Energy} \label{sec:measured_versus_modeled_energy}
+\section{Measured versus modeled energy} \label{sec:measured_versus_modeled_energy}
 
@@
 
-\section{Overheads on Training} \label{subsec:overhead_on_training}
+\section{Overheads on training} \label{subsec:overhead_on_training}
 It is known that machine learning is data-driven; in that sense, the SVR model obtained using only 10 configurations could be improved, but what about the analytical model? 
@@
 
-\section{DVFS and DPM Optimization} \label{sec:dvfs_dpm_optminzation}
+\section{DVFS and DPM optimization} \label{sec:dvfs_dpm_optminzation}
 The effectiveness of the proposed approach during optimization was evaluated with a simple algorithm that finds the optimal frequency and number of active cores from the proposed equation.

### 2022-05-03 thesis/pascalanalyzer/pascalanalyzer.tex
fix section titles

@@
-\section{Pascal Analyzer} \label{sec:pascal_analyzer}
+\section{PascalAnalyzer} \label{sec:pascal_analyzer}
 High-performance computing systems have become increasingly dynamic, complex, and unpredictable.
@@
 
-\section{Pascal architecture} \label{sec:pascal_architecture}
+\section{PascalAnalyzer architecture} \label{sec:pascal_architecture}
 The tool we propose to measure, combine and compare multiple runs of a parallel application implements two main concepts to achieve this goal: actuators and sensors.

### 2022-05-03 thesis/phases/phases.tex
fix section titles

@@
 
-\section{Related work: Phases} \label{sec:related_work_phases}
+\section{Phases related work} \label{sec:related_work_phases}
 Some related work
@@
 
-\section{Data Collection} \label{sec:data_collection}
+\section{Data collection} \label{sec:data_collection}
 Data, power samples, can be gathered by running the application in several possible context configurations.
@@
 
-\section{Results: Phases experiments} \label{sec:results_phases_experiments}
+\section{Phases experiments results} \label{sec:results_phases_experiments}
 

### 2022-05-03 thesis/phases/phases.tex
phases

@@
 
-The energy efficiency of HPC can be achieved through dynamic workload planning.
-Dynamic scaling of frequency and voltage DVFS is the main energy efficiency technique to save power.
-The scaling algorithms can operate dynamically depending on the CPU workload.
-However, the required IT resources and workload vary throughout the applications' execution.
-The division of this workload directly impacts the resulting performance and the complexity of dynamic scaling algorithms.
-This work analyses the impact on the energy efficiency of different dynamic workload partitioning techniques, based on time division of allocated resources and DVFS.
-We evaluate the trade-off between the best time intervals partition and the overhead impact of dynamic scheduling techniques.
-We propose the optimal partition of application executions in time intervals providing the most energy efficient execution parameters, workload, frequency and number of cores, for each time interval.
+Dynamic frequency scaling and dynamic power management are the leading software techniques for energy saving in modern computers.
+These techniques work through dynamic workload planning, and switching from one power state to another is crucial for them.
+This work analyses the impact on the energy efficiency of different dynamic workload partitioning techniques based on the time-division and allocated resources.
+We propose a methodology combining measurement data with a simulation to provide insights into choosing the best time intervals.
+Furthermore, we evaluate the trade-off between the best time intervals partition and the overhead impact of dynamic scheduling techniques identifying that generally exists a limit to the number of divisions that the algorithm can benefit from, giving a lower bound to the minimal time interval.
 
 \section{Introduction to models on time} \label{sec:introduction_to_models_on_time}
-Talk about the importance of energy efficiency.
+Energy efficiency in computing has always been a concern, especially in high-performance computing (HPC), where a small percentage in energy savings can significantly reduce maintenance costs or environmental impact due to its scale of consumption.
+For example, the leading Petaflop supercomputers consume are in a range of 1–18 MW of electrical power, with 1.5 MW on average, which we can estimate to be into a millions dollars range of electricity bills annually \cite{Group2012HandbookSahni}.
+Also,  the International Energy Agency  \cite{iea_2021} estimated that global data center electricity use in 2020 was around 200-250 TWh, or around 1\% of the global electricity demand, which can generate as much pollution as a nation like Argentina \cite{Mathew2012Energy-awareNetworks}.
+
+Given the importance of this topic, modern processors of desktop, mobile phones, and HPC  have internal power management mechanisms, and that is because the processor is one of the main components that drain energy and can reach 50\% of the total system consumption \cite{Fan2007, Barroso2007TheComputing, Malladi2012TowardsDRAM}.
+Within the mechanisms implemented by the processors, the main controllable via software are Dynamic Frequency and Voltage Scaling (DVFS) and Dynamic Power Management (DPM) ~\cite{Rotem2012Power-managementBridge, Brown2005, Hackenberg2015}.
+
+DPM encompasses a set of techniques that take advantage of different power levels on the system, such as active, idle, and disabled~\cite{CARDOSO201717, Shuja2012Energy-efficientCenters, Benini2000AManagement}.
+The basic idea is to shut down devices and wake them up when necessary.
+However, this is not trivial, that are several variables that need to take into account as the switching cost, in terms of time and energy, for how long it will remain in that state,  the power consumption on every state, and a lot more.
+
+The DPM technique is quite interesting on systems where static power is very high or where the system is idle for a long time.
+In these cases, some works report savings of up to 70\% in energy~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.
+
+DVFS allows real-time control of the frequency and voltage depending on the needs.
+This method is motivated by the fact that we can approximate the relationship of the frequency with the power as a cubic equation and frequency and performance as a linear equation \cite{Dayarathna2016DataSurvey, Group2012HandbookSahni}, implying that a reduction in frequency has a cubic impact on power and linear on performance.
+Moreover, reducing the frequency leads to a longer execution time, increasing energy consumption.
+As a result, determining the best voltage and frequency to employ in all conditions is a challenging task.
 

### 2022-06-20 thesis/main.tex
samuel suggestions

@@
 	This chapter will...
+	Describe the context
 	\include{intro/intro}
 
-	\chapter{Application Models} \label{chapter:models}
+	\chapter{Related work*}
+	Move all related work to here
+	Write a small introduction, and group similar works.
+	Derive some conclusion about that.
+
+	\chapter{Theoretical background*}
+
+	\chapter{Parallel Scalability Suite*}
+	This chapter introduces the framework developed, the fingerprint tool and pascal analyzer.
+
+	\include{pascalanalyzer/pascalanalyzer}
+	\include{fingerprint/fingerprint}
+
+	This chapter will describe all the experiments that validate the models constraints.
+	\include{framework/validation}
+
+	\chapter{Application energy and performance models} \label{chapter:models}
 	In this chapter, the proposed power, performance and energy models are derived and validated.
@@
 
-	\chapter{Energy optimization using the models}
+	\chapter{Energy optimization using the models*}
 	\include{experiments/setup}
@@
 
-	\chapter{Extending the models with phases}
+	\chapter{Application-phase energy and performance models}
 	\include{phases/phases}
-
-	\chapter{Framework}
-	This chapter introduces the framework developed, the fingerprint tool and pascal analyzer.
-
-	\include{pascalanalyzer/pascalanalyzer}
-	\include{fingerprint/fingerprint}
-
-	This chapter will describe all the experiments that validate the models constraints.
-	\include{framework/validation}
 

### 2022-08-11 thesis/conclusion.tex
samuel suggestions
update phases chapter

@@
+\section{Contributions*} \label{sec:contributions}
+
 \section{Conclusion Models} \label{sec:conclusion_models}
@@
 
-\section{Conclusion and future work: Phases}
-The main conclusion of this work was that in HPC enviroment we don't need a lot of phases to optimize the energy consumption given that for one phase we can compute the optimal energy.
+\section{Conclusion and future work phases} \label{sec:conclusion}
+The main conclusion of this work was that in the HPC environment, many phases are not necessary to reach optimal energy consumption.
+However, an average of 35 phases was enough for more than three benchmarks covering the most varied HPC applications.
 
-The final idea is to use the information of the best phase to find a performance metric that can provide us this division, in future work we intend to compute this metrics using combination of hardware performance counters enabling optimal phase division in real time.
-This can improve current DVFS algorithms giving them an workload metric that can improve the energy savings.
+It was also possible to notice that there is still a lot to improve in optimizing the DVFS algorithms since we obtained an average of 38\% energy savings compared to Ondemand on Linux.
+
+For future work, the final idea is to use the information from the best phase to find a performance metric that can provide us with this division without needing data.
+In future work, we intend to calculate these metrics using a combination of hardware performance counters that allow real-time optimal phase division.
+This metric can improve current DVFS algorithms by providing them with a workload metric that can improve energy savings.

### 2022-08-11 thesis/main.tex
title thesis
update phases chapter

@@
 
-\title{Title}
+\title{Energy-optimal configurations for high-performance computing applications: automated low-impact characterization and performance optimization of shared-memory applications}
 %\title{My very long two-line PhD thesis title}
@@
 
-	\chapter{Related work*}
-	Move all related work to here
-	Write a small introduction, and group similar works.
-	Derive some conclusion about that.
+%	\chapter{Related work*}
+%	Move all related work to here
+%	Write a small introduction, and group similar works.
+%	Derive some conclusion about that.
 
 	\chapter{Theoretical background*}
+	This chapter will..
+	\include{experiments/setup}
 
@@
 
-	\chapter{Energy optimization using the models*}
-	\include{experiments/setup}
+	%\chapter{Energy optimization using the models*}
 	\include{optmization/optmization}
@@
 	\chapter{Application-phase energy and performance models}
+	Dynamic voltage and frequency scaling (DVFS) and dynamic power management (DPM) are essential software techniques for saving energy in modern computers.
+These techniques control hardware resources such as frequency and number of active processors.
+Although there is much work on the subject, there is an underexplored sub-topic: the impact of the choice of phase division on energy consumption.
+This work analyzes the impact of phase division on the energy efficiency of algorithms by proposing a methodology that combines measurement data with a heuristic to provide insights into choosing the best phase divisions.
+Our heuristic can reduce the scan space from $10^{7000}$ to $10^2$ with an average error of 10\% and up to 38\% reduction in energy consumption using optimal split compared to standard Linux DVFS.
+In addition, we evaluated the trade-off of having too many divisions and the overhead caused.
+Finally, we have identified that there is usually a limit to the number of phases the algorithm can benefit from, giving a lower limit to the minimum number of phases.
 	\include{phases/phases}

### 2022-08-11 thesis/phases/phases.tex
update phases chapter

--- 2022-05-03 thesis/phases/phases.tex
@@
-\section{Model with phases} \label{sec:model_with_phases}
+\section{Introduction} \label{sec:introduction}
 
-The energy efficiency of HPC can be achieved through dynamic workload planning.
-Dynamic scaling of frequency and voltage DVFS is the main energy efficiency technique to save power.
-The scaling algorithms can operate dynamically depending on the CPU workload.
-However, the required IT resources and workload vary throughout the applications' execution.
-The division of this workload directly impacts the resulting performance and the complexity of dynamic scaling algorithms.
-This work analyses the impact on the energy efficiency of different dynamic workload partitioning techniques, based on time division of allocated resources and DVFS.
-We evaluate the trade-off between the best time intervals partition and the overhead impact of dynamic scheduling techniques.
-We propose the optimal partition of application executions in time intervals providing the most energy efficient execution parameters, workload, frequency and number of cores, for each time interval.
+Energy efficiency has been one of the main focuses in computing nowadays, both on mobile and HPC.
+It is essential to increase battery life and reduce heat generation, especially in high-performance computing (HPC), where a small percentage of the savings energy can significantly reduce maintenance costs or environmental impact due to its scale of consumption.
+For example, the leading Petaflop supercomputers consume in the range of 1 to 18 MW of electrical energy, with 1.5 MW on average, which we can estimate in millions of dollars in annual electricity cost \cite{Group2012HandbookSahni}.
+Furthermore, the International Energy Agency \cite{iea_2021} estimated that global data center electricity usage in 2020 was around 200-250 TWh, or around 1\% \cite{Corcoran2017EmergingICT} of global electricity demand, which could generate as much pollution as a nation like Argentina \cite{Mathew2012Energy-awareNetworks}.
 
-\section{Introduction to models on time} \label{sec:introduction_to_models_on_time}
-Talk about the importance of energy efficiency.
+Given the importance of this topic, modern desktops, cell phones, and HPC CPUs have built-in power management mechanisms.
+That is because the processor is one of the main components that drain energy and can reach 50\% of the total system consumption \cite{Fan2007, Barroso2007TheComputing, Malladi2012TowardsDRAM}.
+Among the mechanisms implemented by the CPU, the most impactful and controllable via software are Dynamic Frequency and Voltage Scaling (DVFS), and Dynamic Power Management (DPM) ~\cite{Rotem2012Power-managementBridge, Brown2005, Hackenberg2015}.
 
-\section{Phases related work} \label{sec:related_work_phases}
-Some related work
+DPM encompasses a set of techniques that take advantage of different energy levels in the system, such as active, inactive, and disabled~\cite{CARDOSO201717, Shuja2012Energy-efficientCenters, Benini2000AManagement}.
+The basic idea is to turn off devices and turn them on when needed.
+However, it is not a trivial task, and several variables need to be considered, such as the cost of switching, in terms of time and energy, how long it will remain in that state, the energy consumption in each state, and much more.
 
-\section{Optimal energy configuration} \label{sec:Optimal_energy_configuration}
-Splitting a program into phases, or time intervals with a selected set of execution parameters, has an impact on energy consumption.
-To find the ideal phase division, we need to compute the optimal energy configuration for a given number and duration of phases.
-For that, many different approaches could be applied, each having different trade-offs.
+The DPM technique has the most significant gains when in systems where the static power is very high or where the system is idle for a long time.
+In such cases, some papers report savings of up to 70\% in energy~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.
 
-Using models ref [x], or a static analysis of the application, allows obtaining a very fast but less precise estimate.
-On the other hand, using the application and evaluating its behavior in each possible context would be a more reliable but too expensive method.
-Indeed, since the number of combinations grows exponentially, this is a very time consuming approach.
-Instead, our analysis approach will provide both, speed and precision.
-We will simply run the target application and collect performance data from every possible execution context.
-By context, we can consider, for instance, the frequency and/or number of cores.
-This data will then be used to analyse the behavior of phases in all possible combinations.
-For this, we assume that the number of instructions of the executed program is the same regardless of frequency and number of threads.
+DVFS, on the other hand, allows real-time frequency and voltage control, depending on needs.
+This method is motivated by the fact that we can approximate the relationship of frequency to power as a cubic equation and frequency and performance as a linear equation \cite{Dayarathna2016DataSurvey, Group2012HandbookSahni} implying that a reduction in frequency has a cubic impact on power and linear performance.
+Still, it is not so simple because lowering the frequency leads to longer runtime, increasing power consumption.
+As a result, determining the best voltage and frequency to employ under all conditions is also a challenging task.
 
+Although there are many studies on this topic, most works focus on implementing different optimization algorithms based on workload, and few studies pay attention to phase division.
+Instead, most studies use static phase division to simplify the optimization math and compensate by implementing lightweight algorithms.
+However, this leads to a sub-optimization, as another phase division could result in different frequency values.
+In addition, there is an overhead associated with the moment the algorithm takes action that could be reduced if we knew the ideal moment for it to take action, making it possible even to allow heavier algorithms.
+Therefore, this work proposes a methodology to study the division of time (phases) and analyzing possible gains.
+
+\section{Related work} \label{sec:related_work}
+
+One of the bases works in scheduling algorithms is the work of Irani et al.
+\cite{Irani2007}, where they formalized the problem of scheduling arriving jobs in a way that minimizes total energy use and so that each job is completed after its arrival time and before its deadline.
+Although each problem has been considered separately, this was the first theoretical analysis of systems that can use both mechanisms.
+
+Past work has also used feedback-based approaches, as in Poellabauer et al.
+\cite{Poellabauer2005}, where applications' past CPU utilization's are used to predict future  CPU  requirements.
+However, mispredictions in these approaches can lead to missed deadlines,  sub-optimal energy savings,  or large overheads due to frequent changes to the chosen frequency or voltage.
+One shortcoming of previous approaches is that they ignore other 'indicators' of future CPU requirements, such as the frequency of I/O operations, memory accesses, or interrupts.
+
+In the thesis work of Saha et al.
+\cite{Saha2012}, the timeliness and power consumption behavior of fourteen RT-DVFS schedulers through implementation and actual measurements.
+The schedulers include Static Earliest Deadline First (Static-EDF), Cycle Conserving Earliest Deadline First(CC-EDF), Look-Ahead Earliest Deadline First (LA-EDF), Snowdon-minimum (Snowdon-min), Resource-constrained Energy-Efficient Utility Accrual Algorithm (REUA), DynamicReclaiming Algorithm (DRA) and Aggressive Speed Reduction Algorithm(AGR) among the others.
+They draw attention to the fact that most of these algorithms are based on simulations, which can mislead some results.
+Broadly, RT-DVFS techniques have two objectives: (i) Reduce energy consumption through DVFS; and (ii) Optimize task timeliness behavior through real-time resource management (deadline).
+
+Pietri et al.
+\cite{Pietri2014} used slack reclamation to achieve energy savings.
+The goal is to exploit the idle slots of the processors that occurred due to the earlier completion time of the tasks compared with the latest finish time, which is constrained by the deadline or data dependencies.
+In their model, the user submits a workflow for execution specifying a deadline for completion of the execution.
+
+Mashayekhy et al.
+\cite{Mashayekhy2014} proposed a greedy algorithm, called Energy-aware  MapReduce  Scheduling  Algorithm  (EMRSA), that finds the assignments of the map and reduces tasks to the machine slots in order to minimize the energy consumed when executing the application.
+They consider an extensive data application consisting of a set of map and reduce tasks that need to be completed by deadline D.
+
+Yousefi et al.
+\cite{Yousefi2018}  present a task-based greedy scheduling algorithm, TGSAVE.
+This algorithm selects a slot for each task to minimize the total energy consumption of the MapReduce job for big data applications in heterogeneous environments without significant performance loss while satisfying the service level agreement (SLA).
+TGSAVE  finds solutions under deadlines up to 74\% tighter than the tightest one feasible by EMRSA, and it can meet deadlines as tight as only12\%, on average, more significant than the energy-oblivious minimum makespan.
+
+Crown scheduling is a static scheduling approach for sets of parallelizable tasks with a common deadline.
+Crown schedules are robust, i.e., the runtime prolongation of one task by a moderate percentage does not cause a deadline transgression by the same fraction.
+In addition, by speeding up some tasks scheduled after the lengthy task, the deadline can still be met at moderate additional energy consumption.
+In the work of Kessler et al.
+\cite{Kessler2021} they present a heuristic to perform this re-scaling online and explore the trade-off between additional energy consumption in normal execution and limitation of deadline transgression in delay cases.
+
+%Typically target traditional real-time systems with static deadlines, resulting in conservative energy savings that cannot exploit additional energy optimizations due to dynamic deadlines.
+However, in the work of Yi et al.
+\cite{Yi2021} they present an adaptive system optimization and reconfiguration approach that dynamically adapts the scheduling parameters and processor speeds to satisfy dynamic deadlines while consuming as little energy as possible.
+As their work was for autonomous driving, they used velocities as deadlines.
+
+In the work of \cite{Ajmal2021}, a green cloud computing algorithm named "Cost-based Energy Efficient Scheduling Technique for Dynamic Voltage Frequency Scaling (DVFS) Systems (CEEST)" is proposed.
+The proposed algorithm reduces energy consumption without compromising the quality of service (QoS).
+This algorithm aims to optimize and manage servers in the datacenters by utilizing maximum resources and powering off the underutilized servers.
+Furthermore, CEEST utilizes the scaling of virtual machines to finish jobs within the deadlines to reduce violations of service level agreement (SLA).
+
+All of these works at some point considered a deadline (defined by the user of the system) that defines the phases of the job; this means that choosing a different deadline could result in completed different energy savings, and because this is not the main focus of these works, this is not considered.
+In our paper, we study the impact of this phase division on the application's energy consumption.
+We estimate how much energy could be saved if an algorithm could magically give the ideal deadline.
+
+Although most of the work is considered a deadline, some works include job division as part of the optimization problem.
+For example, in the paper of Agrawal et al.
+\cite{Agrawal2021}, they show that if the jobs can be divided into arbitrary parts, the minimum-energy schedule can be generated in linear time, giving exact scheduling algorithms.
+For the cases where jobs are non-divisible, they claimed a prove that the scheduling problems are NP-hard and also give approximation algorithms for the same along with their bounds.
+
+In this case, where no deadline is defined, the phase division is indirectly considered by solving the optimization problem of minimizing the energy.
+The problem, in this case, is that several constraints restrict the optimization problem while our approach works with measured data, meaning that we can find solutions independent of the constraints.
+
+\section{Phase division approach} \label{sec:application_partition_method}
+To find the optimal phase division, we would have to ensure that the one we choose is the most energy-efficient among all possible divisions and different combinations of power configurations.
+
+This mathematical problem can be modeled in many ways, but always with some concessions to make it solvable.
+Another option would be by brute force, but there are infinite possibilities for splitting phases, although some hardware limitations make our analysis more accessible.
+
+With that in mind, we decided to create a heuristic that comes the closest to brute force using hardware limitations to make it viable.
+
+The first limitation is that the processor has a limited speed at which it can act.
+In other words, a division that can act would make no sense, limiting our analysis to discrete-time intervals at the processor's maximum performance speed.
+
+This already limits our exploration space a lot, but even so, it is still unfeasible.
+For example, considering $d_t$ as the minimum processor action interval, $T$ the total application time, and $C$ the number of power settings.
+The number of possibilities could be estimated by:
 \begin{equation}
-    I=c_{k1}f_1T_1=c_{k2}f_2T_2
+	\left(\frac{T}{d_t}\right)^C,
 \end{equation}
+This number grows astronomically because $d_t$ is in a time range of microseconds, $T$ seconds or minutes, and C of hundreds.
+To give an idea, we can estimate this number to be in a range of $10^7000$.
 
+Another thing that we can see is that each configuration gives the power profile for a given application, and phase division is just a combination of these configurations.
+This allows us to reduce further our exploration of space.
+For example, if we ran each power setting once and we had a way of combining them, our problem now would be feasible.
+
+To make this possible, we assume that the configuration change only affects the program speed and that the power consumption is independent of the previous settings.
+
+\iffalse %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% commented 
+This means that in a certain part of the program $x_1$\% to $x_2$\% it will always execute the same instructions and the energy consumption would depend only on the configuration.
+
+That said, if we treat the program as a percentage of execution, and we have the power consumption profile for all configurations, we can estimate the power consumption for any combination of phases.
+
+For example if we have the phase division, 0-x_1\%, x_1\%-100\%, with the settings c_1 and c_2, the estimated energy would be:
+P_{c_1}*u(0,x_1)*T_{c_1}+P_{c_2}*u(x_1, 100)*T_{c_2}
+
+We can see that we just need a power profile P and the total runtime for each configuration and that we can divide the application into as many phases as we want.
+
+For that we are going to assume that the number of instructions to complete a program is roughly the same independent on the frequency, thus we can write the number of instructions as:
+\begin{equation}
+	I=c_{k1}f_1T_1=c_{k2}f_2T_2,
+\end{equation}
 where $I$ is the number of instructions, $c_{k1}$ the rate of instructions per cycle, $f$ the frequency of the processor and $T$ the execution time.
@@
 The extent of the hypotheses will later be validated, however, by taking advantage of both principles, we can now, based on real measurements, estimate any phase division and possible context combination.
+\fi %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Here a need to show that the time is proportional to the number of executed instructions.
-% WHY ?
-% I am trying to say here that we can split the program using a percentage of the time because the time is proportional to the number of executed instructions.
-If we complete x% of the time, that will be x% of instructions.
-This implies that we can resume and continue the program from any point ensuring that we executed all instructions.
-If we have a previous execution for each configuration, we can compute any combination with minimal error.
+This approach aims to bring the best of both worlds.
+We want simulation speed but with actual measurement data for greater accuracy.
+Our algorithm runs with measured data from a real system.
+The idea is to run the application with all possible configurations (e.g., frequency and number of active cores) of the machine without time division and estimate power consumption for different phase divisions.
 
-\section{Data collection} \label{sec:data_collection}
-Data, power samples, can be gathered by running the application in several possible context configurations.
-In a system with 32 cores that can run at 13 frequencies, we can have 416 (32*13) possible configurations.
-Power samples can be taken from dedicated sensors in the IPMI system [REF].
-To avoid overload, a sample rate of 0.5 seconds is more than sufficient.
-Additionally, the behavior of an application can be affected by the workload, so it is also interesting to run applications with different inputs and assess their impact on the collected data [REF victor alex paper].
+\section{Building applications power profile} \label{sec:building_a_database}
 
-% We can consider different input sizes as an new application...
+%\subsection{Case-Study Architecture} \label{subsec:casestudyarchitecture}
+%We executed the experiments in a system equipped with two Intel Xeon E5-2698 v3 processors with sixteen cores and two hardware threads for each core, with a total physical memory of the node is 128GB (8$\times$16GB).
+We disabled turbo frequency and hardware multi-threading during all experiments.
+The operating system used was Linux CentOS 6.5 and kernel 4.16.
+The overall view of the architecture is shown in \cref{fig:architecture}.
+%
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=0.8\columnwidth]{models/figures/architecture.png}
+%	\caption{Node architecture (the image was made with the lstop application).}
+%	\label{fig:architecture}
+%\end{figure}
+%
+%To perform the frequency control, we used the acpi-cpufreq driver using the Userspace governor, which allows the user or any userspace program to set the CPU to a specific frequency.
+While the cores control was accomplished by modifying the appropriate system files with the default CPU-hotplug driver.
+%
+%The architecture is also equipped with the Intelligent Platform Management Interface (IPMI), a set of interfaces allowing out-of-band management of computer systems and platform-status monitoring via the local network~\cite{Schwenkler2006IntelligentInterface}.
+In addition, it can monitor variables and resources such as the system's temperature, voltage, fans, and power supplies, with independent sensors attached to the hardware.
+%
+%\subsection{Case-Study Applications} \label{subsec:casestudyapplication}
+%The applications Black-Scholes, Bodytrack, Canneal, Dedup, Fluidanimate, Freqmine, Raytrace, Swaptions, Vips and x264 from the PARSEC \footnote{\url{https://parsec.cs.princeton.edu/download.htm}} parallel benchmark suite, version 3.0~\cite{Bienia2008TheSuite}, OpenMC \cite{Romano2015OpenMC:Development} and LINPACK (HPL) \cite{Dongarra1988TheExplanation}, were chosen as case studies.
+The PARSEC benchmark focused on emerging workloads and was designed to represent the next-generation shared-memory programs for chip-multiprocessors.
+It covers many areas such as financial analysis, computer vision, engineering, enterprise storage, animation, similarity search, data mining, machine learning, and media processing.
+The OpenMC and the LINPACK are two classic HPC programs.
 
-\section{The simulator} \label{sec:the_simulator}
-To analyse the impact of configurations and identify optimal time-intervals, phases, we propose an integrator, which computes the energy on an interval for a given configuration, and a reference, a brute force test of all possible combinations.
-The integrator has order zero, and computers the energy between two timestamps, as shown in the figure \ref{fig:zero_order}, where the red and yellow parts area are the energy value.
+\subsection{Data gathering} \label{sec:data_gathering}
+
+Using the Pascal framework \cite{electronics11050689} the data was collected by running the application in all possible configurations.
+In this system, that is 32 cores and 13 frequencies, a total of 416 (32*13) possible configurations.
+Power samples were taken from dedicated sensors in the IPMI system with a sample rate of 0.5 seconds.
+
+%\subsection{Application fingerprint}
+%\textcolor{red}{Introduce the power profile as fingerprint briefly...}
+
+\section{The phase optimization algorithm} \label{sec:heuristic}
+To analyze the impact of configurations and identify the optimal phase division, we propose a zero-order integrator, which calculates the energy in an interval for a given configuration.
+The integrator uses the data previously described in \cref{sec:data_gathering}.
+
+Figure \ref{fig:zero_order} shows an example energy computation for an application with four different power configurations.
+For that, first, the phases are defined in terms of the percentage of execution.
+Then, the integrator will run over each phase using the selected profile of power obtained from the measurements and accumulate the total energy consumption.
 
@@
 	\centering
-    \includegraphics[width=\columnwidth]{phases/figures/zero_order.png}
-    \caption{Power vs time intervals for a given application. The energy associated area is colored.}
+	\includegraphics[width=\columnwidth]{phases/figures/integrator.pdf}
+	\caption{Power vs percentage of execution for a given application in several configurations of power. The energy is represented by the hatched area.}
 	\label{fig:zero_order}
@@
 
-Power and energy can be computed for a given time interval based on power samples. The pseudo-algorithm for this computation is described bellow:
+The pseudo-algorithm for the integrator is described below:
 
-\begin{lstlisting}
-timeslice_energy(info, start, stop)
+\begin{lstlisting}[language=c++]
+	integrator(total_time, nsamples,
+	times, powers
+	begin_phase, end_phase)
 	{
-   power = info.power;
-   time = info.time;
+		energy = 0;
+		ti = total_time * begin_phase; // initial time
+		tf = total_time * end_phase; // final time
 
-   t1 = start_time + (elapsed_time) * start;
-   t2 = start_time + (elapsed_time) * stop;
-   en = 0;
+		idx_i = 0, idx_f = 0;
 
-   // first time that is greater than t1
-   for(i=0; i<nsamples; i++)
-       if(time[i] >= t1)
+		// find the index of first sample in the range
+		for (idx_i= 0; idx_i<nsamples; idx_i++)
+		if (times[idx_i] >= ti)
 		break;
-   v1 = (i == 0)?0:i-1;
-   // first time that is greater than t2
-   for(i=v1; i<nsamples; i++)
-       if(time[i] >= t2)
+		idx_i = (idx_i != 0) ? idx_i - 1 : 0;
+
+		// find the index of last sample in the range
+		for (idx_f= idx_i; i<nsamples; idx_f++)
+		if (times[idx_f] >= tf)
 		break;
-   v2 = (i == 0)?0:i-1;
+		idx_f = (idx_f != 0) ? idx_f - 1 : 0;
 
-   // interpolate and integrates the power samples
-   if(v1 == v2)
-       return (t2-t1)*power[v1];
+		// in case there is only one sample
+		if (idx_i == idx_f) 
+		return (tf-ti)*powers[idx_i];
 
-   en = (time[v1+1]-t1)*power[v1];
-   en += (t2-time[v2])*power[v2];
-   // integrate samples in between
-   for(i=v1+1; i<v2; i++)
-       en += (time[i+1]-time[i])*power[i];
-   return en;
+		// handle the edges
+		energy=(times[idx_i+1]-ti)*powers[idx_i]
+		energy+=(tf-times[idx_f])*powers[idx_f];
+
+		// handle sample in between
+		for (int i= idx_i+1; i<idx_f; i++)
+		energy += (times[i+1]-times[i])*powers[i];
+
+		return energy;
 	}
@@
 
-The reference can simply be obtained by applying brute-force, which cycles through all configurations for all divisions and compute the minimal energy.
-Other more effective, but less accurate, methods can be applied.
+To obtain an optimal energy configuration for a given phase division, we can run all possible configurations and select the one that leads to a minimal consumption, as shown in the pseudo-algorithm below:
 
 \begin{lstlisting}
-bruteforce(data, phases)
+	phase_optimization(configs, phases)
+	{
+		total_en = 0;
+		for (i = 0; i<phases.size(); i++)
 		{
 			min_en = -1;
-   total_en = 0;
-   for (i = 0; i < nphases; i++)
+			for (j = 0; j<configs.size(); j++)
 			{
-       min_en = -1;
-       for (j = 0; j<data.nconfigs; j++)
-       {
-           en = timeslice_energy(data.info[j], 
+				en = estimate_energy_phase(
+				configs[j].total_time, 
+				configs[j].nsamples,
+				configs[j].time_samples,
+				configs[j].power_samples,
 				phases[i], phases[i + 1]);
@@
 					min_en = en;
-               min_conf = data.config[j];
+					min_conf = configs[j];
 				}
@@
 
-\section{Optimizing phases} \label{sec:optimizing_phases}
-The main goal of this section is to identify the optimal number and duration of time intervals.
-Indeed, several short phases can have an impact on switching overhead as well as few long phases on the overall energy.
+This method can now obtain an optimal energy configuration for a given phase. The next step is to find the best phase, division.
 
-Using the objective function, brute force can identify the optimal energy on each phase and any minimization method can provide the best phase division.
-By computing in a range of a number of phases, we can find the one that brings the lowest energy value.
-A genetic algorithm GA can be used to analyse a range of, for instance, 3 to 100 phases.
-A GA for a population of 10K individuals with a 0.1 elitism will ensure that 10\% of the best individuals will survive and 0.1 mutations will avoid local minimums within a maximum of 300 steps.
-A reproduction function will combine the half-phases of one individual with another, and the mutation will randomly change a division.
+\section{Results} \label{sec:results}
 
-\section{Phases experiments results} \label{sec:results_phases_experiments}
+Now that we can evaluate phase divisions, we need to answer two questions, what is the best number of phases, and what is the best division within this number?
 
-% Normalization
-% Show that number of instructions is the same...
+For that, we used the algorithm described previously in \cref{sec:heuristic} as the objective of a minimization problem where we want to find the vector of phases that minimizes the total energy.
+Moreover, by using a genetic algorithm, we can accelerate the process by controlling mutation and initial population generation since we already know some information about our problem.
 
-In order to validate the approach, HPC benchmarks where chosen, 13 applications from PARSEC 3.0, HPC, and Openmc.
+For instance, we decide to start studying phases in the range of 3 to 100 divisions.
+We used a GA with a population of $10^3$ individuals.
+We kept the best 10\% individuals on each generation while having a 10\% of mutation rate, running for 300 generations.
+The reproduction function for the GA combines half-phases of one individual with another, and the mutation will randomly change a division.
+
+We use the applications described in \cref{subsec:casestudyapplication}, to compute the optimal phase selection.
+The \cref{fig:relative_energy} shows the relative energy consumption compared to an optimized single phase.
 
 \begin{figure}[H]
-\includegraphics[width=\columnwidth]{fingerprint/figures/energy_per_phase.png}
-    \caption{Relative energy vs number of phases using 13 applications from PARSEC 3.0, HPC, and Openmc benchmarks. Energy compared to the single-phase optimal configuration.}
+	\includegraphics[width=\columnwidth]{fingerprint/figures/energy_per_phase.pdf}
+	\caption{Relative energy vs number of phases using applications from PARSEC 3.0, HPC, and Openmc benchmarks with different inputs. Energy compared to the single-phase optimal configuration.}
 	\label{fig:relative_energy}
@@
 
-This evaluation shows very interesting results, as shown in the figure \ref{fig:relative_energy}.
-Indeed, for all cases, an optimal single-phase configuration gives the highest energy draw.
-On the contrary, it is not necessary to split the application into more than 35 phases.
-On average, the energy consumption stood still after 35 phases.
+This evaluation shows very interesting results, as shown in \cref{fig:relative_energy}.
+An ideal single-phase configuration provides the highest energy consumption for all cases.
+Furthermore, dividing the application into more than 35 phases is not necessary.
+On average, power consumption stopped after 35 phases.
 
-The results attend our expectation that the optimal with n divisions will always be better or equal than n-1 divisions.
-This can be explained by the fact that after we find the ideal frequency and core curves, increasing the number of phases will not change the control signals.
-Therefore the energy will remain the same, and there is no reason to increase the number of phases beyond that actually.
+% \begin{figure}[H]
+% \includegraphics[width=\columnwidth]{figures/min_phases_distribution.pdf}
+%     \caption{Histogram of minimal number of phases to optimal energy}
+%     \label{fig:hist_optimal_phase}
+% \end{figure}
 
-It is important to note that the additional energy savings that theoretically may be achieved when running with a greater number of phases will have a negative impact on performance and context switching.
-Moreover, the maximal energy saving compared to the single-phase case was just 2.6\%.
-This may be due to the type of applications, where a fine-grain fast control will not save more energy.
-On the contrary, this switching will cause to consume a lot more energy, and this is likely to happen with online DVFS algorithms, very sensitive to fast workload variations [REF].
+The results meet our expectation that the optimum of n-divisions phases will always be better than or equal to n-1 divisions.
+This becomes clearer when we look at the number of cores, and frequency curve \cref{fig:freq_control} and \cref{fig:cores_control}, since starting from a certain number of divisions does not change the control signals since dividing an ideal phase into two would result in two phases with the same control signals.
+Therefore, the energy will remain the same, and there is no reason to increase the number of phases beyond that.
 
-We evaluated separately the impact on energy when varying the number of cores.
-Energy-cores figures doesn't change fundamentally varying the number of phases.
-Indeed, bellow we can see some examples with optimal energy cores selections for 60 and 70 phases (figure \ref{fig:cores_control}).
-Note the very small differences that will not change significantly the energy figures.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_1_freq_signals_cmp.pdf}
+	\caption{Frequency signal vs percentage of execution for the Bodytrack application, comparing the difference of 35 and 99 phases.}
+\end{figure}%
 
-\begin{figure}[h!]
-\includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_3_cores_signals_cmp.png}
-\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_3_cores_signals_cmp.png}
-%\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_cores_signals_cmp.png}  
-\caption{Energy and number of cores with 37 and 99 phases.}
-    \label{fig:cores_control}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\columnwidth]{phases/figures/signals/completo_freq_freq_signals_cmp.pdf}
+	\caption{Frequency signal vs percentage of execution for the Freqmine application, comparing the difference of 35 and 99 phases.}
+\end{figure}%
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_1_cores_signals_cmp.pdf}
+	\caption{Cores signal vs percentage of execution for the Bodytrack application, comparing the difference of 35 and 99 phases.}
+\end{figure}%
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\columnwidth]{phases/figures/signals/completo_freq_cores_signals_cmp.pdf}
+	\caption{Cores signal vs percentage of execution for the Freqmine application, comparing the difference of 35 and 99 phases.}
+\end{figure}%
+
+When we compare this method with the default DVFS algorithm in Linux, we observer an average saving of 38\% as shown in \cref{fig:cmp_ondemand} the relative energy per application.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\columnwidth]{phases/figures/comparison_ondemand.pdf}
+	\caption{Relative energy comparison, optimal phase division energy compared to governor Ondemand on Linux.}
+	\label{fig:cmp_ondemand}
 \end{figure}
 
-We observed a similar behavior just considering frequency control energy optimisation (figure \ref{fig:freq_control}).
+To illustrate the phases division, the \cref{fig:phase_division_cmap_35} and \cref{fig:phase_division_cmap_35} shows a heatmap with the divisions by application and the percentage of total energy spent in that phase.
+There we can see a pattern of 3 phases which for these applications are generally setup phase where it loads data from the disk, a computation phase where the data is processed, and a finalization phase where data is written back to the disk or displayed to the user.
 
-\begin{figure}[h!]
-\includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_3_freq_signals_cmp.png}
-\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_3_freq_signals_cmp.png}
-%\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_freq_signals_cmp.png}
-\caption{Energy and frequency with 37 and 99 phases.}
-    \label{fig:freq_control}
+\begin{figure}[H]
+	\includegraphics[width=\columnwidth]{phases/figures/phase_division_cmap_3.pdf}
+	\caption{Phase division heatmap (3 divisions) showing the energy consumption per phase for all applications.}
+	\label{fig:phase_division_cmap_3}
 \end{figure}
 
-The images below show the phases for all applications:
 \begin{figure}[H]
-\includegraphics[width=\columnwidth]{phases/figures/phase_division_35.png}
+	\includegraphics[width=\columnwidth]{phases/figures/phase_division_cmap_35.pdf}
+	\caption{Phase division heatmap (35 divisions) showing the energy consumption per phase for all applications.}
+	\label{fig:phase_division_cmap_35}
 \end{figure}
-In most applications, we can see a pattern of 3 phases wich for this applications are generally, and setup phase where it loads data from the disk, a computation phase where the data is processed, and a finalization phase where data is written back to the disk or display to the user.
-
-\begin{figure}[H]
-\includegraphics[width=\columnwidth]{phases/figures/phase_division_3.png}
-\end{figure}
-Plotting for 3 phases, we can observe that better.

### 2022-08-11 thesis/references.tex
fix references

@@
 
-	%%%%%%%%%%%%%%
+	%%%%%%%%%%%
 
@@
 
+	%%%%%%%%%%% PHASES %%%%%%%%%%%
+
+	\bibitem{Group2012HandbookSahni}
+	I.~Ahmad and S.~Ranka, \emph{Handbook of Energy-Aware and Green
+		Computing}.\hskip 1em plus 0.5em minus 0.4em\relax Chapman \& Hall/CRC, 2012.
+
+	\bibitem{iea_2021}
+	Iea, ``Data centres and data transmission networks – analysis,'' Nov 2021.
+	[Online]. Available:
+	\url{https://www.iea.org/reports/data-centres-and-data-transmission-networks}
+
+	\bibitem{Corcoran2017EmergingICT}
+	P.~Corcoran and A.~Andrae, ``{Emerging Trends in Electricity Consumption for
+		Consumer ICT},'' no. July 2013, pp. 1--56, 2017.
+
+	\bibitem{Mathew2012Energy-awareNetworks}
+	V.~Mathew, R.~K. Sitaraman, and P.~Shenoy, ``{Energy-aware load balancing in
+		content delivery networks},'' \emph{Proceedings - IEEE INFOCOM}, no.
+	September 2011, pp. 954--962, 2012.
+
+	\bibitem{Fan2007}
+	X.~Fan, W.-D. Weber, and L.~A. Barr¨oso, ``{Power provisioning for a
+		warehouse-sized computer},'' \emph{ACM SIGARCH Computer Architecture News},
+	vol.~35, no.~2, p.~13, 2007.
+
+	\bibitem{Barroso2007TheComputing}
+	L.~A. Barroso and U.~H{\"{o}}lzle, ``{The case for energy-proportional
+		computing},'' \emph{Computer}, vol.~40, no.~12, pp. 33--37, 2007.
+
+	\bibitem{Malladi2012TowardsDRAM}
+	K.~T. Malladi, F.~A. Nothaft, K.~Periyathambi, B.~C. Lee, C.~Kozyrakis, and
+	M.~Horowitz, ``{Towards energy-proportional datacenter memory with mobile
+		DRAM},'' \emph{Proceedings - International Symposium on Computer
+		Architecture}, no. Figure 1, pp. 37--48, 2012.
+
+	\bibitem{Rotem2012Power-managementBridge}
+	E.~Rotem, A.~Naveh, A.~Ananthakrishnan, E.~Weissmann, and D.~Rajwan,
+	``{Power-management architecture of the intel microarchitecture code-named
+		Sandy Bridge},'' \emph{IEEE Micro}, vol.~32, no.~2, pp. 20--27, 2012.
+
+	\bibitem{Brown2005}
+	L.~Brown, R.~Moore, D.~S. Li, L.~Yu, A.~Keshavamurthy, and V.~Pallipadi,
+	``{ACPI in Linux},'' \emph{Symposium A Quarterly Journal In Modern Foreign
+		Literatures}, vol.~51, p.~51, 2005.
+
+	\bibitem{Hackenberg2015}
+	D.~Hackenberg, R.~Sch{\"{o}}ne, T.~Ilsche, D.~Molka, J.~Schuchart, and
+	R.~Geyer, ``{An Energy Efficiency Feature Survey of the Intel Haswell
+		Processor},'' \emph{Proceedings - 2015 IEEE 29th International Parallel and
+		Distributed Processing Symposium Workshops, IPDPSW 2015}, pp. 896--904, 2015.
+
+	\bibitem{CARDOSO201717}
+	J.~M. Cardoso, J.~G.~F. Coutinho, and P.~C. Diniz, ``Chapter 2 -
+	high-performance embedded computing,'' in \emph{Embedded Computing for High
+		Performance}, J.~M. Cardoso, J.~G.~F. Coutinho, and P.~C. Diniz, Eds.\hskip
+	1em plus 0.5em minus 0.4em\relax Boston: Morgan Kaufmann, 2017, pp. 17--56.
+	[Online]. Available:
+	\url{https://www.sciencedirect.com/science/article/pii/B9780128041895000028}
+
+	\bibitem{Shuja2012Energy-efficientCenters}
+	J.~Shuja, S.~A. Madani, K.~Bilal, K.~Hayat, S.~U. Khan, and S.~Sarwar,
+	``{Energy-efficient data centers},'' \emph{Computing}, vol.~94, no.~12, pp.
+	973--994, 2012.
+
+	\bibitem{Benini2000AManagement}
+	L.~Benini, A.~Bogliolo, and G.~De~Micheli, ``{A survey of design techniques for
+		system-level dynamic power management},'' \emph{IEEE Transactions on Very
+		Large Scale Integration (VLSI) Systems}, vol.~8, no.~3, pp. 299--316, 2000.
+
+	\bibitem{Dayarathna2016DataSurvey}
+	M.~Dayarathna, Y.~Wen, and R.~Fan, ``{Data center energy consumption modeling:
+		A survey},'' \emph{IEEE Communications Surveys and Tutorials}, vol.~18,
+	no.~1, pp. 732--794, 2016.
+
+	\bibitem{Irani2007}
+	S.~Irani, S.~Shukla, and R.~Gupta, ``Algorithms for power savings,'' \emph{ACM
+		Transactions on Algorithms}, vol.~3, p.~41, 11 2007. [Online]. Available:
+	\url{https://dl.acm.org/doi/10.1145/1290672.1290678}
+
+	\bibitem{Poellabauer2005}
+	C.~Poellabauer, L.~Singleton, and K.~Schwan, ``Feedback-based dynamic voltage
+	and frequency scaling for memory-bound real-time applications,'' 2005, pp.
+	234--243.
+
+	\bibitem{Saha2012}
+	S.~Saha and B.~Ravindran, ``An experimental evaluation of real-time dvfs
+	scheduling algorithms.''\hskip 1em plus 0.5em minus 0.4em\relax ACM Press,
+	2012, pp. 1--12. [Online]. Available:
+	\url{http://dl.acm.org/citation.cfm?doid=2367589.2367604}
+
+	\bibitem{Pietri2014}
+	I.~Pietri and R.~Sakellariou, ``Energy-aware workflow scheduling using
+	frequency scaling.''\hskip 1em plus 0.5em minus 0.4em\relax IEEE, 9 2014, pp.
+	104--113. [Online]. Available:
+	\url{http://ieeexplore.ieee.org/document/7103444/}
+
+	\bibitem{Mashayekhy2014}
+	L.~Mashayekhy, M.~M. Nejad, D.~Grosu, D.~Lu, and W.~Shi, ``Energy-aware
+	scheduling of mapreduce jobs.''\hskip 1em plus 0.5em minus 0.4em\relax
+	Institute of Electrical and Electronics Engineers Inc., 9 2014, pp. 32--39.
+
+	\bibitem{Yousefi2018}
+	M.~H.~N. Yousefi and M.~Goudarzi, ``A task-based greedy scheduling algorithm
+	for minimizing energy of mapreduce jobs,'' \emph{Journal of Grid Computing},
+	vol.~16, pp. 535--551, 12 2018.
+
+	\bibitem{Kessler2021}
+	C.~Kessler, S.~Litzinger, and J.~Keller, ``Crown-scheduling of sets of
+	parallelizable tasks for robustness and energy-elasticity on many-core
+	systems with discrete dynamic voltage and frequency scaling,'' \emph{Journal
+		of Systems Architecture}, vol. 115, p. 101999, 5 2021. [Online]. Available:
+	\url{https://linkinghub.elsevier.com/retrieve/pii/S1383762121000175}
+
+	\bibitem{Ajmal2021}
+	M.~S. Ajmal, Z.~Iqbal, F.~Z. Khan, M.~Bilal, and R.~M. Mehmood, ``Cost-based
+	energy efficient scheduling technique for dynamic voltage and frequency
+	scaling system in cloud computing,'' \emph{Sustainable Energy Technologies
+		and Assessments}, vol.~45, p. 101210, 6 2021. [Online]. Available:
+	\url{https://linkinghub.elsevier.com/retrieve/pii/S2213138821002204}
+
+	\bibitem{Agrawal2021}
+	P.~Agrawal and S.~Rao, ``Energy-efficient scheduling: classification, bounds,
+	and algorithms,'' \emph{Sādhanā}, vol.~46, p.~46, 12 2021. [Online].
+	Available: \url{http://link.springer.com/10.1007/s12046-021-01564-w}
+
+	\bibitem{Schwenkler2006IntelligentInterface}
+	T.~Schwenkler, ``{Intelligent Platform Management Interface},'' pp. 169--207,
+	2006.
+
+	\bibitem{Bienia2008TheSuite}
+	C.~Bienia, S.~Kumar, J.~P. Singh, and K.~Li, ``{The PARSEC benchmark suite},''
+	in \emph{Proceedings of the 17th international conference on Parallel
+		architectures and compilation techniques - PACT '08}, no. January.\hskip 1em
+	plus 0.5em minus 0.4em\relax New York, New York, USA: ACM Press, 2008, p.~72.
+	[Online]. Available:
+	\url{http://portal.acm.org/citation.cfm?doid=1454115.1454128}
+
+	\bibitem{Romano2015OpenMC:Development}
+	P.~K. Romano, N.~E. Horelik, B.~R. Herman, A.~G. Nelson, B.~Forget, and
+	K.~Smith, ``{OpenMC: A state-of-the-art Monte Carlo code for research and
+		development},'' \emph{Annals of Nuclear Energy}, vol.~82, pp. 90--97, 8 2015.
+	[Online]. Available:
+	\url{https://linkinghub.elsevier.com/retrieve/pii/S030645491400379X}
+
+	\bibitem{Dongarra1988TheExplanation}
+	J.~J. Dongarra, ``{The LINPACK Benchmark: An explanation},'' in \emph{Lecture
+		Notes in Computer Science (including subseries Lecture Notes in Artificial
+		Intelligence and Lecture Notes in Bioinformatics)}, 1988, vol. 297 LNCS, pp.
+	456--474. [Online]. Available:
+	\url{http://link.springer.com/10.1007/3-540-18991-2_27}
+
+	\bibitem{electronics11050689}
+	V.~R.~G. da~Silva, A.~B.~N. da~Silva, C.~Valderrama, P.~Manneback, and
+	S.~Xavier-de Souza, ``A minimally intrusive approach for automatic assessment
+	of parallel performance scalability of shared-memory hpc applications,''
+	\emph{Electronics}, vol.~11, no.~5, 2022. [Online]. Available:
+	\url{https://www.mdpi.com/2079-9292/11/5/689}
+
 \end{thebibliography}

### 2022-08-15 thesis/conclusion.tex
organazing the chapters and trying to connect

@@
-\section{Contributions*} \label{sec:contributions}
+\section{Conclusion}
 
-\section{Conclusion Models} \label{sec:conclusion_models}
-This paper proposes an energy model based on the operating frequency and the number of cores for a shared memory system. This model serves as a reference for DVFS and DPM optimization problems.
+\subsection{Pascal Suite} \label{sec:conclusions_pascal}
+
+This work introduces a practical and easy-to-use tool for measuring and analyzing the efficiency of parallel applications.
+The proposed tool focuses on observing scalability and energy consumption and implements features that enables analysis at hierarchical levels of the program's inner parts.
+It also simplifies the comparison of the application runs in different configurations, helping developers target software optimization efforts.
+
+The tool has a low level of intrusion added to the program's performance measurement under analysis, which is a fundamental aspect of understanding the program's behavior and scalability capacity. 
+Although it does not offer graphic elements to visualize the collected metrics, the tool itself makes it possible to examine the data from the command terminal.
+It also allows this data to be analyzed by an interface provided in Python or through .json files.
+
+\textls[-20]{In future work, we intend to evolve this tool to include the ability to predict speedup and efficiency from a few samples using state-of-the-art prediction models present in the \mbox{literature \cite{Alex2020WhenParallelSpeedups, Vitor2022AnalyticalEnergyModel}}.
+The idea is to present the general behavior of the program and its scalability trend and reduce the execution time necessary to compose a comprehensive analysis.
+Furthermore, we intend to include features that allow the observation of parallel applications in distributed environments that use the Message Passing Interface (MPI) standard.}
+
+\subsubsection{fingerprint tool}
+
+From the results, we observe that the API present overhead similar or lower to other low-level APIs, with the advantage of being in high abstraction and simplified configuration with a few lines of code, is possible to configure and gather counter data.
+
+The tool developed provide also provided a way to fingerprint programs and compute similarities between different programs or the same program with different inputs.
+This can be useful to reduce applications spaces for benchmarks as was done in Polybench clustering but also to analyze the behavior of a parameter providing insights to the programmer to find a possible bottleneck.
+
+We also provide a precise definition to input size that made possible fingerprint the programs, but it also can help programmers of benchmark applications to better create inputs with more precise growth of a particular parameter.
+
+\subsection{Application model} \label{sec:conclusion_models}
+The proposed energy model based on the operating frequency and the number of cores for a shared memory system can serve as a reference for DVFS and DPM optimization problems.
 
@@
 
-%Future work will demonstrate all possible analyzes achievable with the proposed equation as more advanced DVFS models can be derived from it.
-For instance, identifying the different phases of a target program will allow more subtle changes in frequency and, perhaps, in the number of active cores to improve further the results presented here.
-
-This work proposes an energy model based on the frequency and number of cores for a full shared memory system.
-This model can serve as a base for DVFS and DPM optimization problems that include both frequency and active cores, as well for analysis of the contribution of each parameter (ex: parallelism level) to the energy consumption.
+The proposed energy model based on the frequency and number of cores for a full shared memory system can serve as a base for DVFS and DPM optimization problems that include both frequency and active cores.
+As well for analysis of the contribution of each parameter (ex: parallelism level) to the energy consumption.
 
@@
 
-\subsection{Conclusions and Future Works Pascal} \label{sec:conclusions_pascal}
-
-This work introduces a practical and easy-to-use tool for measuring and analyzing the efficiency of parallel applications.
-The proposed tool focuses on observing scalability and energy consumption and implements features that enables analysis at hierarchical levels of the program's inner parts.
-It also simplifies the comparison of the application runs in different configurations, helping developers target software optimization efforts.
-
-The tool has a low level of intrusion added to the program's performance measurement under analysis, which is a fundamental aspect of understanding the program's behavior and scalability capacity. 
-Although it does not offer graphic elements to visualize the collected metrics, the tool itself makes it possible to examine the data from the command terminal.
-It also allows this data to be analyzed by an interface provided in Python or through .json files.
-%Other tools, such as PaScal Viewer, can compensate for the absence of these graphic resources.
-Indeed, PaScal Viewer is a web-based visualization tool focused on analyzing the scalability trend of parallel applications, which has been integrated with PaScal Analyzer in a complementary way.
-
-\textls[-20]{In future work, we intend to evolve this tool to include the ability to predict speedup and efficiency from a few samples using state-of-the-art prediction models present in the \mbox{literature \cite{Alex2020WhenParallelSpeedups, Vitor2022AnalyticalEnergyModel}}.
-The idea is to present the general behavior of the program and its scalability trend and reduce the execution time necessary to compose a comprehensive analysis.
-Furthermore, we intend to include features that allow the observation of parallel applications in distributed environments that use the Message Passing Interface (MPI) standard.}
-
-\subsection{Conclusion fingerprint tool}
-
-%\cite{Processors2012, Gregg2017}
-From the results, we observe that the API present overhead similar or lower to other low-level APIs, with the advantage of being in high abstraction and simplified configuration with a few lines of code, is possible to configure and gather counter data.
-
-The tool developed provide also provided a way to fingerprint programs and compute similarities between different programs or the same program with different inputs.
-This can be useful to reduce applications spaces for benchmarks as was done in Polybench clustering but also to analyze the behavior of a parameter providing insights to the programmer to find a possible bottleneck.
-
-We also provide a precise definition to input size that made possible fingerprint the programs, but it also can help programmers of benchmark applications to better create inputs with more precise growth of a particular parameter.
-
-%\subsection{FUTURE WORK}
-%
-%We pretend to use this tool to create a data set of applications behavior and automatically classify any program as belonging to a cluster or a set of clusters.
-The idea is to have a set of clusters that can describe most applications in this way we can know specific behaviors of the applications.
-This can be applied to various areas such as voltage scaling and frequency of the processor, once knowing the behavior of a particular program we can have an idea of what is the best strategy to control the frequency in order to save energy or increase performance.
-
-\section{Conclusion and future work phases} \label{sec:conclusion}
-The main conclusion of this work was that in the HPC environment, many phases are not necessary to reach optimal energy consumption.
-However, an average of 35 phases was enough for more than three benchmarks covering the most varied HPC applications.
-
+\subsection{Application-phase heuristic} \label{sec:phases_conclusion}
+The main conclusion of this work was that in the HPC environment, in addition to the advantages of finding the optimal phase divisions, not many phases are needed to achieve the optimal energy consumption.
+Moreover, an average maximum of 35 phases was sufficient for the three benchmarks used, covering the most varied HPC applications.
 It was also possible to notice that there is still a lot to improve in optimizing the DVFS algorithms since we obtained an average of 38\% energy savings compared to Ondemand on Linux.
+We also noticed that there is a relation between the application behavior and the phases locations, independently of their number.
 

### 2022-08-15 thesis/intro/intro.tex
change subsection to section and labels
samuel suggestions
organazing the chapters and trying to connect

@@
 
-\section{Objectives}
+\section{Objectives} \label{sec:objectives}
+This thesis aims to propose solutions that use the full potential of the systems involved, based on knowledge of the architectures and the applications, as follow:
 
-\section{Contributions}
+Develop a framework that helps and automates the process of modeling applications and architecture without interfering with the application.
 
-Although much work has been done on DVFS, the focus is still on the consumer electronics and laptop markets. 
-For HPC, the notion of energy perception is relatively new~\cite{Feng2003MakingSupercomputing}. 
-Moreover, the operational characteristics of non-HPC and HPC systems are significantly different. 
-First, the workload on non-HPC systems is very interactive with the end-user, but the workload on the HPC platform is not.
-Second, activities conducted on a non-HPC platform tend to share more machine resources.
-In contrast, in HPC, each job often runs with dedicated resources. 
-Third, an HPC system is usually much larger than a non-HPC system, making it more challenging to gather information, organize, and execute global decisions.
-Therefore, it is worthwhile to investigate whether a DVFS scheduling algorithm, which works well for conventional computing, remains effective for HPC.
+Propose an analytical energy model for energy consumption in high-performance computing systems applications that both optimize energy consumption in these systems and perform deeper analysis using analytical equations.
 
-Our paper proposes a full-system energy model based on the CPU frequency and the number of cores.
+Propose an heuristic algorithm, taking into account changes in the behavior of applications and the systems they run.
+
+\section{Contributions} \label{sec:contributions}
+
+In the \cref{chapter:pascal_suite}, we present a tool for automatically measuring and comparing several executions of a parallel application according to various scenarios characterized by the input arrangements, the number of threads, number of cores, and frequencies.
+Unlike other existing performance analysis tools, the proposed work covers some gaps in specialized features necessary to better understand computational resources scalability trends across configurations while keeping minimal intrusion (less than 1\% overhead).
+
+The \cref{chapter:models}, proposes a full-system energy model based on the CPU frequency and the number of cores.
 The model aims to understand and optimize the energy behavior of parallel applications in HPC systems according to application parameters, such as the degree of parallelism and CPU parameters related to dynamic and static power.
 The proposed model differs from existing ones, including the frequency and number of cores in the same equation for estimating the energy for a specific application in a given configuration.
-This model can serve as a base for considering DVFS and DPM optimization problems, including frequency and active cores.
-It can also be used to analyze the contribution of each parameter (ex: level of parallelism) to energy consumption.
-Furthermore, the number of cores is essential in HPC since applications are designed to run on multiple cores.
+This model can serve as a base for DVFS and DPM optimization problems.
+Including frequency and active cores it can also be used to analyze the contribution of each parameter (ex: level of parallelism) to energy consumption.
 
-The proposed energy model is the product of an application-agnostic power model and an architecture-specific application performance model.
-The power model is based on the CMOS logic gates power draw as a function of the frequency~\cite{Sarwar1997CmosCalculation, Butzen2007LeakageGates} augmented to include the number of cores.
-The performance model is based on Amdahl's law \cite{Amdahl1967ValidityCapabilities, Eyerman2010ModelingDesign, Gustafson1988ReevaluatingLaw}, which can be used to estimate runtime in multi-core systems.
-In addition, this model has been extended to include execution frequency and input size, characterizing the application on the target architecture.
-
-\cref{tab:related_work} summarizes the models comparing the system dependencies and the controllable variables.
-\begin{table}[H]
-	\footnotesize
-	\caption{Related work summary.\label{tab1}}
-	\setlength{\tabcolsep}{2.3mm}\begin{tabular}{cccc}
-		\toprule
-		%\begin{adjustwidth}{-\extralength}{0 cm}
-
-		\textbf{Model}  & \textbf{System Dependency}       & \textbf{Variable}                     & \textbf{Controllable Variables}         \\ \midrule
-		Merkel et al. \cite{Merkel2006BalancingSystems} & performance counters    & number of activities & - \\
-		Roy et al. \cite{Roy2013AnAlgorithms}&performance counters &io operations, total time    & - \\
-		Yakun et al. \cite{Shao2013EnergyProcessor}&number of instructions &frequency&frequency \\
-		Lewis et al. \cite{Lewis2008Run-timeSystems}&energy of subcomponents&energy of subcomponents & - \\
-		Mills et al. \cite{Mills2014EnergySystems}&power of subcomponents&total time, frequency & frequency \\ 
-		Our model & - & frequency, cores, input size & frequency, cores, input size\\ \bottomrule
-	\end{tabular}
-	%\caption{Related work summary}
-	\label{tab:related_work}
-	%\bottomrule
-\end{table}
-
-The main contributions of the proposed model are:
-
-% Contribuições por capitulo...
-\begin{itemize}
-	\item Simple model: faster to fit and compute, good for DVFS and DPM optimization.
-	\item Parameters with logical meaning: helps to understand the contribution of each specific term.
-	\item Analytical analysis: several analyses can be derived from the equation.
-	\item Controllable variables: the equation is in the function of parameters that we can control directly.
-\end{itemize}
+\cref{chapter:phases} proposes a methodology that combines measurement data with a heuristic to provide insights into choosing the best phase divisions.
+Our heuristic can reduce the scan space from $10^{7000}$ to $10^2$ with an average error of 10\% and up to 38\% reduction in energy consumption using optimal distribution compared to standard Linux DVFS.
+Moreover, we evaluated the trade-off of having too many divisions and the overhead caused.
+Finally, we have identified that there is usually a limit to the number of phases a running application can benefit from, giving a lower limit to the minimum number of phases.
 
 \section{Organization}
+
+In the \cref{chapter:theoretical_background} points that are common to the whole thesis are clarified, such as definitions, configurations and benchmarks used.
+In the next chapter \cref{chapter:pascal_suite} the PascalSuite framework is presented in detail and validated.
+The energy model is proposed in sequence in \cref{chapter:models}.
+Finally a heuristic algorithm is presented in \cref{chapter:phases} followed by the conclusion taken from this thesis in \cref{chapter:conclusions}.

### 2022-08-15 thesis/main.tex
Updates from Overleaf
organazing the chapters and trying to connect

--- 2022-05-03 thesis/main.tex
@@
 	\begin{umonsThesisAbstract}
-		Energy consumption is a key to enable exascale High-performance Computing (HPC). Energy-optimized hardware and software combinations still could be inefficient if software operates poorly. 
-		Software operation relies on dynamic scaling of frequency and voltage (DVFS) as well as dynamic power management (DPM), but none have priority information on the application, which leads to inefficient software operation.
-This work proposes an analytical model that can be used to estimate energy-optimal software operation.
-		This work proposes a new energy model that can be used as a basis to optimize DVFS and DPM, as well as to assess the contribution of parameters such as parallelism level and dynamic power to the energy consumption.
-		Since in HPC, applications can run on multiple cores and associated frequencies, both of which can be controlled for most systems, the novelty of the model includes such parameters as well as the workload, two application parameters and three system parameters.
+		Energy consumption is a key to enabling exascale High performance Computing (HPC).
+However, energy-optimized hardware and software combinations could still be inefficient if the software operates poorly.
+
+		Software operation relies on dynamic scaling of frequency and voltage (DVFS) and dynamic power management (DPM), but none have priority information on the application, leading to inefficient software operation.
+
+		This work proposes a set of tools, models, and algorithms for energy optimization aimed at high-performance computing based on knowledge of the application and the specific architecture.
+The main contributions of this work are:
+
+		A framework called PascalSuite automatically measures and compares multiple executions of a parallel application according to various scenarios characterized by input arrangements, number of threads, number of cores, and frequencies.
+As a result, PascalSuite can automate designing application models with an overhead of less than 1\%.
+
+		A full-system energy model based on the CPU frequency and the number of cores.
+The model aims to understand and optimize the energy behavior of parallel applications in HPC systems according to application parameters, such as the degree of parallelism and CPU parameters related to dynamic and static power.
+
+		A methodology that combines measurement data with a heuristic algorithm to provide insights into choosing the best phase divisions.
+Our heuristic can reduce the scan space from $10^{7000}$ to $10^2$ with an average error of 10\% and up to 38\% reduction in energy consumption using optimal distribution compared to standard Linux DVFS.
 	\end{umonsThesisAbstract}
@@
 
-	\chapter{Introduction}
-	This chapter will...
+	\chapter{Introduction} \label{chapter:introduction}
+	In this chapter, we present the motivations for this work, as well as its objectives and main contributions.
 	\include{intro/intro}
 
-	\chapter{Application Models} \label{chapter:models}
-	In this chapter, the proposed power, performance and energy models are derived and validated.
+	\chapter{Theoretical background} \label{chapter:theoretical_background}
+	This chapter will give the reader a better understanding of some terms used throughout the thesis, as well as details of configurations, the chosen benchmarks, and the architecture of the case study.
+	\include{experiments/setup}
+
+	\chapter{Parallel Scalability Suite} \label{chapter:pascal_suite}
+	High-performance computing systems have become increasingly dynamic, complex, and unpredictable.
+To help build software that uses full-system capabilities, performance measurement and analysis tools exploit extensive execution analysis focusing on single-run results.
+Despite being effective in identifying performance hotspots and bottlenecks, these tools are not sufficiently suitable to evaluate the overall scalability trends of parallel applications.
+Either they lack the support for combining data from multiple runs or collect excessive data, causing unnecessary overhead.
+	In this chapter, we present a tool for automatically measuring and comparing several executions of a parallel application according to various scenarios characterized by the input arrangements, the number of threads, number of cores, and frequencies.
+	Unlike other existing performance analysis tools, the proposed work covers some gaps in specialized features necessary to better understand computational resources scalability trends across configurations.
+	In order to improve scalability analysis and productivity over the vast spectrum of possible configurations, the proposed tool features automatic instrumentation, direct mapping of parallel regions, accuracy-preserving data reductions, and ease of use.
+	As it aims at accurately understanding scalability trends of parallel applications, detailed single-run performance analyses show minimal intrusion (less than 1\% overhead).
+
+	\include{pascalanalyzer/pascalanalyzer}
+	\include{fingerprint/fingerprint}
+	\include{framework/validation}
+
+	\chapter{Application energy and performance models} \label{chapter:models}
 	Energy consumption is crucial in high-performance computing (HPC), especially to enable the next exascale generation.
@@
 Nonetheless, due to numerous different implementations, we can always push the limits of software to achieve the most efficient use of our hardware.
-To be
-	energy efficient, the software relies on dynamic voltage and frequency scaling (DVFS), as well as dynamic power management (DPM).
+To be energy efficient, the software relies on dynamic voltage and frequency scaling (DVFS), as well as dynamic power management (DPM).
 Yet, none have privileged information on the hardware architecture and application behavior, which may lead to energy-inefficient software operation.
-	This study  proposes analytical modeling for architecture and application behavior that can be used to estimate energy-optimal software configurations and provide knowledgeable hints to improve DVFS and DPM techniques for single-node HPC applications.
+	This chapter proposes analytical modeling for architecture and application behavior that can be used to estimate energy-optimal software configurations and provide knowledgeable hints to improve DVFS and DPM techniques for single-node HPC applications.
 	Additionally, model parameters, such as the level of parallelism and dynamic power, provide insights into how the modeled application consumes energy, which can be helpful for energy-efficient software development and operation.
 	This novel analytical model takes the number of active cores, the operating frequencies, and the input size as inputs to provide energy consumption estimation.
-	We present the modeling of 13 parallel applications employed to determine energy-optimal configurations for several different input sizes.
+	In this chapter we present the modeling of 13 parallel applications employed to determine energy-optimal configurations for several different input sizes.
 	The results show that up to 70\% of energy could be saved in the best scenario compared to the default Linux choice and 14\% on average.
@@
 	%\include{optmizers/optmizers}
-
-	\chapter{Energy optimization using the models}
-	\include{experiments/setup}
 	\include{optmization/optmization}
 
-	\chapter{Extending the models with phases}
+	\chapter{Application-phase heuristic energy optimization} \label{chapter:phases}
+	Dynamic voltage and frequency scaling (DVFS) and dynamic power management (DPM) are essential techniques for saving energy in modern computers.
+These techniques control hardware resources such as frequency and number of active processors.
+Although there is much work on the subject, there is an underexplored sub-topic: the impact of the choice of phase division on energy consumption.
+	This chapter analyzes the impact of phase division on the energy efficiency of algorithms and proposes a methodology that combines measurement data with a heuristic to provide insights into choosing the best phase divisions.
+	Our heuristic can reduce the scan space from $10^{7000}$ to $10^2$ with an average error of 10\% and up to 38\% reduction in energy consumption using optimal distribution compared to standard Linux DVFS.
+Moreover, we evaluated the trade-off of having too many divisions and the overhead caused.
+Finally, we have identified that there is usually a limit to the number of phases a running application can benefit from, giving a lower limit to the minimum number of phases.
 	\include{phases/phases}
 
-	\chapter{Framework}
-	This chapter introduces the framework developed, the fingerprint tool and pascal analyzer.
-
-	\include{pascalanalyzer/pascalanalyzer}
-	\include{fingerprint/fingerprint}
-
-	This chapter will describe all the experiments that validate the models constraints.
-	\include{framework/validation}
-
-	\chapter{Conclusions}
-	This chapter will...
+	\chapter{Conclusions and future work*} \label{chapter:conclusions}
+	In this chapter, we deal with the conclusions about the effectiveness of the frameworks and the advantages and disadvantages of using models and algorithms with additional information on applications and architectures for energy optimization.
+	We also discussed what could be improved in each approach and framework.
 	\include{conclusion}

### 2022-08-15 thesis/pascalanalyzer/pascalanalyzer.tex
organazing the chapters and trying to connect

@@
-\section{PascalAnalyzer} \label{sec:pascal_analyzer}
-High-performance computing systems have become increasingly dynamic, complex, and unpredictable.
-To help build software that uses full-system capabilities, performance measurement and analysis tools exploit extensive execution analysis focusing on single-run results.
-Despite being effective in identifying performance hotspots and bottlenecks, these tools are not sufficiently suitable to evaluate the overall scalability trends of parallel applications.
-Either they lack the support for combining data from multiple runs or collect excessive data, causing unnecessary overhead.
-In this work, we present a tool for automatically measuring and comparing several executions of a parallel application according to various scenarios characterized by the input arrangements, the number of threads, number of cores, and frequencies.
-%In this work, we present a tool for measuring and comparing several executions of a parallel application automatically with different input arrangements, number of threads, cores, and frequencies to understand its scalability by observing how it uses the available computational resources across configurations.
-Unlike other existing performance analysis tools, the proposed work covers some gaps in specialized features necessary to better understand computational resources scalability trends across configurations.
-%The proposed work covers some gaps and unspecialized features of existing performance analysis tools. 
-%The tool features automatic instrumentation and direct mapping of parallel regions, accuracy-preserving data reductions, and ease of use to compare multiple configurations for improving analysis and productivity.
-In order to improve scalability analysis and productivity over the vast spectrum of possible configurations, the proposed tool features automatic instrumentation, direct mapping of parallel regions, accuracy-preserving data reductions, and ease of use.
-%In addition, it aims at parallel scalability, instead focus on single-run performance, aggregating a minimal level of intrusion (<1\% on average), which is fundamental for accurately understanding the behavior and scalability trends of the parallel application.
-As it aims at accurately understanding scalability trends of parallel applications, detailed single-run performance analyses show minimal intrusion (less than 1\% overhead).
-
 \section{Introduction to profiling tools} \label{sec:introduction_to_profiling_tools}
@@
 This investigative process tends to be tedious and complex, requiring in-depth knowledge of the problem domain, parallel systems, and measurement and analysis tools.
-%Mastering the analysis tools becomes necessary mainly because they differ from each other in terms of measurement strategy, the collected metrics, and the many points and focus of observation. 
 Mastering the analysis tools is necessary because they differ from each other in terms of measurement strategy, metrics collected, and the many points and focus of observation.
@@
 
-%Despite their differences, every existing tool has its own contributions in helping to understand the inner working of a program, and some can present a vast and varied set of metrics with a high degree of detail.
 Every existing tool has its own contributions in helping to understand the inner working of a program. Some can present a vast and varied set of metrics with a high degree of detail. 
 However, developers may have difficulty using these tools when their interest is in analyzing parallel scalability.
-%The main challenge arises from the focus on investigating a single run, because scalability analysis requires contrasting runs of different configurations. 
 The main challenge arises from the emphasis on studying a single run because scalability analysis requires contrasting runs of different configurations.
-%In this sense, it tends to be more significant to measure and compare only a few data across executions of multiple configurations than to collect a vast amount of finer-grain data of single-configuration runs.
 In terms of productivity, measuring and comparing only a few data between runs of multiple configurations tends to be more significant to scalability analysis than collecting a vast amount of finer-grain data of single-configuration runs.
@@
 First, because measuring and collecting only the necessary data limits the degree of intrusion and, second, because this strategy makes the analysis tool simpler and easier to use.
-%Finally, the tool's availability can also be a drawback as some apply to specific compilers and execution frameworks.
-Indeed, the framework version is a limitation for some traditional tools like HPCToolkit~\cite{Adhianto2010} and TAU (Tuning and Analysis Utilities)
-%MDPI: Is this an abbreviation? If affirmative, please add the explanation. Please carefully check and confirm -- INCLUDED ON THE NEXT PARAGRAPH
 ~\cite{Shende2006}.
@@
 These features allow one to automate the process of running and merging data collected from multiple configurations, bypassing manual efforts and avoiding the drawbacks of non-computer-oriented analysis.
-%The tool does not display visuals elements natively.
-Still, its interface allows the collected data to be exported and viewed from the control terminal in tabular form.
-Data exported via the interface are organized into data frames that can be interpreted by other tools or even rendered directly from graphic libraries.
 The collected data, available from the terminal in tabular form, can be exported as data blocks that can be interpreted by other tools or even rendered by graphics libraries to facilitate the visualization and data analysis.
 
-%The proposed tool is an alternative for realizing parallel scalability analysis more efficiently than single-run-centric performance measurement and analysis tools. 
 We offer an alternative tool for realizing parallel scalability analysis more efficiently than single-run-centric performance measurement and analysis tools.

### 2022-08-15 thesis/phases/phases.tex
fix references
organazing the chapters and trying to connect

@@
-\section{Introduction} \label{sec:introduction}
+\section{Introduction} \label{sec:phases_introduction}
 
-Energy efficiency has been one of the main focuses in computing nowadays, both on mobile and HPC.
-It is essential to increase battery life and reduce heat generation, especially in high-performance computing (HPC), where a small percentage of the savings energy can significantly reduce maintenance costs or environmental impact due to its scale of consumption.
+Energy efficiency has been one of the main focuses in computing nowadays, both in the mobile world and in high performance computing (HPC).
+Indeed, it is essential to increase battery life and reduce heat generation, especially in HPC, where a small percentage of the savings energy can significantly reduce maintenance costs or environmental impact due to its scale of consumption.
 For example, the leading Petaflop supercomputers consume in the range of 1 to 18 MW of electrical energy, with 1.5 MW on average, which we can estimate in millions of dollars in annual electricity cost \cite{Group2012HandbookSahni}.
-Furthermore, the International Energy Agency \cite{iea_2021} estimated that global data center electricity usage in 2020 was around 200-250 TWh, or around 1\% \cite{Corcoran2017EmergingICT} of global electricity demand, which could generate as much pollution as a nation like Argentina \cite{Mathew2012Energy-awareNetworks}.
+Furthermore, the International Energy Agency \cite{iea_2021} estimated that global data center electricity usage in 2020 was around 200-250 TWh, or around 1\% \cite{Corcoran2017EmergingICT} of global electricity demand  \cite{Mathew2012Energy-awareNetworks}.
 
 Given the importance of this topic, modern desktops, cell phones, and HPC CPUs have built-in power management mechanisms.
-That is because the processor is one of the main components that drain energy and can reach 50\% of the total system consumption \cite{Fan2007, Barroso2007TheComputing, Malladi2012TowardsDRAM}.
-Among the mechanisms implemented by the CPU, the most impactful and controllable via software are Dynamic Frequency and Voltage Scaling (DVFS), and Dynamic Power Management (DPM) ~\cite{Rotem2012Power-managementBridge, Brown2005, Hackenberg2015}.
+That is because the processor is one of the main components that drain energy and can reach 50\% of the total system consumption \cite{Fan2007PowerComputer, Barroso2007TheComputing, Malladi2012TowardsDRAM}.
+Among the mechanisms implemented by the CPU, the most impactful and controllable via software are Dynamic Frequency and Voltage Scaling (DVFS), and Dynamic Power Management (DPM) ~\cite{Rotem2012Power-managementBridge, Brown2005ACPILinux, Hackenberg2015AnProcessor}.
 
@@
 However, it is not a trivial task, and several variables need to be considered, such as the cost of switching, in terms of time and energy, how long it will remain in that state, the energy consumption in each state, and much more.
-
-The DPM technique has the most significant gains when in systems where the static power is very high or where the system is idle for a long time.
-In such cases, some papers report savings of up to 70\% in energy~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.
+This technique has the most significant gains when in systems where the static power is very high or where the system is idle for a long time.
+In such cases, some papers report energy savings of up to 70\%~\cite{Shuja2012Energy-efficientCenters, Benini2000AManagement}.
 
 DVFS, on the other hand, allows real-time frequency and voltage control, depending on needs.
-This method is motivated by the fact that we can approximate the relationship of frequency to power as a cubic equation and frequency and performance as a linear equation \cite{Dayarathna2016DataSurvey, Group2012HandbookSahni} implying that a reduction in frequency has a cubic impact on power and linear performance.
-Still, it is not so simple because lowering the frequency leads to longer runtime, increasing power consumption.
-As a result, determining the best voltage and frequency to employ under all conditions is also a challenging task.
+This method is motivated by the fact that we can approximate the relationship of frequency to power as a cubic equation and frequency and performance as a linear equation \cite{Dayarathna2016DataSurvey, Group2012HandbookSahni}, implying that a reduction in frequency has a cubic impact on power and linear performance.
+However, it is not so simple because lowering the frequency leads to longer execution time, increasing power consumption.
+Therefore, determining the best voltage and frequency to employ under all conditions is a daunting task.
 
-Although there are many studies on this topic, most works focus on implementing different optimization algorithms based on workload, and few studies pay attention to phase division.
+Although there are many studies on this topic, most works focus on implementing different optimization algorithms based on workload, and few studies pay attention to phase division, i.e.
+splitting the application in time intervals.
 Instead, most studies use static phase division to simplify the optimization math and compensate by implementing lightweight algorithms.
@@
 In addition, there is an overhead associated with the moment the algorithm takes action that could be reduced if we knew the ideal moment for it to take action, making it possible even to allow heavier algorithms.
-Therefore, this work proposes a methodology to study the division of time (phases) and analyzing possible gains.
 
-\section{Related work} \label{sec:related_work}
+Therefore, this work proposes a methodology to study the division of time (phases) and analyzing possible gains, this approach is better justified in \cref{sec:proposed_approach}
 
-One of the bases works in scheduling algorithms is the work of Irani et al.
-\cite{Irani2007}, where they formalized the problem of scheduling arriving jobs in a way that minimizes total energy use and so that each job is completed after its arrival time and before its deadline.
-Although each problem has been considered separately, this was the first theoretical analysis of systems that can use both mechanisms.
+The main contributions of this work are:
+\begin{itemize}
+	\item Heuristic algorithm for phase division
+	\item Phase division optimization method
+	\item Ideal number of phases for HPC applications
+\end{itemize}
+The proposed algorithm for phase division can be integrated in DVFS and DPM solutions to further improve them and archive better energy savings, described in \cref{sec:phase_division_algorithm}.
 
-Past work has also used feedback-based approaches, as in Poellabauer et al.
-\cite{Poellabauer2005}, where applications' past CPU utilization's are used to predict future  CPU  requirements.
-However, mispredictions in these approaches can lead to missed deadlines,  sub-optimal energy savings,  or large overheads due to frequent changes to the chosen frequency or voltage.
-One shortcoming of previous approaches is that they ignore other 'indicators' of future CPU requirements, such as the frequency of I/O operations, memory accesses, or interrupts.
+The phase optimization method can be used as an alternative for DVFS and DPM in energy saving optimization and also helps in the understanding of ideal phase division for a group of applications this is further detailed in \cref{sec:phases_optimization}.
 
-In the thesis work of Saha et al.
-\cite{Saha2012}, the timeliness and power consumption behavior of fourteen RT-DVFS schedulers through implementation and actual measurements.
-The schedulers include Static Earliest Deadline First (Static-EDF), Cycle Conserving Earliest Deadline First(CC-EDF), Look-Ahead Earliest Deadline First (LA-EDF), Snowdon-minimum (Snowdon-min), Resource-constrained Energy-Efficient Utility Accrual Algorithm (REUA), DynamicReclaiming Algorithm (DRA) and Aggressive Speed Reduction Algorithm(AGR) among the others.
-They draw attention to the fact that most of these algorithms are based on simulations, which can mislead some results.
-Broadly, RT-DVFS techniques have two objectives: (i) Reduce energy consumption through DVFS; and (ii) Optimize task timeliness behavior through real-time resource management (deadline).
+The ideal number of phases is discussed in the results in \cref{sec:results} as well a comparison of the proposed phase optimization gains with the default method in Linux.
+
+The conclusion of this work and proposed improvements are found in \cref{sec:conclusion}.
+
+\section{Related work} \label{sec:phases_related_work}
+
+One of the reference works in scheduling algorithms is the work of Irani et al.
+\cite{Irani2007}, where they formalized the problem of scheduling incoming jobs in a way that minimizes the total energy used, so that each job is completed after its arrival time and before its deadline.
+Although each problem has been considered separately, this was the first theoretical analysis of systems that could use both mechanisms.
+
+Previous work has also used feedback-based approaches, as in Poellabauer et al.
+\cite{Poellabauer2005}, where past CPU behavior when running the application is used to predict future  CPU  requirements.
+However, mispredictions in these approaches can lead to missed deadlines,  sub-optimal energy savings,  or significant overheads due to frequent changes in the chosen frequency or voltage.
+One shortcoming of previous approaches is that they ignore other "indicators" of future CPU requirements, such as the frequency of I/O operations, memory accesses, or interrupts.
+
+The thesis work of Saha et al.
+\cite{Saha2012} shows the timeliness and power consumption behavior of fourteen RT-DVFS schedulers through implementation and actual measurements.
+The schedulers include Static Earliest Deadline First (Static-EDF), Cycle Conserving Earliest Deadline First(CC-EDF), Look-Ahead Earliest Deadline First (LA-EDF), Snowdon-minimum (Snowdon-min), Resource-constrained Energy-Efficient Utility Accrual Algorithm (REUA), DynamicReclaiming Algorithm (DRA) and Aggressive Speed Reduction Algorithm(AGR) among others.
+They draw attention to the fact that most of the evaluations were based on simulations, which can mislead some results.
+Broadly, RT-DVFS techniques have two objectives: (i) Reduce energy consumption through DVFS, and (ii) Optimize task timeliness behavior through real-time resource management (deadline).
 
@@
 \cite{Pietri2014} used slack reclamation to achieve energy savings.
-The goal is to exploit the idle slots of the processors that occurred due to the earlier completion time of the tasks compared with the latest finish time, which is constrained by the deadline or data dependencies.
+The goal is to exploit the idle slots of the processors that occurred due to the earlier completion time of the tasks compared with the latest finish time, which is constrained by deadline or data dependencies.
 In their model, the user submits a workflow for execution specifying a deadline for completion of the execution.
@@
 This algorithm selects a slot for each task to minimize the total energy consumption of the MapReduce job for big data applications in heterogeneous environments without significant performance loss while satisfying the service level agreement (SLA).
-TGSAVE  finds solutions under deadlines up to 74\% tighter than the tightest one feasible by EMRSA, and it can meet deadlines as tight as only12\%, on average, more significant than the energy-oblivious minimum makespan.
+TGSAVE  finds solutions under deadlines up to 74\% tighter than the tightest one feasible by EMRSA, and it can meet deadlines as tight as only 12\%, on average, more significant than the makespan algorithm.
 
@@
 
-%Typically target traditional real-time systems with static deadlines, resulting in conservative energy savings that cannot exploit additional energy optimizations due to dynamic deadlines.
-However, in the work of Yi et al.
-\cite{Yi2021} they present an adaptive system optimization and reconfiguration approach that dynamically adapts the scheduling parameters and processor speeds to satisfy dynamic deadlines while consuming as little energy as possible.
-As their work was for autonomous driving, they used velocities as deadlines.
-
 In the work of \cite{Ajmal2021}, a green cloud computing algorithm named "Cost-based Energy Efficient Scheduling Technique for Dynamic Voltage Frequency Scaling (DVFS) Systems (CEEST)" is proposed.
@@
 
-All of these works at some point considered a deadline (defined by the user of the system) that defines the phases of the job; this means that choosing a different deadline could result in completed different energy savings, and because this is not the main focus of these works, this is not considered.
+All of these works at some point considered a deadline (defined by the user or the system) that defines the phases of the job; this means that choosing a different deadline could result in completely different energy savings.
 In our paper, we study the impact of this phase division on the application's energy consumption.
-We estimate how much energy could be saved if an algorithm could magically give the ideal deadline.
+In this context, we estimate how much energy could be saved if an algorithm could magically give the ideal deadline.
 
-Although most of the work is considered a deadline, some works include job division as part of the optimization problem.
-For example, in the paper of Agrawal et al.
+Although most of the works consider the deadline, some works also take job division as part of the optimization problem.
+In the paper of Agrawal et al.
 \cite{Agrawal2021}, they show that if the jobs can be divided into arbitrary parts, the minimum-energy schedule can be generated in linear time, giving exact scheduling algorithms.
-For the cases where jobs are non-divisible, they claimed a prove that the scheduling problems are NP-hard and also give approximation algorithms for the same along with their bounds.
-
+For the cases where jobs are non-divisible, they claimed a proof that the scheduling problems are NP-hard, while giving approximation algorithms for these with their bounds.
 In this case, where no deadline is defined, the phase division is indirectly considered by solving the optimization problem of minimizing the energy.
@@
 
-\section{Phase division approach} \label{sec:application_partition_method}
+\section{Proposed approach} \label{sec:proposed_approach}
 To find the optimal phase division, we would have to ensure that the one we choose is the most energy-efficient among all possible divisions and different combinations of power configurations.
-
 This mathematical problem can be modeled in many ways, but always with some concessions to make it solvable.
 Another option would be by brute force, but there are infinite possibilities for splitting phases, although some hardware limitations make our analysis more accessible.
-
 With that in mind, we decided to create a heuristic that comes the closest to brute force using hardware limitations to make it viable.
@@
 The number of possibilities could be estimated by:
-\begin{equation}
-	\left(\frac{T}{d_t}\right)^C,
-\end{equation}
+$ \left(\frac{T}{d_t}\right)^C.$
 This number grows astronomically because $d_t$ is in a time range of microseconds, $T$ seconds or minutes, and C of hundreds.
-To give an idea, we can estimate this number to be in a range of $10^7000$.
+To give an idea, we can estimate this number to be in a range of $10^{7000}$.
 
@@
 For example, if we ran each power setting once and we had a way of combining them, our problem now would be feasible.
-
 To make this possible, we assume that the configuration change only affects the program speed and that the power consumption is independent of the previous settings.
-
-\iffalse %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% commented 
-This means that in a certain part of the program $x_1$\% to $x_2$\% it will always execute the same instructions and the energy consumption would depend only on the configuration.
-
-That said, if we treat the program as a percentage of execution, and we have the power consumption profile for all configurations, we can estimate the power consumption for any combination of phases.
-
-For example if we have the phase division, 0-x_1\%, x_1\%-100\%, with the settings c_1 and c_2, the estimated energy would be:
-P_{c_1}*u(0,x_1)*T_{c_1}+P_{c_2}*u(x_1, 100)*T_{c_2}
-
-We can see that we just need a power profile P and the total runtime for each configuration and that we can divide the application into as many phases as we want.
-
-For that we are going to assume that the number of instructions to complete a program is roughly the same independent on the frequency, thus we can write the number of instructions as:
-\begin{equation}
-	I=c_{k1}f_1T_1=c_{k2}f_2T_2,
-\end{equation}
-where $I$ is the number of instructions, $c_{k1}$ the rate of instructions per cycle, $f$ the frequency of the processor and $T$ the execution time.
-All the required parameters can be collected by simply running the application with the available frequencies.
-
-From this equation we can split the execution of the program into several different phases $i$ as follows:
-% \begin{equation}
-%     I=xI+(1-x)I=xc_{k1}f_1+(1-x)c_{k2}f_2
-% \end{equation}
-\begin{equation}
-	I=\sum_{i}{a_iI} %=\sum_i{a_ic_{ki}f_i}
-\end{equation}
-
-with $a_i$ as the percentages of instructions executed with frequency $f_i$, we can compute any combination of phases. For convenience, we write this as a function of the time:
-% \begin{equation}
-% T=\frac{I}{c_{k1}f_1}
-% \end{equation}
-\begin{equation}
-	I=Tc_{k}f
-\end{equation}
-\begin{equation}
-	I=\sum_{i}{a_iI}=\sum_i{a_iTc_{ki}f_i}
-\end{equation}
-% \begin{equation}
-% I=xI+(1-x)I=xTc_{k1}f_1+(1-x)Tc_{k2}f_2
-% \end{equation}
-
-the application being divided into phases of different duration, $a_i$  becomes the duration of the phase as a percentage of the total execution time.
-This general equation does not limit the number and duration of phases, and several phases can have the same frequency.
-
-In the event that we decide to include the number of cores as a context parameter, the equation becomes more complex.
-Indeed, the number of cores can completely change the application's behavior but, depending on the type of application, it's still possible to find similarities.
-Many HPC applications use data parallelism, and the Amdahl's law can model most of it.
-For this kind of applications, we can estimate that:
-\begin{equation}
-	T=\frac{T_s}{S}
-\end{equation}
-\begin{equation}
-	T_s=\frac{I}{c_kf}
-\end{equation}
-\begin{equation}
-	S=\frac{1}{1-w+\frac{w}{p}}
-\end{equation}
-\begin{equation}
-	T=\frac{I(1-w+\frac{w}{p})}{c_kf}=\frac{I}{c_kf}-\frac{Iw}{c_kf}+\frac{Iw}{c_kfp}
-\end{equation}
-
-Looking at a fixed frequency, we arrive at a very similar structure as before:
-\begin{equation}
-	T=c_1I+\frac{c_2I}{p}
-\end{equation}
-
-As previously, we can use $I$ as the number of reference instructions, and compute $a$ as the percentage of single and multi-core instructions, or express it as a percentage of the total time T:
-\begin{equation}
-	aT=c_1aI+\frac{c_2aI}{p}
-\end{equation}
-
-The extent of the hypotheses will later be validated, however, by taking advantage of both principles, we can now, based on real measurements, estimate any phase division and possible context combination.
-\fi %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\section{Building applications power profile} \label{sec:building_a_database}
-
-%\subsection{Case-Study Architecture} \label{subsec:casestudyarchitecture}
-%We executed the experiments in a system equipped with two Intel Xeon E5-2698 v3 processors with sixteen cores and two hardware threads for each core, with a total physical memory of the node is 128GB (8$\times$16GB).
-We disabled turbo frequency and hardware multi-threading during all experiments.
-The operating system used was Linux CentOS 6.5 and kernel 4.16.
-The overall view of the architecture is shown in \cref{fig:architecture}.
-%
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=0.8\columnwidth]{models/figures/architecture.png}
-%	\caption{Node architecture (the image was made with the lstop application).}
-%	\label{fig:architecture}
-%\end{figure}
-%
-%To perform the frequency control, we used the acpi-cpufreq driver using the Userspace governor, which allows the user or any userspace program to set the CPU to a specific frequency.
-While the cores control was accomplished by modifying the appropriate system files with the default CPU-hotplug driver.
-%
-%The architecture is also equipped with the Intelligent Platform Management Interface (IPMI), a set of interfaces allowing out-of-band management of computer systems and platform-status monitoring via the local network~\cite{Schwenkler2006IntelligentInterface}.
-In addition, it can monitor variables and resources such as the system's temperature, voltage, fans, and power supplies, with independent sensors attached to the hardware.
-%
-%\subsection{Case-Study Applications} \label{subsec:casestudyapplication}
-%The applications Black-Scholes, Bodytrack, Canneal, Dedup, Fluidanimate, Freqmine, Raytrace, Swaptions, Vips and x264 from the PARSEC \footnote{\url{https://parsec.cs.princeton.edu/download.htm}} parallel benchmark suite, version 3.0~\cite{Bienia2008TheSuite}, OpenMC \cite{Romano2015OpenMC:Development} and LINPACK (HPL) \cite{Dongarra1988TheExplanation}, were chosen as case studies.
-The PARSEC benchmark focused on emerging workloads and was designed to represent the next-generation shared-memory programs for chip-multiprocessors.
-It covers many areas such as financial analysis, computer vision, engineering, enterprise storage, animation, similarity search, data mining, machine learning, and media processing.
-The OpenMC and the LINPACK are two classic HPC programs.
-
-\subsection{Data gathering} \label{sec:data_gathering}
-
-Using the Pascal framework \cite{electronics11050689} the data was collected by running the application in all possible configurations.
-In this system, that is 32 cores and 13 frequencies, a total of 416 (32*13) possible configurations.
-Power samples were taken from dedicated sensors in the IPMI system with a sample rate of 0.5 seconds.
-
-%\subsection{Application fingerprint}
-%\textcolor{red}{Introduce the power profile as fingerprint briefly...}
-
-\section{The phase optimization algorithm} \label{sec:heuristic}
+\section{Phase division algorithm} \label{sec:phase_division_algorithm}
 To analyze the impact of configurations and identify the optimal phase division, we propose a zero-order integrator, which calculates the energy in an interval for a given configuration.
-The integrator uses the data previously described in \cref{sec:data_gathering}.
+The integrator uses the data gathering described in \cref{subsec:data_gathering}.
 
 Figure \ref{fig:zero_order} shows an example energy computation for an application with four different power configurations.
-For that, first, the phases are defined in terms of the percentage of execution.
-Then, the integrator will run over each phase using the selected profile of power obtained from the measurements and accumulate the total energy consumption.
+The phases are, first, defined in terms of the percentage of execution.
+Then, the integrator will run over each phase using the selected power profile obtained from the measurements, and accumulate the total energy consumption.
 
@@
 	\includegraphics[width=\columnwidth]{phases/figures/integrator.pdf}
-	\caption{Power vs percentage of execution for a given application in several configurations of power. The energy is represented by the hatched area.}
+	\caption{Power vs percentage of execution for a given application in several configurations (1 to 4). For each configuration, the energy in each phase is represented by the hatched area.}
 	\label{fig:zero_order}
@@
 
-		// find the index of first sample in the range
+		// find the index of the first sample in the range
 		for (idx_i= 0; idx_i<nsamples; idx_i++)
@@
 
-		// handle sample in between
+		// handle samples in between
 		for (int i= idx_i+1; i<idx_f; i++)
@@
 
+This algorithm works by...
+
+\section{Phases optimization} \label{sec:phases_optimization}
+
+\subsection{Data gathering} \label{subsec:_phases_data_gathering}
+
+Using the Pascal framework \cite{electronics11050689}, the data was collected by running applications in all possible configurations.
+In this system, that is 32 cores and 13 frequencies, means a total of 416 (32*13) possible configurations per application.
+Power samples were taken from dedicated sensors in the IPMI system with a sample rate of 0.5 seconds.
+
+\subsection{Phase optimization algorithm} \label{subsec:phase_optimization_algorithm}
 To obtain an optimal energy configuration for a given phase division, we can run all possible configurations and select the one that leads to a minimal consumption, as shown in the pseudo-algorithm below:
@@
 			{
-				en = estimate_energy_phase(
+				en = integrator(
 				configs[j].total_time, 
@@
 
-This method can now obtain an optimal energy configuration for a given phase. The next step is to find the best phase, division.
+This method can now obtain an optimal energy configuration for a given phase. The next step is to find the best phase division.
 
 \section{Results} \label{sec:results}
+Now that we can evaluate phase divisions, we need to answer two questions, what is the best number of phases, and what is the best division within this number?
+For that, we used the algorithm described previously in \cref{sec:heuristic} as the objective of a minimization problem where we want to find the vector of phases that minimizes the total energy. 
+A genetic algorithm will be used as a solver, whose execution process can be accelerated by controlling the mutation and the initial population since some constraints are already known.
+For instance, to study the phases in the range of 3 to 100 divisions, we use a GA with a population of $10^3$ individuals.
+We keep the best 10\% individuals from each generation, while having a 10\% of mutation rate, over 300 generations.
+The reproduction function for the GA combines half of the phases of one individual with another, and the mutation will randomly change one division.
 
-Now that we can evaluate phase divisions, we need to answer two questions, what is the best number of phases, and what is the best division within this number?
-
-For that, we used the algorithm described previously in \cref{sec:heuristic} as the objective of a minimization problem where we want to find the vector of phases that minimizes the total energy.
-Moreover, by using a genetic algorithm, we can accelerate the process by controlling mutation and initial population generation since we already know some information about our problem.
-
-For instance, we decide to start studying phases in the range of 3 to 100 divisions.
-We used a GA with a population of $10^3$ individuals.
-We kept the best 10\% individuals on each generation while having a 10\% of mutation rate, running for 300 generations.
-The reproduction function for the GA combines half-phases of one individual with another, and the mutation will randomly change a division.
-
-We use the applications described in \cref{subsec:casestudyapplication}, to compute the optimal phase selection.
-The \cref{fig:relative_energy} shows the relative energy consumption compared to an optimized single phase.
+We used the applications described in \cref{sec:casestudyapplication}, to compute the optimal phase selection.
+The \cref{fig:relative_energy} indicates, for each application, the energy consumption according to the number of phases, compared to an optimized single-phase.
 
@@
 	\includegraphics[width=\columnwidth]{fingerprint/figures/energy_per_phase.pdf}
-	\caption{Relative energy vs number of phases using applications from PARSEC 3.0, HPC, and Openmc benchmarks with different inputs. Energy compared to the single-phase optimal configuration.}
+	\caption{Relative energy (compared to the single-phase optimal configuration) vs number of phases using applications from PARSEC 3.0, HPC, and Openmc benchmarks with different sizes' inputs.}
 	\label{fig:relative_energy}
@@
 
-This evaluation shows very interesting results, as shown in \cref{fig:relative_energy}.
-An ideal single-phase configuration provides the highest energy consumption for all cases.
-Furthermore, dividing the application into more than 35 phases is not necessary.
-On average, power consumption stopped after 35 phases.
+The \cref{fig:relative_energy} shows very interesting results.
+The energy consumption of the optimal single-phase configuration is always higher for all cases.
+Furthermore, it is not necessary to divide the application into more than 35 phases.
+Indeed, the energy gain is negligible beyond 35 phases.
 
-% \begin{figure}[H]
-% \includegraphics[width=\columnwidth]{figures/min_phases_distribution.pdf}
-%     \caption{Histogram of minimal number of phases to optimal energy}
-%     \label{fig:hist_optimal_phase}
-% \end{figure}
+The \cref{fig:hist_optimal_phase} shows in more detail the ideal number of phases for all of our tests, there we can see that most applications reach the ideal configuration with less than 20 phases, while there is still a uniform distribution to a higher number of phases.
+
+\begin{figure}[H]
+	\includegraphics[width=\columnwidth]{phases/figures/min_phases_distribution.pdf}
+	\caption{Histogram showing the frequency of the optimal number of phases for all applications}
+	\label{fig:hist_optimal_phase}
+\end{figure}
 
 The results meet our expectation that the optimum of n-divisions phases will always be better than or equal to n-1 divisions.
-This becomes clearer when we look at the number of cores, and frequency curve \cref{fig:freq_control} and \cref{fig:cores_control}, since starting from a certain number of divisions does not change the control signals since dividing an ideal phase into two would result in two phases with the same control signals.
+This becomes clearer when we look at the number of cores, in \cref{fig:cores_control_1} and \cref{fig:cores_control_2}, and frequencies, in  \cref{fig:black_control_1} and \cref{fig:black_control_2} for two particular applications, Bodytrack and Blackscholes.
+These applications demonstrate the optimal frequency and cores do not fundamentally change between 35 and 99 phases.
+%Since starting from a certain number of divisions does not change the control signals since dividing an ideal phase into two would result in two phases with the same control signals.
 Therefore, the energy will remain the same, and there is no reason to increase the number of phases beyond that.
@@
 	\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_1_freq_signals_cmp.pdf}
-	\caption{Frequency signal vs percentage of execution for the Bodytrack application, comparing the difference of 35 and 99 phases.}
+	\caption{execution frequency vs percentage of execution for the Bodytrack application, comparison when using 35 and 99 phases.}
+	\label{fig:black_control_1}
 \end{figure}%
@@
 	\centering
-	\includegraphics[width=\columnwidth]{phases/figures/signals/completo_freq_freq_signals_cmp.pdf}
-	\caption{Frequency signal vs percentage of execution for the Freqmine application, comparing the difference of 35 and 99 phases.}
+	\includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_1_freq_signals_cmp.pdf}
+	\caption{execution frequency vs percentage of execution for the Blackscholes application, comparison when using 35 and 99 phases.}
+	\label{fig:black_control_2}
 \end{figure}%
@@
 	\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_1_cores_signals_cmp.pdf}
-	\caption{Cores signal vs percentage of execution for the Bodytrack application, comparing the difference of 35 and 99 phases.}
+	\caption{number of cores vs percentage of execution for the Bodytrack application, comparison when using 35 and 99 phases.}
+	\label{fig:cores_control_1}
 \end{figure}%
@@
 	\centering
-	\includegraphics[width=\columnwidth]{phases/figures/signals/completo_freq_cores_signals_cmp.pdf}
-	\caption{Cores signal vs percentage of execution for the Freqmine application, comparing the difference of 35 and 99 phases.}
+	\includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_1_cores_signals_cmp.pdf}
+	\caption{number of cores vs percentage of execution for the Blackscholes application, comparison when using 35 and 99 phases.}
+	\label{fig:cores_control_2}
 \end{figure}%
 
-When we compare this method with the default DVFS algorithm in Linux, we observer an average saving of 38\% as shown in \cref{fig:cmp_ondemand} the relative energy per application.
+When we compare this method with the default DVFS algorithm in Linux, we observer an average saving of 38\%, as shown in \cref{fig:cmp_ondemand}, the relative energy per application with different input sizes.
 
@@
 	\includegraphics[width=\columnwidth]{phases/figures/comparison_ondemand.pdf}
-	\caption{Relative energy comparison, optimal phase division energy compared to governor Ondemand on Linux.}
+	\caption{Optimal phase splitting energy vs. on-demand governor on Linux: relative energy comparison for applications with different input sizes.}
 	\label{fig:cmp_ondemand}
@@
 
-To illustrate the phases division, the \cref{fig:phase_division_cmap_35} and \cref{fig:phase_division_cmap_35} shows a heatmap with the divisions by application and the percentage of total energy spent in that phase.
-There we can see a pattern of 3 phases which for these applications are generally setup phase where it loads data from the disk, a computation phase where the data is processed, and a finalization phase where data is written back to the disk or displayed to the user.
+To illustrate the phases division, the \cref{fig:phase_division_cmap_3} and \cref{fig:phase_division_cmap_35} show the heatmap for applications when divided with the same number of phases and the percentage of total energy spent in each phase.
+The 3 phases division can already show interesting program behavior, such as application setup, loading data, computation and data storage.
 

### 2022-08-15 thesis/references.tex
organazing the chapters and trying to connect

--- 2022-04-14 thesis/references.tex
@@
 
+	%%%%%%%%%%% PHASES %%%%%%%%%%%
+
+	\bibitem[Iea(2021)]{iea_2021}
+	Iea, ``Data centres and data transmission networks – analysis,'' Nov 2021.
+	[Online]. Available:
+	\url{https://www.iea.org/reports/data-centres-and-data-transmission-networks}
+
+	\bibitem[Cardoso(2017)]{CARDOSO201717}
+	J.~M. Cardoso, J.~G.~F. Coutinho, and P.~C. Diniz, ``Chapter 2 -
+	high-performance embedded computing,'' in \emph{Embedded Computing for High Performance}, J.~M. Cardoso, J.~G.~F. Coutinho, and P.~C. Diniz, Eds.\hskip
+	1em plus 0.5em minus 0.4em\relax Boston: Morgan Kaufmann, 2017, pp. 17--56.
+	[Online]. Available:
+	\url{https://www.sciencedirect.com/science/article/pii/B9780128041895000028}
+
+	\bibitem[Irani(2007)]{Irani2007}
+	S.~Irani, S.~Shukla, and R.~Gupta, ``Algorithms for power savings,'' \emph{ACM Transactions on Algorithms}, vol.~3, p.~41, 11 2007. [Online]. Available:
+	\url{https://dl.acm.org/doi/10.1145/1290672.1290678}
+
+	\bibitem[Poellabauer(2005)]{Poellabauer2005}
+	C.~Poellabauer, L.~Singleton, and K.~Schwan, ``Feedback-based dynamic voltage
+	and frequency scaling for memory-bound real-time applications,'' 2005, pp.
+	234--243.
+
+	\bibitem[Saha(2012)]{Saha2012}
+	S.~Saha and B.~Ravindran, ``An experimental evaluation of real-time dvfs
+	scheduling algorithms.''\hskip 1em plus 0.5em minus 0.4em\relax ACM Press,
+	2012, pp. 1--12. [Online]. Available:
+	\url{http://dl.acm.org/citation.cfm?doid=2367589.2367604}
+
+	\bibitem[Pietri(2014)]{Pietri2014}
+	I.~Pietri and R.~Sakellariou, ``Energy-aware workflow scheduling using
+	frequency scaling.''\hskip 1em plus 0.5em minus 0.4em\relax IEEE, 9 2014, pp.
+	104--113. [Online]. Available:
+	\url{http://ieeexplore.ieee.org/document/7103444/}
+
+	\bibitem[Mashayekhy(2014)]{Mashayekhy2014}
+	L.~Mashayekhy, M.~M. Nejad, D.~Grosu, D.~Lu, and W.~Shi, ``Energy-aware
+	scheduling of mapreduce jobs.''\hskip 1em plus 0.5em minus 0.4em\relax
+	Institute of Electrical and Electronics Engineers Inc., 9 2014, pp. 32--39.
+
+	\bibitem[Yousefi(2018)]{Yousefi2018}
+	M.~H.~N. Yousefi and M.~Goudarzi, ``A task-based greedy scheduling algorithm
+	for minimizing energy of mapreduce jobs,'' \emph{Journal of Grid Computing},
+	vol.~16, pp. 535--551, 12 2018.
+
+	\bibitem[Kessler(2021)]{Kessler2021}
+	C.~Kessler, S.~Litzinger, and J.~Keller, ``Crown-scheduling of sets of
+	parallelizable tasks for robustness and energy-elasticity on many-core
+	systems with discrete dynamic voltage and frequency scaling,'' \emph{Journal
+		of Systems Architecture}, vol. 115, p. 101999, 5 2021. [Online]. Available:
+	\url{https://linkinghub.elsevier.com/retrieve/pii/S1383762121000175}
+
+	\bibitem[Ajmal(2021)]{Ajmal2021}
+	M.~S. Ajmal, Z.~Iqbal, F.~Z. Khan, M.~Bilal, and R.~M. Mehmood, ``Cost-based
+	energy efficient scheduling technique for dynamic voltage and frequency
+	scaling system in cloud computing,'' \emph{Sustainable Energy Technologies
+		and Assessments}, vol.~45, p. 101210, 6 2021. [Online]. Available:
+	\url{https://linkinghub.elsevier.com/retrieve/pii/S2213138821002204}
+
+	\bibitem[Agrawal(2021)]{Agrawal2021}
+	P.~Agrawal and S.~Rao, ``Energy-efficient scheduling: classification, bounds,
+	and algorithms,'' \emph{Sādhanā}, vol.~46, p.~46, 12 2021. [Online].
+	Available: \url{http://link.springer.com/10.1007/s12046-021-01564-w}
+
+	\bibitem[Silva(2022)]{electronics11050689}
+	V.~R.~G. da~Silva, A.~B.~N. da~Silva, C.~Valderrama, P.~Manneback, and
+	S.~Xavier-de Souza, ``A minimally intrusive approach for automatic assessment
+	of parallel performance scalability of shared-memory hpc applications,''
+	\emph{Electronics}, vol.~11, no.~5, 2022. [Online]. Available:
+	\url{https://www.mdpi.com/2079-9292/11/5/689}
+
 \end{thebibliography}

### 2022-08-15 thesis/pascalanalyzer/pascalanalyzer.tex
Updates from Overleaf

--- 2022-05-03 thesis/pascalanalyzer/pascalanalyzer.tex
@@
-\section{PascalAnalyzer} \label{sec:pascal_analyzer}
+\section{PascalAnalyzer (pascalina)} \label{sec:pascal_analyzer}
 High-performance computing systems have become increasingly dynamic, complex, and unpredictable.

### 2022-08-19 thesis/conclusion.tex
conclusion

@@
 \section{Conclusion}
+
+This thesis presents a complete way to model HPC applications.
+With a low overhead framework to collect the data, an analytical model of energy consumption and a heuristic algorithm for multi-phase applications.
+Each of the methods presented has possible improvements that will be discussed below.
 
@@
 
-This work introduces a practical and easy-to-use tool for measuring and analyzing the efficiency of parallel applications.
-The proposed tool focuses on observing scalability and energy consumption and implements features that enables analysis at hierarchical levels of the program's inner parts.
+The Pascal Suite framework introduces a practical and easy-to-use tools for measuring and analyzing the efficiency of parallel applications.
+The proposed tool focuses on observing energy consumption and scalability and implements features that enables analysis at hierarchical levels of the program's inner parts.
 It also simplifies the comparison of the application runs in different configurations, helping developers target software optimization efforts.
 
-The tool has a low level of intrusion added to the program's performance measurement under analysis, which is a fundamental aspect of understanding the program's behavior and scalability capacity. 
-Although it does not offer graphic elements to visualize the collected metrics, the tool itself makes it possible to examine the data from the command terminal.
-It also allows this data to be analyzed by an interface provided in Python or through .json files.
+The tool has a low level of intrusion added to the program's performance measurement under analysis, which is a fundamental aspect of understanding the program's behavior.
+
+From the results, we observe that the fingerprint module API present overhead similar or lower to other low-level APIs, with the advantage of being in high abstraction and simplified configuration with a few lines of code, is possible to configure and gather performance counter data.
+
+The fingerprint module developed provide also provided a way to compute similarities between different programs or the same program with different inputs.
+This can be useful to reduce applications spaces for benchmarks as was done in Polybench clustering but also to analyze the behavior of a parameter providing insights to the programmer to find a possible bottleneck.
 
 \textls[-20]{In future work, we intend to evolve this tool to include the ability to predict speedup and efficiency from a few samples using state-of-the-art prediction models present in the \mbox{literature \cite{Alex2020WhenParallelSpeedups, Vitor2022AnalyticalEnergyModel}}.
-The idea is to present the general behavior of the program and its scalability trend and reduce the execution time necessary to compose a comprehensive analysis.
+The idea is to present the general behavior of the program and reduce the execution time necessary to compose a comprehensive analysis.
 Furthermore, we intend to include features that allow the observation of parallel applications in distributed environments that use the Message Passing Interface (MPI) standard.}
 
-\subsubsection{fingerprint tool}
-
-From the results, we observe that the API present overhead similar or lower to other low-level APIs, with the advantage of being in high abstraction and simplified configuration with a few lines of code, is possible to configure and gather counter data.
-
-The tool developed provide also provided a way to fingerprint programs and compute similarities between different programs or the same program with different inputs.
-This can be useful to reduce applications spaces for benchmarks as was done in Polybench clustering but also to analyze the behavior of a parameter providing insights to the programmer to find a possible bottleneck.
-
-We also provide a precise definition to input size that made possible fingerprint the programs, but it also can help programmers of benchmark applications to better create inputs with more precise growth of a particular parameter.
-
-\subsection{Application model} \label{sec:conclusion_models}
-The proposed energy model based on the operating frequency and the number of cores for a shared memory system can serve as a reference for DVFS and DPM optimization problems.
-
-Results from three different HPC benchmarks demonstrate the potential of the proposed model while consuming 10 times less energy than a machine learning approach, such as SVR, to characterize applications.
-Moreover, it can provide knowledge-based hints to improve DVFS and DPM algorithms by enabling analysis of the contribution of each model parameter (e.g., level of parallelism) to the energy consumption.
-Indeed, as shown in Section \ref{sec:model_validation}, when no oracle is available to choose the frequency and the number of cores the application should use, the proposed model can save around 12\% of energy for a random choice and up to 70\% for the worse possible choice.
-Considering the job history of our own HPC center, which shows the prevalence of worse possible choices made by users, the potential energy savings are very significant and encourage further research.
-
-Although the model is promising, it still has some limitations.
-The main one is related to the input size, which needs to be estimated to create the application model and optimize the application.
-Another limitation concerns the power model, which does not consider the load variation, so our model ends up using an average of the energy consumption, which is enough to obtain good results but limits its implementation in real-time optimization.
-Future research is intended to solve both problems, first adapting the model to use the ratio of executed instructions as input size, something which is more tangible and easy to measure in modern systems without much overhead, and adding  new parameters to the power model to account for the load.
-This would allow us to develop more advanced DVFS models that could identify different phases of a target program with more subtle changes in frequency, and, perhaps, in the number of active cores to  further improve the results presented here.
+\subsection{Application-energy model} \label{sec:conclusion_models}
 
@@
 Another important aspect that is typically not taken into account is the number of processing cores to be used by a parallel program.
-This choice is left to the user, which often is not trivial as shown in this paper.
+This choice is left to the user, which often is not trivial as shown in this thesis.
 
-\subsection{Application-phase heuristic} \label{sec:phases_conclusion}
+\subsection{Application-phase heuristic algorithm} \label{sec:phases_conclusion}
 The main conclusion of this work was that in the HPC environment, in addition to the advantages of finding the optimal phase divisions, not many phases are needed to achieve the optimal energy consumption.

### 2022-08-19 thesis/framework/validation.tex
conclusion

@@
-\section{Pascal framework validation} \label{sec:pascal_framework_validation}
+\section{Pascal Analyzer validation} \label{sec:pascal_framework_validation}
 In this Section, we discuss the measurements and simulation results for three distinct cases, including two real applications.
@@
 % \end{figure}
-\section{Fingerprint Tool validation} \label{sec:fingerprint_tool_validation}
+\section{PMC Module validation} \label{sec:fingerprint_tool_validation}
 

### 2022-08-19 thesis/main.tex
conclusion

@@
 Our heuristic can reduce the scan space from $10^{7000}$ to $10^2$ with an average error of 10\% and up to 38\% reduction in energy consumption using optimal distribution compared to standard Linux DVFS.
+		A novel normalised time representation of the application serving to characterize the application parameters and model, named application finger print. 
 	\end{umonsThesisAbstract}
@@
 
-	\chapter{Conclusions and future work*} \label{chapter:conclusions}
+	\chapter{Conclusions and future work} \label{chapter:conclusions}
 	In this chapter, we deal with the conclusions about the effectiveness of the frameworks and the advantages and disadvantages of using models and algorithms with additional information on applications and architectures for energy optimization.
